### 2024-07-30 src/introduction.tex
Creating the repository

\section{Introduction}
\label{introduction:sec}

\subparagraph{Motivation}

### 2024-07-30 src/metadata.tex
Creating the repository

% Description: Metadata for the plain template
%
% Warning: this file is generated automatically
% from the paper-meta.yaml file. If you really want to edit
% this file, please mirror the changes to paper-meta.yaml.

\title{Well-quasi-orderings on word languages}
\author{%
        Roland Guttenberg \and
        Nathan Lhote \and
        Aliaume Lopez \and
        Corto Mascle \and
        Vincent Michielini \and
        Lia Schütze \and
        Omid Yaghoubi}

### 2024-07-30 wqo-on-words.tex
Creating the repository

\documentclass[11pt]{article}

\input{lib/packages}
\input{lib/aliaume}

\input{src/metadata.tex}

\begin{document}
\maketitle

\input{src/introduction}

\end{document}

### 2024-07-30 src/infixes.tex
starting to write a bit

% lang: en-US

\section{Infixes}
\label{infixes:sec}

In this section, we study the well-quasi-ordering of languages under the infix
relation.

\subsection{Lower bound}

Approximative theorem.
\begin{theorem}
    Let $(X, \preceq)$ be a quasi-ordered set.
    Then the following are equivalent
    \begin{itemize}
        \item $X$ embeds into $(\Sigma^*, \preceq_{\text{infix}})$
        \item $X$ is countable, and for every $x \in X$,
            $\{y \in X \mid y \preceq x\}$ is finite.
    \end{itemize}

    Furthermore, if $X$ is recursively enumerable and $\preceq$ decidable,
    then
    the embedding is effective.
\end{theorem}

\subsection{Upper bound}

This section will use so-called \emph{amalgamation systems}.

\begin{itemize}
    \item Define amalgamation systems
    \item Pumping argument
    \item Characterization
    \item Decision procedure
\end{itemize}

### 2024-07-30 src/introduction.tex
starting to write a bit

@@
 
-\subparagraph{Motivation}
+A \intro{well-quasi-ordered} set is a set $X$ equipped with a quasi-order
+$\preceq$ such that every infinite sequence $x_0 \preceq x_1 \preceq x_2
+\preceq \ldots$ contains a pair $x_i \preceq x_j$ with $i < j$.
+Well-quasi-orderings serve as a core combinatorial tool powering many
+termination arguments, and in particular applies to the verification of
+infinite state transition systems \cite{ABDU96,ABDU98}.
+
+One of the appealing properties of well-quasi-orderings is that they are closed
+under many operations, such as taking products, finite unions, and finite
+powersets \cite{SCSC12}. More surprisingly, the class of well-quasi-ordered
+sets is also stable under the operation of taking finite words and finite trees
+labelled by elements of a well-quasi-ordered set \cite{HIG52,KRU72}.
+
+Note that in these latter cases, the precise choice of ordering on words and
+trees is crucial to ensure that the resulting structure is well-quasi-ordered.
+A celebrated result of Higman implies that the set of finite words over a
+finite alphabet is well-quasi-ordered by the so-called \kl{subword relation}
+\cite{HIG52}. The subword relation for words over $(X, \preceq)$ is defined as
+follows: a word $u$ is a \intro{subword} of a word $v$ if there exists an
+increasing function $f \colon \{1, \ldots, |u|\} \to \{1, \ldots, |v|\}$ such
+that $u_i \preceq v_{f(i)}$ for all $i \in \{1, \ldots, |u|\}$.
+
+There are many other natural orderings on words that could be considered in the
+context of well-quasi-orderings. For instance, one could consider the
+\kl{prefix relation}, \kl{suffix relation} or the \kl{infix relation}, all of
+which are not well-quasi-ordered as soon as the alphabet contains two distinct
+letters. Let us remark that these other word orderings are however
+well-founded, and therefore the (only) obstruction to being well-quasi-ordered
+is not due to the existence of infinite sets of incomparable elements.
+
+\subparagraph{Contribution} In this paper, we characterize for which languages
+$L \subseteq \Sigma^*$ the other natural orderins on word are
+well-quasi-ordered. When languages are decided by simple enough machines (e.g.,
+finite automata or pushdown automata), we will also provide \emph{decision
+procedures} to check whether the language is well-quasi-ordered by the
+considered relation.
+
+\subparagraph{Outline}

### 2024-07-30 src/prefixes.tex
starting to write a bit

% lang: en-US
\section{Prefixes and Suffixes}
\label{prefixes-suffixes:sec}

In this section, we study the well-quasi-ordering of languages under the prefix
relation. Let us first remark that the case of the suffix relation is
completely symmetric, and therefore that all our results will apply to the
suffix relation as well.

\begin{itemize}
    \item examples: $a^n$, $ab^n$, $b^na$. 
    \item the tree of prefixes and existence antichains
    \item decision procedure for regular languages
\end{itemize}

### 2024-08-05 src/introduction.tex
update intro

@@
+% LTeX: language=en-GB 
 \section{Introduction}
@@
 A \intro{well-quasi-ordered} set is a set $X$ equipped with a quasi-order
-$\preceq$ such that every infinite sequence $x_0 \preceq x_1 \preceq x_2
-\preceq \ldots$ contains a pair $x_i \preceq x_j$ with $i < j$.
-Well-quasi-orderings serve as a core combinatorial tool powering many
-termination arguments, and in particular applies to the verification of
-infinite state transition systems \cite{ABDU96,ABDU98}.
+$\preceq$ such that every infinite sequence $x_0, x_1, x_2 , \ldots$ contains a
+pair $x_i \preceq x_j$ with $i < j$. Well-quasi-orderings serve as a core
+combinatorial tool powering many termination arguments, and in particular
+applies to the verification of infinite state transition systems
+\cite{ABDU96,ABDU98}. 
+In this paper, we will use an equivalent definition of
+well-quasi-orderings based on the non-existence of infinite \intro{antichains}
+-- sets of pairwise incomparable elements -- as well as the non-existence of
+infinite \intro{decreasing sequences} -- sequences of elements $x_0, x_1, x_2,
+\ldots$ such that $x_{i+1} \prec x_{i}$ --.
 
@@
 \subparagraph{Contribution} In this paper, we characterize for which languages
-$L \subseteq \Sigma^*$ the other natural orderins on word are
+$L \subseteq \Sigma^*$ the other natural orderings on word are
 well-quasi-ordered. When languages are decided by simple enough machines (e.g.,
@@
 procedures} to check whether the language is well-quasi-ordered by the
-considered relation.
+considered relation. Quite surprisingly, for the \kl{prefix relation} and the
+\kl{suffix relation}, the only languages that are \kl{well-quasi-ordered} are
+the \emph{finite} unions of \kl{chains}, that is, the simplest possible
+well-quasi-ordered sets.
 

### 2024-08-05 src/infixes.tex
adding monoid questions...
Adding automata based approach + theorem

@@
 In this section, we study the well-quasi-ordering of languages under the infix
-relation.
+relation. As opposed to the prefix and suffix relations, the infix relation can
+yield to very complicated well-quasi-ordered languages. As a matter of fact,
+the following theorem shows that \emph{any} countable quasi-ordering with
+finite initial segments can be embedded into the infix relation of a language.
 
-\subsection{Lower bound}
-
-Approximative theorem.
 \begin{theorem}
+    \label{thm:infix-embedding}
     Let $(X, \preceq)$ be a quasi-ordered set.
@@
 
-\subsection{Upper bound}
+As a consequence of \cref{thm:infix-embedding}, we cannot replay proofs of
+\cref{prefixes-suffixes:sec}, and will actually need to leverage some
+regularity of the languages to obtain a characterization of well-quasi-ordered
+languages under the infix relation. Let us first play this game for languages
+that are recognized by finite automata.
+
+\begin{theorem}
+    \label{thm:infix-finite-automata}
+    Let $L \subseteq \Sigma^*$ be a language recognized by a finite automaton.
+    Then $L$ is well-quasi-ordered by the infix relation if and only if $L$ is
+    a finite union of chains for the \kl{infix relation}.
+\end{theorem}
+
+\begin{corollary}
+    Decidability?
+\end{corollary}
+
+\subsection{Amalgamation systems}
 
@@
 \end{itemize}
+
+\subsection{Monoid equations?}
+
+\begin{itemize}
+    \item What are the finite monoids $(M, \cdot)$ such that
+    every language recognized by $M$ is well-quasi-ordered by the infix relation?
+    \item What are the equations on $M$ and $P \subseteq M$ characterizing 
+    the well-quasi-ordering of $L(M, P)$?
+\end{itemize}

### 2024-08-05 src/prefixes.tex
Add statements and proof scheme

@@
-% lang: en-US
+% LTeX: language=en-GB 
 \section{Prefixes and Suffixes}
@@
 In this section, we study the well-quasi-ordering of languages under the prefix
-relation. Let us first remark that the case of the suffix relation is
-completely symmetric, and therefore that all our results will apply to the
-suffix relation as well.
+relation. A word $u$ is a \intro{prefix} of a word $w$ if there exists a word
+$v$ such that $w = uv$. We denote this relation by $u \intro*\prefleq w$.
+Similarly, a word $u$ is a \intro{suffix} of a word $w$ if there exists a word
+$v$ such that $w = vu$. We denote this relation by $u \intro*\suffleq w$.
 
-\begin{itemize}
-    \item examples: $a^n$, $ab^n$, $b^na$. 
-    \item the tree of prefixes and existence antichains
-    \item decision procedure for regular languages
-\end{itemize}
+Let us immediately remark that the map $u \mapsto u^R$ that reverses a word is
+an order-bijection between $(X^*, \prefleq)$ and $(X^*, \suffleq)$. Therefore,
+we will focus on the prefix relation in the following.
+
+As mentioned in the introduction, the prefix relation is not a well-quasi-order
+on words over a two-letter alphabet, as witnessed by the infinite
+\kl{antichain} $\setof{a^nb}{n \in \Nat}$.
+
+Let us order the words in $\Sigma^*$ by the \kl{prefix relation} to build an
+infinite tree $T$. The root of the tree is the empty word $\varepsilon$, and
+the children of a word $w$ are the words $wa$ for all $a \in \Sigma$. This tree
+is finitely branching. Let $L \subseteq \Sigma^*$ be a language. Assume that
+$L$ contains an infinite \kl{antichain} $A$ for the \kl{prefix relation}. In
+particular, no two word of $A$ belong to the same branch of the tree $T$. Let
+us prove the following: there exists an infinite branch $B$ of $T$ such that from
+every point of the branch, one can reach a word of $A$ that is not in the
+branch.
+
+\begin{definition}
+    An \intro{antichain branch} for a language $L$ is an infinite 
+    branch $B$ of the tree $T$ such that from every point of the branch, 
+    one can reach a word of $A$ that is not in the branch.
+\end{definition}
+
+\begin{lemma}
+    Let $L \subseteq \Sigma^*$ be a language. Then, $L$ contains an infinite
+    \kl{antichain} if and only if there exists an \kl{antichain branch} for $L$.
+\end{lemma}
+
+\begin{corollary}
+    If $L$ is regular, then the existence of an infinite antichain is decidable.
+\end{corollary}
+
+\begin{theorem}
+    A language $L \subseteq \Sigma^*$ is \kl{well-quasi-ordered} by the
+    \kl{prefix relation} if and only if $L$ is a union of \kl{chains}.
+\end{theorem}
+
+\begin{corollary}
+    Let $L \subseteq \Sigma^*$ be a language that is \kl{well-quasi-ordered} by
+    the \kl{prefix relation}. Then, there exists a finite $k \in \Nat$ such that
+    the
+    \kl{maximal order type} of $L$ is $k \cdot \omega$,
+    the \kl{ordinal height} of $L$ is $\omega$, and its
+    \kl{ordinal width} is $k$.
+\end{corollary}

### 2024-08-06 src/infixes.tex
rewriting introduction

@@
 
-In this section, we study the well-quasi-ordering of languages under the infix
-relation. As opposed to the prefix and suffix relations, the infix relation can
-yield to very complicated well-quasi-ordered languages. As a matter of fact,
-the following theorem shows that \emph{any} countable quasi-ordering with
-finite initial segments can be embedded into the infix relation of a language.
+In this section, we study the well-quasi-ordering of languages under the
+\kl{infix relation}. As opposed to the \kl{prefix} and \kl{suffix} relations,
+the \kl{infix relation} can yield to very complicated \kl{well-quasi-ordered}
+languages. Formally, the following theorem shows that \emph{any} countable
+quasi-ordering with finite initial segments can be embedded into the infix
+relation of a language.
 

### 2024-08-06 src/introduction.tex
rewriting introduction

--- 2024-07-30 src/introduction.tex
@@
 
-\subparagraph{Motivation}
+A \intro{well-quasi-ordered} set is a set $X$ equipped with a quasi-order
+$\preceq$ such that every infinite sequence $\seqof{x_n}$ of elements taken in
+$X$ contains a pair $x_i \preceq x_j$ with $i < j$. Well-quasi-orderings serve
+as a core combinatorial tool powering many termination arguments, and was
+successfully applied to the verification of infinite state transition systems
+\cite{ABDU96,ABDU98}. One of the appealing properties of well-quasi-orderings
+is that they are closed under many operations, such as taking products, finite
+unions, and finite powersets \cite{SCSC12}. More surprisingly, the class of
+well-quasi-ordered sets is also stable under the operation of taking finite
+words and finite trees labelled by elements of a well-quasi-ordered set
+\cite{HIG52,KRU72}.
+
+Note that in the case of finite words and finite trees, the precise choice of
+ordering on words and trees is crucial to ensure that the resulting structure
+is \kl{well-quasi-ordered}. The celebrated result of Higman states that the set
+of finite words over an ordered alphabet $(X, \preceq)$ is
+\kl{well-quasi-ordered} by the so-called \kl{subword embedding relation}
+\cite{HIG52}. Let us recall that the \kl{subword relation} for words over $(X,
+\preceq)$ is defined as follows: a word $u$ is a \intro{subword} of a word $v$
+if there exists an increasing function $f \colon \{1, \ldots, |u|\} \to \{1,
+\ldots, |v|\}$ such that $u_i \preceq v_{f(i)}$ for all $i \in \{1, \ldots,
+|u|\}$.
+
+However, there are many other natural orderings on words that could be
+considered in the context of \kl{well-quasi-orderings}. The three alternatives
+we consider are the \kl{prefix relation}, the \kl{suffix relation}, and the
+\kl{infix relation}. Unfortunately, it is an easy check that all of these
+constructions are not well-quasi-ordered as soon as the alphabet contains two
+non-equivalent letters. This is not the end of the road, because some
+\emph{subsets} of $\Sigma^*$ might still be \kl{well-quasi-ordered} by these
+relations, the simplest example being the case of finite sets of (finite) words
+which are all \kl{well-quasi-ordered} regardless of the ordering considered.
+This raises the question of characterizing the subsets $L \subseteq \Sigma^*$
+that are \kl{well-quasi-ordered} with respect to the \kl{prefix relation}
+(respectively, the \kl{suffix relation} or the \kl{infix relation}), and
+providing suitable decision procedures.
+
+\subparagraph{Contribution} In this paper, we focus on words over a finite
+alphabet $\Sigma$, and characterize subsets $L \subseteq \Sigma^*$ that are
+\kl{well-quasi-ordered} by the \kl{prefix relation}, the \kl{suffix relation},
+and the \kl{infix relation}. Let us remark that these word orderings are all
+\intro{well-founded}, i.e., they do not contain infinite \intro{decreasing
+sequences} $\seqof{x_n}$ such that $x_{n+1} \prec x_n$ for all $n \in \Nat$.
+Let us recall that the key difference between \kl{well-quasi-orderings} and the
+more usual notion of well-founded orderings is that the latter allows the
+existence of infinite \intro{antichains}, that is, sets of pairwise
+incomparable elements. In fact, a set $X$ is \kl{well-quasi-ordered} if and
+only if it is \kl{well-founded} and does not contain an infinite \kl{antichain}
+\cite{KRU72}.  As a consequence, our study focuses on detecting infinite
+\kl{antichains} in languages $L \subseteq \Sigma^*$.
+
+In the case of the \kl{prefix relation}, we provide a complete characterization
+of languages $L \subseteq \Sigma^*$ that are \kl{well-quasi-ordered}, which are
+the simplest possible languages: they are precisely the \emph{finite} unions of
+\intro{chains}, that is, sets of words that are all comparable for the
+\kl{prefix relation}. Note that \kl{chains} are the simplest possible
+\kl{well-quasi-ordered} sets (they are totally ordered and well-founded) and it
+is known that finite unions of \kl{well-quasi-ordered} sets are
+\kl{well-quasi-ordered}. As a consequence, the above characterization states
+that only the simplest possible \kl{well-quasi-ordered} sets are
+\kl{well-quasi-ordered} by the \kl{prefix relation}. Let us furthermore
+highlight that this characterization holds without any restriction on the
+decidability of the language $L$ itself, but heavily relies on the assumption
+that $\Sigma$ is finite. By symmetry, we obtain an analogue statement for the
+\kl{suffix relation}.
+
+In the case of the \kl{infix relation}, the situation is more complex, and we
+show that \emph{any} countable quasi-ordering with finite initial segments can
+be embedded into the \kl{infix relation} of a language over a binary alphabet
+$\Sigma \defined \set{a,b}$. As an example, on can embed $(\Sigma^*, \higleq)$
+into $(\Sigma^*, \infleq)$ where $\higleq$ is the \kl{subword relation} and
+$\infleq$ is the \kl{infix relation}. This strongly contrasts with the
+\kl{prefix relation}, and suggests that a meaningful characterization of
+\kl{well-quasi-ordered} languages under the \kl{infix relation} is unlikely to
+be found.
+
+We circumvent this issue by adding some regularity assumptions on the languages
+$L \subseteq \Sigma^*$ that we consider. This can be done by assuming that $L$
+is recognized by a finite automaton, or more generally that $L$ is the language
+of an \kl{amalgamation system}. Under these relatively mild assumptions, we
+also prove that \kl{well-quasi-ordered} languages under the \kl{infix relation}
+are precisely the finite unions of \kl{chains}.
+
+\subparagraph{Related work} The study of alternative \kl{well-quasi-ordered}
+relations over finite words is far from new. For instance, orders obtained by
+so-called \emph{derivation relations} where already analysed by Bucher,
+Ehrenfeucht, and Haussler \cite{BUEUD85}. These results were later extended by
+D'Alessandro and Varricchio \cite{ALVA03,ALVA06}. However, in all those cases
+the orderings are \emph{multiplicative}, that is, if $u_1 \preceq v_1$ and $u_2
+\preceq v_2$ then $u_1u_2 \preceq v_1v_2$. This assumption does not hold for
+the \kl{prefix}, \kl{suffix}, and \kl{infix} relations.
+
+\subparagraph{Outline} In \cref{prefixes:sec}, which is relatively
+self-contained, we study the \kl{prefix relation} prove in \cref{prefixes:thm}
+the characterization of \kl{well-quasi-ordered} languages by the \kl{prefix
+relation}. In \cref{infixes:sec}, we prove that any countable quasi-ordering
+with finite initial segments can be embedded into the \kl{infix relation} of a
+language (\cref{count-embed-infix:thm}), we obtain the analogue of
+\cref{prefixes:thm} for languages recognized by finite automata
+(\cref{infix-finite-automata:thm}). Finally, in \cref{amalgamation:sec}, we
+introduce \kl{amalgamation systems} and generalize
+\cref{infix-finite-automata:thm} to this setting
+(\cref{infix-amalgamation:thm}).
+
+\subparagraph{Acknowledgements} We would like to thank participants of the 2024
+edition of Autobóz for their helpful comments and discussions.

### 2024-08-06 src/prefixes.tex
rewriting introduction

@@
 Similarly, a word $u$ is a \intro{suffix} of a word $w$ if there exists a word
-$v$ such that $w = vu$. We denote this relation by $u \intro*\suffleq w$.
+$v$ such that $w = vu$. We denote this relation by $u \intro*\suffleq w$. Let
+us immediately remark that the map $u \mapsto u^R$ that reverses a word is an
+order-bijection between $(X^*, \prefleq)$ and $(X^*, \suffleq)$. Therefore, we
+will focus on the prefix relation in the rest of this section.
 
-Let us immediately remark that the map $u \mapsto u^R$ that reverses a word is
-an order-bijection between $(X^*, \prefleq)$ and $(X^*, \suffleq)$. Therefore,
-we will focus on the prefix relation in the following.
+Let us briefly restate the fact that some (even regular) languages 
+are not well-quasi-ordered by the prefix relation.
 
-As mentioned in the introduction, the prefix relation is not a well-quasi-order
-on words over a two-letter alphabet, as witnessed by the infinite
-\kl{antichain} $\setof{a^nb}{n \in \Nat}$.
+\begin{example}
+    The set $L = \setof{a^nb}{n \in \Nat}$ is an infinite antichain for the
+    prefix relation.
+\end{example}
 
-Let us order the words in $\Sigma^*$ by the \kl{prefix relation} to build an
-infinite tree $T$. The root of the tree is the empty word $\varepsilon$, and
-the children of a word $w$ are the words $wa$ for all $a \in \Sigma$. This tree
-is finitely branching. Let $L \subseteq \Sigma^*$ be a language. Assume that
-$L$ contains an infinite \kl{antichain} $A$ for the \kl{prefix relation}. In
-particular, no two word of $A$ belong to the same branch of the tree $T$. Let
-us prove the following: there exists an infinite branch $B$ of $T$ such that from
-every point of the branch, one can reach a word of $A$ that is not in the
-branch.
+In order to characterize the existence of infinite antichains for the prefix
+relation, we will introduce the following tree construction that
+will be useful in the rest of this section.
+
+\begin{definition}
+    The \intro{tree of prefixes} over a finite alphabet $\Sigma$
+    is the infinite tree $T$ whose nodes are the words of $\Sigma^*$, and
+    such that the children of a word $w$ are the words $wa$ for all $a \in
+    \Sigma$. 
+\end{definition}
+
+Notice that the tree of prefixes is finitely branching. Let us now
+observe how antichains in the prefix relation can be witnessed
+by infinite branches in the tree of prefixes.
 
@@
 \end{definition}
+
+It is clear that an \kl{antichain branch} for a language $L$ yields an infinite
+antichain, and the converse is quite easy to prove. Because the notion of
+antichain branch is \kl{$\MSO$-definable} whenever $L$ is regular, we
+immediately obtain decidability as a corollary of this lemma.
 
@@
 
+Let us now go further and fully characterize languages $L$ such that the
+prefix relation is well-quasi-ordered, even without any restriction on the
+decidability of $L$ itself. Let us remark that finite unions of \kl{chains} are
+always \kl{well-quasi-ordered} by the \kl{prefix relation} because they lack
+infinite \kl{antichains} by definition. The following theorem states that this
+is the only possible reason for a language $L$ to be \kl{well-quasi-ordered} by
+the \kl{prefix relation}.
+
 \begin{theorem}
@@
 
+As an immediate consequence, we have a very fine-grained understanding of the
+\kl{ordinal invariants} of such \kl{well-quasi-ordered} languages, which can be
+leveraged in bounding the complexity of algorithms working on such languages.
+
 \begin{corollary}
+    \textbf{false: some chains can be finite}
     Let $L \subseteq \Sigma^*$ be a language that is \kl{well-quasi-ordered} by
@@
 \end{corollary}
+
+Let us conclude by noting that it is unsurprisingly not possible to decide
+whether a decidable language is \kl{well-quasi-ordered} by the \kl{prefix
+relation}.
+
+\begin{lemma}
+    todo.
+\end{lemma}

### 2024-08-06 src/prefixes.tex
adding some proofs

@@
 \begin{example}
-    The set $L = \setof{a^nb}{n \in \Nat}$ is an infinite antichain for the
-    prefix relation.
+    The set $L = \setof{a^nb}{n \in \Nat} \subseteq \set{a,b}^*$ is an infinite \kl{antichain} for the
+    \kl{prefix relation}.
 \end{example}
@@
 
+\begin{figure}
+    \centering
+    \caption{The tree of prefixes over the alphabet $\set{a,b}$,
+        with an \kl{antichain} in red, and
+        an \kl{antichain branch} in blue.}
+\end{figure}
+
 Notice that the tree of prefixes is finitely branching. Let us now
@@
     An \intro{antichain branch} for a language $L$ is an infinite 
-    branch $B$ of the tree $T$ such that from every point of the branch, 
-    one can reach a word of $A$ that is not in the branch.
+    branch $B$ of the \kl{tree of prefixes} such that from every point of the branch, 
+    one can reach a word of $L$ that is not in the branch.
 \end{definition}
@@
 \end{lemma}
+\begin{proof}
+    Assume that $L$ contains an \kl{antichain branch}. Let us construct an
+    infinite \kl{antichain} as follows. We start with a set $A_0 \defined
+    \emptyset$ and a node $v_0$ at the root of the tree. At step $i$, we
+    consider a word $w_i$ such that $v_i$ is a \kl{prefix} of $w_i$, and $w_i
+    \in L \setminus B$, which exists by definition of \kl{antichain branches}.
+    We then set $A_{i+1} \defined A_i \cup \set{w_i}$. To compute $v_{i+1}$, we
+    consider the largest prefix of $w_i$ that belongs to $B$, and set $v_{i+1}$
+    to be the successor of this prefix in $B$. By an immediate induction, we
+    conclude that for all $i \in \Nat$, $A_i$ is an \kl{antichain}, and that
+    $v_i$ is a node in the \kl{antichain branch} $B$ such that $v_i$ is not a
+    prefix of any word in $A_i$.
+
+    Conversely, assume that $L$ contains an infinite \kl{antichain} $A$. Let us
+    construct an \kl{antichain branch}. Let us consider the subtree of the
+    \kl{tree of prefixes} that consists in words that are \kl{prefixes} of
+    words in $A$. This subtree is infinite, and by König's lemma, it contains
+    an infinite branch. By definition this is an \kl{antichain branch}.
+\end{proof}
 
@@
 \end{corollary}
+\begin{proof}
+    If $L$ is regular, then it is \kl{$\MSO$-definable}, and there 
+    exists a formula $\varphi(x)$ in \kl{$\MSO$} that selects nodes 
+    of the \kl{tree of prefixes} that belong to $L$. Now, to decide whether there
+    exists an \kl{antichain branch} for $L$, we can simply check whether
+    the following formula is satisfied:
+    \begin{equation*}
+        \exists B. 
+        B \text{ is a branch } \land
+        \forall x \in B, \exists y. y \text{ is a child of } x \land \varphi(y) \land y \not\in B
+        \quad .
+    \end{equation*}
+    Because the above formula is an $\MSO$-formula over the infinite
+    binary tree, whether it is satisfied is decidable.
+    \textbf{TODO: cite}.
+\end{proof}
 
@@
 \end{theorem}
+\begin{proof}
+    Assume that $L$ is a finite union of \kl{chains}.
+    Because the \kl{prefix relation} is \kl{well-founded},
+    and that finite unions of \kl{chains} have finite \kl{antichains},
+    we conclude that $L$ is \kl{well-quasi-ordered}.
+
+    Conversely, assume that $L$ is not a finite union of \kl{chains}. Then
+    there exists an infinite \kl{antichain branch} $B$ for $L$. Indeed
+    \textbf{todo}. In particular, $L$ is not \kl{well-quasi-ordered}.
+\end{proof}
 

### 2024-08-07 src/conclusion.tex
writing proofs

% LTeX: language=en-GB 
\section{Conclusion}
\label{conclusion:sec}

\subparagraph{Towards infinite alphabets} Can we go to
alphabets that are infinite but equipped with a well-quasi order?

\subparagraph{Monoid equations}

### 2024-08-07 src/infixes.tex
writing proofs

@@
-% lang: en-US
-
-\section{Infixes}
-\label{infixes:sec}
+% LTeX: language=en-GB 
+\section{Infixes and Regular Languages}
+\label{infixes-regular:sec}
 
@@
 
+Let us recall that an \intro{order embedding} from a quasi-ordered set $(X,
+\preceq)$ into a quasi-ordered set $(Y, \preceq')$ is a function $f \colon X
+\to Y$ such that for all $x, y \in X$, $x \preceq y$ if and only if $f(x)
+\preceq' f(y)$. When such an embedding exists, we say that $X$ \reintro{embeds
+into} $Y$.
+
+We say that a quasi-ordered set $(X, \preceq)$ is a \intro{partial ordering}
+whenever the relation $\preceq$ is antisymmetric, that is $x \preceq y$ and $y
+\preceq x$ implies $x = y$.
+
 \begin{theorem}
-    \label{thm:infix-embedding}
-    Let $(X, \preceq)$ be a quasi-ordered set.
-    Then the following are equivalent
-    \begin{itemize}
-        \item $X$ embeds into $(\Sigma^*, \preceq_{\text{infix}})$
-        \item $X$ is countable, and for every $x \in X$,
-            $\{y \in X \mid y \preceq x\}$ is finite.
-    \end{itemize}
-
+    \label{infix-embedding:thm}
+    Let $(X, \preceq)$ be a \kl{partially ordered} set,
+    and $\Sigma \defined \set{a,b,c}$.
+    Then the following are equivalent:
+    \begin{enumerate}
+        \item \label{infix-embedding-embeds:item} 
+            $X$ \kl{embeds into} $(\Sigma^*, \infleq)$,
+        \item \label{infix-embedding-count:item}
+            $X$ is countable, and for every $x \in X$,
+        $\setof{y \in X}{y \preceq x}$ is finite.
+    \end{enumerate}
     Furthermore, if $X$ is recursively enumerable and $\preceq$ decidable,
     then
-    the embedding is effective.
+    the \kl{embedding} is computable.
 \end{theorem}
+\begin{proof}
+    Let us first prove that \cref{infix-embedding-embeds:item} implies
+    \cref{infix-embedding-count:item}. Let $f \colon X \to \Sigma^*$ be an
+    \kl{embedding}. Note that $f$ is injective, because $X$ and $\Sigma^*$
+    are \kl{partially
+    ordered}. In particular, because $\dwset[\infleq]{w}$ is finite
+    for every $w \in \Sigma^*$, then so is $\setof{y \in X}{y \preceq x}$
+    for every $x \in X$. Similarly, $X$ must be countable
+    because $\Sigma^*$ is.
 
-As a consequence of \cref{thm:infix-embedding}, we cannot replay proofs of
-\cref{prefixes-suffixes:sec}, and will actually need to leverage some
+    Conversely, let us prove that \cref{infix-embedding-count:item} implies
+    \cref{infix-embedding-embeds:item}. To that end, let us first define
+    inductively a sequence $\seqof{X_n}$ of subsets of $X$ as follows: $X_0$ is
+    the set of minimal elements of $X$, and $X_{n+1}$ is the set of minimal
+    elements of $X \setminus \bigcup_{i \leq n} X_i$. Because initial segments
+    of $X$ are finite, it is clear that $X = \bigcup_{n \in \Nat} X_n$. To
+    simplify notations, we will also use $Y_n \defined \bigcup_{i \leq n} X_i$.
+    We also let $L_0 \defined a b^* a$, and 
+    $L_{n+1} \defined L_n \cup c^{n+1} a^{n+1} b^* a^{n+1} c^{n+1} (L_n c^{n+1})^*$.
+    Now, let us inductively define \kl{embeddings} $f_n \colon Y_n \to
+    L_n$.
+
+    For the base case, let $\seqof{x_i}[i \in \Nat]$ be an enumeration of
+    $X_0$, which is possible because $X$ is countable. We define $f_0 \colon
+    X_0 \to \Sigma^*$ as follows: $f_0(x_i) \defined a b^i a$. It is immediate
+    that $f_0$ is an \kl{embedding}. For the inductive step, let $\seqof{x_i}[i
+    \in \Nat]$ be an enumeration of $X_{n+1}$, and assume that $f_n$ is an
+    \kl{embedding} from $Y_n$ to $\Sigma^*$.
+    Let us define $D_i \defined \setof{ y \in Y_n }{ y \preceq x_i }$
+    for $i \in \Nat$, all of which are finite by assumption.
+    We then let for all $i \in \Nat$:
+    \begin{equation*}
+        f_{n+1}(x_i) \defined 
+        \underbrace{c^{n+1} a^{n+1} b^i a^{n+1} c^{n+1} \prod_{y \in D_i} \overbrace{f_n(y)}^{\in L_n} c^{n+1}}_{\in L_{n+1}}
+        \quad .
+    \end{equation*}
+    And let $f_{n+1}(y) \defined f_n(y)$ for all $y \in Y_n$.
+
+    To check that $f_{n+1}$ is an \kl{embedding}, let us first notice that
+    words $f_{n+1}(x_i)$ and $f_{n+1}(x_j)$ are incomparable for the \kl{infix
+    relation} when $i \neq j$, because the only factor of the form $c^{n+1}
+    a^{n+1} b^t a^{n+1} c^{n+1}$ appears at the beginning of these words, and
+    precisely encode $i$ (resp. $j$) in the value of $t$. Now, let us consider
+    $y \in Y_n$ and $x_i \in X_{n+1}$ such that $y \preceq x_i$. Because $y \in
+    D_i$, we have $f_n(y) \infleq f_{n+1}(x_i)$ by construction, and since
+    $f_n(y) = f_{n+1}(y)$, we conclude that $f_{n+1}(y) \infleq f_{n+1}(x_i)$. 
+    Conversely, assume that
+    $f_{n+1}(y) \infleq f_{n+1}(x_i)$. Again, this means
+    that $f_n(y) \infleq f_{n+1}(x_i)$. Now, the encoding
+    is robust enough so that we can conclude
+    $f_{n}(y) \infleq f_{n}(z)$ for some $z \in D_i$.
+    By induction hypothesis, $y \preceq z$, hence, $y \preceq x_i$.
+
+    We conclude by noticing that $\bigcup_{n \in \Nat} f_n$ is an \kl{embedding}
+    of $X$ into $\Sigma^*$.
+\end{proof}
+
+As a consequence of \cref{infix-embedding:thm}, we cannot replay proofs of
+\cref{prefixes:sec}, and will actually need to leverage some
 regularity of the languages to obtain a characterization of well-quasi-ordered
@@
 
+\textbf{TODO: introduce finite automata and refer to the book.}.
+
 \begin{theorem}
-    \label{thm:infix-finite-automata}
+    \label{infix-finite-automata:thm}
     Let $L \subseteq \Sigma^*$ be a language recognized by a finite automaton.
@@
 
+In order to prove \cref{infix-finite-automata:thm}, we will perform
+some preliminary analysis on the structure of an automaton recognizing a
+well-quasi-ordered language under the infix relation.
+
+\begin{lemma}
+    Let $A = (Q, \Sigma, \delta, q_0, F)$ be a finite automaton recognizing a
+    language $L \subseteq \Sigma^*$. Assume furthermore that
+    $L$ is \kl{well-quasi-ordered} by the \kl{infix relation}.
+    Then, for all words $u,w,v \in \Sigma^*$, and states
+    $q, q_f \in Q \times F$ such that
+    $q_0 \xrightarrow{u} q \xrightarrow{w} q \xrightarrow{v} q_f$,
+    one of the following holds:
+    \begin{itemize}
+        \item There exists $k \in \Nat$
+            such that $u$ is a \kl{suffix} of $w^k$,
+        \item There exists $k \in \Nat$
+            such that $v$ is a \kl{prefix} of $w^k$.
+    \end{itemize}
+\end{lemma}
+
+\begin{lemma}
+    Let $A = (Q, \Sigma, \delta, q_0, F)$ be a finite automaton recognizing a
+    language $L \subseteq \Sigma^*$. Assume furthermore that
+    $L$ is \kl{well-quasi-ordered} by the \kl{infix relation}.
+    Then, for all words $w_1, w_2 \in \Sigma^*$, and states
+    $q, q_f \in Q \times F$ such that
+    $q_0 \rightarrow^* q \xrightarrow{w} q \rightarrow^* q_f$
+    and 
+    $q_0 \rightarrow^* q \xrightarrow{w_2} q \rightarrow^* q_f$,
+    there exists a word $w \in \Sigma^*$
+    such that $w_1$ and $w_2$ are powers of $w$.
+\end{lemma}
+
+\begin{lemma}
+    Let $A = (Q, \Sigma, \delta, q_0, F)$ be a finite automaton recognizing a
+    language $L \subseteq \Sigma^*$. Assume furthermore that
+    $L$ is \kl{well-quasi-ordered} by the \kl{infix relation}.
+    Then, for all words $u, w_1, w_2 \in \Sigma^*$, and states
+    $q_1, q_2, q_f \in Q^2 \times F$,
+    such that
+    $q_0 \rightarrow^* q_1 \xrightarrow{w_1} q_1 \xrightarrow{u} q_2 \xrightarrow{w_2} q_2
+    \rightarrow^* q_f$,
+    there exists a $w \in \Sigma^*$ such that $w_1$, $w_2$, and $u$ are powers of $w$.
+\end{lemma}
+
+We are now ready to prove the main result.
+
+\begin{proofof}{infix-finite-automata:thm}
+    Let $w \in L$. Consider a run of the automaton on $w$. If $w$ is too long,
+    then there exists a loop, and we can prove that $w = u x^p v$ for some
+    short period $x$. By repeating the process for $u$ and $v$, we conclude
+    that $w$ has the form $w = u_0 x_0^{p_0} u_1 x_1^{p_1} \cdots u_k x_k^{p_k}
+    v_k$. Now, applying the loop merging lemma, we conclude that $x_0 = x_1 =
+    \cdots = x_k$. Hence, $w = u_0 x^{p} v_k$. Finally, we know that either
+    $u_0$ is a prefix of $x^{p}$ or that $v_k$ is a suffix of $x^{p}$. We
+    conclude in particular that $\dwset[\infleq]{ \setof{ u x^p v }{ p \in \Nat
+    }}$ is a finite union of \kl{chains}. Because $u,v,x$ are of bounded
+    length, we conclude that $L$ itself included in a finite union of
+    \kl{chains}, hence is a finite union of \kl{chains}.
+\end{proofof}
+
 \begin{corollary}
@@
 
-\subsection{Amalgamation systems}
+\section{Infixes and Amalgamation systems}
+\label{infixes-amalgamation:sec}
 
 This section will use so-called \emph{amalgamation systems}.
+
+\begin{theorem}
+    \label{infix-amalgamation:thm}
+    Let $L \subseteq \Sigma^*$ be a language recognized by an 
+    \kl{amalgamation system}.
+    Then $L$ is well-quasi-ordered by the infix relation if and only if $L$ is
+    a finite union of chains for the \kl{infix relation}.
+\end{theorem}
 
@@
 \end{itemize}
-
-\subsection{Monoid equations?}
-
-\begin{itemize}
-    \item What are the finite monoids $(M, \cdot)$ such that
-    every language recognized by $M$ is well-quasi-ordered by the infix relation?
-    \item What are the equations on $M$ and $P \subseteq M$ characterizing 
-    the well-quasi-ordering of $L(M, P)$?
-\end{itemize}

### 2024-08-07 src/introduction.tex
writing proofs

@@
 the characterization of \kl{well-quasi-ordered} languages by the \kl{prefix
-relation}. In \cref{infixes:sec}, we prove that any countable quasi-ordering
+relation}. In \cref{infixes-regular:sec}, we prove that any countable quasi-ordering
 with finite initial segments can be embedded into the \kl{infix relation} of a
-language (\cref{count-embed-infix:thm}), we obtain the analogue of
+language (\cref{infix-embedding:thm}), we obtain the analogue of
 \cref{prefixes:thm} for languages recognized by finite automata
-(\cref{infix-finite-automata:thm}). Finally, in \cref{amalgamation:sec}, we
+(\cref{infix-finite-automata:thm}). Finally, in \cref{infixes-amalgamation:sec}, we
 introduce \kl{amalgamation systems} and generalize

### 2024-08-07 src/metadata.tex
writing proofs

@@
         Omid Yaghoubi}
+
+\newcommand{\makeabstract}{
+\begin{abstract}
+    The set of finite words over a well-quasi-ordered set is itself
+    well-quasi-ordered. This seminal result by Higman is a cornerstone
+    of the theory of well-quasi-orderings and has found numerous
+    applications in computer science. However, this result is based on a
+    specific choice of ordering on words, the (scattered) subword
+    ordering. In this paper, to what extent other natural orderings
+    (prefix, suffix, infix) on words can be used to derive Higman-like
+    theorems. More specifically, we are interested in characterizing
+    \emph{languages} of words that are well-quasi-ordered under these
+    orderings. We show that a simple characterization is possible for
+    the prefix and suffix orderings, and that under extra regularity
+    assumptions, this also lifts to the infix ordering.
+\end{abstract}
+}

### 2024-08-07 src/prefixes.tex
writing proofs

@@
 \section{Prefixes and Suffixes}
-\label{prefixes-suffixes:sec}
+\label{prefixes:sec}
 
@@
 
-Notice that the tree of prefixes is finitely branching. Let us now
-observe how antichains in the prefix relation can be witnessed
-by infinite branches in the tree of prefixes.
+Because the alphabet $\Sigma$ is assumed to be finite, the \kl{tree of
+prefixes} is finitely branching. Let us now observe how \kl{antichains} in the
+prefix relation can be witnessed by infinite branches in the \kl{tree of
+prefixes}.
 
@@
 \begin{theorem}
+    \label{prefixes:thm}
     A language $L \subseteq \Sigma^*$ is \kl{well-quasi-ordered} by the
@@
 
-    Conversely, assume that $L$ is not a finite union of \kl{chains}. Then
-    there exists an infinite \kl{antichain branch} $B$ for $L$. Indeed
-    \textbf{todo}. In particular, $L$ is not \kl{well-quasi-ordered}.
+    Conversely, assume that $L$ is \kl{well-quasi-ordered} by the \kl{prefix
+    relation}. Let us define $S$ the set of words $w$ such that there exists
+    two words $wu$ and $wv$ both in $L$ that are not comparable for the
+    \kl{prefix relation}. Assume by contradiction that $S$ is infinite.
+    Then, $S$ equipped with the \kl{prefix relation} is an infinite
+    tree with finite branching, and therefore contains an infinite
+    branch, which is by definition an \kl{antichain branch} for $L$.
+    This contradicts the assumption that $L$ is \kl{well-quasi-ordered}.
+    Now, consider $w$ be a maximal element for the \kl{prefix ordering}
+    in $S$. By definition, all words in $L$ that contain $w$ as a prefix
+    must be comparable for the \kl{prefix relation}. Therefore
+    $\upset[\prefleq]{w} \cap L$ is a \kl{chain} for the \kl{prefix relation}.
+    In particular, letting $S_{\max}$ be the set of all maximal elements
+    of $S$,
+    we conclude that 
+    \begin{equation*}
+        L \subseteq S \cup \bigcup_{w \in S_{\max}} \upset[\prefleq]{w} \cap L
+        \quad .
+    \end{equation*}
+    Hence, that $L$ is finite union of \kl{chains}.
 \end{proof}
@@
 \begin{corollary}
-    \textbf{false: some chains can be finite}
-    Let $L \subseteq \Sigma^*$ be a language that is \kl{well-quasi-ordered} by
-    the \kl{prefix relation}. Then, there exists a finite $k \in \Nat$ such that
+    Let $L \subseteq \Sigma^*$ be an infinite language that is \kl{well-quasi-ordered} by
+    the \kl{prefix relation}. Then, there exists finite numbers $k,\ell \in \Nat$ such that
     the
@@
     the \kl{ordinal height} of $L$ is $\omega$, and its
-    \kl{ordinal width} is $k$.
+    \kl{ordinal width} is $\ell$.
 \end{corollary}
+\begin{proof}
+    \textbf{todo.}
+\end{proof}
 

### 2024-08-09 src/infixes.tex
proving the regular language case

@@
     We also let $L_0 \defined a b^* a$, and 
-    $L_{n+1} \defined L_n \cup c^{n+1} a^{n+1} b^* a^{n+1} c^{n+1} (L_n c^{n+1})^*$.
+    $L_{n+1} \defined L_n \cup c^{n+1} b^* c^{n+1} (L_n c^{n+1})^*$.
     Now, let us inductively define \kl{embeddings} $f_n \colon Y_n \to
@@
         f_{n+1}(x_i) \defined 
-        \underbrace{c^{n+1} a^{n+1} b^i a^{n+1} c^{n+1} \prod_{y \in D_i} \overbrace{f_n(y)}^{\in L_n} c^{n+1}}_{\in L_{n+1}}
+        \underbrace{c^{n+1} b^i c^{n+1} \prod_{y \in D_i} \overbrace{f_n(y)}^{\in L_n} c^{n+1}}_{\in L_{n+1}}
         \quad .
@@
     $f_n(y) = f_{n+1}(y)$, we conclude that $f_{n+1}(y) \infleq f_{n+1}(x_i)$.
-    Conversely, assume that
-    $f_{n+1}(y) \infleq f_{n+1}(x_i)$. Again, this means
-    that $f_n(y) \infleq f_{n+1}(x_i)$. Now, the encoding
-    is robust enough so that we can conclude
-    $f_{n}(y) \infleq f_{n}(z)$ for some $z \in D_i$.
-    By induction hypothesis, $y \preceq z$, hence, $y \preceq x_i$.
+    Conversely, assume that $f_{n+1}(y) \infleq f_{n+1}(x_i)$. Again, this
+    means that $f_n(y) \infleq f_{n+1}(x_i)$. Now, the encoding is robust
+    enough so that we can conclude $f_{n}(y) \infleq f_{n}(z)$ for some $z \in
+    D_i$ By induction hypothesis, $y \preceq z$, hence, $y \preceq x_i$.
 
@@
 As a consequence of \cref{infix-embedding:thm}, we cannot replay proofs of
-\cref{prefixes:sec}, and will actually need to leverage some
-regularity of the languages to obtain a characterization of well-quasi-ordered
-languages under the infix relation. Let us first play this game for languages
-that are recognized by finite automata.
+\cref{prefixes:sec}, and will actually need to leverage some regularity of the
+languages to obtain a characterization of \kl{well-quasi-ordered} languages
+under the \kl{infix relation}. Let us first play this game for languages that
+are recognized by finite automata. We assume that the reader is familiar with
+the notions of deterministic finite automaton and regular languages, and refer
+to the book of Thomas for a comprehensive introduction to the subject
+\cite{THOM97}. The main goal of the remainder of this section is to prove the
+following \cref{infix-finite-automata:thm}.
 
-\textbf{TODO: introduce finite automata and refer to the book.}.
-
-\begin{theorem}
-    \label{infix-finite-automata:thm}
+\begin{theorem}[restate=infix-finite-automata:thm,label=infix-finite-automata:thm]
     Let $L \subseteq \Sigma^*$ be a language recognized by a finite automaton.
@@
 
-In order to prove \cref{infix-finite-automata:thm}, we will perform
-some preliminary analysis on the structure of an automaton recognizing a
-well-quasi-ordered language under the infix relation.
+\AP In order to prove \cref{infix-finite-automata:thm}, we will perform some
+preliminary analysis on the structure of an automaton recognizing a
+well-quasi-ordered language under the infix relation, which will be powered by
+folklore results on \emph{periodic} words. Let us recall that a non-empty word
+$w \in \Sigma^+$ is \intro(word){periodic} with period $x \in \Sigma^*$ if
+there exists a $p \in \Nat$ such that $w \infleq x^p$. The \intro{periodic
+length} of a word $u$ is the minimal length of a word $x$ such that $u$ is an
+\kl{infix} of $x^p$ for some $p \in \Nat$ and $x \in \Sigma^+$. We will
+essentially rely on the following result on periodic words.
 
 \begin{lemma}
-    Let $A = (Q, \Sigma, \delta, q_0, F)$ be a finite automaton recognizing a
-    language $L \subseteq \Sigma^*$. Assume furthermore that
-    $L$ is \kl{well-quasi-ordered} by the \kl{infix relation}.
-    Then, for all words $u,w,v \in \Sigma^*$, and states
-    $q, q_f \in Q \times F$ such that
-    $q_0 \xrightarrow{u} q \xrightarrow{w} q \xrightarrow{v} q_f$,
-    one of the following holds:
-    \begin{itemize}
-        \item There exists $k \in \Nat$
-            such that $u$ is a \kl{suffix} of $w^k$,
-        \item There exists $k \in \Nat$
-            such that $v$ is a \kl{prefix} of $w^k$.
-    \end{itemize}
+    \label{periodic-infixes:lem}
+    Let $u,v \in \Sigma^*$ be two (non-empty) \kl{periodic words}
+    having \kl{periodic lengths} $p$ and $q$ respectively.
+    Then, if $u \infleq v$ and $\card{u} \geq \factorial[p]{p \times q}$,
+    then $u$ and $v$ share the same \kl{periodic length}
+    $p = q$.
 \end{lemma}
+\begin{proof}
+    The fact that $u$ and $v$ are \kl{periodic length}
+    respectively $p$ and $q$ translates into the fact that $u_{i+p} = u_i$ and
+    $v_{i+q} = v_i$ for all indices $i \in \Nat$ such that those letters are
+    well-defined.
+
+    Now, assume that $u$ is an \kl{infix} of $v$, this provides the existence
+    of a $k \in \Nat$ such that $u = v_{k} \cdots v_{k + \card{u} - 1}$. In
+    particular, $v_{k+i+p} = v_{k+i}$ for all $i \in \Nat$ such that $k+i+p < k
+    + \card{u}$. Since we also have $v_{k+i+q} = v_{k+i}$ for all $1 \leq i
+    \leq \card{v} - k - q$. We conclude that both $u$ and $v$ are of
+    \kl{periodic length} the greatest common divisor of $p$ and $q$, and by
+    minimality of $q$ this must be equal to $q$ and to $p$.
+\end{proof}
+
+The reason why \kl{periodic words} built using a given period $x \in \Sigma^+$
+are interesting for the \kl{infix relation} is that they naturally create
+\kl{chains}. Indeed, if $x \in \Sigma^+$ is a finite word, then $\setof{x^p}{p
+\in \Nat}$ is a \kl{chain} for the \kl{infix relation}. Note that in general,
+the downwards closure of a chain is \emph{not} a chain. However, for the chains
+generated using periodic words, the downwards closure
+$\dwset[\infleq]{\setof{x^p}{p \in \Nat}}$ is a \emph{finite union} of
+\kl{chains}. Because this set will appear in bigger equations,
+we introduce the shorter notation 
+$\intro*\InfPeriodChain{x}$ for the set of 
+\kl{infixes} of words of the form $x^p$, where $p$ ranges over $\Nat$.
 
 \begin{lemma}
-    Let $A = (Q, \Sigma, \delta, q_0, F)$ be a finite automaton recognizing a
-    language $L \subseteq \Sigma^*$. Assume furthermore that
-    $L$ is \kl{well-quasi-ordered} by the \kl{infix relation}.
-    Then, for all words $w_1, w_2 \in \Sigma^*$, and states
-    $q, q_f \in Q \times F$ such that
-    $q_0 \rightarrow^* q \xrightarrow{w} q \rightarrow^* q_f$
-    and 
-    $q_0 \rightarrow^* q \xrightarrow{w_2} q \rightarrow^* q_f$,
-    there exists a word $w \in \Sigma^*$
-    such that $w_1$ and $w_2$ are powers of $w$.
+    \label{inf-period-chain:lem}
+    Let $x \in \Sigma^+$ be a word, and
+    Then $\InfPeriodChain{x}$ is a finite union of \kl{chains}
+    for the \kl{infix}, \kl{prefix} and \kl{suffix} relations 
+    simultaneously.
 \end{lemma}
+\begin{proof}
+    Let $x \in \Sigma^+$ be a word, and let $P_x$ be the (finite) set 
+    of all \kl{prefixes} of $x$, and $S_x$ be the (finite)
+    set of all \kl{suffixes} of $x$.
+    Assume that $w \in \InfPeriodChain{x}$, then $w = u x^p v$ for some
+    $u \in S_x$, $v \in P_x$, and $p \in \Nat$.
+    We have proven that
+    \begin{equation*}
+        \InfPeriodChain{x} \subseteq \bigcup_{u \in P_x} \bigcup_{v \in S_x} u x^* v
+        \quad .
+    \end{equation*}
+
+    Let us now demonstrate that for all $(u,v) \in S_x \times P_x$, the
+    language $u x^* v$ is a \kl{chain} for the \kl{infix}, \kl{suffix} and \kl{prefix} relations.
+    To that end,
+    let $(u,v) \in S_x \times P_x$ and $\ell, k \in \Nat$ be such that $\ell <
+    k$, let us prove that $u x^\ell v \infleq u x^k  v$. Because $v \prefleq
+    x$, we know that there exists $w$ such that $vw = x$. In particular,
+    $ux^\ell vw = u x^{\ell + 1}$, and because $\ell < k$, we conclude that $u
+    x^{\ell + 1} \prefleq u x^k v$. By transitivity, $u x^\ell v \prefleq u x^k
+    v$, and \emph{a fortiori}, $u x^\ell v \infleq u x^k v$. 
+    Similarly, because $u \suffleq x$,  there exists $w$ such that $wu  = x$, 
+    and we conclude that $u x^{\ell} v \suffleq w u x^\ell v = x^{\ell + 1} v \suffleq u x^k v$.
+    \qedhere
+\end{proof}
+
+The following combinatorial lemma connects the property of being
+\kl{well-quasi-ordered} to a property of the \kl{periodic lengths} of words in
+a language, based on the assumption that some factors can be iterated.
 
 \begin{lemma}
-    Let $A = (Q, \Sigma, \delta, q_0, F)$ be a finite automaton recognizing a
-    language $L \subseteq \Sigma^*$. Assume furthermore that
-    $L$ is \kl{well-quasi-ordered} by the \kl{infix relation}.
-    Then, for all words $u, w_1, w_2 \in \Sigma^*$, and states
-    $q_1, q_2, q_f \in Q^2 \times F$,
+    \label{pumping-periods:lem}
+    Let $L \subseteq \Sigma^*$ be a language
+    that is \kl{well-quasi-ordered} by the \kl{infix relation}.
+    Let $k \in \Nat$, $u_1, \cdots, u_{k+1} \in \Sigma^*$,
+    and $v_1, \cdots, v_{k} \in \Sigma^+$
+    be such that
+    $w[\vec{X}] \defined (\prod_{i = 1}^k u_k v_k^{X_k}) u_{k+1}$
+    belongs to $L$
+    for arbitrarily large values of $\vec{X} \in \Nat^k$.
+    Then, 
+    there exists $x,y \in \Sigma^+$ of size 
+    at most $\max \setof{\card{v_i}}{1 \leq i \leq k}$
     such that 
-    $q_0 \rightarrow^* q_1 \xrightarrow{w_1} q_1 \xrightarrow{u} q_2 \xrightarrow{w_2} q_2
-    \rightarrow^* q_f$,
-    there exists a $w \in \Sigma^*$ such that $w_1$, $w_2$, and $u$ are powers of $w$.
+    one of the following holds for all
+    $\vec{X} \in \Nat^{k}$:
+    \begin{enumerate}
+        \item $w[\vec{X}] \in u_1 \InfPeriodChain{x}$,
+        \item $w[\vec{X}] \in \InfPeriodChain{x} u_{k+1}$,
+        \item $w[\vec{X}] \in \InfPeriodChain{x} u_i \InfPeriodChain{y}$
+            for some $1 \leq i \leq k + 1$.
+
+    \end{enumerate}
 \end{lemma}
+\begin{proof}
+    Let us construct a sequence of words $\seqof{w_i}[i \in \Nat]$, where $w_i
+    = w[n_i, m_i]$ for some well-chosen indices $n_i, m_i \in \Nat$. The idea
+    being that 
+    if $w[n_i, m_i]$ is an \kl{infix} of $w[n_{j}, m_{j}]$,
+    then one of $u_1$ or $u_3$ must be left untouched by the embedding,
+    and the overlapping parts must be long enough to apply 
+    \cref{periodic-infixes:lem}.
+    In order to achieve this,
+    we define
+    $n_0$ and $m_0$ to be numbers greater than $\factorial[p]{\card{v_1} \times
+    \card{v_2}}$ and such that $w[n_0, m_0]$ belongs to $L$. 
+    Then, we define 
+    $n_{i+1}$ and $m_{i+1}$ to be the smallest numbers greater than 
+    $2 \times \card{w[n_i,m_i]}$ such that $w[n_{i+1}, m_{i+1}]$ belongs to $L$.
 
-We are now ready to prove the main result.
+    Because $L$ is \kl{well-quasi-ordered} by the \kl{infix relation}, there
+    exists $i < j$ such that $w[n_i, m_i]$ is an \kl{infix} of $w[n_j, m_j]$.
+    \textbf{TODO:} assume that we have a large intersection of $v_1$ and $v_2$,
+    then we are in the latter cases. Otherwise, we are in the first case.
+\end{proof}
 
-\begin{proofof}{infix-finite-automata:thm}
-    Let $w \in L$. Consider a run of the automaton on $w$. If $w$ is too long,
-    then there exists a loop, and we can prove that $w = u x^p v$ for some
-    short period $x$. By repeating the process for $u$ and $v$, we conclude
-    that $w$ has the form $w = u_0 x_0^{p_0} u_1 x_1^{p_1} \cdots u_k x_k^{p_k}
-    v_k$. Now, applying the loop merging lemma, we conclude that $x_0 = x_1 =
-    \cdots = x_k$. Hence, $w = u_0 x^{p} v_k$. Finally, we know that either
-    $u_0$ is a prefix of $x^{p}$ or that $v_k$ is a suffix of $x^{p}$. We
-    conclude in particular that $\dwset[\infleq]{ \setof{ u x^p v }{ p \in \Nat
-    }}$ is a finite union of \kl{chains}. Because $u,v,x$ are of bounded
-    length, we conclude that $L$ itself included in a finite union of
-    \kl{chains}, hence is a finite union of \kl{chains}.
+\begin{proofof}{infix-finite-automata:thm}[main]
+    Let $w \in L$, because $Q$ is finite, there exists
+    a factorization of $w$
+    into words $(\prod_{i = 1}^k u_i v_i) u_{k+1}$
+    such that 
+    for all $1 \leq i \leq k+1$, $\card{u_i} \leq \card{Q}$,
+    for all $1 \leq i \leq k$, $1 \leq \card{v_i} \leq \card{Q}$,
+    and satisfying 
+    that $w[\vec{X}] \defined 
+    (\prod_{i = 1}^k u_i v_i^{X_i}) u_{k+1}$
+    belongs to $L$ for all choices of values $\vec{X} \in \Nat^k$.
+
+    Applying \cref{pumping-periods:lem}, we conclude that 
+    there exists $x,y \in \Sigma^+$ of size at most $\card{Q}$
+    such that 
+    $w \in u_1 \InfPeriodChain{x} \cup \InfPeriodChain{y} u_{k+1}
+    \cup \bigcup_{1 \leq i \leq k+1} \InfPeriodChain{x} u_i \InfPeriodChain{y}$.
+    In particular, we conclude that
+    \begin{equation}
+        \label{infix-automata:eq}
+        L
+        \subseteq
+        \bigcup_{x,y,u\in \Sigma^{\leq \card{Q}}}
+        u \InfPeriodChain{x}
+        \cup 
+        \InfPeriodChain{x} u
+        \cup
+        \InfPeriodChain{x} u \InfPeriodChain{y}
+        \quad .
+    \end{equation}
+
+    Now, thanks to \cref{inf-period-chain:lem}
+    this happens to be a finite union of \kl{chains}
+    for the \kl{infix relation}.
 \end{proofof}
@@
 \begin{corollary}
-    Decidability?
+    Given a regular language $L$, it is decidable whether
+    $L$ is \kl{well-quasi-ordered} for the \kl{infix relation}.
 \end{corollary}
+\begin{proof}
+    It suffices to decide whether the equation 
+    \cref{infix-automata:eq} holds for the language $L$.
+    This is an inclusion of regular languages, which is decidable.
+\end{proof}
 
@@
 
-This section will use so-called \emph{amalgamation systems}.
+When proving \cref{infix-finite-automata:thm}, we have leveraged a powerful
+combinatorial argument from \cref{pumping-periods:lem}, which made the proof
+almost trivial. It turns out that there is a rather large family of systems for
+which pumping arguments based on so-called \emph{minimal runs} exist: they are
+called \emph{amalgamation systems}. Having done the heavy lifting on finite
+automata, the rest of this section is mostly devoted to the introduction of
+amalgamation systems and collecting the necessary pumping argument they enjoy.
+Using this meta-proof, we will generalize \cref{infix-finite-automata:thm} to
+context-free grammars, languages recognized by VASS, and more.
 

### 2024-08-09 src/infixes.tex
fixing proofs

@@
 
+\begin{corollary}
+    \label{inf-period-union-chains:lem}
+    Let $x,u,y \in \Sigma^*$ be words.  The following 
+    are finite unions of \kl{chains} for the infix relation:
+    $\InfPeriodChain{x}u$, $u \InfPeriodChain{x}$,
+    and $\InfPeriodChain{x} u \InfPeriodChain{y}$.
+\end{corollary}
+\begin{proof}
+    Because $\InfPeriodChain{x}$ is a finite union of \kl{chains} for the \kl{suffix}
+    relation (using \cref{inf-period-chain:lem}), we conclude that $\InfPeriodChain{x}u$ is one too for
+    the \kl{infix} relation. This holds similarly for $u \InfPeriodChain{x}$.
+    In the case of $\InfPeriodChain{x} u \InfPeriodChain{y}$,
+    we remark that $\InfPeriodChain{x} u$ is a finite union of chains
+    for the \kl{suffix relation},
+    and that $\InfPeriodChain{y}$ is one for the \kl{prefix relation}
+    (using again \cref{inf-period-chain:lem}).
+    As a consequence,
+    $\InfPeriodChain{x}u \InfPeriodChain{y}$ is a finite union 
+    of \kl{chains} for the \kl{infix} relation.
+\end{proof}
+
 The following combinatorial lemma connects the property of being
@@
     be such that
-    $w[\vec{X}] \defined (\prod_{i = 1}^k u_k v_k^{X_k}) u_{k+1}$
+    $w[\vec{n}] \defined (\prod_{i = 1}^k u_i v_i^{n_i}) u_{k+1}$
     belongs to $L$
-    for arbitrarily large values of $\vec{X} \in \Nat^k$.
+    for arbitrarily large values of $\vec{n} \in \Nat^k$.
     Then, 
@@
     one of the following holds for all
-    $\vec{X} \in \Nat^{k}$:
+    $\vec{n} \in \Nat^{k}$:
     \begin{enumerate}
-        \item $w[\vec{X}] \in u_1 \InfPeriodChain{x}$,
-        \item $w[\vec{X}] \in \InfPeriodChain{x} u_{k+1}$,
-        \item $w[\vec{X}] \in \InfPeriodChain{x} u_i \InfPeriodChain{y}$
+        \item $w[\vec{n}] \in u_1 \InfPeriodChain{x}$,
+        \item $w[\vec{n}] \in \InfPeriodChain{x} u_{k+1}$,
+        \item $w[\vec{n}] \in \InfPeriodChain{x} u_i \InfPeriodChain{y}$
             for some $1 \leq i \leq k + 1$.
@@
 \begin{proof}
+    Note that the result is obvious if $k = 0$, and therefore
+    we assume $k \geq 1$ in the following proof.
+
     Let us construct a sequence of words $\seqof{w_i}[i \in \Nat]$, where $w_i
-    = w[n_i, m_i]$ for some well-chosen indices $n_i, m_i \in \Nat$. The idea
+    \defined w[\vec{n_i}]$ for some well-chosen indices $\vec{n_i} \in \Nat^k$. The goal
     being that 
-    if $w[n_i, m_i]$ is an \kl{infix} of $w[n_{j}, m_{j}]$,
-    then one of $u_1$ or $u_3$ must be left untouched by the embedding,
-    and the overlapping parts must be long enough to apply 
+    if $w[\vec{n_i}]$ is an \kl{infix} of $w[\vec{n_j}]$,
+    then it can intersect at most \emph{two} iterated words,
+    with an intersection that is long enough to successfully apply
     \cref{periodic-infixes:lem}.
     In order to achieve this,
-    we define
-    $n_0$ and $m_0$ to be numbers greater than $\factorial[p]{\card{v_1} \times
-    \card{v_2}}$ and such that $w[n_0, m_0]$ belongs to $L$. 
-    Then, we define 
-    $n_{i+1}$ and $m_{i+1}$ to be the smallest numbers greater than 
-    $2 \times \card{w[n_i,m_i]}$ such that $w[n_{i+1}, m_{i+1}]$ belongs to $L$.
+    let us first define $s$ as the maximal size of a word $v_i$
+    ($1 \leq i \leq k$) and $u_j$ ($1 \leq j \leq k+1$).
+    Then,
+    we consider $\vec{n_0} \in \Nat^k$ such that $\vec{n_0}$ has all 
+    its components greater than $\factorial{s}$ and such that
+    $w[\vec{n_0}]$ belongs to $L$. 
+    Then, we inductively define 
+    $\vec{n_{i+1}}$  as the smallest vector of numbers greater than $\vec{n_i}$,
+    such that $w[\vec{n_{i+1}}]$ belongs to $L$, 
+    and with $\vec{n_i}$ having all components greater than
+    $2\card{w[\vec{n_i}]}$.
 
+    Let us assume that $k \geq 2$ in the following proof for symmetry purposes,
+    and argue later on that when $k = 1$ the same argument goes through.
     Because $L$ is \kl{well-quasi-ordered} by the \kl{infix relation}, there
-    exists $i < j$ such that $w[n_i, m_i]$ is an \kl{infix} of $w[n_j, m_j]$.
-    \textbf{TODO:} assume that we have a large intersection of $v_1$ and $v_2$,
-    then we are in the latter cases. Otherwise, we are in the first case.
+    exists $i < j$ such that $w[\vec{n_i}]$ is an \kl{infix} of $w[\vec{n_j}]$.
+    Now, because of the values of $\vec{n_j}$, there exists $1 \leq \ell \leq
+    k-1$ such that $w[\vec{n_i}]$ is actually an \kl{infix} of $u_{\ell}
+    v_{\ell}^{n_{j,\ell}} u_{\ell+1} v_{\ell+1}^{n_{j,\ell+1}} u_{\ell+2}$.
+    Even more,
+    one of the three following equations holds:
+    \begin{itemize}
+        \item $w[\vec{n_i}] \infleq v_{\ell}^{n_{j,\ell}} u_{\ell+1} v_{\ell+1}^{n_{j,\ell+1}}$,
+        \item $w[\vec{n_i}] \infleq u_{\ell}
+            v_{\ell}^{n_{j,\ell}}$,
+        \item $w[\vec{n_i}] \infleq
+            v_{\ell+1}^{n_{j,\ell+1}} u_{\ell+2}$.
+    \end{itemize}
+    In all those cases, we conclude using \cref{periodic-infixes:lem}.
+
+    When $k = 1$, the situation is a bit more specific since we only have two
+    cases: either $w_i \infleq u_1 v_1^{n_j}$ or $w_i \infleq v_1^{n_j} u_2$,
+    and we conclude with an identical reasoning.
 \end{proof}
+
+We are now ready to restate and proof our main theorem.
 
@@
 
-    Now, thanks to \cref{inf-period-chain:lem}
+    Now, thanks to \cref{inf-period-union-chains:lem},
     this happens to be a finite union of \kl{chains}

### 2024-08-09 src/conclusion.tex
starting to add amalgamation systems

@@
 
-\subparagraph{Towards infinite alphabets} Can we go to
-alphabets that are infinite but equipped with a well-quasi order?
+We provided concretes statements that justify why the \kl{subword relation}
+is used when defining \kl{well-quasi-orders} on finite words. Even if
+\kl{prefix}, \kl{suffix} or \kl{infix} relations are meaningful, they are
+\kl{well-quasi-ordered} if and only if they behave similarly to disjoint copies
+of $\Nat$ or $\Nat^2$. However, our approach suffers some limitations 
+and opens the road to natural continuation of this line of work.
 
-\subparagraph{Monoid equations}
+\subparagraph{Towards infinite alphabets} In this paper, we restricted our
+attention to \emph{finite} alphabets, having in mind the application to regular
+languages. However, the conclusions of \cref{infix-finite-automata:thm} and
+\cref{prefixes:thm} could be conjecture to hold in the case of infinite
+alphabets (themselves equipped with a \kl{well-quasi-ordering}). This would
+require new techniques, as the finiteness of the alphabet is crucial to all of
+our positive results.
+
+\subparagraph{Monoid equations}  It could be interesting to understand which
+monoids $M$ recognize languages that are \kl{well-quasi-ordered} by the
+\kl{infix}, \kl{prefix} or \kl{suffix} relations. This is connected to finding
+which classes of graphs of \emph{bounded clique-width} are
+\kl{well-quasi-ordered} with respect to the \emph{induced subgraph relation}, a
+subject that is recently gaining traction.

### 2024-08-09 src/infixes.tex
starting to add amalgamation systems

@@
 
-\begin{theorem}
-    \label{infix-amalgamation:thm}
+\begin{theorem}[label=infix-amalgamation:thm,restate=infix-amalgamation:thm]
     Let $L \subseteq \Sigma^*$ be a language recognized by an 
@@
 
-\begin{itemize}
-    \item Define amalgamation systems
-    \item Pumping argument
-    \item Characterization
-    \item Decision procedure
-\end{itemize}
+\begin{definition}
+    \textbf{TODO}. An \intro{amalgamation system}
+    is a triple such that.
+\end{definition}
+
+Let us recall some examples of \kl{amalgamation systems}
+arising in the literature.
+
+\begin{example}
+    Regular languages, context-free languages, and languages recognized by
+    VASS are all recognized by amalgamation systems.
+\end{example}
+
+\begin{lemma}
+    Let $L$ be a language recognized by an \kl{amalgamation system}
+    and assume that $L$ is \kl{well-quasi-ordered}
+    by the \kl{infix relation}. Then,
+    for every minimal run $\rho$,
+    the \kl{gap languages} of $\rho$
+    are \kl{subsemigroup languages}.
+\end{lemma}
+
+\begin{proofof}{infix-amalgamation:thm}
+    TODO.
+\end{proofof}
+
+\begin{lemma}
+    Let $L$ be given by an \kl{effective amalgamation system},
+    then it is decidable whether $L$ 
+    is \kl{well-quasi-ordered} by the \kl{infix relation}.
+\end{lemma}

### 2024-08-09 src/introduction.tex
starting to add amalgamation systems

@@
 \cite{HIG52}. Let us recall that the \kl{subword relation} for words over $(X,
-\preceq)$ is defined as follows: a word $u$ is a \intro{subword} of a word $v$
+\preceq)$ is defined as follows: a word $u$ is a \intro{subword} of a word $v$,
+written $u \intro*\higleq v$,
 if there exists an increasing function $f \colon \{1, \ldots, |u|\} \to \{1,
@@
 However, there are many other natural orderings on words that could be
-considered in the context of \kl{well-quasi-orderings}. The three alternatives
-we consider are the \kl{prefix relation}, the \kl{suffix relation}, and the
-\kl{infix relation}. Unfortunately, it is an easy check that all of these
-constructions are not well-quasi-ordered as soon as the alphabet contains two
-non-equivalent letters. This is not the end of the road, because some
-\emph{subsets} of $\Sigma^*$ might still be \kl{well-quasi-ordered} by these
-relations, the simplest example being the case of finite sets of (finite) words
-which are all \kl{well-quasi-ordered} regardless of the ordering considered.
-This raises the question of characterizing the subsets $L \subseteq \Sigma^*$
-that are \kl{well-quasi-ordered} with respect to the \kl{prefix relation}
-(respectively, the \kl{suffix relation} or the \kl{infix relation}), and
-providing suitable decision procedures.
+considered in the context of \kl{well-quasi-orderings}. In our, case we
+consider alphabets $\Sigma$ that are finite and equipped with the equality
+relation. In this setting, the three alternatives we consider are the
+\intro{prefix relation} --- $u \intro*\prefleq v$ if there exists $w$ with $uw
+= v$ ---, the \intro{suffix relation} --- $u \intro*\suffleq v$ if there exists
+$w$ such that $wu = v$ ---, and the \intro{infix relation} --- $u
+\intro*\infleq v$ if there exists $w_1,w_2$ such that $w_1 u w_2 = v$ ---.
+Unfortunately, it is an easy check that all of these constructions are not
+well-quasi-ordered as soon as the alphabet contains two non-equivalent letters.
+This is not the end of the road, because some \emph{subsets} of $\Sigma^*$
+might still be \kl{well-quasi-ordered} by these relations, the simplest example
+being the case of finite sets of (finite) words which are all
+\kl{well-quasi-ordered} regardless of the ordering considered. This raises the
+question of characterizing the subsets $L \subseteq \Sigma^*$ that are
+\kl{well-quasi-ordered} with respect to the \kl{prefix relation} (respectively,
+the \kl{suffix relation} or the \kl{infix relation}), and providing suitable
+decision procedures.
 

### 2024-08-09 src/prefixes.tex
starting to add amalgamation systems

@@
 In this section, we study the well-quasi-ordering of languages under the prefix
-relation. A word $u$ is a \intro{prefix} of a word $w$ if there exists a word
-$v$ such that $w = uv$. We denote this relation by $u \intro*\prefleq w$.
-Similarly, a word $u$ is a \intro{suffix} of a word $w$ if there exists a word
-$v$ such that $w = vu$. We denote this relation by $u \intro*\suffleq w$. Let
-us immediately remark that the map $u \mapsto u^R$ that reverses a word is an
-order-bijection between $(X^*, \prefleq)$ and $(X^*, \suffleq)$. Therefore, we
-will focus on the prefix relation in the rest of this section.
+relation. Let us immediately remark that the map $u \mapsto u^R$ that reverses
+a word is an order-bijection between $(X^*, \prefleq)$ and $(X^*, \suffleq)$.
+Therefore, we will focus on the prefix relation in the rest of this section.
 
-Let us briefly restate the fact that some (even regular) languages 
-are not well-quasi-ordered by the prefix relation.
+Let us briefly restate the fact that some (even regular) languages are not
+well-quasi-ordered by the prefix relation.
 
@@
 \begin{corollary}
+    \label{prefix-wqo-reg-decidable:cor}
     If $L$ is regular, then the existence of an infinite antichain is decidable.
@@
 \end{corollary}
-\begin{proof}
-    \textbf{todo.}
-\end{proof}
 
@@
 whether a decidable language is \kl{well-quasi-ordered} by the \kl{prefix
-relation}.
+relation}. This is a very easy result whose sole purpose is to contrast with
+the decidability result of \cref{prefix-wqo-reg-decidable:cor}.
 
 \begin{lemma}
-    todo.
+    The following problem is undecidable: given a language $L$
+    decided by a Turing machine, answer whether 
+    $L$ is \kl{well-quasi-ordered} for the \kl{prefix relation}.
 \end{lemma}
+\begin{proof}
+    We reduce the halting problem on the empty string $\varepsilon$.
+    Let $M$ be a Turing Machine, we write the languages $L$ of finite runs
+    of $M$ starting on the empty string,
+    that we surround by special markers. This language is decidable,
+    and 
+    is
+    \kl{well-quasi-ordered} if and only if it is finite
+    if and only if $M$ terminates on $\varepsilon$.
+\end{proof}

### 2024-08-11 src/infixes.tex
amalgamation?

@@
 the \kl{infix relation} can yield to very complicated \kl{well-quasi-ordered}
-languages. Formally, the following theorem shows that \emph{any} countable
+languages. Formally, the upcoming \cref{infix-embedding:thm} shows that \emph{any} countable
 quasi-ordering with finite initial segments can be embedded into the infix
@@
 
+\AP
 Let us recall that an \intro{order embedding} from a quasi-ordered set $(X,
@@
 into} $Y$.
-
 We say that a quasi-ordered set $(X, \preceq)$ is a \intro{partial ordering}
@@
             $X$ is countable, and for every $x \in X$,
-        $\setof{y \in X}{y \preceq x}$ is finite.
+            its \kl{downwards closure}
+            $\dwset[\preceq]{x}$ is finite.
     \end{enumerate}
-    Furthermore, if $X$ is recursively enumerable and $\preceq$ decidable,
-    then
-    the \kl{embedding} is computable.
 \end{theorem}
@@
     ordered}. In particular, because $\dwset[\infleq]{w}$ is finite
-    for every $w \in \Sigma^*$, then so is $\setof{y \in X}{y \preceq x}$
+    for every $w \in \Sigma^*$, then so is $\dwset[\preceq]{x}$
     for every $x \in X$. Similarly, $X$ must be countable
@@
 \in \Nat}$ is a \kl{chain} for the \kl{infix relation}. Note that in general,
-the downwards closure of a chain is \emph{not} a chain. However, for the chains
-generated using periodic words, the downwards closure
+the \kl{downwards closure} of a chain is \emph{not} a chain. However, for the chains
+generated using periodic words, the \kl{downwards closure}
 $\dwset[\infleq]{\setof{x^p}{p \in \Nat}}$ is a \emph{finite union} of
@@
 
-\section{Infixes and Amalgamation systems}
+\section{Infixes and Amalgamation Systems}
 \label{infixes-amalgamation:sec}
@@
 When proving \cref{infix-finite-automata:thm}, we have leveraged a powerful
-combinatorial argument from \cref{pumping-periods:lem}, which made the proof
-almost trivial. It turns out that there is a rather large family of systems for
-which pumping arguments based on so-called \emph{minimal runs} exist: they are
-called \emph{amalgamation systems}. Having done the heavy lifting on finite
-automata, the rest of this section is mostly devoted to the introduction of
-amalgamation systems and collecting the necessary pumping argument they enjoy.
-Using this meta-proof, we will generalize \cref{infix-finite-automata:thm} to
-context-free grammars, languages recognized by VASS, and more.
+combinatorial argument from \cref{pumping-periods:lem}. It turns out that there
+is a rather large family of systems for which pumping arguments based on
+so-called \emph{minimal runs} exist: they are called \emph{amalgamation
+systems} and were recently introduced by Anad, Schmitz, Sch\"{u}tze, and
+Zetzsche \cite{ASZZ24}. Having done the heavy lifting on finite automata, the
+rest of this section is mostly devoted to the introduction of \kl{amalgamation
+systems} and collecting the necessary pumping argument they enjoy. Using this
+meta-proof, we will generalize \cref{infix-finite-automata:thm} to context-free
+grammars, languages recognized by vector addition systems with state (VASS),
+and more. The goal of this section is not to introduce all of these systems, or
+justify their usefulness, but to state and prove the following theorem.
 
@@
 
-Let us recall some examples of \kl{amalgamation systems}
-arising in the literature.
-
-\begin{example}
-    Regular languages, context-free languages, and languages recognized by
-    VASS are all recognized by amalgamation systems.
-\end{example}
+Let us recall some examples of \kl{amalgamation systems} given by the authors
+of the original paper: regular languages \cite[Theorem 5.3]{ASZZ24}, VASS as a
+consequence of \cite[Theorem 5.5]{ASZZ24}, context-free languages as a
+consequence of \cite[Theorem 5.10]{ASZZ24}.
 

### 2024-08-11 src/prefixes.tex
amalgamation?

@@
 
-In this section, we study the well-quasi-ordering of languages under the prefix
-relation. Let us immediately remark that the map $u \mapsto u^R$ that reverses
-a word is an order-bijection between $(X^*, \prefleq)$ and $(X^*, \suffleq)$.
-Therefore, we will focus on the prefix relation in the rest of this section.
+In this section, we study the well-quasi-ordering of languages under the
+\kl{prefix relation}. Let us immediately remark that the map $u \mapsto u^R$
+that reverses a word is an order-bijection between $(X^*, \prefleq)$ and $(X^*,
+\suffleq)$, that is, $u \prefleq v$ if and only if $u^R \suffleq v^R$.
+Therefore, we will focus on the prefix relation in the rest of this section, as
+$(L, \prefleq)$ is \kl{well-quasi-ordered} if and only if $(L^R, \suffleq)$ is.
 
-Let us briefly restate the fact that some (even regular) languages are not
-well-quasi-ordered by the prefix relation.
+The next remark that we make is that $\Sigma^*$ is not \kl{well-quasi-ordered}
+by the \kl{prefix relation} as soon as $\Sigma$ contains two distinct letters
+$a$ and $b$. As an example of infinite \kl{antichain}, we can consider the set
+of words $a^n b$ for $n \in \Nat$. As mentioned in the introduction, there are
+some languages however that are \kl{well-quasi-ordered} by the \kl{prefix
+relation}. A simple example being the (regular) language $a^* \subseteq
+\set{a,b}^*$, which is order-isomorphic to natural numbers with their usual
+orderings $(\Nat, \leq)$.
 
-\begin{example}
-    The set $L = \setof{a^nb}{n \in \Nat} \subseteq \set{a,b}^*$ is an infinite \kl{antichain} for the
-    \kl{prefix relation}.
-\end{example}
-
-In order to characterize the existence of infinite antichains for the prefix
-relation, we will introduce the following tree construction that
+In order to characterize the existence of infinite \kl{antichains} for the
+\kl{prefix relation}, we will introduce the following tree construction that
 will be useful in the rest of this section.
@@
 
-Because the alphabet $\Sigma$ is assumed to be finite, the \kl{tree of
-prefixes} is finitely branching. Let us now observe how \kl{antichains} in the
-prefix relation can be witnessed by infinite branches in the \kl{tree of
-prefixes}.
+We will use this \kl{tree of prefixes} to find simple witnesses
+of the existence of infinite \kl{antichains} in the \kl{prefix relation}
+for a given language $L$, namely by introducing \kl{antichain branches}.
 
@@
 
-It is clear that an \kl{antichain branch} for a language $L$ yields an infinite
-antichain, and the converse is quite easy to prove. Because the notion of
-antichain branch is \kl{$\MSO$-definable} whenever $L$ is regular, we
-immediately obtain decidability as a corollary of this lemma.
+Let us illustrate the notion of \kl{antichain branch} over the alphabet $\Sigma
+= \set{a,b}$, and the language $L = a^* b$. In this case, the set $a^*$ (which
+is a branch of the \kl{tree of prefixes}) is an \kl{antichain branch} for $L$.
+This holds because for any $a^k$ in the branch, $a \prefleq a^kb$ which is in
+$L$ and not in the branch. In general, the existence of an \kl{antichain
+branch} for a language $L$ implies that $L$ contains an infinite
+\kl{antichain}, and because the alphabet $\Sigma$ is assumed to be finite, one
+can leverage the fact that the \kl{tree of prefixes} is finitely branching to
+prove that the converse holds as well.
 
 \begin{lemma}
+    \label{antichain-branches-prefix:lem}
     Let $L \subseteq \Sigma^*$ be a language. Then, $L$ contains an infinite
@@
 
+One immediate application of \cref{antichain-branches-prefix:lem} is that
+\kl{antichain branches} can be described inside the \kl{tree of prefixes} by a
+monadic second order formula (\kl{$\MSO$-formula}), allowing us to leverage the
+decidability of $\MSO$ over infinite binary trees \cite[Theorem 1.1]{RAB69}.
+
 \begin{corollary}
@@
     Because the above formula is an $\MSO$-formula over the infinite
-    binary tree, whether it is satisfied is decidable.
-    \textbf{TODO: cite}.
+    $\Sigma$-branching tree, whether it is satisfied is decidable
+    as an easy consequence of the decidability of $\MSO$ over infinite binary
+    trees
+    \cite[Theorem 1.1]{RAB69}.
 \end{proof}
@@
 the \kl{prefix relation}.
+
+For the proof of the following theorem, we will use special notations to
+describe the \intro{upwards closure} of a set $S$ for a relation $\preceq$,
+which is defined as $\upset[\preceq]{S} \defined \setof{y \in \Sigma^*}{
+\exists x \in S. x \preceq y}$. Anticipating the use of the symmetric notion of
+\intro{downwards closure}, let us introduce its notation 
+straight away as follows: $\dwset[\preceq]{S} \defined \setof{y
+\in \Sigma^*}{ \exists x \in S. y \preceq x}$. Abusing notations, we will
+write $\upset{w}$ and $\dwset{w}$ for the upwards and downwards closure of a
+single element $w$, and omit the ordering relation when it is clear from the
+context.
 
@@
     in $S$. By definition, all words in $L$ that contain $w$ as a prefix
-    must be comparable for the \kl{prefix relation}. Therefore
-    $\upset[\prefleq]{w} \cap L$ is a \kl{chain} for the \kl{prefix relation}.
+    must be comparable for the \kl{prefix relation}. Therefore,
+    $(\upset[\prefleq]{w}) \cap L$ is a \kl{chain} for the \kl{prefix relation}.
     In particular, letting $S_{\max}$ be the set of all maximal elements
@@
     \begin{equation*}
-        L \subseteq S \cup \bigcup_{w \in S_{\max}} \upset[\prefleq]{w} \cap L
+        L \subseteq S \cup \bigcup_{w \in S_{\max}} (\upset[\prefleq]{w}) \cap L
         \quad .

### 2024-08-12 src/infixes.tex
trying to add amalgamation

@@
 
-\AP
-Let us recall that an \intro{order embedding} from a quasi-ordered set $(X,
+\AP Let us recall that an \intro{order embedding} from a quasi-ordered set $(X,
 \preceq)$ into a quasi-ordered set $(Y, \preceq')$ is a function $f \colon X
@@
 \preceq' f(y)$. When such an embedding exists, we say that $X$ \reintro{embeds
-into} $Y$.
-We say that a quasi-ordered set $(X, \preceq)$ is a \intro{partial ordering}
-whenever the relation $\preceq$ is antisymmetric, that is $x \preceq y$ and $y
-\preceq x$ implies $x = y$.
+into} $Y$. We say that a quasi-ordered set $(X, \preceq)$ is a \intro{partial
+ordering} whenever the relation $\preceq$ is antisymmetric, that is $x \preceq
+y$ and $y \preceq x$ implies $x = y$.
 
@@
 
+\AP Let us now formally introduce the notion of \kl{amalgamation systems}, and
+recall some results from \cite{ASZZ24} that will be useful for the proof of
+\cref{infix-amalgamation:thm}. Let us recall that over an alphabet $(\Sigma,
+=)$ a \kl{subword embedding} between two words $u \in \Sigma^*$ and $v \in
+\Sigma^*$ is a function $\rho \colon \range{\card{u}} \to \range{\card{v}}$
+such that $u_i = v_{\rho(i)}$ for all $i \in \range{\card{u}}$. We write
+$\intro*\HigEmb(u,v)$ the set of all \kl{subword embeddings} between $u$ and
+$v$. It may be useful to notice that the set of finite words over $\Sigma$
+forms a category when we consider \kl{subword embeddings} as morphisms.
+
+\AP Given a \kl{subword embedding} $f \colon u \to v$ between two words $u$ and
+$v$, there exists a unique decomposition $v = \GapWord{f}{0} u_1 \GapWord{f}{1}
+\cdots \GapWord{f}{k-1} u_k \GapWord{f}{k}$ where $\GapWord{f}{i} =
+v[f(i)+1:f(i+1)-1]$ for all $1 \leq i \leq k-1$, $\GapWord{f}{k} =
+v[f(k)+1:\card{v}]$, and $\GapWord{f}{0}   = v[1: f(1)-1]$.  We say that
+$\intro*\GapWord{f}{i}$ is the $i$-th \intro{gap word} of $f$. Given a run $r
+\in R$, and $i \in \range[0]{\card{\canrun(r)}}$, we define the \intro{gap
+language} of $r$ at position $i$ $\GapLanguage{r}{i} \defined
+\setof{\GapWord{f}{i}}{\exists s \in R. \exists f \in E(r,s) }$.
+
 \begin{definition}
-    \textbf{TODO}. An \intro{amalgamation system}
-    is a triple such that.
+    An \intro{amalgamation system}
+    is a triple $(\Sigma, R, E, \canrun)$ where
+    $\Sigma$ is a finite alphabet,
+    $R$ is a set of so-called \emph{runs},
+    and 
+    $E$ is a set of so-called \emph{run embeddings}.
+    \begin{itemize}
+        \item For all $r, s \in R$,
+            $E(r,s) \subseteq \HigEmb(\canrun(r), \canrun(s))$.
+        \item $(R, E)$ forms a category.
+        \item $(R, \leq_E)$ is a well-quasi-ordered set.
+        \item concatenative amalagamation
+    \end{itemize}
+
+    The \intro{amalgamation language} of such a system
+    is the set of all words $w \in \Sigma^*$ such that
+    there exists a run $r \in R$
+    such that the concatenation of the letters (and possibly the empty word) in
+    $\canrun(r)$ equals $w$.
 \end{definition}
 
-Let us recall some examples of \kl{amalgamation systems} given by the authors
-of the original paper: regular languages \cite[Theorem 5.3]{ASZZ24}, VASS as a
-consequence of \cite[Theorem 5.5]{ASZZ24}, context-free languages as a
-consequence of \cite[Theorem 5.10]{ASZZ24}.
+The notion of \kl{amalgamation system} is tailored to produce \emph{pumping
+arguments}, which is exactly what our \cref{pumping-periods:lem} talks about.
+Because the notion is quite abstract, it may be useful to replicate some
+intuitions developed by the authors of \cite{ASZZ24}.
 
-\begin{lemma}
-    Let $L$ be a language recognized by an \kl{amalgamation system}
-    and assume that $L$ is \kl{well-quasi-ordered}
-    by the \kl{infix relation}. Then,
-    for every minimal run $\rho$,
-    the \kl{gap languages} of $\rho$
-    are \kl{subsemigroup languages}.
-\end{lemma}
+Let us recall some examples of languages that can be recognized by
+\kl{amalgamation systems} given by the authors of the original paper: regular
+languages \cite[Theorem 5.3]{ASZZ24}, VASS as a consequence of \cite[Theorem
+5.5]{ASZZ24}, context-free languages as a consequence of \cite[Theorem
+5.10]{ASZZ24}.
 
 \begin{proofof}{infix-amalgamation:thm}
-    TODO.
+    Assume that $L$ is \kl{well-quasi-ordered} by the \kl{infix relation},
+    and obtained by an \kl{amalgamation system}
+    $(\Sigma, R, E, \canrun)$.
+
+    Let us consider the set $M$ of minimal runs for the relation $\leq_E$,
+    which is finite because the latter is a \kl{well-quasi-ordering}. By
+    definition, for every word $w \in L$, there exists $r \in M$, $s \in R$,
+    and $f \in E(r,s)$ such that $w = \canrun(s)$.
+    In particular, we conclude that
+    \begin{equation*}
+        L \subseteq \bigcup_{r \in M} 
+        \left(
+        \GapLanguage{r}{0}
+        \prod_{i = 1}^{\card{\canrun(r)}-1} \canrun(r)_i \GapLanguage{r}{i} 
+        \right)
+        \quad .
+    \end{equation*}
+
+    Let us first argue that languages $\GapLanguage{r}{i}$ are all \kl{chains}
+    for the \kl{prefix ordering}. According to \cite[Lemma 4.1]{ASZZ24}, either
+    $\GapLanguage{r}{i}$ is a \kl{chain} for the \kl{prefix relation}, or it is
+    an \kl{unbounded language}. Note that \kl{unbounded languages}
+    are \kl{antichains} for the \kl{infix relation}, and therefore
+    \textbf{unclear}.
+
+    Then, we can leverage \cref{pumping-periods:lem}
+    to conclude that $L$ is a finite union of \kl{chains}
+    for the \kl{infix relation}.
 \end{proofof}
 
-\begin{lemma}
+\begin{conjecture}
     Let $L$ be given by an \kl{effective amalgamation system},
@@
     is \kl{well-quasi-ordered} by the \kl{infix relation}.
-\end{lemma}
+\end{conjecture}

### 2024-08-12 src/introduction.tex
amalgamation?
trying to add amalgamation

@@
 
+\AP
 Note that in the case of finite words and finite trees, the precise choice of
@@
 
+\AP
 However, there are many other natural orderings on words that could be
@@
 
-\subparagraph{Contribution} In this paper, we focus on words over a finite
+\subparagraph{Contribution} 
+\AP
+In this paper, we focus on words over a finite
 alphabet $\Sigma$, and characterize subsets $L \subseteq \Sigma^*$ that are

### 2024-08-12 src/conclusion.tex
typos + amalgamation + conclusion

@@
 monoids $M$ recognize languages that are \kl{well-quasi-ordered} by the
-\kl{infix}, \kl{prefix} or \kl{suffix} relations. This is connected to finding
-which classes of graphs of \emph{bounded clique-width} are
-\kl{well-quasi-ordered} with respect to the \emph{induced subgraph relation}, a
-subject that is recently gaining traction.
+\kl{infix}, \kl{prefix} or \kl{suffix} relations. This research direction is
+connected to finding which classes of graphs of \emph{bounded clique-width} are
+\kl{well-quasi-ordered} with respect to the \emph{induced subgraph relation},
+as shown in \cite{DRT10}, and recently revisited by one of the authors in
+\cite{L24:arxiv:v2}.
+
+\subparagraph{Complexity} We have chosen to disregard complexity considerations
+when proving decidability results, as we do not believe that a fined grained
+complexity analysis would be particularly enlightening. However, we conjecture
+that for regular languages, the complexity should be polynomial in the size of
+the minimal automaton recognizing the language.

### 2024-08-12 src/infixes.tex
typos + amalgamation + conclusion

@@
 
+\begin{corollary}
+    \label{powers-infixes:cor}
+    Let $u,v \in \Sigma^*$ and $k, \ell \in \Nat$
+    such that $k \geq \factorial[p]{\card{u} \times \card{v}}$,
+    $\ell \geq \factorial[p]{\card{v} \times \card{u}}$,
+    and $u^k \infleq v^\ell$.
+    Then, there exists $w \in \Sigma^*$ of size at most
+    $\min \set{\card{u}, \card{v}}$ and a $p \in \Nat$
+    such that
+    $u^k \infleq v^\ell \infleq w^p$.
+\end{corollary}
+
 The reason why \kl{periodic words} built using a given period $x \in \Sigma^+$
@@
 \in \Nat}$ is a \kl{chain} for the \kl{infix relation}. Note that in general,
-the \kl{downwards closure} of a chain is \emph{not} a chain. However, for the chains
-generated using periodic words, the \kl{downwards closure}
+the \kl{downwards closure} of a chain is \emph{not} a chain. However, for the
+chains generated using periodic words, the \kl{downwards closure}
 $\dwset[\infleq]{\setof{x^p}{p \in \Nat}}$ is a \emph{finite union} of
-\kl{chains}. Because this set will appear in bigger equations,
-we introduce the shorter notation 
-$\intro*\InfPeriodChain{x}$ for the set of 
-\kl{infixes} of words of the form $x^p$, where $p$ ranges over $\Nat$.
+\kl{chains}. Because this set will appear in bigger equations, we introduce the
+shorter notation $\intro*\InfPeriodChain{x}$ for the set of \kl{infixes} of
+words of the form $x^p$, where $p$ ranges over $\Nat$.
 
@@
     Because $\InfPeriodChain{x}$ is a finite union of \kl{chains} for the \kl{suffix}
-    relation (using \cref{inf-period-chain:lem}), we conclude that $\InfPeriodChain{x}u$ is one too for
+    relation (rsing \cref{inf-period-chain:lem}), we conclude that $\InfPeriodChain{x}u$ is one too for
     the \kl{infix} relation. This holds similarly for $u \InfPeriodChain{x}$.
@@
     \end{itemize}
-    In all those cases, we conclude using \cref{periodic-infixes:lem}.
+    In all those cases, we conclude using \cref{powers-infixes:cor}
+    that there exists $x,y \in \Sigma^+$ of size at most $s$, and 
+    a number $1 \leq t \leq k$ such that
+    $v_i^{n_i} \in \InfPeriodChain{x}$ for all $1 \leq i \leq t$,
+    and
+    $v_i^{n_i} \in \InfPeriodChain{y}$ for all $t < i \leq k$.
+    In particular,
+    $w[\vec{n_i}] \in \InfPeriodChain{x} u_{t} \InfPeriodChain{y}$.
 
@@
 
-We are now ready to restate and proof our main theorem.
+We are now ready to restate and prove our main theorem.
 
@@
 
-\AP Let us now formally introduce the notion of \kl{amalgamation systems}, and
+Let us now formally introduce the notion of \kl{amalgamation systems}, and
 recall some results from \cite{ASZZ24} that will be useful for the proof of
-\cref{infix-amalgamation:thm}. Let us recall that over an alphabet $(\Sigma,
-=)$ a \kl{subword embedding} between two words $u \in \Sigma^*$ and $v \in
-\Sigma^*$ is a function $\rho \colon \range{\card{u}} \to \range{\card{v}}$
-such that $u_i = v_{\rho(i)}$ for all $i \in \range{\card{u}}$. We write
-$\intro*\HigEmb(u,v)$ the set of all \kl{subword embeddings} between $u$ and
-$v$. It may be useful to notice that the set of finite words over $\Sigma$
-forms a category when we consider \kl{subword embeddings} as morphisms.
+\cref{infix-amalgamation:thm}. The notion of \kl{amalgamation system} is
+tailored to produce \emph{pumping arguments}, which is exactly what our
+\cref{pumping-periods:lem} talks about. At the core of a pumping argument,
+there is a notion of \emph{run}, which could for instance be a sequence of
+transitions taken in a finite state automaton. Continuing on the analogy with
+finite automata, there is a natural ordering between runs, i.e., a run is
+smaller than another one if one can ``unroll loops" of the first run to obtain
+the second one. Typical pumping arguments then rely on the fact that
+\emph{minimal} runs are of finite size, and that all other runs are
+obtained by ``gluing" minimal runs together. This is exactly what
+\kl{amalgamation systems} are about.
 
-\AP Given a \kl{subword embedding} $f \colon u \to v$ between two words $u$ and
-$v$, there exists a unique decomposition $v = \GapWord{f}{0} u_1 \GapWord{f}{1}
-\cdots \GapWord{f}{k-1} u_k \GapWord{f}{k}$ where $\GapWord{f}{i} =
-v[f(i)+1:f(i+1)-1]$ for all $1 \leq i \leq k-1$, $\GapWord{f}{k} =
-v[f(k)+1:\card{v}]$, and $\GapWord{f}{0}   = v[1: f(1)-1]$.  We say that
-$\intro*\GapWord{f}{i}$ is the $i$-th \intro{gap word} of $f$. Given a run $r
-\in R$, and $i \in \range[0]{\card{\canrun(r)}}$, we define the \intro{gap
-language} of $r$ at position $i$ $\GapLanguage{r}{i} \defined
-\setof{\GapWord{f}{i}}{\exists s \in R. \exists f \in E(r,s) }$.
+\AP Let us recall that over an alphabet $(\Sigma, =)$ a \kl{subword embedding}
+between two words $u \in \Sigma^*$ and $v \in \Sigma^*$ is a function $\rho
+\colon \range{\card{u}} \to \range{\card{v}}$ such that $u_i = v_{\rho(i)}$ for
+all $i \in \range{\card{u}}$. We write $\intro*\HigEmb(u,v)$ the set of all
+\kl{subword embeddings} between $u$ and $v$. It may be useful to notice that
+the set of finite words over $\Sigma$ forms a category when we consider
+\kl{subword embeddings} as morphisms. Given a \kl{subword embedding} $f \colon
+u \to v$ between two words $u$ and $v$, there exists a unique decomposition $v
+= \GapWord{f}{0} u_1 \GapWord{f}{1} \cdots \GapWord{f}{k-1} u_k \GapWord{f}{k}$
+where $\GapWord{f}{i} = v[f(i)+1:f(i+1)-1]$ for all $1 \leq i \leq k-1$,
+$\GapWord{f}{k} = v[f(k)+1:\card{v}]$, and $\GapWord{f}{0}   = v[1: f(1)-1]$.
+We say that $\intro*\GapWord{f}{i}$ is the $i$-th \intro{gap word} of $f$.
 
@@
     and 
-    $E$ is a set of so-called \emph{run embeddings}.
-    \begin{itemize}
+    $E$ is a set of so-called \emph{run embeddings},
+    and $\canrun \colon R \to (\Sigma \cup \set{\varepsilon})^*$ is a 
+    function computing a \emph{canonical decompostion} of a run.
+    Given a run $r \in R$, and $i \in \range[0]{\card{\canrun(r)}}$, we define
+    the \intro{gap language} of $r$ at position $i$ as $\GapLanguage{r}{i} \defined
+    \setof{\GapWord{f}{i}}{\exists s \in R. \exists f \in E(r,s) }$.
+    An \kl{amalgamation system} furthermore satisfies the following 
+    properties:
+    \begin{enumerate}
         \item For all $r, s \in R$,
@@
         \item $(R, \leq_E)$ is a well-quasi-ordered set.
-        \item concatenative amalagamation
-    \end{itemize}
+        \item For all $r, s, t \in R$,
+            for all $f \in E(r,s)$,
+            for all $g \in E(r,t)$,
+            and for all $0 \leq i_0 \leq \card{\canrun(r)}$,
+            there exists a run $z \in R$ and morphisms
+            $h \in E(s,z)$ and $k \in E(t,z)$
+            satisfying
+            $h \circ f = k \circ g$,
+            $\GapLanguage{h \circ f}{i} = \GapLanguage{f}{i} \GapLanguage{g}{i}$
+            or 
+            $\GapLanguage{h \circ f}{i} = \GapLanguage{g}{i} \GapLanguage{f}{i}$
+            for every $0 \leq j \leq n$,
+            and
+            $\GapLanguage{h \circ f}{i_0} = \GapLanguage{f}{i_0} \GapLanguage{g}{i_0}$.
+    \end{enumerate}
 
@@
 
-The notion of \kl{amalgamation system} is tailored to produce \emph{pumping
-arguments}, which is exactly what our \cref{pumping-periods:lem} talks about.
-Because the notion is quite abstract, it may be useful to replicate some
-intuitions developed by the authors of \cite{ASZZ24}.
+Intuitively, the definition of an amalgamation system allows the comparison of
+runs, and the proper ``gluing" of runs together to obtain new runs.
 
@@
 
-    Let us first argue that languages $\GapLanguage{r}{i}$ are all \kl{chains}
-    for the \kl{prefix ordering}. According to \cite[Lemma 4.1]{ASZZ24}, either
-    $\GapLanguage{r}{i}$ is a \kl{chain} for the \kl{prefix relation}, or it is
-    an \kl{unbounded language}. Note that \kl{unbounded languages}
-    are \kl{antichains} for the \kl{infix relation}, and therefore
-    \textbf{unclear}.
+    Let $u$ and $v$ be two words in the \kl{gap language}
+    $\GapLanguage{r}{\ell}$ for some $r \in M$ and $0 \leq \ell \leq
+    \card{\canrun(r)}$. We have corresponding runs $s_1, s_2 \in R$ and
+    embeddings $f_1 \in E(r,s_1)$ and $f_2 \in E(r,s_2)$ such that
+    $\GapWord{f_1}{\ell} = u$ and $\GapWord{f_2}{\ell} = v$. Using the
+    amalgamation property, we can inductively construct runs $r_k \in R$ and
+    embeddings $f_k \in E(r,s_k)$ such that $\GapWord{f_k}{\ell} = u v^k u$,
+    and $\GapWord{f_k}{i}$ is obtained by concatenating in some order two
+    occurrences of $\GapWord{f_1}{i}$ and $k$ occurrences of
+    $\GapWord{f_2}{i}$. Because $L$ is \kl{well-quasi-ordered} for the
+    \kl{infix relation}, there exists $i < j$ such that $\yieldrun(r_i) \infleq
+    \yieldrun(r_j)$. Assuming that $\card{\canrun(r)}$ is greater than $2$, and
+    that we have extracted an increasing sequence of indices so that
+    $\alpha(i+1) > \factorial{\card{\yieldrun(r_i)}}$, we conclude that $u
+    v^{\alpha(i)} u$ is an \kl{infix} of $\yieldrun(r_j)$.
+    Because of the number of repetitions, a \emph{large} infix of $v^{\alpha(i)}$ must be
+    an infix of some $\GapWord{f_2}{p}^{\alpha(j)}$.
+    We conclude that $\Gapword{f_2}{p}^{\alpha(j)}$ and $v$
+    have a common period. Refining the analysis, we also conclude
+    that $uv^{q}$ and $\GapWord{f_2}{p}^{\alpha(j)}$ have the same period,
+    which proves that $uv^{q}u$ belongs to some $\InfPeriodChain{x}$
+    \textbf{todo}.
+
+    \begin{quote}
+        Idea: when pumping like this, we can always find
+        large blocks of $\GapWord{f_2}{p}$, so this looks like
+        something suitable for \cref{pumping-periods:lem}.
+        This should allow us to conclude that every pair of words
+        $u,v$ in a given gap language
+        are infixes of \emph{the same} periodic word repeated.
+    \end{quote}
 

### 2024-08-12 src/introduction.tex
typos + amalgamation + conclusion

@@
 is recognized by a finite automaton, or more generally that $L$ is the language
-of an \kl{amalgamation system}. Under these relatively mild assumptions, we
-also prove that \kl{well-quasi-ordered} languages under the \kl{infix relation}
-are precisely the finite unions of \kl{chains}.
+of an \kl{amalgamation system}, a recently introduced notion that abstracts the
+notion of systems whose runs can be suitably ``pumped'' \cite{ASZZ24}. Under
+this relatively mild assumption, we also prove that \kl{well-quasi-ordered}
+languages under the \kl{infix relation} are precisely the finite unions of
+\kl{chains}.
 

### 2024-08-15 src/infixes.tex
typos

@@
 
-In this section, we study the well-quasi-ordering of languages under the
-\kl{infix relation}. As opposed to the \kl{prefix} and \kl{suffix} relations,
-the \kl{infix relation} can yield to very complicated \kl{well-quasi-ordered}
-languages. Formally, the upcoming \cref{infix-embedding:thm} shows that \emph{any} countable
-quasi-ordering with finite initial segments can be embedded into the infix
-relation of a language.
-
-\AP Let us recall that an \intro{order embedding} from a quasi-ordered set $(X,
+\AP
+In this section, we study languages equipped with the \kl{infix relation}. As
+opposed to the \kl{prefix} and \kl{suffix} relations, the \kl{infix relation}
+can yield to very complicated \kl{well-quasi-ordered} languages. Formally, the
+upcoming \cref{infix-embedding:thm} shows that \emph{any} countable
+partial-ordering with finite initial segments can be embedded into the infix
+relation of a language. To make the former statement precise,
+Let us recall that an \intro{order embedding} from a quasi-ordered set $(X,
 \preceq)$ into a quasi-ordered set $(Y, \preceq')$ is a function $f \colon X
@@
     elements of $X \setminus \bigcup_{i \leq n} X_i$. Because initial segments
-    of $X$ are finite, it is clear that $X = \bigcup_{n \in \Nat} X_n$. To
+    of $X$ are finite, we know that $X = \bigcup_{n \in \Nat} X_n$. To
     simplify notations, we will also use $Y_n \defined \bigcup_{i \leq n} X_i$.
@@
     $L_{n+1} \defined L_n \cup c^{n+1} b^* c^{n+1} (L_n c^{n+1})^*$.
-    Now, let us inductively define \kl{embeddings} $f_n \colon Y_n \to
-    L_n$.
+    Now, let us inductively define \kl{embeddings} $f_n \colon (Y_n,\preceq) \to
+    (L_n, \infleq)$.
 
@@
     Let $L \subseteq \Sigma^*$ be a language recognized by a finite automaton.
-    Then $L$ is well-quasi-ordered by the infix relation if and only if $L$ is
-    a finite union of chains for the \kl{infix relation}.
+    Then $L$ is \kl{well-quasi-ordered} by the \kl{infix relation} if and only if $L$ is
+    a finite union of \kl{chains} for the \kl{infix relation}.
 \end{theorem}
@@
 \end{proof}
+
+The upcoming \cref{powers-infixes:cor} is based on the observation that if $S$
+is a \kl{chain} for the \kl{suffix relation} and $P$ is a \kl{chain} for the
+\kl{prefix relation}, then $SP$ is a \kl{chain} for the \kl{infix relation}.
 
@@
 \kl{well-quasi-ordered} to a property of the \kl{periodic lengths} of words in
-a language, based on the assumption that some factors can be iterated.
+a language, based on the assumption that some factors can be iterated. It is
+the core result that powers the analysis done in the upcoming
+\cref{infix-finite-automata:thm,infix-amalgamation:thm}.
 
@@
     exists $i < j$ such that $w[\vec{n_i}]$ is an \kl{infix} of $w[\vec{n_j}]$.
-    Now, because of the values of $\vec{n_j}$, there exists $1 \leq \ell \leq
+    Now, because of the chosen values for $\vec{n_j}$, there exists $1 \leq \ell \leq
     k-1$ such that $w[\vec{n_i}]$ is actually an \kl{infix} of $u_{\ell}
@@
 
-We are now ready to restate and prove our main theorem.
+We are now ready to restate and prove our main theorem, proven by combining
+\cref{pumping-periods:lem} together with classical pumping arguments on finite
+state automata.
 
@@
 \begin{proof}
-    It suffices to decide whether the equation 
+    It suffices to decide whether 
     \cref{infix-automata:eq} holds for the language $L$.
@@
 \end{proof}
+
+Let $(X,\preceq)$ be a quasi-ordered set, and $L \subseteq X$ be such that $(L,
+\preceq)$ is \kl{well-quasi-ordered}. It is not true in general that
+$(\dwset{L}, \preceq)$ is \kl{well-quasi-ordered}. In the case of $(\Sigma^*,
+\infleq)$ a typical example is to start from an infinite \kl{antichain} $A$,
+together with an enumeration $\seqof{w_i}$ of $A$, and build the language $L
+\defined \setof{ \prod_{i = 0}^n w_i }{ i \in \Nat }$. By definition, $L$ is a
+\kl{chain} for the \kl{infix} ordering, hence \kl{well-quasi-ordered}. However,
+$\dwset[\infleq]{L}$ contains $A$, and is therefore not
+\kl{well-quasi-ordered}. As a fun corollary of our analysis, we conclude that
+this particular example cannot be realized by a regular language $A$.
+
+\begin{corollary}
+    Given a regular language $L$, it is \kl{well-quasi-ordered} 
+    for the \kl{infix relation} if and only if 
+    $\dwset[\infleq] L$ is.
+\end{corollary}
 
@@
 the set of finite words over $\Sigma$ forms a category when we consider
-\kl{subword embeddings} as morphisms. Given a \kl{subword embedding} $f \colon
-u \to v$ between two words $u$ and $v$, there exists a unique decomposition $v
-= \GapWord{f}{0} u_1 \GapWord{f}{1} \cdots \GapWord{f}{k-1} u_k \GapWord{f}{k}$
-where $\GapWord{f}{i} = v[f(i)+1:f(i+1)-1]$ for all $1 \leq i \leq k-1$,
-$\GapWord{f}{k} = v[f(k)+1:\card{v}]$, and $\GapWord{f}{0}   = v[1: f(1)-1]$.
-We say that $\intro*\GapWord{f}{i}$ is the $i$-th \intro{gap word} of $f$.
+\kl{subword embeddings} as morphisms, which is a fancy way to state that
+$\mathrm{id} \in \HigEmb(u,u)$ and that $f \circ g \in \HigEmb(u,w)$ whenever
+$g \in \HigEmb(u,v)$ and $f \in \HigEmb(v,w)$, for any choice of words
+$u,v,w \in \Sigma^*$.
+
+\AP Given a \kl{subword embedding} $f \colon u \to v$ between two words $u$ and
+$v$, there exists a unique decomposition $v = \GapWord{f}{0} u_1 \GapWord{f}{1}
+\cdots \GapWord{f}{k-1} u_k \GapWord{f}{k}$ where $\GapWord{f}{i} =
+v[f(i)+1:f(i+1)-1]$ for all $1 \leq i \leq k-1$, $\GapWord{f}{k} =
+v[f(k)+1:\card{v}]$, and $\GapWord{f}{0}   = v[1: f(1)-1]$. We say that
+$\intro*\GapWord{f}{i}$ is the $i$-th \intro{gap word} of $f$. These \kl{gap
+words} will be useful to describe how and where runs of a system (described by
+words) can be combined.
 
@@
     $E$ is a set of so-called \emph{run embeddings},
-    and $\canrun \colon R \to (\Sigma \cup \set{\varepsilon})^*$ is a 
-    function computing a \emph{canonical decompostion} of a run.
-    Given a run $r \in R$, and $i \in \range[0]{\card{\canrun(r)}}$, we define
-    the \intro{gap language} of $r$ at position $i$ as $\GapLanguage{r}{i} \defined
+    and $\canrun \colon R \to (\Sigma \uplus \set{\varepsilon})^*$ is a 
+    function computing a \emph{canonical decomposition} of a run.
+    Given a run $r \in R$, and $i \in \range[0]{\card{\canrun(r)}}$, 
+    the \intro{gap language} of $r$ at position $i$ is $\GapLanguage{r}{i} \defined
     \setof{\GapWord{f}{i}}{\exists s \in R. \exists f \in E(r,s) }$.
@@
     \begin{enumerate}
-        \item For all $r, s \in R$,
+        \item \emph{Word Embeddings as Morphisms.} For all $r, s \in R$,
             $E(r,s) \subseteq \HigEmb(\canrun(r), \canrun(s))$.
-        \item $(R, E)$ forms a category.
-        \item $(R, \leq_E)$ is a well-quasi-ordered set.
-        \item For all $r, s, t \in R$,
+        \item \emph{$(R, E)$ Forms a Category.}
+            For all $r,s,t \in R$,
+            $\mathrm{id} \in E(r,r)$,
+            and whenever $f \in E(r,s)$ and $g \in E(s,t)$,
+            then $g \circ f \in E(r,t)$.
+        \item \emph{Well-Quasi-Ordered System.}
+            $(R, \leq_E)$ is a well-quasi-ordered set,
+            where $r \leq_E s$ if and only if $E(r,s) \neq \emptyset$.
+        \item \emph{Controlled Amalgamation.}
+            For all $r, s, t \in R$,
             for all $f \in E(r,s)$,
@@
     there exists a run $r \in R$
-    such that the concatenation of the letters (and possibly the empty word) in
-    $\canrun(r)$ equals $w$.
+    such that the concatenation of the letters of its
+    canonical decomposition, written $\intro*\yieldrun(r)$,
+    equals $w$.
 \end{definition}
@@
 Intuitively, the definition of an amalgamation system allows the comparison of
-runs, and the proper ``gluing" of runs together to obtain new runs.
+runs, and the proper ``gluing" of runs together to obtain new runs. Let us
+recall some examples of languages that can be recognized by \kl{amalgamation
+systems} given by the authors of the original paper: regular languages
+\cite[Theorem 5.3]{ASZZ24}, VASS as a consequence of \cite[Theorem
+5.5]{ASZZ24}, context-free languages as a consequence of \cite[Theorem
+5.10]{ASZZ24}. For this paper to be self-contained, we will also recall how
+runs of a finite state automaton can be understood as an \kl{amalgamation
+system}.
 
-Let us recall some examples of languages that can be recognized by
-\kl{amalgamation systems} given by the authors of the original paper: regular
-languages \cite[Theorem 5.3]{ASZZ24}, VASS as a consequence of \cite[Theorem
-5.5]{ASZZ24}, context-free languages as a consequence of \cite[Theorem
-5.10]{ASZZ24}.
+\begin{example}
+    Let $A = (Q, \delta, q_0, F)$ be a finite state automaton over a finite
+    alphabet $\Sigma$. Let $\Delta$ be the set of transitions $(q_1, a, q_2)
+    \in Q \times \Sigma \times Q$,
+    and $R \subseteq \Delta^*$ be the set of 
+    words over transitions that start with the initial state $q_0$,
+    end in a final state $q_f \in F$, and such that the end state of a
+    letter is the start state of the following one.
+
+    The canonical decomposition $\canrun \colon R \to \Sigma^*$
+    is defined as a morphism from $\Delta^*$ to $\Sigma^*$
+    that maps $(q,a,p)$ to $a$.
+
+    Finally, given two runs $r$ and $s$ of the automaton,
+    let $E(r,s) = \HigEmb(\canrun(r), \canrun(s))$, that is, we do not
+    remove morphisms.
+
+    The system $(\Sigma, R, E, \canrun)$ is an \kl{amalgamation system},
+    whose language is precisely the language of words recognized
+    by the automaton $A$.
+\end{example}
+\begin{proof}
+    TODO.
+\end{proof}
 
@@
         L \subseteq \bigcup_{r \in M} 
-        \left(
+        \yieldrun \left(
         \GapLanguage{r}{0}
@@
     Because of the number of repetitions, a \emph{large} infix of $v^{\alpha(i)}$ must be
-    an infix of some $\GapWord{f_2}{p}^{\alpha(j)}$.
-    We conclude that $\Gapword{f_2}{p}^{\alpha(j)}$ and $v$
+    an infix of some $(\GapWord{f_2}{p})^{\alpha(j)}$.
+    We conclude that $(\GapWord{f_2}{p})^{\alpha(j)}$ and $v$
     have a common period. Refining the analysis, we also conclude
-    that $uv^{q}$ and $\GapWord{f_2}{p}^{\alpha(j)}$ have the same period,
+    that $uv^{q}$ and $(\GapWord{f_2}{p})^{\alpha(j)}$ have the same period,
     which proves that $uv^{q}u$ belongs to some $\InfPeriodChain{x}$

### 2024-08-15 src/introduction.tex
typos

@@
 is that they are closed under many operations, such as taking products, finite
-unions, and finite powersets \cite{SCSC12}. More surprisingly, the class of
-well-quasi-ordered sets is also stable under the operation of taking finite
-words and finite trees labelled by elements of a well-quasi-ordered set
+unions, and finite powersets \cite{SCSC12}. Perhaps more surprisingly, the
+class of well-quasi-ordered sets is also stable under the operation of taking
+finite words and finite trees labelled by elements of a well-quasi-ordered set
 \cite{HIG52,KRU72}.
@@
 
-\AP
-However, there are many other natural orderings on words that could be
-considered in the context of \kl{well-quasi-orderings}. In our, case we
-consider alphabets $\Sigma$ that are finite and equipped with the equality
-relation. In this setting, the three alternatives we consider are the
-\intro{prefix relation} --- $u \intro*\prefleq v$ if there exists $w$ with $uw
-= v$ ---, the \intro{suffix relation} --- $u \intro*\suffleq v$ if there exists
-$w$ such that $wu = v$ ---, and the \intro{infix relation} --- $u
-\intro*\infleq v$ if there exists $w_1,w_2$ such that $w_1 u w_2 = v$ ---.
-Unfortunately, it is an easy check that all of these constructions are not
-well-quasi-ordered as soon as the alphabet contains two non-equivalent letters.
-This is not the end of the road, because some \emph{subsets} of $\Sigma^*$
-might still be \kl{well-quasi-ordered} by these relations, the simplest example
-being the case of finite sets of (finite) words which are all
-\kl{well-quasi-ordered} regardless of the ordering considered. This raises the
-question of characterizing the subsets $L \subseteq \Sigma^*$ that are
-\kl{well-quasi-ordered} with respect to the \kl{prefix relation} (respectively,
-the \kl{suffix relation} or the \kl{infix relation}), and providing suitable
-decision procedures.
+\AP However, there are many other natural orderings on words that could be
+considered in the context of \kl{well-quasi-orderings}. Let us restrict
+ourselves to a simplified case where the alphabet is finite and equipped with
+the equality relation. In this setting, the three alternatives we consider are
+the \intro{prefix relation} --- $u \intro*\prefleq v$ if there exists $w$ with
+$uw = v$ ---, the \intro{suffix relation} --- $u \intro*\suffleq v$ if there
+exists $w$ such that $wu = v$ ---, and the \intro{infix relation} --- $u
+\intro*\infleq v$ if there exists $w_1,w_2$ such that $w_1 u w_2 = v$ ---. Note
+that these three relations straightforwardly generalize to infinite
+quasi-ordered alphabets. Unfortunately, it is an easy check that all of these
+constructions are not well-quasi-ordered as soon as the alphabet contains two
+non-equivalent letters. This is not the end of the road, because some
+\emph{subsets} of $\Sigma^*$ might still be \kl{well-quasi-ordered} by these
+relations, the simplest example being the case of finite sets of (finite) words
+which are all \kl{well-quasi-ordered} regardless of the ordering considered.
+This raises the question of characterizing the subsets $L \subseteq \Sigma^*$
+that are \kl{well-quasi-ordered} with respect to the \kl{prefix relation}
+(respectively, the \kl{suffix relation} or the \kl{infix relation}), and
+providing suitable decision procedures.
 
@@
 Let us recall that the key difference between \kl{well-quasi-orderings} and the
-more usual notion of well-founded orderings is that the latter allows the
+maybe more usual notion of well-founded orderings is that the latter allows the
 existence of infinite \intro{antichains}, that is, sets of pairwise
 incomparable elements. In fact, a set $X$ is \kl{well-quasi-ordered} if and
-only if it is \kl{well-founded} and does not contain an infinite \kl{antichain}
+only if it is \kl{well-founded} and does not contain infinite \kl{antichains}
 \cite{KRU72}.  As a consequence, our study focuses on detecting infinite
@@
 show that \emph{any} countable quasi-ordering with finite initial segments can
-be embedded into the \kl{infix relation} of a language over a binary alphabet
-$\Sigma \defined \set{a,b}$. As an example, on can embed $(\Sigma^*, \higleq)$
-into $(\Sigma^*, \infleq)$ where $\higleq$ is the \kl{subword relation} and
-$\infleq$ is the \kl{infix relation}. This strongly contrasts with the
-\kl{prefix relation}, and suggests that a meaningful characterization of
-\kl{well-quasi-ordered} languages under the \kl{infix relation} is unlikely to
-be found.
+be \kl{embedded} into the \kl{infix relation} of a language over a binary
+alphabet $\Sigma \defined \set{a,b}$. As an example, on can identify
+$(\Sigma^*, \higleq)$ with a subset of $(\Sigma^*, \infleq)$ where $\higleq$ is
+the \kl{subword relation} and $\infleq$ is the \kl{infix relation}. This
+strongly contrasts with the \kl{prefix relation}, and suggests that a
+meaningful characterization of \kl{well-quasi-ordered} languages under the
+\kl{infix relation} is unlikely to be found.
 
@@
 of an \kl{amalgamation system}, a recently introduced notion that abstracts the
-notion of systems whose runs can be suitably ``pumped'' \cite{ASZZ24}. Under
-this relatively mild assumption, we also prove that \kl{well-quasi-ordered}
-languages under the \kl{infix relation} are precisely the finite unions of
-\kl{chains}.
+notion of systems whose runs can be suitably ``pumped'' \cite{ASZZ24}. As
+examples \kl{amalgamation systems} covers the case of regular languages,
+context free languages, or even languages recognized by ``vector addition
+systems with state" \cite{ASZZ24}. Under this relatively mild assumption, we
+also prove that \kl{well-quasi-ordered} languages under the \kl{infix relation}
+are precisely the finite unions of \kl{chains}.
 
@@
 so-called \emph{derivation relations} where already analysed by Bucher,
-Ehrenfeucht, and Haussler \cite{BUEUD85}. These results were later extended by
+Ehrenfeucht, and Haussler \cite{BUEUD85}, and were later extended by
 D'Alessandro and Varricchio \cite{ALVA03,ALVA06}. However, in all those cases
@@
 \subparagraph{Acknowledgements} We would like to thank participants of the 2024
-edition of Autobóz for their helpful comments and discussions.
+edition of \kl{Autobóz} for their helpful comments and discussions.

### 2024-08-15 src/metadata.tex
typos

@@
 \author{%
-        Roland Guttenberg \and
-        Nathan Lhote \and
-        Aliaume Lopez \and
-        Corto Mascle \and
-        Vincent Michielini \and
-        Lia Schütze \and
-        Omid Yaghoubi}
+        Roland Guttenberg
+    \thanks{Technical University of Munich, Germany}
+     \and
+        Nathan Lhote
+    \thanks{Aix-Marseille University}
+     \and
+        Aliaume Lopez
+    \thanks{University of Warsaw}
+     \and
+        Corto Mascle
+    \thanks{LaBRI, University of Bordeaux}
+     \and
+        Vincent Michielini
+    \thanks{University of Warsaw}
+     \and
+        Lia Schütze
+    \thanks{MPI-SWS, Germany}
+     \and
+        Omid Yaghoubi
+    \thanks{University of Warsaw}
+    }
 
@@
     specific choice of ordering on words, the (scattered) subword
-    ordering. In this paper, to what extent other natural orderings
-    (prefix, suffix, infix) on words can be used to derive Higman-like
-    theorems. More specifically, we are interested in characterizing
-    \emph{languages} of words that are well-quasi-ordered under these
-    orderings. We show that a simple characterization is possible for
-    the prefix and suffix orderings, and that under extra regularity
-    assumptions, this also lifts to the infix ordering.
+    ordering. In this paper, we describe to what extent other natural
+    orderings (prefix, suffix, infix) on words can be used to derive
+    Higman-like theorems. More specifically, we are interested in
+    characterizing \emph{languages} of words that are well-quasi-ordered
+    under these orderings. We show that a simple characterization is
+    possible for the prefix and suffix orderings, and that under extra
+    regularity assumptions, this also lifts to the infix ordering.
 \end{abstract}

### 2024-08-15 src/prefixes.tex
typos + amalgamation + conclusion
typos

@@
 
-The next remark that we make is that $\Sigma^*$ is not \kl{well-quasi-ordered}
-by the \kl{prefix relation} as soon as $\Sigma$ contains two distinct letters
-$a$ and $b$. As an example of infinite \kl{antichain}, we can consider the set
-of words $a^n b$ for $n \in \Nat$. As mentioned in the introduction, there are
-some languages however that are \kl{well-quasi-ordered} by the \kl{prefix
+The next remark we make is that $\Sigma^*$ is not \kl{well-quasi-ordered} by
+the \kl{prefix relation} as soon as $\Sigma$ contains two distinct letters $a$
+and $b$. As an example of infinite \kl{antichain}, we can consider the set of
+words $a^n b$ for $n \in \Nat$. As mentioned in the introduction, there are
+however some languages  that are \kl{well-quasi-ordered} by the \kl{prefix
 relation}. A simple example being the (regular) language $a^* \subseteq
@@
 
-\begin{figure}
-    \centering
-    \caption{The tree of prefixes over the alphabet $\set{a,b}$,
-        with an \kl{antichain} in red, and
-        an \kl{antichain branch} in blue.}
-\end{figure}
-
 We will use this \kl{tree of prefixes} to find simple witnesses
@@
     branch $B$ of the \kl{tree of prefixes} such that from every point of the branch, 
-    one can reach a word of $L$ that is not in the branch.
+    one can reach a word in $L \setminus B$.
 \end{definition}
@@
 is a branch of the \kl{tree of prefixes}) is an \kl{antichain branch} for $L$.
-This holds because for any $a^k$ in the branch, $a \prefleq a^kb$ which is in
-$L$ and not in the branch. In general, the existence of an \kl{antichain
-branch} for a language $L$ implies that $L$ contains an infinite
-\kl{antichain}, and because the alphabet $\Sigma$ is assumed to be finite, one
-can leverage the fact that the \kl{tree of prefixes} is finitely branching to
-prove that the converse holds as well.
+This holds because for any $a^k$, the word $a \prefleq a^kb$ belongs to $L
+\setminus a^*$. In general, the existence of an \kl{antichain branch} for a
+language $L$ implies that $L$ contains an infinite \kl{antichain}, and because
+the alphabet $\Sigma$ is assumed to be finite, one can leverage the fact that
+the \kl{tree of prefixes} is finitely branching to prove that the converse
+holds as well.
 
@@
 \exists x \in S. x \preceq y}$. Anticipating the use of the symmetric notion of
-\intro{downwards closure}, let us introduce its notation 
-straight away as follows: $\dwset[\preceq]{S} \defined \setof{y
+\intro{downwards closure}, let us introduce it as follows: $\dwset[\preceq]{S} \defined \setof{y
 \in \Sigma^*}{ \exists x \in S. y \preceq x}$. Abusing notations, we will
 write $\upset{w}$ and $\dwset{w}$ for the upwards and downwards closure of a
-single element $w$, and omit the ordering relation when it is clear from the
+single element $w$, omitting the ordering relation when it is clear from the
 context.
@@
 As an immediate consequence, we have a very fine-grained understanding of the
-\kl{ordinal invariants} of such \kl{well-quasi-ordered} languages, which can be
+\kl{ordinal invariants} of such \kl{well-quasi-ordered} languages, which could be
 leveraged in bounding the complexity of algorithms working on such languages.

### 2024-08-16 src/infixes.tex
adding regular languages as example

@@
             $h \circ f = k \circ g$,
-            $\GapLanguage{h \circ f}{i} = \GapLanguage{f}{i} \GapLanguage{g}{i}$
+            $\GapWord{h \circ f}{i} = \GapWord{f}{i} \GapWord{g}{i}$
             or 
-            $\GapLanguage{h \circ f}{i} = \GapLanguage{g}{i} \GapLanguage{f}{i}$
+            $\GapWord{h \circ f}{i} = \GapWord{g}{i} \GapWord{f}{i}$
             for every $0 \leq j \leq n$,
             and
-            $\GapLanguage{h \circ f}{i_0} = \GapLanguage{f}{i_0} \GapLanguage{g}{i_0}$.
+            $\GapWord{h \circ f}{i_0} = \GapWord{f}{i_0} \GapWord{g}{i_0}$.
     \end{enumerate}
@@
 
-\begin{example}
+\begin{example}[{\cite[Section 3.2]{ASZZ24}}]
     Let $A = (Q, \delta, q_0, F)$ be a finite state automaton over a finite
@@
     letter is the start state of the following one.
-
     The canonical decomposition $\canrun \colon R \to \Sigma^*$
@@
     that maps $(q,a,p)$ to $a$.
-
     Finally, given two runs $r$ and $s$ of the automaton,
-    let $E(r,s) = \HigEmb(\canrun(r), \canrun(s))$, that is, we do not
-    remove morphisms.
+    we say that an embedding $f \in \HigEmb(\canrun(r), \canrun(s))$
+    belongs to $E(r,s)$ when
+    $f$ is also defining an embedding from $r$ to $s$ as words in $\Delta^*$,
+    where $\Delta$ is equipped with the equality relation.
 
@@
 \begin{proof}
-    TODO.
+    By definition, elements of $E(r,s)$ are defined as elements
+    of $\HigEmb(\canrun(r), \canrun(s))$, and they compose properly.
+    Because $\Delta = Q \times \Sigma \times Q$ is finite, it is 
+    a \kl{well-quasi-ordering} when equipped with the equality relation, and 
+    we conclude that $\Delta^*$ with $\higleq$ is a \kl{well-quasi-order}
+    according to Higman’s Lemma \cite{HIG52}.
+
+    Let us now move to proving that the system satisfies the amalgamation
+    property. Given three runs $r,s,t \in R$, and two embeddings $f \in E(r,s)$
+    and $g \in E(r,t)$, we want to construct an amalgamated run $s \vee t$.
+    Because letters in the run $r$ respect the transitions of the automaton
+    (i.e., if the letter $i$ ends in state $q$, then the letter $i+1$ starts in
+    state $q$), then the \kl{gap word} at position $i$ starts in state $q$ and
+    ends in state $q$ too. This means that for both embeddings
+    $f$ and $g$, the \kl{gap words} are read by the automaton by looping
+    on a state. In particular, these loops can be taken in any order
+    and continue to represent a valid run. That is, we can even select
+    the order of concatenation in the amalgamation for \emph{all} 
+    $0 \leq i \leq \card{\canrun(r)}$ and not just for one separately.
+
+    Let us now remark that 
+    the language of this amalgamation system is
+    the set of $\yieldrun(r)$ when $r$ ranges over $R$,
+    and because $R$ is the set of valid runs of the automaton,
+    and $\yieldrun(r)$ is the word read along this run,
+    we immediately conclude.
 \end{proof}
@@
         L \subseteq \bigcup_{r \in M} 
-        \yieldrun \left(
+        \left(
         \GapLanguage{r}{0}

### 2024-08-16 src/infixes.tex
adding decidability

@@
     Let $u$ and $v$ be two words in the \kl{gap language}
-    $\GapLanguage{r}{\ell}$ for some $r \in M$ and $0 \leq \ell \leq
+    $\GapLanguage{r}{\ell}$ for some $r \in M$ and $0 < \ell <
     \card{\canrun(r)}$. We have corresponding runs $s_1, s_2 \in R$ and
@@
     occurrences of $\GapWord{f_1}{i}$ and $k$ occurrences of
-    $\GapWord{f_2}{i}$. Because $L$ is \kl{well-quasi-ordered} for the
-    \kl{infix relation}, there exists $i < j$ such that $\yieldrun(r_i) \infleq
-    \yieldrun(r_j)$. Assuming that $\card{\canrun(r)}$ is greater than $2$, and
-    that we have extracted an increasing sequence of indices so that
-    $\alpha(i+1) > \factorial{\card{\yieldrun(r_i)}}$, we conclude that $u
-    v^{\alpha(i)} u$ is an \kl{infix} of $\yieldrun(r_j)$.
-    Because of the number of repetitions, a \emph{large} infix of $v^{\alpha(i)}$ must be
-    an infix of some $(\GapWord{f_2}{p})^{\alpha(j)}$.
-    We conclude that $(\GapWord{f_2}{p})^{\alpha(j)}$ and $v$
-    have a common period. Refining the analysis, we also conclude
-    that $uv^{q}$ and $(\GapWord{f_2}{p})^{\alpha(j)}$ have the same period,
-    which proves that $uv^{q}u$ belongs to some $\InfPeriodChain{x}$
-    \textbf{todo}.
+    $\GapWord{f_2}{i}$. 
+    This means that $\canrun(r_i)$ has the shape
+    \begin{equation*}
+        v_0^{x_1} u_0 v_0^{x_2} u_0 v_0^{x_3} \canrun(r_i)_1 
+        v_1^{x_4} u_1 v_1^{x_5} u_1 v_1^{x_6} \canrun(r_i)_2 
+        \cdots
+        v_{\ell}^{x_k} u_{\ell}
+        v_{\ell}^{x_{k+1}}  u_{\ell}
+        v_{\ell}^{x_{k+2}}
+        \canrun(r_i)_{\ell}
+        u v^i u
+        \canrun(r_i)_{\ell+1}
+        \cdots
+        \quad .
+    \end{equation*}
 
-    \begin{quote}
-        Idea: when pumping like this, we can always find
-        large blocks of $\GapWord{f_2}{p}$, so this looks like
-        something suitable for \cref{pumping-periods:lem}.
-        This should allow us to conclude that every pair of words
-        $u,v$ in a given gap language
-        are infixes of \emph{the same} periodic word repeated.
-    \end{quote}
+    Now, leveraging \cref{pumping-periods:lem}, we get that $u v^k u$ is an
+    infix of some $x^p$ where $x$ has size smaller or equal than $u$ and $v$.
+    This proves  that $u$ and $v$ are comparable for the prefix, infix, and
+    suffix relations. In particular, we have proven that $\GapLanguage{r}{i}$
+    is bounded, because of \cite[Lemma 4.1]{ASZZ24}, when $0 < i <
+    \card{\canrun(r)}$.
 
-    Then, we can leverage \cref{pumping-periods:lem}
-    to conclude that $L$ is a finite union of \kl{chains}
-    for the \kl{infix relation}.
+    For the edge cases, we do a very similar construction. For $i = 0$, we let
+    $\canrun(r_i) = v^k u \canrun(r_1) \dots$, we conclude similarly. For $i =
+    \card{\canrun(r)}$, we let $\canrun(r_i) = \dots
+    \canrun(r)_{\card{\canrun(r) - 1}} u v^k$.
 \end{proofof}
+
+The abstract definition of \kl{amalgamation system} can be enriched by
+effectiveness assumptions, as explained in \cite[Section 3.1.4]{ASZZ24}. In
+this setting, we conjecture that the following decidability result holds.
 

### 2024-08-19 src/infixes.tex
finish proofs

@@
 \begin{theorem}[restate=infix-finite-automata:thm,label=infix-finite-automata:thm]
+    \proofref{infix-finite-automata:thm}
     Let $L \subseteq \Sigma^*$ be a language recognized by a finite automaton.
@@
 
-\begin{proofof}{infix-finite-automata:thm}[main]
+\begin{proofof}{infix-finite-automata:thm}
     Let $w \in L$, because $Q$ is finite, there exists
@@
 \begin{theorem}[label=infix-amalgamation:thm,restate=infix-amalgamation:thm]
+    \proofref{infix-amalgamation:thm}
     Let $L \subseteq \Sigma^*$ be a language recognized by an 
@@
 
-\begin{proofof}{infix-amalgamation:thm}
+Let us now introduce a combinatorial lemma that explains how \kl{gap languages}
+can be pumped. Note that a single \kl{gap language} is always stable under
+concatenation, and our concern will be on a potential description of
+\emph{simultaneously} pumping different \kl{gap languages} to produce valid
+runs.
+
+\begin{lemma}
+    \label{pumping-gap-languages:lem}
+    Let $(\Sigma, R, E, \canrun)$ be an \kl{amalgamation system}, let $r \in R$
+    be a run of size $n \in \Nat$, and let $s, t \in R$ be two runs
+    with embeddings $f \in E(r,s)$ and $g \in E(r,t)$.
+    Let $\seqof{u_i}[0 \leq i \leq n]$ be the sequence of \kl{gap words}
+    for $f$, that is, $u_i =
+    \GapWord{f}{i}$ for all $0 \leq i \leq n$. Similarly, let
+    $\seqof{v_i}[0 \leq i \leq n]$ be the sequence of \kl{gap words} for $g$.
+
+    Then, for all $k \in \Nat$, and $0 < \ell < n$, the following words belong to 
+    the language of the \kl{amalgamation system}.
+    \begin{enumerate}
+        \item $(\prod_{i = 0}^{\ell - 1} u_i^{x_i} v_i u_i^{y_i} v_i u_i^{z_i}) 
+               v_\ell u_\ell^k v_\ell
+               (\prod_{i = \ell+1}^{n} u_i^{x_i} v_i u_i^{y_i} v_i u_i^{z_i})$,
+        \item $(\prod_{i = 0}^{n - 1} u_i^{x_i} v_i u_i^{y_i}) 
+               v_n u_n^k$,
+        \item $u_0^k v_0 
+            (\prod_{i = 1}^{n} u_i^{x_i} v_i u_i^{y_i})$.
+    \end{enumerate}
+    Where, for all $0 \leq i \leq n$, $x_i + y_i + z_i = k$, with $z_i = 0$
+    in the two last items.
+\end{lemma}
+\begin{proof}
+    The result immediately follows from the embedding property
+    applied inductively on runs obtained by gluing 
+    $k$ times $s$ with one or two times $t$ (depending on the item number).
+\end{proof}
+
+In the upcoming proof of \cref{infix-amalgamation:thm}, we will use the notion
+of \intro{bounded language} for languages $L \subseteq \Sigma^*$ such that
+there exists words $w_1, \dots, w_n$ satisfying $L \subseteq w_1^* \cdots
+w_n^*$. This will be useful for us to derive the inclusion of a language $L$
+into a union of \kl{chains}. Note that we isolate this intermediate result in a
+separate theorem that could be applied beyond \kl{amalgamation systems}.
+
+\begin{theorem}
+    \label{bounded-language:thm}
+    Let $L$ be a \kl{bounded language} of $\Sigma^*$. Then,
+    $L$ is a \kl{well-quasi-order} when endowed with the 
+    \kl{infix relation} if and only if it is a finite union of \kl{chains}.
+\end{theorem}
+\begin{proof}
+    Let $w_1, \dots, w_n$ be such that
+    $L \subseteq w_1^* \cdots w_n^*$.
+    Let us define $m \defined \max \setof{\card{w_i}}{1 \leq i \leq n}$
+
+    Let $w[\vec{k}] \defined w_1^{k_1} \cdots w_n^{k_n}$ be a map from $\Nat^k$
+    to $\Sigma^*$. We are interested in the intersection of the image of $w$
+    with $L$. Let us assume for instance that for all $\vec{k} \in \Nat^n$,
+    there exists $\vec{\ell} \geq \vec{k}$ such that $w[\vec{\ell}] \in L$.
+    Then, leveraging \cref{pumping-periods:lem}, we conclude that there exists
+    $x,y$ of size at most $\max\setof{\card{w_i}}{1 \leq i \leq n}$ such that
+    $w[\vec{k}] \in \InfPeriodChain{x} \cup \InfPeriodChain{x}
+    \InfPeriodChain{y}$, and we conclude that $L \subseteq \InfPeriodChain{x}
+    \cup \InfPeriodChain{x} \InfPeriodChain{y}$.
+
+    Now, it may be the case that one cannot simultaneously assume that two
+    component of the vector $\vec{k}$ are unbounded. In general, given a set $S
+    \subseteq \set{1, \dots, n}$ of indices, we say that $S$ is admissible if
+    there exists a bound $N_0$ such that for all $\vec{b} \in \Nat^S$, there
+    exists a vector $\vec{k} \in \Nat^n$, such that $\vec{k}$ is greater than
+    $\vec{b}$ on the $S$ components, and the other components are below the
+    bound $N_0$. The language of an admissible set $S$ is the set of words
+    obtained by repeating $w_i$ at most $N_0$ times if it is not in $S$
+    ($w_i^{\leq N_0}$) and arbitrarily many times otherwise ($w_i^*$).
+    Note that $L \subseteq \bigcup_{S \text{ admissible }} L(S)$.
+
+    Now, admissible languages are ready to be pumped according to
+    \cref{pumping-periods:lem}. For every admissible language,
+    the size of a word that is not iterated is at most
+    $N_0 \times m$ by definition, and we conclude that:
+    \begin{equation*}
+        L \subseteq 
+        \bigcup_{x,y \in \Sigma^{\leq n}}
+        \bigcup_{u \in \Sigma^{\leq m \times N_0}}
+        \InfPeriodChain{x} u \InfPeriodChain{y}
+        \cup
+        \InfPeriodChain{x} u
+        \cup
+        u \InfPeriodChain{x}
+        \quad .
+    \end{equation*}
+\end{proof}
+
+To conclude this section and prove our main \cref{infix-amalgamation:thm}, it
+suffices to demonstrate that languages recognized by \kl{amalgamation systems}
+that are \kl{well-quasi-ordered} for the \kl{infix relation} are \kl{bounded
+languages}.
+
+\begin{lemma}
+    \label{infix-amalgamation-bounded:lem}
+    Let $L \subseteq \Sigma^*$ be a language recognized
+    by an \kl{amalgamation system} that is \kl{well-quasi-ordered}
+    for the \kl{infix relation}. Then $L$ is a \kl{bounded language}.
+\end{lemma}
+\begin{proof}
     Assume that $L$ is \kl{well-quasi-ordered} by the \kl{infix relation},
@@
     $\GapLanguage{r}{\ell}$ for some $r \in M$ and $0 < \ell <
-    \card{\canrun(r)}$. We have corresponding runs $s_1, s_2 \in R$ and
-    embeddings $f_1 \in E(r,s_1)$ and $f_2 \in E(r,s_2)$ such that
-    $\GapWord{f_1}{\ell} = u$ and $\GapWord{f_2}{\ell} = v$. Using the
-    amalgamation property, we can inductively construct runs $r_k \in R$ and
-    embeddings $f_k \in E(r,s_k)$ such that $\GapWord{f_k}{\ell} = u v^k u$,
-    and $\GapWord{f_k}{i}$ is obtained by concatenating in some order two
-    occurrences of $\GapWord{f_1}{i}$ and $k$ occurrences of
-    $\GapWord{f_2}{i}$. 
-    This means that $\canrun(r_i)$ has the shape
+    \card{\canrun(r)} \defined n$. One can apply \cref{pumping-gap-languages:lem}
+    to conclude that  for all $k \geq 1$,
     \begin{equation*}
-        v_0^{x_1} u_0 v_0^{x_2} u_0 v_0^{x_3} \canrun(r_i)_1 
-        v_1^{x_4} u_1 v_1^{x_5} u_1 v_1^{x_6} \canrun(r_i)_2 
-        \cdots
-        v_{\ell}^{x_k} u_{\ell}
-        v_{\ell}^{x_{k+1}}  u_{\ell}
-        v_{\ell}^{x_{k+2}}
-        \canrun(r_i)_{\ell}
-        u v^i u
-        \canrun(r_i)_{\ell+1}
-        \cdots
-        \quad .
+       \left(\prod_{i = 0}^{\ell - 1} u_i^{x_i} v_i u_i^{y_i} v_i u_i^{z_i}\right) 
+       v u^k v
+       \left(\prod_{i = \ell+1}^{n} u_i^{x_i} v_i u_i^{y_i} v_i u_i^{z_i}\right)
+       \in 
+       L
     \end{equation*}
+    Where $x_i + y_i + z_i = k$ for all $0 \leq i \leq n$
+    Let us assume without loss of generality that the sequence
+    $\seqof{(x_i, y_i, n_i)}[i \geq 1]$ is pointwise increasing,
+    and even increasing coordinate wise. This is possible because
+    at least one of the coordinates tends to $+\infty$ and in the case
+    of a variable that is bounded, one can extract the sequence to
+    simply not use this variable.
 
-    Now, leveraging \cref{pumping-periods:lem}, we get that $u v^k u$ is an
-    infix of some $x^p$ where $x$ has size smaller or equal than $u$ and $v$.
-    This proves  that $u$ and $v$ are comparable for the prefix, infix, and
-    suffix relations. In particular, we have proven that $\GapLanguage{r}{i}$
-    is bounded, because of \cite[Lemma 4.1]{ASZZ24}, when $0 < i <
-    \card{\canrun(r)}$.
+    Now, leveraging \cref{pumping-periods:lem}, we get that for infinitely many
+    $k \geq 1$, $u v^k u$ is an infix of some $x^p$ where $x$ has size smaller
+    or equal than $u$ and $v$. This proves  that $u$ and $v$ are comparable for
+    the prefix, infix, and suffix relations. In particular, we have can
+    conclude that $\GapLanguage{r}{i}$ is a \kl{bounded language}, because of
+    \cite[Lemma 4.1]{ASZZ24}, when $0 < i < \card{\canrun(r)}$. The proof goes
+    on similarly for proving that $\GapLanguage{r}{0}$ and $\GapLanguage{r}{n}$
+    are \kl{bounded languages}, using the other items of
+    \cref{pumping-gap-languages:lem}.
 
-    For the edge cases, we do a very similar construction. For $i = 0$, we let
-    $\canrun(r_i) = v^k u \canrun(r_1) \dots$, we conclude similarly. For $i =
-    \card{\canrun(r)}$, we let $\canrun(r_i) = \dots
-    \canrun(r)_{\card{\canrun(r) - 1}} u v^k$.
+    Because \kl{bounded languages} are stable under concatenation and finite
+    unions, we conclude that $L$ itself must be a \kl{bounded language}.
+\end{proof}
+
+\begin{proofof}{infix-amalgamation:thm}
+    We first apply \cref{infix-amalgamation-bounded:lem},
+    and conclude with \cref{bounded-language:thm}.
 \end{proofof}

### 2024-08-19 src/infixes.tex
characterizing chains for the infix relation

@@
 \end{conjecture}
+
+Let us conclude this section by fully characterizing what are
+actually the chains for the \kl{infix} relation, without any 
+regularity assumption.
+
+\begin{lemma}
+    \label{chains-infix:lem}
+    Let $L$ be a \kl{chain} for the \kl{infix relation}. Then, there exists
+    a \kl{chain} $P$ for the \kl{prefix relation} and a \kl{chain}
+    $S$ for the \kl{suffix relation} such that $L \subseteq SP$.
+\end{lemma}
+\begin{proof}
+    Because $L$ is a chain,
+    we have an enumeration $L = \seqof{w_n}$,
+    where $s_n \defined \card{w_n}$ is increasing, 
+    and functions
+    $f_{i,j} \colon \set{1, \dots, s_i} \to \set{1, \dots s_{j}}$
+    satisfying
+    that $f_{j,k} \circ f_{i, j} = f_{i, k}$,
+    $f_{i,i} = \mathrm{id}$, and
+    $w_i = w_j[f_{i,j}(1):f_{i,j}(s_i)]$ for all $i \leq j$.
+    Now, we can build the infinite word $w_\infty$ indexed by $\Zel$, and
+    embeddings $f_{i, \infty} \colon \set{1, \dots, s_i} \to \Rel$ such that
+    $w_i = w_\infty[f_{i, \infty}(1):f_{i, \infty}(s_i)]$, and $f_{i, \infty}
+    \circ f_{j, i} = f_{j, \infty}$ for all $i \leq j$.
+
+    Let us define $w_l$ as the (infinite) word $w_\infty[-\infty: f_{0,
+    \infty}(s_0)]$ and $w_r$ as the (infinite) word $w_\infty[f_{0,
+    \infty}(s_0): +\infty]$. It is now clear that every word $w_i$ can be
+    written as $w_i = w_{i,l} w_{i,r}$, where $w_{i,l}$ is a prefix of $w_r$
+    and $w_{i,r}$ is a suffix of $w_l$. Furthermore, the functions $f_{i,
+    \infty}$ can be used to notice that $\seqof{w_{i,r}}$ is a \kl{chain} for
+    the \kl{prefix relation}, and $\seqof{w_{i,l}}$ is a \kl{chain} for the
+    \kl{suffix relation}, which allows us to conclude.
+\end{proof}

### 2024-08-19 src/infixes.tex
typo

@@
     $w_i = w_j[f_{i,j}(1):f_{i,j}(s_i)]$ for all $i \leq j$.
-    Now, we can build the infinite word $w_\infty$ indexed by $\Zel$, and
+    Now, we can build the infinite word $w_\infty$ indexed by $\Rel$, and
     embeddings $f_{i, \infty} \colon \set{1, \dots, s_i} \to \Rel$ such that

### 2024-08-19 src/introduction.tex
tests and all

@@
 
+\begin{figure}
+    \centering
+    \includestandalone{src/prefix-embedding-standalone}
+
+    \includestandalone{src/infix-embedding-standalone}
+
+    \includestandalone{src/subword-embedding-standalone}
+    \caption{A simple representation of the \kl{subword relation},
+        \kl{prefix relation},
+        and \kl{infix relation},
+        on the alphabet $\set{a,b}$ for words of
+        length at most $3$.}
+    \label{word-embeddings:fig}
+\end{figure}
+
 \AP However, there are many other natural orderings on words that could be
@@
 
-\subparagraph{Contribution} 
-\AP
-In this paper, we focus on words over a finite
+\subparagraph{Contribution} \AP In this paper, we focus on words over a finite
 alphabet $\Sigma$, and characterize subsets $L \subseteq \Sigma^*$ that are
@@
 \cite{KRU72}.  As a consequence, our study focuses on detecting infinite
-\kl{antichains} in languages $L \subseteq \Sigma^*$.
+\kl{antichains} in languages $L \subseteq \Sigma^*$. Remark that for the
+relations we consider (including the \kl{subword relation}), the set of all
+words of a fixed size $k$ is always an \kl{antichain}, and there are therefore
+no bound on the maximal length of an antichain.
 

### 2024-08-19 src/prefixes.tex
tests and all

@@
 will be useful in the rest of this section.
+
+\begin{figure}
+    \centering
+    \includestandalone{src/antichain-branch-standalone}
+    \caption{An \kl{antichain branch} for the language $a^* b$,
+        represented in the \kl{tree of prefixes} over the alphabet $\set{a,b}$.
+        The branch is represented with dashed lines in turquoise, and the
+        \kl{antichain} is represented in dotted lines in blood-red.
+    }
+    \label{antichain-branch:fig}
+\end{figure}
 
@@
 
+\begin{figure}
+    \centering
+    TODO.
+    \caption{Illustrating the construction of 
+        \cref{prefixes:thm} consisting in taking the words
+        that lead to prefix-incomparable elements in $S$,
+        and then the upwards closure of the maximal such elements.
+    }
+    \label{prefixes-theorem:fig}
+\end{figure}
+
 \begin{theorem}

### 2024-08-19 src/word-embeddings-figure.py
tests and all

@@ -1,0 +1,248 @@
+#!/usr/bin/env python3
+#
+# Aliaume Lopez
+#
+# This script generates figures representing some relation
+# between words using Hasse diagrams.
+# 
+
+import itertools as it
+from functools import cmp_to_key
+
+def repr_word(l):
+    return "".join(l)
+
+
+def test_repr_word():
+    assert repr_word("") == ""
+    assert repr_word(["a", "b", "a"]) == "aba"
+    assert repr_word(()) == ""
+    assert repr_word(("a", "b", "a")) == "aba"
+
+def repr_tikz(l):
+    if l == "":
+        return "epsilon"
+    else:
+        return l
+
+def test_repr_tikz():
+    assert repr_tikz("") == "epsilon"
+    assert repr_tikz("aba") == "aba"
+    assert repr_tikz("ab") == repr_word("ab")
+
+def repr_math(l):
+    if l == "":
+        return "$\\varepsilon$" 
+    else:
+        return l    
+
+def test_repr_math():
+    assert repr_math("") == "$\\varepsilon$"
+    assert repr_math("aba") == "aba"
+    assert repr_math("ab") == repr_word("ab")
+
+
+def subword(u,v):
+    if u == "":
+        return True
+
+    if len(u) > len(v):
+        return False
+
+    if u[0] == v[0]:
+        return subword(u[1:], v[1:]) or subword(u, v[1:])
+    else:
+        return subword(u, v[1:])
+
+
+def prefix(u,v):
+    return v.startswith(u)
+
+def suffix(u,v):
+    return v.endswith(u)
+
+def infix(u,v):
+    for i in range(len(v)-len(u)+1):
+        if v[i:i+len(u)] == u:
+            return True
+    return False
+
+
+
+
+
+def generate_figure(size = 4, relation = subword):
+    # We create a "triangle-like" 
+    # figure with the empty word ε at the bottom
+    # and each layer containing all the words of size k
+    # on top of eachother, and centered based on their size
+
+    layers = []
+    for k in range(size):
+        words = list(it.product("ab", repeat=k))
+        layers.append(words)
+    layers[0] = [""]
+
+    # The "nodes", are then dicts 
+    # with { "word": "abab", "x": x, "y": y}      
+    nodes = {}
+    for i in range(size):
+        layer_size = 2**i # =  len(layers[i])
+        for j, word in enumerate(layers[i]):
+            x = j - layer_size/2
+            y = i
+            w = repr_word(word)
[…]

### 2024-08-20 src/infixes.tex
tests and all
better figures

@@
 \end{proof}
+
+\begin{figure}
+    \centering
+    \includestandalone[width=\linewidth]{fig/infix-encoding-standalone}
+    \caption{Representation of the \kl{subword relation} for $\set{a,b}^*$
+        inside the \kl{infix relation} for $\set{a,b,c}^*$
+        using the encoding of \cref{infix-embedding:thm}, restricted to words
+        of length at most $3$.}
+    \label{infix-embedding:fig}
+\end{figure}
 
@@
 
+\begin{figure}
+    \centering
+    TODO
+    \caption{The gap words of an embedding between two 
+    finite words}
+    \label{gap-word-embedding:fig}
+\end{figure}
+
+\begin{figure}
+    \centering
+    TODO.
+    \caption{Amalgamation of two runs on a picture}
+    \label{amalgamation-runs:fig}
+\end{figure}
+
 \begin{definition}

### 2024-08-20 src/introduction.tex
better figures

@@
     \centering
-    \includestandalone{src/prefix-embedding-standalone}
-
-    \includestandalone{src/infix-embedding-standalone}
-
-    \includestandalone{src/subword-embedding-standalone}
+    \begin{tabular}{p{3cm}p{8cm}}
+    Prefix Relation
+    &
+    \includestandalone[width=\linewidth]{fig/prefix-embedding-standalone}
+    \\
+    Infix relation
+    &
+    \includestandalone[width=\linewidth]{fig/infix-embedding-standalone}
+    \\
+    Subword relation
+    &
+    \includestandalone[width=\linewidth]{fig/subword-embedding-standalone}
+    \end{tabular}
     \caption{A simple representation of the \kl{subword relation},
@@
         on the alphabet $\set{a,b}$ for words of
-        length at most $3$.}
+        length at most $3$. The figures are Hasse Diagrams,
+        representing the successor relation of the order.
+        Furthermore, we highlight in red relations that are added
+        when moving from the prefix relation to the infix one,
+        and to the infix relation to the subword one.}
     \label{word-embeddings:fig}

### 2024-08-20 src/word-embeddings-figure.py

@@ -13,11 +13,6 @@ def repr_word(l):
     return "".join(l)
 
 
-def test_repr_word():
-    assert repr_word("") == ""
-    assert repr_word(["a", "b", "a"]) == "aba"
-    assert repr_word(()) == ""
-    assert repr_word(("a", "b", "a")) == "aba"
 
 def repr_tikz(l):
     if l == "":
@@ -25,10 +20,6 @@ def repr_tikz(l):
     else:
         return l
 
-def test_repr_tikz():
-    assert repr_tikz("") == "epsilon"
-    assert repr_tikz("aba") == "aba"
-    assert repr_tikz("ab") == repr_word("ab")
 
 def repr_math(l):
     if l == "":
@@ -36,10 +27,6 @@ def repr_math(l):
     else:
         return l    
 
-def test_repr_math():
-    assert repr_math("") == "$\\varepsilon$"
-    assert repr_math("aba") == "aba"
-    assert repr_math("ab") == repr_word("ab")
 
 
 def subword(u,v):
@@ -66,6 +53,71 @@ def infix(u,v):
         if v[i:i+len(u)] == u:
             return True
     return False
+
+def print_power(u,n):
+    if n <= 1:
+        return u*n
+    else:
+        return f"{{{u}^{n}}}"
+
+def compress_word(u):
+    # group u by letters
+    # and replace each block
+    # by the print_power(x,n)
+    if u == "":
+        return ""
+    else:
+        l = it.groupby(u, lambda x: x)
+        return "".join( print_power(u, len(list(v))) for (u,v) in l if len(u) == 1)
+
+
+def encode_item_in_infix(item):
+    layer_num, item_num, dwset = item["layer_num"], item["item_num"], item["dwset"]
+    # f(0, i, _) = a b^i a
+    # f(n+1, i, D) = c^{n+1} b^i c^{n+1} \prod_{d \in D} (f(n, i, d) c^{n+1})
+    if layer_num == 0:
+        return "a" + "b"* item_num + "a"
+    else:
+        sep = "c"* layer_num
+        dwset_enc = sep.join(encode_item_in_infix(dw) for dw in dwset)
+        return "".join([sep, "b"*item_num, sep, dwset_enc, sep])
+
+
+def encode_in_infix(nodes, edges):
+    # we compute layers
+    # based on the minimal elements
+    # minimal elements
+    layer_num = 0
+    item_num = 1
+
+    for n in nodes.values():
[…]

### 2024-08-20 src/infixes.tex
better figures

@@
 y$ and $y \preceq x$ implies $x = y$.
+
+\marginpar{TODO: change the encoding to take the maximal elements of the
+downwards closure.}
 
@@
         using the encoding of \cref{infix-embedding:thm}, restricted to words
-        of length at most $3$.}
+        of length at most $3$. To obtain smaller words,
+        we replaced $D_i$ by $\max D_i$ in this construction.
+    }
     \label{infix-embedding:fig}
 \end{figure}
+
+The encoding using three distinct letters is clearer to use and to
+illustrate, but it is not surprising that one can encode a three
+letter alphabet into a two letter one, immediately
+leading to the following \cref{infix-embedding-binary:cor}.
+
+\begin{corollary}
+    \label{infix-embedding-binary:cor}
+    Let $(X, \preceq)$ be a \kl{partially ordered} set,
+    and $\Sigma \defined \set{a,b}$.
+    Then the following are equivalent:
+    \begin{enumerate}
+        \item \label{infix-embedding-embeds:item} 
+            $X$ \kl{embeds into} $(\Sigma^*, \infleq)$,
+        \item \label{infix-embedding-count:item}
+            $X$ is countable, and for every $x \in X$,
+            its \kl{downwards closure}
+            $\dwset[\preceq]{x}$ is finite.
+    \end{enumerate}
+\end{corollary}
+\begin{proof}
+    Let us construct an \kl{order embedding} from 
+    $(\set{a,b,c}^*, \infleq)$ to $(\set{0,1}, \infleq)$.
+    To that end, consider
+    the morphism $\mu$ that maps $a$ to $101$, $b$ to $1001$
+    and $c$ to $10001$.
+    Assume that $u \infleq v$ in $\set{a,b,c}^*$,
+    then $\mu(u) \infleq \mu(v)$ because $\mu$ is a morphism.
+    Conversely, assume that $\mu(u) \infleq \mu(v)$.
+    That, is, there exists an \kl{infix} $w$ of $\mu(v)$
+    such that $\mu(u) = w$.
+    Because $\mu(u)$ starts with $10$ and ends with $01$, so does
+    $w$. In particular, the word $w$ is compatible with $\mu$
+    in the sense that $w = \mu(s)$ for some $s \infleq v$.
+    Because the map $\mu$ is injective,
+    we conclude that $\mu(s) = \mu(u)$ implies $s = u$
+    and $u \infleq v$ as expected.
+\end{proof}
 

### 2024-08-20 src/prefixes.tex
better figures

@@
     \centering
-    \includestandalone{src/antichain-branch-standalone}
+    \includestandalone{fig/antichain-branch-standalone}
     \caption{An \kl{antichain branch} for the language $a^* b$,
@@
     \centering
-    TODO.
+    \includestandalone[width=\linewidth]{fig/prefix-core-and-branches-standalone}
     \caption{Illustrating the construction of 
-        \cref{prefixes:thm} consisting in taking the words
-        that lead to prefix-incomparable elements in $S$,
-        and then the upwards closure of the maximal such elements.
+        \cref{prefixes:thm}. We represent the Hasse Diagram of a
+        \kl{partial order}, with elements in $S$ being bigger and 
+        in desert sand, elements of $S_{\max}$ being in
+        blood-red, and \kl{chains} of elements in sky blue.
+        The upwards closure of an element in $S_{\max}$ 
+        is depicted in pale green, and cannot intersect two distinct
+        branches, as it does not contain two incomparable elements.
     }

### 2024-08-20 src/word-embeddings-figure.py

@@ -360,7 +360,7 @@ def test_compress_word():
     assert compress_word("caabbc") == "c{a^2}{b^2}c"
 
 def test_infix_encoding():
-    nodes, edges = generate_figure(4, subword)
+    nodes, edges = generate_figure(6, subword)
     nodes, edges = encode_in_infix(nodes, edges)
 
     for n in nodes.values():

### 2024-08-20 src/infixes.tex
better figures

@@
 v[f(k)+1:\card{v}]$, and $\GapWord{f}{0}   = v[1: f(1)-1]$. We say that
-$\intro*\GapWord{f}{i}$ is the $i$-th \intro{gap word} of $f$. These \kl{gap
-words} will be useful to describe how and where runs of a system (described by
-words) can be combined.
+$\intro*\GapWord{f}{i}$ is the $i$-th \intro{gap word} of $f$. We encourage the
+reader to look at \cref{gap-word-embedding:fig} to see an example of the
+\kl{gap words} resulting from a \kl{subword embedding} between two words. These
+\kl{gap words} will be useful to describe how and where runs of a system
+(described by words) can be combined.
 
@@
     \centering
-    TODO
-    \caption{The gap words of an embedding between two 
-    finite words}
+    \includestandalone[width=\linewidth]{fig/gap-word-embedding-standalone}
+    \caption{The \kl{gap words} resulting from a \kl{subword embedding} between two 
+    finite words.}
     \label{gap-word-embedding:fig}
@@
     \centering
-    TODO.
-    \caption{Amalgamation of two runs on a picture}
+    \includestandalone[width=\linewidth]{fig/run-amalgamation-standalone}
+    \caption{We illustrate how 
+        embeddings $f$ and $g$ between runs of an
+        \kl{amalgamation system} can be glued
+        together, seen on their \kl{canonical decomposition}.
+    }
     \label{amalgamation-runs:fig}

### 2024-08-20 src/infixes.tex
almost ready

@@
 y$ and $y \preceq x$ implies $x = y$.
-
-\marginpar{TODO: change the encoding to take the maximal elements of the
-downwards closure.}
 
@@
         \kl{amalgamation system} can be glued
-        together, seen on their \kl{canonical decomposition}.
+        together, seen on their canonical decomposition.
     }
@@
             $\GapWord{h \circ f}{i_0} = \GapWord{f}{i_0} \GapWord{g}{i_0}$.
+            We refer to \cref{amalgamation-runs:fig} for an illustration 
+            of this property.
     \end{enumerate}

### 2024-09-19 src/introduction.tex
work on intro/contribution

@@
 
-\AP
-Note that in the case of finite words and finite trees, the precise choice of
-ordering on words and trees is crucial to ensure that the resulting structure
-is \kl{well-quasi-ordered}. The celebrated result of Higman states that the set
-of finite words over an ordered alphabet $(X, \preceq)$ is
-\kl{well-quasi-ordered} by the so-called \kl{subword embedding relation}
-\cite{HIG52}. Let us recall that the \kl{subword relation} for words over $(X,
-\preceq)$ is defined as follows: a word $u$ is a \intro{subword} of a word $v$,
-written $u \intro*\higleq v$,
-if there exists an increasing function $f \colon \{1, \ldots, |u|\} \to \{1,
-\ldots, |v|\}$ such that $u_i \preceq v_{f(i)}$ for all $i \in \{1, \ldots,
-|u|\}$.
+\AP Note that in the case of finite words and finite trees, the precise choice
+of ordering is crucial to ensure that the resulting structure is
+\kl{well-quasi-ordered}. The celebrated result of Higman states that the set of
+finite words over an ordered alphabet $(X, \preceq)$ is \kl{well-quasi-ordered}
+by the so-called \kl{subword embedding relation} \cite{HIG52}. Let us recall
+that the \kl{subword relation} for words over $(X, \preceq)$ is defined as
+follows: a word $u$ is a \intro{subword} of a word $v$, written $u
+\intro*\higleq v$, if there exists an increasing function $f \colon \{1,
+\ldots, |u|\} \to \{1, \ldots, |v|\}$ such that $u_i \preceq v_{f(i)}$ for all
+$i \in \{1, \ldots, |u|\}$.
 
@@
     \centering
-    \begin{tabular}{p{3cm}p{8cm}}
-    Prefix Relation
-    &
+    \begin{subfigure}[t]{0.48\textwidth}
+    	\centering
     	\includestandalone[width=\linewidth]{fig/prefix-embedding-standalone}
-    \\
-    Infix relation
-    &
+    	\caption{Prefix Relation}
+   	\end{subfigure}%
+   	\hfill%
+   	\begin{subfigure}[t]{0.48\textwidth}
+   		\centering
    		\includestandalone[width=\linewidth]{fig/infix-embedding-standalone}
-    \\
-    Subword relation
-    &
+   		\caption{Infix Relation}
+   	\end{subfigure}
+   	\begin{subfigure}[t]{0.48\textwidth}
+   		\centering
    		\includestandalone[width=\linewidth]{fig/subword-embedding-standalone}
-    \end{tabular}
+   		\caption{Subword Relation}
+   	\end{subfigure}
+
    	\caption{A simple representation of the \kl{subword relation},
@@
 the equality relation. In this setting, the three alternatives we consider are
-the \intro{prefix relation} --- $u \intro*\prefleq v$ if there exists $w$ with
-$uw = v$ ---, the \intro{suffix relation} --- $u \intro*\suffleq v$ if there
-exists $w$ such that $wu = v$ ---, and the \intro{infix relation} --- $u
-\intro*\infleq v$ if there exists $w_1,w_2$ such that $w_1 u w_2 = v$ ---. Note
-that these three relations straightforwardly generalize to infinite
-quasi-ordered alphabets. Unfortunately, it is an easy check that all of these
-constructions are not well-quasi-ordered as soon as the alphabet contains two
-non-equivalent letters. This is not the end of the road, because some
-\emph{subsets} of $\Sigma^*$ might still be \kl{well-quasi-ordered} by these
-relations, the simplest example being the case of finite sets of (finite) words
+the \intro{prefix relation} ($u \intro*\prefleq v$ if there exists $w$ with $uw
+= v$), the \intro{suffix relation} ($u \intro*\suffleq v$ if there exists $w$
+such that $wu = v$), and the \intro{infix relation} ($u \intro*\infleq v$ if
+there exists $w_1,w_2$ such that $w_1 u w_2 = v$). Note that these three
+relations straightforwardly generalize to infinite quasi-ordered alphabets.
+Unfortunately, it is easy to see that all of these constructions are not
+well-quasi-ordered as soon as the alphabet contains two non-equivalent letters:
+for instance, the infinite sequence $a^n b^n a^n$ is \kl{well-quasi-ordered} by
+the \kl{subword relation} but by neither the \kl{prefix relation}, nor the
+\kl{suffix relation}, nor the \kl{infix relation}.
+
+\AP While this dooms \kl[well-quasi-ordered]{well-quasi-orderedness} of these relations in the general case, there may be
+\emph{subsets} of $\Sigma^*$ which are \kl{well-quasi-ordered} by these
+relations. As a simple example, take the case of finite sets of (finite) words
 which are all \kl{well-quasi-ordered} regardless of the ordering considered.
-This raises the question of characterizing the subsets $L \subseteq \Sigma^*$
-that are \kl{well-quasi-ordered} with respect to the \kl{prefix relation}
+This raises the question of characterizing exactly which subsets $L \subseteq \Sigma^*$
+are \kl{well-quasi-ordered} with respect to the \kl{prefix relation}
 (respectively, the \kl{suffix relation} or the \kl{infix relation}), and
@@
 
-\subparagraph{Contribution} \AP In this paper, we focus on words over a finite
+\subparagraph{Contribution} In this paper, we focus on words over a finite
 alphabet $\Sigma$, and characterize subsets $L \subseteq \Sigma^*$ that are
 \kl{well-quasi-ordered} by the \kl{prefix relation}, the \kl{suffix relation},
-and the \kl{infix relation}. Let us remark that these word orderings are all
-\intro{well-founded}, i.e., they do not contain infinite \intro{decreasing
-sequences} $\seqof{x_n}$ such that $x_{n+1} \prec x_n$ for all $n \in \Nat$.
-Let us recall that the key difference between \kl{well-quasi-orderings} and the
-maybe more usual notion of well-founded orderings is that the latter allows the
-existence of infinite \intro{antichains}, that is, sets of pairwise
-incomparable elements. In fact, a set $X$ is \kl{well-quasi-ordered} if and
-only if it is \kl{well-founded} and does not contain infinite \kl{antichains}
-\cite{KRU72}.  As a consequence, our study focuses on detecting infinite
-\kl{antichains} in languages $L \subseteq \Sigma^*$. Remark that for the
-relations we consider (including the \kl{subword relation}), the set of all
-words of a fixed size $k$ is always an \kl{antichain}, and there are therefore
-no bound on the maximal length of an antichain.
+and the \kl{infix relation}.
+Specifically, for the \kl{prefix relation} and the \kl{suffix relation}, we are able to show the following decomposition into \intro{chains}, that is, totally ordered ascending sequences.
 
-In the case of the \kl{prefix relation}, we provide a complete characterization
-of languages $L \subseteq \Sigma^*$ that are \kl{well-quasi-ordered}, which are
-the simplest possible languages: they are precisely the \emph{finite} unions of
-\intro{chains}, that is, sets of words that are all comparable for the
-\kl{prefix relation}. Note that \kl{chains} are the simplest possible
+{
+\renewcommand{\labelenumi}{R\arabic{enumi}}
+\begin{enumerate}
+	\item A language $L \subseteq \Sigma^*$ is \kl{well-quasi-ordered} by the \kl{prefix relation} if and only if it is a finite union of \kl{chains} of the \kl{prefix relation}.
+	\item A language $L \subseteq \Sigma^*$ is \kl{well-quasi-ordered} by the \kl{suffix relation} if and only if it is a finite union of \kl{chains} of the \kl{suffix relation}.
+\end{enumerate}
+}
+
+Note that \kl{chains} are the simplest possible
 \kl{well-quasi-ordered} sets (they are totally ordered and well-founded) and it
@@
 decidability of the language $L$ itself, but heavily relies on the assumption
-that $\Sigma$ is finite. By symmetry, we obtain an analogue statement for the
-\kl{suffix relation}.
+that $\Sigma$ is finite.
 
-In the case of the \kl{infix relation}, the situation is more complex, and we
-show that \emph{any} countable quasi-ordering with finite initial segments can
-be \kl{embedded} into the \kl{infix relation} of a language over a binary
-alphabet $\Sigma \defined \set{a,b}$. As an example, on can identify
-$(\Sigma^*, \higleq)$ with a subset of $(\Sigma^*, \infleq)$ where $\higleq$ is
-the \kl{subword relation} and $\infleq$ is the \kl{infix relation}. This
-strongly contrasts with the \kl{prefix relation}, and suggests that a
-meaningful characterization of \kl{well-quasi-ordered} languages under the
-\kl{infix relation} is unlikely to be found.
+\AP One way to measure the complexity of a \kl{well-quasi-ordered} set is to
+consider its so-called \intro{ordinal invariants}. Several of these ordinal
+measures have been developed for well-quasi-orders. The \intro{maximal order
+type} (or m.o.t.) was originally defined by \cite{dejongh77} as the order type
+of the maximal linearization of a well-quasi-ordered set. Then,
+\cite{schmidt81} introduced the \intro{ordinal height} as the order type of a
+maximal chain of a well-quasi-ordered set. Finally, \cite{kriz90b} crafted the
+notion of \intro{ordinal width}, standardized the definitions of these three
+\kl{ordinal invariants}, and proved numerical relationships between them. For
+the \kl{prefix relation}, the \kl{suffix relation}, and the \kl{infix
+relation}, the \kl{ordinal height} is always bounded by $\omega$, and the key
+parameter to describe its complexity will be the \kl{ordinal width}, i.e.,
+``how large the antichains are.''
 
-We circumvent this issue by adding some regularity assumptions on the languages
-$L \subseteq \Sigma^*$ that we consider. This can be done by assuming that $L$
-is recognized by a finite automaton, or more generally that $L$ is the language
-of an \kl{amalgamation system}, a recently introduced notion that abstracts the
-notion of systems whose runs can be suitably ``pumped'' \cite{ASZZ24}. As
-examples \kl{amalgamation systems} covers the case of regular languages,
-context free languages, or even languages recognized by ``vector addition
-systems with state" \cite{ASZZ24}. Under this relatively mild assumption, we
-also prove that \kl{well-quasi-ordered} languages under the \kl{infix relation}
-are precisely the finite unions of \kl{chains}.
+\todo{rephrase}
+\AP In the case of the \kl{prefix relation} and the \kl{suffix relation}, we
+conclude from our characterization that the \kl{ordinal width} of a language
+$L$ is either finite or $L$ is not \kl{well-quasi-ordered}. The straightforward
+generalization of the results for the \kl{prefix relation} and the \kl{suffix
+relation} to the \kl{infix relation} is not possible. For example, Kuske has
+shown that any countable partial order with finite downward closures can be
+embedded into it~\cite{DBLP:journals/ita/Kuske06}. In particular, this means
+that the \kl{subword relation} can be embedded into the \kl{infix relation}.
+The former being a \kl{well-quasi-ordered} with a very large \kl{ordinal width}
+(see for example https://arxiv.org/pdf/2312.14587), this implies that such
+"easy characterizations" will not be possible in general. However, not all is
+lost: we show that with some mild restrictions we can recover much of the
+simple structure of the \kl[prefix relation]{prefix} and \kl{suffix relation}:
+
+{
+\renewcommand{\labelenumi}{R\arabic{enumi}}
+\begin{enumerate}
+	\setcounter{enumi}{2}
+    \item A \emph{bounded} language $L \subseteq \Sigma^*$ is \kl{well-quasi-ordered} by the \kl{infix relation} if and only if it is a finite union of languages $S_i \cdot P_i$, where each $S_i$ is \kl{well-quasi-ordered} by the \kl{suffix relation} and each $P_i$ is \kl{well-quasi-ordered} by the \kl{prefix relation}.
+\end{enumerate}
+}
+
+We illustrate how this result can be used to study languages recognized by
+finite automata, context-free grammars, petri nets, and other systems. This is
+possible by leveraging the recently introduced notion of \kl{amalgamation
+systems} \cite{ASZZ24}.
+
+{
+\renewcommand{\labelenumi}{R\arabic{enumi}}
+\begin{enumerate}
+	\setcounter{enumi}{2}
+	\item A language $L \subseteq \Sigma^*$ recognized by an \kl{amalgamation system} is \kl{well-quasi-ordered} by the \kl{infix relation} if and only if it is a finite union of languages $S_i \cdot P_i$, where each $S_i$ is \kl{well-quasi-ordered} by the \kl{suffix relation} and each $P_i$ is \kl{well-quasi-ordered} by the \kl{prefix relation}.
+\end{enumerate}
+}
+
+Finally, there has been recent interest in deciding whether a given order is a well-quasi-order.
+These questions have been investigated for example in the context of the verification of counter machines~\cite{DBLP:conf/fsttcs/FinkelG19} and in the context of logic~\cite{DBLP:journals/pacmpl/BergstrasserGLZ24}.
+We conclude our results with a contribution in this continuity.
+
+{
+\renewcommand{\labelenumi}{R\arabic{enumi}}
+\begin{enumerate}
+	\setcounter{enumi}{3}
+	\item For languages of amalgamation systems, it is decidable whether they are \kl{well-quasi-ordered} by the \kl[prefix relation]{prefix}, \kl[suffix relation]{suffix}, or \kl{infix relation}.
+\end{enumerate}
+}
 
@@
 
-\subparagraph{Outline} In \cref{prefixes:sec}, which is relatively
-self-contained, we study the \kl{prefix relation} prove in \cref{prefixes:thm}
-the characterization of \kl{well-quasi-ordered} languages by the \kl{prefix
-relation}. In \cref{infixes-regular:sec}, we prove that any countable quasi-ordering
-with finite initial segments can be embedded into the \kl{infix relation} of a
-language (\cref{infix-embedding:thm}), we obtain the analogue of
-\cref{prefixes:thm} for languages recognized by finite automata
-(\cref{infix-finite-automata:thm}). Finally, in \cref{infixes-amalgamation:sec}, we
-introduce \kl{amalgamation systems} and generalize
-\cref{infix-finite-automata:thm} to this setting
+\subparagraph{Outline} In
+\cref{prefixes:sec}, which is relatively
+self-contained, we study the \kl{prefix relation} and prove in
+\cref{prefixes:thm} the characterization of \kl{well-quasi-ordered}
+languages by the \kl{prefix relation}. In
+\cref{infixes-regular:sec}, we
+obtain the \kl[infix relation]{infix} analogue of \cref{prefixes:thm}
+specifically for regular languages
+(\cref{infix-finite-automata:thm}). Finally, we generalize this result to all
+\kl{amalgamation systems} in \cref{infixes-amalgamation:sec}
+in
 (\cref{infix-amalgamation:thm}).

### 2024-09-19 src/introduction.tex
work on intro/contribution

@@
 
-\AP Note that in the case of finite words and finite trees, the precise choice
-of ordering is crucial to ensure that the resulting structure is
-\kl{well-quasi-ordered}. The celebrated result of Higman states that the set of
-finite words over an ordered alphabet $(X, \preceq)$ is \kl{well-quasi-ordered}
-by the so-called \kl{subword embedding relation} \cite{HIG52}. Let us recall
-that the \kl{subword relation} for words over $(X, \preceq)$ is defined as
-follows: a word $u$ is a \intro{subword} of a word $v$, written $u
-\intro*\higleq v$, if there exists an increasing function $f \colon \{1,
-\ldots, |u|\} \to \{1, \ldots, |v|\}$ such that $u_i \preceq v_{f(i)}$ for all
-$i \in \{1, \ldots, |u|\}$.
+\AP
+Note that in the case of finite words and finite trees, the precise choice of
+ordering on words and trees is crucial to ensure that the resulting structure
+is \kl{well-quasi-ordered}. The celebrated result of Higman states that the set
+of finite words over an ordered alphabet $(X, \preceq)$ is
+\kl{well-quasi-ordered} by the so-called \kl{subword embedding relation}
+\cite{HIG52}. Let us recall that the \kl{subword relation} for words over $(X,
+\preceq)$ is defined as follows: a word $u$ is a \intro{subword} of a word $v$,
+written $u \intro*\higleq v$,
+if there exists an increasing function $f \colon \{1, \ldots, |u|\} \to \{1,
+\ldots, |v|\}$ such that $u_i \preceq v_{f(i)}$ for all $i \in \{1, \ldots,
+|u|\}$.
 
@@
     \centering
-    \begin{subfigure}[t]{0.48\textwidth}
-    	\centering
+    \begin{tabular}{p{3cm}p{8cm}}
+    Prefix Relation
+    &
     \includestandalone[width=\linewidth]{fig/prefix-embedding-standalone}
-    	\caption{Prefix Relation}
-   	\end{subfigure}%
-   	\hfill%
-   	\begin{subfigure}[t]{0.48\textwidth}
-   		\centering
+    \\
+    Infix relation
+    &
     \includestandalone[width=\linewidth]{fig/infix-embedding-standalone}
-   		\caption{Infix Relation}
-   	\end{subfigure}
-   	\begin{subfigure}[t]{0.48\textwidth}
-   		\centering
+    \\
+    Subword relation
+    &
     \includestandalone[width=\linewidth]{fig/subword-embedding-standalone}
-   		\caption{Subword Relation}
-   	\end{subfigure}
-
+    \end{tabular}
     \caption{A simple representation of the \kl{subword relation},
@@
 the equality relation. In this setting, the three alternatives we consider are
-the \intro{prefix relation} ($u \intro*\prefleq v$ if there exists $w$ with $uw
-= v$), the \intro{suffix relation} ($u \intro*\suffleq v$ if there exists $w$
-such that $wu = v$), and the \intro{infix relation} ($u \intro*\infleq v$ if
-there exists $w_1,w_2$ such that $w_1 u w_2 = v$). Note that these three
-relations straightforwardly generalize to infinite quasi-ordered alphabets.
-Unfortunately, it is easy to see that all of these constructions are not
-well-quasi-ordered as soon as the alphabet contains two non-equivalent letters:
-for instance, the infinite sequence $a^n b^n a^n$ is \kl{well-quasi-ordered} by
-the \kl{subword relation} but by neither the \kl{prefix relation}, nor the
-\kl{suffix relation}, nor the \kl{infix relation}.
+the \intro{prefix relation} ($u \intro*\prefleq v$ if there exists $w$ with
+$uw = v$), the \intro{suffix relation} ($u \intro*\suffleq v$ if there
+exists $w$ such that $wu = v$), and the \intro{infix relation} ($u
+\intro*\infleq v$ if there exists $w_1,w_2$ such that $w_1 u w_2 = v$). Note
+that these three relations straightforwardly generalize to infinite
+quasi-ordered alphabets. Unfortunately, it is easy to see that all of these
+constructions are not well-quasi-ordered as soon as the alphabet contains two
+non-equivalent letters.
 
@@
 
-\AP One way to measure the complexity of a \kl{well-quasi-ordered} set is to
-consider its so-called \intro{ordinal invariants}. Several of these ordinal
-measures have been developed for well-quasi-orders. The \intro{maximal order
-type} (or m.o.t.) was originally defined by \cite{dejongh77} as the order type
-of the maximal linearization of a well-quasi-ordered set. Then,
-\cite{schmidt81} introduced the \intro{ordinal height} as the order type of a
-maximal chain of a well-quasi-ordered set. Finally, \cite{kriz90b} crafted the
-notion of \intro{ordinal width}, standardized the definitions of these three
-\kl{ordinal invariants}, and proved numerical relationships between them. For
-the \kl{prefix relation}, the \kl{suffix relation}, and the \kl{infix
-relation}, the \kl{ordinal height} is always bounded by $\omega$, and the key
-parameter to describe its complexity will be the \kl{ordinal width}, i.e.,
-``how large the antichains are.''
-
-\todo{rephrase}
-\AP In the case of the \kl{prefix relation} and the \kl{suffix relation}, we
-conclude from our characterization that the \kl{ordinal width} of a language
-$L$ is either finite or $L$ is not \kl{well-quasi-ordered}. The straightforward
-generalization of the results for the \kl{prefix relation} and the \kl{suffix
-relation} to the \kl{infix relation} is not possible. For example, Kuske has
-shown that any countable partial order with finite downward closures can be
-embedded into it~\cite{DBLP:journals/ita/Kuske06}. In particular, this means
-that the \kl{subword relation} can be embedded into the \kl{infix relation}.
-The former being a \kl{well-quasi-ordered} with a very large \kl{ordinal width}
-(see for example https://arxiv.org/pdf/2312.14587), this implies that such
-"easy characterizations" will not be possible in general. However, not all is
-lost: we show that with some mild restrictions we can recover much of the
-simple structure of the \kl[prefix relation]{prefix} and \kl{suffix relation}:
-
-{
-\renewcommand{\labelenumi}{R\arabic{enumi}}
-\begin{enumerate}
-	\setcounter{enumi}{2}
-    \item A \emph{bounded} language $L \subseteq \Sigma^*$ is \kl{well-quasi-ordered} by the \kl{infix relation} if and only if it is a finite union of languages $S_i \cdot P_i$, where each $S_i$ is \kl{well-quasi-ordered} by the \kl{suffix relation} and each $P_i$ is \kl{well-quasi-ordered} by the \kl{prefix relation}.
-\end{enumerate}
-}
-
-We illustrate how this result can be used to study languages recognized by
-finite automata, context-free grammars, petri nets, and other systems. This is
-possible by leveraging the recently introduced notion of \kl{amalgamation
-systems} \cite{ASZZ24}.
+Such a straightforward characterization will however remain elusive for the \kl{infix relation}.
+For example, Kuske~\cite{DBLP:journals/ita/Kuske06} has shown that any countable partial order with finite downward closures can be embedded into it.
+Therefore, the structure of the \kl{infix relation} must be as least as complex as the \kl{subword relation}.
+However, not all is lost: we show that with some mild restrictions we can recover much of the simple structure of the \kl[prefix relation]{prefix} and \kl{suffix relation}:
 
@@
 
+This is not an overly onerous restriction: \kl[amalgamation system]{Amalgamation systems} are a recently introduced notion that 
+abstracts the notion of systems whose runs can be suitably ``pumped'' and include common language classes such as regular languages,
+context free languages, and even languages recognized by petri nets \cite{ASZZ24}.
+
 Finally, there has been recent interest in deciding whether a given order is a well-quasi-order.
@@
 
+\todo{verify this result :)}
 {
@@
 
-\subparagraph{Outline} In
-\cref{prefixes:sec}, which is relatively
-self-contained, we study the \kl{prefix relation} and prove in
-\cref{prefixes:thm} the characterization of \kl{well-quasi-ordered}
-languages by the \kl{prefix relation}. In
-\cref{infixes-regular:sec}, we
-obtain the \kl[infix relation]{infix} analogue of \cref{prefixes:thm}
-specifically for regular languages
+\subparagraph{Outline} In \cref{prefixes:sec}, which is relatively
+self-contained, we study the \kl{prefix relation} and prove in \cref{prefixes:thm}
+the characterization of \kl{well-quasi-ordered} languages by the \kl{prefix
+relation}. In \cref{infixes-regular:sec}, we obtain the \kl[infix relation]{infix} analogue of
+\cref{prefixes:thm} specifically for regular languages
 (\cref{infix-finite-automata:thm}). Finally, we generalize this result to all 
-\kl{amalgamation systems} in \cref{infixes-amalgamation:sec}
-in
+\kl{amalgamation systems} in \cref{infixes-amalgamation:sec} in 
 (\cref{infix-amalgamation:thm}).

### 2024-09-19 src/metadata.tex
metadata
adding git metadata to the generation
work on intro/contribution

@@
 \author{%
-        Roland Guttenberg
-    \thanks{Technical University of Munich, Germany}
+        Roland Guttenberg\thanks{Technical University of Munich,
+Germany}
      \and
-        Nathan Lhote
-    \thanks{Aix-Marseille University}
+        Nathan Lhote\thanks{Aix-Marseille University}
      \and
-        Aliaume Lopez
-    \thanks{University of Warsaw}
+        Aliaume Lopez\thanks{University of Warsaw}
      \and
-        Corto Mascle
-    \thanks{LaBRI, University of Bordeaux}
+        Corto Mascle\thanks{LaBRI, University of Bordeaux}
      \and
-        Vincent Michielini
-    \thanks{University of Warsaw}
+        Vincent Michielini\thanks{University of Warsaw}
      \and
-        Lia Schütze
-    \thanks{MPI-SWS, Germany}
+        Lia Schütze\thanks{MPI-SWS, Germany}
      \and
-        Omid Yaghoubi
-    \thanks{University of Warsaw}
+        Omid Yaghoubi\thanks{University of Warsaw}
     }
+
+\date{2024-09-05 17:28:15
++0200 -- 9cb13df663297a4b9b193d3efde15c3e328c4902 (main) at https://github.com/AliaumeL/word-ordering-wqo.git}
 
@@
     ordering. In this paper, we describe to what extent other natural
-    orderings (prefix, suffix, infix) on words can be used to derive
+    orderings (prefix, suffix, and infix) on words can be used to derive
     Higman-like theorems. More specifically, we are interested in
@@
     possible for the prefix and suffix orderings, and that under extra
-    regularity assumptions, this also lifts to the infix ordering.
+    regularity assumptions, this also extends to the infix ordering.
 \end{abstract}

### 2024-10-14 src/infixes.tex
adding recent proofs

@@
 \begin{corollary}
+    \label{reg-wqo-decidable:cor}
     Given a regular language $L$, it is decidable whether
@@
 \begin{corollary}
+    \label{reg-wqo-iff-dwclosed-wqo:cor}
     Given a regular language $L$, it is \kl{well-quasi-ordered} 
@@
 
-\section{Infixes and Amalgamation Systems}
+\section{Bounded Languages and Amalgamation Systems}
 \label{infixes-amalgamation:sec}
@@
 When proving \cref{infix-finite-automata:thm}, we have leveraged a powerful
-combinatorial argument from \cref{pumping-periods:lem}. It turns out that there
-is a rather large family of systems for which pumping arguments based on
-so-called \emph{minimal runs} exist: they are called \emph{amalgamation
-systems} and were recently introduced by Anad, Schmitz, Sch\"{u}tze, and
-Zetzsche \cite{ASZZ24}. Having done the heavy lifting on finite automata, the
-rest of this section is mostly devoted to the introduction of \kl{amalgamation
-systems} and collecting the necessary pumping argument they enjoy. Using this
-meta-proof, we will generalize \cref{infix-finite-automata:thm} to context-free
-grammars, languages recognized by vector addition systems with state (VASS),
-and more. The goal of this section is not to introduce all of these systems, or
-justify their usefulness, but to state and prove the following theorem.
+combinatorial argument on regular langugaes stated in
+\cref{pumping-periods:lem}. Our first remark is that this idea can
+be generalized to the larger class of \intro{bounded languages}, i.e.,
+languages $L \subseteq \Sigma^*$ such that there exists words $w_1, \dots, w_n$
+satisfying $L \subseteq w_1^* \cdots w_n^*$.
+
+\begin{theorem}
+    \label{bounded-language:thm}
+    Let $L$ be a \kl{bounded language} of $\Sigma^*$. Then,
+    $L$ is a \kl{well-quasi-order} when endowed with the 
+    \kl{infix relation} if and only if it is a finite union of \kl{chains}.
+\end{theorem}
+\begin{proof}
+    Let $w_1, \dots, w_n$ be such that
+    $L \subseteq w_1^* \cdots w_n^*$.
+    Let us define $m \defined \max \setof{\card{w_i}}{1 \leq i \leq n}$
+
+    Let $w[\vec{k}] \defined w_1^{k_1} \cdots w_n^{k_n}$ be a map from $\Nat^k$
+    to $\Sigma^*$. We are interested in the intersection of the image of $w$
+    with $L$. Let us assume for instance that for all $\vec{k} \in \Nat^n$,
+    there exists $\vec{\ell} \geq \vec{k}$ such that $w[\vec{\ell}] \in L$.
+    Then, leveraging \cref{pumping-periods:lem}, we conclude that there exists
+    $x,y$ of size at most $\max\setof{\card{w_i}}{1 \leq i \leq n}$ such that
+    $w[\vec{k}] \in \InfPeriodChain{x} \cup \InfPeriodChain{x}
+    \InfPeriodChain{y}$, and we conclude that $L \subseteq \InfPeriodChain{x}
+    \cup \InfPeriodChain{x} \InfPeriodChain{y}$.
+
+    Now, it may be the case that one cannot simultaneously assume that two
+    component of the vector $\vec{k}$ are unbounded. In general, given a set $S
+    \subseteq \set{1, \dots, n}$ of indices, we say that $S$ is admissible if
+    there exists a bound $N_0$ such that for all $\vec{b} \in \Nat^S$, there
+    exists a vector $\vec{k} \in \Nat^n$, such that $\vec{k}$ is greater than
+    $\vec{b}$ on the $S$ components, and the other components are below the
+    bound $N_0$. The language of an admissible set $S$ is the set of words
+    obtained by repeating $w_i$ at most $N_0$ times if it is not in $S$
+    ($w_i^{\leq N_0}$) and arbitrarily many times otherwise ($w_i^*$).
+    Note that $L \subseteq \bigcup_{S \text{ admissible }} L(S)$.
+
+    Now, admissible languages are ready to be pumped according to
+    \cref{pumping-periods:lem}. For every admissible language,
+    the size of a word that is not iterated is at most
+    $N_0 \times m$ by definition, and we conclude that:
+    \begin{equation*}
+        L \subseteq 
+        \bigcup_{x,y \in \Sigma^{\leq n}}
+        \bigcup_{u \in \Sigma^{\leq m \times N_0}}
+        \InfPeriodChain{x} u \InfPeriodChain{y}
+        \cup
+        \InfPeriodChain{x} u
+        \cup
+        u \InfPeriodChain{x}
+        \quad .
+    \end{equation*}
+\end{proof}
+
+\subsection{Amalgameation Systems}
+\label{infixes-amalgamation:subsec}
+
+It turns out that there is a rather large family of systems for which pumping
+arguments based on so-called \emph{minimal runs} exist: they are called
+\emph{amalgamation systems} and were recently introduced by Anad, Schmitz,
+Sch\"{u}tze, and Zetzsche \cite{ASZZ24}. Having done the heavy lifting on
+\kl{bounded languages}, the rest of this section is mostly devoted to the
+introduction of \kl{amalgamation systems} and collecting the necessary pumping
+argument they enjoy. Using this meta-proof, we will generalize
+\cref{infix-finite-automata:thm} to context-free grammars, languages recognized
+by vector addition systems with state (VASS), and more. The goal of this
+section is not to introduce all of these systems, or justify their usefulness,
+but to state and prove the following theorem.
 
@@
 
-In the upcoming proof of \cref{infix-amalgamation:thm}, we will use the notion
-of \intro{bounded language} for languages $L \subseteq \Sigma^*$ such that
-there exists words $w_1, \dots, w_n$ satisfying $L \subseteq w_1^* \cdots
-w_n^*$. This will be useful for us to derive the inclusion of a language $L$
-into a union of \kl{chains}. Note that we isolate this intermediate result in a
-separate theorem that could be applied beyond \kl{amalgamation systems}.
-
-\begin{theorem}
-    \label{bounded-language:thm}
-    Let $L$ be a \kl{bounded language} of $\Sigma^*$. Then,
-    $L$ is a \kl{well-quasi-order} when endowed with the 
-    \kl{infix relation} if and only if it is a finite union of \kl{chains}.
-\end{theorem}
-\begin{proof}
-    Let $w_1, \dots, w_n$ be such that
-    $L \subseteq w_1^* \cdots w_n^*$.
-    Let us define $m \defined \max \setof{\card{w_i}}{1 \leq i \leq n}$
-
-    Let $w[\vec{k}] \defined w_1^{k_1} \cdots w_n^{k_n}$ be a map from $\Nat^k$
-    to $\Sigma^*$. We are interested in the intersection of the image of $w$
-    with $L$. Let us assume for instance that for all $\vec{k} \in \Nat^n$,
-    there exists $\vec{\ell} \geq \vec{k}$ such that $w[\vec{\ell}] \in L$.
-    Then, leveraging \cref{pumping-periods:lem}, we conclude that there exists
-    $x,y$ of size at most $\max\setof{\card{w_i}}{1 \leq i \leq n}$ such that
-    $w[\vec{k}] \in \InfPeriodChain{x} \cup \InfPeriodChain{x}
-    \InfPeriodChain{y}$, and we conclude that $L \subseteq \InfPeriodChain{x}
-    \cup \InfPeriodChain{x} \InfPeriodChain{y}$.
-
-    Now, it may be the case that one cannot simultaneously assume that two
-    component of the vector $\vec{k}$ are unbounded. In general, given a set $S
-    \subseteq \set{1, \dots, n}$ of indices, we say that $S$ is admissible if
-    there exists a bound $N_0$ such that for all $\vec{b} \in \Nat^S$, there
-    exists a vector $\vec{k} \in \Nat^n$, such that $\vec{k}$ is greater than
-    $\vec{b}$ on the $S$ components, and the other components are below the
-    bound $N_0$. The language of an admissible set $S$ is the set of words
-    obtained by repeating $w_i$ at most $N_0$ times if it is not in $S$
-    ($w_i^{\leq N_0}$) and arbitrarily many times otherwise ($w_i^*$).
-    Note that $L \subseteq \bigcup_{S \text{ admissible }} L(S)$.
-
-    Now, admissible languages are ready to be pumped according to
-    \cref{pumping-periods:lem}. For every admissible language,
-    the size of a word that is not iterated is at most
-    $N_0 \times m$ by definition, and we conclude that:
-    \begin{equation*}
-        L \subseteq 
-        \bigcup_{x,y \in \Sigma^{\leq n}}
-        \bigcup_{u \in \Sigma^{\leq m \times N_0}}
-        \InfPeriodChain{x} u \InfPeriodChain{y}
-        \cup
-        \InfPeriodChain{x} u
-        \cup
-        u \InfPeriodChain{x}
-        \quad .
-    \end{equation*}
-\end{proof}
-
 To conclude this section and prove our main \cref{infix-amalgamation:thm}, it
@@
 \end{proof}
+
+\subsection{Downwards Closed Subsets}
+\label{infixes-downwards:subsec}
+
+In the light of \cref{reg-wqo-iff-dwclosed-wqo:cor} one may
+wonder whether the \kl{downwards closed} languages for the \kl{infix relation}
+satisfy a similar characterization in terms of union of \kl{chains} and
+\kl{grids} as stated in
+\cref{infix-finite-automata:thm,bounded-language:thm,infix-amalgamation:thm}.
+Let us first remark that from our previous results we obtain the following
+lemma that illustrates how all the previous characterizations of
+\kl{well-quasi-ordered} languages for the \kl{infix relation} are collapsing in
+the case of \kl{downwards closed} languages.
+
+\begin{lemma}
+    \label{dwclosed-infixes-wqo:lem}
+    Let $L \subseteq \Sigma^*$ be a \kl{downwards closed} language for the
+    \kl{infix relation} that is \kl{well-quasi-ordered}. Then, the following
+    are equivalent:
+    \begin{enumerate}
+        \item $L$ is \kl{bounded},
+        \item $L$ is \kl{regular},
+        \item $L$ is a finite union of \kl{chains} and \kl{grids}.
+    \end{enumerate}
+\end{lemma}
+\begin{proof}
+    TODO.
+\end{proof}
+
+One may think that all \kl{downwards closed} languages for the \kl{infix
+relation} that are \kl{well-quasi-ordered} are \kl{regular}. Note that this is
+what happens in the case of the \kl{subword embedding}, where any \kl{downwards
+closed} language for this relation is characterized by finitely many excluded
+patterns, hence \kl{regular} \textbf{TODO cite}.
+However, this is not the case for the \kl{infix relation}, as we will now
+illustrate with the following two examples.
+
+\begin{example}
+    Let $L \defined a^* b^* \cup b^* a^*$. This language is \kl{downwards
+    closed} for the \kl{infix relation}, is \kl{well-quasi-ordered} for the
+    \kl{infix relation}, but is characterized by an \emph{infinite} number 
+    of excluded infixes of the form $ab^ka$ and $ba^kb$ where $k \geq 1$.
+\end{example}
+
+Let $w$ be the Thue-Morse word \textbf{CITE}, and $L$ be the set of infixes of
+$w$. By construction $L$ is an \emph{infinite} \kl{downwards closed} for the
+\kl{infix relation}. Let us argue that $L$ is \kl{well-quasi-ordered} for the
+\kl{infix relation}, but is not \kl{regular}. First, $L$ cannot be \kl{bounded}
+because $w$ is \emph{cube free}, that is, no infix of $w$ can be written $uuu$
+for some word $u$ \textbf{CITE}. Furthermore, $w$ is a \intro{uniformly
+recurrent word} \textbf{CITE}, that is, for every infix $u$ of $w$, there
+exists $k \geq 1$ such that $u$ occurs in every infix $v$ of $w$ of size at
+least $k$. This implies that $L$ is \kl{well-quasi-ordered} for the \kl{infix
+relation}. We conclude thanks to \cref{dwclosed-infixes-wqo:lem},
+that $L$ is not \kl{regular}.

### 2024-10-14 src/metadata.tex
adding recent proofs

@@
      \and
-        Corto Mascle\thanks{LaBRI, University of Bordeaux}
-     \and
-        Vincent Michielini\thanks{University of Warsaw}
-     \and
         Lia Schütze\thanks{MPI-SWS, Germany}
-     \and
-        Omid Yaghoubi\thanks{University of Warsaw}
     }
 
-\date{2024-09-05 17:28:15
-+0200 -- 9cb13df663297a4b9b193d3efde15c3e328c4902 (main) at https://github.com/AliaumeL/word-ordering-wqo.git}
+\date{2024-10-14 10:28:01
++0200 -- 9f829fa27689f65f659ce09cd9273d1f894838fa (main) at git@github.com:AliaumeL/word-ordering-wqo.git}
 

### 2024-10-14 src/infixes.tex
Refactor + todos

@@
 \label{infixes-regular:sec}
+
+\todo{Delete this section: the \cref{infix-embedding:thm} is already
+    known, we should just restate it. The
+    \cref{infix-finite-automata:thm} is a direct consequence of
+    \cref{infix-amalgamation:thm} and \cref{bounded-language:thm}.
+    I don't know if we should keep it for clarity, but it does not 
+    seems to be necessary to me.
+}
 
@@
     Let $x,u,y \in \Sigma^*$ be words.  The following 
-    are finite unions of \kl{chains} for the infix relation:
+    are finite unions of \kl{grids} for the infix relation:
     $\InfPeriodChain{x}u$, $u \InfPeriodChain{x}$,
@@
 When proving \cref{infix-finite-automata:thm}, we have leveraged a powerful
-combinatorial argument on regular langugaes stated in
+combinatorial argument on regular langugages stated in
 \cref{pumping-periods:lem}. Our first remark is that this idea can
@@
     $L$ is a \kl{well-quasi-order} when endowed with the 
-    \kl{infix relation} if and only if it is a finite union of \kl{chains}.
+    \kl{infix relation} if and only if it is a finite union of \kl{grids}.
 \end{theorem}
@@
 
-\subsection{Amalgameation Systems}
+Note that this immediately generalizes \cref{reg-wqo-iff-dwclosed-wqo:cor}
+to all \kl{bounded languages}.
+
+\begin{corollary}
+    \label{bounded-wqo-dwclosed:cor}
+    Let $L$ be a \kl{bounded language} of $\Sigma^*$. Then,
+    $L$ is a \kl{well-quasi-order} when endowed with the
+    \kl{infix relation} if and only if $\dwset[\infleq]{L}$ is.
+\end{corollary}
+
+In the light of \cref{reg-wqo-iff-dwclosed-wqo:cor,bounded-wqo-dwclosed:cor}
+one may wonder whether the \kl{downwards closed}
+languages for the \kl{infix relation} satisfy a similar characterization in
+terms of union of \kl{grids} as stated in
+\cref{infix-finite-automata:thm,bounded-language:thm}. Let us
+first remark that from our previous results we obtain the following lemma that
+illustrates how all the previous characterizations of \kl{well-quasi-ordered}
+languages for the \kl{infix relation} are collapsing in the case of
+\kl{downwards closed} languages.
+
+\begin{lemma}
+    \label{dwclosed-infixes-wqo:lem}
+    Let $L \subseteq \Sigma^*$ be a \kl{downwards closed} language for the
+    \kl{infix relation} that is \kl{well-quasi-ordered}. Then, the following
+    are equivalent:
+    \begin{enumerate}
+        \item $L$ is \kl{bounded},
+        \item $L$ is \kl{regular},
+        \item $L$ is a finite union of \kl{chains} and \kl{grids}.
+    \end{enumerate}
+\end{lemma}
+\begin{proof}
+    TODO.
+\end{proof}
+
+One may think that all \kl{downwards closed} languages for the \kl{infix
+relation} that are \kl{well-quasi-ordered} are \kl{regular}. Note that this is
+what happens in the case of the \kl{subword embedding}, where any \kl{downwards
+closed} language for this relation is characterized by finitely many excluded
+patterns, hence \kl{regular}. However, this is not the case for the \kl{infix
+relation}, as we will now illustrate with the following two examples.
+
+\begin{example}
+    \label{dwclosed-wqo-not-finite-excl:ex}
+    Let $L \defined a^* b^* \cup b^* a^*$. This language is \kl{downwards
+    closed} for the \kl{infix relation}, is \kl{well-quasi-ordered} for the
+    \kl{infix relation}, but is characterized by an \emph{infinite} number 
+    of excluded infixes, respectively of the form $ab^ka$ and $ba^kb$ where $k \geq 1$.
+\end{example}
+
+To strengthen \cref{dwclosed-wqo-not-finite-excl:ex}, we will
+leverage the \intro{Thue-Morse sequence} $\intro*\ThueMorse \in
+\set{0,1}^{\Nat}$, which we will use as a black-box for its two main
+characteristics: it is \kl{cube-free} and \kl{uniformly recurrent}. Being
+\intro{cube-free} means that no (finite) word of the form $uuu$ is an
+\kl{infix} of $\ThueMorse$, and being \intro{uniformly recurrent} means that
+for every word $u$ that is an \kl{infix} of $\ThueMorse$, there exists $k \geq
+1$ such that $u$ is an \kl{infix} of every word $v$ of size at least $k$. We
+refer the reader to a nice survey of Allouche and Shallit for more information
+on this sequence and its properties \cite{ALSHA99}.
+
+\begin{lemma}
+    \label{thue-morse:lemma}
+    There exists a language $L$ that is \kl{downwards closed} for the \kl{infix
+    relation}, \kl{well-quasi-ordered} for the \kl{infix relation}, but is not
+    \kl{regular}.
+\end{lemma}
+\begin{proof}
+    Let $\ThueMorse$ be the \kl{Thue-Morse sequence}
+    $L$ be the set of infixes of $w$. By construction $L$ is an \emph{infinite}
+    \kl{downwards closed} for the \kl{infix relation}. Let us argue that $L$ is
+    \kl{well-quasi-ordered} for the \kl{infix relation}, but is not \kl{regular}.
+
+    Assume by contradiction that $L$ is \kl{bounded}. In this case, there exist
+    words $w_1, \dots, w_k \in \Sigma^*$ such that $L \subseteq w_1^* \cdots
+    w_k^*$. Since $L$ is infinite and \kl{downwards closed}, there exists a
+    word $u \in L$ such that $u = w_i^3$ for some $1 \leq i \leq k$. This is absurd
+    because $u \infleq \ThueMorse$, which is \kl{cube-free}.
+
+    Furthermore, $L$ is \kl{well-quasi-ordered} for the \kl{infix relation}.
+    Indeed, consider a sequence $\seqof{u_i}$ of words in $L$. Without loss of
+    generality, we may consider a subsequence such that $\card{u_i} <
+    \card{u_{i+1}}$ for all $i \in \Nat$. Because $\ThueMorse$ is \kl{uniformly
+    recurrent}, there exists $k \geq 1$ such that $u_1$ is an \kl{infix} of
+    every word $v$ of size at least $k$. In particular, $u_1$ is an \kl{infix}
+    of $u_k$, hence the sequence $\seqof{u_i}$ is \kl(wqo){good}.
+
+    We conclude thanks to
+    \cref{dwclosed-infixes-wqo:lem}, that $L$ is not \kl{regular}.
+\end{proof}
+
+\subsection{Amalgamation Systems}
 \label{infixes-amalgamation:subsec}
@@
 \begin{proof}
-    By definition, elements of $E(r,s)$ are defined as elements
+    By definition, the embeddings inside $E(r,s)$ are defined as elements
     of $\HigEmb(\canrun(r), \canrun(s))$, and they compose properly.
@@
 \end{proof}
-
-\subsection{Downwards Closed Subsets}
-\label{infixes-downwards:subsec}
-
-In the light of \cref{reg-wqo-iff-dwclosed-wqo:cor} one may
-wonder whether the \kl{downwards closed} languages for the \kl{infix relation}
-satisfy a similar characterization in terms of union of \kl{chains} and
-\kl{grids} as stated in
-\cref{infix-finite-automata:thm,bounded-language:thm,infix-amalgamation:thm}.
-Let us first remark that from our previous results we obtain the following
-lemma that illustrates how all the previous characterizations of
-\kl{well-quasi-ordered} languages for the \kl{infix relation} are collapsing in
-the case of \kl{downwards closed} languages.
-
-\begin{lemma}
-    \label{dwclosed-infixes-wqo:lem}
-    Let $L \subseteq \Sigma^*$ be a \kl{downwards closed} language for the
-    \kl{infix relation} that is \kl{well-quasi-ordered}. Then, the following
-    are equivalent:
-    \begin{enumerate}
-        \item $L$ is \kl{bounded},
-        \item $L$ is \kl{regular},
-        \item $L$ is a finite union of \kl{chains} and \kl{grids}.
-    \end{enumerate}
-\end{lemma}
-\begin{proof}
-    TODO.
-\end{proof}
-
-One may think that all \kl{downwards closed} languages for the \kl{infix
-relation} that are \kl{well-quasi-ordered} are \kl{regular}. Note that this is
-what happens in the case of the \kl{subword embedding}, where any \kl{downwards
-closed} language for this relation is characterized by finitely many excluded
-patterns, hence \kl{regular} \textbf{TODO cite}.
-However, this is not the case for the \kl{infix relation}, as we will now
-illustrate with the following two examples.
-
-\begin{example}
-    Let $L \defined a^* b^* \cup b^* a^*$. This language is \kl{downwards
-    closed} for the \kl{infix relation}, is \kl{well-quasi-ordered} for the
-    \kl{infix relation}, but is characterized by an \emph{infinite} number 
-    of excluded infixes of the form $ab^ka$ and $ba^kb$ where $k \geq 1$.
-\end{example}
-
-Let $w$ be the Thue-Morse word \textbf{CITE}, and $L$ be the set of infixes of
-$w$. By construction $L$ is an \emph{infinite} \kl{downwards closed} for the
-\kl{infix relation}. Let us argue that $L$ is \kl{well-quasi-ordered} for the
-\kl{infix relation}, but is not \kl{regular}. First, $L$ cannot be \kl{bounded}
-because $w$ is \emph{cube free}, that is, no infix of $w$ can be written $uuu$
-for some word $u$ \textbf{CITE}. Furthermore, $w$ is a \intro{uniformly
-recurrent word} \textbf{CITE}, that is, for every infix $u$ of $w$, there
-exists $k \geq 1$ such that $u$ occurs in every infix $v$ of $w$ of size at
-least $k$. This implies that $L$ is \kl{well-quasi-ordered} for the \kl{infix
-relation}. We conclude thanks to \cref{dwclosed-infixes-wqo:lem},
-that $L$ is not \kl{regular}.

### 2024-10-14 src/introduction.tex
work on intro/contribution (#1)
* work on intro/contribution

* Subfigures for intro graphics

---------

Co-authored-by: lschuetze <lschuetze@mpi-sws.org>
Refactor + todos

--- 2024-09-19 src/introduction.tex
@@
 \AP One way to measure the complexity of a \kl{well-quasi-ordered} set is to
-consider its so-called \intro{ordinal invariants}. Several of these ordinal
-measures have been developed for well-quasi-orders. The \intro{maximal order
-type} (or m.o.t.) was originally defined by \cite{dejongh77} as the order type
-of the maximal linearization of a well-quasi-ordered set. Then,
-\cite{schmidt81} introduced the \intro{ordinal height} as the order type of a
-maximal chain of a well-quasi-ordered set. Finally, \cite{kriz90b} crafted the
-notion of \intro{ordinal width}, standardized the definitions of these three
-\kl{ordinal invariants}, and proved numerical relationships between them. For
-the \kl{prefix relation}, the \kl{suffix relation}, and the \kl{infix
-relation}, the \kl{ordinal height} is always bounded by $\omega$, and the key
-parameter to describe its complexity will be the \kl{ordinal width}, i.e.,
-``how large the antichains are.''
+consider its so-called \intro{ordinal invariants}. For instance, the
+\intro{order type} of a \kl{well-quasi-ordered} set is the smallest ordinal
+that cannot be embedded into it. For the \kl{prefix relation} and the \kl{infix
+relation}, we conclude that the order type of a \kl{well-quasi-ordered} set is
+at most $k \times \omega$, and its width is finite.\todo{check this + cite}.
 
-\todo{rephrase}
-\AP In the case of the \kl{prefix relation} and the \kl{suffix relation}, we
-conclude from our characterization that the \kl{ordinal width} of a language
-$L$ is either finite or $L$ is not \kl{well-quasi-ordered}. The straightforward
-generalization of the results for the \kl{prefix relation} and the \kl{suffix
-relation} to the \kl{infix relation} is not possible. For example, Kuske has
-shown that any countable partial order with finite downward closures can be
-embedded into it~\cite{DBLP:journals/ita/Kuske06}. In particular, this means
-that the \kl{subword relation} can be embedded into the \kl{infix relation}.
-The former being a \kl{well-quasi-ordered} with a very large \kl{ordinal width}
-(see for example https://arxiv.org/pdf/2312.14587), this implies that such
-"easy characterizations" will not be possible in general. However, not all is
-lost: we show that with some mild restrictions we can recover much of the
-simple structure of the \kl[prefix relation]{prefix} and \kl{suffix relation}:
+\AP The straightforward generalization of the results for the \kl{prefix
+relation} and the \kl{suffix relation} to the \kl{infix relation} is not
+possible. For example, Kuske has shown that any countable partial order with
+finite downward closures can be embedded into
+it~\cite{DBLP:journals/ita/Kuske06}. In particular, this means that the
+\kl{subword relation} can be embedded into the \kl{infix relation}. The former
+being a \kl{well-quasi-ordered} with \kl{order type} $\omega^\omega$
+\todo{cite}. However, not all is lost: we show that with some mild restrictions
+we can recover much of the simple structure of the \kl[prefix relation]{prefix}
+and \kl{suffix relation}:
 
@@
 possible by leveraging the recently introduced notion of \kl{amalgamation
-systems} \cite{ASZZ24}.
+systems} \cite{ASZZ24}, which abstracts the notion of systems whose runs can be
+suitably ``pumped'' and include common language classes such as regular
+languages, context-free languages, and even languages recognized by petri nets.
 
@@
 
-Finally, there has been recent interest in deciding whether a given order is a well-quasi-order.
-These questions have been investigated for example in the context of the verification of counter machines~\cite{DBLP:conf/fsttcs/FinkelG19} and in the context of logic~\cite{DBLP:journals/pacmpl/BergstrasserGLZ24}.
-We conclude our results with a contribution in this continuity.
+Finally, there has been recent interest in deciding whether a given order is a
+well-quasi-order. These questions have been investigated for example in the
+context of the verification of counter
+machines~\cite{DBLP:conf/fsttcs/FinkelG19} and in the context of
+logic~\cite{DBLP:journals/pacmpl/BergstrasserGLZ24}. We conclude our results
+with a contribution in this continuity.
 
+\todo{verify this result :)}
 {

### 2024-10-16 src/infixes.tex
Draft: proof of decidability for infix

@@
 
-\begin{theorem}
+The following result due to Kuske shows the structural complexity of the \kl{infix relation}
+
+\begin{lemma}{\cite[Lemma 5.1]{DBLP:journals/ita/Kuske06}}
     \label{infix-embedding:thm}
     Let $(X, \preceq)$ be a \kl{partially ordered} set,
-    and $\Sigma \defined \set{a,b,c}$.
+    and $\Sigma$ be an alphabet with at least two letters.
     Then the following are equivalent:
@@
     \end{enumerate}
-\end{theorem}
-\begin{proof}
-    Let us first prove that \cref{infix-embedding-embeds:item} implies
-    \cref{infix-embedding-count:item}. Let $f \colon X \to \Sigma^*$ be an
-    \kl{embedding}. Note that $f$ is injective, because $X$ and $\Sigma^*$
-    are \kl{partially
-    ordered}. In particular, because $\dwset[\infleq]{w}$ is finite
-    for every $w \in \Sigma^*$, then so is $\dwset[\preceq]{x}$
-    for every $x \in X$. Similarly, $X$ must be countable
-    because $\Sigma^*$ is.
-
-    Conversely, let us prove that \cref{infix-embedding-count:item} implies
-    \cref{infix-embedding-embeds:item}. To that end, let us first define
-    inductively a sequence $\seqof{X_n}$ of subsets of $X$ as follows: $X_0$ is
-    the set of minimal elements of $X$, and $X_{n+1}$ is the set of minimal
-    elements of $X \setminus \bigcup_{i \leq n} X_i$. Because initial segments
-    of $X$ are finite, we know that $X = \bigcup_{n \in \Nat} X_n$. To
-    simplify notations, we will also use $Y_n \defined \bigcup_{i \leq n} X_i$.
-    We also let $L_0 \defined a b^* a$, and 
-    $L_{n+1} \defined L_n \cup c^{n+1} b^* c^{n+1} (L_n c^{n+1})^*$.
-    Now, let us inductively define \kl{embeddings} $f_n \colon (Y_n,\preceq) \to
-    (L_n, \infleq)$.
-
-    For the base case, let $\seqof{x_i}[i \in \Nat]$ be an enumeration of
-    $X_0$, which is possible because $X$ is countable. We define $f_0 \colon
-    X_0 \to \Sigma^*$ as follows: $f_0(x_i) \defined a b^i a$. It is immediate
-    that $f_0$ is an \kl{embedding}. For the inductive step, let $\seqof{x_i}[i
-    \in \Nat]$ be an enumeration of $X_{n+1}$, and assume that $f_n$ is an
-    \kl{embedding} from $Y_n$ to $\Sigma^*$.
-    Let us define $D_i \defined \setof{ y \in Y_n }{ y \preceq x_i }$
-    for $i \in \Nat$, all of which are finite by assumption.
-    We then let for all $i \in \Nat$:
-    \begin{equation*}
-        f_{n+1}(x_i) \defined 
-        \underbrace{c^{n+1} b^i c^{n+1} \prod_{y \in D_i} \overbrace{f_n(y)}^{\in L_n} c^{n+1}}_{\in L_{n+1}}
-        \quad .
-    \end{equation*}
-    And let $f_{n+1}(y) \defined f_n(y)$ for all $y \in Y_n$.
-
-    To check that $f_{n+1}$ is an \kl{embedding}, let us first notice that
-    words $f_{n+1}(x_i)$ and $f_{n+1}(x_j)$ are incomparable for the \kl{infix
-    relation} when $i \neq j$, because the only factor of the form $c^{n+1}
-    a^{n+1} b^t a^{n+1} c^{n+1}$ appears at the beginning of these words, and
-    precisely encode $i$ (resp. $j$) in the value of $t$. Now, let us consider
-    $y \in Y_n$ and $x_i \in X_{n+1}$ such that $y \preceq x_i$. Because $y \in
-    D_i$, we have $f_n(y) \infleq f_{n+1}(x_i)$ by construction, and since
-    $f_n(y) = f_{n+1}(y)$, we conclude that $f_{n+1}(y) \infleq f_{n+1}(x_i)$.
-    Conversely, assume that $f_{n+1}(y) \infleq f_{n+1}(x_i)$. Again, this
-    means that $f_n(y) \infleq f_{n+1}(x_i)$. Now, the encoding is robust
-    enough so that we can conclude $f_{n}(y) \infleq f_{n}(z)$ for some $z \in
-    D_i$ By induction hypothesis, $y \preceq z$, hence, $y \preceq x_i$.
-
-    We conclude by noticing that $\bigcup_{n \in \Nat} f_n$ is an \kl{embedding}
-    of $X$ into $\Sigma^*$.
-\end{proof}
-
+\end{lemma}
 \begin{figure}
@@
     \caption{Representation of the \kl{subword relation} for $\set{a,b}^*$
-        inside the \kl{infix relation} for $\set{a,b,c}^*$
+        inside the \kl{infix relation} for $\set{a,b,\#}^*$
         using the encoding of \cref{infix-embedding:thm}, restricted to words
@@
 \end{figure}
-
-The encoding using three distinct letters is clearer to use and to
-illustrate, but it is not surprising that one can encode a three
-letter alphabet into a two letter one, immediately
-leading to the following \cref{infix-embedding-binary:cor}.
-
-\begin{corollary}
-    \label{infix-embedding-binary:cor}
-    Let $(X, \preceq)$ be a \kl{partially ordered} set,
-    and $\Sigma \defined \set{a,b}$.
-    Then the following are equivalent:
-    \begin{enumerate}
-        \item \label{infix-embedding-embeds:item} 
-            $X$ \kl{embeds into} $(\Sigma^*, \infleq)$,
-        \item \label{infix-embedding-count:item}
-            $X$ is countable, and for every $x \in X$,
-            its \kl{downwards closure}
-            $\dwset[\preceq]{x}$ is finite.
-    \end{enumerate}
-\end{corollary}
-\begin{proof}
-    Let us construct an \kl{order embedding} from 
-    $(\set{a,b,c}^*, \infleq)$ to $(\set{0,1}, \infleq)$.
-    To that end, consider
-    the morphism $\mu$ that maps $a$ to $101$, $b$ to $1001$
-    and $c$ to $10001$.
-    Assume that $u \infleq v$ in $\set{a,b,c}^*$,
-    then $\mu(u) \infleq \mu(v)$ because $\mu$ is a morphism.
-    Conversely, assume that $\mu(u) \infleq \mu(v)$.
-    That, is, there exists an \kl{infix} $w$ of $\mu(v)$
-    such that $\mu(u) = w$.
-    Because $\mu(u)$ starts with $10$ and ends with $01$, so does
-    $w$. In particular, the word $w$ is compatible with $\mu$
-    in the sense that $w = \mu(s)$ for some $s \infleq v$.
-    Because the map $\mu$ is injective,
-    we conclude that $\mu(s) = \mu(u)$ implies $s = u$
-    and $u \infleq v$ as expected.
-\end{proof}
 
@@
 
+In the upcoming proof of \cref{infix-amalgamation:thm}, we will use the notion
+of \intro{bounded language} for languages $L \subseteq \Sigma^*$ such that
+there exists words $w_1, \dots, w_n$ satisfying $L \subseteq w_1^* \cdots
+w_n^*$. This will be useful for us to derive the inclusion of a language $L$
+into a union of \kl{chains}. Note that we isolate this intermediate result in a
+separate theorem that could be applied beyond \kl{amalgamation systems}.
+
+\begin{theorem}
+    \label{bounded-language:thm}
+    Let $L$ be a \kl{bounded language} of $\Sigma^*$. Then,
+    $L$ is a \kl{well-quasi-order} when endowed with the 
+    \kl{infix relation} if and only if it is a finite union of \kl{chains}.
+\end{theorem}
+\begin{proof}
+    Let $w_1, \dots, w_n$ be such that
+    $L \subseteq w_1^* \cdots w_n^*$.
+    Let us define $m \defined \max \setof{\card{w_i}}{1 \leq i \leq n}$
+
+    Let $w[\vec{k}] \defined w_1^{k_1} \cdots w_n^{k_n}$ be a map from $\Nat^k$
+    to $\Sigma^*$. We are interested in the intersection of the image of $w$
+    with $L$. Let us assume for instance that for all $\vec{k} \in \Nat^n$,
+    there exists $\vec{\ell} \geq \vec{k}$ such that $w[\vec{\ell}] \in L$.
+    Then, leveraging \cref{pumping-periods:lem}, we conclude that there exists
+    $x,y$ of size at most $\max\setof{\card{w_i}}{1 \leq i \leq n}$ such that
+    $w[\vec{k}] \in \InfPeriodChain{x} \cup \InfPeriodChain{x}
+    \InfPeriodChain{y}$, and we conclude that $L \subseteq \InfPeriodChain{x}
+    \cup \InfPeriodChain{x} \InfPeriodChain{y}$.
+
+    Now, it may be the case that one cannot simultaneously assume that two
+    component of the vector $\vec{k}$ are unbounded. In general, given a set $S
+    \subseteq \set{1, \dots, n}$ of indices, we say that $S$ is admissible if
+    there exists a bound $N_0$ such that for all $\vec{b} \in \Nat^S$, there
+    exists a vector $\vec{k} \in \Nat^n$, such that $\vec{k}$ is greater than
+    $\vec{b}$ on the $S$ components, and the other components are below the
+    bound $N_0$. The language of an admissible set $S$ is the set of words
+    obtained by repeating $w_i$ at most $N_0$ times if it is not in $S$
+    ($w_i^{\leq N_0}$) and arbitrarily many times otherwise ($w_i^*$).
+    Note that $L \subseteq \bigcup_{S \text{ admissible }} L(S)$.
+
+    Now, admissible languages are ready to be pumped according to
+    \cref{pumping-periods:lem}. For every admissible language,
+    the size of a word that is not iterated is at most
+    $N_0 \times m$ by definition, and we conclude that:
+    \begin{equation}
+    	\label{amalgamation-infix-wqo-inclusion:eqn}
+        L \subseteq 
+        \bigcup_{x,y \in \Sigma^{\leq n}}
+        \bigcup_{u \in \Sigma^{\leq m \times N_0}}
+        \InfPeriodChain{x} u \InfPeriodChain{y}
+        \cup
+        \InfPeriodChain{x} u
+        \cup
+        u \InfPeriodChain{x}
+        \quad .
+    \end{equation}
+\end{proof}
+
 To conclude this section and prove our main \cref{infix-amalgamation:thm}, it
@@
 
-The abstract definition of \kl{amalgamation system} can be enriched by
-effectiveness assumptions, as explained in \cite[Section 3.1.4]{ASZZ24}. In
-this setting, we conjecture that the following decidability result holds.
+For the next result, we impose a mild restriction by requiring the language class to be closed under rational transductions.
+This property holds for most common language classes, including all known \kl{amalgamation systems}.
 
-\begin{conjecture}
-    Let $L$ be given by an \kl{effective amalgamation system},
-    then it is decidable whether $L$ 
-    is \kl{well-quasi-ordered} by the \kl{infix relation}.
-\end{conjecture}
+\begin{theorem}
+	\label{infix-wqo-is-emptiness:thm}
+	Let $\mathcal{C}$ be a class of languages recognized by amalgamation systems and closed under rational transductions. Then the following are equivalent:
+
+	\begin{enumerate}
+		\item\label{wqo-infix-decidable} \kl[well-quasi-order]{Well-quasi-orderedness} of the \kl{infix relation} is decidable.
+		\item\label{emptiness-decidable} Emptiness is decidable.
+	\end{enumerate}
+\end{theorem}
+
+\begin{proof}
+	\cref{emptiness-decidable} $\Rightarrow$ \cref{wqo-infix-decidable}. We aim to make the inclusion test of \cref{amalgamation-infix-wqo-inclusion:eqn} effective.
+
+	Let $R = \bigcup_{x,y \in \Sigma^{\leq n}} \bigcup_{u \in \Sigma^{\leq m \times N_0}} \InfPeriodChain{x} u \InfPeriodChain{y} \cup \InfPeriodChain{x}u \cup u\InfPeriodChain{x}$.
+For any concrete values of the bounds $n$, $m$, and $N_0$, this language is regular.
+As $\mathcal{C}$ is closed under rational transductions, we can therefore reduce the inclusion to emptiness of $L \cap \overline{R}$.
+However, we need to find these bounds first.
+
+	To determine values for $n$ and $m$, we first test if $L$ is bounded.
+As the algorithm in \cite[Section 4.2]{ASZZ24} is effective, this yields words $w_1, \ldots w_n$ such that $L \subseteq w_1^* \cdots w_n^*$ and therefore yields also the bounds in questions.
+
+	To determine the value for $N_0$, we first computer maximal subsets $S \subseteq [n]$ such that $w_i$ for $i \in S$ are simultaneously unbounded.
+We do this by applying a transduction mapping $w_i$ to some fresh letter $a_i$ if $i \in S$ and $\varepsilon$ otherwise, and testing for simultaneous unboundedness \cite[Section 4.1]{ASZZ24}.
+Given a candidate bound $n_0$ for $N_0$, we construct the languages $L_{S,n_0} = w_1^{\circ_1}\cdots w_n^{\circ_n}$ where $\circ_i = *$ if $i \in S$ and $\circ_i = \leq n_0$ otherwise.
+We then test if $L \subseteq \bigcup_{S \text{ maximal}} L_{S,n_0}$.
+If yes, $n_0$ is our value for $N_0$, otherwise, we increase it and repeat the construction.
+As we chose maximal sets $S$, this procedure will eventually terminate.
+
+	\cref{wqo-infix-decidable} $\Rightarrow$ \cref{emptiness-decidable}. We 
+	consider the transduction $\mathcal T = \Sigma^* \times \{a, b\}$. Then 
+	for any language $L \in \mathcal C$, $\mathcal TL$ is well-quasi-ordered 
+	by infix if and only if $L$ is empty.
+\end{proof}
 

### 2024-10-17 src/introduction.tex
trying to update intro, in a bad way
new intro héhé

@@
 
-A \intro{well-quasi-ordered} set is a set $X$ equipped with a quasi-order
+\AP A \intro{well-quasi-ordered} set is a set $X$ equipped with a quasi-order
 $\preceq$ such that every infinite sequence $\seqof{x_n}$ of elements taken in
@@
 as a core combinatorial tool powering many termination arguments, and was
-successfully applied to the verification of infinite state transition systems
-\cite{ABDU96,ABDU98}. One of the appealing properties of well-quasi-orderings
-is that they are closed under many operations, such as taking products, finite
-unions, and finite powersets \cite{SCSC12}. Perhaps more surprisingly, the
-class of well-quasi-ordered sets is also stable under the operation of taking
-finite words and finite trees labelled by elements of a well-quasi-ordered set
+successfully applied to the verification of infinite state transition
+systems~\cite{ABDU96,ABDU98}. One of the appealing properties of
+well-quasi-orderings is that they are closed under many operations, such as
+taking products, finite unions, and finite powerset
+constructions~\cite{SCSC12}. Perhaps more surprisingly, the class of
+well-quasi-ordered sets is also stable under the operation of taking finite
+words and finite trees labelled by elements of a well-quasi-ordered set
 \cite{HIG52,KRU72}.
@@
 finite words over an ordered alphabet $(X, \preceq)$ is \kl{well-quasi-ordered}
-by the so-called \kl{subword embedding relation} \cite{HIG52}. Let us recall
+by the so-called \kl{subword embedding relation}~\cite{HIG52}. Let us recall
 that the \kl{subword relation} for words over $(X, \preceq)$ is defined as
@@
 \AP However, there are many other natural orderings on words that could be
-considered in the context of \kl{well-quasi-orderings}. Let us restrict
-ourselves to a simplified case where the alphabet is finite and equipped with
-the equality relation. In this setting, the three alternatives we consider are
-the \intro{prefix relation} ($u \intro*\prefleq v$ if there exists $w$ with $uw
-= v$), the \intro{suffix relation} ($u \intro*\suffleq v$ if there exists $w$
-such that $wu = v$), and the \intro{infix relation} ($u \intro*\infleq v$ if
-there exists $w_1,w_2$ such that $w_1 u w_2 = v$). Note that these three
-relations straightforwardly generalize to infinite quasi-ordered alphabets.
+considered in the context of \kl{well-quasi-orderings}, even in the simplified
+setting of a finite alphabet $\Sigma$ equipped with the equality relation. In
+this setting, the three alternatives we consider are the \intro{prefix
+relation} ($u \intro*\prefleq v$ if there exists $w$ with $uw = v$), the
+\intro{suffix relation} ($u \intro*\suffleq v$ if there exists $w$ such that
+$wu = v$), and the \intro{infix relation} ($u \intro*\infleq v$ if there exists
+$w_1,w_2$ such that $w_1 u w_2 = v$). Note that these three relations
+straightforwardly generalize to infinite quasi-ordered alphabets.
 Unfortunately, it is easy to see that all of these constructions are not
-well-quasi-ordered as soon as the alphabet contains two non-equivalent letters:
+well-quasi-ordered as soon as the alphabet contains two distinct letters:
 for instance, the infinite sequence $a^n b^n a^n$ is \kl{well-quasi-ordered} by
@@
 
-\AP While this dooms \kl[well-quasi-ordered]{well-quasi-orderedness} of these relations in the general case, there may be
-\emph{subsets} of $\Sigma^*$ which are \kl{well-quasi-ordered} by these
-relations. As a simple example, take the case of finite sets of (finite) words
-which are all \kl{well-quasi-ordered} regardless of the ordering considered.
-This raises the question of characterizing exactly which subsets $L \subseteq \Sigma^*$
-are \kl{well-quasi-ordered} with respect to the \kl{prefix relation}
-(respectively, the \kl{suffix relation} or the \kl{infix relation}), and
-providing suitable decision procedures.
+\AP While this dooms \kl[well-quasi-ordered]{well-quasi-orderedness} of these
+relations in the general case, there may be \emph{subsets} of $\Sigma^*$ which
+are \kl{well-quasi-ordered} by these relations. As a simple example, take the
+case of finite sets of (finite) words which are all \kl{well-quasi-ordered}
+regardless of the ordering considered. This raises the question of
+characterizing exactly which subsets $L \subseteq \Sigma^*$ are
+\kl{well-quasi-ordered} with respect to the \kl{prefix relation} (respectively,
+the \kl{suffix relation} or the \kl{infix relation}), and together with
+suitable decision procedures.
+
+\AP Let us argue that this question is not only of theoretical interest.
+Indeed, there has been recent breakthroughs in deciding whether a given order
+is a \kl{well-quasi-order}, for instance in the context of the verification of
+infinite state transition systems~\cite{DBLP:conf/fsttcs/FinkelG19} or in the
+context of logic~\cite{DBLP:journals/pacmpl/BergstrasserGLZ24}. Our decision
+procedures clearly contribute to this line of research in the context of words.
+Furthermore, a previous work by Kuske shows that any
+\emph{reasonable}\footnote{ This will be made precise in
+\cref{infix-embedding:thm}. } partially ordered set $(X, \leq)$ can
+be embedded into $\set{a,b}^*$ with the \kl{infix relation}~\cite[Lemma
+5.1]{DBLP:journals/ita/Kuske06}. Phrased differently, one can encode a large
+class of partially ordered sets as subsets of $\set{a,b}^*$. As a consequence,
+the following decision problem provides a reasonable abstract framework for
+deciding whether a given partially ordered set is \kl{well-quasi-ordered}:
+given a language $L \subseteq \Sigma^*$, decide whether $L$ is
+\kl{well-quasi-ordered} by the \kl{suffix relation}, or the \kl{infix
+relation}.
+
+\AP When considering an algorithm based on a \kl{well-quasi-ordering}, the
+runtime of the algorithm is deeply related to the ``complexity'' of the
+underlying quasi-order \todo{CITE???}. One way to measure this complexity is to
+consider its so-called \intro{ordinal invariants}: for instance, the
+\intro{maximal order type} (or m.o.t.), originally defined by De Jongh and
+Parikh \cite{dejongh77}, is the order type of the maximal linearization of a
+well-quasi-ordered set. In the case of a finite set, the m.o.t. is precisely
+the size of the set. Better runtime bounds were obtained by considering two
+other parameters\todo{cite sylvain width/ HDR?}: the \intro{ordinal height}
+introduced by Schmidt \cite{schmidt81}, and the \intro{ordinal width} of Kříž
+and Thomas \cite{kriz90b}. Therefore, when characterizing
+\kl{well-quasi-ordered} languages, we will also be interested in computing
+bounds on their \kl{ordinal invariants}. We refer to the
+\cref{ordinal-invariants:subsec} for a more detailed introduction to these
+parameters and ordinal computations in general. For the moment, let us just
+remark that all languages $L \subseteq \Sigma^*$ have an \kl{ordinal height} of
+at most $\omega$.
 
@@
 \kl{well-quasi-ordered} by the \kl{prefix relation}, the \kl{suffix relation},
-and the \kl{infix relation}.
-Specifically, for the \kl{prefix relation} and the \kl{suffix relation}, we are able to show the following decomposition into \intro{chains}, that is, totally ordered ascending sequences.
+and the \kl{infix relation}. Furthermore, we devise decision algorithms
+whenever the languages are given by reasonable computational models. Finally,
+we provide relatively low bounds on the possible \kl{ordinal invariants} of
+such languages.
+
+Specifically, for the \kl{prefix relation} and the
+\kl{suffix relation}, we are able to show the following decomposition into
+\intro{chains}, that is, totally ordered ascending sequences.
 
@@
 
-Note that \kl{chains} are the simplest possible
-\kl{well-quasi-ordered} sets (they are totally ordered and well-founded) and it
-is known that finite unions of \kl{well-quasi-ordered} sets are
-\kl{well-quasi-ordered}. As a consequence, the above characterization states
-that only the simplest possible \kl{well-quasi-ordered} sets are
-\kl{well-quasi-ordered} by the \kl{prefix relation}. Let us furthermore
-highlight that this characterization holds without any restriction on the
-decidability of the language $L$ itself, but heavily relies on the assumption
-that $\Sigma$ is finite.
+Note that \kl{chains} are the simplest possible \kl{well-quasi-ordered} sets
+(they are totally ordered and well-founded) and it is known that finite unions
+of \kl{well-quasi-ordered} sets are \kl{well-quasi-ordered}. As a consequence,
+the above characterization states that only the simplest possible
+\kl{well-quasi-ordered} sets are \kl{well-quasi-ordered} by the \kl{prefix
+relation}. Let us furthermore highlight that this characterization holds
+without any restriction on the decidability of the language $L$ itself, but
+heavily relies on the assumption that $\Sigma$ is finite. This allows us to
+derive the following bounds on the \kl{ordinal invariants} of such
+\kl{well-quasi-ordered} languages, which can be interpreted in two dual ways:
+first, these languages are extremely simple, which means that one could have
+hoped for using directly another well-quasi-ordered set without resorting to
+finite words; second, any time one encounters a \kl{well-quasi-ordered}
+language, it is going to provide relatively efficient algorithms.
 
-\AP One way to measure the complexity of a \kl{well-quasi-ordered} set is to
-consider its so-called \intro{ordinal invariants}. For instance, the
-\intro{order type} of a \kl{well-quasi-ordered} set is the smallest ordinal
-that cannot be embedded into it. For the \kl{prefix relation} and the \kl{infix
-relation}, we conclude that the order type of a \kl{well-quasi-ordered} set is
-at most $k \times \omega$, and its width is finite.\todo{check this + cite}.
+\renewcommand{\labelenumi}{R\arabic{enumi}}
+\begin{enumerate}
+    \setcounter{enumi}{2}
+    \item All languages $L \subseteq \Sigma^*$ are 
+        \kl{well-quasi-ordered} by the \kl{prefix relation} 
+        (resp. \kl{suffix relation}) have an \kl{ordinal height} of at most $\omega$,
+        a finite \kl{ordinal width}, and a \kl{maximal order type} strictly
+        below $\omega^2$.
+\end{enumerate}
+}
 
@@
 relation} and the \kl{suffix relation} to the \kl{infix relation} is not
-possible. For example, Kuske has shown that any countable partial order with
-finite downward closures can be embedded into
-it~\cite{DBLP:journals/ita/Kuske06}. In particular, this means that the
-\kl{subword relation} can be embedded into the \kl{infix relation}. The former
-being a \kl{well-quasi-ordered} with \kl{order type} $\omega^\omega$
-\todo{cite}. However, not all is lost: we show that with some mild restrictions
-we can recover much of the simple structure of the \kl[prefix relation]{prefix}
-and \kl{suffix relation}:
+possible. Indeed, it follows from Kuske's result that $\Sigma^*$ equipped with
+the \kl{subword relation} can be embedded into $\set{a,b}^*$ with the \kl{infix
+relation}~\cite[Lemma 5.1]{DBLP:journals/ita/Kuske06}. This implies that there
+are \kl{well-quasi-ordered} languages for the \kl{infix relation} that have
+very large \kl{ordinal invariants}: for instance, the \kl{maximal order type}
+of the \kl{subword relation} is $\omega^{\omega^{\card{\Sigma} - 1}}$, which
+equals its \kl{ordinal width} \cite[Corollary 3.9, Theorem 4.21]{DZSCSC20}.
+However, not all is lost if we restrict ourselves to \kl{bounded languages},
+i.e., languages that are included in some $w_1^* \cdots w_k^*$ for some finite
+choice of words $w_1, \ldots, w_k$. We are able to show the following:
 
@@
 \begin{enumerate}
-	\setcounter{enumi}{2}
-    \item A \emph{bounded} language $L \subseteq \Sigma^*$ is \kl{well-quasi-ordered} by the \kl{infix relation} if and only if it is a finite union of languages $S_i \cdot P_i$, where each $S_i$ is \kl{well-quasi-ordered} by the \kl{suffix relation} and each $P_i$ is \kl{well-quasi-ordered} by the \kl{prefix relation}.
+	\setcounter{enumi}{3}
+    \item A \kl{bounded language} $L \subseteq \Sigma^*$ is \kl{well-quasi-ordered} by the \kl{infix relation} if and only if it is a finite union of languages $S_i \cdot P_i$, where each $S_i$ 
+        is a \kl{chain} for the \kl{suffix relation},
+        and where
+        each $P_i$ is a \kl{chain} for the \kl{prefix relation}.
 \end{enumerate}
@@
 
-We illustrate how this result can be used to study languages recognized by
-finite automata, context-free grammars, petri nets, and other systems. This is
-possible by leveraging the recently introduced notion of \kl{amalgamation
-systems} \cite{ASZZ24}, which abstracts the notion of systems whose runs can be
-suitably ``pumped'' and include common language classes such as regular
-languages, context-free languages, and even languages recognized by petri nets.
-
+This result directly translates into bounds on the possible \kl{ordinal invariants}
+of such languages similarly as for the \kl{prefix relation}.
 {
@@
 \begin{enumerate}
-	\setcounter{enumi}{2}
-	\item A language $L \subseteq \Sigma^*$ recognized by an \kl{amalgamation system} is \kl{well-quasi-ordered} by the \kl{infix relation} if and only if it is a finite union of languages $S_i \cdot P_i$, where each $S_i$ is \kl{well-quasi-ordered} by the \kl{suffix relation} and each $P_i$ is \kl{well-quasi-ordered} by the \kl{prefix relation}.
+    \setcounter{enumi}{4}
+    \item All \kl{bounded languages}
+            $L \subseteq \Sigma^*$ that are \kl{well-quasi-ordered} by the \kl{infix relation}
+            have an \kl{ordinal height} of at most $\omega$,
+            an \kl{ordinal width} strictly below $\omega^2$, and
+            a \kl{maximal order type} strictly below $\omega^3$.
 \end{enumerate}
@@
 
-Finally, there has been recent interest in deciding whether a given order is a
-well-quasi-order. These questions have been investigated for example in the
-context of the verification of counter
-machines~\cite{DBLP:conf/fsttcs/FinkelG19} and in the context of
-logic~\cite{DBLP:journals/pacmpl/BergstrasserGLZ24}. We conclude our results
-with a contribution in this continuity.
-
-\todo{verify this result :)}
+\AP
+Finally, we provide a decision procedure for checking whether a language is
+\kl{well-quasi-ordered}, which takes as inputs languages recognized
+by \kl{amalgamation systems} \cite{ASZZ24}. Such systems will be
+formally introduced in \cref{amalgamation-systems:subsec}, but for the moment
+let us just say that they include many classical computational models such as
+finite automata, context-free grammars, and petri nets \cite{ASZZ24}. 
 {
@@
 \begin{enumerate}
-	\setcounter{enumi}{3}
-	\item For languages of amalgamation systems, it is decidable whether they are \kl{well-quasi-ordered} by the \kl[prefix relation]{prefix}, \kl[suffix relation]{suffix}, or \kl{infix relation}.
+    \setcounter{enumi}{5}
+    \item There is a decision procedure for checking whether a language $L$ recognized by an \kl{amalgamation system} is \kl{well-quasi-ordered} by the \kl{prefix relation}, the \kl{suffix relation}, or the \kl{infix relation}.
+Furthermore, if it is \kl{well-quasi-ordered}, then $L$ is a \kl{bounded language}.
 \end{enumerate}
 }
+Notice that the above result implies that the hypothesis of
+\kl{bounded languages} on the theoretical side is not a restriction at all in
+practice.
 

### 2024-10-17 src/amalgam.tex
restructuring paper

--- 2024-10-16 src/infixes.tex
@@
-% LTeX: language=en-GB 
-\section{Infixes and Regular Languages}
-\label{infixes-regular:sec}
-
-\todo{Delete this section: the \cref{infix-embedding:thm} is already
-    known, we should just restate it. The
-    \cref{infix-finite-automata:thm} is a direct consequence of
-    \cref{infix-amalgamation:thm} and \cref{bounded-language:thm}.
-    I don't know if we should keep it for clarity, but it does not 
-    seems to be necessary to me.
-}
-
-\AP
-In this section, we study languages equipped with the \kl{infix relation}. As
-opposed to the \kl{prefix} and \kl{suffix} relations, the \kl{infix relation}
-can yield to very complicated \kl{well-quasi-ordered} languages. Formally, the
-upcoming \cref{infix-embedding:thm} shows that \emph{any} countable
-partial-ordering with finite initial segments can be embedded into the infix
-relation of a language. To make the former statement precise,
-Let us recall that an \intro{order embedding} from a quasi-ordered set $(X,
-\preceq)$ into a quasi-ordered set $(Y, \preceq')$ is a function $f \colon X
-\to Y$ such that for all $x, y \in X$, $x \preceq y$ if and only if $f(x)
-\preceq' f(y)$. When such an embedding exists, we say that $X$ \reintro{embeds
-into} $Y$. We say that a quasi-ordered set $(X, \preceq)$ is a \intro{partial
-ordering} whenever the relation $\preceq$ is antisymmetric, that is $x \preceq
-y$ and $y \preceq x$ implies $x = y$.
-
-The following result due to Kuske shows the structural complexity of the \kl{infix relation}
-
-\begin{lemma}{\cite[Lemma 5.1]{DBLP:journals/ita/Kuske06}}
-    \label{infix-embedding:thm}
-    Let $(X, \preceq)$ be a \kl{partially ordered} set,
-    and $\Sigma$ be an alphabet with at least two letters.
-    Then the following are equivalent:
-    \begin{enumerate}
-        \item \label{infix-embedding-embeds:item} 
-            $X$ \kl{embeds into} $(\Sigma^*, \infleq)$,
-        \item \label{infix-embedding-count:item}
-            $X$ is countable, and for every $x \in X$,
-            its \kl{downwards closure}
-            $\dwset[\preceq]{x}$ is finite.
-    \end{enumerate}
-\end{lemma}
-\begin{figure}
-    \centering
-    \includestandalone[width=\linewidth]{fig/infix-encoding-standalone}
-    \caption{Representation of the \kl{subword relation} for $\set{a,b}^*$
-        inside the \kl{infix relation} for $\set{a,b,\#}^*$
-        using the encoding of \cref{infix-embedding:thm}, restricted to words
-        of length at most $3$. To obtain smaller words,
-        we replaced $D_i$ by $\max D_i$ in this construction.
-    }
-    \label{infix-embedding:fig}
-\end{figure}
-
-As a consequence of \cref{infix-embedding:thm}, we cannot replay proofs of
-\cref{prefixes:sec}, and will actually need to leverage some regularity of the
-languages to obtain a characterization of \kl{well-quasi-ordered} languages
-under the \kl{infix relation}. Let us first play this game for languages that
-are recognized by finite automata. We assume that the reader is familiar with
-the notions of deterministic finite automaton and regular languages, and refer
-to the book of Thomas for a comprehensive introduction to the subject
-\cite{THOM97}. The main goal of the remainder of this section is to prove the
-following \cref{infix-finite-automata:thm}.
-
-\begin{theorem}[restate=infix-finite-automata:thm,label=infix-finite-automata:thm]
-    \proofref{infix-finite-automata:thm}
-    Let $L \subseteq \Sigma^*$ be a language recognized by a finite automaton.
-    Then $L$ is \kl{well-quasi-ordered} by the \kl{infix relation} if and only if $L$ is
-    a finite union of \kl{chains} for the \kl{infix relation}.
-\end{theorem}
-
-\AP In order to prove \cref{infix-finite-automata:thm}, we will perform some
-preliminary analysis on the structure of an automaton recognizing a
-well-quasi-ordered language under the infix relation, which will be powered by
-folklore results on \emph{periodic} words. Let us recall that a non-empty word
-$w \in \Sigma^+$ is \intro(word){periodic} with period $x \in \Sigma^*$ if
-there exists a $p \in \Nat$ such that $w \infleq x^p$. The \intro{periodic
-length} of a word $u$ is the minimal length of a word $x$ such that $u$ is an
-\kl{infix} of $x^p$ for some $p \in \Nat$ and $x \in \Sigma^+$. We will
-essentially rely on the following result on periodic words.
-
-\begin{lemma}
-    \label{periodic-infixes:lem}
-    Let $u,v \in \Sigma^*$ be two (non-empty) \kl{periodic words}
-    having \kl{periodic lengths} $p$ and $q$ respectively.
-    Then, if $u \infleq v$ and $\card{u} \geq \factorial[p]{p \times q}$,
-    then $u$ and $v$ share the same \kl{periodic length}
-    $p = q$.
-\end{lemma}
-\begin{proof}
-    The fact that $u$ and $v$ are \kl{periodic length}
-    respectively $p$ and $q$ translates into the fact that $u_{i+p} = u_i$ and
-    $v_{i+q} = v_i$ for all indices $i \in \Nat$ such that those letters are
-    well-defined.
-
-    Now, assume that $u$ is an \kl{infix} of $v$, this provides the existence
-    of a $k \in \Nat$ such that $u = v_{k} \cdots v_{k + \card{u} - 1}$. In
-    particular, $v_{k+i+p} = v_{k+i}$ for all $i \in \Nat$ such that $k+i+p < k
-    + \card{u}$. Since we also have $v_{k+i+q} = v_{k+i}$ for all $1 \leq i
-    \leq \card{v} - k - q$. We conclude that both $u$ and $v$ are of
-    \kl{periodic length} the greatest common divisor of $p$ and $q$, and by
-    minimality of $q$ this must be equal to $q$ and to $p$.
-\end{proof}
-
-The upcoming \cref{powers-infixes:cor} is based on the observation that if $S$
-is a \kl{chain} for the \kl{suffix relation} and $P$ is a \kl{chain} for the
-\kl{prefix relation}, then $SP$ is a \kl{chain} for the \kl{infix relation}.
-
-\begin{corollary}
-    \label{powers-infixes:cor}
-    Let $u,v \in \Sigma^*$ and $k, \ell \in \Nat$
-    such that $k \geq \factorial[p]{\card{u} \times \card{v}}$,
-    $\ell \geq \factorial[p]{\card{v} \times \card{u}}$,
-    and $u^k \infleq v^\ell$.
-    Then, there exists $w \in \Sigma^*$ of size at most
-    $\min \set{\card{u}, \card{v}}$ and a $p \in \Nat$
-    such that
-    $u^k \infleq v^\ell \infleq w^p$.
-\end{corollary}
-
-The reason why \kl{periodic words} built using a given period $x \in \Sigma^+$
-are interesting for the \kl{infix relation} is that they naturally create
-\kl{chains}. Indeed, if $x \in \Sigma^+$ is a finite word, then $\setof{x^p}{p
-\in \Nat}$ is a \kl{chain} for the \kl{infix relation}. Note that in general,
-the \kl{downwards closure} of a chain is \emph{not} a chain. However, for the
-chains generated using periodic words, the \kl{downwards closure}
-$\dwset[\infleq]{\setof{x^p}{p \in \Nat}}$ is a \emph{finite union} of
-\kl{chains}. Because this set will appear in bigger equations, we introduce the
-shorter notation $\intro*\InfPeriodChain{x}$ for the set of \kl{infixes} of
-words of the form $x^p$, where $p$ ranges over $\Nat$.
-
-\begin{lemma}
-    \label{inf-period-chain:lem}
-    Let $x \in \Sigma^+$ be a word, and
-    Then $\InfPeriodChain{x}$ is a finite union of \kl{chains}
-    for the \kl{infix}, \kl{prefix} and \kl{suffix} relations 
-    simultaneously.
-\end{lemma}
-\begin{proof}
-    Let $x \in \Sigma^+$ be a word, and let $P_x$ be the (finite) set 
-    of all \kl{prefixes} of $x$, and $S_x$ be the (finite)
-    set of all \kl{suffixes} of $x$.
-    Assume that $w \in \InfPeriodChain{x}$, then $w = u x^p v$ for some
-    $u \in S_x$, $v \in P_x$, and $p \in \Nat$.
-    We have proven that
-    \begin{equation*}
-        \InfPeriodChain{x} \subseteq \bigcup_{u \in P_x} \bigcup_{v \in S_x} u x^* v
-        \quad .
-    \end{equation*}
-
-    Let us now demonstrate that for all $(u,v) \in S_x \times P_x$, the
-    language $u x^* v$ is a \kl{chain} for the \kl{infix}, \kl{suffix} and \kl{prefix} relations.
-    To that end,
-    let $(u,v) \in S_x \times P_x$ and $\ell, k \in \Nat$ be such that $\ell <
-    k$, let us prove that $u x^\ell v \infleq u x^k  v$. Because $v \prefleq
-    x$, we know that there exists $w$ such that $vw = x$. In particular,
-    $ux^\ell vw = u x^{\ell + 1}$, and because $\ell < k$, we conclude that $u
-    x^{\ell + 1} \prefleq u x^k v$. By transitivity, $u x^\ell v \prefleq u x^k
-    v$, and \emph{a fortiori}, $u x^\ell v \infleq u x^k v$. 
-    Similarly, because $u \suffleq x$,  there exists $w$ such that $wu  = x$, 
-    and we conclude that $u x^{\ell} v \suffleq w u x^\ell v = x^{\ell + 1} v \suffleq u x^k v$.
-    \qedhere
-\end{proof}
-
-\begin{corollary}
-    \label{inf-period-union-chains:lem}
-    Let $x,u,y \in \Sigma^*$ be words.  The following 
-    are finite unions of \kl{grids} for the infix relation:
-    $\InfPeriodChain{x}u$, $u \InfPeriodChain{x}$,
-    and $\InfPeriodChain{x} u \InfPeriodChain{y}$.
-\end{corollary}
-\begin{proof}
-    Because $\InfPeriodChain{x}$ is a finite union of \kl{chains} for the \kl{suffix}
-    relation (rsing \cref{inf-period-chain:lem}), we conclude that $\InfPeriodChain{x}u$ is one too for
-    the \kl{infix} relation. This holds similarly for $u \InfPeriodChain{x}$.
-    In the case of $\InfPeriodChain{x} u \InfPeriodChain{y}$,
-    we remark that $\InfPeriodChain{x} u$ is a finite union of chains
-    for the \kl{suffix relation},
-    and that $\InfPeriodChain{y}$ is one for the \kl{prefix relation}
-    (using again \cref{inf-period-chain:lem}).
-    As a consequence,
-    $\InfPeriodChain{x}u \InfPeriodChain{y}$ is a finite union 
-    of \kl{chains} for the \kl{infix} relation.
-\end{proof}
-
-The following combinatorial lemma connects the property of being
-\kl{well-quasi-ordered} to a property of the \kl{periodic lengths} of words in
-a language, based on the assumption that some factors can be iterated. It is
-the core result that powers the analysis done in the upcoming
-\cref{infix-finite-automata:thm,infix-amalgamation:thm}.
-
-\begin{lemma}
-    \label{pumping-periods:lem}
-    Let $L \subseteq \Sigma^*$ be a language
-    that is \kl{well-quasi-ordered} by the \kl{infix relation}.
-    Let $k \in \Nat$, $u_1, \cdots, u_{k+1} \in \Sigma^*$,
-    and $v_1, \cdots, v_{k} \in \Sigma^+$
-    be such that
-    $w[\vec{n}] \defined (\prod_{i = 1}^k u_i v_i^{n_i}) u_{k+1}$
-    belongs to $L$
-    for arbitrarily large values of $\vec{n} \in \Nat^k$.
-    Then, 
-    there exists $x,y \in \Sigma^+$ of size 
-    at most $\max \setof{\card{v_i}}{1 \leq i \leq k}$
-    such that 
-    one of the following holds for all
-    $\vec{n} \in \Nat^{k}$:
-    \begin{enumerate}
-        \item $w[\vec{n}] \in u_1 \InfPeriodChain{x}$,
-        \item $w[\vec{n}] \in \InfPeriodChain{x} u_{k+1}$,
-        \item $w[\vec{n}] \in \InfPeriodChain{x} u_i \InfPeriodChain{y}$
-            for some $1 \leq i \leq k + 1$.
-
-    \end{enumerate}
-\end{lemma}
-\begin{proof}
-    Note that the result is obvious if $k = 0$, and therefore
-    we assume $k \geq 1$ in the following proof.
-
-    Let us construct a sequence of words $\seqof{w_i}[i \in \Nat]$, where $w_i
-    \defined w[\vec{n_i}]$ for some well-chosen indices $\vec{n_i} \in \Nat^k$. The goal
-    being that 
-    if $w[\vec{n_i}]$ is an \kl{infix} of $w[\vec{n_j}]$,
-    then it can intersect at most \emph{two} iterated words,
-    with an intersection that is long enough to successfully apply
-    \cref{periodic-infixes:lem}.
-    In order to achieve this,
-    let us first define $s$ as the maximal size of a word $v_i$
-    ($1 \leq i \leq k$) and $u_j$ ($1 \leq j \leq k+1$).
-    Then,
-    we consider $\vec{n_0} \in \Nat^k$ such that $\vec{n_0}$ has all 
-    its components greater than $\factorial{s}$ and such that
-    $w[\vec{n_0}]$ belongs to $L$. 
-    Then, we inductively define 
-    $\vec{n_{i+1}}$  as the smallest vector of numbers greater than $\vec{n_i}$,
-    such that $w[\vec{n_{i+1}}]$ belongs to $L$, 
-    and with $\vec{n_i}$ having all components greater than
-    $2\card{w[\vec{n_i}]}$.
-
-    Let us assume that $k \geq 2$ in the following proof for symmetry purposes,
-    and argue later on that when $k = 1$ the same argument goes through.
-    Because $L$ is \kl{well-quasi-ordered} by the \kl{infix relation}, there
-    exists $i < j$ such that $w[\vec{n_i}]$ is an \kl{infix} of $w[\vec{n_j}]$.
-    Now, because of the chosen values for $\vec{n_j}$, there exists $1 \leq \ell \leq
-    k-1$ such that $w[\vec{n_i}]$ is actually an \kl{infix} of $u_{\ell}
-    v_{\ell}^{n_{j,\ell}} u_{\ell+1} v_{\ell+1}^{n_{j,\ell+1}} u_{\ell+2}$.
-    Even more,
-    one of the three following equations holds:
-    \begin{itemize}
-        \item $w[\vec{n_i}] \infleq v_{\ell}^{n_{j,\ell}} u_{\ell+1} v_{\ell+1}^{n_{j,\ell+1}}$,
-        \item $w[\vec{n_i}] \infleq u_{\ell}
-            v_{\ell}^{n_{j,\ell}}$,
-        \item $w[\vec{n_i}] \infleq
-            v_{\ell+1}^{n_{j,\ell+1}} u_{\ell+2}$.
-    \end{itemize}
-    In all those cases, we conclude using \cref{powers-infixes:cor}
-    that there exists $x,y \in \Sigma^+$ of size at most $s$, and 
-    a number $1 \leq t \leq k$ such that
-    $v_i^{n_i} \in \InfPeriodChain{x}$ for all $1 \leq i \leq t$,
-    and
-    $v_i^{n_i} \in \InfPeriodChain{y}$ for all $t < i \leq k$.
-    In particular,
-    $w[\vec{n_i}] \in \InfPeriodChain{x} u_{t} \InfPeriodChain{y}$.
-
-    When $k = 1$, the situation is a bit more specific since we only have two
-    cases: either $w_i \infleq u_1 v_1^{n_j}$ or $w_i \infleq v_1^{n_j} u_2$,
-    and we conclude with an identical reasoning.
-\end{proof}
-
-We are now ready to restate and prove our main theorem, proven by combining
-\cref{pumping-periods:lem} together with classical pumping arguments on finite
-state automata.
-
-\begin{proofof}{infix-finite-automata:thm}
-    Let $w \in L$, because $Q$ is finite, there exists
-    a factorization of $w$
-    into words $(\prod_{i = 1}^k u_i v_i) u_{k+1}$
-    such that 
-    for all $1 \leq i \leq k+1$, $\card{u_i} \leq \card{Q}$,
-    for all $1 \leq i \leq k$, $1 \leq \card{v_i} \leq \card{Q}$,
-    and satisfying 
-    that $w[\vec{X}] \defined 
-    (\prod_{i = 1}^k u_i v_i^{X_i}) u_{k+1}$
-    belongs to $L$ for all choices of values $\vec{X} \in \Nat^k$.
-
-    Applying \cref{pumping-periods:lem}, we conclude that 
-    there exists $x,y \in \Sigma^+$ of size at most $\card{Q}$
-    such that 
-    $w \in u_1 \InfPeriodChain{x} \cup \InfPeriodChain{y} u_{k+1}
-    \cup \bigcup_{1 \leq i \leq k+1} \InfPeriodChain{x} u_i \InfPeriodChain{y}$.
-    In particular, we conclude that
-    \begin{equation}
-        \label{infix-automata:eq}
-        L
-        \subseteq
-        \bigcup_{x,y,u\in \Sigma^{\leq \card{Q}}}
-        u \InfPeriodChain{x}
-        \cup 
-        \InfPeriodChain{x} u
-        \cup
-        \InfPeriodChain{x} u \InfPeriodChain{y}
-        \quad .
-    \end{equation}
-
-    Now, thanks to \cref{inf-period-union-chains:lem},
-    this happens to be a finite union of \kl{chains}
-    for the \kl{infix relation}.
-\end{proofof}
-
-\begin{corollary}
-    \label{reg-wqo-decidable:cor}
-    Given a regular language $L$, it is decidable whether
-    $L$ is \kl{well-quasi-ordered} for the \kl{infix relation}.
-\end{corollary}
-\begin{proof}
-    It suffices to decide whether 
-    \cref{infix-automata:eq} holds for the language $L$.
-    This is an inclusion of regular languages, which is decidable.
-\end{proof}
-
-Let $(X,\preceq)$ be a quasi-ordered set, and $L \subseteq X$ be such that $(L,
-\preceq)$ is \kl{well-quasi-ordered}. It is not true in general that
-$(\dwset{L}, \preceq)$ is \kl{well-quasi-ordered}. In the case of $(\Sigma^*,
-\infleq)$ a typical example is to start from an infinite \kl{antichain} $A$,
-together with an enumeration $\seqof{w_i}$ of $A$, and build the language $L
-\defined \setof{ \prod_{i = 0}^n w_i }{ i \in \Nat }$. By definition, $L$ is a
-\kl{chain} for the \kl{infix} ordering, hence \kl{well-quasi-ordered}. However,
-$\dwset[\infleq]{L}$ contains $A$, and is therefore not
-\kl{well-quasi-ordered}. As a fun corollary of our analysis, we conclude that
-this particular example cannot be realized by a regular language $A$.
-
-\begin{corollary}
-    \label{reg-wqo-iff-dwclosed-wqo:cor}
-    Given a regular language $L$, it is \kl{well-quasi-ordered} 
-    for the \kl{infix relation} if and only if 
-    $\dwset[\infleq] L$ is.
-\end{corollary}
-
-\section{Bounded Languages and Amalgamation Systems}
+\section{Amalgamation Systems and Decidability}
 \label{infixes-amalgamation:sec}
@@
 satisfying $L \subseteq w_1^* \cdots w_n^*$.
-
-\begin{theorem}
-    \label{bounded-language:thm}
-    Let $L$ be a \kl{bounded language} of $\Sigma^*$. Then,
-    $L$ is a \kl{well-quasi-order} when endowed with the 
-    \kl{infix relation} if and only if it is a finite union of \kl{grids}.
-\end{theorem}
-\begin{proof}
-    Let $w_1, \dots, w_n$ be such that
-    $L \subseteq w_1^* \cdots w_n^*$.
-    Let us define $m \defined \max \setof{\card{w_i}}{1 \leq i \leq n}$
-
-    Let $w[\vec{k}] \defined w_1^{k_1} \cdots w_n^{k_n}$ be a map from $\Nat^k$
-    to $\Sigma^*$. We are interested in the intersection of the image of $w$
-    with $L$. Let us assume for instance that for all $\vec{k} \in \Nat^n$,
-    there exists $\vec{\ell} \geq \vec{k}$ such that $w[\vec{\ell}] \in L$.
-    Then, leveraging \cref{pumping-periods:lem}, we conclude that there exists
-    $x,y$ of size at most $\max\setof{\card{w_i}}{1 \leq i \leq n}$ such that
-    $w[\vec{k}] \in \InfPeriodChain{x} \cup \InfPeriodChain{x}
-    \InfPeriodChain{y}$, and we conclude that $L \subseteq \InfPeriodChain{x}
-    \cup \InfPeriodChain{x} \InfPeriodChain{y}$.
-
-    Now, it may be the case that one cannot simultaneously assume that two
-    component of the vector $\vec{k}$ are unbounded. In general, given a set $S
-    \subseteq \set{1, \dots, n}$ of indices, we say that $S$ is admissible if
-    there exists a bound $N_0$ such that for all $\vec{b} \in \Nat^S$, there
-    exists a vector $\vec{k} \in \Nat^n$, such that $\vec{k}$ is greater than
-    $\vec{b}$ on the $S$ components, and the other components are below the
-    bound $N_0$. The language of an admissible set $S$ is the set of words
-    obtained by repeating $w_i$ at most $N_0$ times if it is not in $S$
-    ($w_i^{\leq N_0}$) and arbitrarily many times otherwise ($w_i^*$).
-    Note that $L \subseteq \bigcup_{S \text{ admissible }} L(S)$.
-
-    Now, admissible languages are ready to be pumped according to
-    \cref{pumping-periods:lem}. For every admissible language,
-    the size of a word that is not iterated is at most
-    $N_0 \times m$ by definition, and we conclude that:
-    \begin{equation*}
-        L \subseteq 
-        \bigcup_{x,y \in \Sigma^{\leq n}}
-        \bigcup_{u \in \Sigma^{\leq m \times N_0}}
-        \InfPeriodChain{x} u \InfPeriodChain{y}
-        \cup
-        \InfPeriodChain{x} u
-        \cup
-        u \InfPeriodChain{x}
-        \quad .
-    \end{equation*}
-\end{proof}
-
-Note that this immediately generalizes \cref{reg-wqo-iff-dwclosed-wqo:cor}
-to all \kl{bounded languages}.
-
-\begin{corollary}
-    \label{bounded-wqo-dwclosed:cor}
-    Let $L$ be a \kl{bounded language} of $\Sigma^*$. Then,
-    $L$ is a \kl{well-quasi-order} when endowed with the
-    \kl{infix relation} if and only if $\dwset[\infleq]{L}$ is.
-\end{corollary}
-
-In the light of \cref{reg-wqo-iff-dwclosed-wqo:cor,bounded-wqo-dwclosed:cor}
-one may wonder whether the \kl{downwards closed}
-languages for the \kl{infix relation} satisfy a similar characterization in
-terms of union of \kl{grids} as stated in
-\cref{infix-finite-automata:thm,bounded-language:thm}. Let us
-first remark that from our previous results we obtain the following lemma that
-illustrates how all the previous characterizations of \kl{well-quasi-ordered}
-languages for the \kl{infix relation} are collapsing in the case of
-\kl{downwards closed} languages.
-
-\begin{lemma}
-    \label{dwclosed-infixes-wqo:lem}
-    Let $L \subseteq \Sigma^*$ be a \kl{downwards closed} language for the
-    \kl{infix relation} that is \kl{well-quasi-ordered}. Then, the following
-    are equivalent:
-    \begin{enumerate}
-        \item $L$ is \kl{bounded},
-        \item $L$ is \kl{regular},
-        \item $L$ is a finite union of \kl{chains} and \kl{grids}.
-    \end{enumerate}
-\end{lemma}
-\begin{proof}
-    TODO.
-\end{proof}
-
-One may think that all \kl{downwards closed} languages for the \kl{infix
-relation} that are \kl{well-quasi-ordered} are \kl{regular}. Note that this is
-what happens in the case of the \kl{subword embedding}, where any \kl{downwards
-closed} language for this relation is characterized by finitely many excluded
-patterns, hence \kl{regular}. However, this is not the case for the \kl{infix
-relation}, as we will now illustrate with the following two examples.
-
-\begin{example}
-    \label{dwclosed-wqo-not-finite-excl:ex}
-    Let $L \defined a^* b^* \cup b^* a^*$. This language is \kl{downwards
-    closed} for the \kl{infix relation}, is \kl{well-quasi-ordered} for the
-    \kl{infix relation}, but is characterized by an \emph{infinite} number 
-    of excluded infixes, respectively of the form $ab^ka$ and $ba^kb$ where $k \geq 1$.
-\end{example}
-
-To strengthen \cref{dwclosed-wqo-not-finite-excl:ex}, we will
-leverage the \intro{Thue-Morse sequence} $\intro*\ThueMorse \in
-\set{0,1}^{\Nat}$, which we will use as a black-box for its two main
-characteristics: it is \kl{cube-free} and \kl{uniformly recurrent}. Being
-\intro{cube-free} means that no (finite) word of the form $uuu$ is an
-\kl{infix} of $\ThueMorse$, and being \intro{uniformly recurrent} means that
-for every word $u$ that is an \kl{infix} of $\ThueMorse$, there exists $k \geq
-1$ such that $u$ is an \kl{infix} of every word $v$ of size at least $k$. We
-refer the reader to a nice survey of Allouche and Shallit for more information
-on this sequence and its properties \cite{ALSHA99}.
-
-\begin{lemma}
-    \label{thue-morse:lemma}
-    There exists a language $L$ that is \kl{downwards closed} for the \kl{infix
-    relation}, \kl{well-quasi-ordered} for the \kl{infix relation}, but is not
-    \kl{regular}.
-\end{lemma}
-\begin{proof}
-    Let $\ThueMorse$ be the \kl{Thue-Morse sequence}
-    $L$ be the set of infixes of $w$. By construction $L$ is an \emph{infinite}
-    \kl{downwards closed} for the \kl{infix relation}. Let us argue that $L$ is
-    \kl{well-quasi-ordered} for the \kl{infix relation}, but is not \kl{regular}.
-
-    Assume by contradiction that $L$ is \kl{bounded}. In this case, there exist
-    words $w_1, \dots, w_k \in \Sigma^*$ such that $L \subseteq w_1^* \cdots
-    w_k^*$. Since $L$ is infinite and \kl{downwards closed}, there exists a
-    word $u \in L$ such that $u = w_i^3$ for some $1 \leq i \leq k$. This is absurd
-    because $u \infleq \ThueMorse$, which is \kl{cube-free}.
-
-    Furthermore, $L$ is \kl{well-quasi-ordered} for the \kl{infix relation}.
-    Indeed, consider a sequence $\seqof{u_i}$ of words in $L$. Without loss of
-    generality, we may consider a subsequence such that $\card{u_i} <
-    \card{u_{i+1}}$ for all $i \in \Nat$. Because $\ThueMorse$ is \kl{uniformly
-    recurrent}, there exists $k \geq 1$ such that $u_1$ is an \kl{infix} of
-    every word $v$ of size at least $k$. In particular, $u_1$ is an \kl{infix}
-    of $u_k$, hence the sequence $\seqof{u_i}$ is \kl(wqo){good}.
-
-    We conclude thanks to
-    \cref{dwclosed-infixes-wqo:lem}, that $L$ is not \kl{regular}.
-\end{proof}
 
@@
 
-In the upcoming proof of \cref{infix-amalgamation:thm}, we will use the notion
-of \intro{bounded language} for languages $L \subseteq \Sigma^*$ such that
-there exists words $w_1, \dots, w_n$ satisfying $L \subseteq w_1^* \cdots
-w_n^*$. This will be useful for us to derive the inclusion of a language $L$
-into a union of \kl{chains}. Note that we isolate this intermediate result in a
-separate theorem that could be applied beyond \kl{amalgamation systems}.
-
-\begin{theorem}
-    \label{bounded-language:thm}
-    Let $L$ be a \kl{bounded language} of $\Sigma^*$. Then,
-    $L$ is a \kl{well-quasi-order} when endowed with the 
-    \kl{infix relation} if and only if it is a finite union of \kl{chains}.
-\end{theorem}
-\begin{proof}
-    Let $w_1, \dots, w_n$ be such that
-    $L \subseteq w_1^* \cdots w_n^*$.
-    Let us define $m \defined \max \setof{\card{w_i}}{1 \leq i \leq n}$
-
-    Let $w[\vec{k}] \defined w_1^{k_1} \cdots w_n^{k_n}$ be a map from $\Nat^k$
-    to $\Sigma^*$. We are interested in the intersection of the image of $w$
-    with $L$. Let us assume for instance that for all $\vec{k} \in \Nat^n$,
-    there exists $\vec{\ell} \geq \vec{k}$ such that $w[\vec{\ell}] \in L$.
-    Then, leveraging \cref{pumping-periods:lem}, we conclude that there exists
-    $x,y$ of size at most $\max\setof{\card{w_i}}{1 \leq i \leq n}$ such that
-    $w[\vec{k}] \in \InfPeriodChain{x} \cup \InfPeriodChain{x}
-    \InfPeriodChain{y}$, and we conclude that $L \subseteq \InfPeriodChain{x}
-    \cup \InfPeriodChain{x} \InfPeriodChain{y}$.
-
-    Now, it may be the case that one cannot simultaneously assume that two
-    component of the vector $\vec{k}$ are unbounded. In general, given a set $S
-    \subseteq \set{1, \dots, n}$ of indices, we say that $S$ is admissible if
-    there exists a bound $N_0$ such that for all $\vec{b} \in \Nat^S$, there
-    exists a vector $\vec{k} \in \Nat^n$, such that $\vec{k}$ is greater than
-    $\vec{b}$ on the $S$ components, and the other components are below the
-    bound $N_0$. The language of an admissible set $S$ is the set of words
-    obtained by repeating $w_i$ at most $N_0$ times if it is not in $S$
-    ($w_i^{\leq N_0}$) and arbitrarily many times otherwise ($w_i^*$).
-    Note that $L \subseteq \bigcup_{S \text{ admissible }} L(S)$.
-
-    Now, admissible languages are ready to be pumped according to
-    \cref{pumping-periods:lem}. For every admissible language,
-    the size of a word that is not iterated is at most
-    $N_0 \times m$ by definition, and we conclude that:
-    \begin{equation}
-    	\label{amalgamation-infix-wqo-inclusion:eqn}
-        L \subseteq 
-        \bigcup_{x,y \in \Sigma^{\leq n}}
-        \bigcup_{u \in \Sigma^{\leq m \times N_0}}
-        \InfPeriodChain{x} u \InfPeriodChain{y}
-        \cup
-        \InfPeriodChain{x} u
-        \cup
-        u \InfPeriodChain{x}
-        \quad .
-    \end{equation}
-\end{proof}
-
 To conclude this section and prove our main \cref{infix-amalgamation:thm}, it
@@
 \end{proof}
-
-Let us conclude this section by fully characterizing what are
-actually the chains for the \kl{infix} relation, without any 
-regularity assumption.
-
-\begin{lemma}
-    \label{chains-infix:lem}
-    Let $L$ be a \kl{chain} for the \kl{infix relation}. Then, there exists
-    a \kl{chain} $P$ for the \kl{prefix relation} and a \kl{chain}
-    $S$ for the \kl{suffix relation} such that $L \subseteq SP$.
-\end{lemma}
-\begin{proof}
-    Because $L$ is a chain,
-    we have an enumeration $L = \seqof{w_n}$,
-    where $s_n \defined \card{w_n}$ is increasing, 
-    and functions
-    $f_{i,j} \colon \set{1, \dots, s_i} \to \set{1, \dots s_{j}}$
-    satisfying
-    that $f_{j,k} \circ f_{i, j} = f_{i, k}$,
-    $f_{i,i} = \mathrm{id}$, and
-    $w_i = w_j[f_{i,j}(1):f_{i,j}(s_i)]$ for all $i \leq j$.
-    Now, we can build the infinite word $w_\infty$ indexed by $\Rel$, and
-    embeddings $f_{i, \infty} \colon \set{1, \dots, s_i} \to \Rel$ such that
-    $w_i = w_\infty[f_{i, \infty}(1):f_{i, \infty}(s_i)]$, and $f_{i, \infty}
-    \circ f_{j, i} = f_{j, \infty}$ for all $i \leq j$.
-
-    Let us define $w_l$ as the (infinite) word $w_\infty[-\infty: f_{0,
-    \infty}(s_0)]$ and $w_r$ as the (infinite) word $w_\infty[f_{0,
-    \infty}(s_0): +\infty]$. It is now clear that every word $w_i$ can be
-    written as $w_i = w_{i,l} w_{i,r}$, where $w_{i,l}$ is a prefix of $w_r$
-    and $w_{i,r}$ is a suffix of $w_l$. Furthermore, the functions $f_{i,
-    \infty}$ can be used to notice that $\seqof{w_{i,r}}$ is a \kl{chain} for
-    the \kl{prefix relation}, and $\seqof{w_{i,l}}$ is a \kl{chain} for the
-    \kl{suffix relation}, which allows us to conclude.
-\end{proof}

### 2024-10-17 src/infixes.tex
restructuring paper

@@
 % LTeX: language=en-GB 
-\section{Infixes and Regular Languages}
+\section{Infixes and Bounded Languages}
 \label{infixes-regular:sec}
@@
 
-\section{Bounded Languages and Amalgamation Systems}
-\label{infixes-amalgamation:sec}
-
-When proving \cref{infix-finite-automata:thm}, we have leveraged a powerful
-combinatorial argument on regular langugages stated in
-\cref{pumping-periods:lem}. Our first remark is that this idea can
-be generalized to the larger class of \intro{bounded languages}, i.e.,
-languages $L \subseteq \Sigma^*$ such that there exists words $w_1, \dots, w_n$
-satisfying $L \subseteq w_1^* \cdots w_n^*$.
-
 \begin{theorem}
@@
     every word $v$ of size at least $k$. In particular, $u_1$ is an \kl{infix}
-    of $u_k$, hence the sequence $\seqof{u_i}$ is \kl(wqo){good}.
+    of $u_k$, hence the sequence $\seqof{u_i}$ is \kl(sequence){good}.
 
@@
 \end{proof}
-
-\subsection{Amalgamation Systems}
-\label{infixes-amalgamation:subsec}
-
-It turns out that there is a rather large family of systems for which pumping
-arguments based on so-called \emph{minimal runs} exist: they are called
-\emph{amalgamation systems} and were recently introduced by Anad, Schmitz,
-Sch\"{u}tze, and Zetzsche \cite{ASZZ24}. Having done the heavy lifting on
-\kl{bounded languages}, the rest of this section is mostly devoted to the
-introduction of \kl{amalgamation systems} and collecting the necessary pumping
-argument they enjoy. Using this meta-proof, we will generalize
-\cref{infix-finite-automata:thm} to context-free grammars, languages recognized
-by vector addition systems with state (VASS), and more. The goal of this
-section is not to introduce all of these systems, or justify their usefulness,
-but to state and prove the following theorem.
-
-\begin{theorem}[label=infix-amalgamation:thm,restate=infix-amalgamation:thm]
-    \proofref{infix-amalgamation:thm}
-    Let $L \subseteq \Sigma^*$ be a language recognized by an 
-    \kl{amalgamation system}.
-    Then $L$ is well-quasi-ordered by the infix relation if and only if $L$ is
-    a finite union of chains for the \kl{infix relation}.
-\end{theorem}
-
-Let us now formally introduce the notion of \kl{amalgamation systems}, and
-recall some results from \cite{ASZZ24} that will be useful for the proof of
-\cref{infix-amalgamation:thm}. The notion of \kl{amalgamation system} is
-tailored to produce \emph{pumping arguments}, which is exactly what our
-\cref{pumping-periods:lem} talks about. At the core of a pumping argument,
-there is a notion of \emph{run}, which could for instance be a sequence of
-transitions taken in a finite state automaton. Continuing on the analogy with
-finite automata, there is a natural ordering between runs, i.e., a run is
-smaller than another one if one can ``unroll loops" of the first run to obtain
-the second one. Typical pumping arguments then rely on the fact that
-\emph{minimal} runs are of finite size, and that all other runs are
-obtained by ``gluing" minimal runs together. This is exactly what
-\kl{amalgamation systems} are about.
-
-\AP Let us recall that over an alphabet $(\Sigma, =)$ a \kl{subword embedding}
-between two words $u \in \Sigma^*$ and $v \in \Sigma^*$ is a function $\rho
-\colon \range{\card{u}} \to \range{\card{v}}$ such that $u_i = v_{\rho(i)}$ for
-all $i \in \range{\card{u}}$. We write $\intro*\HigEmb(u,v)$ the set of all
-\kl{subword embeddings} between $u$ and $v$. It may be useful to notice that
-the set of finite words over $\Sigma$ forms a category when we consider
-\kl{subword embeddings} as morphisms, which is a fancy way to state that
-$\mathrm{id} \in \HigEmb(u,u)$ and that $f \circ g \in \HigEmb(u,w)$ whenever
-$g \in \HigEmb(u,v)$ and $f \in \HigEmb(v,w)$, for any choice of words
-$u,v,w \in \Sigma^*$.
-
-\AP Given a \kl{subword embedding} $f \colon u \to v$ between two words $u$ and
-$v$, there exists a unique decomposition $v = \GapWord{f}{0} u_1 \GapWord{f}{1}
-\cdots \GapWord{f}{k-1} u_k \GapWord{f}{k}$ where $\GapWord{f}{i} =
-v[f(i)+1:f(i+1)-1]$ for all $1 \leq i \leq k-1$, $\GapWord{f}{k} =
-v[f(k)+1:\card{v}]$, and $\GapWord{f}{0}   = v[1: f(1)-1]$. We say that
-$\intro*\GapWord{f}{i}$ is the $i$-th \intro{gap word} of $f$. We encourage the
-reader to look at \cref{gap-word-embedding:fig} to see an example of the
-\kl{gap words} resulting from a \kl{subword embedding} between two words. These
-\kl{gap words} will be useful to describe how and where runs of a system
-(described by words) can be combined.
-
-\begin{figure}
-    \centering
-    \includestandalone[width=\linewidth]{fig/gap-word-embedding-standalone}
-    \caption{The \kl{gap words} resulting from a \kl{subword embedding} between two 
-    finite words.}
-    \label{gap-word-embedding:fig}
-\end{figure}
-
-\begin{figure}
-    \centering
-    \includestandalone[width=\linewidth]{fig/run-amalgamation-standalone}
-    \caption{We illustrate how 
-        embeddings $f$ and $g$ between runs of an
-        \kl{amalgamation system} can be glued
-        together, seen on their canonical decomposition.
-    }
-    \label{amalgamation-runs:fig}
-\end{figure}
-
-\begin{definition}
-    An \intro{amalgamation system}
-    is a triple $(\Sigma, R, E, \canrun)$ where
-    $\Sigma$ is a finite alphabet,
-    $R$ is a set of so-called \emph{runs},
-    and 
-    $E$ is a set of so-called \emph{run embeddings},
-    and $\canrun \colon R \to (\Sigma \uplus \set{\varepsilon})^*$ is a 
-    function computing a \emph{canonical decomposition} of a run.
-    Given a run $r \in R$, and $i \in \range[0]{\card{\canrun(r)}}$, 
-    the \intro{gap language} of $r$ at position $i$ is $\GapLanguage{r}{i} \defined
-    \setof{\GapWord{f}{i}}{\exists s \in R. \exists f \in E(r,s) }$.
-    An \kl{amalgamation system} furthermore satisfies the following 
-    properties:
-    \begin{enumerate}
-        \item \emph{Word Embeddings as Morphisms.} For all $r, s \in R$,
-            $E(r,s) \subseteq \HigEmb(\canrun(r), \canrun(s))$.
-        \item \emph{$(R, E)$ Forms a Category.}
-            For all $r,s,t \in R$,
-            $\mathrm{id} \in E(r,r)$,
-            and whenever $f \in E(r,s)$ and $g \in E(s,t)$,
-            then $g \circ f \in E(r,t)$.
-        \item \emph{Well-Quasi-Ordered System.}
-            $(R, \leq_E)$ is a well-quasi-ordered set,
-            where $r \leq_E s$ if and only if $E(r,s) \neq \emptyset$.
-        \item \emph{Controlled Amalgamation.}
-            For all $r, s, t \in R$,
-            for all $f \in E(r,s)$,
-            for all $g \in E(r,t)$,
-            and for all $0 \leq i_0 \leq \card{\canrun(r)}$,
-            there exists a run $z \in R$ and morphisms
-            $h \in E(s,z)$ and $k \in E(t,z)$
-            satisfying
-            $h \circ f = k \circ g$,
-            $\GapWord{h \circ f}{i} = \GapWord{f}{i} \GapWord{g}{i}$
-            or 
-            $\GapWord{h \circ f}{i} = \GapWord{g}{i} \GapWord{f}{i}$
-            for every $0 \leq j \leq n$,
-            and
-            $\GapWord{h \circ f}{i_0} = \GapWord{f}{i_0} \GapWord{g}{i_0}$.
-            We refer to \cref{amalgamation-runs:fig} for an illustration 
-            of this property.
-    \end{enumerate}
-
-    The \intro{amalgamation language} of such a system
-    is the set of all words $w \in \Sigma^*$ such that
-    there exists a run $r \in R$
-    such that the concatenation of the letters of its
-    canonical decomposition, written $\intro*\yieldrun(r)$,
-    equals $w$.
-\end{definition}
-
-Intuitively, the definition of an amalgamation system allows the comparison of
-runs, and the proper ``gluing" of runs together to obtain new runs. Let us
-recall some examples of languages that can be recognized by \kl{amalgamation
-systems} given by the authors of the original paper: regular languages
-\cite[Theorem 5.3]{ASZZ24}, VASS as a consequence of \cite[Theorem
-5.5]{ASZZ24}, context-free languages as a consequence of \cite[Theorem
-5.10]{ASZZ24}. For this paper to be self-contained, we will also recall how
-runs of a finite state automaton can be understood as an \kl{amalgamation
-system}.
-
-\begin{example}[{\cite[Section 3.2]{ASZZ24}}]
-    Let $A = (Q, \delta, q_0, F)$ be a finite state automaton over a finite
-    alphabet $\Sigma$. Let $\Delta$ be the set of transitions $(q_1, a, q_2)
-    \in Q \times \Sigma \times Q$,
-    and $R \subseteq \Delta^*$ be the set of 
-    words over transitions that start with the initial state $q_0$,
-    end in a final state $q_f \in F$, and such that the end state of a
-    letter is the start state of the following one.
-    The canonical decomposition $\canrun \colon R \to \Sigma^*$
-    is defined as a morphism from $\Delta^*$ to $\Sigma^*$
-    that maps $(q,a,p)$ to $a$.
-    Finally, given two runs $r$ and $s$ of the automaton,
-    we say that an embedding $f \in \HigEmb(\canrun(r), \canrun(s))$
-    belongs to $E(r,s)$ when
-    $f$ is also defining an embedding from $r$ to $s$ as words in $\Delta^*$,
-    where $\Delta$ is equipped with the equality relation.
-
-    The system $(\Sigma, R, E, \canrun)$ is an \kl{amalgamation system},
-    whose language is precisely the language of words recognized
-    by the automaton $A$.
-\end{example}
-\begin{proof}
-    By definition, the embeddings inside $E(r,s)$ are defined as elements
-    of $\HigEmb(\canrun(r), \canrun(s))$, and they compose properly.
-    Because $\Delta = Q \times \Sigma \times Q$ is finite, it is 
-    a \kl{well-quasi-ordering} when equipped with the equality relation, and 
-    we conclude that $\Delta^*$ with $\higleq$ is a \kl{well-quasi-order}
-    according to Higman’s Lemma \cite{HIG52}.
-
-    Let us now move to proving that the system satisfies the amalgamation
-    property. Given three runs $r,s,t \in R$, and two embeddings $f \in E(r,s)$
-    and $g \in E(r,t)$, we want to construct an amalgamated run $s \vee t$.
-    Because letters in the run $r$ respect the transitions of the automaton
-    (i.e., if the letter $i$ ends in state $q$, then the letter $i+1$ starts in
-    state $q$), then the \kl{gap word} at position $i$ starts in state $q$ and
-    ends in state $q$ too. This means that for both embeddings
-    $f$ and $g$, the \kl{gap words} are read by the automaton by looping
-    on a state. In particular, these loops can be taken in any order
-    and continue to represent a valid run. That is, we can even select
-    the order of concatenation in the amalgamation for \emph{all} 
-    $0 \leq i \leq \card{\canrun(r)}$ and not just for one separately.
-
-    Let us now remark that 
-    the language of this amalgamation system is
-    the set of $\yieldrun(r)$ when $r$ ranges over $R$,
-    and because $R$ is the set of valid runs of the automaton,
-    and $\yieldrun(r)$ is the word read along this run,
-    we immediately conclude.
-\end{proof}
-
-Let us now introduce a combinatorial lemma that explains how \kl{gap languages}
-can be pumped. Note that a single \kl{gap language} is always stable under
-concatenation, and our concern will be on a potential description of
-\emph{simultaneously} pumping different \kl{gap languages} to produce valid
-runs.
-
-\begin{lemma}
-    \label{pumping-gap-languages:lem}
-    Let $(\Sigma, R, E, \canrun)$ be an \kl{amalgamation system}, let $r \in R$
-    be a run of size $n \in \Nat$, and let $s, t \in R$ be two runs
-    with embeddings $f \in E(r,s)$ and $g \in E(r,t)$.
-    Let $\seqof{u_i}[0 \leq i \leq n]$ be the sequence of \kl{gap words}
-    for $f$, that is, $u_i =
-    \GapWord{f}{i}$ for all $0 \leq i \leq n$. Similarly, let
-    $\seqof{v_i}[0 \leq i \leq n]$ be the sequence of \kl{gap words} for $g$.
-
-    Then, for all $k \in \Nat$, and $0 < \ell < n$, the following words belong to 
-    the language of the \kl{amalgamation system}.
-    \begin{enumerate}
-        \item $(\prod_{i = 0}^{\ell - 1} u_i^{x_i} v_i u_i^{y_i} v_i u_i^{z_i}) 
-               v_\ell u_\ell^k v_\ell
-               (\prod_{i = \ell+1}^{n} u_i^{x_i} v_i u_i^{y_i} v_i u_i^{z_i})$,
-        \item $(\prod_{i = 0}^{n - 1} u_i^{x_i} v_i u_i^{y_i}) 
-               v_n u_n^k$,
-        \item $u_0^k v_0 
-            (\prod_{i = 1}^{n} u_i^{x_i} v_i u_i^{y_i})$.
-    \end{enumerate}
-    Where, for all $0 \leq i \leq n$, $x_i + y_i + z_i = k$, with $z_i = 0$
-    in the two last items.
-\end{lemma}
-\begin{proof}
-    The result immediately follows from the embedding property
-    applied inductively on runs obtained by gluing 
-    $k$ times $s$ with one or two times $t$ (depending on the item number).
-\end{proof}
-
-In the upcoming proof of \cref{infix-amalgamation:thm}, we will use the notion
-of \intro{bounded language} for languages $L \subseteq \Sigma^*$ such that
-there exists words $w_1, \dots, w_n$ satisfying $L \subseteq w_1^* \cdots
-w_n^*$. This will be useful for us to derive the inclusion of a language $L$
-into a union of \kl{chains}. Note that we isolate this intermediate result in a
-separate theorem that could be applied beyond \kl{amalgamation systems}.
-
-\begin{theorem}
-    \label{bounded-language:thm}
-    Let $L$ be a \kl{bounded language} of $\Sigma^*$. Then,
-    $L$ is a \kl{well-quasi-order} when endowed with the 
-    \kl{infix relation} if and only if it is a finite union of \kl{chains}.
-\end{theorem}
-\begin{proof}
-    Let $w_1, \dots, w_n$ be such that
-    $L \subseteq w_1^* \cdots w_n^*$.
-    Let us define $m \defined \max \setof{\card{w_i}}{1 \leq i \leq n}$
-
-    Let $w[\vec{k}] \defined w_1^{k_1} \cdots w_n^{k_n}$ be a map from $\Nat^k$
-    to $\Sigma^*$. We are interested in the intersection of the image of $w$
-    with $L$. Let us assume for instance that for all $\vec{k} \in \Nat^n$,
-    there exists $\vec{\ell} \geq \vec{k}$ such that $w[\vec{\ell}] \in L$.
-    Then, leveraging \cref{pumping-periods:lem}, we conclude that there exists
-    $x,y$ of size at most $\max\setof{\card{w_i}}{1 \leq i \leq n}$ such that
-    $w[\vec{k}] \in \InfPeriodChain{x} \cup \InfPeriodChain{x}
-    \InfPeriodChain{y}$, and we conclude that $L \subseteq \InfPeriodChain{x}
-    \cup \InfPeriodChain{x} \InfPeriodChain{y}$.
-
-    Now, it may be the case that one cannot simultaneously assume that two
-    component of the vector $\vec{k}$ are unbounded. In general, given a set $S
-    \subseteq \set{1, \dots, n}$ of indices, we say that $S$ is admissible if
-    there exists a bound $N_0$ such that for all $\vec{b} \in \Nat^S$, there
-    exists a vector $\vec{k} \in \Nat^n$, such that $\vec{k}$ is greater than
-    $\vec{b}$ on the $S$ components, and the other components are below the
-    bound $N_0$. The language of an admissible set $S$ is the set of words
-    obtained by repeating $w_i$ at most $N_0$ times if it is not in $S$
-    ($w_i^{\leq N_0}$) and arbitrarily many times otherwise ($w_i^*$).
-    Note that $L \subseteq \bigcup_{S \text{ admissible }} L(S)$.
-
-    Now, admissible languages are ready to be pumped according to
-    \cref{pumping-periods:lem}. For every admissible language,
-    the size of a word that is not iterated is at most
-    $N_0 \times m$ by definition, and we conclude that:
-    \begin{equation}
-    	\label{amalgamation-infix-wqo-inclusion:eqn}
-        L \subseteq 
-        \bigcup_{x,y \in \Sigma^{\leq n}}
-        \bigcup_{u \in \Sigma^{\leq m \times N_0}}
-        \InfPeriodChain{x} u \InfPeriodChain{y}
-        \cup
-        \InfPeriodChain{x} u
-        \cup
-        u \InfPeriodChain{x}
-        \quad .
-    \end{equation}
-\end{proof}
-
-To conclude this section and prove our main \cref{infix-amalgamation:thm}, it
-suffices to demonstrate that languages recognized by \kl{amalgamation systems}
-that are \kl{well-quasi-ordered} for the \kl{infix relation} are \kl{bounded
-languages}.
-
-\begin{lemma}
-    \label{infix-amalgamation-bounded:lem}
-    Let $L \subseteq \Sigma^*$ be a language recognized
-    by an \kl{amalgamation system} that is \kl{well-quasi-ordered}
-    for the \kl{infix relation}. Then $L$ is a \kl{bounded language}.
-\end{lemma}
-\begin{proof}
-    Assume that $L$ is \kl{well-quasi-ordered} by the \kl{infix relation},
-    and obtained by an \kl{amalgamation system}
-    $(\Sigma, R, E, \canrun)$.
-
-    Let us consider the set $M$ of minimal runs for the relation $\leq_E$,
-    which is finite because the latter is a \kl{well-quasi-ordering}. By
-    definition, for every word $w \in L$, there exists $r \in M$, $s \in R$,
-    and $f \in E(r,s)$ such that $w = \canrun(s)$.
-    In particular, we conclude that
-    \begin{equation*}
-        L \subseteq \bigcup_{r \in M} 
-        \left(
-        \GapLanguage{r}{0}
-        \prod_{i = 1}^{\card{\canrun(r)}-1} \canrun(r)_i \GapLanguage{r}{i} 
-        \right)
-        \quad .
-    \end{equation*}
-
-    Let $u$ and $v$ be two words in the \kl{gap language}
-    $\GapLanguage{r}{\ell}$ for some $r \in M$ and $0 < \ell <
-    \card{\canrun(r)} \defined n$. One can apply \cref{pumping-gap-languages:lem}
-    to conclude that  for all $k \geq 1$,
-    \begin{equation*}
-       \left(\prod_{i = 0}^{\ell - 1} u_i^{x_i} v_i u_i^{y_i} v_i u_i^{z_i}\right) 
-       v u^k v
-       \left(\prod_{i = \ell+1}^{n} u_i^{x_i} v_i u_i^{y_i} v_i u_i^{z_i}\right)
-       \in 
-       L
-    \end{equation*}
-    Where $x_i + y_i + z_i = k$ for all $0 \leq i \leq n$
-    Let us assume without loss of generality that the sequence
-    $\seqof{(x_i, y_i, n_i)}[i \geq 1]$ is pointwise increasing,
-    and even increasing coordinate wise. This is possible because
-    at least one of the coordinates tends to $+\infty$ and in the case
-    of a variable that is bounded, one can extract the sequence to
-    simply not use this variable.
-
-    Now, leveraging \cref{pumping-periods:lem}, we get that for infinitely many
-    $k \geq 1$, $u v^k u$ is an infix of some $x^p$ where $x$ has size smaller
-    or equal than $u$ and $v$. This proves  that $u$ and $v$ are comparable for
-    the prefix, infix, and suffix relations. In particular, we have can
-    conclude that $\GapLanguage{r}{i}$ is a \kl{bounded language}, because of
-    \cite[Lemma 4.1]{ASZZ24}, when $0 < i < \card{\canrun(r)}$. The proof goes
-    on similarly for proving that $\GapLanguage{r}{0}$ and $\GapLanguage{r}{n}$
-    are \kl{bounded languages}, using the other items of
-    \cref{pumping-gap-languages:lem}.
-
-    Because \kl{bounded languages} are stable under concatenation and finite
-    unions, we conclude that $L$ itself must be a \kl{bounded language}.
-\end{proof}
-
-\begin{proofof}{infix-amalgamation:thm}
-    We first apply \cref{infix-amalgamation-bounded:lem},
-    and conclude with \cref{bounded-language:thm}.
-\end{proofof}
-
-For the next result, we impose a mild restriction by requiring the language class to be closed under rational transductions.
-This property holds for most common language classes, including all known \kl{amalgamation systems}.
-
-\begin{theorem}
-	\label{infix-wqo-is-emptiness:thm}
-	Let $\mathcal{C}$ be a class of languages recognized by amalgamation systems and closed under rational transductions. Then the following are equivalent:
-
-	\begin{enumerate}
-		\item\label{wqo-infix-decidable} \kl[well-quasi-order]{Well-quasi-orderedness} of the \kl{infix relation} is decidable.
-		\item\label{emptiness-decidable} Emptiness is decidable.
-	\end{enumerate}
-\end{theorem}
-
-\begin{proof}
-	\cref{emptiness-decidable} $\Rightarrow$ \cref{wqo-infix-decidable}. We aim to make the inclusion test of \cref{amalgamation-infix-wqo-inclusion:eqn} effective.
-
-	Let $R = \bigcup_{x,y \in \Sigma^{\leq n}} \bigcup_{u \in \Sigma^{\leq m \times N_0}} \InfPeriodChain{x} u \InfPeriodChain{y} \cup \InfPeriodChain{x}u \cup u\InfPeriodChain{x}$.
-For any concrete values of the bounds $n$, $m$, and $N_0$, this language is regular.
-As $\mathcal{C}$ is closed under rational transductions, we can therefore reduce the inclusion to emptiness of $L \cap \overline{R}$.
-However, we need to find these bounds first.
-
-	To determine values for $n$ and $m$, we first test if $L$ is bounded.
-As the algorithm in \cite[Section 4.2]{ASZZ24} is effective, this yields words $w_1, \ldots w_n$ such that $L \subseteq w_1^* \cdots w_n^*$ and therefore yields also the bounds in questions.
-
-	To determine the value for $N_0$, we first computer maximal subsets $S \subseteq [n]$ such that $w_i$ for $i \in S$ are simultaneously unbounded.
-We do this by applying a transduction mapping $w_i$ to some fresh letter $a_i$ if $i \in S$ and $\varepsilon$ otherwise, and testing for simultaneous unboundedness \cite[Section 4.1]{ASZZ24}.
-Given a candidate bound $n_0$ for $N_0$, we construct the languages $L_{S,n_0} = w_1^{\circ_1}\cdots w_n^{\circ_n}$ where $\circ_i = *$ if $i \in S$ and $\circ_i = \leq n_0$ otherwise.
-We then test if $L \subseteq \bigcup_{S \text{ maximal}} L_{S,n_0}$.
-If yes, $n_0$ is our value for $N_0$, otherwise, we increase it and repeat the construction.
-As we chose maximal sets $S$, this procedure will eventually terminate.
-
-	\cref{wqo-infix-decidable} $\Rightarrow$ \cref{emptiness-decidable}. We 
-	consider the transduction $\mathcal T = \Sigma^* \times \{a, b\}$. Then 
-	for any language $L \in \mathcal C$, $\mathcal TL$ is well-quasi-ordered 
-	by infix if and only if $L$ is empty.
-\end{proof}
-
-Let us conclude this section by fully characterizing what are
-actually the chains for the \kl{infix} relation, without any 
-regularity assumption.
-
-\begin{lemma}
-    \label{chains-infix:lem}
-    Let $L$ be a \kl{chain} for the \kl{infix relation}. Then, there exists
-    a \kl{chain} $P$ for the \kl{prefix relation} and a \kl{chain}
-    $S$ for the \kl{suffix relation} such that $L \subseteq SP$.
-\end{lemma}
-\begin{proof}
-    Because $L$ is a chain,
-    we have an enumeration $L = \seqof{w_n}$,
-    where $s_n \defined \card{w_n}$ is increasing, 
-    and functions
-    $f_{i,j} \colon \set{1, \dots, s_i} \to \set{1, \dots s_{j}}$
-    satisfying
-    that $f_{j,k} \circ f_{i, j} = f_{i, k}$,
-    $f_{i,i} = \mathrm{id}$, and
-    $w_i = w_j[f_{i,j}(1):f_{i,j}(s_i)]$ for all $i \leq j$.
-    Now, we can build the infinite word $w_\infty$ indexed by $\Rel$, and
-    embeddings $f_{i, \infty} \colon \set{1, \dots, s_i} \to \Rel$ such that
-    $w_i = w_\infty[f_{i, \infty}(1):f_{i, \infty}(s_i)]$, and $f_{i, \infty}
-    \circ f_{j, i} = f_{j, \infty}$ for all $i \leq j$.
-
-    Let us define $w_l$ as the (infinite) word $w_\infty[-\infty: f_{0,
-    \infty}(s_0)]$ and $w_r$ as the (infinite) word $w_\infty[f_{0,
-    \infty}(s_0): +\infty]$. It is now clear that every word $w_i$ can be
-    written as $w_i = w_{i,l} w_{i,r}$, where $w_{i,l}$ is a prefix of $w_r$
-    and $w_{i,r}$ is a suffix of $w_l$. Furthermore, the functions $f_{i,
-    \infty}$ can be used to notice that $\seqof{w_{i,r}}$ is a \kl{chain} for
-    the \kl{prefix relation}, and $\seqof{w_{i,l}}$ is a \kl{chain} for the
-    \kl{suffix relation}, which allows us to conclude.
-\end{proof}

### 2024-10-17 src/prefixes.tex
restructuring paper

@@
     branch $B$ of the \kl{tree of prefixes} such that from every point of the branch, 
-    one can reach a word in $L \setminus B$.
+    one can reach a word in $L \setminus B$. Formally:
+    $\forall u \in B, \exists v \in \Sigma^*, uv \in L \setminus B$.
 \end{definition}
@@
 is a branch of the \kl{tree of prefixes}) is an \kl{antichain branch} for $L$.
-This holds because for any $a^k$, the word $a \prefleq a^kb$ belongs to $L
+This holds because for any $a^k$, the word $a^k \prefleq a^kb$ belongs to $L
 \setminus a^*$. In general, the existence of an \kl{antichain branch} for a
@@
 
-One immediate application of \cref{antichain-branches-prefix:lem} is that
-\kl{antichain branches} can be described inside the \kl{tree of prefixes} by a
-monadic second order formula (\kl{$\MSO$-formula}), allowing us to leverage the
-decidability of $\MSO$ over infinite binary trees \cite[Theorem 1.1]{RAB69}.
+One immediate application of \cref{antichain-branches-prefix:lem} is
+that \kl{antichain branches} can be described inside the \kl{tree of prefixes}
+by a monadic second order formula (\kl{$\MSO$-formula}), allowing us to
+leverage the decidability of $\MSO$ over infinite binary trees \cite[Theorem
+1.1]{RAB69}. This result will follow from our general decidability result in
+\cref{amalgamation:sec} but serves as a good illustration.
 
@@
 \begin{corollary}
-    Let $L \subseteq \Sigma^*$ be an infinite language that is \kl{well-quasi-ordered} by
-    the \kl{prefix relation}. Then, there exists finite numbers $k,\ell \in \Nat$ such that
-    the
-    \kl{maximal order type} of $L$ is $k \cdot \omega$,
-    the \kl{ordinal height} of $L$ is $\omega$, and its
-    \kl{ordinal width} is $\ell$.
+    \label{prefixes-ordinal-invariants:cor}
+    Let $L \subseteq \Sigma^*$ be a language that is
+    \kl{well-quasi-ordered} by the \kl{prefix relation}. Then,
+    \kl{maximal order type} of $L$ strictly smaller than $\omega^2$,
+    the \kl{ordinal height} of $L$ is at most $\omega$, and
+    its \kl{ordinal width} is finite.
 \end{corollary}

### 2024-10-17 src/prelims.tex
restructuring paper

\section{Preliminaries}
\label{prelims:sec}

\paragraph*{Finite words.} In this paper, we use letters $\Sigma, \Gamma$ to
denote finite alphabets, $\Sigma^*$ to denote the set of finite words over
$\Sigma$, and $\varepsilon$ for the empty word in $\Sigma^*$. In order to give
some intuition on the decision problems, we will sometimes use the notion of
finite automata, regular languages, and Monadic Second Order ($\intro*\MSO$)
logic on finite words, and assume is familiar with them. We refer to the
textbook of \cite{THOM97} for a detailed introduction to this topic.

\paragraph*{Well-quasi-orders.} Let us introduce some notations for
\kl{well-quasi-orders}. A sequence $\seqof{x_i}$ in a set $X$ is
\intro(sequence){good} if there exist $i < j$ such that $x_i \leq x_j$. It is
\intro(sequence){bad} otherwise. Therefore, a \kl{well-quasi-ordered} set is a
set where every infinite sequence is \kl(sequence){good}. A \intro{decreasing
sequence} is a sequence $\seqof{x_i}$ such that $x_{i+1} < x_i$ for all $i$,
and an \intro{antichain} is a set of pairwise incomparable elements. An
equivalent definition of a \kl{well-quasi-ordered} set is that it contains no
infinite \kl{decreasing sequences}, nor infinite \kl{antichains}. We refer to
\cite{SCSC12} for a detailed survey on well-quasi-orders.

Let us point out that the \kl{prefix relation} (resp. the \kl{suffix relation}
and the \kl{infix relation}) on $\Sigma^*$ are always \kl{well-founded}, i.e.,
there are no infinite \kl{decreasing sequences} for this ordering. In
particular, for a language $L \subseteq \Sigma^*$ to be
\kl{well-quasi-ordered}, it suffices to prove that it contains no infinite
\kl{antichain}.

\paragraph*{Ordinal Invariants.} An ordinal is a well-founded ordered set. We
use $\alpha, \beta, \gamma$ to denote ordinals, and use $\omega$ to denote the
first infinite ordinal, i.e., the set of natural numbers with the usual
ordering. We assume superficial familiarity with ordinal arithmetic, and refer
to \cite{KUNEN80} for a detailed introduction to this domain. Given a tree $T$
whose branches are all finite we can define an ordinal $\alpha_T$ inductively
as follows: if $T$ is a leaf then $\alpha_T = 0$, if $T$ has children
$\seqof{T_i}$ then $\alpha_T = \sup \setof{\alpha_{T_i} + 1}{i \in \Nat}$. We
say that $\alpha_T$ is the \emph{rank} of $T$.

Let $(X, \leq)$ be a \kl{well-quasi-ordered} set. One can define three
well-founded trees from $X$: the tree of \kl{bad sequences}, the tree of
decreasing sequences, and the tree of \kl{antichains}. The rank of these
respective trees are called respectively the \kl{maximal order type}
\cite{dejongh77}, the \kl{ordinal height} \cite{schmidt81}, and the \kl{ordinal
width} of $X$ \cite{kriz90b}. We refer to the survey of \cite{DZSCSC20} for a
detail discussion on these concepts and their computation on specific classes
of well-quasi-ordered sets.

### 2024-10-17 src/conclusion.tex
Add lex

@@
 the minimal automaton recognizing the language.
+
+\subparagraph{Lexicographic orderings} There is another natural ordering on
+words, the \emph{lexicographic ordering}, which does not fit well in our
+current framework because it is always of \kl{ordinal width} $0$. However, the
+\kl{order-type} of the lexicographic ordering over \kl{regular languages} has
+already been investigated in the context of infinite words, and it would be
+interesting to see if one can extend these results to decide whether such an
+ordering is \kl{well-founded} for languages recognized by \kl{amalgamation
+systems}.

### 2024-10-17 src/trash.tex
Adding a trash

--- 2024-10-17 src/infixes.tex
@@
-% LTeX: language=en-GB 
-\section{Infixes and Bounded Languages}
-\label{infixes-regular:sec}
+\section{Trash}
 
-\todo{Delete this section: the \cref{infix-embedding:thm} is already
-    known, we should just restate it. The
-    \cref{infix-finite-automata:thm} is a direct consequence of
-    \cref{infix-amalgamation:thm} and \cref{bounded-language:thm}.
-    I don't know if we should keep it for clarity, but it does not 
-    seems to be necessary to me.
-}
-
-\AP
-In this section, we study languages equipped with the \kl{infix relation}. As
-opposed to the \kl{prefix} and \kl{suffix} relations, the \kl{infix relation}
-can yield to very complicated \kl{well-quasi-ordered} languages. Formally, the
-upcoming \cref{infix-embedding:thm} shows that \emph{any} countable
-partial-ordering with finite initial segments can be embedded into the infix
-relation of a language. To make the former statement precise,
-Let us recall that an \intro{order embedding} from a quasi-ordered set $(X,
-\preceq)$ into a quasi-ordered set $(Y, \preceq')$ is a function $f \colon X
-\to Y$ such that for all $x, y \in X$, $x \preceq y$ if and only if $f(x)
-\preceq' f(y)$. When such an embedding exists, we say that $X$ \reintro{embeds
-into} $Y$. We say that a quasi-ordered set $(X, \preceq)$ is a \intro{partial
-ordering} whenever the relation $\preceq$ is antisymmetric, that is $x \preceq
-y$ and $y \preceq x$ implies $x = y$.
-
-The following result due to Kuske shows the structural complexity of the \kl{infix relation}
-
-\begin{lemma}{\cite[Lemma 5.1]{DBLP:journals/ita/Kuske06}}
-    \label{infix-embedding:thm}
-    Let $(X, \preceq)$ be a \kl{partially ordered} set,
-    and $\Sigma$ be an alphabet with at least two letters.
-    Then the following are equivalent:
-    \begin{enumerate}
-        \item \label{infix-embedding-embeds:item} 
-            $X$ \kl{embeds into} $(\Sigma^*, \infleq)$,
-        \item \label{infix-embedding-count:item}
-            $X$ is countable, and for every $x \in X$,
-            its \kl{downwards closure}
-            $\dwset[\preceq]{x}$ is finite.
-    \end{enumerate}
-\end{lemma}
-\begin{figure}
-    \centering
-    \includestandalone[width=\linewidth]{fig/infix-encoding-standalone}
-    \caption{Representation of the \kl{subword relation} for $\set{a,b}^*$
-        inside the \kl{infix relation} for $\set{a,b,\#}^*$
-        using the encoding of \cref{infix-embedding:thm}, restricted to words
-        of length at most $3$. To obtain smaller words,
-        we replaced $D_i$ by $\max D_i$ in this construction.
-    }
-    \label{infix-embedding:fig}
-\end{figure}
-
-As a consequence of \cref{infix-embedding:thm}, we cannot replay proofs of
-\cref{prefixes:sec}, and will actually need to leverage some regularity of the
-languages to obtain a characterization of \kl{well-quasi-ordered} languages
-under the \kl{infix relation}. Let us first play this game for languages that
-are recognized by finite automata. We assume that the reader is familiar with
-the notions of deterministic finite automaton and regular languages, and refer
-to the book of Thomas for a comprehensive introduction to the subject
-\cite{THOM97}. The main goal of the remainder of this section is to prove the
-following \cref{infix-finite-automata:thm}.
+Let us first play this game for languages that are recognized by
+finite automata. We assume that the reader is familiar with the notions of
+deterministic finite automaton and regular languages, and refer to the book of
+Thomas for a comprehensive introduction to the subject \cite{THOM97}. The main
+goal of the remainder of this section is to prove the following
+\cref{infix-finite-automata:thm}.
 
@@
 \end{theorem}
-
-\AP In order to prove \cref{infix-finite-automata:thm}, we will perform some
-preliminary analysis on the structure of an automaton recognizing a
-well-quasi-ordered language under the infix relation, which will be powered by
-folklore results on \emph{periodic} words. Let us recall that a non-empty word
-$w \in \Sigma^+$ is \intro(word){periodic} with period $x \in \Sigma^*$ if
-there exists a $p \in \Nat$ such that $w \infleq x^p$. The \intro{periodic
-length} of a word $u$ is the minimal length of a word $x$ such that $u$ is an
-\kl{infix} of $x^p$ for some $p \in \Nat$ and $x \in \Sigma^+$. We will
-essentially rely on the following result on periodic words.
-
-\begin{lemma}
-    \label{periodic-infixes:lem}
-    Let $u,v \in \Sigma^*$ be two (non-empty) \kl{periodic words}
-    having \kl{periodic lengths} $p$ and $q$ respectively.
-    Then, if $u \infleq v$ and $\card{u} \geq \factorial[p]{p \times q}$,
-    then $u$ and $v$ share the same \kl{periodic length}
-    $p = q$.
-\end{lemma}
-\begin{proof}
-    The fact that $u$ and $v$ are \kl{periodic length}
-    respectively $p$ and $q$ translates into the fact that $u_{i+p} = u_i$ and
-    $v_{i+q} = v_i$ for all indices $i \in \Nat$ such that those letters are
-    well-defined.
-
-    Now, assume that $u$ is an \kl{infix} of $v$, this provides the existence
-    of a $k \in \Nat$ such that $u = v_{k} \cdots v_{k + \card{u} - 1}$. In
-    particular, $v_{k+i+p} = v_{k+i}$ for all $i \in \Nat$ such that $k+i+p < k
-    + \card{u}$. Since we also have $v_{k+i+q} = v_{k+i}$ for all $1 \leq i
-    \leq \card{v} - k - q$. We conclude that both $u$ and $v$ are of
-    \kl{periodic length} the greatest common divisor of $p$ and $q$, and by
-    minimality of $q$ this must be equal to $q$ and to $p$.
-\end{proof}
-
-The upcoming \cref{powers-infixes:cor} is based on the observation that if $S$
-is a \kl{chain} for the \kl{suffix relation} and $P$ is a \kl{chain} for the
-\kl{prefix relation}, then $SP$ is a \kl{chain} for the \kl{infix relation}.
-
-\begin{corollary}
-    \label{powers-infixes:cor}
-    Let $u,v \in \Sigma^*$ and $k, \ell \in \Nat$
-    such that $k \geq \factorial[p]{\card{u} \times \card{v}}$,
-    $\ell \geq \factorial[p]{\card{v} \times \card{u}}$,
-    and $u^k \infleq v^\ell$.
-    Then, there exists $w \in \Sigma^*$ of size at most
-    $\min \set{\card{u}, \card{v}}$ and a $p \in \Nat$
-    such that
-    $u^k \infleq v^\ell \infleq w^p$.
-\end{corollary}
-
-The reason why \kl{periodic words} built using a given period $x \in \Sigma^+$
-are interesting for the \kl{infix relation} is that they naturally create
-\kl{chains}. Indeed, if $x \in \Sigma^+$ is a finite word, then $\setof{x^p}{p
-\in \Nat}$ is a \kl{chain} for the \kl{infix relation}. Note that in general,
-the \kl{downwards closure} of a chain is \emph{not} a chain. However, for the
-chains generated using periodic words, the \kl{downwards closure}
-$\dwset[\infleq]{\setof{x^p}{p \in \Nat}}$ is a \emph{finite union} of
-\kl{chains}. Because this set will appear in bigger equations, we introduce the
-shorter notation $\intro*\InfPeriodChain{x}$ for the set of \kl{infixes} of
-words of the form $x^p$, where $p$ ranges over $\Nat$.
-
-\begin{lemma}
-    \label{inf-period-chain:lem}
-    Let $x \in \Sigma^+$ be a word, and
-    Then $\InfPeriodChain{x}$ is a finite union of \kl{chains}
-    for the \kl{infix}, \kl{prefix} and \kl{suffix} relations 
-    simultaneously.
-\end{lemma}
-\begin{proof}
-    Let $x \in \Sigma^+$ be a word, and let $P_x$ be the (finite) set 
-    of all \kl{prefixes} of $x$, and $S_x$ be the (finite)
-    set of all \kl{suffixes} of $x$.
-    Assume that $w \in \InfPeriodChain{x}$, then $w = u x^p v$ for some
-    $u \in S_x$, $v \in P_x$, and $p \in \Nat$.
-    We have proven that
-    \begin{equation*}
-        \InfPeriodChain{x} \subseteq \bigcup_{u \in P_x} \bigcup_{v \in S_x} u x^* v
-        \quad .
-    \end{equation*}
-
-    Let us now demonstrate that for all $(u,v) \in S_x \times P_x$, the
-    language $u x^* v$ is a \kl{chain} for the \kl{infix}, \kl{suffix} and \kl{prefix} relations.
-    To that end,
-    let $(u,v) \in S_x \times P_x$ and $\ell, k \in \Nat$ be such that $\ell <
-    k$, let us prove that $u x^\ell v \infleq u x^k  v$. Because $v \prefleq
-    x$, we know that there exists $w$ such that $vw = x$. In particular,
-    $ux^\ell vw = u x^{\ell + 1}$, and because $\ell < k$, we conclude that $u
-    x^{\ell + 1} \prefleq u x^k v$. By transitivity, $u x^\ell v \prefleq u x^k
-    v$, and \emph{a fortiori}, $u x^\ell v \infleq u x^k v$. 
-    Similarly, because $u \suffleq x$,  there exists $w$ such that $wu  = x$, 
-    and we conclude that $u x^{\ell} v \suffleq w u x^\ell v = x^{\ell + 1} v \suffleq u x^k v$.
-    \qedhere
-\end{proof}
-
-\begin{corollary}
-    \label{inf-period-union-chains:lem}
-    Let $x,u,y \in \Sigma^*$ be words.  The following 
-    are finite unions of \kl{grids} for the infix relation:
-    $\InfPeriodChain{x}u$, $u \InfPeriodChain{x}$,
-    and $\InfPeriodChain{x} u \InfPeriodChain{y}$.
-\end{corollary}
-\begin{proof}
-    Because $\InfPeriodChain{x}$ is a finite union of \kl{chains} for the \kl{suffix}
-    relation (rsing \cref{inf-period-chain:lem}), we conclude that $\InfPeriodChain{x}u$ is one too for
-    the \kl{infix} relation. This holds similarly for $u \InfPeriodChain{x}$.
-    In the case of $\InfPeriodChain{x} u \InfPeriodChain{y}$,
-    we remark that $\InfPeriodChain{x} u$ is a finite union of chains
-    for the \kl{suffix relation},
-    and that $\InfPeriodChain{y}$ is one for the \kl{prefix relation}
-    (using again \cref{inf-period-chain:lem}).
-    As a consequence,
-    $\InfPeriodChain{x}u \InfPeriodChain{y}$ is a finite union 
-    of \kl{chains} for the \kl{infix} relation.
-\end{proof}
-
-The following combinatorial lemma connects the property of being
-\kl{well-quasi-ordered} to a property of the \kl{periodic lengths} of words in
-a language, based on the assumption that some factors can be iterated. It is
-the core result that powers the analysis done in the upcoming
-\cref{infix-finite-automata:thm,infix-amalgamation:thm}.
-
-\begin{lemma}
-    \label{pumping-periods:lem}
-    Let $L \subseteq \Sigma^*$ be a language
-    that is \kl{well-quasi-ordered} by the \kl{infix relation}.
-    Let $k \in \Nat$, $u_1, \cdots, u_{k+1} \in \Sigma^*$,
-    and $v_1, \cdots, v_{k} \in \Sigma^+$
-    be such that
-    $w[\vec{n}] \defined (\prod_{i = 1}^k u_i v_i^{n_i}) u_{k+1}$
-    belongs to $L$
-    for arbitrarily large values of $\vec{n} \in \Nat^k$.
-    Then, 
-    there exists $x,y \in \Sigma^+$ of size 
-    at most $\max \setof{\card{v_i}}{1 \leq i \leq k}$
-    such that 
-    one of the following holds for all
-    $\vec{n} \in \Nat^{k}$:
-    \begin{enumerate}
-        \item $w[\vec{n}] \in u_1 \InfPeriodChain{x}$,
-        \item $w[\vec{n}] \in \InfPeriodChain{x} u_{k+1}$,
-        \item $w[\vec{n}] \in \InfPeriodChain{x} u_i \InfPeriodChain{y}$
-            for some $1 \leq i \leq k + 1$.
-
-    \end{enumerate}
-\end{lemma}
-\begin{proof}
-    Note that the result is obvious if $k = 0$, and therefore
-    we assume $k \geq 1$ in the following proof.
-
-    Let us construct a sequence of words $\seqof{w_i}[i \in \Nat]$, where $w_i
-    \defined w[\vec{n_i}]$ for some well-chosen indices $\vec{n_i} \in \Nat^k$. The goal
-    being that 
-    if $w[\vec{n_i}]$ is an \kl{infix} of $w[\vec{n_j}]$,
-    then it can intersect at most \emph{two} iterated words,
-    with an intersection that is long enough to successfully apply
-    \cref{periodic-infixes:lem}.
-    In order to achieve this,
-    let us first define $s$ as the maximal size of a word $v_i$
-    ($1 \leq i \leq k$) and $u_j$ ($1 \leq j \leq k+1$).
-    Then,
-    we consider $\vec{n_0} \in \Nat^k$ such that $\vec{n_0}$ has all 
-    its components greater than $\factorial{s}$ and such that
-    $w[\vec{n_0}]$ belongs to $L$. 
-    Then, we inductively define 
-    $\vec{n_{i+1}}$  as the smallest vector of numbers greater than $\vec{n_i}$,
-    such that $w[\vec{n_{i+1}}]$ belongs to $L$, 
-    and with $\vec{n_i}$ having all components greater than
-    $2\card{w[\vec{n_i}]}$.
-
-    Let us assume that $k \geq 2$ in the following proof for symmetry purposes,
-    and argue later on that when $k = 1$ the same argument goes through.
-    Because $L$ is \kl{well-quasi-ordered} by the \kl{infix relation}, there
-    exists $i < j$ such that $w[\vec{n_i}]$ is an \kl{infix} of $w[\vec{n_j}]$.
-    Now, because of the chosen values for $\vec{n_j}$, there exists $1 \leq \ell \leq
-    k-1$ such that $w[\vec{n_i}]$ is actually an \kl{infix} of $u_{\ell}
-    v_{\ell}^{n_{j,\ell}} u_{\ell+1} v_{\ell+1}^{n_{j,\ell+1}} u_{\ell+2}$.
-    Even more,
-    one of the three following equations holds:
-    \begin{itemize}
-        \item $w[\vec{n_i}] \infleq v_{\ell}^{n_{j,\ell}} u_{\ell+1} v_{\ell+1}^{n_{j,\ell+1}}$,
-        \item $w[\vec{n_i}] \infleq u_{\ell}
-            v_{\ell}^{n_{j,\ell}}$,
-        \item $w[\vec{n_i}] \infleq
-            v_{\ell+1}^{n_{j,\ell+1}} u_{\ell+2}$.
-    \end{itemize}
-    In all those cases, we conclude using \cref{powers-infixes:cor}
-    that there exists $x,y \in \Sigma^+$ of size at most $s$, and 
-    a number $1 \leq t \leq k$ such that
-    $v_i^{n_i} \in \InfPeriodChain{x}$ for all $1 \leq i \leq t$,
-    and
-    $v_i^{n_i} \in \InfPeriodChain{y}$ for all $t < i \leq k$.
-    In particular,
-    $w[\vec{n_i}] \in \InfPeriodChain{x} u_{t} \InfPeriodChain{y}$.
-
-    When $k = 1$, the situation is a bit more specific since we only have two
-    cases: either $w_i \infleq u_1 v_1^{n_j}$ or $w_i \infleq v_1^{n_j} u_2$,
-    and we conclude with an identical reasoning.
-\end{proof}
-
-We are now ready to restate and prove our main theorem, proven by combining
-\cref{pumping-periods:lem} together with classical pumping arguments on finite
-state automata.
-
 \begin{proofof}{infix-finite-automata:thm}
@@
 \end{corollary}
-
-\begin{theorem}
-    \label{bounded-language:thm}
-    Let $L$ be a \kl{bounded language} of $\Sigma^*$. Then,
-    $L$ is a \kl{well-quasi-order} when endowed with the 
-    \kl{infix relation} if and only if it is a finite union of \kl{grids}.
-\end{theorem}
-\begin{proof}
-    Let $w_1, \dots, w_n$ be such that
-    $L \subseteq w_1^* \cdots w_n^*$.
-    Let us define $m \defined \max \setof{\card{w_i}}{1 \leq i \leq n}$
-
-    Let $w[\vec{k}] \defined w_1^{k_1} \cdots w_n^{k_n}$ be a map from $\Nat^k$
-    to $\Sigma^*$. We are interested in the intersection of the image of $w$
-    with $L$. Let us assume for instance that for all $\vec{k} \in \Nat^n$,
-    there exists $\vec{\ell} \geq \vec{k}$ such that $w[\vec{\ell}] \in L$.
-    Then, leveraging \cref{pumping-periods:lem}, we conclude that there exists
-    $x,y$ of size at most $\max\setof{\card{w_i}}{1 \leq i \leq n}$ such that
-    $w[\vec{k}] \in \InfPeriodChain{x} \cup \InfPeriodChain{x}
-    \InfPeriodChain{y}$, and we conclude that $L \subseteq \InfPeriodChain{x}
-    \cup \InfPeriodChain{x} \InfPeriodChain{y}$.
-
-    Now, it may be the case that one cannot simultaneously assume that two
-    component of the vector $\vec{k}$ are unbounded. In general, given a set $S
-    \subseteq \set{1, \dots, n}$ of indices, we say that $S$ is admissible if
-    there exists a bound $N_0$ such that for all $\vec{b} \in \Nat^S$, there
-    exists a vector $\vec{k} \in \Nat^n$, such that $\vec{k}$ is greater than
-    $\vec{b}$ on the $S$ components, and the other components are below the
-    bound $N_0$. The language of an admissible set $S$ is the set of words
-    obtained by repeating $w_i$ at most $N_0$ times if it is not in $S$
-    ($w_i^{\leq N_0}$) and arbitrarily many times otherwise ($w_i^*$).
-    Note that $L \subseteq \bigcup_{S \text{ admissible }} L(S)$.
-
-    Now, admissible languages are ready to be pumped according to
-    \cref{pumping-periods:lem}. For every admissible language,
-    the size of a word that is not iterated is at most
-    $N_0 \times m$ by definition, and we conclude that:
-    \begin{equation*}
-        L \subseteq 
-        \bigcup_{x,y \in \Sigma^{\leq n}}
-        \bigcup_{u \in \Sigma^{\leq m \times N_0}}
-        \InfPeriodChain{x} u \InfPeriodChain{y}
-        \cup
-        \InfPeriodChain{x} u
-        \cup
-        u \InfPeriodChain{x}
-        \quad .
-    \end{equation*}
-\end{proof}
-
-Note that this immediately generalizes \cref{reg-wqo-iff-dwclosed-wqo:cor}
-to all \kl{bounded languages}.
-
-\begin{corollary}
-    \label{bounded-wqo-dwclosed:cor}
-    Let $L$ be a \kl{bounded language} of $\Sigma^*$. Then,
-    $L$ is a \kl{well-quasi-order} when endowed with the
-    \kl{infix relation} if and only if $\dwset[\infleq]{L}$ is.
-\end{corollary}
-
-In the light of \cref{reg-wqo-iff-dwclosed-wqo:cor,bounded-wqo-dwclosed:cor}
-one may wonder whether the \kl{downwards closed}
-languages for the \kl{infix relation} satisfy a similar characterization in
-terms of union of \kl{grids} as stated in
-\cref{infix-finite-automata:thm,bounded-language:thm}. Let us
-first remark that from our previous results we obtain the following lemma that
-illustrates how all the previous characterizations of \kl{well-quasi-ordered}
-languages for the \kl{infix relation} are collapsing in the case of
-\kl{downwards closed} languages.
-
-\begin{lemma}
-    \label{dwclosed-infixes-wqo:lem}
-    Let $L \subseteq \Sigma^*$ be a \kl{downwards closed} language for the
-    \kl{infix relation} that is \kl{well-quasi-ordered}. Then, the following
-    are equivalent:
-    \begin{enumerate}
-        \item $L$ is \kl{bounded},
-        \item $L$ is \kl{regular},
-        \item $L$ is a finite union of \kl{chains} and \kl{grids}.
-    \end{enumerate}
-\end{lemma}
-\begin{proof}
-    TODO.
-\end{proof}
-
-One may think that all \kl{downwards closed} languages for the \kl{infix
-relation} that are \kl{well-quasi-ordered} are \kl{regular}. Note that this is
-what happens in the case of the \kl{subword embedding}, where any \kl{downwards
-closed} language for this relation is characterized by finitely many excluded
-patterns, hence \kl{regular}. However, this is not the case for the \kl{infix
-relation}, as we will now illustrate with the following two examples.
-
-\begin{example}
-    \label{dwclosed-wqo-not-finite-excl:ex}
-    Let $L \defined a^* b^* \cup b^* a^*$. This language is \kl{downwards
-    closed} for the \kl{infix relation}, is \kl{well-quasi-ordered} for the
-    \kl{infix relation}, but is characterized by an \emph{infinite} number 
-    of excluded infixes, respectively of the form $ab^ka$ and $ba^kb$ where $k \geq 1$.
-\end{example}
-
-To strengthen \cref{dwclosed-wqo-not-finite-excl:ex}, we will
-leverage the \intro{Thue-Morse sequence} $\intro*\ThueMorse \in
-\set{0,1}^{\Nat}$, which we will use as a black-box for its two main
-characteristics: it is \kl{cube-free} and \kl{uniformly recurrent}. Being
-\intro{cube-free} means that no (finite) word of the form $uuu$ is an
-\kl{infix} of $\ThueMorse$, and being \intro{uniformly recurrent} means that
-for every word $u$ that is an \kl{infix} of $\ThueMorse$, there exists $k \geq
-1$ such that $u$ is an \kl{infix} of every word $v$ of size at least $k$. We
-refer the reader to a nice survey of Allouche and Shallit for more information
-on this sequence and its properties \cite{ALSHA99}.
-
-\begin{lemma}
-    \label{thue-morse:lemma}
-    There exists a language $L$ that is \kl{downwards closed} for the \kl{infix
-    relation}, \kl{well-quasi-ordered} for the \kl{infix relation}, but is not
-    \kl{regular}.
-\end{lemma}
-\begin{proof}
-    Let $\ThueMorse$ be the \kl{Thue-Morse sequence}
-    $L$ be the set of infixes of $w$. By construction $L$ is an \emph{infinite}
-    \kl{downwards closed} for the \kl{infix relation}. Let us argue that $L$ is
-    \kl{well-quasi-ordered} for the \kl{infix relation}, but is not \kl{regular}.
-
-    Assume by contradiction that $L$ is \kl{bounded}. In this case, there exist
-    words $w_1, \dots, w_k \in \Sigma^*$ such that $L \subseteq w_1^* \cdots
-    w_k^*$. Since $L$ is infinite and \kl{downwards closed}, there exists a
-    word $u \in L$ such that $u = w_i^3$ for some $1 \leq i \leq k$. This is absurd
-    because $u \infleq \ThueMorse$, which is \kl{cube-free}.
-
-    Furthermore, $L$ is \kl{well-quasi-ordered} for the \kl{infix relation}.
-    Indeed, consider a sequence $\seqof{u_i}$ of words in $L$. Without loss of
-    generality, we may consider a subsequence such that $\card{u_i} <
-    \card{u_{i+1}}$ for all $i \in \Nat$. Because $\ThueMorse$ is \kl{uniformly
-    recurrent}, there exists $k \geq 1$ such that $u_1$ is an \kl{infix} of
-    every word $v$ of size at least $k$. In particular, $u_1$ is an \kl{infix}
-    of $u_k$, hence the sequence $\seqof{u_i}$ is \kl(sequence){good}.
-
-    We conclude thanks to
-    \cref{dwclosed-infixes-wqo:lem}, that $L$ is not \kl{regular}.
-\end{proof}

### 2024-10-17 src/prelims.tex
Format

@@
 some intuition on the decision problems, we will sometimes use the notion of
-finite automata, regular languages, and Monadic Second Order ($\intro*\MSO$)
-logic on finite words, and assume is familiar with them. We refer to the
-textbook of \cite{THOM97} for a detailed introduction to this topic.
+\intro{finite automata}, \intro{regular languages}, and \intro{Monadic Second
+Order} ($\intro*\MSO$) logic on finite words, and assume is familiar with them.
+We refer to the textbook of \cite{THOM97} for a detailed introduction to this
+topic.
 
@@
 ordering. We assume superficial familiarity with ordinal arithmetic, and refer
-to \cite{KUNEN80} for a detailed introduction to this domain. Given a tree $T$
-whose branches are all finite we can define an ordinal $\alpha_T$ inductively
-as follows: if $T$ is a leaf then $\alpha_T = 0$, if $T$ has children
-$\seqof{T_i}$ then $\alpha_T = \sup \setof{\alpha_{T_i} + 1}{i \in \Nat}$. We
-say that $\alpha_T$ is the \emph{rank} of $T$.
+to \cite{KUNEN80} for a detailed introduction to this domain.
+Given a tree $T$ whose branches are all finite we can define
+an ordinal $\alpha_T$ inductively as follows: if $T$ is a leaf then $\alpha_T =
+0$, if $T$ has children $\seqof{T_i}$ then $\alpha_T = \sup \setof{\alpha_{T_i}
++ 1}{i \in \Nat}$. We say that $\alpha_T$ is the \emph{rank} of $T$.
 
@@
 decreasing sequences, and the tree of \kl{antichains}. The rank of these
-respective trees are called respectively the \kl{maximal order type}
-\cite{dejongh77}, the \kl{ordinal height} \cite{schmidt81}, and the \kl{ordinal
-width} of $X$ \cite{kriz90b}. We refer to the survey of \cite{DZSCSC20} for a
-detail discussion on these concepts and their computation on specific classes
-of well-quasi-ordered sets.
+respective trees are called respectively the \intro{maximal order type} of $X$
+written $\intro*{\oType{X}}$ \cite{dejongh77}, the \intro{ordinal height} of
+$X$ written $\intro*{\oHeight{X}}$ \cite{schmidt81}, and the \intro{ordinal
+width} of $X$ written $\intro*{\oWidth{X}}$ \cite{kriz90b}. We refer to the
+survey of \cite{DZSCSC20} for a detail discussion on these concepts and their
+computation on specific classes of well-quasi-ordered sets.

### 2024-10-17 src/introduction.tex
restructuring paper
Double knowledge

@@
 underlying quasi-order \todo{CITE???}. One way to measure this complexity is to
-consider its so-called \intro{ordinal invariants}: for instance, the
-\intro{maximal order type} (or m.o.t.), originally defined by De Jongh and
+consider its so-called \kl{ordinal invariants}: for instance, the
+\kl{maximal order type} (or m.o.t.), originally defined by De Jongh and
 Parikh \cite{dejongh77}, is the order type of the maximal linearization of a
@@
 the size of the set. Better runtime bounds were obtained by considering two
-other parameters\todo{cite sylvain width/ HDR?}: the \intro{ordinal height}
-introduced by Schmidt \cite{schmidt81}, and the \intro{ordinal width} of Kříž
+other parameters\todo{cite sylvain width/ HDR?}: the \kl{ordinal height}
+introduced by Schmidt \cite{schmidt81}, and the \kl{ordinal width} of Kříž
 and Thomas \cite{kriz90b}. Therefore, when characterizing
@@
 	\item A language $L \subseteq \Sigma^*$ is \kl{well-quasi-ordered} by the \kl{prefix relation} if and only if it is a finite union of \kl{chains} of the \kl{prefix relation}.
+    \hfill (\cref{prefixes:thm} p.\pageref{prefixes:thm})
 	\item A language $L \subseteq \Sigma^*$ is \kl{well-quasi-ordered} by the \kl{suffix relation} if and only if it is a finite union of \kl{chains} of the \kl{suffix relation}.
+    \hfill (\cref{prefixes:thm} p.\pageref{prefixes:thm})
 \end{enumerate}
@@
 language, it is going to provide relatively efficient algorithms.
-
+{
 \renewcommand{\labelenumi}{R\arabic{enumi}}
@@
         below $\omega^2$.
+        \hfill (\cref{prefixes-ordinal-invariants:cor} p.\pageref{prefixes-ordinal-invariants:cor})
 \end{enumerate}
@@
     \setcounter{enumi}{5}
-    \item There is a decision procedure for checking whether a language $L$ recognized by an \kl{amalgamation system} is \kl{well-quasi-ordered} by the \kl{prefix relation}, the \kl{suffix relation}, or the \kl{infix relation}.
-Furthermore, if it is \kl{well-quasi-ordered}, then $L$ is a \kl{bounded language}.
+    \item There is a decision procedure for checking whether a language $L$ 
+        recognized by an \kl{amalgamation system} is \kl{well-quasi-ordered}
+        by the \kl{prefix relation}, the \kl{suffix relation},
+        or the \kl{infix relation}. Furthermore, if it is \kl{well-quasi-ordered}, 
+        then $L$ is a \kl{bounded language}.
+        \hfill (\cref{infix-wqo-is-emptiness:thm} p.\pageref{infix-wqo-is-emptiness:thm})
 \end{enumerate}
@@
 
-\subparagraph{Outline} In
+\subparagraph{Outline} We introduce in \cref{prelims:sec} the
+necessary background on \kl{well-quasi-orders} and \kl{ordinal invariants}.
+In
 \cref{prefixes:sec}, which is relatively

### 2024-10-17 src/amalgam.tex
moving from amalgam to infixes

@@
-\section{Amalgamation Systems and Decidability}
+\section{Decision Prodecure and Amalgamation Systems}
 \label{infixes-amalgamation:sec}
 
-When proving \cref{infix-finite-automata:thm}, we have leveraged a powerful
-combinatorial argument on regular langugages stated in
-\cref{pumping-periods:lem}. Our first remark is that this idea can
-be generalized to the larger class of \intro{bounded languages}, i.e.,
-languages $L \subseteq \Sigma^*$ such that there exists words $w_1, \dots, w_n$
-satisfying $L \subseteq w_1^* \cdots w_n^*$.
+\AP
+In this section, we are going to design an effective decision procedure for
+\kl{well-quasi-ordering} by the \kl{infix relation}. To that end, the first
+requirement is to fix a way to represent languages $L \subseteq \Sigma^*$.
+Traditionally, one would use finite automata, context-free grammars, or for the
+more adventurous, vector addition systems with states. However, our proof
+technique will only require us to have a way to ``glue'' together runs of the
+system to ``pump'' them and produce new runs: this is the usual pumping lemma
+in automata theory, and Ogden's lemma for context-free grammars.
+\todo{Cite the pumping lemma for VASS. Cite Ogden's lemma.}
 
-\subsection{Amalgamation Systems}
-\label{infixes-amalgamation:subsec}
-
-It turns out that there is a rather large family of systems for which pumping
-arguments based on so-called \emph{minimal runs} exist: they are called
+\AP It turns out that there is a rather large family of systems for which
+pumping arguments based on so-called \emph{minimal runs} exist: they are called
 \emph{amalgamation systems} and were recently introduced by Anad, Schmitz,
@@
 \cref{infix-finite-automata:thm} to context-free grammars, languages recognized
-by vector addition systems with state (VASS), and more. The goal of this
+by vector addition systems with state (VASS), and more.
+\todo{unclear formulation}
+The goal of this
 section is not to introduce all of these systems, or justify their usefulness,
-but to state and prove the following theorem.
+but to state and prove the following two theorems.
 
@@
     Then $L$ is well-quasi-ordered by the infix relation if and only if $L$ is
-    a finite union of chains for the \kl{infix relation}.
+    \kl(language){bounded}.
 \end{theorem}
+
+\begin{theorem}[restated=infix-wqo-is-emptiness:thm,label=infix-wqo-is-emptiness:thm]
+% 	\label{infix-wqo-is-emptiness:thm}
+	Let $\mathcal{C}$ be a class of languages recognized by 
+    \kl{amalgamation systems}
+    and closed under
+    \kl{rational transductions}. 
+    Then the following are equivalent:
+	\begin{enumerate}
+        \item\label{wqo-infix-decidable} \kl[wqo]{Well-quasi-orderedness} of the \kl{infix relation} is decidable for languages in $\mathcal{C}$.
+        \item\label{wqo-prefix-decidable} \kl[wqo]{Well-quasi-orderedness} of the \kl{prefix relation} is decidable for languages in $\mathcal{C}$.
+        \item\label{emptiness-decidable} Emptiness is decidable for languages in $\mathcal{C}$.
+	\end{enumerate}
+\end{theorem}
+
+\subsection{Amalgamation Systems}
+\label{infixes-amalgamation:subsec}
 
@@
 
+\subsection{Amalgamation Systems, WQOs, and Bounded Languages}
+
 Let us now introduce a combinatorial lemma that explains how \kl{gap languages}
@@
 
-\begin{lemma}
-    \label{infix-amalgamation-bounded:lem}
-    Let $L \subseteq \Sigma^*$ be a language recognized
-    by an \kl{amalgamation system} that is \kl{well-quasi-ordered}
-    for the \kl{infix relation}. Then $L$ is a \kl{bounded language}.
-\end{lemma}
-\begin{proof}
+\begin{proofof}{infix-amalgamation:thm}
     Assume that $L$ is \kl{well-quasi-ordered} by the \kl{infix relation},
@@
     unions, we conclude that $L$ itself must be a \kl{bounded language}.
+\end{proofof}
+
+One can leverage the same kind of reasoning to take the opposite direction, and
+notice that the notions of being a \kl{bounded language}, a \kl{regular
+language}, or being included in a finite union of products of \kl{chains} all
+collapse for \kl{well-quasi-ordered} and \kl{downwards closed} languages for
+the \kl{infix relation}.
+
+\begin{lemma}
+    \label{dwclosed-infixes-wqo:lem}
+    Let $L \subseteq \Sigma^*$ be a \kl{downwards closed} language for the
+    \kl{infix relation} that is \kl{well-quasi-ordered}. Then, the following
+    are equivalent:
+    {\renewcommand{\theenumi}{\roman{enumi}}
+     \renewcommand{\labelenumi}{(\theenumi)}
+    \begin{enumerate}
+        \item\label{dwci-reg:item} $L$ is a \kl{regular language},
+        \item\label{dwci-aml:item} $L$ is recognized by \emph{some} \kl{amalgamation system},
+        \item\label{dwci-bod:item} $L$ is a \kl{bounded language},
+        \item\label{dwci-uoc:item} $L$ is included in a finite union of products of \kl{chains}
+    \end{enumerate}
+    }
+\end{lemma}
+\begin{proof}
+    It is clear that \cref{dwci-reg:item} $\Rightarrow$ \cref{dwci-aml:item}
+    because regular languages are recognized by finite automata, and finite
+    automata are a particular case of \kl{amalgamation systems}.
+    The implication \cref{dwci-aml:item} $\Rightarrow$ \cref{dwci-bod:item}
+    is the content of \cref{infix-amalgamation:thm}.
+    The implication \cref{dwci-bod:item} $\Rightarrow$ \cref{dwci-uoc:item}
+    is \cref{bounded-language:thm}.
+    Finally, the implication \cref{dwci-uoc:item} $\Rightarrow$ \cref{dwci-reg:item}
+    is simply because a \kl{downwards closed} language 
+    that is a finite union of products of \kl{chains} is a regular language.
 \end{proof}
 
-\begin{proofof}{infix-amalgamation:thm}
-    We first apply \cref{infix-amalgamation-bounded:lem},
-    and conclude with \cref{bounded-language:thm}.
+\subsection{Effective Decision Procedures}
+\label{infixes-amalgamation-effective:subsec}
+
+For the next result, we impose a mild restriction by requiring the language
+class to be closed under \kl{rational transductions}. This property holds for
+most common language classes, including all known \kl{amalgamation systems}.
+
+\AP A class of languages $\mathcal{C}$ is \intro{recognized by amalgamation
+systems} when, for every language $L \in \mathcal{C}$, there exists an
+\kl{amalgamation system} $(\Sigma, R, E, \canrun)$ such that $L$ is the
+language of this system.
+
+\AP A \intro{rational transduction} is a function $f \colon \Sigma^* \to
+\Gamma^*$ that can be defined by a finite automaton with output, or
+equivalently by a finite set of rules of the form $u \to v$ where $u \in
+\Sigma^*$ and $v \in \Gamma^*$. We write $f(L) = \setof{f(w)}{w \in L}$ for the
+image of a language $L$ by a rational transduction $f$. A class of languages
+$\mathcal{C}$ is \intro{closed under rational transductions} when, for every
+language $L \in \mathcal{C}$, and every rational transduction $f \colon
+\Sigma^* \to \Gamma^*$, the language $f(L)$ is in $\mathcal{C}$.\todo{check these definitions ...}
+
+\begin{proofof}{infix-wqo-is-emptiness:thm}
+	\cref{emptiness-decidable} $\Rightarrow$ \cref{wqo-infix-decidable}. We aim to make the inclusion test of \cref{bounded-language:eq} effective.
+
+    Let $R = \bigcup_{x,y \in \Sigma^{\leq n}} \bigcup_{u \in \Sigma^{\leq m
+    \times N_0}} \InfPeriodChain{x} u \InfPeriodChain{y} \cup
+    \InfPeriodChain{x}u \cup u\InfPeriodChain{x}$. For any concrete values of
+    the bounds $n$, $m$, and $N_0$, this language is regular. As $\mathcal{C}$
+    is \kl{closed under rational transductions}, we can therefore reduce the
+    inclusion to emptiness of $L \cap \overline{R}$.\todo{explain that $L
+    \mapsto L \cap R$ is a rational  transduction?} However, we need to find
+    these bounds first.
+
+    To determine values for $n$ and $m$, we first test if $L$ is bounded. As
+    the algorithm in \cite[Section 4.2]{ASZZ24} is effective, this yields words
+    $w_1, \ldots w_n$ such that $L \subseteq w_1^* \cdots w_n^*$ and therefore
+    yields also the bounds in questions.
+
+    To determine the value for $N_0$, we first computer maximal subsets $S
+    \subseteq [n]$ such that $w_i$ for $i \in S$ are simultaneously unbounded.\todo{define this}
+    We do this by applying a transduction mapping $w_i$ to some fresh letter
+    $a_i$ if $i \in S$ and $\varepsilon$ otherwise, and testing for
+    simultaneous unboundedness \cite[Section 4.1]{ASZZ24}. Given a candidate
+    bound $n_0$ for $N_0$, we construct the languages $L_{S,n_0} =
+    w_1^{\circ_1}\cdots w_n^{\circ_n}$ where $\circ_i = *$ if $i \in S$ and
+    $\circ_i = \leq n_0$ otherwise. We then test if $L \subseteq \bigcup_{S
+    \text{ maximal}} L_{S,n_0}$. If yes, $n_0$ is our value for $N_0$,
+    otherwise, we increase it and repeat the construction. As we chose maximal
+    sets $S$, this procedure will eventually terminate.
+
+    \cref{wqo-infix-decidable} $\Rightarrow$ \cref{wqo-prefix-decidable}. We
+    just consider the transduction
+
+    \cref{wqo-prefix-decidable} $\Rightarrow$
+    \cref{emptiness-decidable}. We consider the transduction $\mathcal T
+    = \Sigma^* \times \{a, b\}$.\todo{Define this transduction explicitly} Then
+    for any language $L \in \mathcal C$, $\mathcal TL$ is well-quasi-ordered by
+    infix if and only if $L$ is empty.
 \end{proofof}
-
-For the next result, we impose a mild restriction by requiring the language class to be closed under rational transductions.
-This property holds for most common language classes, including all known \kl{amalgamation systems}.
-
-\begin{theorem}
-	\label{infix-wqo-is-emptiness:thm}
-	Let $\mathcal{C}$ be a class of languages recognized by amalgamation systems and closed under rational transductions. Then the following are equivalent:
-
-	\begin{enumerate}
-		\item\label{wqo-infix-decidable} \kl[well-quasi-order]{Well-quasi-orderedness} of the \kl{infix relation} is decidable.
-		\item\label{emptiness-decidable} Emptiness is decidable.
-	\end{enumerate}
-\end{theorem}
-
-\begin{proof}
-	\cref{emptiness-decidable} $\Rightarrow$ \cref{wqo-infix-decidable}. We aim to make the inclusion test of \cref{amalgamation-infix-wqo-inclusion:eqn} effective.
-
-	Let $R = \bigcup_{x,y \in \Sigma^{\leq n}} \bigcup_{u \in \Sigma^{\leq m \times N_0}} \InfPeriodChain{x} u \InfPeriodChain{y} \cup \InfPeriodChain{x}u \cup u\InfPeriodChain{x}$.
-For any concrete values of the bounds $n$, $m$, and $N_0$, this language is regular.
-As $\mathcal{C}$ is closed under rational transductions, we can therefore reduce the inclusion to emptiness of $L \cap \overline{R}$.
-However, we need to find these bounds first.
-
-	To determine values for $n$ and $m$, we first test if $L$ is bounded.
-As the algorithm in \cite[Section 4.2]{ASZZ24} is effective, this yields words $w_1, \ldots w_n$ such that $L \subseteq w_1^* \cdots w_n^*$ and therefore yields also the bounds in questions.
-
-	To determine the value for $N_0$, we first computer maximal subsets $S \subseteq [n]$ such that $w_i$ for $i \in S$ are simultaneously unbounded.
-We do this by applying a transduction mapping $w_i$ to some fresh letter $a_i$ if $i \in S$ and $\varepsilon$ otherwise, and testing for simultaneous unboundedness \cite[Section 4.1]{ASZZ24}.
-Given a candidate bound $n_0$ for $N_0$, we construct the languages $L_{S,n_0} = w_1^{\circ_1}\cdots w_n^{\circ_n}$ where $\circ_i = *$ if $i \in S$ and $\circ_i = \leq n_0$ otherwise.
-We then test if $L \subseteq \bigcup_{S \text{ maximal}} L_{S,n_0}$.
-If yes, $n_0$ is our value for $N_0$, otherwise, we increase it and repeat the construction.
-As we chose maximal sets $S$, this procedure will eventually terminate.
-
-	\cref{wqo-infix-decidable} $\Rightarrow$ \cref{emptiness-decidable}. We 
-	consider the transduction $\mathcal T = \Sigma^* \times \{a, b\}$. Then 
-	for any language $L \in \mathcal C$, $\mathcal TL$ is well-quasi-ordered 
-	by infix if and only if $L$ is empty.
-\end{proof}

### 2024-10-17 src/infixes.tex
moving from amalgam to infixes

@@
 
-\todo{Delete this section: the \cref{infix-embedding:thm} is already
-    known, we should just restate it. The
-    \cref{infix-finite-automata:thm} is a direct consequence of
-    \cref{infix-amalgamation:thm} and \cref{bounded-language:thm}.
-    I don't know if we should keep it for clarity, but it does not 
-    seems to be necessary to me.
-}
-
-\AP
-In this section, we study languages equipped with the \kl{infix relation}. As
-opposed to the \kl{prefix} and \kl{suffix} relations, the \kl{infix relation}
-can yield to very complicated \kl{well-quasi-ordered} languages. Formally, the
-upcoming \cref{infix-embedding:thm} shows that \emph{any} countable
-partial-ordering with finite initial segments can be embedded into the infix
-relation of a language. To make the former statement precise,
-Let us recall that an \intro{order embedding} from a quasi-ordered set $(X,
-\preceq)$ into a quasi-ordered set $(Y, \preceq')$ is a function $f \colon X
-\to Y$ such that for all $x, y \in X$, $x \preceq y$ if and only if $f(x)
+\AP In this section, we study languages equipped with the \kl{infix relation}.
+As opposed to the \kl{prefix} and \kl{suffix} relations, the \kl{infix
+relation} can lead to very complicated \kl{well-quasi-ordered} languages.
+Formally, the upcoming \cref{infix-embedding:thm} shows that
+\emph{any} countable partial-ordering with finite initial segments can be
+embedded into the infix relation of a language. To make the former statement
+precise, let us recall that an \intro{order embedding} from a quasi-ordered set
+$(X, \preceq)$ into a quasi-ordered set $(Y, \preceq')$ is a function $f \colon
+X \to Y$ such that for all $x, y \in X$, $x \preceq y$ if and only if $f(x)
 \preceq' f(y)$. When such an embedding exists, we say that $X$ \reintro{embeds
@@
 y$ and $y \preceq x$ implies $x = y$. 
-
-The following result due to Kuske shows the structural complexity of the \kl{infix relation}
-
 \begin{lemma}{\cite[Lemma 5.1]{DBLP:journals/ita/Kuske06}}
@@
 
-As a consequence of \cref{infix-embedding:thm}, we cannot replay proofs of
-\cref{prefixes:sec}, and will actually need to leverage some regularity of the
-languages to obtain a characterization of \kl{well-quasi-ordered} languages
-under the \kl{infix relation}. Let us first play this game for languages that
-are recognized by finite automata. We assume that the reader is familiar with
-the notions of deterministic finite automaton and regular languages, and refer
-to the book of Thomas for a comprehensive introduction to the subject
-\cite{THOM97}. The main goal of the remainder of this section is to prove the
-following \cref{infix-finite-automata:thm}.
+\AP As a consequence of \cref{infix-embedding:thm}, we cannot replay
+proofs of \cref{prefixes:sec}, and will
+actually need to leverage some regularity of the languages to obtain a
+characterization of \kl{well-quasi-ordered} languages under the \kl{infix
+relation}. This regularity will be imposed through the notion of \intro{bounded
+languages}, i.e., languages $L \subseteq \Sigma^*$ such that there exists words
+$w_1, \dots, w_n$ satisfying $L \subseteq w_1^* \cdots w_n^*$.
 
-\begin{theorem}[restate=infix-finite-automata:thm,label=infix-finite-automata:thm]
-    \proofref{infix-finite-automata:thm}
-    Let $L \subseteq \Sigma^*$ be a language recognized by a finite automaton.
-    Then $L$ is \kl{well-quasi-ordered} by the \kl{infix relation} if and only if $L$ is
-    a finite union of \kl{chains} for the \kl{infix relation}.
+\begin{theorem}[restate=bounded-language:thm,label=bounded-language:thm]
+    %\label{bounded-language:thm}
+    Let $L$ be a \kl{bounded language} of $\Sigma^*$. Then,
+    $L$ is a \kl{well-quasi-order} when endowed with the 
+    \kl{infix relation} if and only if it included in a finite union of 
+    products $S_i \cdot P_i$ where 
+    $S_i$ is a \kl{chain} for the \kl{suffix relation}, and 
+    $P_i$ is a \kl{chain} for the \kl{prefix relation},
+    for all $1 \leq i \leq n$.
 \end{theorem}
 
-\AP In order to prove \cref{infix-finite-automata:thm}, we will perform some
-preliminary analysis on the structure of an automaton recognizing a
-well-quasi-ordered language under the infix relation, which will be powered by
-folklore results on \emph{periodic} words. Let us recall that a non-empty word
-$w \in \Sigma^+$ is \intro(word){periodic} with period $x \in \Sigma^*$ if
-there exists a $p \in \Nat$ such that $w \infleq x^p$. The \intro{periodic
-length} of a word $u$ is the minimal length of a word $x$ such that $u$ is an
-\kl{infix} of $x^p$ for some $p \in \Nat$ and $x \in \Sigma^+$. We will
-essentially rely on the following result on periodic words.
+Let us first remark that if $S$ is a \kl{chain} for the \kl{suffix relation}
+and $P$ is a \kl{chain} for the \kl{prefix relation}, then $SP$ is
+\kl{well-quasi-ordered} for the \kl{infix relation}. This proves the (easy)
+right-to-left implication of \cref{bounded-language:thm}. Furthermore, any
+language $L$ included in a finite union of products $S_i \cdot P_i$ where $S_i$
+is in fact a \kl{bounded language}.
+
+\AP In order to prove the (difficult) left-to-right implication of
+\cref{bounded-language:thm}, we will rely heavily on the
+combinatorics of periodic words. Let us recall that a non-empty word $w \in
+\Sigma^+$ is \intro(word){periodic} with period $x \in \Sigma^*$ if there
+exists a $p \in \Nat$ such that $w \infleq x^p$. The \intro{periodic length} of
+a word $u$ is the minimal length of a word $x$ such that $u$ is an \kl{infix}
+of $x^p$ for some $p \in \Nat$ and $x \in \Sigma^+$.
 
@@
 
-The upcoming \cref{powers-infixes:cor} is based on the observation that if $S$
-is a \kl{chain} for the \kl{suffix relation} and $P$ is a \kl{chain} for the
-\kl{prefix relation}, then $SP$ is a \kl{chain} for the \kl{infix relation}.
-
 \begin{corollary}
@@
 are interesting for the \kl{infix relation} is that they naturally create
-\kl{chains}. Indeed, if $x \in \Sigma^+$ is a finite word, then $\setof{x^p}{p
-\in \Nat}$ is a \kl{chain} for the \kl{infix relation}. Note that in general,
-the \kl{downwards closure} of a chain is \emph{not} a chain. However, for the
-chains generated using periodic words, the \kl{downwards closure}
-$\dwset[\infleq]{\setof{x^p}{p \in \Nat}}$ is a \emph{finite union} of
-\kl{chains}. Because this set will appear in bigger equations, we introduce the
-shorter notation $\intro*\InfPeriodChain{x}$ for the set of \kl{infixes} of
-words of the form $x^p$, where $p$ ranges over $\Nat$.
+\kl{chains} for the \kl{prefix} and \kl{suffix} relations. Indeed, if $x \in
+\Sigma^+$ is a finite word, then $\setof{x^p}{p \in \Nat}$ is a \kl{chain} for
+the \kl{infix relation}. Note that in general, the \kl{downwards closure} of a
+chain is \emph{not} a chain. However, for the chains generated using periodic
+words, the \kl{downwards closure} $\dwset[\infleq]{\setof{x^p}{p \in \Nat}}$ is
+a \emph{finite union} of \kl{chains}. Because this set will appear in bigger
+equations, we introduce the shorter notation $\intro*\InfPeriodChain{x}$ for
+the set of \kl{infixes} of words of the form $x^p$, where $p$ ranges over
+$\Nat$.
 
@@
     $\InfPeriodChain{x}u \InfPeriodChain{y}$ is a finite union 
-    of \kl{chains} for the \kl{infix} relation.
+    of \kl{grids} for the \kl{infix} relation.
 \end{proof}
@@
 
-We are now ready to restate and prove our main theorem, proven by combining
-\cref{pumping-periods:lem} together with classical pumping arguments on finite
-state automata.
-
-\begin{proofof}{infix-finite-automata:thm}
-    Let $w \in L$, because $Q$ is finite, there exists
-    a factorization of $w$
-    into words $(\prod_{i = 1}^k u_i v_i) u_{k+1}$
-    such that 
-    for all $1 \leq i \leq k+1$, $\card{u_i} \leq \card{Q}$,
-    for all $1 \leq i \leq k$, $1 \leq \card{v_i} \leq \card{Q}$,
-    and satisfying 
-    that $w[\vec{X}] \defined 
-    (\prod_{i = 1}^k u_i v_i^{X_i}) u_{k+1}$
-    belongs to $L$ for all choices of values $\vec{X} \in \Nat^k$.
-
-    Applying \cref{pumping-periods:lem}, we conclude that 
-    there exists $x,y \in \Sigma^+$ of size at most $\card{Q}$
-    such that 
-    $w \in u_1 \InfPeriodChain{x} \cup \InfPeriodChain{y} u_{k+1}
-    \cup \bigcup_{1 \leq i \leq k+1} \InfPeriodChain{x} u_i \InfPeriodChain{y}$.
-    In particular, we conclude that
-    \begin{equation}
-        \label{infix-automata:eq}
-        L
-        \subseteq
-        \bigcup_{x,y,u\in \Sigma^{\leq \card{Q}}}
-        u \InfPeriodChain{x}
-        \cup 
-        \InfPeriodChain{x} u
-        \cup
-        \InfPeriodChain{x} u \InfPeriodChain{y}
-        \quad .
-    \end{equation}
-
-    Now, thanks to \cref{inf-period-union-chains:lem},
-    this happens to be a finite union of \kl{chains}
-    for the \kl{infix relation}.
-\end{proofof}
-
-\begin{corollary}
-    \label{reg-wqo-decidable:cor}
-    Given a regular language $L$, it is decidable whether
-    $L$ is \kl{well-quasi-ordered} for the \kl{infix relation}.
-\end{corollary}
-\begin{proof}
-    It suffices to decide whether 
-    \cref{infix-automata:eq} holds for the language $L$.
-    This is an inclusion of regular languages, which is decidable.
-\end{proof}
-
-Let $(X,\preceq)$ be a quasi-ordered set, and $L \subseteq X$ be such that $(L,
-\preceq)$ is \kl{well-quasi-ordered}. It is not true in general that
-$(\dwset{L}, \preceq)$ is \kl{well-quasi-ordered}. In the case of $(\Sigma^*,
-\infleq)$ a typical example is to start from an infinite \kl{antichain} $A$,
-together with an enumeration $\seqof{w_i}$ of $A$, and build the language $L
-\defined \setof{ \prod_{i = 0}^n w_i }{ i \in \Nat }$. By definition, $L$ is a
-\kl{chain} for the \kl{infix} ordering, hence \kl{well-quasi-ordered}. However,
-$\dwset[\infleq]{L}$ contains $A$, and is therefore not
-\kl{well-quasi-ordered}. As a fun corollary of our analysis, we conclude that
-this particular example cannot be realized by a regular language $A$.
-
-\begin{corollary}
-    \label{reg-wqo-iff-dwclosed-wqo:cor}
-    Given a regular language $L$, it is \kl{well-quasi-ordered} 
-    for the \kl{infix relation} if and only if 
-    $\dwset[\infleq] L$ is.
-\end{corollary}
-
-\begin{theorem}
-    \label{bounded-language:thm}
-    Let $L$ be a \kl{bounded language} of $\Sigma^*$. Then,
-    $L$ is a \kl{well-quasi-order} when endowed with the 
-    \kl{infix relation} if and only if it is a finite union of \kl{grids}.
-\end{theorem}
-\begin{proof}
+\begin{proofof}{bounded-language:thm}
     Let $w_1, \dots, w_n$ be such that
@@
     $N_0 \times m$ by definition, and we conclude that:
-    \begin{equation*}
+    \begin{equation}
+        \label{bounded-language:eq}
         L \subseteq 
@@
         \quad .
-    \end{equation*}
-\end{proof}
+    \end{equation}
+\end{proofof}
 
-Note that this immediately generalizes \cref{reg-wqo-iff-dwclosed-wqo:cor}
-to all \kl{bounded languages}.
+Let us now discuss the implications of this characterization in terms of
+\kl{downwards closures}: if $L$ is a \kl{bounded language}, then considering
+$L$ or its \kl{downwards closure} is equivalent with respect to being
+\kl{well-quasi-ordered} by the \kl{infix relation}.
 
@@
 \end{corollary}
-
-In the light of \cref{reg-wqo-iff-dwclosed-wqo:cor,bounded-wqo-dwclosed:cor}
-one may wonder whether the \kl{downwards closed}
-languages for the \kl{infix relation} satisfy a similar characterization in
-terms of union of \kl{grids} as stated in
-\cref{infix-finite-automata:thm,bounded-language:thm}. Let us
-first remark that from our previous results we obtain the following lemma that
-illustrates how all the previous characterizations of \kl{well-quasi-ordered}
-languages for the \kl{infix relation} are collapsing in the case of
-\kl{downwards closed} languages.
-
-\begin{lemma}
-    \label{dwclosed-infixes-wqo:lem}
-    Let $L \subseteq \Sigma^*$ be a \kl{downwards closed} language for the
-    \kl{infix relation} that is \kl{well-quasi-ordered}. Then, the following
-    are equivalent:
-    \begin{enumerate}
-        \item $L$ is \kl{bounded},
-        \item $L$ is \kl{regular},
-        \item $L$ is a finite union of \kl{chains} and \kl{grids}.
-    \end{enumerate}
-\end{lemma}
 \begin{proof}
-    TODO.
+    Because $L \subseteq \dwset[\infleq]{L}$, the right-to-left implication
+    is trivial.
+    For the left-to-right implication, let us assume that $L$ is a
+    \kl{well-quasi-ordered} language for the \kl{infix relation}.
+    Then $L$ is included in a finite union 
+    of products of \kl{chains}:
+    \begin{equation*}
+        L \subseteq \bigcup_{i = 1}^n S_i \cdot P_i \quad .
+    \end{equation*}
+    Remark that the \kl{downwards closure} of a product of two \kl{chains}
+    is a product of two chains.
+    As a consequence, we conclude that $\dwset[\infleq]{L}$ is itself included
+    in a finite union of products of \kl{chains}.
+    Note that this also proves that $\dwset[\infleq]{L}$ is a \kl{bounded language},
+    hence that it is \kl{well-quasi-ordered} by the \kl{infix relation} 
+    via
+    \cref{bounded-language:thm}.
 \end{proof}
@@
 One may think that all \kl{downwards closed} languages for the \kl{infix
-relation} that are \kl{well-quasi-ordered} are \kl{regular}. Note that this is
-what happens in the case of the \kl{subword embedding}, where any \kl{downwards
-closed} language for this relation is characterized by finitely many excluded
-patterns, hence \kl{regular}. However, this is not the case for the \kl{infix
-relation}, as we will now illustrate with the following two examples.
+relation} that are \kl{well-quasi-ordered} are \kl(language){bounded}. Note
+that this is what happens in the case of the \kl{subword embedding}, where any
+\kl{downwards closed} language for this relation is characterized by finitely
+many excluded \kl{subwords}, hence provides a \kl{bounded language}. However, this
+is not the case for the \kl{infix relation}, as we will now illustrate with the
+following two examples.
 
@@
     relation}, \kl{well-quasi-ordered} for the \kl{infix relation}, but is not
-    \kl{regular}.
+    \kl(language){bounded}.
 \end{lemma}
@@
     \kl{downwards closed} for the \kl{infix relation}. Let us argue that $L$ is
-    \kl{well-quasi-ordered} for the \kl{infix relation}, but is not \kl{regular}.
+    \kl{well-quasi-ordered} for the \kl{infix relation}, but is not \kl(language){bounded}.
 
@@
     of $u_k$, hence the sequence $\seqof{u_i}$ is \kl(sequence){good}.
+\end{proof}
 
-    We conclude thanks to
-    \cref{dwclosed-infixes-wqo:lem}, that $L$ is not \kl{regular}.
+One may refine our analysis of the \kl{Thue-Morse sequence} to obtain 
+precise bounds on the \kl{ordinal invariants} of its language of \kl{infixes}.
+
+\begin{lemma}
+    \label{thue-morse-ordinal:lemma}
+    Let $L$ be the language of \kl{infixes} of the \kl{Thue-Morse sequence}.
+    Then, the \kl{ordinal invariants} of $L$ are as follows:
+    the \kl{maximal order type} of $L$ is $\omega$,
+    the \kl{ordinal height} of $L$ is $\omega$,
+    the \kl{ordinal width} of $L$ is $\omega$.
+\end{lemma}
+\begin{proof}
+    Let us prove that these are upper bounds for the \kl{ordinal invariants} of
+    $L$. For the \kl{ordinal height}, it is true for any language $L$.
+    For the \kl{maximal order type}, we remark that
+    the maximal length of a \kl{bad sequence} is determined by its first element,
+    hence that it is at most $\omega$.
+    Finally, because the \kl{ordinal width} is at most the \kl{maximal order type},
+    we conclude that the \kl{ordinal width} is also at most $\omega$.
+
+    Now, let us prove that these bounds are tight. It is clear that
+    $\oHeight{L} = \omega$: given any number $n \in \Nat$, one can construct a
+    \kl{decreasing sequence} of words in $L$ of length $n$, for instance by
+    considering the first $n$ prefixes of the \kl{Thue-Morse sequence} by
+    decreasing size.
+    Let us now prove that $\oWidth{L} = \omega$. Assume by contradiction that
+    $\oWidth{L}$ is finite. Then, $L$ can be written as a finite union of
+    \kl{chains} for the \kl{infix relation}, and in particular, $L$ is
+    \kl{bounded}, which is absurd by \cref{thue-morse:lemma}.
+    Finally, because the \kl{ordinal width} is at most the \kl{maximal order
+    type}, we conclude that the \kl{maximal order type} of $L$ is also $\omega$.
 \end{proof}
+
+This suggests the following conjecture which states that for \kl{downwards
+closed} languages, the \kl{ordinal invariants} are relatively small.
+
+\begin{conjecture}
+    \label{small-ordinal-invariants:conj}
+    Let $L$ be a \kl{well-quasi-ordered} language for the \kl{infix relation}
+    that is \kl{downwards closed}.
+    Then, the \kl{maximal order type} of $L$ is at most $\omega^3$,
+    its \kl{ordinal height} is at most $\omega$,
+    and its \kl{ordinal width} is at most $\omega^2$.
+\end{conjecture}

### 2024-10-18 src/amalgam.tex
adding more and more todos

@@
     \kl{amalgamation system}.
-    Then $L$ is well-quasi-ordered by the infix relation if and only if $L$ is
+    If $L$ is well-quasi-ordered by the infix relation then $L$ is
     \kl(language){bounded}.
@@
 
+\todo[inline]{There are missing effective restrictions in the statement of the theorem: we should
+have descriptions of the languages, (effective) stability under transductions etc.}
+
 \begin{theorem}[restated=infix-wqo-is-emptiness:thm,label=infix-wqo-is-emptiness:thm]
 % 	\label{infix-wqo-is-emptiness:thm}
-	Let $\mathcal{C}$ be a class of languages recognized by 
-    \kl{amalgamation systems}
-    and closed under
+	Let $\mathcal{C}$ be an \kl{effective amalgamative class} of languages
+    closed under
     \kl{rational transductions}. 
@@
 \subsection{Amalgamation Systems}
-\label{infixes-amalgamation:subsec}
+\label{amalgamation-systems:subsec}
 
@@
 
-For the next result, we impose a mild restriction by requiring the language
+\AP For the next result, we impose a mild restriction by requiring the language
 class to be closed under \kl{rational transductions}. This property holds for
-most common language classes, including all known \kl{amalgamation systems}.
+most common language classes, including all known \kl{amalgamation systems}. A
+\intro{rational transduction} is a function $f \colon \Sigma^* \to \Gamma^*$
+that can be defined by a finite automaton with output, or equivalently by a
+finite set of rules of the form $u \to v$ where $u \in \Sigma^*$ and $v \in
+\Gamma^*$. We write $f(L) = \setof{f(w)}{w \in L}$ for the image of a language
+$L$ by a rational transduction $f$. A class of languages $\mathcal{C}$ is
+\intro{closed under rational transductions} when, for every language $L \in
+\mathcal{C}$, and every rational transduction $f \colon \Sigma^* \to \Gamma^*$,
+the language $f(L)$ is in $\mathcal{C}$.\todo{check these definitions ... Do we allow
+the alphabet to vary?}
 
-\AP A class of languages $\mathcal{C}$ is \intro{recognized by amalgamation
-systems} when, for every language $L \in \mathcal{C}$, there exists an
+\AP A class $\mathcal{C}$ of languages is an \intro{effective amalgamative
+class} whenever there exists a finite presentation of the languages \dots\todo{
+define formally}. And for every $L \in \mathcal{C}$, there exists an
 \kl{amalgamation system} $(\Sigma, R, E, \canrun)$ such that $L$ is the
@@
 
-\AP A \intro{rational transduction} is a function $f \colon \Sigma^* \to
-\Gamma^*$ that can be defined by a finite automaton with output, or
-equivalently by a finite set of rules of the form $u \to v$ where $u \in
-\Sigma^*$ and $v \in \Gamma^*$. We write $f(L) = \setof{f(w)}{w \in L}$ for the
-image of a language $L$ by a rational transduction $f$. A class of languages
-$\mathcal{C}$ is \intro{closed under rational transductions} when, for every
-language $L \in \mathcal{C}$, and every rational transduction $f \colon
-\Sigma^* \to \Gamma^*$, the language $f(L)$ is in $\mathcal{C}$.\todo{check these definitions ...}
+\AP A class $\mathcal{C}$ of languages is \intro(amalg){strongly effective}
+whenever the emptiness problem for languages in $\mathcal{C}$ is decidable.
+Note that this implies boundedness and simultaneous boundedness of languages in
+$\mathcal{C}$ is decidable \cite{ASZZ24}. \todo{define formally}
 
@@
     \cref{wqo-infix-decidable} $\Rightarrow$ \cref{wqo-prefix-decidable}. We
-    just consider the transduction
+    just consider the transduction \todo{fresh symbols? can we assume we have those??}
 

### 2024-10-18 src/conclusion.tex
adding more and more todos

--- 2024-08-12 src/conclusion.tex
@@
 the minimal automaton recognizing the language.
+
+\subparagraph{Lexicographic orderings} There is another natural ordering on
+words, the \emph{lexicographic ordering}, which does not fit well in our
+current framework because it is always of \kl{ordinal width} $0$. However, the
+order-type of the lexicographic ordering over \kl{regular languages} has
+already been investigated in the context of infinite words, and it would be
+interesting to see if one can extend these results to decide whether such an
+ordering is \kl{well-founded} for languages recognized by \kl{amalgamation
+systems}. \todo{Find references for that?}

### 2024-10-18 src/infixes.tex
adding more and more todos

@@
 y$ and $y \preceq x$ implies $x = y$. 
+The embedding defined in \cref{infix-embedding:thm} is illustrated
+for the \kl{subword relation} in \cref{infix-embedding:fig}.
 \begin{lemma}{\cite[Lemma 5.1]{DBLP:journals/ita/Kuske06}}
@@
     \begin{enumerate}
-        \item \label{infix-embedding-embeds:item} 
+        \item 
             $X$ \kl{embeds into} $(\Sigma^*, \infleq)$,
-        \item \label{infix-embedding-count:item}
+        \item 
             $X$ is countable, and for every $x \in X$,
@@
     \qedhere
-\end{proof}
-
-\begin{corollary}
-    \label{inf-period-union-chains:lem}
-    Let $x,u,y \in \Sigma^*$ be words.  The following 
-    are finite unions of \kl{grids} for the infix relation:
-    $\InfPeriodChain{x}u$, $u \InfPeriodChain{x}$,
-    and $\InfPeriodChain{x} u \InfPeriodChain{y}$.
-\end{corollary}
-\begin{proof}
-    Because $\InfPeriodChain{x}$ is a finite union of \kl{chains} for the \kl{suffix}
-    relation (rsing \cref{inf-period-chain:lem}), we conclude that $\InfPeriodChain{x}u$ is one too for
-    the \kl{infix} relation. This holds similarly for $u \InfPeriodChain{x}$.
-    In the case of $\InfPeriodChain{x} u \InfPeriodChain{y}$,
-    we remark that $\InfPeriodChain{x} u$ is a finite union of chains
-    for the \kl{suffix relation},
-    and that $\InfPeriodChain{y}$ is one for the \kl{prefix relation}
-    (using again \cref{inf-period-chain:lem}).
-    As a consequence,
-    $\InfPeriodChain{x}u \InfPeriodChain{y}$ is a finite union 
-    of \kl{grids} for the \kl{infix} relation.
 \end{proof}
@@
 
-    Assume by contradiction that $L$ is \kl{bounded}. In this case, there exist
+    Assume by contradiction that $L$ is \kl(language){bounded}. In this case, there exist
     words $w_1, \dots, w_k \in \Sigma^*$ such that $L \subseteq w_1^* \cdots
@@
     \kl{chains} for the \kl{infix relation}, and in particular, $L$ is
-    \kl{bounded}, which is absurd by \cref{thue-morse:lemma}.
+    \kl(language){bounded}, which is absurd by \cref{thue-morse:lemma}.
     Finally, because the \kl{ordinal width} is at most the \kl{maximal order

### 2024-10-18 src/introduction.tex
adding more and more todos

@@
 runtime of the algorithm is deeply related to the ``complexity'' of the
-underlying quasi-order \todo{CITE???}. One way to measure this complexity is to
-consider its so-called \kl{ordinal invariants}: for instance, the
-\kl{maximal order type} (or m.o.t.), originally defined by De Jongh and
-Parikh \cite{dejongh77}, is the order type of the maximal linearization of a
+underlying quasi-order~\cite{SCHMITZ17}. One way to measure this complexity is
+to consider its so-called \kl{ordinal invariants}: for instance, the
+\kl{maximal order type} (or m.o.t.), originally defined by De Jongh and Parikh
+\cite{dejongh77}, is the order type of the maximal linearization of a
 well-quasi-ordered set. In the case of a finite set, the m.o.t. is precisely
 the size of the set. Better runtime bounds were obtained by considering two
-other parameters\todo{cite sylvain width/ HDR?}: the \kl{ordinal height}
-introduced by Schmidt \cite{schmidt81}, and the \kl{ordinal width} of Kříž
-and Thomas \cite{kriz90b}. Therefore, when characterizing
-\kl{well-quasi-ordered} languages, we will also be interested in computing
-bounds on their \kl{ordinal invariants}. We refer to the
-\cref{ordinal-invariants:subsec} for a more detailed introduction to these
-parameters and ordinal computations in general. For the moment, let us just
-remark that all languages $L \subseteq \Sigma^*$ have an \kl{ordinal height} of
-at most $\omega$.
+other parameters~\cite{SCHMITZ19}: the \kl{ordinal height} introduced by
+Schmidt \cite{schmidt81}, and the \kl{ordinal width} of Kříž and Thomas
+\cite{kriz90b}. Therefore, when characterizing \kl{well-quasi-ordered}
+languages, we will also be interested in computing bounds on their \kl{ordinal
+invariants}. We refer to \cref{ordinal-invariants:subsec}
+for a more detailed introduction to these
+parameters and ordinal computations in general.
 
@@
 	\setcounter{enumi}{3}
-    \item A \kl{bounded language} $L \subseteq \Sigma^*$ is \kl{well-quasi-ordered} by the \kl{infix relation} if and only if it is a finite union of languages $S_i \cdot P_i$, where each $S_i$ 
+    \item A \kl{bounded language} $L \subseteq \Sigma^*$ is \kl{well-quasi-ordered} by the \kl{infix relation} if and only if it is (included in) a finite union of languages $S_i \cdot P_i$, where each $S_i$
         is a \kl{chain} for the \kl{suffix relation},
@@
         each $P_i$ is a \kl{chain} for the \kl{prefix relation}.
+
+        \hfill
+        (\cref{bounded-language:thm} p.\pageref{bounded-language:thm})
 \end{enumerate}
@@
 
-This result directly translates into bounds on the possible \kl{ordinal invariants}
-of such languages similarly as for the \kl{prefix relation}.
+This result directly translates into upper bounds on the possible \kl{ordinal invariants}
+of such languages similarly as for the \kl{prefix relation}. Notice that these
+upper bounds are significantly smaller than \kl{ordinal invariants} of the \kl{subword
+relation}.
 {
@@
 
-\AP
-Finally, we provide a decision procedure for checking whether a language is
-\kl{well-quasi-ordered}, which takes as inputs languages recognized
-by \kl{amalgamation systems} \cite{ASZZ24}. Such systems will be
-formally introduced in \cref{amalgamation-systems:subsec}, but for the moment
-let us just say that they include many classical computational models such as
-finite automata, context-free grammars, and petri nets \cite{ASZZ24}. 
+\AP Finally, we provide a decision procedure for checking whether a language is
+\kl{well-quasi-ordered}, which takes as inputs languages recognized by
+\kl{amalgamation systems} \cite{ASZZ24}. Such systems will be formally
+introduced in
+\cref{amalgamation-systems:subsec}, but
+for the moment let us just say that they include many classical computational
+models such as finite automata, context-free grammars, and Petri nets
+\cite{ASZZ24}. This provides us with a \emph{meta-algorithm} for deciding
+whether a given language is \kl{well-quasi-ordered} by the \kl{prefix
+relation}, the \kl{suffix relation}, or the \kl{infix relation} under mild
+effective restrictions on the computational model that we call an \kl{effective
+amalgamative class}, the formal statement of which we defer to
+\cref{infixes-amalgamation-effective:subsec}.
+
 {
@@
     \setcounter{enumi}{5}
-    \item There is a decision procedure for checking whether a language $L$ 
-        recognized by an \kl{amalgamation system} is \kl{well-quasi-ordered}
-        by the \kl{prefix relation}, the \kl{suffix relation},
-        or the \kl{infix relation}. Furthermore, if it is \kl{well-quasi-ordered}, 
+    \item 
+        Let $L$ be a language recognized by an \kl{amalgamation system}.
+        If $L$ is \kl{well-quasi-ordered} by the \kl{infix relation},
         then $L$ is a \kl{bounded language}.
+
+        \hfill (\cref{infix-amalgamation:thm} p.\pageref{infix-amalgamation:thm})
+    \item 
+        Let $\mathcal{C}$ be a \kl{strongly effective amalgamative class} of languages.
+        There is a decision procedure that takes as input a language $L \in \mathcal{C}$,
+        and decides whether $L$ is \kl{well-quasi-ordered} by the \kl{prefix relation},
+        the \kl{suffix relation}, or the \kl{infix relation}. 
         \hfill (\cref{infix-wqo-is-emptiness:thm} p.\pageref{infix-wqo-is-emptiness:thm})
@@
 }
-Notice that the above result implies that the hypothesis of
+Let us point out that the above result implies that the hypothesis of
 \kl{bounded languages} on the theoretical side is not a restriction at all in
-practice.
+practice. As a practical and more easily understandable 
+consequence,
+our generic decision procedure applies 
+to the class $\mathcal{C}_\text{aut}$ of all languages recognized by finite automata,
+and to the class $\mathcal{C}_\text{cfg}$ of all languages recognized by context-free grammars,
+which are both \kl{effective amalgamative classes}.
 
@@
 
-\subparagraph{Outline} We introduce in \cref{prelims:sec} the
+\subparagraph{Outline} 
+We introduce in \cref{prelims:sec} the
 necessary background on \kl{well-quasi-orders} and \kl{ordinal invariants}.
@@
 obtain the \kl[infix relation]{infix} analogue of \cref{prefixes:thm}
-specifically for regular languages
-(\cref{infix-finite-automata:thm}). Finally, we generalize this result to all
+specifically for \kl{bounded languages}
+(\cref{bounded-language:thm}). Finally, 
+we generalize this result to all
 \kl{amalgamation systems} in \cref{infixes-amalgamation:sec}
 in
-(\cref{infix-amalgamation:thm}).
+(\cref{infix-amalgamation:thm}),
+and provide a decision procedure for checking whether a language is
+\kl{well-quasi-ordered} by the \kl{infix relation} in
+this context (\cref{infix-wqo-is-emptiness:thm}).
 

### 2024-10-18 src/prelims.tex
adding more and more todos

@@
 some intuition on the decision problems, we will sometimes use the notion of
-\intro{finite automata}, \intro{regular languages}, and \intro{Monadic Second
-Order} ($\intro*\MSO$) logic on finite words, and assume is familiar with them.
-We refer to the textbook of \cite{THOM97} for a detailed introduction to this
+\intro{finite automata}, \intro{regular languages}, and Monadic Second Order
+logic ($\intro*\MSO$) over finite words, and assume is familiar with them. We
+refer to the textbook of \cite{THOM97} for a detailed introduction to this
 topic.
@@
 
-\paragraph*{Ordinal Invariants.} An ordinal is a well-founded ordered set. We
+\paragraph*{Ordinal Invariants.} 
+\label{ordinal-invariants:subsec} An ordinal is a well-founded ordered set. We
 use $\alpha, \beta, \gamma$ to denote ordinals, and use $\omega$ to denote the

### 2024-10-18 src/trash.tex
adding more and more todos

@@
 
-    Now, thanks to \cref{inf-period-union-chains:lem},
-    this happens to be a finite union of \kl{chains}
-    for the \kl{infix relation}.
+    Because $\InfPeriodChain{x}$ is a finite union of \kl{chains} for the \kl{suffix}
+    relation (using \cref{inf-period-chain:lem}), we conclude that $\InfPeriodChain{x}u$ is one too for
+    the \kl{infix} relation. This holds similarly for $u \InfPeriodChain{x}$.
+    In the case of $\InfPeriodChain{x} u \InfPeriodChain{y}$,
+    we remark that $\InfPeriodChain{x} u$ is a finite union of chains
+    for the \kl{suffix relation},
+    and that $\InfPeriodChain{y}$ is one for the \kl{prefix relation}
+    (using again \cref{inf-period-chain:lem}).
+    As a consequence,
+    $\InfPeriodChain{x}u \cdot \InfPeriodChain{y}$ is a finite union 
+    of products of \kl{chains}, as desired.
 \end{proofof}

### 2024-10-18 src/amalgam.tex
new proofs

--- 2024-10-17 src/amalgam.tex
@@
 
-\AP
-In this section, we are going to design an effective decision procedure for
+\AP In this section, we are going to design an effective decision procedure for
 \kl{well-quasi-ordering} by the \kl{infix relation}. To that end, the first
@@
 system to ``pump'' them and produce new runs: this is the usual pumping lemma
-in automata theory, and Ogden's lemma for context-free grammars.
-\todo{Cite the pumping lemma for VASS. Cite Ogden's lemma.}
+in automata theory, and Ogden's lemma for context-free grammars \cite{OGDEN68}.
+It turns out that there is a rather large family of systems for which pumping
+arguments based on so-called \emph{minimal runs} exist: they are called
+\emph{amalgamation systems} and were recently introduced by Anad, Schmitz,
+Schütze, and Zetzsche \cite{ASZZ24}.
 
-\AP It turns out that there is a rather large family of systems for which
-pumping arguments based on so-called \emph{minimal runs} exist: they are called
-\emph{amalgamation systems} and were recently introduced by Anad, Schmitz,
-Sch\"{u}tze, and Zetzsche \cite{ASZZ24}. Having done the heavy lifting on
-\kl{bounded languages}, the rest of this section is mostly devoted to the
-introduction of \kl{amalgamation systems} and collecting the necessary pumping
-argument they enjoy. Using this meta-proof, we will generalize
-\cref{infix-finite-automata:thm} to context-free grammars, languages recognized
-by vector addition systems with state (VASS), and more.
-\todo{unclear formulation}
-The goal of this
-section is not to introduce all of these systems, or justify their usefulness,
-but to state and prove the following two theorems.
+\AP Our first result, of theoretical nature, is that \kl{amalgamation systems}
+cannot define \kl{well-quasi-ordered} languages that are not
+\kl(language){bounded}. This implies that all the results of
+\cref{infixes-regular:sec}, and in
+particular \cref{bounded-language:thm}, can safely be transferred to
+\kl{amalgamation systems}.
 
@@
     \kl{amalgamation system}.
-    Then $L$ is well-quasi-ordered by the infix relation if and only if $L$ is
+    If $L$ is \kl{well-quasi-ordered} by the \kl{infix relation} then $L$ is
     \kl(language){bounded}.
@@
 
+\AP Our second focus is of practical nature: we want to give a decision
+procedure for being \kl{well-quasi-ordered}. This will require us to introduce
+\emph{effectiveness assumptions} on the \kl{amalgamation systems}. While most
+of them will be innocuous, an important consequence is that we have to consider
+\emph{classes of languages} rather than individual ones. Such classes will be
+called \kl{effective amalgamative classes} (\kcref{effective amalgamative
+classes}). In the following theorem, we prove that under such assumptions,
+testing \kl{well-quasi-ordering} is inter-reducible to testing whether a
+language of the class is empty.
+
 \begin{theorem}[restated=infix-wqo-is-emptiness:thm,label=infix-wqo-is-emptiness:thm]
 % 	\label{infix-wqo-is-emptiness:thm}
-	Let $\mathcal{C}$ be a class of languages recognized by 
-    \kl{amalgamation systems}
-    and closed under
-    \kl{rational transductions}. 
+	Let $\mathcal{C}$ be an \kl{effective amalgamative class} of languages.
     Then the following are equivalent:
@@
 
+\AP We say that a \kl{strongly effective amalgamative class} is an
+\kl{effective amalgamative class} for which the emptiness problem is decidable.
+Let us immediately remark that the class of \kl{regular languages} is
+\kl(amalg){strongly effective}, and so is the class of context-free languages.
+Therefore, \cref{infix-wqo-is-emptiness:thm} provides a concrete decision
+procedure for these classes.
+
 \subsection{Amalgamation Systems}
-\label{infixes-amalgamation:subsec}
+\label{amalgamation-systems:subsec}
 
@@
 \emph{minimal} runs are of finite size, and that all other runs are
-obtained by ``gluing" minimal runs together. This is exactly what
+obtained by ``gluing'' minimal runs together. This is exactly what
 \kl{amalgamation systems} are about.
@@
 
-One can leverage the same kind of reasoning to take the opposite direction, and
-notice that the notions of being a \kl{bounded language}, a \kl{regular
-language}, or being included in a finite union of products of \kl{chains} all
-collapse for \kl{well-quasi-ordered} and \kl{downwards closed} languages for
-the \kl{infix relation}.
+Leveraging a similar reasoning, we conclude that being a \kl{bounded language},
+a \kl{regular language}, or being included in a finite union of products of
+\kl{chains} all collapse for \kl{well-quasi-ordered} and \kl{downwards closed}
+languages for the \kl{infix relation}.
 
@@
 
+\AP Combining
+\cref{thue-morse:lemma,dwclosed-infixes-wqo:lem}, we can
+conclude that the collection of \kl{infixes} of the \kl{Thue-Morse sequence}
+cannot be recognized by \emph{any} \kl{amalgamation system}. To put this result
+in context, it was proven that the complement of the set $T$ of prefixes of the
+\kl{Thue-Morse sequence} is context-free, and conjectured that the same holds
+for the complement of $\LMorse$ \cite{BERST86}. To the knowledge of the
+authors, this conjecture remains open. Our remark provides a negative answer to
+this conjecture for the language $\LMorse$.
+
 \subsection{Effective Decision Procedures}
@@
 
-For the next result, we impose a mild restriction by requiring the language
-class to be closed under \kl{rational transductions}. This property holds for
-most common language classes, including all known \kl{amalgamation systems}.
+\AP Let us now introduce our effectiveness assumptions on \kl{amalgamation
+systems}. We follow the approach of \cite{ASZZ24} and require that an
+\kl{amalgamation system} $(\Sigma, R, E, \canrun)$ is effective when $R$ is
+recursively enumerable, the function $\canrun(\cdot)$ is computable, and for
+any two runs $r, s \in R$, the set $E(r,s)$ is computable.
 
-\AP A class of languages $\mathcal{C}$ is \intro{recognized by amalgamation
-systems} when, for every language $L \in \mathcal{C}$, there exists an
-\kl{amalgamation system} $(\Sigma, R, E, \canrun)$ such that $L$ is the
-language of this system.
+\todo[inline]{In this definition, the alphabet $\Sigma$ can be changed 
+(and we seem to actually use it in the proofs), it is strange.}
 
-\AP A \intro{rational transduction} is a function $f \colon \Sigma^* \to
-\Gamma^*$ that can be defined by a finite automaton with output, or
-equivalently by a finite set of rules of the form $u \to v$ where $u \in
-\Sigma^*$ and $v \in \Gamma^*$. We write $f(L) = \setof{f(w)}{w \in L}$ for the
-image of a language $L$ by a rational transduction $f$. A class of languages
-$\mathcal{C}$ is \intro{closed under rational transductions} when, for every
-language $L \in \mathcal{C}$, and every rational transduction $f \colon
-\Sigma^* \to \Gamma^*$, the language $f(L)$ is in $\mathcal{C}$.\todo{check these definitions ...}
+\AP We say that a class $\mathcal{C}$ of languages is an \kl{effective
+amalgamative class} whenever for every $L \in \mathcal{C}$, there exists an
+effective \kl{amalgamation system} recognizing $L$, and such that $\mathcal{C}$
+is \kl{effectively closed under rational transductions}. Recall that a
+\intro{rational transduction} is a function $f \colon \Sigma^* \to \Gamma^*$
+that can be defined by a (nondeterministic) finite automaton with output. A
+class of languages $\mathcal{C}$ is \intro{effectively closed under rational
+transductions} when, for every language $L \in \mathcal{C}$, and every rational
+transduction $f \colon \Sigma^* \to \Gamma^*$, the image of $L$ through $f$
+belongs to $\mathcal{C}$ and is effectively computable. We refer the reader to
+\cite{BERST79} for a primer on \kl{rational transductions}.
+
+\AP A class $\mathcal{C}$ of languages is \intro(amalg){strongly effective}
+whenever the emptiness problem for languages in $\mathcal{C}$ is decidable.
+This notion is interesting because usual language problems such as boundedness
+or simultaneous boundedness are decidable for \kl{strongly effective
+amalgamative classes}~\cite{ASZZ24}.
 
@@
     \InfPeriodChain{x}u \cup u\InfPeriodChain{x}$. For any concrete values of
-    the bounds $n$, $m$, and $N_0$, this language is regular. As $\mathcal{C}$
-    is \kl{closed under rational transductions}, we can therefore reduce the
-    inclusion to emptiness of $L \cap \overline{R}$.\todo{explain that $L
-    \mapsto L \cap R$ is a rational  transduction?} However, we need to find
+    the bounds $n$, $m$, and $N_0$, this language is regular. The map $L
+    \mapsto L \cap \Sigma^* \setminus R$  is a \kl{rational transduction}
+    because $\Sigma^* \setminus R$ is regular. Since $\mathcal{C}$ is
+    \kl{closed under rational transductions}, we can therefore reduce the
+    inclusion to emptiness of $L \cap \overline{R}$. However, we need to find
     these bounds first.
 
-    To determine values for $n$ and $m$, we first test if $L$ is bounded. As
-    the algorithm in \cite[Section 4.2]{ASZZ24} is effective, this yields words
-    $w_1, \ldots w_n$ such that $L \subseteq w_1^* \cdots w_n^*$ and therefore
-    yields also the bounds in questions.
+    To determine values for $n$ and $m$, we first test if $L$ is bounded. Since
+    emptiness is decidable, we can apply the algorithm in~\cite[Section
+    4.2]{ASZZ24} to decide if $L$ is bounded. If $L$ is bounded, this algorithm
+    yields words $w_1, \ldots w_n$ such that $L \subseteq w_1^* \cdots w_n^*$
+    and therefore yields also the bounds in questions. If $L$ is not bounded,
+    then $L$ cannot be \kl{well-quasi-ordered} by the \kl{infix relation}
+    because of \cref{infix-amalgamation:thm} and we immediately return false.
 
@@
 
-    \cref{wqo-infix-decidable} $\Rightarrow$ \cref{wqo-prefix-decidable}. We
-    just consider the transduction
+    \cref{wqo-infix-decidable} $\Rightarrow$
+    \cref{wqo-prefix-decidable}. We just consider the transduction $f$
+    that maps every word $w$ to $\# w$ where $\# $ is a fresh symbol. Then, for
+    any language $L \in \mathcal C$, $L$ is \kl{well-quasi-ordered} by
+    \kl{prefix} if and only if $f(L)$ is \kl{well-quasi-ordered} by \kl{infix}.
+    \todo{fresh symbols? can we assume we have those??}
 
     \cref{wqo-prefix-decidable} $\Rightarrow$
-    \cref{emptiness-decidable}. We consider the transduction $\mathcal T
-    = \Sigma^* \times \{a, b\}$.\todo{Define this transduction explicitly} Then
-    for any language $L \in \mathcal C$, $\mathcal TL$ is well-quasi-ordered by
-    infix if and only if $L$ is empty.
+    \cref{emptiness-decidable}. We consider the transduction $f$ that 
+    maps every word $w$ to the word $a b^{|w|} a$.
+    Then, for any language $L \in \mathcal C$, $f(L)$ is \kl{well-quasi-ordered}
+    by \kl{prefix} if and only if $L$ is finite if and only if $L$ is finite.
+    Hence, we can decide if $L$ is finite, and therefore if $L$ is empty 
+    by enumeration.
+    \todo{I don't understand the notation $\Sigma^* \times \set{a,b}^*$, so I did 
+    a worse but correct proof for me.}
 \end{proofof}

### 2024-10-18 src/infixes.tex
new proofs

@@
     \label{thue-morse:lemma}
-    There exists a language $L$ that is \kl{downwards closed} for the \kl{infix
+    The language $\intro*\LMorse$ of \kl{infixes} of the \kl{Thue-Morse sequence}
+    is \kl{downwards closed} for the \kl{infix
     relation}, \kl{well-quasi-ordered} for the \kl{infix relation}, but is not
@@
 \begin{proof}
-    Let $\ThueMorse$ be the \kl{Thue-Morse sequence}
-    $L$ be the set of infixes of $w$. By construction $L$ is an \emph{infinite}
-    \kl{downwards closed} for the \kl{infix relation}. Let us argue that $L$ is
+    By construction $\LMorse$ is an \emph{infinite}
+    \kl{downwards closed} for the \kl{infix relation}. Let us argue that $\LMorse$ is
     \kl{well-quasi-ordered} for the \kl{infix relation}, but is not \kl(language){bounded}.
 
-    Assume by contradiction that $L$ is \kl(language){bounded}. In this case, there exist
+    Assume by contradiction that $\LMorse$ is \kl(language){bounded}. In this case, there exist
     words $w_1, \dots, w_k \in \Sigma^*$ such that $L \subseteq w_1^* \cdots
@@
     \label{thue-morse-ordinal:lemma}
-    Let $L$ be the language of \kl{infixes} of the \kl{Thue-Morse sequence}.
-    Then, the \kl{ordinal invariants} of $L$ are as follows:
-    the \kl{maximal order type} of $L$ is $\omega$,
-    the \kl{ordinal height} of $L$ is $\omega$,
-    the \kl{ordinal width} of $L$ is $\omega$.
+    The \kl{maximal order type} of $\LMorse$ is $\omega$,
+    the \kl{ordinal height} of $\LMorse$ is $\omega$,
+    the \kl{ordinal width} of $\LMorse$ is $\omega$.
 \end{lemma}
@@
     Let us prove that these are upper bounds for the \kl{ordinal invariants} of
-    $L$. For the \kl{ordinal height}, it is true for any language $L$.
+    $L$. For the \kl{ordinal height}, it is true for any language $\LMorse$.
     For the \kl{maximal order type}, we remark that
@@
     Now, let us prove that these bounds are tight. It is clear that
-    $\oHeight{L} = \omega$: given any number $n \in \Nat$, one can construct a
+    $\oHeight{\LMorse} = \omega$: given any number $n \in \Nat$, one can construct a
     \kl{decreasing sequence} of words in $L$ of length $n$, for instance by
@@
     decreasing size.
-    Let us now prove that $\oWidth{L} = \omega$. Assume by contradiction that
-    $\oWidth{L}$ is finite. Then, $L$ can be written as a finite union of
+    Let us now prove that $\oWidth{\LMorse} = \omega$. Assume by contradiction that
+    $\oWidth{\LMorse}$ is finite. Then, $L$ can be written as a finite union of
     \kl{chains} for the \kl{infix relation}, and in particular, $L$ is
@@
     Finally, because the \kl{ordinal width} is at most the \kl{maximal order
-    type}, we conclude that the \kl{maximal order type} of $L$ is also $\omega$.
+    type}, we conclude that the \kl{maximal order type} of $\LMorse$ is also $\omega$.
 \end{proof}

### 2024-10-21 src/amalgam.tex
Rework paper organisation

@@
 of them will be innocuous, an important consequence is that we have to consider
-\emph{classes of languages} rather than individual ones. Such classes will be
-called \kl{effective amalgamative classes} (\kcref{effective amalgamative
-classes}). In the following theorem, we prove that under such assumptions,
-testing \kl{well-quasi-ordering} is inter-reducible to testing whether a
-language of the class is empty.
+\emph{classes of languages} rather than individual ones, for instance: the
+class of all regular language, or the class of all context-free languages. Such
+classes will be called \kl{effective amalgamative classes} (\kcref{effective
+amalgamative classes}). In the following theorem, we prove that under such
+assumptions, testing \kl{well-quasi-ordering} is inter-reducible to testing
+whether a language of the class is empty.
 
-\begin{theorem}[restated=infix-wqo-is-emptiness:thm,label=infix-wqo-is-emptiness:thm]
+\begin{theorem}[restate=infix-wqo-is-emptiness:thm,label=infix-wqo-is-emptiness:thm]
 % 	\label{infix-wqo-is-emptiness:thm}
@@
 finite automata, there is a natural ordering between runs, i.e., a run is
-smaller than another one if one can ``unroll loops" of the first run to obtain
+smaller than another one if one can ``unroll loops'' of the first run to obtain
 the second one. Typical pumping arguments then rely on the fact that
@@
 Intuitively, the definition of an amalgamation system allows the comparison of
-runs, and the proper ``gluing" of runs together to obtain new runs. Let us
+runs, and the proper ``gluing'' of runs together to obtain new runs. Let us
 recall some examples of languages that can be recognized by \kl{amalgamation
-systems} given by the authors of the original paper: regular languages
-\cite[Theorem 5.3]{ASZZ24}, VASS as a consequence of \cite[Theorem
-5.5]{ASZZ24}, context-free languages as a consequence of \cite[Theorem
-5.10]{ASZZ24}. For this paper to be self-contained, we will also recall how
-runs of a finite state automaton can be understood as an \kl{amalgamation
-system}.
+systems}: regular languages \cite[Theorem 5.3]{ASZZ24}, VASS as a consequence
+of \cite[Theorem 5.5]{ASZZ24}, context-free languages as a consequence of
+\cite[Theorem 5.10]{ASZZ24}. For this paper to be self-contained, we will also
+recall how runs of a finite state automaton can be understood as an
+\kl{amalgamation system}.
 
@@
     or equal than $u$ and $v$. This proves  that $u$ and $v$ are comparable for
-    the prefix, infix, and suffix relations. In particular, we have can
+    the prefix, infix, and suffix relations. In particular, we
     conclude that $\GapLanguage{r}{i}$ is a \kl{bounded language}, because of
@@
 
-\AP We say that a class $\mathcal{C}$ of languages is an \kl{effective
+\AP We say that a class $\mathcal{C}$ of languages is an \intro{effective
 amalgamative class} whenever for every $L \in \mathcal{C}$, there exists an
@@
 is \kl{effectively closed under rational transductions}. Recall that a
-\intro{rational transduction} is a function $f \colon \Sigma^* \to \Gamma^*$
-that can be defined by a (nondeterministic) finite automaton with output. A
-class of languages $\mathcal{C}$ is \intro{effectively closed under rational
-transductions} when, for every language $L \in \mathcal{C}$, and every rational
-transduction $f \colon \Sigma^* \to \Gamma^*$, the image of $L$ through $f$
-belongs to $\mathcal{C}$ and is effectively computable. We refer the reader to
-\cite{BERST79} for a primer on \kl{rational transductions}.
+\intro{rational transduction} is a \emph{realtion} $R \subseteq \Sigma^* \times
+\Gamma^*$ that can be defined by a (nondeterministic) finite automaton with
+output \cite[Chapter 5, page 64]{BERST79}. A class of languages $\mathcal{C}$
+is \intro{effectively closed under rational transductions} when, for every
+language $L \in \mathcal{C}$, and every rational transduction $R \subseteq
+\Sigma^* \times \Gamma^*$, the image of $L$ through $R$ --- that is,
+ $\setof{v \in \Gamma^*}{\exists u \in L. (u,v) \in R}$ --- is in $\mathcal{C}$
+and effectively computable.
 
@@
 
-    Let $R = \bigcup_{x,y \in \Sigma^{\leq n}} \bigcup_{u \in \Sigma^{\leq m
-    \times N_0}} \InfPeriodChain{x} u \InfPeriodChain{y} \cup
+    Let $R(n,m,N_0) \defined \bigcup_{x,y \in \Sigma^{\leq n}} \bigcup_{u \in
+    \Sigma^{\leq m \times N_0}} \InfPeriodChain{x} u \InfPeriodChain{y} \cup
     \InfPeriodChain{x}u \cup u\InfPeriodChain{x}$. For any concrete values of
     the bounds $n$, $m$, and $N_0$, this language is regular. The map $L
-    \mapsto L \cap \Sigma^* \setminus R$  is a \kl{rational transduction}
-    because $\Sigma^* \setminus R$ is regular. Since $\mathcal{C}$ is
-    \kl{closed under rational transductions}, we can therefore reduce the
-    inclusion to emptiness of $L \cap \overline{R}$. However, we need to find
-    these bounds first.
+    \mapsto L \cap \Sigma^* \setminus R(n,m,N_0)$  is a \kl{rational
+    transduction} because $\Sigma^* \setminus R(n,m,N_0)$ is regular. Since
+    $\mathcal{C}$ is \kl{closed under rational transductions}, we can therefore
+    reduce the inclusion to emptiness of this language. However, we need to
+    find these bounds first.
 
-    To determine values for $n$ and $m$, we first test if $L$ is bounded. Since
-    emptiness is decidable, we can apply the algorithm in~\cite[Section
-    4.2]{ASZZ24} to decide if $L$ is bounded. If $L$ is bounded, this algorithm
+    To determine values for $n$ and $m$, we first test if $L$ is
+    \kl(language){bounded}. Since emptiness is decidable, we can apply the
+    algorithm in~\cite[Section 4.2]{ASZZ24} to decide if $L$ is
+    \kl(language){bounded}. If $L$ is \kl(language){bounded}, this algorithm
     yields words $w_1, \ldots w_n$ such that $L \subseteq w_1^* \cdots w_n^*$
-    and therefore yields also the bounds in questions. If $L$ is not bounded,
-    then $L$ cannot be \kl{well-quasi-ordered} by the \kl{infix relation}
-    because of \cref{infix-amalgamation:thm} and we immediately return false.
+    and therefore yields also the bounds in questions: $n$ is the number of
+    words, and $m$ is the maximal length of a word $w_i$ where $1 \leq i \leq
+    n$. If $L$ is not bounded, then $L$ cannot be \kl{well-quasi-ordered} by
+    the \kl{infix relation} because of \cref{infix-amalgamation:thm} and we
+    immediately return false.
 
     To determine the value for $N_0$, we first computer maximal subsets $S
-    \subseteq [n]$ such that $w_i$ for $i \in S$ are simultaneously unbounded.\todo{define this}
-    We do this by applying a transduction mapping $w_i$ to some fresh letter
-    $a_i$ if $i \in S$ and $\varepsilon$ otherwise, and testing for
-    simultaneous unboundedness \cite[Section 4.1]{ASZZ24}. Given a candidate
-    bound $n_0$ for $N_0$, we construct the languages $L_{S,n_0} =
-    w_1^{\circ_1}\cdots w_n^{\circ_n}$ where $\circ_i = *$ if $i \in S$ and
-    $\circ_i = \leq n_0$ otherwise. We then test if $L \subseteq \bigcup_{S
-    \text{ maximal}} L_{S,n_0}$. If yes, $n_0$ is our value for $N_0$,
-    otherwise, we increase it and repeat the construction. As we chose maximal
-    sets $S$, this procedure will eventually terminate.
+    \subseteq [n]$ such that $w_i$ for $i \in S$ are simultaneously
+    unbounded.\todo{define this} We do this by applying a transduction mapping
+    $w_i$ to some fresh letter $a_i$ if $i \in S$ and $\varepsilon$ otherwise,
+    and testing for simultaneous unboundedness \cite[Section 4.1]{ASZZ24}.
+    Given a candidate bound $n_0$ for $N_0$, we construct the languages
+    $L_{S,n_0} = w_1^{\circ_1}\cdots w_n^{\circ_n}$ where $\circ_i = *$ if $i
+    \in S$ and $\circ_i = \leq n_0$ otherwise. We then test if $L \subseteq
+    \bigcup_{S \text{ maximal}} L_{S,n_0}$. If yes, $n_0$ is our value for
+    $N_0$, otherwise, we increase it and repeat the construction. As we chose
+    maximal sets $S$, this procedure will eventually terminate.
 
@@
     \cref{wqo-prefix-decidable} $\Rightarrow$
-    \cref{emptiness-decidable}. We consider the transduction $f$ that 
-    maps every word $w$ to the word $a b^{|w|} a$.
-    Then, for any language $L \in \mathcal C$, $f(L)$ is \kl{well-quasi-ordered}
-    by \kl{prefix} if and only if $L$ is finite if and only if $L$ is finite.
-    Hence, we can decide if $L$ is finite, and therefore if $L$ is empty 
-    by enumeration.
-    \todo{I don't understand the notation $\Sigma^* \times \set{a,b}^*$, so I did 
-    a worse but correct proof for me.}
+    \cref{emptiness-decidable}. 
+	We 
+	consider the transduction $R \defined \Sigma^* \times \set{a, b}^*$. Then 
+	for any language $L \in \mathcal C$,
+    the image of $L$ through $R$ is \kl{well-quasi-ordered}
+    by \kl{infix} if and only if $L$ is empty.
 \end{proofof}
+
+The class $\mathcal{C}_\text{aut}$ of \kl{regular languages} and the class
+$\mathcal{C}_{\text{cfg}}$ of context-free languages are examples of
+\kl{effective amalgamative classes}, hence the following corollary.
+
+\begin{corollary}
+    \label{aut-cfg-infix:cor}
+    Let $\mathcal{C} \in \set{ \mathcal{C}_\text{aut}, \mathcal{C}_{\text{cfg}}}$.
+    It is decidable whether a language in $\mathcal{C}$ is \kl{well-quasi-ordered}
+    by the \kl{infix relation}.
+    Furthermore, whenever it is \kl{well-quasi-ordered} by the \kl{infix relation},
+    it is a \kl{bounded language}.
+\end{corollary}
+
+Let us conclude by noting that it is unsurprisingly not possible to decide
+whether a decidable language is \kl{well-quasi-ordered} by the \kl{prefix
+relation}. This is a very easy result whose sole purpose is to contrast with
+the decidability result of \cref{aut-cfg-infix:cor}.
+
+\begin{remark}
+    The following problem is undecidable: given a language $L$
+    decided by a Turing machine, answer whether 
+    $L$ is \kl{well-quasi-ordered} for the \kl{prefix relation}.
+\end{remark}
+\begin{proof}
+    We reduce the halting problem on the empty string $\varepsilon$.
+    Let $M$ be a Turing Machine, we write the languages $L$ of finite runs
+    of $M$ starting on the empty string,
+    that we surround by special markers. This language is decidable,
+    and 
+    is
+    \kl{well-quasi-ordered} if and only if it is finite
+    if and only if $M$ terminates on $\varepsilon$.
+\end{proof}

### 2024-10-21 src/infixes.tex
Rework paper organisation

@@
 y$ and $y \preceq x$ implies $x = y$. 
-The embedding defined in \cref{infix-embedding:thm} is illustrated
+A simplified version of the embedding defined in \cref{infix-embedding:thm} is illustrated
 for the \kl{subword relation} in \cref{infix-embedding:fig}.
@@
         inside the \kl{infix relation} for $\set{a,b,\#}^*$
-        using the encoding of \cref{infix-embedding:thm}, restricted to words
-        of length at most $3$. To obtain smaller words,
-        we replaced $D_i$ by $\max D_i$ in this construction.
+        using a simplified version of \cref{infix-embedding:thm}, restricted to words
+        of length at most $3$. 
     }
@@
 the \kl{infix relation}. Note that in general, the \kl{downwards closure} of a
-chain is \emph{not} a chain. However, for the chains generated using periodic
+chain is \emph{not} a chain (see \cref{dw-closure-not-wqo:rem}). However, for the chains generated using periodic
 words, the \kl{downwards closure} $\dwset[\infleq]{\setof{x^p}{p \in \Nat}}$ is
@@
 $\Nat$.
+
+\begin{remark}
+    \label{dw-closure-not-wqo:rem}
+    Let $(X,\preceq)$ be a quasi-ordered set, and $L \subseteq X$ be such that $(L,
+    \preceq)$ is \kl{well-quasi-ordered}. It is not true in general that
+    $(\dwset{L}, \preceq)$ is \kl{well-quasi-ordered}. In the case of $(\Sigma^*,
+    \infleq)$ a typical example is to start from an infinite \kl{antichain} $A$,
+    together with an enumeration $\seqof{w_i}$ of $A$, and build the language $L
+    \defined \setof{ \prod_{i = 0}^n w_i }{ i \in \Nat }$. By definition, $L$ is a
+    \kl{chain} for the \kl{infix} ordering, hence \kl{well-quasi-ordered}. However,
+    $\dwset[\infleq]{L}$ contains $A$, and is therefore not
+    \kl{well-quasi-ordered}. 
+\end{remark}
 
@@
 the core result that powers the analysis done in the upcoming
-\cref{infix-finite-automata:thm,infix-amalgamation:thm}.
+\cref{bounded-language:thm,infix-amalgamation:thm}.
 

### 2024-10-21 src/introduction.tex
Rework paper organisation

@@
         representing the successor relation of the order.
-        Furthermore, we highlight in red relations that are added
+        Furthermore, we highlight in dashed red relations that are added
         when moving from the prefix relation to the infix one,
@@
 straightforwardly generalize to infinite quasi-ordered alphabets.
-Unfortunately, it is easy to see that all of these constructions are not
+Unfortunately, it is easy to see that none of these constructions are 
 well-quasi-ordered as soon as the alphabet contains two distinct letters:
@@
 given a language $L \subseteq \Sigma^*$, decide whether $L$ is
-\kl{well-quasi-ordered} by the \kl{suffix relation}, or the \kl{infix
-relation}.
+\kl{well-quasi-ordered} by the \kl{infix relation}.
 
@@
 Note that \kl{chains} are the simplest possible \kl{well-quasi-ordered} sets
-(they are totally ordered and well-founded) and it is known that finite unions
+(they are \kl{totally ordered} and \kl{well-founded}) and it is known that finite unions
 of \kl{well-quasi-ordered} sets are \kl{well-quasi-ordered}. As a consequence,
@@
 relation}. Let us furthermore highlight that this characterization holds
-without any restriction on the decidability of the language $L$ itself, but
-heavily relies on the assumption that $\Sigma$ is finite. This allows us to
+without any restriction on the decidability of the language $L$ itself.\footnote{But
+heavily relies on the assumption that $\Sigma$ is finite} This allows us to
 derive the following bounds on the \kl{ordinal invariants} of such
@@
 effective restrictions on the computational model that we call an \kl{effective
-amalgamative class}, the formal statement of which we defer to
+amalgamative class}, the formal definition of which we defer to
 \cref{infixes-amalgamation-effective:subsec}.
@@
 Let us point out that the above result implies that the hypothesis of
-\kl{bounded languages} on the theoretical side is not a restriction at all in
-practice. As a practical and more easily understandable 
+\kl{bounded languages} on the theoretical side is not a restriction in
+practice. As a down-to-earth and more easily understandable 
 consequence,
@@
 and to the class $\mathcal{C}_\text{cfg}$ of all languages recognized by context-free grammars,
-which are both \kl{effective amalgamative classes}.
+which are both \kl{effective amalgamative classes} \cite{aut-cfg-infix:cor}.
 

### 2024-10-21 src/metadata.tex
Rework paper organisation

@@
 \author{%
-        Roland Guttenberg\thanks{Technical University of Munich,
-Germany}
-     \and
         Nathan Lhote\thanks{Aix-Marseille University}
@@
 
-\date{2024-10-14 10:28:01
-+0200 -- 9f829fa27689f65f659ce09cd9273d1f894838fa (main) at git@github.com:AliaumeL/word-ordering-wqo.git}
+\date{2024-10-18 15:05:24
++0200 -- a05cd36b54b8b7b587b7b2a6ab7c1cde3f66a997 (new-intro) at git@github.com:AliaumeL/word-ordering-wqo.git}
 

### 2024-10-21 src/prefixes.tex
Rework paper organisation

@@
 leverage the decidability of $\MSO$ over infinite binary trees \cite[Theorem
-1.1]{RAB69}. This result will follow from our general decidability result in
-\cref{amalgamation:sec} but serves as a good illustration.
+1.1]{RAB69}. This result will follow from our general decidability result
+(\cref{infix-wqo-is-emptiness:thm}) but is worth stating on its own for its
+simplicity.
 
@@
 Let us now go further and fully characterize languages $L$ such that the
-prefix relation is well-quasi-ordered, even without any restriction on the
+prefix relation is well-quasi-ordered, without any restriction on the
 decidability of $L$ itself. Let us remark that finite unions of \kl{chains} are
@@
 \end{corollary}
-
-Let us conclude by noting that it is unsurprisingly not possible to decide
-whether a decidable language is \kl{well-quasi-ordered} by the \kl{prefix
-relation}. This is a very easy result whose sole purpose is to contrast with
-the decidability result of \cref{prefix-wqo-reg-decidable:cor}.
-
-\begin{lemma}
-    The following problem is undecidable: given a language $L$
-    decided by a Turing machine, answer whether 
-    $L$ is \kl{well-quasi-ordered} for the \kl{prefix relation}.
-\end{lemma}
-\begin{proof}
-    We reduce the halting problem on the empty string $\varepsilon$.
-    Let $M$ be a Turing Machine, we write the languages $L$ of finite runs
-    of $M$ starting on the empty string,
-    that we surround by special markers. This language is decidable,
-    and 
-    is
-    \kl{well-quasi-ordered} if and only if it is finite
-    if and only if $M$ terminates on $\varepsilon$.
-\end{proof}

### 2024-10-21 src/prelims.tex
Rework paper organisation

@@
 
-\paragraph*{Well-quasi-orders.} Let us introduce some notations for
+\paragraph*{Orderings and Well-Quasi-Orderings.} A \intro{quasi-order} is a
+reflexive and transitive binary relation, it is a \intro{partial order} if it
+is furthermore antisymmetric. A \intro{total order} is a \kl{partial order}
+where any two elements are comparable. Let now us introduce some notations for
 \kl{well-quasi-orders}. A sequence $\seqof{x_i}$ in a set $X$ is
@@
 
-Let us point out that the \kl{prefix relation} (resp. the \kl{suffix relation}
-and the \kl{infix relation}) on $\Sigma^*$ are always \kl{well-founded}, i.e.,
-there are no infinite \kl{decreasing sequences} for this ordering. In
-particular, for a language $L \subseteq \Sigma^*$ to be
-\kl{well-quasi-ordered}, it suffices to prove that it contains no infinite
-\kl{antichain}.
+The \kl{prefix relation} (resp. the \kl{suffix relation} and the \kl{infix
+relation}) on $\Sigma^*$ are always \kl{well-founded}, i.e., there are no
+infinite \kl{decreasing sequences} for this ordering. In particular, for a
+language $L \subseteq \Sigma^*$ to be \kl{well-quasi-ordered}, it suffices to
+prove that it contains no infinite \kl{antichain}.
 
-\paragraph*{Ordinal Invariants.} 
-\label{ordinal-invariants:subsec} An ordinal is a well-founded ordered set. We
-use $\alpha, \beta, \gamma$ to denote ordinals, and use $\omega$ to denote the
-first infinite ordinal, i.e., the set of natural numbers with the usual
-ordering. We assume superficial familiarity with ordinal arithmetic, and refer
-to \cite{KUNEN80} for a detailed introduction to this domain.
-Given a tree $T$ whose branches are all finite we can define
-an ordinal $\alpha_T$ inductively as follows: if $T$ is a leaf then $\alpha_T =
-0$, if $T$ has children $\seqof{T_i}$ then $\alpha_T = \sup \setof{\alpha_{T_i}
-+ 1}{i \in \Nat}$. We say that $\alpha_T$ is the \emph{rank} of $T$.
+\paragraph*{Ordinal Invariants.} \label{ordinal-invariants:subsec}
+An \intro{ordinal} is a \kl{well-founded} \kl{totally ordered}
+set. We use $\alpha, \beta, \gamma$ to denote ordinals, and use $\omega$ to
+denote the first infinite ordinal, i.e., the set of natural numbers with the
+usual ordering. We assume superficial familiarity with ordinal arithmetic, and
+refer to the books of Kunen \cite{KUNEN80} and Krivine~\cite[Chapter
+II]{KRIVINE71} for a detailed introduction to this domain. Given a tree $T$
+whose branches are all finite we can define an ordinal $\alpha_T$ inductively
+as follows: if $T$ is a leaf then $\alpha_T = 0$, if $T$ has children
+$\seqof{T_i}$ then $\alpha_T = \sup \setof{\alpha_{T_i} + 1}{i \in \Nat}$. We
+say that $\alpha_T$ is the \emph{rank} of $T$.
 

### 2024-10-21 src/conclusion.tex
Big Mic Mac (#6)
- new introduction 
- removing automata based approach
- add the lexicographic ordering in conclusion 
- splitting `infixes` into `bounded` and `amalgamation systems`
- new bibliographic entries + preliminaries
- add the counter-example using the Thue-Morse sequence
- update figures to add dashes
- remove Roland from co-authors
- small knowledge typos correction

--- 2024-08-12 src/conclusion.tex
@@
 the minimal automaton recognizing the language.
+
+\subparagraph{Lexicographic orderings} There is another natural ordering on
+words, the \emph{lexicographic ordering}, which does not fit well in our
+current framework because it is always of \kl{ordinal width} $0$. However, the
+order-type of the lexicographic ordering over \kl{regular languages} has
+already been investigated in the context of infinite words \cite{CACOPU18}, and
+it would be interesting to see if one can extend these results to decide
+whether such an ordering is \kl{well-founded} for languages recognized by
+\kl{amalgamation systems}.

### 2024-10-21 src/infixes.tex
Big Mic Mac (#6)
- new introduction 
- removing automata based approach
- add the lexicographic ordering in conclusion 
- splitting `infixes` into `bounded` and `amalgamation systems`
- new bibliographic entries + preliminaries
- add the counter-example using the Thue-Morse sequence
- update figures to add dashes
- remove Roland from co-authors
- small knowledge typos correction

@@
 \preceq' f(y)$. When such an embedding exists, we say that $X$ \reintro{embeds
-into} $Y$. We say that a quasi-ordered set $(X, \preceq)$ is a \intro{partial
+into} $Y$. Recall that a quasi-ordered set $(X, \preceq)$ is a \kl{partial
 ordering} whenever the relation $\preceq$ is antisymmetric, that is $x \preceq

### 2024-10-21 src/prelims.tex
Big Mic Mac (#6)
- new introduction 
- removing automata based approach
- add the lexicographic ordering in conclusion 
- splitting `infixes` into `bounded` and `amalgamation systems`
- new bibliographic entries + preliminaries
- add the counter-example using the Thue-Morse sequence
- update figures to add dashes
- remove Roland from co-authors
- small knowledge typos correction

@@
 The \kl{prefix relation} (resp. the \kl{suffix relation} and the \kl{infix
-relation}) on $\Sigma^*$ are always \kl{well-founded}, i.e., there are no
+relation}) on $\Sigma^*$ are always \intro{well-founded}, i.e., there are no
 infinite \kl{decreasing sequences} for this ordering. In particular, for a
@@
 $X$ written $\intro*{\oHeight{X}}$ \cite{schmidt81}, and the \intro{ordinal
-width} of $X$ written $\intro*{\oWidth{X}}$ \cite{kriz90b}. We refer to the
-survey of \cite{DZSCSC20} for a detail discussion on these concepts and their
-computation on specific classes of well-quasi-ordered sets.
+width} of $X$ written $\intro*{\oWidth{X}}$ \cite{kriz90b}. These three
+parameters are called the \intro{ordinal invariants} of a
+\kl{well-quasi-ordered} set $X$. We refer to the survey of \cite{DZSCSC20} for
+a detail discussion on these concepts and their computation on specific classes
+of well-quasi-ordered sets.

### 2024-10-21 src/word-embeddings-figure.py

@@ -221,7 +221,7 @@ def render_figure_tikz(nodes, edges, compare_with=None, standalone=True, infix_e
         isInfixEnc/.style={},
         isRealWord/.style={A4,opacity=0.2},
         isRealEmbedding/.style={A4,opacity=0.2},
-        isNewEdge/.style={B2},
+        isNewEdge/.style={B2,dashed},
         ]
 
         """)

### 2024-10-21 src/prefixes.tex
correct knowledge usage

@@
 that \kl{antichain branches} can be described inside the \kl{tree of prefixes}
-by a monadic second order formula (\kl{$\MSO$-formula}), allowing us to
+by a monadic second order formula ($\MSO$-formula), allowing us to
 leverage the decidability of $\MSO$ over infinite binary trees \cite[Theorem
@@
 \begin{proof}
-    If $L$ is regular, then it is \kl{$\MSO$-definable}, and there 
-    exists a formula $\varphi(x)$ in \kl{$\MSO$} that selects nodes 
+    If $L$ is regular, then it is $\MSO$-definable, and there 
+    exists a formula $\varphi(x)$ in $\MSO$ that selects nodes 
     of the \kl{tree of prefixes} that belong to $L$. Now, to decide whether there
@@
 single element $w$, omitting the ordering relation when it is clear from the
-context.
+context. A set $S$ is called \intro{downwards closed} if $\dwset{S} = S$.
 

### 2024-10-29 src/amalgam.tex
Amalgamation systems: Fix definitions and simplify proofs (#8)
Co-authored-by: lschuetze <lschuetze@mpi-sws.org>

@@
 arguments based on so-called \emph{minimal runs} exist: they are called
-\emph{amalgamation systems} and were recently introduced by Anad, Schmitz,
+\emph{amalgamation systems} and were recently introduced by Anand, Schmitz,
 Schütze, and Zetzsche \cite{ASZZ24}.
@@
 \cref{pumping-periods:lem} talks about. At the core of a pumping argument,
-there is a notion of \emph{run}, which could for instance be a sequence of
+%there is a notion of a \emph{run}, which could for instance be a sequence of
 transitions taken in a finite state automaton. Continuing on the analogy with
 finite automata, there is a natural ordering between runs, i.e., a run is
-smaller than another one if one can ``unroll loops'' of the first run to obtain
-the second one. Typical pumping arguments then rely on the fact that
+smaller than another one if one can ``delete'' loops of the larger run to obtain
+the other. Typical pumping arguments then rely on the fact that
 \emph{minimal} runs are of finite size, and that all other runs are
-obtained by ``gluing'' minimal runs together. This is exactly what
-\kl{amalgamation systems} are about.
+obtained by ``gluing'' loops to minimal runs. Generalizing this notion yields the 
+notion of \kl{amalgamation systems}.
 
@@
 \cdots \GapWord{f}{k-1} u_k \GapWord{f}{k}$ where $\GapWord{f}{i} =
-v[f(i)+1:f(i+1)-1]$ for all $1 \leq i \leq k-1$, $\GapWord{f}{k} =
-v[f(k)+1:\card{v}]$, and $\GapWord{f}{0}   = v[1: f(1)-1]$. We say that
+v_{f(i)+1} \cdots v_{f(i+1)-1}$ for all $1 \leq i \leq k-1$, $\GapWord{f}{k} =
+v_{[f(k)+1} \cdots v_{\card v}$, and $\GapWord{f}{0}   = v_1 \cdots v_{f(1)-1}$. We say that
 $\intro*\GapWord{f}{i}$ is the $i$-th \intro{gap word} of $f$. We encourage the
@@
     An \intro{amalgamation system}
-    is a triple $(\Sigma, R, E, \canrun)$ where
+    is a tuple $(\Sigma, R, \canrun, E)$ where
     $\Sigma$ is a finite alphabet,
     $R$ is a set of so-called \emph{runs},
-    and 
-    $E$ is a set of so-called \emph{run embeddings},
-    and $\canrun \colon R \to (\Sigma \uplus \set{\varepsilon})^*$ is a 
-    function computing a \emph{canonical decomposition} of a run.
+    $\canrun \colon R \to (\Sigma \uplus \set{\cansep})^*$ is a 
+    function computing a \emph{canonical decomposition} of a run,
+    and $E$ describes the so-called \emph{admissible embeddings} between runs: If $\rho$ and $\sigma$ are runs from $R$, then $E(\rho, \sigma)$ is a subset of the subword embeddings between $\canrun[\rho]$ and $\canrun[\sigma]$.
+We write $\rho \intro*\runleq \sigma$ if $E(\rho, \sigma)$ is non-empty.
+If we want to refer to a specific embedding $f \in E(\rho, \sigma)$, we also write $\rho \runleq_f \sigma$.
     Given a run $r \in R$, and $i \in \range[0]{\card{\canrun(r)}}$, 
-    the \intro{gap language} of $r$ at position $i$ is $\GapLanguage{r}{i} \defined
+    the \intro{gap language} of $r$ at position $i$ is $\intro*\GapLanguage{r}{i} \defined
     \setof{\GapWord{f}{i}}{\exists s \in R. \exists f \in E(r,s) }$.
@@
     \begin{enumerate}
-        \item \emph{Word Embeddings as Morphisms.} For all $r, s \in R$,
-            $E(r,s) \subseteq \HigEmb(\canrun(r), \canrun(s))$.
         \item \emph{$(R, E)$ Forms a Category.}
-            For all $r,s,t \in R$,
-            $\mathrm{id} \in E(r,r)$,
-            and whenever $f \in E(r,s)$ and $g \in E(s,t)$,
-            then $g \circ f \in E(r,t)$.
+            For all $\rho, \sigma, \tau \in R$,
+            $\mathrm{id} \in E(\rho,\rho)$,
+            and whenever $f \in E(\rho,\sigma)$ and $g \in E(\sigma,\tau)$,
+            then $g \circ f \in E(\rho,\tau)$.
         \item \emph{Well-Quasi-Ordered System.}
-            $(R, \leq_E)$ is a well-quasi-ordered set,
-            where $r \leq_E s$ if and only if $E(r,s) \neq \emptyset$.
-        \item \emph{Controlled Amalgamation.}
-            For all $r, s, t \in R$,
-            for all $f \in E(r,s)$,
-            for all $g \in E(r,t)$,
-            and for all $0 \leq i_0 \leq \card{\canrun(r)}$,
-            there exists a run $z \in R$ and morphisms
-            $h \in E(s,z)$ and $k \in E(t,z)$
-            satisfying
-            $h \circ f = k \circ g$,
-            $\GapWord{h \circ f}{i} = \GapWord{f}{i} \GapWord{g}{i}$
-            or 
-            $\GapWord{h \circ f}{i} = \GapWord{g}{i} \GapWord{f}{i}$
-            for every $0 \leq j \leq n$,
-            and
-            $\GapWord{h \circ f}{i_0} = \GapWord{f}{i_0} \GapWord{g}{i_0}$.
+            $(R, \runleq)$ is a well-quasi-ordered set.
+        \item \emph{Concatenative Amalgamation.}
+            Let $\rho_0, \rho_1, \rho_2$ be runs 
+            with $\rho_0 \runleq_f \rho_1$ 
+            and $\rho_0 \runleq_g \rho_2$.
+            Then for all $0 \leq i \leq \card{\canrun(\rho_0)}$,
+            there exists a run $\rho_3 \in R$ 
+            and embeddings $\rho_1 \runleq_{g'} \rho_3$
+            and $\rho_2 \runleq_{f'} \rho_3$ 
+            satisfying two conditions:
+            (a) $g' \circ f = f' \circ g$ (we write $h$ for this composition) and
+            (b) for every $0 \leq j \leq \card{\canrun[\rho_0]}$, 
+            the gap word $\GapWord{h}{j}$
+            is either $\GapWord{f}{j} \GapWord{g}{j}$
+            or $\GapWord{h}{j} = \GapWord{g}{j} \GapWord{f}{j}$. 
+            Specifically, for $i$ we may fix $G_{h}{i} = \GapWord{f}{j} \GapWord{g}{j}$.
             We refer to \cref{amalgamation-runs:fig} for an illustration 
@@
 
-    The \intro{amalgamation language} of such a system
-    is the set of all words $w \in \Sigma^*$ such that
-    there exists a run $r \in R$
-    such that the concatenation of the letters of its
-    canonical decomposition, written $\intro*\yieldrun(r)$,
-    equals $w$.
+	The \emph{yield} of a run is obtained by projecting away the separator symbol \cansep~from the canonical decomposition, i.e.
+$\intro*s\yieldrun(\rho) = \project_\Sigma( \canrun[\rho])$.
+The language recognized by an \kl{amalgamation system} is $\yieldrun(R)$.
+
+    We say a language $L$ is an \intro*{amalgamation language} if there exists an \kl{amalgamation system} recognizing it. 
 \end{definition}
@@
 Intuitively, the definition of an amalgamation system allows the comparison of
-runs, and the proper ``gluing'' of runs together to obtain new runs. Let us
-recall some examples of languages that can be recognized by \kl{amalgamation
-systems}: regular languages \cite[Theorem 5.3]{ASZZ24}, VASS as a consequence
-of \cite[Theorem 5.5]{ASZZ24}, context-free languages as a consequence of
-\cite[Theorem 5.10]{ASZZ24}. For this paper to be self-contained, we will also
+runs, and the proper ``gluing'' of runs together to obtain new runs. A number of well-known language classes can be seen to be recognized by \kl{amalgamation
+systems}, e.g., regular languages \cite[Theorem 5.3]{ASZZ24}, reachability and coverability languages of VASS \cite[Theorem 5.5]{ASZZ24}, and context-free languages \cite[Theorem 5.10]{ASZZ24}.
+For this paper to be self-contained, we will also
 recall how runs of a finite state automaton can be understood as an
@@
     letter is the start state of the following one.
-    The canonical decomposition $\canrun \colon R \to \Sigma^*$
+    The canonical decomposition $\canrun$
     is defined as a morphism from $\Delta^*$ to $\Sigma^*$
     that maps $(q,a,p)$ to $a$. 
-    Finally, given two runs $r$ and $s$ of the automaton,
-    we say that an embedding $f \in \HigEmb(\canrun(r), \canrun(s))$
-    belongs to $E(r,s)$ when
-    $f$ is also defining an embedding from $r$ to $s$ as words in $\Delta^*$,
-    where $\Delta$ is equipped with the equality relation.
+    Because of the one-to-one correspondence of steps of a run $\rho$ and letters in its \kl{canonical decomposition}, 
+    we may treat the two interchangeably.
+    Finally, given two runs $\rho$ and $\sigma$ of the automaton,
+    we say that an embedding $f \in \HigEmb(\canrun(\rho), \canrun(\sigma))$
+    belongs to $E(\rho,\sigma)$ when
+    $f$ is also defining an embedding from $\rho$ to $\sigma$ as words in $\Delta^*$.
 
@@
 \begin{proof}
-    By definition, the embeddings inside $E(r,s)$ are defined as elements
-    of $\HigEmb(\canrun(r), \canrun(s))$, and they compose properly.
+    By definition, the embeddings inside $E(\rho,\sigma)$ are defined as elements
+    of $\HigEmb(\canrun(\rho), \canrun(\sigma))$, and they compose properly.
     Because $\Delta = Q \times \Sigma \times Q$ is finite, it is 
@@
     Let us now move to proving that the system satisfies the amalgamation
-    property. Given three runs $r,s,t \in R$, and two embeddings $f \in E(r,s)$
-    and $g \in E(r,t)$, we want to construct an amalgamated run $s \vee t$.
-    Because letters in the run $r$ respect the transitions of the automaton
+    property. Given three runs $\rho,\sigma,\tau \in R$, and two embeddings $f \in E(\rho,\sigma)$
+    and $g \in E(\rho,\tau)$, we want to construct an amalgamated run $\sigma \vee \tau$.
+    Because letters in the run $\rho$ respect the transitions of the automaton
     (i.e., if the letter $i$ ends in state $q$, then the letter $i+1$ starts in
@@
     the order of concatenation in the amalgamation for \emph{all} 
-    $0 \leq i \leq \card{\canrun(r)}$ and not just for one separately.
+    $0 \leq i \leq \card{\canrun(\rho)}$ and not just for one separately.
 
-    Let us now remark that 
+    We conclude by remarking that 
     the language of this amalgamation system is
-    the set of $\yieldrun(r)$ when $r$ ranges over $R$,
-    and because $R$ is the set of valid runs of the automaton,
-    and $\yieldrun(r)$ is the word read along this run,
-    we immediately conclude.
+    the set of $\yieldrun(R)$, 
+    because $R$ is the set of valid runs of the automaton,
+    and $\yieldrun(\rho)$ is the word read along a run $\rho$.
 \end{proof}
 
-\subsection{Amalgamation Systems, WQOs, and Bounded Languages}
+\subsection{$\infleq$-Well-Quasi-Ordered Amalgamation Systems}
 
-Let us now introduce a combinatorial lemma that explains how \kl{gap languages}
-can be pumped. Note that a single \kl{gap language} is always stable under
-concatenation, and our concern will be on a potential description of
-\emph{simultaneously} pumping different \kl{gap languages} to produce valid
-runs.
+We can now show a simple lemma that illuminates much of the structure of amalgamation systems whose language is well-quasi-ordered by $\infleq$.
 
 \begin{lemma}
-    \label{pumping-gap-languages:lem}
-    Let $(\Sigma, R, E, \canrun)$ be an \kl{amalgamation system}, let $r \in R$
-    be a run of size $n \in \Nat$, and let $s, t \in R$ be two runs
-    with embeddings $f \in E(r,s)$ and $g \in E(r,t)$.
-    Let $\seqof{u_i}[0 \leq i \leq n]$ be the sequence of \kl{gap words}
-    for $f$, that is, $u_i =
-    \GapWord{f}{i}$ for all $0 \leq i \leq n$. Similarly, let
-    $\seqof{v_i}[0 \leq i \leq n]$ be the sequence of \kl{gap words} for $g$.
+	\label{gap-words-prefix-ordered:lem}
+	Let $L$ by an \kl{amalgamation language} recognized by $(\Sigma, R, E, \canrun)$ that is well-quasi-ordered by $\infleq$.
+Let $\rho$ be a run with $\canrun[\rho] = a_1 \cdots a_n$, and let $\sigma, \tau$ be runs with $\rho \runleq_f \sigma$ and $\rho \runleq_g \sigma$.
 
-    Then, for all $k \in \Nat$, and $0 < \ell < n$, the following words belong to 
-    the language of the \kl{amalgamation system}.
-    \begin{enumerate}
-        \item $(\prod_{i = 0}^{\ell - 1} u_i^{x_i} v_i u_i^{y_i} v_i u_i^{z_i}) 
-               v_\ell u_\ell^k v_\ell
-               (\prod_{i = \ell+1}^{n} u_i^{x_i} v_i u_i^{y_i} v_i u_i^{z_i})$,
-        \item $(\prod_{i = 0}^{n - 1} u_i^{x_i} v_i u_i^{y_i}) 
-               v_n u_n^k$,
-        \item $u_0^k v_0 
-            (\prod_{i = 1}^{n} u_i^{x_i} v_i u_i^{y_i})$.
-    \end{enumerate}
-    Where, for all $0 \leq i \leq n$, $x_i + y_i + z_i = k$, with $z_i = 0$
-    in the two last items.
+	For any $0 \leq \ell \leq n$, we have $\GapWord{f}{\ell} \infleq \GapWord{g}{\ell}$ or vice versa.
 \end{lemma}
+
 \begin{proof}
-    The result immediately follows from the embedding property
-    applied inductively on runs obtained by gluing 
-    $k$ times $s$ with one or two times $t$ (depending on the item number).
+	Write $u$ for $\GapWord{f}{\ell}$ and $v$ for $\GapWord{g}{\ell}$. 
+	We may assume that both $u$ and $v$ are non-empty, as otherwise the lemma holds trivially.
+	Then, for all $k \in \Nat$, there exists a run with canonical decomposition
+	$$
+	w_k = L_0 a_1 \cdots a_n L_n,
+	$$
+	where $L_i \in \set{vv u^k, vu^kv, u^k vv}$ and specifically $L_\ell = vu^kv$.
+
+	From \cref{pumping-periods:lem}, we may conclude that there are a finite number of words $x, y,$ and $w$ 
+	such that each $w_k$ is contained in a language 
+	$\InfPeriodChain{x}w\InfPeriodChain{y}$.
+
+	As there is an infinite number of words $w_k$, 
+	we may fix $x, y,$ and $w$ and an infinite subset $I \subseteq \Nat$ 
+	such that $\set{w_i \mid i \in I} \subseteq \InfPeriodChain{x}w\InfPeriodChain{y}$. 
+	This implies that either for infinitely many $m \in \Nat$, $u^m v \in \InfPeriodChain{y}$ 
+	or for infinitely many $m$, $v u^m \in \InfPeriodChain{x}$. 
+	In either case, we may conclude that either $u \infleq v$ or $v \infleq u$.
 \end{proof}
-
-To conclude this section and prove our main \cref{infix-amalgamation:thm}, it
-suffices to demonstrate that languages recognized by \kl{amalgamation systems}
-that are \kl{well-quasi-ordered} for the \kl{infix relation} are \kl{bounded
-languages}.
 
@@
     Let us consider the set $M$ of minimal runs for the relation $\leq_E$,
-    which is finite because the latter is a \kl{well-quasi-ordering}. By
-    definition, for every word $w \in L$, there exists $r \in M$, $s \in R$,
-    and $f \in E(r,s)$ such that $w = \canrun(s)$.
-    In particular, we conclude that
-    \begin{equation*}
-        L \subseteq \bigcup_{r \in M} 
-        \left(
-        \GapLanguage{r}{0}
-        \prod_{i = 1}^{\card{\canrun(r)}-1} \canrun(r)_i \GapLanguage{r}{i} 
-        \right)
-        \quad .
-    \end{equation*}
-
-    Let $u$ and $v$ be two words in the \kl{gap language}
-    $\GapLanguage{r}{\ell}$ for some $r \in M$ and $0 < \ell <
-    \card{\canrun(r)} \defined n$. One can apply \cref{pumping-gap-languages:lem}
-    to conclude that  for all $k \geq 1$,
-    \begin{equation*}
-       \left(\prod_{i = 0}^{\ell - 1} u_i^{x_i} v_i u_i^{y_i} v_i u_i^{z_i}\right) 
-       v u^k v
-       \left(\prod_{i = \ell+1}^{n} u_i^{x_i} v_i u_i^{y_i} v_i u_i^{z_i}\right)
-       \in 
-       L
-    \end{equation*}
-    Where $x_i + y_i + z_i = k$ for all $0 \leq i \leq n$
-    Let us assume without loss of generality that the sequence
-    $\seqof{(x_i, y_i, n_i)}[i \geq 1]$ is pointwise increasing,
-    and even increasing coordinate wise. This is possible because
-    at least one of the coordinates tends to $+\infty$ and in the case
-    of a variable that is bounded, one can extract the sequence to
-    simply not use this variable.
-
-    Now, leveraging \cref{pumping-periods:lem}, we get that for infinitely many
-    $k \geq 1$, $u v^k u$ is an infix of some $x^p$ where $x$ has size smaller
-    or equal than $u$ and $v$. This proves  that $u$ and $v$ are comparable for
-    the prefix, infix, and suffix relations. In particular, we
-    conclude that $\GapLanguage{r}{i}$ is a \kl{bounded language}, because of
-    \cite[Lemma 4.1]{ASZZ24}, when $0 < i < \card{\canrun(r)}$. The proof goes
-    on similarly for proving that $\GapLanguage{r}{0}$ and $\GapLanguage{r}{n}$
-    are \kl{bounded languages}, using the other items of
-    \cref{pumping-gap-languages:lem}.
-
-    Because \kl{bounded languages} are stable under concatenation and finite
-    unions, we conclude that $L$ itself must be a \kl{bounded language}.
+    which is finite because the latter is a \kl{well-quasi-ordering}. 
+    By \cref{gap-words-prefix-ordered:lem}, we know that for each minimal run $\rho \in M$,
+    each gap language $\GapLanguage{\rho}{i}$ of $\rho$ is totally ordered by $\infleq$.
+    Adapting the proof of language boundedness from $\cite[Section 4.2]{ASZZ24}$, we may conclude that $\GapLanguage{\rho}{i} \subseteq \InfPeriodChain{w}$ for some $w \in \GapLanguage{\rho}{i}$.
+    As $\InfPeriodChain{w}$ is language bounded and this property is stable under subsets, concatenation and finite union,
+    we can conclude that $L$ is bounded as well.
 \end{proofof}
 
-Leveraging a similar reasoning, we conclude that being a \kl{bounded language},
-a \kl{regular language}, or being included in a finite union of products of
-\kl{chains} all collapse for \kl{well-quasi-ordered} and \kl{downwards closed}
-languages for the \kl{infix relation}.
+If we additionally assume that such a language is closed under infixes, we obtain an even stronger structure: All such languages are regular:
 
@@
     that is a finite union of products of \kl{chains} is a regular language.
+    \todo[inline]{maybe elaborate on this point}
 \end{proof}
@@
 for the complement of $\LMorse$ \cite{BERST86}. To the knowledge of the
-authors, this conjecture remains open. Our remark provides a negative answer to
-this conjecture for the language $\LMorse$.
+authors, this conjecture remained open so far. With \cref{dwclosed-infixes-wqo:lem} we are able to show it false.
 
@@
 recursively enumerable, the function $\canrun(\cdot)$ is computable, and for
-any two runs $r, s \in R$, the set $E(r,s)$ is computable.
+any two runs $\rho, \sigma \in R$, the set $E(\rho,\sigma)$ is computable.
 
@@
 is \kl{effectively closed under rational transductions}. Recall that a
-\intro{rational transduction} is a \emph{realtion} $R \subseteq \Sigma^* \times
+\intro{rational transduction} is a \emph{relation} $R \subseteq \Sigma^* \times
 \Gamma^*$ that can be defined by a (nondeterministic) finite automaton with
@@
 \begin{proofof}{infix-wqo-is-emptiness:thm}
-	\cref{emptiness-decidable} $\Rightarrow$ \cref{wqo-infix-decidable}. We aim to make the inclusion test of \cref{bounded-language:eq} effective.
-
+	\cref{emptiness-decidable} We first show $\Rightarrow$ \cref{wqo-infix-decidable}. We aim to make the inclusion test of \cref{bounded-language:eq} of \cref{bounded-language:thm} effective. 
 	Let $R(n,m,N_0) \defined \bigcup_{x,y \in \Sigma^{\leq n}} \bigcup_{u \in
@@
 
-    To determine the value for $N_0$, we first computer maximal subsets $S
-    \subseteq [n]$ such that $w_i$ for $i \in S$ are simultaneously
-    unbounded.\todo{define this} We do this by applying a transduction mapping
-    $w_i$ to some fresh letter $a_i$ if $i \in S$ and $\varepsilon$ otherwise,
-    and testing for simultaneous unboundedness \cite[Section 4.1]{ASZZ24}.
-    Given a candidate bound $n_0$ for $N_0$, we construct the languages
-    $L_{S,n_0} = w_1^{\circ_1}\cdots w_n^{\circ_n}$ where $\circ_i = *$ if $i
-    \in S$ and $\circ_i = \leq n_0$ otherwise. We then test if $L \subseteq
-    \bigcup_{S \text{ maximal}} L_{S,n_0}$. If yes, $n_0$ is our value for
-    $N_0$, otherwise, we increase it and repeat the construction. As we chose
-    maximal sets $S$, this procedure will eventually terminate.
+    To determine the value for $N_0$, we then compute the downward closure (with respect to subwords) of $L$. 
+    This is effective and yields a finite-state automaton. 
+    Recall that $N_0$ is the maximum number of repetitions of a word $w_i$ that can not be iterated arbitrarily often. 
+    This value is therefore bounded above by the length of the longest simple path in this automaton.
 
@@
     the image of $L$ through $R$ is \kl{well-quasi-ordered}
-    by \kl{infix} if and only if $L$ is empty.
+    by \kl{prefix} if and only if $L$ is empty.
 \end{proofof}

### 2024-11-13 src/amalgam.tex
Revert "Amalgamation systems: Fix definitions and simplify proofs (#8)"
This reverts commit 6d855952072ff4d444f8e773fd1c2d144f342661.
Fix lowhanging fruits of #8

@@
             or $\GapWord{h}{j} = \GapWord{g}{j} \GapWord{f}{j}$. 
-            Specifically, for $i$ we may fix $G_{h}{i} = \GapWord{f}{j} \GapWord{g}{j}$.
+            Specifically, for $i$ we may fix $\GapWord_{h}{i} = \GapWord{f}{i} \GapWord{g}{i}$.
             We refer to \cref{amalgamation-runs:fig} for an illustration 
@@
 	The \emph{yield} of a run is obtained by projecting away the separator symbol \cansep~from the canonical decomposition, i.e.
-$\intro*s\yieldrun(\rho) = \project_\Sigma( \canrun[\rho])$.
+$\intro*\yieldrun(\rho) = \project_\Sigma( \canrun[\rho])$.
 The language recognized by an \kl{amalgamation system} is $\yieldrun(R)$.
@@
     each gap language $\GapLanguage{\rho}{i}$ of $\rho$ is totally ordered by $\infleq$.
-    Adapting the proof of language boundedness from $\cite[Section 4.2]{ASZZ24}$, we may conclude that $\GapLanguage{\rho}{i} \subseteq \InfPeriodChain{w}$ for some $w \in \GapLanguage{\rho}{i}$.
+    Adapting the proof of language boundedness from \cite[Section 4.2]{ASZZ24}, we may conclude that $\GapLanguage{\rho}{i} \subseteq \InfPeriodChain{w}$ for some $w \in \GapLanguage{\rho}{i}$.
     As $\InfPeriodChain{w}$ is language bounded and this property is stable under subsets, concatenation and finite union,
@@
 conclude that the collection of \kl{infixes} of the \kl{Thue-Morse sequence}
-cannot be recognized by \emph{any} \kl{amalgamation system}. To put this result
-in context, it was proven that the complement of the set $T$ of prefixes of the
-\kl{Thue-Morse sequence} is context-free, and conjectured that the same holds
-for the complement of $\LMorse$ \cite{BERST86}. To the knowledge of the
-authors, this conjecture remained open so far. With \cref{dwclosed-infixes-wqo:lem} we are able to show it false.
+cannot be recognized by \emph{any} \kl{amalgamation system}.
 

### 2024-11-14 src/amalgam.tex

@@
 	or for infinitely many $m$, $v u^m \in \InfPeriodChain{x}$.
-	In either case, we may conclude that either $u \infleq v$ or $v \infleq u$.
+
+	In either case, we may conclude that either $u \infleq v$ or $v \infleq u$: Let $m, n \in \Nat$
+	such that $m < n$ and $u^m v, u^n v \in \InfPeriodChain{y}$ (the case for $v u^m$ and $v u^n$ 
+	proceeding analogously). Without loss of generality, assume that $\card{u^m}$ and $\card{u^n}$ are
+	multiples of $\card{y}$. We therefore find $p \prefleq y, s \suffleq y$ such that $u^m, u^n \in sy^*p$, 
+	ergo $ps = y$.
+	In other words, we can write $u^m = (sp)^{m'}, u^n = (sp)^{n'}$. As $u^mv \in \InfPeriodChain{y}$, it 
+	follows that $v$ is a prefix of some word in $(sp)^*$. Hence either $v$ is a prefix of $u$ or $u$ vice versa.
 \end{proof}

### 2024-12-01 src/amalgam.tex
minor changes

@@
 \cref{infixes-regular:sec}, and in
-particular \cref{bounded-language:thm}, can safely be transferred to
+particular \cref{bounded-language:thm}, can safely be applied to
 \kl{amalgamation systems}.
@@
 assumptions, testing \kl{well-quasi-ordering} is inter-reducible to testing
-whether a language of the class is empty.
+whether a language of the class is empty, which is usually the simplest
+problem for a computational model.
+
+\todo[inline]{ What is the complexity of the reduction? Logspace? }
 
@@
             or $\GapWord{h}{j} = \GapWord{g}{j} \GapWord{f}{j}$. 
-            Specifically, for $i$ we may fix $\GapWord_{h}{i} = \GapWord{f}{i} \GapWord{g}{i}$.
+            Specifically, for $i$ we may fix $\GapWord{h}{i} = \GapWord{f}{i} \GapWord{g}{i}$.
             We refer to \cref{amalgamation-runs:fig} for an illustration 

### 2024-12-01 src/infixes.tex
add root file annotations
minor changes

@@
 % LTeX: language=en-GB 
+% !TeX root=../wqo-on-words.tex
 \section{Infixes and Bounded Languages}
@@
 relation} can lead to very complicated \kl{well-quasi-ordered} languages.
-Formally, the upcoming \cref{infix-embedding:thm} shows that
+Formally, the upcoming \cref{infix-embedding:thm} due to Kuske shows that
 \emph{any} countable partial-ordering with finite initial segments can be
@@
     Remark that the \kl{downwards closure} of a product of two \kl{chains}
-    is a product of two chains.
+    is a finite union of products of two chains.
     As a consequence, we conclude that $\dwset[\infleq]{L}$ is itself included

### 2024-12-01 src/introduction.tex
add root file annotations
minor changes

@@
 % LTeX: language=en-GB 
+% !TeX root=../wqo-on-words.tex
 \section{Introduction}
@@
 $\preceq$ such that every infinite sequence $\seqof{x_n}$ of elements taken in
-$X$ contains a pair $x_i \preceq x_j$ with $i < j$. Well-quasi-orderings serve
+$X$ contains an increasing pair $x_i \preceq x_j$ with $i < j$. Well-quasi-orderings serve
 as a core combinatorial tool powering many termination arguments, and was
@@
 \kl{well-quasi-ordered} with respect to the \kl{prefix relation} (respectively,
-the \kl{suffix relation} or the \kl{infix relation}), and together with
+the \kl{suffix relation} or the \kl{infix relation}), and designing
 suitable decision procedures.
 
-\AP Let us argue that this question is not only of theoretical interest.
+\AP Let us argue that these decision procedures fit a larger picture in the
+research area of well-quasi-orderings.
 Indeed, there has been recent breakthroughs in deciding whether a given order
@@
 infinite state transition systems~\cite{DBLP:conf/fsttcs/FinkelG19} or in the
-context of logic~\cite{DBLP:journals/pacmpl/BergstrasserGLZ24}. Our decision
-procedures clearly contribute to this line of research in the context of words.
+context of logic~\cite{DBLP:journals/pacmpl/BergstrasserGLZ24}.
 Furthermore, a previous work by Kuske shows that any
@@
 \cite{kriz90b}. Therefore, when characterizing \kl{well-quasi-ordered}
-languages, we will also be interested in computing bounds on their \kl{ordinal
-invariants}. We refer to \cref{ordinal-invariants:subsec}
-for a more detailed introduction to these
-parameters and ordinal computations in general.
+languages, we will also be interested in deriving upper bounds on their
+\kl{ordinal invariants}. We refer to \cref{ordinal-invariants:subsec} for a
+more detailed introduction to these parameters and ordinal computations in
+general.
 
@@
 \kl{suffix relation}, we are able to show the following decomposition into
-\intro{chains}, that is, totally ordered ascending sequences.
+\intro{chains}, that are totally ordered ascending sequences.
 
@@
 and to the class $\mathcal{C}_\text{cfg}$ of all languages recognized by context-free grammars,
-which are both \kl{effective amalgamative classes} \cite{aut-cfg-infix:cor}.
+which are both \kl{effective amalgamative classes} \cref{aut-cfg-infix:cor}.
 

### 2024-12-01 src/prefixes.tex
add root file annotations
minor changes

@@
 % LTeX: language=en-GB 
+% !TeX root=../wqo-on-words.tex
 \section{Prefixes and Suffixes}
@@
 In order to characterize the existence of infinite \kl{antichains} for the
-\kl{prefix relation}, we will introduce the following tree construction that
-will be useful in the rest of this section.
+\kl{prefix relation}, we will introduce the following tree.
 
@@
 
-\begin{figure}
-    \centering
-    \includestandalone[width=\linewidth]{fig/prefix-core-and-branches-standalone}
-    \caption{Illustrating the construction of 
-        \cref{prefixes:thm}. We represent the Hasse Diagram of a
-        \kl{partial order}, with elements in $S$ being bigger and 
-        in desert sand, elements of $S_{\max}$ being in
-        blood-red, and \kl{chains} of elements in sky blue.
-        The upwards closure of an element in $S_{\max}$ 
-        is depicted in pale green, and cannot intersect two distinct
-        branches, as it does not contain two incomparable elements.
-    }
-    \label{prefixes-theorem:fig}
-\end{figure}
+% \begin{figure}
+%     \centering
+%     \includestandalone[width=\linewidth]{fig/prefix-core-and-branches-standalone}
+%     \caption{Illustrating the construction of 
+%         \cref{prefixes:thm}. We represent the Hasse Diagram of a
+%         \kl{partial order}, with elements in $S$ being bigger and 
+%         in desert sand, elements of $S_{\max}$ being in
+%         blood-red, and \kl{chains} of elements in sky blue.
+%         The upwards closure of an element in $S_{\max}$ 
+%         is depicted in pale green, and cannot intersect two distinct
+%         branches, as it does not contain two incomparable elements.
+%     }
+%     \label{prefixes-theorem:fig}
+% \end{figure}
 
@@
 As an immediate consequence, we have a very fine-grained understanding of the
-\kl{ordinal invariants} of such \kl{well-quasi-ordered} languages, which could be
+\kl{ordinal invariants} of such \kl{well-quasi-ordered} languages, which can be
 leveraged in bounding the complexity of algorithms working on such languages.
@@
 \end{corollary}
+
+\todo[inline]{
+    We actually can \emph{compute} the ordinal invariants
+    in the regular case.
+}

### 2024-12-16 src/infixes.tex
Restructure
Ordinals

@@
 
+\section{Infixes and Downwards Closed Languages}
+
 One may think that all \kl{downwards closed} languages for the \kl{infix
@@
     \label{thue-morse-ordinal:lemma}
-    The \kl{maximal order type} of $\LMorse$ is $\omega$,
-    the \kl{ordinal height} of $\LMorse$ is $\omega$,
-    the \kl{ordinal width} of $\LMorse$ is $\omega$.
+    The \kl{maximal order type} of $\LMorse$ is $\omegaOrd$,
+    the \kl{ordinal height} of $\LMorse$ is $\omegaOrd$,
+    the \kl{ordinal width} of $\LMorse$ is $\omegaOrd$.
 \end{lemma}
@@
     the maximal length of a \kl{bad sequence} is determined by its first element,
-    hence that it is at most $\omega$.
+    hence that it is at most $\omegaOrd$.
     Finally, because the \kl{ordinal width} is at most the \kl{maximal order type},
-    we conclude that the \kl{ordinal width} is also at most $\omega$.
+    we conclude that the \kl{ordinal width} is also at most $\omegaOrd$.
 
     Now, let us prove that these bounds are tight. It is clear that
-    $\oHeight{\LMorse} = \omega$: given any number $n \in \Nat$, one can construct a
+    $\oHeight{\LMorse} = \omegaOrd$: given any number $n \in \Nat$, one can construct a
     \kl{decreasing sequence} of words in $L$ of length $n$, for instance by
@@
     decreasing size.
-    Let us now prove that $\oWidth{\LMorse} = \omega$. Assume by contradiction that
+    Let us now prove that $\oWidth{\LMorse} = \omegaOrd$. Assume by contradiction that
     $\oWidth{\LMorse}$ is finite. Then, $L$ can be written as a finite union of
@@
     Finally, because the \kl{ordinal width} is at most the \kl{maximal order
-    type}, we conclude that the \kl{maximal order type} of $\LMorse$ is also $\omega$.
+    type}, we conclude that the \kl{maximal order type} of $\LMorse$ is also $\omegaOrd$.
 \end{proof}

### 2024-12-16 src/introduction.tex
Ordinals

@@
         \kl{well-quasi-ordered} by the \kl{prefix relation} 
-        (resp. \kl{suffix relation}) have an \kl{ordinal height} of at most $\omega$,
+        (resp. \kl{suffix relation}) have an \kl{ordinal height} of at most $\omegaOrd$,
         a finite \kl{ordinal width}, and a \kl{maximal order type} strictly
-        below $\omega^2$.
+        below $\omegaOrd^2$.
         \hfill (\cref{prefixes-ordinal-invariants:cor} p.\pageref{prefixes-ordinal-invariants:cor})
@@
 very large \kl{ordinal invariants}: for instance, the \kl{maximal order type}
-of the \kl{subword relation} is $\omega^{\omega^{\card{\Sigma} - 1}}$, which
+of the \kl{subword relation} is $\omegaOrd^{\omegaOrd^{\card{\Sigma} - 1}}$, which
 equals its \kl{ordinal width} \cite[Corollary 3.9, Theorem 4.21]{DZSCSC20}.
@@
 \begin{enumerate}
-    \setcounter{enumi}{5}
+    \setcounter{enumi}{6}
     \item 

### 2024-12-16 src/prelims.tex
Ordinals

@@
 An \intro{ordinal} is a \kl{well-founded} \kl{totally ordered}
-set. We use $\alpha, \beta, \gamma$ to denote ordinals, and use $\omega$ to
-denote the first infinite ordinal, i.e., the set of natural numbers with the
+set. We use $\alpha, \beta, \gamma$ to denote ordinals, and use $\intro*\omegaOrd$ to
+denote the first infinite \kl{ordinal}, i.e., the set of natural numbers with the
 usual ordering. We assume superficial familiarity with ordinal arithmetic, and
@@
 II]{KRIVINE71} for a detailed introduction to this domain. Given a tree $T$
-whose branches are all finite we can define an ordinal $\alpha_T$ inductively
+whose branches are all finite we can define an \kl{ordinal} $\alpha_T$ inductively
 as follows: if $T$ is a leaf then $\alpha_T = 0$, if $T$ has children

### 2024-12-16 src/infixes.tex
New theorems

@@
     that is \kl{downwards closed}.
-    Then, the \kl{maximal order type} of $L$ is at most $\omega^3$,
-    its \kl{ordinal height} is at most $\omega$,
-    and its \kl{ordinal width} is at most $\omega^2$.
+    Then, the \kl{maximal order type} of $L$ is at most $\omegaOrd^3$,
+    its \kl{ordinal height} is at most $\omegaOrd$,
+    and its \kl{ordinal width} is at most $\omegaOrd^2$.
 \end{conjecture}
+
+\todo[inline]{
+    Write this in a nicer way.
+    Maybe talk a bit more about automatic sequences?
+    This could be a section ``Infixes and downwards closed languages''. Before the
+    ``infixes and amalgamation systems''.
+}
+
+\begin{theorem}
+    \label{small-ordinal-invariants:thm}
+    Let $L$ be a \kl{well-quasi-ordered} language for the \kl{infix relation}
+    that is \kl{downwards closed}.
+    Then, the \kl{maximal order type} of $L$ is strictly less than $\omegaOrd^3$,
+    its \kl{ordinal height} is at most $\omegaOrd$,
+    and its \kl{ordinal width} is at most $\omegaOrd^2$.
+\end{theorem}
+
+The \cref{small-ordinal-invariants:thm} is a direct consequence of the following
+remarks.
+
+\begin{itemize}
+    \item If $L$ is \kl{wqo} and \kl{downwards closed}, then $L = \bigcup_{i = 1}^n L_i$ where $L_i$
+        is an \kl{order ideal}.
+    \item If $L$ is an \kl{order ideal} that is \kl{well-quasi-ordered}, then
+        $L$ is the set of infixes of some \emph{bi-infinite} word $w \in \Sigma^{\Rel}$.
+    \item The set of infixes of a \emph{bi-infinte} word can be understood 
+        as the set of \kl{infixes} of two infinite words $u,v \in \Sigma^{\Nat}$.
+    \item The set of infixes of an infinite word $w \in \Sigma^{\Nat}$ is 
+        \kl{well-quasi-ordered} if and only if $w$ is \kl{ultimately uniformly recurrent},
+        that is, if there exists an infinite suffix of $w$ that is \kl{uniformly recurrent}.
+    \item The ordinal invariants of a \kl{uniformly recurrent} word are small.
+\end{itemize}
+
+Note that being \kl{uniformly recurrent} still leaves open a lot of potential behaviours,
+for instance, Sturmian words are \kl{uniformly recurrent}, but are not \kl{automatic}.
+\todo{Find citations for this.}
+
+\begin{theorem}
+    Given an \kl{automatic sequence} $w \in \Sigma^{\Nat}$, one can decide
+    whether the set of \kl{infixes} of $w$ is \kl{well-quasi-ordered} for the
+    \kl{infix relation}, and if so, compute the associated \kl{ordinal
+    invariants}.
+\end{theorem}
+\begin{proof}
+    One can check if $w$ is \kl{ultimately uniformly recurrent} by checking if
+    $\exists N_0 \in \Nat, \dots$. Because $w$ is \kl{automatic}, this is decidable.
+    Now, to decide whether $w$ has width $\omegaOrd$, then one can check if $w$ is 
+    \kl{ultimately periodic}. If not, then $w$ has width $\omegaOrd$.
+    If it is \kl{ultimately periodic}, one can compute the smallest size of
+    the period, which gives us the width.
+    Using the width, and because we know the height is at most $\omegaOrd$, we can
+    compute the maximal order type.
+\end{proof}

### 2024-12-16 src/introduction.tex
New theorems

@@
 relation}.
+\todo[inline]{Rephrase: all downwards closed languages have understandable invariants.
+    And bounded languages are WQO if and only if their downward clsosure is!}
 {
@@
             $L \subseteq \Sigma^*$ that are \kl{well-quasi-ordered} by the \kl{infix relation}
-            have an \kl{ordinal height} of at most $\omega$,
-            an \kl{ordinal width} strictly below $\omega^2$, and
-            a \kl{maximal order type} strictly below $\omega^3$.
+            have an \kl{ordinal height} of at most $\omegaOrd$,
+            an \kl{ordinal width} strictly below $\omegaOrd^2$, and
+            a \kl{maximal order type} strictly below $\omegaOrd^3$.
+            \todo{Add reference to the corollary}
+    \item All \kl{downwards closed} languages $L$
+            that are \kl{well-quasi-ordered} by the \kl{infix relation}
+            have an \kl{ordinal height} of at most $\omegaOrd$,
+            an \kl{ordinal width} strictly below $\omegaOrd^2$, and
+            a \kl{maximal order type} strictly below $\omegaOrd^3$.
+            \todo{Add reference to the corollary}
 \end{enumerate}

### 2024-12-16 src/amalgam.tex
typos
Restructure
Todos
adding a lot of todos before arxiv

@@
-\section{Decision Prodecure and Amalgamation Systems}
+\section{Infixes and Amalgamation Systems}
 \label{infixes-amalgamation:sec}
@@
 
-\todo[inline]{ What is the complexity of the reduction? Logspace? }
+\todo[inline]{What is the complexity of the reduction? Logspace? }
 
@@
     \centering
+    \todo[inline]{simplify the figure, only two letter initial word}
     \includestandalone[width=\linewidth]{fig/run-amalgamation-standalone}
@@
     that is a finite union of products of \kl{chains} is a regular language.
-    \todo[inline]{maybe elaborate on this point}
+
+    \todo[inline]{This proof is not very satisfying.}
+    Indeed, assume that
+    $L$ is \kl{downwards closed} and included in a finite union of sets of the form
+    $\InfPeriodChain{x} u \InfPeriodChain{y}$ where $x,y,u$ are possibly empty words.
+    We can assume without loss of generality that
+    for every $n$, $x^n u y^n$ is in $L$, otherwise, we have a bound on the maximal $n$ such that
+    $x^n u y^n$ is in $L$, and we can increase the number of languages in the union, replacing $x$ or $y$
+    with the empty word as necessary.
+    Let us write $L' \defined \bigcup_{i = 1}^k x_i^* u_i y_i^*$. Then, $L'
+    \subseteq L$ by construction. Furthermore, $L \subseteq \dwset{L'}$, also
+    by construction. Finally, we conclude that $L = \dwset{L'}$ because $L$ is
+    \kl{downwards closed}. Now, because $L'$ is a \kl{regular language}, and 
+    \kl{regular languges} are closed under \kl{downwards closure}, we conclude
+    that $L$ is a \kl{regular language}.
 \end{proof}
@@
 \end{proof}
+
+\todo[inline]{
+    Add computing the \kl{ordinal invariants}
+    of the \kl{infix relation} for languages 
+    recognized by \kl{amalgamation systems}.
+}

### 2024-12-16 src/introduction.tex
adding a lot of todos before arxiv

@@
 to consider its so-called \kl{ordinal invariants}: for instance, the
-\kl{maximal order type} (or m.o.t.), originally defined by De Jongh and Parikh
+\kl{maximal order type} (or \kl{m.o.t.}), originally defined by De Jongh and Parikh
 \cite{dejongh77}, is the order type of the maximal linearization of a
-well-quasi-ordered set. In the case of a finite set, the m.o.t. is precisely
+well-quasi-ordered set. In the case of a finite set, the \kl{m.o.t.} is precisely
 the size of the set. Better runtime bounds were obtained by considering two
@@
 such languages.
+
+\todo[inline]{Add a table with the results. But we have to be able to say
+    that bounds are \emph{tight}, and also distinguish between 
+    hypotheses (bounded, downwards closed) and representations
+    (amalgamation system, automatic etc).
+    Plus, it would be good to give \emph{names} to the characterisations,
+    so that we can write them directly in the table.
+    Plus it would be nice to have the corresponding theorem number in the table.
+    \cref{summary:tab}.
+}
+
+\begin{table}
+    \caption{Summary of the results on the \kl{well-quasi-ordered} languages
+    with respect to the \kl{prefix relation}, the \kl{suffix relation}, and the
+    \kl{infix relation}.}
+    \label{summary:tab}
+\begin{center}
+    \begin{tabular}{lp{6cm}l}
+        \toprule
+        \textbf{Hypothesis} & \textbf{Characterization} & $\oWidth{ \cdot }$ \\
+        \midrule
+        \multicolumn{3}{c}{\textbf{Prefix and Suffix Relations}} \\
+        \midrule
+        \addlinespace
+        None & finite union of \kl{chains} & $< \omegaOrd$ \\
+        None & finite union of \kl{chains} & $< \omegaOrd$ \\
+        \midrule
+        % infix for the next four rows in the first column
+        \multicolumn{3}{c}{\textbf{Infix Relation}} \\
+        \midrule
+        \addlinespace
+        None & None & $< \omegaOne$ \\
+        \kl{Downwards Closed} & finite unions of infixes of \kl{ultimately uniformly recurrent} bi-infinite words & $< \omegaOrd^3$ \\
+        \kl{Automatic} & infixes of \kl{ultimately uniformly recurrent} bi-infinite words & $< \omegaOrd^2$ \\
+        \kl{Bounded} & finite union of products of \kl{chains} & $< \omegaOrd^2$ \\
+        \kl{Amalgamation System} & infixes of \kl{ultimately uniformly recurrent} bi-infinite words & $< \omegaOrd^2$ \\
+        \bottomrule
+    \end{tabular}
+\end{center}
+\end{table}
 
@@
 edition of \kl{Autobóz} for their helpful comments and discussions.
+We would also like to thank Vincent Jugé for his pointers on word combinatorics.

### 2024-12-16 src/prelims.tex
adding a lot of todos before arxiv

@@
 denote the first infinite \kl{ordinal}, i.e., the set of natural numbers with the
-usual ordering. We assume superficial familiarity with ordinal arithmetic, and
+usual ordering. We also use $\intro*\omegaOne$ to denote the first \emph{uncountable}
+ordinal.
+We only assume superficial familiarity with ordinal arithmetic, and
 refer to the books of Kunen \cite{KUNEN80} and Krivine~\cite[Chapter

### 2025-01-10 src/amalgam.tex
fix mismatch between unions of products and actual «power chains»

@@
         \item\label{dwci-bod:item} $L$ is a \kl{bounded language},
-        \item\label{dwci-uoc:item} $L$ is included in a finite union of products of \kl{chains}
+        \item\label{dwci-uoc:item} There exists 
+            a finite set $E \subseteq (\Sigma^*)^3$
+            such that $L = \bigcup_{(x,u,y) \in E} \InfPeriodChain{x} u \InfPeriodChain{y}$.
     \end{enumerate}
@@
     The implication \cref{dwci-bod:item} $\Rightarrow$ \cref{dwci-uoc:item}
-    is \cref{bounded-language:thm}.
+    is \cref{bounded-language:lem}.
     Finally, the implication \cref{dwci-uoc:item} $\Rightarrow$ \cref{dwci-reg:item}
@@
 
-    \todo[inline]{This proof is not very satisfying.}
     Indeed, assume that

### 2025-01-10 src/infixes.tex
adding a lot of todos before arxiv
fix mismatch between unions of products and actual «power chains»

@@
 
-\begin{proofof}{bounded-language:thm}
+\begin{lemma}
+    \label{bounded-language:lem}
+    Let $L \subseteq \Sigma^*$ be a \kl{bounded language}
+    that is \kl{well-quasi-ordered} by the \kl{infix relation}.
+    Then, there exists a finite subset $E \subseteq (\Sigma^*)^3$,
+    such that:
+    \begin{equation*}
+        L \subseteq \bigcup_{(x,u,y) \in E} \InfPeriodChain{x} u \InfPeriodChain{y}
+        \quad .
+    \end{equation*}
+\end{lemma}
+\begin{proof}
     Let $w_1, \dots, w_n$ be such that
@@
     \end{equation}
+\end{proof}
+
+\begin{proofof}{bounded-language:thm}
+    We apply \cref{bounded-language:lem}, and conclude
+    because $\InfPeriodChain{x}$ is a finite union of \kl{chains}
+    for the \kl{prefix}, \kl{suffix} and \kl{infix} relations
+    (\cref{inf-period-chain:lem}).
 \end{proofof}

### 2025-01-10 src/infixes.tex
adding proofs for uniformly recurrent words

@@
 \begin{lemma}
+    \label{uniformly-recurrent:lem}
+    Let $w \in \Sigma^\Nat$ be a \kl{uniformly recurrent} word.
+    Then, the set of finite \kl{infixes} of $w$ is \kl{well-quasi-ordered} for the \kl{infix relation}.
+\end{lemma}
+\begin{proof}
+    Let $L$ be the set of finite \kl{infixes} of $w$.
+    Consider a sequence $\seqof{u_i}$ of words in $L$. Without loss of
+    generality, we may consider a subsequence such that $\card{u_i} <
+    \card{u_{i+1}}$ for all $i \in \Nat$. Because $\ThueMorse$ is \kl{uniformly
+    recurrent}, there exists $k \geq 1$ such that $u_1$ is an \kl{infix} of
+    every word $v$ of size at least $k$. In particular, $u_1$ is an \kl{infix}
+    of $u_k$, hence the sequence $\seqof{u_i}$ is \kl(sequence){good}.
+\end{proof}
+
+\begin{lemma}
     \label{thue-morse:lemma}
@@
     \kl{well-quasi-ordered} for the \kl{infix relation}, but is not \kl(language){bounded}.
+    It is \kl{well-quasi-ordered} because it is the set of finite \kl{infixes}
+    of a \kl{uniformly recurrent} word, and we can apply \cref{uniformly-recurrent:lem}.
 
@@
     because $u \infleq \ThueMorse$, which is \kl{cube-free}.
-
-    Furthermore, $L$ is \kl{well-quasi-ordered} for the \kl{infix relation}.
-    Indeed, consider a sequence $\seqof{u_i}$ of words in $L$. Without loss of
-    generality, we may consider a subsequence such that $\card{u_i} <
-    \card{u_{i+1}}$ for all $i \in \Nat$. Because $\ThueMorse$ is \kl{uniformly
-    recurrent}, there exists $k \geq 1$ such that $u_1$ is an \kl{infix} of
-    every word $v$ of size at least $k$. In particular, $u_1$ is an \kl{infix}
-    of $u_k$, hence the sequence $\seqof{u_i}$ is \kl(sequence){good}.
 \end{proof}
@@
 
-This suggests the following conjecture which states that for \kl{downwards
+This motivated the following theorem, which states that for \kl{downwards
 closed} languages, the \kl{ordinal invariants} are relatively small.
-
-\begin{conjecture}
-    \label{small-ordinal-invariants:conj}
-    Let $L$ be a \kl{well-quasi-ordered} language for the \kl{infix relation}
-    that is \kl{downwards closed}.
-    Then, the \kl{maximal order type} of $L$ is at most $\omegaOrd^3$,
-    its \kl{ordinal height} is at most $\omegaOrd$,
-    and its \kl{ordinal width} is at most $\omegaOrd^2$.
-\end{conjecture}
-
-\todo[inline]{
-    Write this in a nicer way.
-    Maybe talk a bit more about automatic sequences?
-    This could be a section ``Infixes and downwards closed languages''. Before the
-    ``infixes and amalgamation systems''.
-}
 
@@
 
-The \cref{small-ordinal-invariants:thm} is a direct consequence of the following
-remarks.
+\AP A subset $I \subseteq X$ is \intro{directed} if, for every $x,y \in I$,
+there exists $z \in I$ such that $x \leq z$ and $y \leq z$. Given a
+\kl{well-quasi-order} $(X, \leq)$, one can always decompose $X$ into a finite
+union of \intro{order ideals}, that is, non-empty sets $I \subseteq X$ that are
+\kl{downwards closed} and \kl{directed} for the \kl{order relation} $\leq$. In
+our case, a \kl{well-quasi-ordered} \kl{order ideal} for the \kl{infix
+relation} is the set of finite \kl{infixes} of a finite or bi-infinte word $w
+\in \Sigma^\Nat \cup \Sigma^\Rel $ (\cref{bi-infinite:lem}). Therefore, we will
+first study languages that are described as the set of finite \kl{infixes} of
+some (bi-)infinite word.
+
+\begin{lemma}
+    \label{bi-infinite:lem}
+    Let $L \subseteq \Sigma^*$ be an \kl{order ideal} that is \kl{well-quasi-ordered} 
+    for the 
+    \kl{infix relation}. Then $L$ is the set of finite \kl{infixes}
+    of a finite, infinite or bi-infinite word $w$.
+\end{lemma}
+\begin{proof}
+    Let us assume that $L$ is infinite, the case when it is finite 
+    is similar, but leads to a finite word.
+
+    Because the alphabet $\Sigma$ is finite, we can enumerate the words of $L$
+    as $\seqof{w_i}$. Now, let us build the sequence $\seqof{u_i}$ by induction
+    as follows: $u_0 = w_0$, and $u_{i+1}$ is a word that contains $u_i$ and
+    $w_i$, which exists in $L$ because $L$ is \kl{directed}. Since $L$ is
+    \kl{well-quasi-ordered}, one can extract an infinite set of indices $I
+    \subseteq \Nat$ such that $u_i \infleq u_{j}$ for all $i \leq j \in I$.
+
+    We can build a word $w$ as the limit of the sequence $\seqof{u_i}[i \in
+    I]$. This word is infinite or bi-infinite, and contains as infixes all the
+    words $u_i$ for $i \in I$. Because every word of $L$ is an infix of every
+    $u_i$ for a large enough $I$, one concludes that $L$ is contained in the
+    set of finite infixes of $w$. Conversely, every finite infix of $w$ is 
+    an infix of some $u_i$ by definition of the limit construction, hence
+    belongs to $L$ since $u_i \in L$ and $L$ is \kl{downwards closed}.
+\end{proof}
+
+\AP We say that an infinite word $w \in \Sigma^\Nat$ is \intro{ultimately
+uniformly recurrent} if there exists a bound $N_0 \in \Nat$ such that $w_{\geq
+N_0}$ is \kl{uniformly recurrent}.
+
+\begin{lemma}
+    \label{ultimately-uniformly-recurrent:lem}
+    Let $w \in \Sigma^\Nat$ be an infinite word. 
+    Then, the set of finite infixes of $w$ is \kl{well-quasi-ordered} for the \kl{infix relation}
+    if and only if $w$ is \kl{ultimately uniformly recurrent}.
+\end{lemma}
+\begin{proof}
+
+    Assume that $w$ is \kl{ultimately uniformly recurrent}. Consider a sequence
+    of words $\seqof{w_i}$ that are finite \kl{infixes} of $w$. Because $w$ is
+    \kl{ultimately uniformly recurrent}, there exists a bound $N_0$ such that
+    $w_{\geq N_0}$ is \kl{uniformly recurrent}. Let $i < N_0$, we claim that,
+    without loss of generality, only finitely many words in the sequence
+    $\seqof{w_i}$ can be found starting at the position $i$ in $w$. Indeed, if
+    it is not the case, then we have an infinite subsequence of words that are
+    all comparable for the \kl{infix relation}, and therefore a \kl{good
+    sequence}, because the \kl{infix relation} is \kl{well-founded}. We can
+    therefore assume that all words in the sequence $\seqof{w_i}$ are such that
+    they start at a position $i \geq N_0$. But then, they are all finite
+    \kl{infixes} of $w_{\geq N_0}$, which is a \kl{uniformly recurrent} word,
+    whose set of finite \kl{infixes} is \kl{well-quasi-ordered}
+    (\cref{uniformly-recurrent:lem}).
+
+    Conversely, assume that the set of finite infixes of $w$ is
+    \kl{well-quasi-ordered}. Let us write $\mathsf{Rec}(w)$ the set of finite
+    \kl{infixes} of $w$ that appear infinitely often. We can similarly define
+    $\mathsf{Rec}(w_{\geq i})$ for any (infinite) suffix of $w$. The sequence
+    $R_i \defined \mathsf{Rec}(w_{\geq i})$ is a sequence of \kl{downwards
+    closed} sets of finite words, included in the set of finite infixes of $w$
+    by definition. Because the latter is \kl{well-quasi-ordered}, there exists
+    an $N_0 \in \Nat$, such that $\bigcap_{i \in \Nat} R_i = R_{N_0}$.
+    Now, consider $v \defined w_{\geq N_0}$. By construction, 
+    every finite infix of $v$ 
+    appears infinitely often in $v$. Let us consider 
+    some finite infix $u \infleq v$, we claim that there is a bound $N_u$
+    on the distance
+    between two consecutive occurrences of $u$ in $v$.
+    Indeed, if it is not the case, then there exists an infinite sequence
+    $\seqof{u x_i u}[i \in \Nat]$ of infixes of $v$, such that $x_i$ is a word that is
+    of size $\geq i$.
+    Because the finite infixes of $w$ (hence, of $v$) are \kl{well-quasi-ordered},
+    one can extract an infinite set of indices $I \subseteq \Nat$
+    such that $u x_i u \infleq u x_{j} u$ for all $i \leq j \in I$.
+    In particular, $u x_i u \infleq u x_{j} u$ for some $j > |x_i|$, 
+    which contradicts the fact that $u x_j u$ coded two consecutive
+    occurrences of $u$ in $v$.
+
+    We have shown that for every finite infix $u$ of $v$, there exists a bound
+    $N_u$ such that every two occurrences of $u$ in $v$ start at distance at
+    most $N_u$. In particular, there exists a bound $M_u$ such that every infix
+    of $v$ of size at least $M_u$ contains $u$. We have proven that
+    $v$ is \kl{uniformly recurrent}, hence that $w$ is \kl{ultimately uniformly
+    recurrent}.
+\end{proof}
+
+\begin{lemma}
+    \label{small-ordinal-invariants:lem}
+    Let $w \in \Sigma^\Nat$ be an \kl{ultimately uniformly recurrent} word.
+    Then, the set of finite infixes of $w$ has \kl{ordinal width}
+    less than $2 \cdot \omegaOrd$.
+\end{lemma}
+\begin{proof}
+    \todo[inline]{prove this}
+\end{proof}
+
+\begin{lemma}
+    \label{from-bi-to-single:lem}
+    Let $w \in \Sigma^\Rel$ be a bi-infinite word. Then, the set of finite
+    infixes of $w$ is \kl{well-quasi-ordered} for the \kl{infix relation} if
+    and only if $w_+$ and $w_-$ are two \kl{ultimately uniformly recurrent}
+    words. In this case, the \kl{ordinal width}
+    of the set of finite infixes of $w$ is less than $3 \cdot \omegaOrd$.
+\end{lemma}
+
+Before proving \cref{small-ordinal-invariants:thm}, let us restrict 
+our attention to a simpler case:
+
+The \cref{small-ordinal-invariants:thm} is a direct consequence of the
+following remarks.
 

### 2025-01-13 src/amalgam.tex
adding proofs for uniformly recurrent words, end

@@
     $\canrun \colon R \to (\Sigma \uplus \set{\cansep})^*$ is a 
-    function computing a \emph{canonical decomposition} of a run,
-    and $E$ describes the so-called \emph{admissible embeddings} between runs: If $\rho$ and $\sigma$ are runs from $R$, then $E(\rho, \sigma)$ is a subset of the subword embeddings between $\canrun[\rho]$ and $\canrun[\sigma]$.
+    function computing a \intro{canonical decomposition} of a run,
+    and $E$ describes the so-called \intro{admissible embeddings} between runs: If $\rho$ and $\sigma$ are runs from $R$, then $E(\rho, \sigma)$ is a subset of the subword embeddings between $\canrun[\rho]$ and $\canrun[\sigma]$.
 We write $\rho \intro*\runleq \sigma$ if $E(\rho, \sigma)$ is non-empty.
@@
     \kl{downwards closed}. Now, because $L'$ is a \kl{regular language}, and 
-    \kl{regular languges} are closed under \kl{downwards closure}, we conclude
+    \kl{regular languages} are closed under \kl{downwards closure}, we conclude
     that $L$ is a \kl{regular language}.

### 2025-01-13 src/infixes.tex
adding proofs for uniformly recurrent words, end

@@
 
+\begin{corollary}
+    \label{ordinal-invariants-bounded:cor}
+    Let $L$ be a \kl{bounded language} of $\Sigma^*$
+    that is \kl{well-quasi-ordered} by the \kl{infix relation}.
+    Then, the \kl{ordinal width} of $L$ is finite,
+    its \kl{ordinal height} is at most $\omega$,
+    and its \kl{maximal order type} is at most $\omega$.
+\end{corollary}
+
 \section{Infixes and Downwards Closed Languages}
@@
 
-This motivated the following theorem, which states that for \kl{downwards
-closed} languages, the \kl{ordinal invariants} are relatively small.
+We prove if the upcoming theorem that the status of the \kl{Thue-Morse
+sequence} is actually representative of most \kl{downwards closed} languages
+for the \kl{infix relation}. Indeed, we show that the \kl{ordinal invariants}
+of such languages are relatively small, and the proof of this fact relies on
+the connection between \kl{well-quasi-ordered} languages and \kl{uniformly
+recurrent} words.
 
@@
 
-\AP A subset $I \subseteq X$ is \intro{directed} if, for every $x,y \in I$,
-there exists $z \in I$ such that $x \leq z$ and $y \leq z$. Given a
+\AP A subset $I \subseteq X$ is \intro(subset){directed} if, for every $x,y \in
+I$, there exists $z \in I$ such that $x \leq z$ and $y \leq z$. Given a
 \kl{well-quasi-order} $(X, \leq)$, one can always decompose $X$ into a finite
 union of \intro{order ideals}, that is, non-empty sets $I \subseteq X$ that are
-\kl{downwards closed} and \kl{directed} for the \kl{order relation} $\leq$. In
-our case, a \kl{well-quasi-ordered} \kl{order ideal} for the \kl{infix
-relation} is the set of finite \kl{infixes} of a finite or bi-infinte word $w
-\in \Sigma^\Nat \cup \Sigma^\Rel $ (\cref{bi-infinite:lem}). Therefore, we will
+\kl{downwards closed} and \kl(subset){directed} for the relation $\leq$. In our
+case, a \kl{well-quasi-ordered} \kl{order ideal} for the \kl{infix relation} is
+the set of finite \kl{infixes} of a finite or bi-infinte word $w \in
+\Sigma^\Nat \cup \Sigma^\Rel $ (\cref{bi-infinite:lem}). Therefore, we will
 first study languages that are described as the set of finite \kl{infixes} of
-some (bi-)infinite word.
+some (bi-)infinite word (\cref{ultimately-uniformly-recurrent:lem}). To that
+end, let us introduce the notation $\intro*\infset{w}$ for the set of finite
+\kl{infixes} of a (possibly infinite or bi-infinite) word $w$.
 
@@
     as follows: $u_0 = w_0$, and $u_{i+1}$ is a word that contains $u_i$ and
-    $w_i$, which exists in $L$ because $L$ is \kl{directed}. Since $L$ is
+    $w_i$, which exists in $L$ because $L$ is \kl(subset){directed}. Since $L$ is
     \kl{well-quasi-ordered}, one can extract an infinite set of indices $I
@@
 \begin{proof}
-    \todo[inline]{prove this}
+    Let $N_0$ be a bound such that $w_{\geq N_0}$ is \kl{uniformly recurrent}.
+    Let us write $\infset{w}$ the set of finite infixes of $w$.
+    We prove that $\oWidth{\infset{w}} \leq \omegaOrd + N_0$.
+    Let $u_1 \infleq w$ be a finite word.
+
+    If $u_1$ is an \kl{infix} of $w_{\geq N_0}$, then there exists $k \geq 1$
+    such that $u_1$ is an \kl{infix} of every word of size at least $k$. In
+    particular, there is finite bound on the length of every sequence of
+    incomparable elements starting with $u_1$. We conclude in particular that
+    $\infset{w} \setminus \upset{u_1}$ has a finite \kl{ordinal width}.
+
+    Otherwise, $u_1$ can only be found \emph{before} $N_0$. In this case, we
+    consider a second element of a \kl{bad sequence} $u_2 \infleq w$, which is
+    incomparable with $u_1$ for the \kl{infix relation}. If $u_2$ is an
+    \kl{infix} of $w_{\geq N_0}$, then we can conclude as before. Otherwise,
+    notice that $u_1$ and $u_2$ cannot start at the same position in $w$
+    (because they are incomparable). Continuing this argument, we conclude that
+    there are at most $N_0$ elements starting before $N_0$
+    at the start of any sequence of
+    incomparable elements in $\infset{w}$. We conclude that
+    $\oWidth{\infset{w}} \leq \omegaOrd + N_0$.
 \end{proof}
+
+Let us briefly discuss the tightness of the bound in
+\cref{small-ordinal-invariants:lem}. The \kl{Thue-Morse sequence} over a binary
+alphabet $\set{a,b}$ has \kl{ordinal width} $\omegaOrd$. Given a number $N_0
+\in \Nat$, one can construct an arbitrarily long \kl{antichain} of words for
+the \kl{infix relation} by using a new letter $c$. When concatenating this
+(finite) antichain as a prefix of the \kl{Thue-Morse sequence}, one obtains a
+new (infinite) word $w$. It is clear that the \kl{ordinal width} of
+$\infset{w}$ is now at least $\omegaOrd + N_0$. Hence, the bound $\oWidth{X} <
+2 \cdot \omegaOrd$ is tight.
+
+The behaviour of bi-infinite words is very similar to the infinite ones. Given
+a bi-infinite word $w \in \Sigma^{\Rel}$, we can consider $w_+ \in \Sigma^\Nat$
+and $w_- \in \Sigma^\Nat$ the two infinite words obtained as follows: for all
+$i \in \Nat$, $(w_+)_i = w(i)$ and $(w_-)_i = w(-i)$. Note that the two share
+the letter at position $0$.
 
@@
 \end{lemma}
+\begin{proof}
+    Assume that $w_+$ and $w_-$ are \kl{ultimately uniformly recurrent}. Let us
+    write $\infset{w}$ the set of finite infixes of $w$. Consider an infinite
+    sequence of words $\seqof{u_i}[i \in \Nat]$ in $\infset{w}$. If there is an
+    infinite subsequence of words that are all in $\infset{w_+}$, then there
+    exists an increasing pair of indices $i < j$ such that $u_i \infleq
+    u_j$ because \cref{uniformly-recurrent:lem} applies to $w_+$. Similarly, if
+    there is an infinite subsequence of words that are all in $\infset{w_-}$,
+    then there exists an increasing pair of indices $i < j$ such that $u_i
+    \infleq u_j$ because \cref{uniformly-recurrent:lem} applies to $w_-$ (and
+    the \kl{infix relation} is compatible with mirroring). Otherwise, one can
+    assume without loss of generality that all words in the sequence have a
+    starting position in $w_-$ and an ending position in $w_+$. In this case,
+    let us write $(k_i,l_i) \in \Nat^2$ the pair of indices such that $u_i$ is the infix
+    of $w$ that starts at position $-k_i$ of $w$ (i.e., $k_i$ of $w_-$) and
+    ends at position $l_i$ of $w$ (i.e., $l_i$ of $w_+$).
+    Because $\Nat^2$ is a \kl{well-quasi-ordering} with the product ordering,
+    there exists $i < j$ such that $k_i \leq k_j$ and $l_i \leq l_j$, 
+    in particular, $u_i \infleq u_j$. We have proven that every infinite
+    sequence of words in $\infset{w}$ is \kl(wqo){good}, hence $\infset{w}$ is
+    \kl{well-quasi-ordered}.
 
-Before proving \cref{small-ordinal-invariants:thm}, let us restrict 
-our attention to a simpler case:
+    Conversely, assume that $\infset{w}$ is \kl{well-quasi-ordered}. In
+    particular, $\infset{w_+} \subseteq \infset{w}$ is \kl{well-quasi-ordered}.
+    Similarly, $\infset{w_-}$ is \kl{well-quasi-ordered} because the \kl{infix
+    relation} is compatible with mirroring. Applying
+    \cref{ultimately-uniformly-recurrent:lem}, we conclude that both are
+    \kl{ultimately uniformly recurrent} words.
 
-The \cref{small-ordinal-invariants:thm} is a direct consequence of the
-following remarks.
+    To obtain the upper bound of $3 \cdot \omegaOrd$, we can consider the same
+    argument as for \cref{small-ordinal-invariants:lem}. We let $N_0$ be such
+    that $w_{\geq N_0}$ and $(w_-)_{\geq N_0}$ are \kl{uniformly recurrent}
+    words. In any sequence of incomparable elements of $\infset{w}$, there are
+    less than $N_0^2$ elements that are found in $(w_{< N_0})_{> -N_0}$. Then,
+    one has to pick a finite \kl{infix} in either $w_{\geq N_0}$ or $w_{\leq
+    -N_0}$. Because of \cref{small-ordinal-invariants:lem}, any sequence of
+    incomparable elements of these two infinite words has length bounded based
+    on the choice of the first element of that sequence. This means that the
+    \kl{ordinal width} of $\infset{w}$ is at most $\omegaOrd + \omegaOrd +
+    N_0^2$. We conclude that $\oWidth{\infset{w}} < 3 \cdot \omegaOrd$.
+\end{proof}
 
-\begin{itemize}
-    \item If $L$ is \kl{wqo} and \kl{downwards closed}, then $L = \bigcup_{i = 1}^n L_i$ where $L_i$
-        is an \kl{order ideal}.
-    \item If $L$ is an \kl{order ideal} that is \kl{well-quasi-ordered}, then
-        $L$ is the set of infixes of some \emph{bi-infinite} word $w \in \Sigma^{\Rel}$.
-    \item The set of infixes of a \emph{bi-infinte} word can be understood 
-        as the set of \kl{infixes} of two infinite words $u,v \in \Sigma^{\Nat}$.
-    \item The set of infixes of an infinite word $w \in \Sigma^{\Nat}$ is 
-        \kl{well-quasi-ordered} if and only if $w$ is \kl{ultimately uniformly recurrent},
-        that is, if there exists an infinite suffix of $w$ that is \kl{uniformly recurrent}.
-    \item The ordinal invariants of a \kl{uniformly recurrent} word are small.
-\end{itemize}
+As for \cref{small-ordinal-invariants:lem}, let us briefly discuss
+that the bound in \cref{from-bi-to-single:lem} is tight. Indeed, one can
+construct a bi-infinite word $w$ by concatenating a reversed \kl{Thue-Morse
+sequence} on a binary alphabet $\set{a,b}$, a finite antichain of arbitrarily
+large size over a distinct alphabet $\set{c,d}$, and then a \kl{Thue-Morse
+sequence} on a binary alphabet $\set{e,f}$. The \kl{ordinal width} of the set
+of \kl{infixes} of $w$ is then at least $2 \cdot \omegaOrd + K$, where $K$ is the
+size of the chosen antichain.
 
-Note that being \kl{uniformly recurrent} still leaves open a lot of potential behaviours,
-for instance, Sturmian words are \kl{uniformly recurrent}, but are not \kl{automatic}.
-\todo{Find citations for this.}
+We are now ready to conclude the proof of \cref{small-ordinal-invariants:thm}.
+\begin{proofof}{small-ordinal-invariants:thm}
+
+    It is always true that the \kl{ordinal height} of a language over a finite
+    alphabet is at most $\omegaOrd$. Let us now consider a
+    \kl{well-quasi-ordered} language $L$ that is \kl{downwards closed} for the
+    \kl{infix relation}. Because it is a \kl{well-quasi-ordered} set, it can be
+    written as a finite union of \kl{order ideals} $L = \bigcup_{i = 1}^n L_i$.
+
+    For every such \kl(order){ideal} $L_i$, we can apply
+    \cref{bi-infinite:lem}, and conclude that $L_i$ is the set of finite
+    \kl{infixes} of a finite, infinite or bi-infinite word $w_i$. We can then
+    directly conclude that $\oWidth{L_i}$ less than $\omegaOrd$ (in the case of
+    a finite word), less than $2 \cdot \omegaOrd$ (in the case of an infinite
+    word thanks to \cref{small-ordinal-invariants:lem}), or less than $3 \cdot
+    \omegaOrd$ (in the case of a bi-infinite word, thanks to
+    \cref{from-bi-to-single:lem}). In any case,
+    we have the bound $\oWidth{L_i} < 3 \cdot \omegaOrd$.
+
+    Now, $\oWidth{L} \leq \sum_{i = 1}^n \oWidth{L_i} < 3 \cdot \omegaOrd <
+    \omegaOrd^2$. We conclude that the \kl{ordinal width} of $L$ is less than
+    $\omegaOrd^2$. Finally, the inequality $\oType{L} \leq \oWidth{L} \oComProd
+    \oHeight{L} < \omega \oComProd \omega^2 = \omegaOrd^3$ allows us to conclude.
+\end{proofof}
+
+\AP As we have demonstrated, infinite (or bi-infinite words) can be used to
+represent languages that are \kl{well-quasi-ordered} for the \kl{infix
+relation} by considering their set of finite \kl{infixes}. We can leverage the
+theory of \kl{automatic sequences}, that is tailored to study the regularity of
+infinite words to build a decision procedure for \kl{well-quasi-ordered}
+languages for the \kl{infix relation}. A sequence $w \in \Sigma^\Nat$ is
+\intro(sequence){automatic} whenever there exists a base $b \in \Nat$ and a
+finite automaton that can compute the $i$-th letter of $w$ given as input the
+number $i$ in base $b$. The \kl{Thue-Morse sequence} is an example of an
+\kl{automatic sequence}.
 
@@
     whether the set of \kl{infixes} of $w$ is \kl{well-quasi-ordered} for the
-    \kl{infix relation}, and if so, compute the associated \kl{ordinal
-    invariants}.
+    \kl{infix relation}.
 \end{theorem}
 \begin{proof}
-    One can check if $w$ is \kl{ultimately uniformly recurrent} by checking if
-    $\exists N_0 \in \Nat, \dots$. Because $w$ is \kl{automatic}, this is decidable.
-    Now, to decide whether $w$ has width $\omegaOrd$, then one can check if $w$ is 
-    \kl{ultimately periodic}. If not, then $w$ has width $\omegaOrd$.
-    If it is \kl{ultimately periodic}, one can compute the smallest size of
-    the period, which gives us the width.
-    Using the width, and because we know the height is at most $\omegaOrd$, we can
-    compute the maximal order type.
+    Because of \cref{ultimately-uniformly-recurrent:lem}, it suffices
+    to decide whether $w$ is \kl{ultimately uniformly recurrent}.
+    We can rewrite this as a question on the \kl{automatic sequence} $w$
+    as follows:
+    \begin{align*}
+        &\exists N_0,                   &   \text{ultimately} \\
+        &\forall i_s \geq N_0,          &   \text{for every infix (start) } u \\
+        &\forall i_e > i_s,             &   \text{for every infix (end) }   u \\
+        &\exists k \geq 1,              &   \text{there exists a bound} \\
+        &\forall j_s \geq N_0,          &   \text{for every other infix (start) } v \\
+        &\forall j_e \geq j_s + k,      &   \text{of size at least $k$} \\
+        &\exists l \geq 0,              &   \text{there exists a position in } v \\
+        &\forall 0 \leq m < i_e - i_s,  &   \text{where } u \text{ can be found} \\
+        &j_s + m + l < j_e \land
+        w(i_s + m) = w(j_s + m + l) \quad .
+    \end{align*}
+    Because $w$ is computable by a finite automaton, one can reduce the above
+    formula to a regular language, for which it suffices to check emptiness, which
+    is decidable.
 \end{proof}
+
+Note that being \kl{uniformly recurrent} still leaves open a lot of potential
+behaviours, for instance, Sturmian words are \kl{uniformly recurrent}, but are
+not \kl(sequence){automatic}. Interestingly enough, the \kl{ordinal invariants}
+of the \kl{automatic sequences} are not smaller than the ones of \kl{ultimately
+uniformly recurrent} words: this contrasts with the case of languages that are
+\kl{well-quasi-ordered} for the \kl{infix relation} in general, where the
+hypothesis of being representable (by a \kl{regular language}, or an
+\kl{amalgamation system}) will drastically reduce the \kl{ordinal invariants}
+of said language (\cref{ordinal-invariants-bounded:cor}, \cref{infix-amalgamation:thm}).

### 2025-01-13 src/introduction.tex
Rework the intro and prelims

--- 2024-12-16 src/introduction.tex
@@
 whenever the languages are given by reasonable computational models. Finally,
-we provide relatively low bounds on the possible \kl{ordinal invariants} of
-such languages.
+we provide precise bounds on the possible \kl{ordinal invariants} of such
+languages.
 
-Specifically, for the \kl{prefix relation} and the
-\kl{suffix relation}, we are able to show the following decomposition into
-\intro{chains}, that are totally ordered ascending sequences.
-
-{
-\renewcommand{\labelenumi}{R\arabic{enumi}}
-\begin{enumerate}
-	\item A language $L \subseteq \Sigma^*$ is \kl{well-quasi-ordered} by the \kl{prefix relation} if and only if it is a finite union of \kl{chains} of the \kl{prefix relation}.
-    \hfill (\cref{prefixes:thm} p.\pageref{prefixes:thm})
-	\item A language $L \subseteq \Sigma^*$ is \kl{well-quasi-ordered} by the \kl{suffix relation} if and only if it is a finite union of \kl{chains} of the \kl{suffix relation}.
-    \hfill (\cref{prefixes:thm} p.\pageref{prefixes:thm})
-\end{enumerate}
-}
-
-Note that \kl{chains} are the simplest possible \kl{well-quasi-ordered} sets
-(they are \kl{totally ordered} and \kl{well-founded}) and it is known that finite unions
+In the case of the \kl{prefix} and \kl{suffix} relations, we show that a
+language $L \subseteq \Sigma^*$ is \kl{well-quasi-ordered} by the \kl{prefix
+relation} (resp. \kl{suffix}) if and only if it is a finite union of
+\kl{chains} of the \kl{prefix relation} (resp. \kl{prefix}), where a
+\intro{chain} is a totally ordered set that is well-founded. Note that
+\kl{chains} are the simplest possible \kl{well-quasi-ordered} sets (they are
+\kl{totally ordered} and \kl{well-founded}) and it is known that finite unions
 of \kl{well-quasi-ordered} sets are \kl{well-quasi-ordered}. As a consequence,
@@
 \kl{well-quasi-ordered} sets are \kl{well-quasi-ordered} by the \kl{prefix
-relation}. Let us furthermore highlight that this characterization holds
-without any restriction on the decidability of the language $L$ itself.\footnote{But
-heavily relies on the assumption that $\Sigma$ is finite} This allows us to
-derive the following bounds on the \kl{ordinal invariants} of such
-\kl{well-quasi-ordered} languages, which can be interpreted in two dual ways:
-first, these languages are extremely simple, which means that one could have
-hoped for using directly another well-quasi-ordered set without resorting to
-finite words; second, any time one encounters a \kl{well-quasi-ordered}
-language, it is going to provide relatively efficient algorithms.
-{
-\renewcommand{\labelenumi}{R\arabic{enumi}}
-\begin{enumerate}
-    \setcounter{enumi}{2}
-    \item All languages $L \subseteq \Sigma^*$ are 
-        \kl{well-quasi-ordered} by the \kl{prefix relation} 
-        (resp. \kl{suffix relation}) have an \kl{ordinal height} of at most $\omegaOrd$,
-        a finite \kl{ordinal width}, and a \kl{maximal order type} strictly
-        below $\omegaOrd^2$.
-        \hfill (\cref{prefixes-ordinal-invariants:cor} p.\pageref{prefixes-ordinal-invariants:cor})
-\end{enumerate}
-}
+relation} (\cref{prefixes:thm}). Let us furthermore highlight that this
+characterization holds without any restriction on the decidability of the
+language $L$ itself, but heavily relies on the assumption that $\Sigma$ is
+finite. This allows us to derive tight bounds on the \kl{ordinal invariants} of
+such \kl{well-quasi-ordered} languages, which can be interpreted in two dual
+ways: first, these languages are extremely simple, which means that one could
+have hoped for using directly another well-quasi-ordered set without resorting
+to finite words; second, any time one encounters a \kl{well-quasi-ordered}
+language, it is going to provide relatively efficient algorithms. Indeed, we
+prove that all languages $L \subseteq \Sigma^*$ are \kl{well-quasi-ordered} by
+the \kl{prefix relation} (resp. \kl{suffix relation}) have an \kl{ordinal
+height} of at most $\omegaOrd$, a finite \kl{ordinal width}, and a \kl{maximal
+order type} strictly below $\omegaOrd^2$
+(\cref{prefixes-ordinal-invariants:cor}).
 
@@
 very large \kl{ordinal invariants}: for instance, the \kl{maximal order type}
-of the \kl{subword relation} is $\omegaOrd^{\omegaOrd^{\card{\Sigma} - 1}}$, which
-equals its \kl{ordinal width} \cite[Corollary 3.9, Theorem 4.21]{DZSCSC20}.
-However, not all is lost if we restrict ourselves to \kl{bounded languages},
-i.e., languages that are included in some $w_1^* \cdots w_k^*$ for some finite
-choice of words $w_1, \ldots, w_k$. We are able to show the following:
+of the \kl{subword relation} is $\omegaOrd^{\omegaOrd^{\card{\Sigma} - 1}}$,
+which equals its \kl{ordinal width} \cite[Corollary 3.9, Theorem
+4.21]{DZSCSC20}. We show that in two situations, this can be avoided: when the
+language is \kl{downwards closed} (i.e., when it is closed under taking
+\kl{infixes}), and when the language is \kl(language){bounded} (i.e., 
+when it is included in some $w_1^* \cdots w_k^*$ for some finite choice of
+words $w_1, \ldots, w_k$). In those cases, we are able to characterize
+\kl{well-quasi-ordered} languages by the \kl{infix relation} and derive tight
+bounds on their \kl{ordinal invariants}.
 
-{
-\renewcommand{\labelenumi}{R\arabic{enumi}}
-\begin{enumerate}
-	\setcounter{enumi}{3}
-    \item A \kl{bounded language} $L \subseteq \Sigma^*$ is \kl{well-quasi-ordered} by the \kl{infix relation} if and only if it is (included in) a finite union of languages $S_i \cdot P_i$, where each $S_i$
-        is a \kl{chain} for the \kl{suffix relation},
-        and where
-        each $P_i$ is a \kl{chain} for the \kl{prefix relation}.
+In the case of \kl{bounded languages}, we prove that a \kl{bounded language} $L
+\subseteq \Sigma^*$ is \kl{well-quasi-ordered} by the \kl{infix relation} if
+and only if it is (included in) a finite union of languages $S_i \cdot P_i$,
+where each $S_i$ is a \kl{chain} for the \kl{suffix relation}, and where each
+$P_i$ is a \kl{chain} for the \kl{prefix relation}
+(\cref{bounded-language:thm}) This result directly translates into upper bounds
+on the possible \kl{ordinal invariants} of such languages similarly as for the
+\kl{prefix relation}. Notice that these upper bounds are significantly smaller
+than \kl{ordinal invariants} of the \kl{subword relation}: they have an \kl{ordinal
+height} of at most $\omegaOrd$, an \kl{ordinal width} strictly below
+$\omegaOrd^2$, and a \kl{maximal order type} strictly below $\omegaOrd^3$
+(\cref{ordinal-invariants-bounded:cor}).
 
-        \hfill
-        (\cref{bounded-language:thm} p.\pageref{bounded-language:thm})
-\end{enumerate}
-}
+In the case of \kl{downwards closed} languages, we prove that they are deeply
+related to the notion of \kl{uniformly recurrent words}, borrowed from the
+study of word combinatorics. As an intermediate result, we prove that an
+infinite word $w$ is \kl{ultimately uniformly recurrent} if and only if the set
+of all \kl{infixes} of $w$ is \kl{well-quasi-ordered} by the \kl{infix
+relation} (\cref{ultimately-uniformly-recurrent:lem}). We show that every
+language $L \subseteq \Sigma^*$ that is \kl{downwards closed} and
+\kl{well-quasi-ordered} by the \kl{infix relation} is a finite union of the
+sets of finite \kl{infixes} of some \kl{ultimately uniformly recurrent}
+bi-infinite words. This also proves that such languages have an \kl{ordinal
+height} of at most $\omegaOrd$, an \kl{ordinal width} strictly below
+$\omegaOrd^2$, and a \kl{maximal order type} strictly below $\omegaOrd^3$
+(\cref{small-ordinal-invariants:thm}).
 
-This result directly translates into upper bounds on the possible \kl{ordinal invariants}
-of such languages similarly as for the \kl{prefix relation}. Notice that these
-upper bounds are significantly smaller than \kl{ordinal invariants} of the \kl{subword
-relation}.
-{
-\renewcommand{\labelenumi}{R\arabic{enumi}}
-\begin{enumerate}
-    \setcounter{enumi}{4}
-    \item All \kl{bounded languages}
-            $L \subseteq \Sigma^*$ that are \kl{well-quasi-ordered} by the \kl{infix relation}
-            have an \kl{ordinal height} of at most $\omega$,
-            an \kl{ordinal width} strictly below $\omega^2$, and
-            a \kl{maximal order type} strictly below $\omega^3$.
-\end{enumerate}
-}
-
-\AP Finally, we provide a decision procedure for checking whether a language is
-\kl{well-quasi-ordered}, which takes as inputs languages recognized by
-\kl{amalgamation systems} \cite{ASZZ24}. Such systems will be formally
-introduced in
-\cref{amalgamation-systems:subsec}, but
+Then, we turned our attention to decision procedures. To that end, we need to
+choose a computational model representing languages. For \kl{downwards closed}
+languages, because of their close connection with infinite words, we considered
+a model based on \kl{automatic sequences}. Using this model, we can decide
+whether a language is \kl{well-quasi-ordered} by the \kl{infix relation}
+(\cref{automatic-wqo:thm}). We also studied another representation, where
+languages are recognized by \kl{amalgamation systems} \cite{ASZZ24}. Such
+systems will be formally introduced in \cref{amalgamation-systems:subsec}, but
 for the moment let us just say that they include many classical computational
@@
 amalgamative class}, the formal definition of which we defer to
-\cref{infixes-amalgamation-effective:subsec}.
+\cref{infixes-amalgamation-effective:subsec}. Given a class $\mathcal{C}$ that
+is a \kl{strongly effective amalgamative class} of languages, we designed a
+decision procedure that takes as input a language $L \in \mathcal{C}$, and
+decides whether $L$ is \kl{well-quasi-ordered} by the \kl{prefix relation}, the
+\kl{suffix relation}, or the \kl{infix relation}
+(\cref{infix-wqo-is-emptiness:thm}). Quite surprisingly, we also showed that,
+if a language recognized by an \kl{amalgamation system} is
+\kl{well-quasi-ordered} for the \kl{infix relation}, then it is a \kl{bounded
+language} (\cref{infix-amalgamation:thm}), which automatically bounds the
+\kl{ordinal invariants} of the language. Let us point out that the above result
+implies that the hypothesis of \kl{bounded languages} on the theoretical side
+is not a restriction in practice. As a down-to-earth and more easily
+understandable consequence, our generic decision procedure applies to the class
+$\mathcal{C}_\text{aut}$ of all languages recognized by finite automata, and to
+the class $\mathcal{C}_\text{cfg}$ of all languages recognized by context-free
+grammars, which are both \kl{effective amalgamative classes}
+(\cref{aut-cfg-infix:cor}).
 
-{
-\renewcommand{\labelenumi}{R\arabic{enumi}}
-\begin{enumerate}
-    \setcounter{enumi}{6}
-    \item 
-        Let $L$ be a language recognized by an \kl{amalgamation system}.
-        If $L$ is \kl{well-quasi-ordered} by the \kl{infix relation},
-        then $L$ is a \kl{bounded language}.
-
-        \hfill (\cref{infix-amalgamation:thm} p.\pageref{infix-amalgamation:thm})
-    \item 
-        Let $\mathcal{C}$ be a \kl{strongly effective amalgamative class} of languages.
-        There is a decision procedure that takes as input a language $L \in \mathcal{C}$,
-        and decides whether $L$ is \kl{well-quasi-ordered} by the \kl{prefix relation},
-        the \kl{suffix relation}, or the \kl{infix relation}. 
-        \hfill (\cref{infix-wqo-is-emptiness:thm} p.\pageref{infix-wqo-is-emptiness:thm})
-\end{enumerate}
-}
-Let us point out that the above result implies that the hypothesis of
-\kl{bounded languages} on the theoretical side is not a restriction in
-practice. As a down-to-earth and more easily understandable 
-consequence,
-our generic decision procedure applies 
-to the class $\mathcal{C}_\text{aut}$ of all languages recognized by finite automata,
-and to the class $\mathcal{C}_\text{cfg}$ of all languages recognized by context-free grammars,
-which are both \kl{effective amalgamative classes} \cref{aut-cfg-infix:cor}.
+Finally, we noticed that for \kl{downwards closed} languages that are
+\kl{well-quasi-ordered} by the \kl{infix relation}, being
+\kl(language){bounded} is the same as being \kl(language){regular}.
+Furthermore, a \kl{bounded language} is \kl{well-quasi-ordered} by the
+\kl{infix relation} if and only if its \kl{downwards closure} is
+\kl{well-quasi-ordered} by the \kl{infix relation}
+(\cref{bounded-wqo-dwclosed:cor}). This shows that, for \kl{bounded languages}
+(and therefore, for all languages recognized by \kl{amalgamation systems}) that
+are \kl{well-quasi-ordered} by the \kl{infix relation}, their \kl{downwards
+closure} is a \kl{regular language}. This is a weak version of the usual result
+that the \kl{downwards closure} for the \kl{scattered subword relation} is
+always a \kl{regular language}.
 
@@
 languages by the \kl{prefix relation}. In
-\cref{infixes-regular:sec}, we
+\cref{infixes-bounded:sec}, we
 obtain the \kl[infix relation]{infix} analogue of \cref{prefixes:thm}
 specifically for \kl{bounded languages}
-(\cref{bounded-language:thm}). Finally, 
-we generalize this result to all
+(\cref{bounded-language:thm}). 
+In \cref{infixes-dwclosed:sec}, we study the \kl{downwards closed}
+languages, and compute bounds on their \kl{ordinal invariants} in \cref{small-ordinal-invariants:thm}.
+Finally, 
+we generalize these results to all
 \kl{amalgamation systems} in \cref{infixes-amalgamation:sec}
@@
 and provide a decision procedure for checking whether a language is
-\kl{well-quasi-ordered} by the \kl{infix relation} in
+\kl{well-quasi-ordered} by the \kl{infix relation} (resp. \kl{prefix} and \kl{suffix}) in
 this context (\cref{infix-wqo-is-emptiness:thm}).

### 2025-01-13 src/prelims.tex
Rework the intro and prelims

@@
 
-\paragraph*{Finite words.} In this paper, we use letters $\Sigma, \Gamma$ to
+\paragraph*{Finite words.}
+\AP 
+In this paper, we use letters $\Sigma, \Gamma$ to
 denote finite alphabets, $\Sigma^*$ to denote the set of finite words over
@@
 
-\paragraph*{Orderings and Well-Quasi-Orderings.} A \intro{quasi-order} is a
+\paragraph*{Orderings and Well-Quasi-Orderings.}
+\AP
+A \intro{quasi-order} is a
 reflexive and transitive binary relation, it is a \intro{partial order} if it
@@
 \paragraph*{Ordinal Invariants.} \label{ordinal-invariants:subsec}
+\AP
 An \intro{ordinal} is a \kl{well-founded} \kl{totally ordered}
@@
 refer to the books of Kunen \cite{KUNEN80} and Krivine~\cite[Chapter
-II]{KRIVINE71} for a detailed introduction to this domain. Given a tree $T$
+II]{KRIVINE71} for a detailed introduction to this domain.
+Given a tree $T$
 whose branches are all finite we can define an \kl{ordinal} $\alpha_T$ inductively
@@
 
+\AP
 Let $(X, \leq)$ be a \kl{well-quasi-ordered} set. One can define three
@@
 of well-quasi-ordered sets.
+
+\AP
+We will use the following inequality between \kl{ordinal invariants}, due to
+\cite{kriz90b}, and that was recalled in \cite[Theorem 3.8]{DZSCSC20}:
+$\oType{X} \leq \oHeight{X} \oComProd \oWidth{X}$, where $\intro*\oComProd$ is
+the \intro{commutative ordinal product}, also known as the \reintro{Hessenberg
+product}. We will not recall the definition of this product here, and refer to
+\cite[Section 3.5]{DZSCSC20} for a detailed introduction to this concept. The
+only equalities we will use are $\omegaOrd \oComProd \omegaOrd = \omegaOrd^2$
+and $\omegaOrd^2 \oComProd \omegaOrd = \omegaOrd^3$.

### 2025-01-13 src/amalgam.tex
removing todos + polish

@@
 problem for a computational model.
-
-\todo[inline]{What is the complexity of the reduction? Logspace? }
 
@@
     \centering
-    \todo[inline]{simplify the figure, only two letter initial word}
     \includestandalone[width=\linewidth]{fig/run-amalgamation-standalone}
@@
 
-    We say a language $L$ is an \intro*{amalgamation language} if there exists an \kl{amalgamation system} recognizing it. 
+    We say a language $L$ is an \intro{amalgamation language} 
+    if there exists an \kl{amalgamation system} recognizing it. 
 \end{definition}
@@
 \begin{proof}
-    By definition, the embeddings inside $E(\rho,\sigma)$ are defined as elements
+    By definition, the embeddings inside $E(\rho,\sigma)$ are in
     of $\HigEmb(\canrun(\rho), \canrun(\sigma))$, and they compose properly.
@@
 
-\todo[inline]{In this definition, the alphabet $\Sigma$ can be changed 
-(and we seem to actually use it in the proofs), it is strange.}
-
 \AP We say that a class $\mathcal{C}$ of languages is an \intro{effective
@@
 language $L \in \mathcal{C}$, and every rational transduction $R \subseteq
-\Sigma^* \times \Gamma^*$, the image of $L$ through $R$ --- that is,
- $\setof{v \in \Gamma^*}{\exists u \in L. (u,v) \in R}$ --- is in $\mathcal{C}$
-and effectively computable.
+\Sigma^* \times \Gamma^*$, the image of $L$ through $R$ --- that is, $\setof{v
+\in \Gamma^*}{\exists u \in L. (u,v) \in R}$ --- is in $\mathcal{C}$ and
+effectively computable. In particular, it implies that the class contains
+languages over arbitrary (finite) alphabets.
 
@@
     \kl{prefix} if and only if $f(L)$ is \kl{well-quasi-ordered} by \kl{infix}.
-    \todo{fresh symbols? can we assume we have those??}
 
@@
 \end{proof}
-
-\todo[inline]{
-    Add computing the \kl{ordinal invariants}
-    of the \kl{infix relation} for languages 
-    recognized by \kl{amalgamation systems}.
-}

### 2025-01-13 src/conclusion.tex
add root file annotations
Todos
removing todos + polish

@@
 % LTeX: language=en-GB 
+% !TeX root=../wqo-on-words.tex
 \section{Conclusion}
@@
 \subparagraph{Towards infinite alphabets} In this paper, we restricted our
-attention to \emph{finite} alphabets, having in mind the application to regular
-languages. However, the conclusions of \cref{infix-finite-automata:thm} and
+attention to \emph{finite} alphabets, having in mind the application to
+\kl{regular languages}. However, the conclusions of
+\cref{bounded-language:thm}, \cref{small-ordinal-invariants:thm}, and
 \cref{prefixes:thm} could be conjecture to hold in the case of infinite
@@
 words, the \emph{lexicographic ordering}, which does not fit well in our
-current framework because it is always of \kl{ordinal width} $0$. However, the
+current framework because it is always of \kl{ordinal width} $1$. However, the
 order-type of the lexicographic ordering over \kl{regular languages} has

### 2025-01-13 src/infixes.tex
removing todos + polish

@@
 \section{Infixes and Bounded Languages}
-\label{infixes-regular:sec}
+\label{infixes-bounded:sec}
 
@@
 \section{Infixes and Downwards Closed Languages}
+\label{infixes-dwclosed:sec}
 
@@
     Conversely, assume that $\infset{w}$ is \kl{well-quasi-ordered}. In
-    particular, $\infset{w_+} \subseteq \infset{w}$ is \kl{well-quasi-ordered}.
+    particular, the subset $\infset{w_+} \subseteq \infset{w}$ is \kl{well-quasi-ordered}.
     Similarly, $\infset{w_-}$ is \kl{well-quasi-ordered} because the \kl{infix
@@
 \begin{theorem}
+    \label{automatic-wqo:thm}
     Given an \kl{automatic sequence} $w \in \Sigma^{\Nat}$, one can decide

### 2025-03-24 src/infixes.tex
minor changes + funding

@@
     \infleq)$ a typical example is to start from an infinite \kl{antichain} $A$,
-    together with an enumeration $\seqof{w_i}$ of $A$, and build the language $L
+    together with an enumeration $\seqof{w_i}[i \in \Nat]$ of $A$, and build the language $L
     \defined \setof{ \prod_{i = 0}^n w_i }{ i \in \Nat }$. By definition, $L$ is a
@@
     Let $L$ be the set of finite \kl{infixes} of $w$.
-    Consider a sequence $\seqof{u_i}$ of words in $L$. Without loss of
+    Consider a sequence $\seqof{u_i}[i \in \Nat]$ of words in $L$. Without loss of
     generality, we may consider a subsequence such that $\card{u_i} <
@@
     every word $v$ of size at least $k$. In particular, $u_1$ is an \kl{infix}
-    of $u_k$, hence the sequence $\seqof{u_i}$ is \kl(sequence){good}.
+    of $u_k$, hence the sequence $\seqof{u_i}[i \in \Nat]$ is \kl(sequence){good}.
 \end{proof}
@@
     Because the alphabet $\Sigma$ is finite, we can enumerate the words of $L$
-    as $\seqof{w_i}$. Now, let us build the sequence $\seqof{u_i}$ by induction
+    as $\seqof{w_i}[i \in \Nat]$. Now, let us build the sequence $\seqof{u_i}[i \in \Nat]$ by induction
     as follows: $u_0 = w_0$, and $u_{i+1}$ is a word that contains $u_i$ and
@@
     Assume that $w$ is \kl{ultimately uniformly recurrent}. Consider a sequence
-    of words $\seqof{w_i}$ that are finite \kl{infixes} of $w$. Because $w$ is
+    of words $\seqof{w_i}[i \in \Nat]$ that are finite \kl{infixes} of $w$. Because $w$ is
     \kl{ultimately uniformly recurrent}, there exists a bound $N_0$ such that
@@
     without loss of generality, only finitely many words in the sequence
-    $\seqof{w_i}$ can be found starting at the position $i$ in $w$. Indeed, if
+    $\seqof{w_i}[i \in \Nat]$ can be found starting at the position $i$ in $w$. Indeed, if
     it is not the case, then we have an infinite subsequence of words that are
@@
     sequence}, because the \kl{infix relation} is \kl{well-founded}. We can
-    therefore assume that all words in the sequence $\seqof{w_i}$ are such that
+    therefore assume that all words in the sequence $\seqof{w_i}[i \in \Nat]$ are such that
     they start at a position $i \geq N_0$. But then, they are all finite

### 2025-07-31 src/infixes.tex
broken subscript

@@
 % LTeX: language=en-GB 
-% !TeX root=../wqo-on-words.tex
+% !TeX root=../wqo-on-words.lncs.tex
 \section{Infixes and Bounded Languages}
@@
 \begin{theorem}[restate=bounded-language:thm,label=bounded-language:thm]
-    %\label{bounded-language:thm}
+    \label{bounded-language:thm}
     Let $L$ be a \kl{bounded language} of $\Sigma^*$. Then,
@@
 \kl{well-quasi-ordered} for the \kl{infix relation}. This proves the (easy)
-right-to-left implication of \cref{bounded-language:thm}. Furthermore, any
-language $L$ included in a finite union of products $S_i \cdot P_i$ where $S_i$
-is in fact a \kl{bounded language}.
+right-to-left implication of \cref{bounded-language:thm}.
 
@@
     and we conclude that $u x^{\ell} v \suffleq w u x^\ell v = x^{\ell + 1} v \suffleq u x^k v$.
-    \qedhere
 \end{proof}
@@
 
-Let us now discuss the implications of this characterization in terms of
-\kl{downwards closures}: if $L$ is a \kl{bounded language}, then considering
-$L$ or its \kl{downwards closure} is equivalent with respect to being
-\kl{well-quasi-ordered} by the \kl{infix relation}.
+\begin{corollary}
+    \label{ordinal-invariants-bounded:cor}
+    Let $L$ be a \kl{bounded language} of $\Sigma^*$
+    that is \kl{well-quasi-ordered} by the \kl{infix relation}.
+    Then, the \kl{ordinal width} of $L$ less than $\omega^2$,
+    its \kl{ordinal height} is at most $\omega$,
+    and its \kl{maximal order type} less than $\omega^3$.
+    Furthermore, those three bounds are tight.
+\end{corollary}
+\begin{proof}
+  Upper bounds are a direct consequence of \cref{bounded-language:thm},
+  and the tightness is witnessed by the 
+  languages: 
+  $L_k \defined \bigcup_{i = 2}^{k+1} (a b^i a)^* (b a^i b)^*$,
+  that are \kl{bounded languages} of $\set{a,b}^*$,
+  \kl{well-quasi-ordered} by the \kl{infix relation},
+  and have \kl{ordinal width}, \kl{ordinal height} and
+  \kl{maximal order type} respectively equal to $k \cdot \omega$, $\omega$ and $k \cdot \omega^2$.
+\end{proof}
+
+\section{Infixes and Downwards Closed Languages}
+\label{infixes-dwclosed:sec}
+
+Let us now discuss another classical restriction that can be imposed on 
+languages when studying \kl{well-quasi-orders}, that of being
+\kl{downwards closed}. Indeed, the \cref{infix-embedding:thm}
+crucially relies on constructing languages that are \emph{not}
+\kl{downwards closed}, and we have shown 
+in \cref{dw-closure-not-wqo:rem} that the \kl{downwards closure}
+of a \kl{well-quasi-ordered} language is not necessarily
+\kl{well-quasi-ordered}.
+
+\subsection{Characterization of Well-Quasi-Ordered Downwards Closed Languages}
+
+An immediate consequence of \cref{bounded-language:thm} is that   if $L$ is a
+\kl{bounded language}, then considering $L$ or its \kl{downwards closure}
+$\dwset[\infleq]{L}$ is equivalent with respect to being
+\kl{well-quasi-ordered} by the \kl{infix relation}, as opposed to
+the general case illustrated in \cref{dw-closure-not-wqo:rem}.
 
@@
 \end{proof}
-
-\begin{corollary}
-    \label{ordinal-invariants-bounded:cor}
-    Let $L$ be a \kl{bounded language} of $\Sigma^*$
-    that is \kl{well-quasi-ordered} by the \kl{infix relation}.
-    Then, the \kl{ordinal width} of $L$ is finite,
-    its \kl{ordinal height} is at most $\omega$,
-    and its \kl{maximal order type} is at most $\omega$.
-\end{corollary}
-
-\section{Infixes and Downwards Closed Languages}
-\label{infixes-dwclosed:sec}
 
@@
 
-\begin{lemma}
+\begin{theorem}
     \label{uniformly-recurrent:lem}
@@
     Then, the set of finite \kl{infixes} of $w$ is \kl{well-quasi-ordered} for the \kl{infix relation}.
-\end{lemma}
+\end{theorem}
 \begin{proof}
@@
     and its \kl{ordinal width} is at most $\omegaOrd^2$.
+
+    Furthermore, those bounds are tight.
 \end{theorem}
@@
     less than $2 \cdot \omegaOrd$.
+    Furthermore, this bound is tight. 
 \end{lemma}
@@
     $\oWidth{\infset{w}} \leq \omegaOrd + N_0$.
-\end{proof}
 
-Let us briefly discuss the tightness of the bound in
-\cref{small-ordinal-invariants:lem}. The \kl{Thue-Morse sequence} over a binary
-alphabet $\set{a,b}$ has \kl{ordinal width} $\omegaOrd$. Given a number $N_0
+  Let us now justify that this bound is tight.
+  The \kl{Thue-Morse sequence} over a binary
+  alphabet $\set{a,b}$ has \kl{ordinal width} $\omegaOrd$
+  from \cref{thue-morse-ordinal:lemma}.
+    Given a number $N_0
   \in \Nat$, one can construct an arbitrarily long \kl{antichain} of words for
@@
   new (infinite) word $w$. It is clear that the \kl{ordinal width} of
-$\infset{w}$ is now at least $\omegaOrd + N_0$. Hence, the bound $\oWidth{X} <
-2 \cdot \omegaOrd$ is tight.
+  $\infset{w}$ is now at least $\omegaOrd + N_0$.
+\end{proof}
 
-The behaviour of bi-infinite words is very similar to the infinite ones. Given
-a bi-infinite word $w \in \Sigma^{\Rel}$, we can consider $w_+ \in \Sigma^\Nat$
-and $w_- \in \Sigma^\Nat$ the two infinite words obtained as follows: for all
-$i \in \Nat$, $(w_+)_i = w(i)$ and $(w_-)_i = w(-i)$. Note that the two share
-the letter at position $0$.
+The behaviour of bi-infinite words is very similar to the behaviour of infinite
+ones. Given a bi-infinite word $w \in \Sigma^{\Rel}$, we can consider $w_+ \in
+\Sigma^\Nat$ and $w_- \in \Sigma^\Nat$ the two infinite words obtained as
+follows: for all $i \in \Nat$, $(w_+)_i = w(i)$ and $(w_-)_i = w(-i)$. Note
+that the two share the letter at position $0$.
 
@@
     words. In this case, the \kl{ordinal width}
-    of the set of finite infixes of $w$ is less than $3 \cdot \omegaOrd$.
+    of the set of finite infixes of $w$ is less than $3 \cdot \omegaOrd$,
+    and this bound is tight.
 \end{lemma}
@@
     N_0^2$. We conclude that $\oWidth{\infset{w}} < 3 \cdot \omegaOrd$.
-\end{proof}
 
-As for \cref{small-ordinal-invariants:lem}, let us briefly discuss
-that the bound in \cref{from-bi-to-single:lem} is tight. Indeed, one can
+  Let us briefly argue that the bound is tight. Indeed, one can
   construct a bi-infinite word $w$ by concatenating a reversed \kl{Thue-Morse
@@
   of \kl{infixes} of $w$ is then at least $2 \cdot \omegaOrd + K$, where $K$ is the
-size of the chosen antichain.
+  size of the chosen antichain, following the same argument as in the proof of
+  \cref{small-ordinal-invariants:lem}, using \cref{thue-morse-ordinal:lemma}.
+\end{proof}
 
@@
 \begin{proofof}{small-ordinal-invariants:thm}
-
     It is always true that the \kl{ordinal height} of a language over a finite
@@
     \oHeight{L} < \omega \oComProd \omega^2 = \omegaOrd^3$ allows us to conclude.
+
+    The tightness of the bounds is a direct consequence of
+    \cref{from-bi-to-single:lem}, and by considering a finite union of 
+    these examples over disjoint alphabets (or even, by considering a binary 
+    alphabet and using unambiguous codes to separate the different components).
 \end{proofof}
+
+\subsection{Decision Procedures}
+\label{decision-procedures:sec}
 
@@
 represent languages that are \kl{well-quasi-ordered} for the \kl{infix
-relation} by considering their set of finite \kl{infixes}. We can leverage the
-theory of \kl{automatic sequences}, that is tailored to study the regularity of
-infinite words to build a decision procedure for \kl{well-quasi-ordered}
-languages for the \kl{infix relation}. A sequence $w \in \Sigma^\Nat$ is
-\intro(sequence){automatic} whenever there exists a base $b \in \Nat$ and a
-finite automaton that can compute the $i$-th letter of $w$ given as input the
-number $i$ in base $b$. The \kl{Thue-Morse sequence} is an example of an
-\kl{automatic sequence}.
+relation} by considering their set of finite \kl{infixes}.
+Let us formalise the representation of languages by sets of bi-infinite
+words that we will use in this section, following the characterization of
+\cref{bi-infinite:lem}. A \intro{sequence representation} of a language $L
+\subseteq \Sigma^*$ is a finite set of triples $(w_i^-, a_i, w_i^+)_{1 \leq i
+\leq n}$ where $w_i^-, w_i^+ \in \Sigma^\Nat \cup \Sigma^*$ are two potentially
+infinite words, and $a_i \in \Sigma$ is a letter, such that
+\begin{equation*}
+    L = \bigcup_{i = 1}^n \infset{\mathsf{reversed}(w_i^-) a_i w_i^+} \quad .
+\end{equation*}
 
-\begin{theorem}
-    \label{automatic-wqo:thm}
+\AP Given an effective representation of sequences, one obtains an effective
+representation of languages via \kl{sequence representations}. In this section,
+we will rely on definitions originating from the area of symbolic dynamics,
+that precisely study infinite words whose generation follows from a finitely
+described process. However, we will not assume that the reader is familiar with
+this domain, and we will use as black-boxes key results from this area.
+
+\AP A first model that one can use to represent infinite words is the model of
+\intro{automatic sequences}. In this case, the infinite word $w$ is described
+by a finite state automaton, that can compute the $i$-th letter of the word $w$
+given as input the number $i$ written in some base $b \in \Nat$. An example of
+such a sequence is the \kl{Thue-Morse sequence} that can be described by a
+finite automaton using a binary representation of the indices. The good
+algorithmic properties of \kl{automatic sequences} come from the fact that a
+Presburger definable property that uses letters of the sequence can be
+(trivially) translated into a finite automaton that reads the base $b$
+representation of the free variables (that are indices of the sequence). We
+briefly explain in the following lemma the folklore result that one can decide
+if an \kl{automatic sequence} is \kl{ultimately uniformly recurrent}.
+
+\begin{lemma}
+  \label{automatic-uur:lem}
     Given an \kl{automatic sequence} $w \in \Sigma^{\Nat}$, one can decide
-    whether the set of \kl{infixes} of $w$ is \kl{well-quasi-ordered} for the
-    \kl{infix relation}.
-\end{theorem}
+    whether it is \kl{ultimately uniformly recurrent}.
+\end{lemma}
 \begin{proof}
-    Because of \cref{ultimately-uniformly-recurrent:lem}, it suffices
-    to decide whether $w$ is \kl{ultimately uniformly recurrent}.
     We can rewrite this as a question on the \kl{automatic sequence} $w$
@@
 
-Note that being \kl{uniformly recurrent} still leaves open a lot of potential
-behaviours, for instance, Sturmian words are \kl{uniformly recurrent}, but are
-not \kl(sequence){automatic}. Interestingly enough, the \kl{ordinal invariants}
-of the \kl{automatic sequences} are not smaller than the ones of \kl{ultimately
-uniformly recurrent} words: this contrasts with the case of languages that are
-\kl{well-quasi-ordered} for the \kl{infix relation} in general, where the
-hypothesis of being representable (by a \kl{regular language}, or an
-\kl{amalgamation system}) will drastically reduce the \kl{ordinal invariants}
-of said language (\cref{ordinal-invariants-bounded:cor}, \cref{infix-amalgamation:thm}).
+Based on \cref{automatic-uur:lem}, we can now prove the following theorem.
+\begin{theorem}
+    \label{automatic-wqo:thm}
+    Given a \kl{sequence representation} of a language $L \subseteq
+    \Sigma^*$ where all infinite words are \kl{automatic sequences}, one can
+    decide whether $L$ is \kl{well-quasi-ordered} for the \kl{infix relation}.
+\end{theorem}
+\begin{proof}
+    It is easy to see that $L$ is \kl{well-quasi-ordered} for the \kl{infix
+    relation} if and only if for every triple $(w_i^-, a_i, w_i^+)$ in the
+    \kl{sequence representation} of $L$, the (potentially bi-infinite) word
+    $\mathsf{reversed}(w_i^-) a_i w_i^+$ defines a \kl{well-quasi-ordered}
+    language. By \cref{from-bi-to-single:lem}, this is the case if and only if
+    both $w_i^-$ and $w_i^+$ are \kl{ultimately uniformly recurrent}. 
+    Since one can decide whether an \kl{automatic sequence} is
+    \kl{ultimately uniformly recurrent} using \cref{automatic-uur:lem},
+    we conclude the proof.
+\end{proof}
+
+\AP In fact, \kl{automatic sequences} are part of a larger family of sequences
+studied in symbolic dynamics, called \kl{morphic sequences}. Let us first
+recall that a \intro(words){morphism} is a function $f \colon \Sigma^* \to
+\Gamma^*$ such that for every $u,v \in \Sigma^*$, $f(uv) = f(u)f(v)$. A
+\intro{morphic sequence} $w$ is an infinite word obtained by iterating a
+\kl(words){morphism} $f \colon \Sigma^* \to \Sigma^*$ on a letter $a \in
+\Sigma$ such that $f(a)$ starts with $a$, and then applying a
+\kl(words){homomorphism} $h \colon \Sigma^* \to \Gamma^*$. The infinite word
+$f^\omega(a)$ is the limit of the sequence $\seqof{f^n(a)}[n \in \Nat]$, which
+is well-defined because $f(a)$ starts with $a$, and the \kl{morphic sequence}
+is $ w \defined h(f^\omega(a))$.
+
+\AP
+Every \kl{automatic
+sequence} is a \kl{morphic sequence}, but not the other way around. We refer
+the reader to a short survey of \cite{ALSZ17} for more details on the possible
+variations on the definition of \kl{morphic sequences} and their relationships.
+It was relatively recently proven that one can decide whether a \kl{morphic
+sequence} is \kl{uniformly recurrent} \cite[Theorem 1]{DURAND13}. We were not
+able to find in the literature whether one can decide \kl{ultimate uniform
+recurrence}, but we can prove it for a restricted class of \kl{morphic
+sequences}, that are called \intro{pure morphic sequences}, and are precisely
+those that do not use a final \kl(words){homomorphism} $h$. Following the same proof
+as for \cref{automatic-wqo:thm}, one deduces from the upcoming
+\cref{morphic-uur:lem} the following \cref{morphic-wqo:cor}. We actually
+conjecture that \cref{morphic-uur:lem} can be extended to all \kl{morphic
+sequences} \cref{morphic-uur:conj}.
+
+\begin{lemma}
+  \label{morphic-uur:lem}
+    Given a \kl{pure morphic sequence} $w \in \Sigma^{\Nat}$, one can decide
+    whether it is \kl{ultimately uniformly recurrent}.
+\end{lemma}
+\begin{proof}
+  On the one hand, one can guess a bound $N_0$ and check that 
+  the suffix of the word $w_{\geq N_0}$ is \kl{uniformly recurrent},
+  since the suffix of a \kl{morphic sequence} is also a \kl{morphic sequence},
+  and \kl{uniform recurrence} is decidable for \kl{morphic sequences}
+  \cite[Theorem 2]{DURAND13}.
+
+  Let us now prove that a specific choice of $N_0$ is sufficient. Let us write
+  $w \defined f^\omega(a)$ where $f \colon \Sigma^* \to \Sigma^*$ is a
+  \kl(words){morphism} and $a \in \Sigma$ is a letter such that $f(a)$ starts
+  with $a$. Let us also write $f(a) = a s$, where $s \in \Sigma^+$ is a
+  non-empty word. Notice that $w = a s f(s) f^2(s) \cdots$. Without loss of
+  generality, one can assume that $|f^n(s)| \to \infty$ when $n \to \infty$,
+  otherwise $w$ is ultimately periodic, and therefore \kl{ultimately uniformly
+  recurrent}. Since one can algorithmically detect if this is the case, we
+  assume that it is not.
+
+  Because $\Sigma$ is finite, there exists a finite set of letters $A \subseteq
+  \Sigma$ and a suffix $w'$ of $w$ such that $w' \in A^*$ and every letter of
+  $w'$ appears infinitely often in $w'$. Furthermore, one can compute this set
+  $A$ and a bound $N_0$ such that $f^n(s) \in A^*$ for every $n \geq N_0$. We
+  claim that $w$ is \kl{ultimately uniformly recurrent} if and only if
+  $f^{N_0}(s) f^{N_0 + 1}(s) \cdots$ is \kl{uniformly recurrent}.
+  One direction being obvious, we assume that $w$ is \kl{ultimately uniformly
+  recurrent}, and let us prove that $f^{N_0}(s) f^{N_0 + 1}(s) \cdots$ is
+  \kl{uniformly recurrent}.
+
+  Let us assume that $w'' \defined f^{N_1}(s) f^{N_1 + 1}(s) \cdots$ where $N_1
+  \geq N_0$ is \kl{uniformly recurrent}. Let us show that $f^{N_0}(s)$ is an
+  \kl{infix} of $w''$ if and only if there exists a letter $b$ in $s$ such that
+  $f^{N_0}(s) \infleq f^{k} (b)$ for some finite number $k$.
+
+  Assume that $f^{N_0}(s)$ is an \kl{infix} of $w''$. Recall that $|f^n(s)| \to
+  \infty$ when $n \to \infty$, hence, there must exist a letter $b$ in $s$ such
+  that $|f^n(b)| \to \infty$ when $n \to \infty$. Because every letter of
+  $f^{N_0}(s)$ appears infinitely often in $w''$, $f^n(b)$ is a growing
+  sequence of infixes of $w''$. Since $w''$ is \kl{uniformly recurrent}, there
+  exists a bound $K$ such that $f^{N_0}(s)$ is an \kl{infix} of every $f^k(b)$
+  for $k \geq K$. Conversely, if there exists a letter $b$ in $s$ such that
+  $f^{N_0}(s) \infleq f^{k} (b)$ for some finite number $k$, then $f^{N_0}(s)$
+  is an \kl{infix} of $w''$ because $f^n(b)$ is an infix of $w''$ for every $n
+  \in \Nat$.
+
+  Because $w''$ is \kl{uniformly recurrent}, there exists a bound $K$ such that
+  $f^{N_1}(s) \infleq f^k(b)$ for every $k \geq K$.
+  \todo[inline]{Cannot finish the proof...}
+
+\end{proof}
+
+\begin{corollary}
+    \label{morphic-wqo:cor}
+    Given a \kl{sequence representation} of a language $L \subseteq
+    \Sigma^*$ where all infinite words are \kl{pure morphic sequences}, one can
+    decide whether $L$ is \kl{well-quasi-ordered} for the \kl{infix relation}.
+\end{corollary}
+
+\begin{conjecture}
+  \label{morphic-uur:conj}
+    Given a \kl{morphic sequence} $w \in \Sigma^{\Nat}$, one can decide
+    whether it is \kl{ultimately uniformly recurrent}.
+\end{conjecture}

### 2025-08-06 src/prelims.tex
minor

@@
 
-\paragraph*{Finite words.}
-\AP 
-In this paper, we use letters $\Sigma, \Gamma$ to
-denote finite alphabets, $\Sigma^*$ to denote the set of finite words over
-$\Sigma$, and $\varepsilon$ for the empty word in $\Sigma^*$. In order to give
-some intuition on the decision problems, we will sometimes use the notion of
-\intro{finite automata}, \intro{regular languages}, and Monadic Second Order
-logic ($\intro*\MSO$) over finite words, and assume is familiar with them. We
-refer to the textbook of \cite{THOM97} for a detailed introduction to this
-topic.
+\paragraph*{Finite words.} \AP In this paper, we use upper Greek letters
+$\Sigma, \Gamma$ to denote finite alphabets, $\Sigma^*$ to denote the set of
+finite words over $\Sigma$, and $\varepsilon$ for the empty word in $\Sigma^*$.
+In order to give some intuition on the decision problems, we will sometimes use
+the notion of \intro{finite automata}, \intro{regular languages}, and Monadic
+Second Order logic ($\intro*\MSO$) over finite words, and assume the reader to
+be familiar with them. We refer to the textbook of \cite{THOM97} for a detailed
+introduction. However, we will require no prior knowledge on word
+combinatorics.
 
@@
 parameters are called the \intro{ordinal invariants} of a
-\kl{well-quasi-ordered} set $X$. We refer to the survey of \cite{DZSCSC20} for
+\kl{well-quasi-ordered} set $X$. 
+As an example, for $(\Nat, \leq)$, all bad sequences are descending and there are no antichains.
+Therefore the maximal order type and the ordinal height of $\Nat$ are $\omega$ and the ordinal width is 1.
+\todo{Please check this!}
+We refer to the survey of \cite{DZSCSC20} for
 a detail discussion on these concepts and their computation on specific classes

### 2025-08-06 src/prefixes.tex
removing todos + polish
more notes

@@
 prefix relation is well-quasi-ordered, without any restriction on the
-decidability of $L$ itself. Let us remark that finite unions of \kl{chains} are
-always \kl{well-quasi-ordered} by the \kl{prefix relation} because they lack
-infinite \kl{antichains} by definition. The following theorem states that this
-is the only possible reason for a language $L$ to be \kl{well-quasi-ordered} by
-the \kl{prefix relation}.
-
+decidability of $L$ itself.
 For the proof of the following theorem, we will use special notations to
@@
     \end{equation*}
+    \todo[inline]{Because $S$ only contains words with two incomparable suffixes in $L$, I think this is wrong - consider e.g. $L \subseteq a^*$, where $S$ and $S_{max}$ are necessarily empty}
     Hence, that $L$ is finite union of \kl{chains}.
@@
     the \kl{ordinal height} of $L$ is at most $\omega$, and
-    its \kl{ordinal width} is finite.
+    its \kl{ordinal width} is finite. Furthermore, these bounds 
+    are tight.
 \end{corollary}
-
-\todo[inline]{
-    We actually can \emph{compute} the ordinal invariants
-    in the regular case.
-}
+\begin{proof}
+    The upper bounds follow from the fact that $L$ is a finite union of
+    \kl{chains}. The tightness can be obtained by considering the languages
+    $L_k \defined \bigcup_{i=0}^{k-1} a^i b^*$ for $k \in \Nat$, which 
+    are \kl{well-quasi-ordered} by the \kl{prefix relation} (as they are
+    finite unions of \kl{chains}), and satisfy
+    that $\oWidth{L_k} = k$, $\oHeight{L_k} = \omega$,
+    and therefore $\oType{L_k} = k \cdot \omega$.
+\end{proof}

### 2025-08-07 src/infixes.tex
more notes
new: add todos

--- 2025-03-24 src/infixes.tex
@@
 of said language (\cref{ordinal-invariants-bounded:cor}, \cref{infix-amalgamation:thm}).
+
+\todo[inline]{For Nathan: 
+  talk about morphic words / sequences and 
+  decidability here.
+  \url{https://arxiv.org/pdf/1204.5393}
+}
+
+\todo[inline]{Talk about factor complexity and 
+well-quasi-orderings: the two are unrelated. This can be a remark.
+\url{https://arxiv.org/pdf/2206.00376}
+Here example 5 p8 is an explicit example of a uniformly recurrent word with exponential factor complexity.
+}

### 2025-08-07 src/introduction.tex
new: add todos

@@
 
+\todo[inline]{
+  Add a figure to recap our complexity results: 
+  bounded / downwards closed / amalgamative / morphic | width.
+  And only talk about WQO classes, collapse everything and say regular.
+}
+
 \subparagraph{Related work} The study of alternative \kl{well-quasi-ordered}
@@
 the \kl{prefix}, \kl{suffix}, and \kl{infix} relations.
+
+A similar question was studied by Atminas, Lozin, and Moshkov \cite{ALM17}, in
+the hope of finding characterizations of classes of \emph{finite graphs} that
+are \kl{well-quasi-ordered} by the \emph{induced subgraph relation}
+\cite[Section 7]{ALM17}. In this setting, it is common to refer to classes of
+graphs via a list of \emph{forbidden patterns}, which are finite graphs that
+cannot be found as induced subgraphs in the class. Applying this reasoning to
+finite words with the \kl{infix relation}, they provide an efficient decision
+procedure for checking whether a language $L \subseteq \Sigma^*$ is
+\kl{well-quasi-ordered} by the \kl{infix relation} whenever said language is
+given as input via a list of \emph{forbidden factors} \cite[Theorem 1, Theorem
+2]{ALM17}. The key construction of their paper is to study languages $L$ that
+are \emph{regular} (recognized by some finite deterministic automata), for
+which they can decide whether $L$ is \kl{well-quasi-ordered} by the \kl{infix
+relation} \cite[Theorem 1]{ALM17}. Because it is easy to transform a list of
+forbidden factors into a regular language \cite[Theorem 1]{ALM17}, this yields
+the desired decision procedure. Our work extends this result in several ways:
+first, we also consider the \kl{prefix relation} and the \kl{suffix relation},
+then we consider non-regular languages, and finally, we provide very precise
+descriptions of the \kl{well-quasi-ordered} languages, as well as tight bounds
+on their \kl{ordinal invariants}. In particular, we highlight the role of
+\kl{bounded languages} as the one that are well-behaved with respect to being
+\kl{well-quasi-ordering} by the \kl{infix relation} \cref{infixes-bounded:sec}.
+These non-trivial extensions continue to rely
+on classical word combinatorial techniques that are already present in the work
+of Atminas, Lozin, and Moshkov \cite[Section 3]{ALM17}.
 

### 2025-08-07 wqo-on-words.tex
starting to write a bit
import mathematics
rewriting introduction
writing proofs
typos
finish proofs
work on intro/contribution (#1)
* work on intro/contribution

* Subfigures for intro graphics

---------

Co-authored-by: lschuetze <lschuetze@mpi-sws.org>
Add linenumbers
restructuring paper
Big Mic Mac (#6)
- new introduction 
- removing automata based approach
- add the lexicographic ordering in conclusion 
- splitting `infixes` into `bounded` and `amalgamation systems`
- new bibliographic entries + preliminaries
- add the counter-example using the Thue-Morse sequence
- update figures to add

@@
-\documentclass[11pt]{article}
+%! TEX program = pdflatex
+% WARNING: this is a generated file.
+%
+% Please do not edit this file directly. 
+% - If you want to update the medatata of the paper (title, authors, abstract), please
+%   edit the `paper-meta.yaml` file in the root of the repository.
+% - If you want to update the content of the paper, please edit the latex files
+%   in the `src` directory.
+% - If you want to update the template itself (e.g., change the layout), please
+%   edit the `templates/plain-article.tex` file instead.
+\documentclass[11pt,a4paper,twosided]{article}
 
-\input{lib/packages}
+% we setup a custom geometry because the default one is too narrow
+\usepackage{geometry}
+\geometry{margin=3.5cm}
+
+% utf-8 for old systems
+\usepackage[utf8]{inputenc}
+\usepackage[T1]{fontenc}
+
+% babel for language settings
+\usepackage[english]{babel}
+
+% microtype for better typography
+\usepackage{microtype}
+
+\usepackage{todonotes}
+\usepackage{lineno}
+\linenumbers
+
+% math packages
+\usepackage{amsmath,amsthm,amssymb,stmaryrd,thmtools,upgreek}
+
+% configure some theorems
+\newtheorem{theorem}{Theorem}
+\newtheorem{lemma}[theorem]{Lemma}
+\newtheorem{corollary}[theorem]{Corollary}
+\newtheorem{proposition}[theorem]{Proposition}
+\newtheorem{conjecture}[theorem]{Conjecture}
+\newtheorem{assumption}[theorem]{Assumption}
+\theoremstyle{definition}
+\newtheorem{definition}[theorem]{Definition}
+\newtheorem{remark}[theorem]{Remark}
+\newtheorem{example}[theorem]{Example}
+
+% graphics packages
+\usepackage{graphicx}
+\usepackage{subcaption}
+\usepackage[obeyclassoptions,mode=tex]{standalone}
+\usepackage{tikz}
+\usetikzlibrary{backgrounds}
+\usetikzlibrary{shapes.geometric}
+\usetikzlibrary{positioning}
+\usetikzlibrary{automata}
+\usetikzlibrary{tikzmark}
+\usetikzlibrary{patterns}
+\usetikzlibrary{arrows}
+\tikzset{every state/.style={minimum size=1pt}}
+\usepackage{tikz-cd}
+
+% ornaments
+\usepackage{pgfornament}
+
+% links inside the document
+\usepackage{hyperref}
+\usepackage[capitalise,noabbrev,nameinlink]{cleveref}
+\Crefname{assumption}{Assumption}{Assumptions}
+\usepackage[composition,hyperref,xcolor,cleveref]{knowledge}
+\knowledgeconfigure{notion}
+
+% Tables 
+\usepackage{booktabs}
+\usepackage{varwidth}
+
+% Algorithms
+\usepackage{algorithm2e}
+\Crefname{algocfline}{Algorithm}{Algorithms}
+\crefname{algocfline}{Algorithm}{Algorithms}
+\crefname{algocf}{Algorithm}{Algorithms}
+\Crefname{algocf}{Algorithm}{Algorithms}
+
+% Packages for macro definitions
+\usepackage{xparse}
+\usepackage{xpatch}
+\usepackage{tokcycle}
+\usepackage{ifthen}
+
+% Proofs
+\usepackage{bussproofs}
+
+% Colors 
+\usepackage{ensps-colorscheme}
+% parametrize knowledge colors and style
+% intro => A4 + emph 
+% kl    => normal color, normal size not emph
+\knowledgestyle{intro notion}{color={A2}, emphasize}
+\knowledgestyle{notion}{color={B1}}
+\hypersetup{
+    colorlinks=true,
+    breaklinks=true,
+    linkcolor=A4,
+    citecolor=A4,
+    urlcolor=A4,
+    filecolor=A4,
+}
+
+% we include whatever the user wants to include in the header
+
+% we include libraries (tex files) usually written in the `lib` directory
 \input{lib/aliaume}
+\input{lib/maths}
+\input{knowledges.kl}
 
-\input{src/metadata.tex}
+% We include the title and author information based on the 
+% `paper-meta.yaml` file.
 
+\title{Well-quasi-orderings on word languages}
+
+\author{
+Nathan Lhote\thanks{Aix-Marseille University}
+ \and
+Aliaume Lopez\thanks{University of Warsaw}
+ \and
+Lia Schütze\thanks{Max Planck Institute for Software Systems}
+}
+
+% For the date, we first check if the user has provided a date,
+% and otherwise use the git meta inforamtion (if available).
+\date{2025-07-01 13:26:43 +0200\footnote{99064007a226e55157ecfac20bbb40804687f3e2 -- branch main at git@github.com:AliaumeL/word-ordering-wqo.git}}
+
+\newcommand{\repositoryUrl}{\url{https://github.com/AliaumeL/word-ordering-wqo}}
+
+\knowledge{notion}
+ | kl-usage
+
+% Now, we create the document itself.
 \begin{document}
+% Generate the title page
 \maketitle
+% Print the abstract
+\begin{abstract}
+    The set of finite words over a well-quasi-ordered set is itself well-quasi-ordered.
+This seminal result by Higman is a cornerstone of the theory of well-quasi-orderings and has found numerous applications in computer science.
+However, this result is based on a specific choice of ordering on words, the (scattered) subword ordering.
+In this paper, we describe to what extent other natural orderings (prefix, suffix, and infix) on words can be used to derive Higman-like theorems.
+More specifically, we are interested in characterizing \emph{languages} of words that are well-quasi-ordered under these orderings.
+We show that a simple characterization is possible for the prefix and suffix orderings, and that under extra regularity assumptions, this also extends to the infix ordering.
+We furthermore provide decision procedures for a large class of languages, that contains regular and context-free languages.
+    \paragraph{Keywords:}
+    equivariant ideal, Hilbert basis, ideal membership problem, orbit finite, oligomorphic, well-quasi-ordering
+\paragraph{Repository:} \repositoryUrl
+\end{abstract}
 
+\klogo\ This document uses \href{https://ctan.org/pkg/knowledge}{knowledge}:
+\kl[kl-usage]{notion} points to its \intro[kl-usage]{definition}.
+
+% Include the content of the paper
 \input{src/introduction}
+\input{src/prelims.tex}
+\input{src/prefixes}
+\input{src/infixes}
+\input{src/amalgam}
+\input{src/conclusion}
+
+% Include the bibliography
+\bibliographystyle{plainurl}
+\bibliography{papers.bib}
+
+% If there are any appendices, we include them here.
 

### 2025-08-07 fig/periodic-infixed.tex
Add figure to periodic infixes

\documentclass[tikz]{standalone}

\usepackage{tikz}
\usepackage{ensps-colorscheme}
%\usepackage{ifthen}

\usepackage{subcaption}

\pgfdeclarelayer{background}
\pgfsetlayers{background,main}

\begin{document}
	\begin{tikzpicture}[scale=0.9,every node/.style={scale=0.9}]

		\foreach \i in {1,...,11} {
			\pgfmathsetmacro{\x}{0.6 * \i}
			\node [draw,rectangle,minimum size=.6cm] (v\i) at (\x cm,0) {};
			\node [draw,rectangle,minimum size=.6cm] (u\i) at (\x cm,-1.5) {};
		}

		\node at (8cm,0) {$v$};
		\node at (8cm,-1.5) {$u$};

		\draw [dashed] (v1.north west) -- ++(-.5,0);
		\draw [dashed] (v1.south west) -- ++(-.5,0);		
		\draw [dashed] (v11.north east) -- ++(.5,0);
		\draw [dashed] (v11.south east) -- ++(.5,0);

		\path [color=D3,-{latex}] (u1) edge (v1);
		\path [color=D3,-{latex}] (u5) edge (v5);
		\path [color=D3,-{latex}] (u9) edge (v9);

		\path [color=D4,-{latex}] (u1.south) edge [bend right] (u5.south);
		\path [color=D4,-{latex}] (u5.south) edge [bend right] (u9.south);
		%\path [color=C5,-{latex}] (v9.north) edge [bend right] (v3.north);

		\begin{pgfonlayer}{background}
			\node [fill,color=D4bg,minimum size=.6cm] at (u1) {};
			\node [fill,color=D4bg,minimum size=.6cm] at (u5) {};
			\node [fill,color=D4bg,minimum size=.6cm] at (u9) {};
			\node [fill,color=D4hint,minimum size=.6cm] at (v1) {};
			\node [fill,color=D4hint,minimum size=.6cm] at (v5) {};
			\node [fill,color=D4hint,minimum size=.6cm] at (v9) {};	
		\end{pgfonlayer}

		\begin{scope}[shift={(8.8cm,0)}]
			\foreach \i in {1,...,11} {
				\pgfmathsetmacro{\x}{0.6 * \i}
				\node [draw,rectangle,minimum size=.6cm] (v\i) at (\x cm,0) {};
				\node [draw,rectangle,minimum size=.6cm] (u\i) at (\x cm,-1.5) {};
			}

			\draw [dashed] (v1.north west) -- ++(-.5,0);
			\draw [dashed] (v1.south west) -- ++(-.5,0);		
			\draw [dashed] (v11.north east) -- ++(.5,0);
			\draw [dashed] (v11.south east) -- ++(.5,0);

			\path [color=D3,-{latex}] (u1) edge (v1);
			\path [color=D3,-{latex}] (u5) edge (v5);
			\path [color=D3,-{latex}] (u9) edge (v9);

			\path [color=D3,-{latex}] (v3) edge (u3);
			\path [color=D3,-{latex}] (u7) edge (v7);
			\path [color=D3,-{latex}] (u11) edge (v11);

			\path [color=D4,-{latex}] (u1.south) edge [bend right] (u5.south);
			\path [color=D4,-{latex}] (u5.south) edge [bend right] (u9.south);

			\path [color=C5,-{latex}] (v9.north) edge [bend right] (v3.north);

			\path [color=D4,-{latex}] (u3.south) edge [bend right] (u7.south);
			\path [color=D4,-{latex}] (u7.south) edge [bend right] (u11.south);

			\begin{pgfonlayer}{background}
				\node [fill,color=D4bg,minimum size=.6cm] at (u1) {};
				\node [fill,color=D4bg,minimum size=.6cm] at (u5) {};
				\node [fill,color=D4bg,minimum size=.6cm] at (u9) {};
				\node [fill,color=D4hint,minimum size=.6cm] at (v1) {};
				\node [fill,color=D4hint,minimum size=.6cm] at (v5) {};
				\node [fill,color=D4hint,minimum size=.6cm] at (v9) {};

				\node [fill,color=D4bg,minimum size=.6cm] at (v3) {};
				\node [fill,color=D4hint,minimum size=.6cm] at (u3) {};
				\node [fill,color=D4hint,minimum size=.6cm] at (v7) {};
				\node [fill,color=D4hint,minimum size=.6cm] at (u7) {};
				\node [fill,color=D4hint,minimum size=.6cm] at (v11) {};
				\node [fill,color=D4hint,minimum size=.6cm] at (u11) {};
			\end{pgfonlayer}
		\end{scope}

	\end{tikzpicture}
\end{document}

### 2025-08-07 src/infixes.tex
Add figure to periodic infixes

--- 2025-07-31 src/infixes.tex
@@
 \kl{well-quasi-ordered} for the \kl{infix relation}. This proves the (easy)
-right-to-left implication of \cref{bounded-language:thm}.
+right-to-left implication of \cref{bounded-language:thm}. Furthermore, any
+language $L$ included in a finite union of products $S_i \cdot P_i$
+is in fact a \kl{bounded language}. \todo[inline]{I think we can in fact construct chains that are not bounded languages. E.g. $ab, aba^2b, aba^2ba^3b, \ldots$?}
 
@@
 exists a $p \in \Nat$ such that $w \infleq x^p$. The \intro{periodic length} of
-a word $u$ is the minimal length of a word $x$ such that $u$ is an \kl{infix}
-of $x^p$ for some $p \in \Nat$ and $x \in \Sigma^+$.
+a word $u$ is the minimal length of a period $x$ of $u$.
 
@@
     having \kl{periodic lengths} $p$ and $q$ respectively.
-    Then, if $u \infleq v$ and $\card{u} \geq \factorial[p]{p \times q}$,
+    Then, if $u \infleq v$ and $\card{u} \geq pq$,
     then $u$ and $v$ share the same \kl{periodic length}
@@
     particular, $v_{k+i+p} = v_{k+i}$ for all $i \in \Nat$ such that $k+i+p < k
-    + \card{u}$. Since we also have $v_{k+i+q} = v_{k+i}$ for all $1 \leq i
-    \leq \card{v} - k - q$. We conclude that both $u$ and $v$ are of
+    + \card{u}$. Due to the periodicity of $v$, we also have $v_{k+i+q} = v_{k+i}$ for all $i$ with $k + i + q \leq k + \card{u}$. We conclude that both $u$ and $v$ are of
     \kl{periodic length} the greatest common divisor of $p$ and $q$, and by
     minimality of $q$ this must be equal to $q$ and to $p$.
+    \todo{maybe elaborate on why $pq$ as bound on $\card{u}$ (also technically $2\max\{p,q\}$ would be enough, but maybe more cumbersome?}
 \end{proof}
+
+\begin{figure}
+	\begin{center}
+		\input{fig/periodic-infixed.tex}
+	\end{center}
+	\caption{Alignment of two periodic words with periods of length $6$ and $4$, respectively. We can conclude that the periodic length is at most $2$.}
+	\label{fig:periodic-infixes}
+\end{figure}
 

### 2025-08-12 src/amalgam.tex
minor changes + funding
fix: making llncs work

@@
 \kl(language){bounded}. This implies that all the results of
-\cref{infixes-regular:sec}, and in
+\cref{infixes-bounded:sec}, and in
 particular \cref{bounded-language:thm}, can safely be applied to
@@
 
-\begin{theorem}[label=infix-amalgamation:thm,restate=infix-amalgamation:thm]
+\begin{theorem}
+    \label{infix-amalgamation:thm}
     \proofref{infix-amalgamation:thm}
@@
 
-\begin{theorem}[restate=infix-wqo-is-emptiness:thm,label=infix-wqo-is-emptiness:thm]
-% 	\label{infix-wqo-is-emptiness:thm}
+\begin{theorem}
+	\label{infix-wqo-is-emptiness:thm}
 	Let $\mathcal{C}$ be an \kl{effective amalgamative class} of languages.

### 2025-08-12 src/conclusion.tex
new: add todos
fix: making llncs work

@@
 
-\subparagraph{Towards infinite alphabets} In this paper, we restricted our
+\paragraph*{Towards infinite alphabets} In this paper, we restricted our
 attention to \emph{finite} alphabets, having in mind the application to
@@
 
-\subparagraph{Monoid equations}  It could be interesting to understand which
+\paragraph*{Monoid equations}  It could be interesting to understand which
 monoids $M$ recognize languages that are \kl{well-quasi-ordered} by the
@@
 
-\subparagraph{Complexity} We have chosen to disregard complexity considerations
+\paragraph*{Complexity} We have chosen to disregard complexity considerations
 when proving decidability results, as we do not believe that a fined grained
@@
 
-\subparagraph{Lexicographic orderings} There is another natural ordering on
+\paragraph*{Lexicographic orderings} There is another natural ordering on
 words, the \emph{lexicographic ordering}, which does not fit well in our
@@
 \kl{amalgamation systems}.
+
+\paragraph*{Graph Classes}
+\todo[inline]{aliaume: retalk about \cite{ALM17} and \cite{L24:arxiv:v2}}

### 2025-08-12 src/introduction.tex
fix: making llncs work

@@
 
-\subparagraph{Contribution} In this paper, we focus on words over a finite
+\paragraph*{Contribution} In this paper, we focus on words over a finite
 alphabet $\Sigma$, and characterize subsets $L \subseteq \Sigma^*$ that are
@@
 
-\subparagraph{Related work} The study of alternative \kl{well-quasi-ordered}
+\paragraph*{Related work} The study of alternative \kl{well-quasi-ordered}
 relations over finite words is far from new. For instance, orders obtained by
@@
 
-\subparagraph{Outline} 
+\paragraph*{Outline} 
 We introduce in \cref{prelims:sec} the
@@
 
-\subparagraph{Acknowledgements} We would like to thank participants of the 2024
+\paragraph*{Acknowledgements} We would like to thank participants of the 2024
 edition of \kl{Autobóz} for their helpful comments and discussions.

### 2025-09-05 src/introduction.tex
new: cut intro!

@@
 context of logic~\cite{DBLP:journals/pacmpl/BergstrasserGLZ24}.
+In the graph theory community, recent works have studied classes of graphs 
+that are \kl{well-quasi-ordered} by the \kl{induced subgraph relation}
+using similar language theoretic techniques~\cite{DRT10,L24:arxiv:v2,ALM17}.
 Furthermore, a previous work by Kuske shows that any
@@
 
-\AP When considering an algorithm based on a \kl{well-quasi-ordering}, the
-runtime of the algorithm is deeply related to the ``complexity'' of the
-underlying quasi-order~\cite{SCHMITZ17}. One way to measure this complexity is
-to consider its so-called \kl{ordinal invariants}: for instance, the
-\kl{maximal order type} (or \kl{m.o.t.}), originally defined by De Jongh and Parikh
-\cite{dejongh77}, is the order type of the maximal linearization of a
-well-quasi-ordered set. In the case of a finite set, the \kl{m.o.t.} is precisely
-the size of the set. Better runtime bounds were obtained by considering two
-other parameters~\cite{SCHMITZ19}: the \kl{ordinal height} introduced by
-Schmidt \cite{schmidt81}, and the \kl{ordinal width} of Kříž and Thomas
-\cite{kriz90b}. Therefore, when characterizing \kl{well-quasi-ordered}
-languages, we will also be interested in deriving upper bounds on their
-\kl{ordinal invariants}. We refer to \cref{ordinal-invariants:subsec} for a
-more detailed introduction to these parameters and ordinal computations in
-general.
+\AP The runtime of an algorithm based on \kl{well-quasi-orderings} is deeply
+related to the ``complexity'' of the underlying quasi-order~\cite{SCHMITZ17}.
+One way to measure this complexity is to consider its so-called \kl{ordinal
+invariants}: for instance, the \kl{maximal order type} (or \kl{m.o.t.}),
+originally defined by De Jongh and Parikh \cite{dejongh77}, is the order type
+of the maximal linearization of a well-quasi-ordered set. In the case of a
+finite set, the \kl{m.o.t.} is precisely the size of the set. Better runtime
+bounds were obtained by considering two other parameters~\cite{SCHMITZ19}: the
+\kl{ordinal height} introduced by Schmidt \cite{schmidt81}, and the \kl{ordinal
+width} of Kříž and Thomas \cite{kriz90b}. Therefore, when characterizing
+\kl{well-quasi-ordered} languages, we will also be interested in deriving upper
+bounds on their \kl{ordinal invariants}. This analysis also allows us to better
+compare the \kl{well-quasi-orderings}. We refer to
+\cref{ordinal-invariants:subsec} for a more detailed introduction to these
+parameters and ordinal computations in general.
 
-\paragraph*{Contribution} In this paper, we focus on words over a finite
-alphabet $\Sigma$, and characterize subsets $L \subseteq \Sigma^*$ that are
-\kl{well-quasi-ordered} by the \kl{prefix relation}, the \kl{suffix relation},
-and the \kl{infix relation}. Furthermore, we devise decision algorithms
-whenever the languages are given by reasonable computational models. Finally,
-we provide precise bounds on the possible \kl{ordinal invariants} of such
-languages.
+\paragraph*{Contributions} We focus on languages over a finite alphabet
+$\Sigma$. In this setting, we first characterize languages that are
+\kl{well-quasi-ordered} by the \kl{prefix relation} (and symmetrically, by the
+\kl{suffix relation}), and derive tight bounds on their \kl{ordinal
+invariants}. These generic results are then used to devise a decision procedure
+for checking whether a language is \kl{well-quasi-ordered} by the \kl{prefix
+relation}, provided the language is given as input as a finite automaton
+(\cref{prefix-wqo-reg-decidable:cor}). A
+summary of these results can be found in \cref{prefixes-summary:fig}.
 
-In the case of the \kl{prefix} and \kl{suffix} relations, we show that a
-language $L \subseteq \Sigma^*$ is \kl{well-quasi-ordered} by the \kl{prefix
-relation} (resp. \kl{suffix}) if and only if it is a finite union of
-\kl{chains} of the \kl{prefix relation} (resp. \kl{prefix}), where a
-\intro{chain} is a totally ordered set that is well-founded. Note that
-\kl{chains} are the simplest possible \kl{well-quasi-ordered} sets (they are
-\kl{totally ordered} and \kl{well-founded}) and it is known that finite unions
-of \kl{well-quasi-ordered} sets are \kl{well-quasi-ordered}. As a consequence,
-the above characterization states that only the simplest possible
-\kl{well-quasi-ordered} sets are \kl{well-quasi-ordered} by the \kl{prefix
-relation} (\cref{prefixes:thm}). Let us furthermore highlight that this
-characterization holds without any restriction on the decidability of the
-language $L$ itself, but heavily relies on the assumption that $\Sigma$ is
-finite. This allows us to derive tight bounds on the \kl{ordinal invariants} of
-such \kl{well-quasi-ordered} languages, which can be interpreted in two dual
-ways: first, these languages are extremely simple, which means that one could
-have hoped for using directly another well-quasi-ordered set without resorting
-to finite words; second, any time one encounters a \kl{well-quasi-ordered}
-language, it is going to provide relatively efficient algorithms. Indeed, we
-prove that all languages $L \subseteq \Sigma^*$ are \kl{well-quasi-ordered} by
-the \kl{prefix relation} (resp. \kl{suffix relation}) have an \kl{ordinal
-height} of at most $\omegaOrd$, a finite \kl{ordinal width}, and a \kl{maximal
-order type} strictly below $\omegaOrd^2$
-(\cref{prefixes-ordinal-invariants:cor}).
+\begin{figure}[h]
+    \centering
+    \begin{tabular}{c|c|c|c}
+        \toprule
+        $L$ & \textbf{Characterisation} & $\oWidth{L}$ & $\oType{L}$ \\
+        \midrule
+        arbitrary & Finite unions of \kl{chains} \cref{prefixes:thm} & $< \omegaOrd$ & $< \omegaOrd^2$ \\
+        regular   & Finite unions of \kl{regular} \kl{chains} \cref{prefix-wqo-reg-decidable:cor} & $< \omegaOrd$ & $< \omegaOrd^2$ \\
+        \bottomrule
+    \end{tabular}
+    \caption{Summary of results for the \kl{prefix relation} (and symmetrically, for the \kl{suffix relation}).}
+    \label{prefixes-summary:fig}
+\end{figure}
 
-\AP The straightforward generalization of the results for the \kl{prefix
-relation} and the \kl{suffix relation} to the \kl{infix relation} is not
-possible. Indeed, it follows from Kuske's result that $\Sigma^*$ equipped with
-the \kl{subword relation} can be embedded into $\set{a,b}^*$ with the \kl{infix
-relation}~\cite[Lemma 5.1]{DBLP:journals/ita/Kuske06}. This implies that there
-are \kl{well-quasi-ordered} languages for the \kl{infix relation} that have
-very large \kl{ordinal invariants}: for instance, the \kl{maximal order type}
-of the \kl{subword relation} is $\omegaOrd^{\omegaOrd^{\card{\Sigma} - 1}}$,
-which equals its \kl{ordinal width} \cite[Corollary 3.9, Theorem
-4.21]{DZSCSC20}. We show that in two situations, this can be avoided: when the
-language is \kl{downwards closed} (i.e., when it is closed under taking
-\kl{infixes}), and when the language is \kl(language){bounded} (i.e., 
-when it is included in some $w_1^* \cdots w_k^*$ for some finite choice of
-words $w_1, \ldots, w_k$). In those cases, we are able to characterize
-\kl{well-quasi-ordered} languages by the \kl{infix relation} and derive tight
-bounds on their \kl{ordinal invariants}.
+We then turn our attention to the \kl{infix relation}. In this case, we notice
+that Lemma 5.1 from \cite{DBLP:journals/ita/Kuske06} imply that there are
+\kl{well-quasi-ordered} languages for the \kl{infix relation} that have
+arbitrarily large \kl{ordinal invariants} (except for the \kl{ordinal height},
+which is always at most $\omegaOrd$). Therefore, we focus on two natural
+semantic restrictions on languages: on the one hand, we consider \kl{bounded
+languages}, that is, languages included in some $w_1^* \cdots w_k^*$ for some
+finite choice of words $w_1, \ldots, w_k$; on the other hand, we consider
+\kl{downwards closed} languages, that is, languages closed under taking
+\kl{infixes}. In both cases, we provide a very precise characterization of
+\kl{well-quasi-ordered} languages by the \kl{infix relation}, and derive tight
+bounds on their \kl{ordinal invariants}. These results are summarized in
+\cref{infixes-summary:fig}. We furthermore notice that for \kl{downwards
+closed} languages that are \kl{well-quasi-ordered} by the \kl{infix relation},
+being \kl(language){bounded} is the same as being \kl(language){regular}
+(\cref{bounded-wqo-dwclosed:cor}), and that a \kl{bounded language} is
+\kl{well-quasi-ordered} by the \kl{infix relation} if and only if its
+\kl{downwards closure} is \kl{well-quasi-ordered} by the \kl{infix relation}
+(\cref{bounded-wqo-dwclosed:cor}). This shows that, for \kl{bounded languages},
+being \kl{well-quasi-ordered} implies that their \kl{downwards closure} is a
+\kl{regular language}, which is a weakening of the usual result that the
+\kl{downwards closure} of \emph{any language} for the \kl{scattered subword
+relation} is always a \kl{regular language}.
 
-In the case of \kl{bounded languages}, we prove that a \kl{bounded language} $L
-\subseteq \Sigma^*$ is \kl{well-quasi-ordered} by the \kl{infix relation} if
-and only if it is (included in) a finite union of languages $S_i \cdot P_i$,
-where each $S_i$ is a \kl{chain} for the \kl{suffix relation}, and where each
-$P_i$ is a \kl{chain} for the \kl{prefix relation}
-(\cref{bounded-language:thm}) This result directly translates into upper bounds
-on the possible \kl{ordinal invariants} of such languages similarly as for the
-\kl{prefix relation}. Notice that these upper bounds are significantly smaller
-than \kl{ordinal invariants} of the \kl{subword relation}: they have an \kl{ordinal
-height} of at most $\omegaOrd$, an \kl{ordinal width} strictly below
-$\omegaOrd^2$, and a \kl{maximal order type} strictly below $\omegaOrd^3$
-(\cref{ordinal-invariants-bounded:cor}).
+\begin{figure}[h]
+  \centering
+  \begin{tabular}{c|c|c|c}
+      \toprule
+      $L$ & \textbf{Characterisation} & $\oWidth{L}$ & $\oType{L}$ \\
+      \midrule
+      arbitrary & \cref{infix-embedding:thm} & $< \omegaUncount$  & $< \omegaUncount$ \\
+      bounded & \cref{bounded-language:thm} & $< \omegaOrd^2$ & $< \omegaOrd^3$ \\
+      downwards closed & \cref{small-ordinal-invariants:thm} & $< \omegaOrd^2$ & $< \omegaOrd^3$ \\
+      \bottomrule
+    \end{tabular}
+    \caption{Summary of results for the \kl{infix relation}.}
+    \label{infixes-summary:fig}
+\end{figure}
 
-In the case of \kl{downwards closed} languages, we prove that they are deeply
-related to the notion of \kl{uniformly recurrent words}, borrowed from the
-study of word combinatorics. As an intermediate result, we prove that an
-infinite word $w$ is \kl{ultimately uniformly recurrent} if and only if the set
-of all \kl{infixes} of $w$ is \kl{well-quasi-ordered} by the \kl{infix
-relation} (\cref{ultimately-uniformly-recurrent:lem}). We show that every
-language $L \subseteq \Sigma^*$ that is \kl{downwards closed} and
-\kl{well-quasi-ordered} by the \kl{infix relation} is a finite union of the
-sets of finite \kl{infixes} of some \kl{ultimately uniformly recurrent}
-bi-infinite words. This also proves that such languages have an \kl{ordinal
-height} of at most $\omegaOrd$, an \kl{ordinal width} strictly below
-$\omegaOrd^2$, and a \kl{maximal order type} strictly below $\omegaOrd^3$
-(\cref{small-ordinal-invariants:thm}).
+Let us point out that \cref{small-ordinal-invariants:thm} connects
+\kl{well-quasi-orderings} by the \kl{infix relation} with the notion of
+\kl{uniformly recurrent words}, borrowed from the field of word combinatorics.
 
-Then, we turned our attention to decision procedures. To that end, we need to
-choose a computational model representing languages. For \kl{downwards closed}
-languages, because of their close connection with infinite words, we considered
-a model based on \kl{automatic sequences}. Using this model, we can decide
-whether a language is \kl{well-quasi-ordered} by the \kl{infix relation}
-(\cref{automatic-wqo:thm}). We also studied another representation, where
-languages are recognized by \kl{amalgamation systems} \cite{ASZZ24}. Such
-systems will be formally introduced in \cref{amalgamation-systems:subsec}, but
-for the moment let us just say that they include many classical computational
-models such as finite automata, context-free grammars, and Petri nets
-\cite{ASZZ24}. This provides us with a \emph{meta-algorithm} for deciding
-whether a given language is \kl{well-quasi-ordered} by the \kl{prefix
-relation}, the \kl{suffix relation}, or the \kl{infix relation} under mild
-effective restrictions on the computational model that we call an \kl{effective
-amalgamative class}, the formal definition of which we defer to
-\cref{infixes-amalgamation-effective:subsec}. Given a class $\mathcal{C}$ that
-is a \kl{strongly effective amalgamative class} of languages, we designed a
-decision procedure that takes as input a language $L \in \mathcal{C}$, and
-decides whether $L$ is \kl{well-quasi-ordered} by the \kl{prefix relation}, the
-\kl{suffix relation}, or the \kl{infix relation}
-(\cref{infix-wqo-is-emptiness:thm}). Quite surprisingly, we also showed that,
-if a language recognized by an \kl{amalgamation system} is
-\kl{well-quasi-ordered} for the \kl{infix relation}, then it is a \kl{bounded
-language} (\cref{infix-amalgamation:thm}), which automatically bounds the
-\kl{ordinal invariants} of the language. Let us point out that the above result
-implies that the hypothesis of \kl{bounded languages} on the theoretical side
-is not a restriction in practice. As a down-to-earth and more easily
-understandable consequence, our generic decision procedure applies to the class
-$\mathcal{C}_\text{aut}$ of all languages recognized by finite automata, and to
-the class $\mathcal{C}_\text{cfg}$ of all languages recognized by context-free
-grammars, which are both \kl{effective amalgamative classes}
-(\cref{aut-cfg-infix:cor}).
-
-Finally, we noticed that for \kl{downwards closed} languages that are
-\kl{well-quasi-ordered} by the \kl{infix relation}, being
-\kl(language){bounded} is the same as being \kl(language){regular}.
-Furthermore, a \kl{bounded language} is \kl{well-quasi-ordered} by the
-\kl{infix relation} if and only if its \kl{downwards closure} is
-\kl{well-quasi-ordered} by the \kl{infix relation}
-(\cref{bounded-wqo-dwclosed:cor}). This shows that, for \kl{bounded languages}
-(and therefore, for all languages recognized by \kl{amalgamation systems}) that
-are \kl{well-quasi-ordered} by the \kl{infix relation}, their \kl{downwards
-closure} is a \kl{regular language}. This is a weak version of the usual result
-that the \kl{downwards closure} for the \kl{scattered subword relation} is
-always a \kl{regular language}.
+Turning our attention to decision procedures, we consider two computational
+models respectively tailored to \kl{downwards closed} languages and to
+\kl{bounded languages}. For \kl{downwards closed} languages, we consider a
+model based on \kl{morphic words}, for which we provide a decision procedure
+(\cref{automatic-wqo:thm}). For \kl{bounded languages}, we consider the model
+of \kl{amalgamation systems} \cite{ASZZ24}, which is an abstract computational
+model that encompasses many classical ones, such as finite
+automata, context-free grammars, and Petri nets \cite{ASZZ24}. We show that if
+a language recognized by an \kl{amalgamation system} is \kl{well-quasi-ordered}
+by the \kl{infix relation}, then it is a \kl{bounded language}
+(\cref{infix-amalgamation:thm}), and is therefore \kl{regular}. Furthermore, we
+show that we can decide whether a given language recognized by an
+\kl{amalgamation system} is \kl{well-quasi-ordered} by the \kl{infix relation}
+(\cref{infix-wqo-is-emptiness:thm}). We defer the introduction of
+\kl{amalgamation systems} to \cref{amalgamation-systems:subsec}.
 
 \todo[inline]{
-  Add a figure to recap our complexity results: 
-  bounded / downwards closed / amalgamative / morphic | width.
-  And only talk about WQO classes, collapse everything and say regular.
+  rework the content so that cited things exists and/or are 
+  corollaries and theorems (mainly). Also, add results to prove that 
+  bounds are tight!
 }

### 2025-09-05 src/introduction.tex
fix: sentences

@@
       \midrule
-      arbitrary & \cref{infix-embedding:thm} & $< \omegaUncount$  & $< \omegaUncount$ \\
+      arbitrary & \cref{infix-embedding:thm} & $< \omegaOne$  & $< \omegaOne$ \\
       bounded & \cref{bounded-language:thm} & $< \omegaOrd^2$ & $< \omegaOrd^3$ \\

### 2025-09-05 src/prefixes.tex
fix: sentences

--- 2024-12-01 src/prefixes.tex
@@
 \end{corollary}
-
-\todo[inline]{
-    We actually can \emph{compute} the ordinal invariants
-    in the regular case.
-}

### 2025-09-05 src/prelims.tex
fix: sentences

@@
 parameters are called the \intro{ordinal invariants} of a
-\kl{well-quasi-ordered} set $X$. 
-As an example, for $(\Nat, \leq)$, all bad sequences are descending and there are no antichains.
-Therefore the maximal order type and the ordinal height of $\Nat$ are $\omega$ and the ordinal width is 1.
-\todo{Please check this!}
-We refer to the survey of \cite{DZSCSC20} for
+\kl{well-quasi-ordered} set $X$. We refer to the survey of \cite{DZSCSC20} for
 a detail discussion on these concepts and their computation on specific classes

### 2025-09-08 src/prefixes.tex
new: add tightness for prefixes

@@
     the \kl{ordinal height} of $L$ is at most $\omega$, and
-    its \kl{ordinal width} is finite.
+    its \kl{ordinal width} is finite. Furthermore, these bounds 
+    are tight.
 \end{corollary}
+\begin{proof}
+    The upper bounds follow from the fact that $L$ is a finite union of
+    \kl{chains}. The tightness can be obtained by considering the languages
+    $L_k \defined \bigcup_{i=0}^{k-1} a^i b^*$ for $k \in \Nat$, which 
+    are \kl{well-quasi-ordered} by the \kl{prefix relation} (as they are
+    finite unions of \kl{chains}), and satisfy
+    that $\oWidth{L_k} = k$, $\oHeight{L_k} = \omega$,
+    and therefore $\oType{L_k} = \omega \cdot k$.
+\end{proof}

### 2025-09-08 src/amalgam.tex
todos: refactor + add subsections + add tightness

@@
 \label{infixes-amalgamation:sec}
+
+\todo[inline]{%
+  We have studied infinite words, but there are other ways to represent
+  languages, such as finite automata, context-free grammars, or vector
+  addition systems with states. In this section, we show that even for
+  a generalisation of these systems, we can decide whether the language is
+  well-quasi-ordered by the infix relation, and that in turns,
+  they all fit in the bounded languages studied in the previous section.
+}
 

### 2025-09-08 src/infixes.tex
fix: making llncs work
todos: refactor + add subsections + add tightness

--- 2025-08-07 src/infixes.tex
@@
 
-\begin{theorem}[restate=bounded-language:thm,label=bounded-language:thm]
-    %\label{bounded-language:thm}
+\begin{theorem}
+    \label{bounded-language:thm}
     Let $L$ be a \kl{bounded language} of $\Sigma^*$. Then,
@@
 
+\begin{corollary}
+    \label{ordinal-invariants-bounded:cor}
+    Let $L$ be a \kl{bounded language} of $\Sigma^*$
+    that is \kl{well-quasi-ordered} by the \kl{infix relation}.
+    Then, the \kl{ordinal width} of $L$ less than $\omega^2$,
+    its \kl{ordinal height} is at most $\omega$,
+    and its \kl{maximal order type} less than $\omega^3$.
+    Furthermore, those three bounds are tight.
+\end{corollary}
+\begin{proof}
+  Upper bounds are a direct consequence of \cref{bounded-language:thm},
+  and the tightness is witnessed by the 
+  languages: 
+  $L_k \defined \bigcup_{i = 2}^{k+1} (a b^i a)^* (b a^i b)^*$,
+  that are \kl{bounded languages} of $\set{a,b}^*$,
+  \kl{well-quasi-ordered} by the \kl{infix relation},
+  and have \kl{ordinal width}, \kl{ordinal height} and
+  \kl{maximal order type} respectively equal to $\omega \cdot k$, $\omega$ and $\omega^2 \cdot k$.
+\end{proof}
+
+\section{Infixes and Downwards Closed Languages}
+\label{infixes-dwclosed:sec}
+
+\subsection{Downwards Closures and Bounded Languages}
+
 Let us now discuss the implications of this characterization in terms of
@@
 
-\begin{corollary}
-    \label{ordinal-invariants-bounded:cor}
-    Let $L$ be a \kl{bounded language} of $\Sigma^*$
-    that is \kl{well-quasi-ordered} by the \kl{infix relation}.
-    Then, the \kl{ordinal width} of $L$ is finite,
-    its \kl{ordinal height} is at most $\omega$,
-    and its \kl{maximal order type} is at most $\omega$.
-\end{corollary}
+\begin{theorem}
+  \label{bounded-wqo-dw-closed-regular:thm}
+  Let $L$ be a \kl{downwards closed} and \kl{well-quasi-ordered} language
+  of $\Sigma^*$ for the \kl{infix relation}. The following are equivalent:
+  \begin{enumerate}
+    \item $L$ is a \kl{bounded language},
+    \item $L$ is a \kl{regular language},
+    \item There exists a finite set $E \subset_{\mathsf{fin}} (\Sigma^*)^3$ such that
+      $L = \bigcup_{(x,u,y) \in E} \InfPeriodChain{x} u \InfPeriodChain{y}$.
+  \end{enumerate}
+\end{theorem}
 
-\section{Infixes and Downwards Closed Languages}
-\label{infixes-dwclosed:sec}
+\subsection{General Downwards Closed Languages}
 
@@
 
+\todo[inline]{aliaume: write it}
+Let us conclude this section with a few remarks on the notion of
+\kl{factor complexity} of languages.
+
+\begin{lemma}
+  There exists a language $L$ with exponential factor complexity 
+  that is \kl{well-quasi-ordered} for the \kl{infix relation}.
+\end{lemma}
+
+\subsection{Decision Procedures}
+
 \AP As we have demonstrated, infinite (or bi-infinite words) can be used to

### 2025-09-08 src/amalgam.tex
new: add refs + morhpic words + todos

--- 2025-08-12 src/amalgam.tex
@@
 
-\AP In this section, we are going to design an effective decision procedure for
-\kl{well-quasi-ordering} by the \kl{infix relation}. To that end, the first
-requirement is to fix a way to represent languages $L \subseteq \Sigma^*$.
-Traditionally, one would use finite automata, context-free grammars, or for the
-more adventurous, vector addition systems with states. However, our proof
-technique will only require us to have a way to ``glue'' together runs of the
-system to ``pump'' them and produce new runs: this is the usual pumping lemma
-in automata theory, and Ogden's lemma for context-free grammars \cite{OGDEN68}.
-It turns out that there is a rather large family of systems for which pumping
-arguments based on so-called \emph{minimal runs} exist: they are called
-\emph{amalgamation systems} and were recently introduced by Anand, Schmitz,
-Schütze, and Zetzsche \cite{ASZZ24}.
+\AP In the previous section, we have represented languages that are
+\kl{downwards closed} by the \kl{infix relation} as \kl{infixes} of infinite
+words. However, there are many other natural ways to represent languages, such
+as finite automata or context-free grammars. In this section, we are going to
+show that our results on \kl{bounded languages} can be applied to a large class
+of systems, called \kl{amalgamation systems}, that includes as particular
+examples finite automata and context-free grammars.
 
@@
 \kl(language){bounded}. This implies that all the results of
-\cref{infixes-bounded:sec}, and in
-particular \cref{bounded-language:thm}, can safely be applied to
-\kl{amalgamation systems}.
+\cref{infixes-bounded:sec}, and in particular \cref{bounded-language:thm}, can
+safely be applied to \kl{amalgamation systems}.
 

### 2025-09-08 src/infixes.tex
new: add refs + morhpic words + todos

@@
   and have \kl{ordinal width}, \kl{ordinal height} and
-  \kl{maximal order type} respectively equal to $\omega \cdot k$, $\omega$ and $\omega^2 \cdot k$.
+  \kl{maximal order type} respectively equal to $k \cdot \omega$, $\omega$ and $k \cdot \omega^2$.
 \end{proof}
@@
 
-\subsection{Downwards Closures and Bounded Languages}
+Let us now discuss another classical restriction that can be imposed on 
+languages when studying \kl{well-quasi-orders}, that of being
+\kl{downwards closed}. Indeed, the \cref{infix-embedding:thm}
+crucially relies on constructing languages that are \emph{not}
+\kl{downwards closed}, and we have shown 
+in \cref{dw-closure-not-wqo:rem} that the \kl{downwards closure}
+of a \kl{well-quasi-ordered} language is not necessarily
+\kl{well-quasi-ordered}.
 
-Let us now discuss the implications of this characterization in terms of
-\kl{downwards closures}: if $L$ is a \kl{bounded language}, then considering
-$L$ or its \kl{downwards closure} is equivalent with respect to being
-\kl{well-quasi-ordered} by the \kl{infix relation}.
+\subsection{Characterization of Well-Quasi-Ordered Downwards Closed Languages}
+
+An immediate consequence of \cref{bounded-language:thm} is that   if $L$ is a
+\kl{bounded language}, then considering $L$ or its \kl{downwards closure}
+$\dwset[\infleq]{L}$ is equivalent with respect to being
+\kl{well-quasi-ordered} by the \kl{infix relation}, as opposed to
+the general case illustrated in \cref{dw-closure-not-wqo:rem}.
 
@@
 \end{proof}
-
-\begin{theorem}
-  \label{bounded-wqo-dw-closed-regular:thm}
-  Let $L$ be a \kl{downwards closed} and \kl{well-quasi-ordered} language
-  of $\Sigma^*$ for the \kl{infix relation}. The following are equivalent:
-  \begin{enumerate}
-    \item $L$ is a \kl{bounded language},
-    \item $L$ is a \kl{regular language},
-    \item There exists a finite set $E \subset_{\mathsf{fin}} (\Sigma^*)^3$ such that
-      $L = \bigcup_{(x,u,y) \in E} \InfPeriodChain{x} u \InfPeriodChain{y}$.
-  \end{enumerate}
-\end{theorem}
-
-\subsection{General Downwards Closed Languages}
 
@@
 
-\todo[inline]{aliaume: write it}
-Let us conclude this section with a few remarks on the notion of
-\kl{factor complexity} of languages.
+\AP
+Let us conclude this section with a few remarks on the notion of \kl{factor
+complexity} of languages. Recall that the \intro{factor complexity} of a
+language $L \subseteq \Sigma^*$ is the function $f_L: \Nat \to \Nat$ such that
+$f_L(n)$ is the number of distinct words of size $n$ in $L$. A language $L$ has
+\intro{exponential factor complexity} if there exists a constant $C > 1$ such
+that $f_L(n) \geq C^n$ for infinitely many $n \in \Nat$.
 
 \begin{lemma}
-  There exists a language $L$ with exponential factor complexity 
+  \label{exponential-factor-complexity:lem}
+  There exists a language $L$ with \kl{exponential factor complexity}
   that is \kl{well-quasi-ordered} for the \kl{infix relation}.
 \end{lemma}
+\begin{proof}
+  We leverage results from \cite{CAKA97} on Toeplitz words.
+  Namely, the 
+  $(5, 3)$-Toeplitz word is \kl{uniformly recurrent} 
+  \cite[p. 499]{CAKA97},
+  but has an exponential factor complexity 
+  \cite[Theorem 5]{CAKA97}.
+\end{proof}
+
+It is the case that \kl{bounded languages} have at most a polynomial factor
+complexity, and that the \kl{Thue-Morse sequence} has a linear factor
+complexity. More generally, as we will see, most ``concrete'' examples of
+infinite words have a low factor complexity: \kl{automatic sequences} have at
+most a linear factor complexity, and \kl(morphic){morphic sequences} have at
+most a polynomial factor complexity. In that sense, the
+\cref{exponential-factor-complexity:lem} shows that one can still construct
+``pathological'' examples of infinite words that are \kl{well-quasi-ordered}
+for the \kl{infix relation}, but have a high factor complexity.
 
@@
 
-Note that being \kl{uniformly recurrent} still leaves open a lot of potential
-behaviours, for instance, Sturmian words are \kl{uniformly recurrent}, but are
-not \kl(sequence){automatic}. Interestingly enough, the \kl{ordinal invariants}
-of the \kl{automatic sequences} are not smaller than the ones of \kl{ultimately
-uniformly recurrent} words: this contrasts with the case of languages that are
-\kl{well-quasi-ordered} for the \kl{infix relation} in general, where the
-hypothesis of being representable (by a \kl{regular language}, or an
-\kl{amalgamation system}) will drastically reduce the \kl{ordinal invariants}
-of said language (\cref{ordinal-invariants-bounded:cor}, \cref{infix-amalgamation:thm}).
+One can go beyond \cref{automatic-wqo:thm} and consider
+\kl(morphic){morphic sequences}, which are obtained by iterating a
+\kl(substitution){substitution} on a finite alphabet, and then applying a
+\kl(homomorphism){homomorphism}. The \kl{Fibonacci sequence} is an example of
+a \kl(morphic){morphic sequence}. It is known that one can decide whether a
+\kl(morphic){morphic sequence} is \kl{uniformly recurrent}
+\cite[Theorem 1]{DURAND13}.
 
-\todo[inline]{For Nathan: 
-  talk about morphic words / sequences and 
-  decidability here.
-  \url{https://arxiv.org/pdf/1204.5393}
-}
-
-\todo[inline]{Talk about factor complexity and 
-well-quasi-orderings: the two are unrelated. This can be a remark.
-\url{https://arxiv.org/pdf/2206.00376}
-Here example 5 p8 is an explicit example of a uniformly recurrent word with exponential factor complexity.
-}
+\todo[inline]{Define properly morphic words and 
+  the questions associated to them. Also, there is a difference between
+  \kl{ultimately uniformly recurrent} and \kl{uniformly recurrent}
+  that is not handled here.}
+\begin{theorem}
+    \label{morphic-wqo:thm}
+    Given a \kl(morphic){morphic sequence} $w \in \Sigma^{\Nat}$, one can decide
+    whether the set of \kl{infixes} of $w$ is \kl{well-quasi-ordered} for the
+    \kl{infix relation}.
+\end{theorem}

### 2025-09-08 wqo-on-words.lncs.tex
fix: ci/cd

--- 2025-08-07 wqo-on-words.tex
@@
 % - If you want to update the template itself (e.g., change the layout), please
-%   edit the `templates/plain-article.tex` file instead.
-\documentclass[11pt,a4paper,twosided]{article}
+%   edit the `templates/lncs/lncs.tex` file instead.
+\documentclass[runningheads]{llncs}
 
-% we setup a custom geometry because the default one is too narrow
-\usepackage{geometry}
-\geometry{margin=3.5cm}
-
-% utf-8 for old systems
 \usepackage[utf8]{inputenc}
 \usepackage[T1]{fontenc}
+
+\usepackage{lineno}
+\linenumbers
+
+\usepackage{todonotes}
 
@@
 
-\usepackage{todonotes}
-\usepackage{lineno}
-\linenumbers
-
 % math packages
-\usepackage{amsmath,amsthm,amssymb,stmaryrd,thmtools,upgreek}
-
-% configure some theorems
-\newtheorem{theorem}{Theorem}
-\newtheorem{lemma}[theorem]{Lemma}
-\newtheorem{corollary}[theorem]{Corollary}
-\newtheorem{proposition}[theorem]{Proposition}
-\newtheorem{conjecture}[theorem]{Conjecture}
-\newtheorem{assumption}[theorem]{Assumption}
-\theoremstyle{definition}
-\newtheorem{definition}[theorem]{Definition}
-\newtheorem{remark}[theorem]{Remark}
-\newtheorem{example}[theorem]{Example}
+\usepackage{amssymb,amsmath,stmaryrd,thmtools,thm-restate,upgreek}
 
@@
 \usepackage{graphicx}
-\usepackage{subcaption}
 \usepackage[obeyclassoptions,mode=tex]{standalone}
@@
 
-% ornaments
-\usepackage{pgfornament}
+\usepackage{subcaption}
 
@@
 
+% Packages for macro definitions
+\usepackage{xparse}
+\usepackage{xpatch}
+\usepackage{tokcycle}
+\usepackage{ifthen}
+
+% Proof trees
+\usepackage{bussproofs}
+
 % Algorithms
@@
 
-% Packages for macro definitions
-\usepackage{xparse}
-\usepackage{xpatch}
-\usepackage{tokcycle}
-\usepackage{ifthen}
-
-% Proofs
-\usepackage{bussproofs}
-
 % Colors 
 \usepackage{ensps-colorscheme}
-% parametrize knowledge colors and style
-% intro => A4 + emph 
-% kl    => normal color, normal size not emph
 \knowledgestyle{intro notion}{color={A2}, emphasize}
 \knowledgestyle{notion}{color={B1}}
-\hypersetup{
-    colorlinks=true,
-    breaklinks=true,
-    linkcolor=A4,
-    citecolor=A4,
-    urlcolor=A4,
-    filecolor=A4,
-}
 
@@
 
-% We include the title and author information based on the 
-% `paper-meta.yaml` file.
+\newcommand{\repositoryUrl}{\url{https://github.com/AliaumeL/word-ordering-wqo}}
 
-\title{Well-quasi-orderings on word languages}
+\begin{document}
+%
+\title{Well-quasi-orderings on word languages \\
+       main 0121a5929682a015f9ca5ed9399070224366a74c \\
+       2025-08-12 15:10:21 +0200
+}
+\titlerunning{WQO on Words}
 
 \author{
-Nathan Lhote\thanks{Aix-Marseille University}
+        Nathan Lhote\inst{2}\orcidID{0000-0003-3303-5368} \and
+        Aliaume Lopez\inst{1}\orcidID{0000-0002-4205-327X} \and
+        Lia Schütze\inst{3}\orcidID{0000-0003-4002-5491}}
+\authorrunning{N. Lhote and A. Lopez and L. Schütze}
+\institute{
+University of Warsaw
  \and
-Aliaume Lopez\thanks{University of Warsaw}
+Aix-Marseille University
  \and
-Lia Schütze\thanks{Max Planck Institute for Software Systems}
+Max Planck Institute for Software Systems
 }
 
-% For the date, we first check if the user has provided a date,
-% and otherwise use the git meta inforamtion (if available).
-\date{2025-07-01 13:26:43 +0200\footnote{99064007a226e55157ecfac20bbb40804687f3e2 -- branch main at git@github.com:AliaumeL/word-ordering-wqo.git}}
+\date{\today}
 
-\newcommand{\repositoryUrl}{\url{https://github.com/AliaumeL/word-ordering-wqo}}
-
-\knowledge{notion}
- | kl-usage
-
-% Now, we create the document itself.
-\begin{document}
-% Generate the title page
-\maketitle
-% Print the abstract
+%
+\maketitle              % typeset the header of the contribution
+%
 \begin{abstract}
@@
 We furthermore provide decision procedures for a large class of languages, that contains regular and context-free languages.
-    \paragraph{Keywords:}
-    equivariant ideal, Hilbert basis, ideal membership problem, orbit finite, oligomorphic, well-quasi-ordering
-\paragraph{Repository:} \repositoryUrl
 \end{abstract}
-
-\klogo\ This document uses \href{https://ctan.org/pkg/knowledge}{knowledge}:
-\kl[kl-usage]{notion} points to its \intro[kl-usage]{definition}.
 
@@
 
+% Include acknowledgements
+
 % Include the bibliography
-\bibliographystyle{plainurl}
+\bibliographystyle{splncs04}
 \bibliography{papers.bib}

### 2025-09-08 wqo-on-words.lipics.tex
fix: ci/cd

--- 2025-08-07 wqo-on-words.tex
@@
-%! TEX program = pdflatex
+%! TEX program = xelatex
 % WARNING: this is a generated file.
@@
 % - If you want to update the template itself (e.g., change the layout), please
-%   edit the `templates/plain-article.tex` file instead.
-\documentclass[11pt,a4paper,twosided]{article}
+%   edit the `templates/lipics/lipics.tex` file instead.
+\documentclass[
+    a4paper,
+    UKenglish,
+    cleveref,
+    autoref,anonymous,
+    thm-restate]{lipics-v2021}
 
-% we setup a custom geometry because the default one is too narrow
-\usepackage{geometry}
-\geometry{margin=3.5cm}
-
-% utf-8 for old systems
-\usepackage[utf8]{inputenc}
-\usepackage[T1]{fontenc}
-
-% babel for language settings
-\usepackage[english]{babel}
-
-% microtype for better typography
-\usepackage{microtype}
-
-\usepackage{todonotes}
-\usepackage{lineno}
-\linenumbers
-
-% math packages
-\usepackage{amsmath,amsthm,amssymb,stmaryrd,thmtools,upgreek}
-
-% configure some theorems
-\newtheorem{theorem}{Theorem}
-\newtheorem{lemma}[theorem]{Lemma}
-\newtheorem{corollary}[theorem]{Corollary}
-\newtheorem{proposition}[theorem]{Proposition}
-\newtheorem{conjecture}[theorem]{Conjecture}
-\newtheorem{assumption}[theorem]{Assumption}
-\theoremstyle{definition}
-\newtheorem{definition}[theorem]{Definition}
-\newtheorem{remark}[theorem]{Remark}
-\newtheorem{example}[theorem]{Example}
+\hideLIPIcs
 
@@
 
-% ornaments
-\usepackage{pgfornament}
-
 % links inside the document
-\usepackage{hyperref}
-\usepackage[capitalise,noabbrev,nameinlink]{cleveref}
-\Crefname{assumption}{Assumption}{Assumptions}
 \usepackage[composition,hyperref,xcolor,cleveref]{knowledge}
@@
 \usepackage{varwidth}
-
-% Algorithms
-\usepackage{algorithm2e}
-\Crefname{algocfline}{Algorithm}{Algorithms}
-\crefname{algocfline}{Algorithm}{Algorithms}
-\crefname{algocf}{Algorithm}{Algorithms}
-\Crefname{algocf}{Algorithm}{Algorithms}
 
@@
 
-\author{
-Nathan Lhote\thanks{Aix-Marseille University}
- \and
-Aliaume Lopez\thanks{University of Warsaw}
- \and
-Lia Schütze\thanks{Max Planck Institute for Software Systems}
-}
+\titlerunning{WQO on Words}
+\author{Nathan Lhote}
+       {Aix-Marseille University}
+       {nathan.lhote@lis-lab.fr}
+       {https://orcid.org/0000-0003-3303-5368}
+       {}
+\author{Aliaume Lopez}
+       {University of Warsaw}
+       {ad.lopez@uw.edu.pl}
+       {https://orcid.org/0000-0002-4205-327X}
+       {}
+\author{Lia Schütze}
+       {Max Planck Institute for Software Systems}
+       {lschuetze@mpi-sws.org}
+       {https://orcid.org/0000-0003-4002-5491}
+       {}
 
-% For the date, we first check if the user has provided a date,
-% and otherwise use the git meta inforamtion (if available).
-\date{2025-07-01 13:26:43 +0200\footnote{99064007a226e55157ecfac20bbb40804687f3e2 -- branch main at git@github.com:AliaumeL/word-ordering-wqo.git}}
+\authorrunning{N. Lhote and A. Lopez and L. Schütze}
+
+\Copyright{Nathan Lhote and Aliaume Lopez and Lia Schütze}
+
+\ccsdesc[100]{Mathematics of computing\textsubscript{Algebra}Polynomials}
+\ccsdesc[100]{Theory of computation\textasciitilde Logic and verification\textasciitilde Model checking}
+
+\keywords{equivariant ideal, Hilbert basis, ideal membership problem, orbit finite, oligomorphic, well-quasi-ordering} %TODO mandatory; please add comma-separated list of keywords
+
+\category{} %optional, e.g. invited paper
 
@@
 
-\knowledge{notion}
- | kl-usage
+%Editor-only macros:: begin (do not touch as author)%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\EventEditors{John Q. Open and Joan R. Access}
+\EventNoEds{2}
+\EventLongTitle{42nd Conference on Very Important Topics (CVIT 2016)}
+\EventShortTitle{CVIT 2016}
+\EventAcronym{CVIT}
+\EventYear{2016}
+\EventDate{December 24--27, 2016}
+\EventLocation{Little Whinging, United Kingdom}
+\EventLogo{}
+\SeriesVolume{42}
+\ArticleNo{23}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 In this paper, we describe to what extent other natural orderings (prefix, suffix, and infix) on words can be used to derive Higman-like theorems.
-More specifically, we are interested in characterizing \emph{languages} of words that are well-quasi-ordered under these orderings.
-We show that a simple characterization is possible for the prefix and suffix orderings, and that under extra regularity assumptions, this also extends to the infix ordering.
-We furthermore provide decision procedures for a large class of languages, that contains regular and context-free languages.
-    \paragraph{Keywords:}
-    equivariant ideal, Hilbert basis, ideal membership problem, orbit finite, oligomorphic, well-quasi-ordering
-\paragraph{Repository:} \repositoryUrl
+More specifically, we are interested in characterizing \emph{languages} of words that are well-quasi-ordered under these orderings, and explore their properties and connections with other language theoretic notions.
+We furthermore give decition procedures when the languages are given by various comupational models such as automata, context-free grammars, and automatic structures.
 \end{abstract}
-
-\klogo\ This document uses \href{https://ctan.org/pkg/knowledge}{knowledge}:
-\kl[kl-usage]{notion} points to its \intro[kl-usage]{definition}.
 

### 2025-09-29 wqo-on-words.lncs.tex
fix: institutions

@@
 \title{Well-quasi-orderings on word languages \\
-       main 0121a5929682a015f9ca5ed9399070224366a74c \\
-       2025-08-12 15:10:21 +0200
+       main dd2237175323943c3893f9ae93107aba9f6049a5 \\
+       2025-09-08 16:09:49 +0200
 }
@@
 \institute{
-University of Warsaw
+INP Bordeaux, LaBRI, CNRS
  \and
@@
 In this paper, we describe to what extent other natural orderings (prefix, suffix, and infix) on words can be used to derive Higman-like theorems.
-More specifically, we are interested in characterizing \emph{languages} of words that are well-quasi-ordered under these orderings.
-We show that a simple characterization is possible for the prefix and suffix orderings, and that under extra regularity assumptions, this also extends to the infix ordering.
-We furthermore provide decision procedures for a large class of languages, that contains regular and context-free languages.
+More specifically, we are interested in characterizing \emph{languages} of words that are well-quasi-ordered under these orderings, and explore their properties and connections with other language theoretic notions.
+We furthermore give decition procedures when the languages are given by various comupational models such as automata, context-free grammars, and automatic structures.
 \end{abstract}

### 2025-09-30 src/infixes.tex
new: factor complexity uniformly recurrent morphic sequences

@@
 
-\AP
-Let us conclude this section with a few remarks on the notion of \kl{factor
+\subsubsection{Factor Complexity}
+
+\AP Let us conclude this section with a few remarks on the notion of \kl{factor
 complexity} of languages. Recall that the \intro{factor complexity} of a
@@
 \intro{exponential factor complexity} if there exists a constant $C > 1$ such
-that $f_L(n) \geq C^n$ for infinitely many $n \in \Nat$.
+that $f_L(n) \geq C^n$ for infinitely many $n \in \Nat$. A language $L$ has
+\intro{subaffine factor complexity} if $f_L(n) \leq C \cdot n + K$ for some
+constants $C, K > 0$ and all $n \in \Nat$.
 
@@
 
+\todo[inline]{Rewrite this paragraph to cite precise results.}
 It is the case that \kl{bounded languages} have at most a polynomial factor
@@
 infinite words have a low factor complexity: \kl{automatic sequences} have at
-most a linear factor complexity, and \kl(morphic){morphic sequences} have at
-most a polynomial factor complexity. In that sense, the
-\cref{exponential-factor-complexity:lem} shows that one can still construct
-``pathological'' examples of infinite words that are \kl{well-quasi-ordered}
-for the \kl{infix relation}, but have a high factor complexity.
+most a linear factor complexity, and \kl(morphic){morphic sequences} 
+have a polynomial factor complexity. In
+that sense, the \cref{exponential-factor-complexity:lem} shows that one can
+still construct ``pathological'' examples of infinite words that are
+\kl{well-quasi-ordered} for the \kl{infix relation}, but have a high factor
+complexity.
+
+Note that because of \cite[Theorem 24]{NIPR09}, any \kl{uniformly recurrent}
+\kl{morphic sequence} has a linear factor complexity. In particular, we
+conjecture that any \kl{ultimately uniformly recurrent} \kl{morphic sequence}
+has a linear factor complexity too.
 

### 2025-09-30 src/conclusion.tex
new: add refs + morhpic words + todos
new: morphic sequences factor complexity + moving stuff around

@@
 
-\paragraph*{Graph Classes}
-\todo[inline]{aliaume: retalk about \cite{ALM17} and \cite{L24:arxiv:v2}}
+\subsubsection{Factor Complexity}
+
+\AP Let us conclude this section with a few remarks on the notion of \kl{factor
+complexity} of languages. Recall that the \intro{factor complexity} of a
+language $L \subseteq \Sigma^*$ is the function $f_L: \Nat \to \Nat$ such that
+$f_L(n)$ is the number of distinct words of size $n$ in $L$. A language $L$ has
+\intro{exponential factor complexity} if there exists a constant $C > 1$ such
+that $f_L(n) \geq C^n$ for infinitely many $n \in \Nat$. A language $L$ has
+\intro{subaffine factor complexity} if $f_L(n) \leq C \cdot n + K$ for some
+constants $C, K > 0$ and all $n \in \Nat$.
+We extend the notion of \kl{factor complexity} to infinite and bi-infinite words
+as the \kl{factor complexity} of their set of finite \kl{infixes}.
+
+While there clearly are languages with \kl{subaffine factor complexity} that are
+not \kl{well-quasi-ordered} for the \kl{infix relation}, such as 
+the language $L \defined \dwset{ a b^* a }$;
+one would expect that languages that are \kl{well-quasi-ordered} for the
+\kl{infix relation} would have a low \kl{factor complexity}.
+
+\AP It is the case that \kl{bounded languages}, studied in
+\cref{infixes-bounded:sec}, have a polynomial factor complexity: such a
+language is by definition included in $w_1^* \cdots w_n^*$ for some words $w_1,
+\dots, w_n$, and an infix is a choice of the number of occurrences of each
+$w_i$ in the infix, plus a choice of beginning and ending of the infix in the
+leftmost and rightmost $w_i$. We showed in \cref{bounded-language:thm} and
+particularly in \cref{bounded-language:lem}, that \kl{bounded languages} that
+are \kl{well-quasi-ordered} for the \kl{infix relation} are in fact included in
+some finite union of languages of the form $w_1^* w_2 w_3^*$, hence have
+actually a quadratic factor complexity. However,
+\cref{exponential-factor-complexity:lem} shows that there are
+\kl{downwards-closed} languages that are \kl{well-quasi-ordered} for the
+\kl{infix relation} but have an \kl{exponential factor complexity}.
+
+\begin{lemma}
+  \label{exponential-factor-complexity:lem}
+  There exists a language $L$ with \kl{exponential factor complexity}
+  that is \kl{well-quasi-ordered} for the \kl{infix relation}.
+\end{lemma}
+\begin{proof}
+  We leverage results from \cite{CAKA97} on Toeplitz words.
+  Namely, the 
+  $(5, 3)$-Toeplitz word is \kl{uniformly recurrent} 
+  \cite[p. 499]{CAKA97},
+  but has an exponential factor complexity 
+  \cite[Theorem 5]{CAKA97}.
+\end{proof}
+
+Our understanding of the situation is that whenever a computational model is
+fixed to represent languages that are \kl{well-quasi-ordered} for the \kl{infix
+relation}, the \kl{factor complexity} of the represented languages drops to a
+quadratic one. This is the case for languages recognized by \kl{amalgamation
+systems} \cref{amalgamation-systems:subsec}, that are \kl{bounded languages}
+when they are \kl{well-quasi-ordered} (\cref{infix-amalgamation:thm}). 
+This is
+also the case for languages described as the infixes of a finite set of pairs
+of \kl{morphic sequences}. Indeed, the factor complexity of a \kl{morphic
+sequence} that is \kl{uniformly recurrent} is linear \cite[Theorem 24]{NIPR09}.
+and therefore the factor complexity of the set of infixes of a bi-infinite word
+obtained as $\bar{w_1} u w_2$ where $w_1, w_2$ are \kl{morphic sequences} and
+$u$ is a finite word is quadratic.

### 2025-09-30 src/infixes.tex
new: morphic sequences factor complexity + moving stuff around

--- 2025-09-08 src/infixes.tex
@@
     \oHeight{L} < \omega \oComProd \omega^2 = \omegaOrd^3$ allows us to conclude.
+
+    The tightness of the bounds is a direct consequence of
+    \cref{from-bi-to-single:lem}, and by considering a finite union of 
+    these examples over disjoint alphabets (or even, by considering a binary 
+    alphabet and using unambiguous codes to separate the different components).
 \end{proofof}
 
-\AP
-Let us conclude this section with a few remarks on the notion of \kl{factor
-complexity} of languages. Recall that the \intro{factor complexity} of a
-language $L \subseteq \Sigma^*$ is the function $f_L: \Nat \to \Nat$ such that
-$f_L(n)$ is the number of distinct words of size $n$ in $L$. A language $L$ has
-\intro{exponential factor complexity} if there exists a constant $C > 1$ such
-that $f_L(n) \geq C^n$ for infinitely many $n \in \Nat$.
-
-\begin{lemma}
-  \label{exponential-factor-complexity:lem}
-  There exists a language $L$ with \kl{exponential factor complexity}
-  that is \kl{well-quasi-ordered} for the \kl{infix relation}.
-\end{lemma}
-\begin{proof}
-  We leverage results from \cite{CAKA97} on Toeplitz words.
-  Namely, the 
-  $(5, 3)$-Toeplitz word is \kl{uniformly recurrent} 
-  \cite[p. 499]{CAKA97},
-  but has an exponential factor complexity 
-  \cite[Theorem 5]{CAKA97}.
-\end{proof}
-
-It is the case that \kl{bounded languages} have at most a polynomial factor
-complexity, and that the \kl{Thue-Morse sequence} has a linear factor
-complexity. More generally, as we will see, most ``concrete'' examples of
-infinite words have a low factor complexity: \kl{automatic sequences} have at
-most a linear factor complexity, and \kl(morphic){morphic sequences} have at
-most a polynomial factor complexity. In that sense, the
-\cref{exponential-factor-complexity:lem} shows that one can still construct
-``pathological'' examples of infinite words that are \kl{well-quasi-ordered}
-for the \kl{infix relation}, but have a high factor complexity.
-
 \subsection{Decision Procedures}
+\label{decision-procedures:sec}
 
@@
 represent languages that are \kl{well-quasi-ordered} for the \kl{infix
-relation} by considering their set of finite \kl{infixes}. We can leverage the
-theory of \kl{automatic sequences}, that is tailored to study the regularity of
-infinite words to build a decision procedure for \kl{well-quasi-ordered}
-languages for the \kl{infix relation}. A sequence $w \in \Sigma^\Nat$ is
-\intro(sequence){automatic} whenever there exists a base $b \in \Nat$ and a
-finite automaton that can compute the $i$-th letter of $w$ given as input the
-number $i$ in base $b$. The \kl{Thue-Morse sequence} is an example of an
-\kl{automatic sequence}.
+relation} by considering their set of finite \kl{infixes}.
+Let us formalise the representation of languages by sets of bi-infinite
+words that we will use in this section, following the characterization of
+\cref{bi-infinite:lem}. A \intro{sequence representation} of a language $L
+\subseteq \Sigma^*$ is a finite set of triples $(w_i^-, a_i, w_i^+)_{1 \leq i
+\leq n}$ where $w_i^-, w_i^+ \in \Sigma^\Nat \cup \Sigma^*$ are two potentially
+infinite words, and $a_i \in \Sigma$ is a letter, such that
+\begin{equation*}
+    L = \bigcup_{i = 1}^n \infset{\mathsf{reversed}(w_i^-) a_i w_i^+} \quad .
+\end{equation*}
 
-\begin{theorem}
-    \label{automatic-wqo:thm}
+\AP Given an effective representation of sequences, one obtains an effective
+representation of languages via \kl{sequence representations}. We can leverage
+the theory of \kl{automatic sequences}, that is tailored to study the
+regularity of infinite words to build a decision procedure for
+\kl{well-quasi-ordered} languages for the \kl{infix relation}. A sequence $w
+\in \Sigma^\Nat$ is \intro(sequence){automatic} whenever there exists a base $b
+\in \Nat$ and a finite automaton that can compute the $i$-th letter of $w$
+given as input the number $i$ in base $b$. The \kl{Thue-Morse sequence} is an
+example of an \kl{automatic sequence}.
+
+\begin{lemma}
+  \label{automatic-uur:lem}
     Given an \kl{automatic sequence} $w \in \Sigma^{\Nat}$, one can decide
-    whether the set of \kl{infixes} of $w$ is \kl{well-quasi-ordered} for the
-    \kl{infix relation}.
-\end{theorem}
+    whether it is \kl{ultimately uniformly recurrent}.
+\end{lemma}
 \begin{proof}
-    Because of \cref{ultimately-uniformly-recurrent:lem}, it suffices
-    to decide whether $w$ is \kl{ultimately uniformly recurrent}.
     We can rewrite this as a question on the \kl{automatic sequence} $w$
@@
 
-One can go beyond \cref{automatic-wqo:thm} and consider
-\kl(morphic){morphic sequences}, which are obtained by iterating a
-\kl(substitution){substitution} on a finite alphabet, and then applying a
-\kl(homomorphism){homomorphism}. The \kl{Fibonacci sequence} is an example of
-a \kl(morphic){morphic sequence}. It is known that one can decide whether a
-\kl(morphic){morphic sequence} is \kl{uniformly recurrent}
-\cite[Theorem 1]{DURAND13}.
+Based on \cref{automatic-uur:lem}, we can now prove the following theorem.
+\begin{theorem}
+    \label{automatic-wqo:thm}
+    Given a \kl{sequence representation} of a language $L \subseteq
+    \Sigma^*$ where all infinite words are \kl{automatic sequences}, one can
+    decide whether $L$ is \kl{well-quasi-ordered} for the \kl{infix relation}.
+\end{theorem}
+\begin{proof}
+    It is easy to see that $L$ is \kl{well-quasi-ordered} for the \kl{infix
+    relation} if and only if for every triple $(w_i^-, a_i, w_i^+)$ in the
+    \kl{sequence representation} of $L$, the (potentially bi-infinite) word
+    $\mathsf{reversed}(w_i^-) a_i w_i^+$ defines a \kl{well-quasi-ordered}
+    language. By \cref{from-bi-to-single:lem}, this is the case if and only if
+    both $w_i^-$ and $w_i^+$ are \kl{ultimately uniformly recurrent}. 
+    Since one can decide whether an \kl{automatic sequence} is
+    \kl{ultimately uniformly recurrent} from \cref{automatic-uur:lem},
+    we conclude the proof.
+\end{proof}
 
-\todo[inline]{Define properly morphic words and 
-  the questions associated to them. Also, there is a difference between
-  \kl{ultimately uniformly recurrent} and \kl{uniformly recurrent}
-  that is not handled here.}
-\begin{theorem}
-    \label{morphic-wqo:thm}
+\todo[inline]{properly introduce morphic sequences, and maybe talk about
+them in the introduction}
+
+\AP
+For \kl{automatic sequences}, the question of being \kl{ultimately uniformly
+recurrent} is easily seen to be decidable, but symbolic dynamics provides a
+more general framework to study infinite words. A \intro{morphic sequence} is
+obtained by iterating a \kl(substitution){substitution} on a finite alphabet,
+and then applying a \kl(homomorphism){homomorphism}. The \kl{Fibonacci
+sequence} is an example of a \kl(morphic){morphic sequence}.
+It was recently shown that one can decide whether a \kl(morphic){morphic
+sequence} is \kl{uniformly recurrent} \cite[Theorem 1]{DURAND13}.
+We can leverage this result to obtain the following lemma.
+
+\begin{lemma}
+  \label{morphic-uur:lem}
     Given a \kl(morphic){morphic sequence} $w \in \Sigma^{\Nat}$, one can decide
-    whether the set of \kl{infixes} of $w$ is \kl{well-quasi-ordered} for the
-    \kl{infix relation}.
-\end{theorem}
+    whether it is \kl{ultimately uniformly recurrent}.
+\end{lemma}
+
+From \cref{morphic-uur:lem}, we can now prove the following corollary,
+which is a generalization of \cref{automatic-wqo:thm}.
+
+\begin{corollary}
+    \label{morphic-wqo:cor}
+    Given a \kl{sequence representation} of a language $L \subseteq
+    \Sigma^*$ where all infinite words are \kl(morphic){morphic sequences}, one can
+    decide whether $L$ is \kl{well-quasi-ordered} for the \kl{infix relation}.
+\end{corollary}

### 2025-09-30 src/introduction.tex
new: morphic sequences factor complexity + moving stuff around

@@
 bounds on their \kl{ordinal invariants}. These results are summarized in
-\cref{infixes-summary:fig}. We furthermore notice that for \kl{downwards
+\cref{infixes-summary:fig}. \todo{this is not what the corollary says} We furthermore notice that for \kl{downwards
 closed} languages that are \kl{well-quasi-ordered} by the \kl{infix relation},
@@
 
-Let us point out that \cref{small-ordinal-invariants:thm} connects
-\kl{well-quasi-orderings} by the \kl{infix relation} with the notion of
-\kl{uniformly recurrent words}, borrowed from the field of word combinatorics.
-
 Turning our attention to decision procedures, we consider two computational
@@
 \kl{bounded languages}. For \kl{downwards closed} languages, we consider a
-model based on \kl{morphic words}, for which we provide a decision procedure
-(\cref{automatic-wqo:thm}). For \kl{bounded languages}, we consider the model
+model based on representations of infinite words
+(\cref{decision-procedures:sec}), for which we provide a decision procedure
+(\cref{automatic-wqo:thm}). The model used to represent these infinite words is
+based on \kl{automatic sequences} and \kl{morphic sequences} \cite{CAKA97},
+which are well-studied in the context of symbolic dynamics.
+For \kl{bounded languages}, we consider the model
 of \kl{amalgamation systems} \cite{ASZZ24}, which is an abstract computational
-model that encompasses many classical ones, such as finite
-automata, context-free grammars, and Petri nets \cite{ASZZ24}. We show that if
-a language recognized by an \kl{amalgamation system} is \kl{well-quasi-ordered}
-by the \kl{infix relation}, then it is a \kl{bounded language}
+model that encompasses many classical ones, such as finite automata,
+context-free grammars, and Petri nets \cite{ASZZ24}. We show that if a language
+recognized by an \kl{amalgamation system} is \kl{well-quasi-ordered} by the
+\kl{infix relation}, then it is a \kl{bounded language}
 (\cref{infix-amalgamation:thm}), and is therefore \kl{regular}. Furthermore, we

### 2025-10-01 src/conclusion.tex
fix: knowledges + refs + trying to prove uur decidability for pure morphic words

@@
 \cref{exponential-factor-complexity:lem} shows that there are
-\kl{downwards-closed} languages that are \kl{well-quasi-ordered} for the
+\kl{downwards closed} languages that are \kl{well-quasi-ordered} for the
 \kl{infix relation} but have an \kl{exponential factor complexity}.
@@
 systems} \cref{amalgamation-systems:subsec}, that are \kl{bounded languages}
-when they are \kl{well-quasi-ordered} (\cref{infix-amalgamation:thm}). 
-This is
+when they are \kl{well-quasi-ordered} (\cref{infix-amalgamation:thm}). This is
 also the case for languages described as the infixes of a finite set of pairs
 of \kl{morphic sequences}. Indeed, the factor complexity of a \kl{morphic
-sequence} that is \kl{uniformly recurrent} is linear \cite[Theorem 24]{NIPR09}.
-and therefore the factor complexity of the set of infixes of a bi-infinite word
-obtained as $\bar{w_1} u w_2$ where $w_1, w_2$ are \kl{morphic sequences} and
-$u$ is a finite word is quadratic.
+sequence} that is \kl{uniformly recurrent} is linear \cite[Theorem 24]{NIPR09},
+therefore the \kl{factor complexity} of a language given by \kl{sequence
+representation} using \kl{morphic sequences} is at most quadratic.

### 2025-10-01 src/infixes.tex
fix: knowledges + refs + trying to prove uur decidability for pure morphic words

@@
 \AP Given an effective representation of sequences, one obtains an effective
-representation of languages via \kl{sequence representations}. We can leverage
-the theory of \kl{automatic sequences}, that is tailored to study the
-regularity of infinite words to build a decision procedure for
-\kl{well-quasi-ordered} languages for the \kl{infix relation}. A sequence $w
-\in \Sigma^\Nat$ is \intro(sequence){automatic} whenever there exists a base $b
-\in \Nat$ and a finite automaton that can compute the $i$-th letter of $w$
-given as input the number $i$ in base $b$. The \kl{Thue-Morse sequence} is an
-example of an \kl{automatic sequence}.
+representation of languages via \kl{sequence representations}. In this section,
+we will rely on definitions originating from the area of symbolic dynamics,
+that precisely study infinite words whose generation follows from a finitely
+described process. However, we will not assume that the reader is familiar with
+this domain, and we will use as black-boxes key results from this area.
+
+\AP A first model that one can use to represent infinite words is the model of
+\intro{automatic sequences}. In this case, the infinite word $w$ is described
+by a finite state automaton, that can compute the $i$-th letter of the word $w$
+given as input the number $i$ written in some base $b \in \Nat$. An example of
+such a sequence is the \kl{Thue-Morse sequence} that can be described by a
+finite automaton using a binary representation of the indices. The good
+algorithmic properties of \kl{automatic sequences} come from the fact that a
+Presburger definable property that uses letters of the sequence can be
+(trivially) translated into a finite automaton that reads the base $b$
+representation of the free variables (that are indices of the sequence). We
+briefly explain in the following lemma the folklore result that one can decide
+if an \kl{automatic sequence} is \kl{ultimately uniformly recurrent}.
 
@@
     Since one can decide whether an \kl{automatic sequence} is
-    \kl{ultimately uniformly recurrent} from \cref{automatic-uur:lem},
+    \kl{ultimately uniformly recurrent} using \cref{automatic-uur:lem},
     we conclude the proof.
@@
 
-\todo[inline]{properly introduce morphic sequences, and maybe talk about
-them in the introduction}
+\AP In fact, \kl{automatic sequences} are part of a larger family of sequences
+studied in symbolic dynamics, called \kl{morphic sequences}. Let us first
+recall that a \intro(words){morphism} is a function $f \colon \Sigma^* \to
+\Gamma^*$ such that for every $u,v \in \Sigma^*$, $f(uv) = f(u)f(v)$. A
+\intro{morphic sequence} $w$ is an infinite word obtained by iterating a
+\kl(words){morphism} $f \colon \Sigma^* \to \Sigma^*$ on a letter $a \in
+\Sigma$ such that $f(a)$ starts with $a$, and then applying a
+\kl(words){homomorphism} $h \colon \Sigma^* \to \Gamma^*$. The infinite word
+$f^\omega(a)$ is the limit of the sequence $\seqof{f^n(a)}[n \in \Nat]$, which
+is well-defined because $f(a)$ starts with $a$, and the \kl{morphic sequence}
+is $ w \defined h(f^\omega(a))$.
 
 \AP
-For \kl{automatic sequences}, the question of being \kl{ultimately uniformly
-recurrent} is easily seen to be decidable, but symbolic dynamics provides a
-more general framework to study infinite words. A \intro{morphic sequence} is
-obtained by iterating a \kl(substitution){substitution} on a finite alphabet,
-and then applying a \kl(homomorphism){homomorphism}. The \kl{Fibonacci
-sequence} is an example of a \kl(morphic){morphic sequence}.
-It was recently shown that one can decide whether a \kl(morphic){morphic
-sequence} is \kl{uniformly recurrent} \cite[Theorem 1]{DURAND13}.
-We can leverage this result to obtain the following lemma.
+Every \kl{automatic
+sequence} is a \kl{morphic sequence}, but not the other way around. We refer
+the reader to a short survey of \cite{ALSZ17} for more details on the possible
+variations on the definition of \kl{morphic sequences} and their relationships.
+It was relatively recently proven that one can decide whether a \kl{morphic
+sequence} is \kl{uniformly recurrent} \cite[Theorem 1]{DURAND13}. We were not
+able to find in the literature whether one can decide \kl{ultimate uniform
+recurrence}, but we can prove it for a restricted class of \kl{morphic
+sequences}, that are called \intro{pure morphic sequences}, and are precisely
+those that do not use a final \kl(words){homomorphism} $h$. Following the same proof
+as for \cref{automatic-wqo:thm}, one deduces from the upcoming
+\cref{morphic-uur:lem} the following \cref{morphic-wqo:cor}. We actually
+conjecture that \cref{morphic-uur:lem} can be extended to all \kl{morphic
+sequences} \cref{morphic-uur:conj}.
 
@@
   \label{morphic-uur:lem}
-    Given a \kl(morphic){morphic sequence} $w \in \Sigma^{\Nat}$, one can decide
+    Given a \kl{pure morphic sequence} $w \in \Sigma^{\Nat}$, one can decide
     whether it is \kl{ultimately uniformly recurrent}.
 \end{lemma}
+\begin{proof}
+  On the one hand, one can guess a bound $N_0$ and check that 
+  the suffix of the word $w_{\geq N_0}$ is \kl{uniformly recurrent},
+  since the suffix of a \kl{morphic sequence} is also a \kl{morphic sequence},
+  and \kl{uniform recurrence} is decidable for \kl{morphic sequences}
+  \cite[Theorem 2]{DURAND13}.
 
-From \cref{morphic-uur:lem}, we can now prove the following corollary,
-which is a generalization of \cref{automatic-wqo:thm}.
+  On the other hand, we can also provide an upper bound on the possible values
+  of $N_0$ if $w$ is \kl{ultimately uniformly recurrent}. Let us write $f(a) =
+  a s$ for some finite word $s \in \Sigma^*$. If $s$ is $f$-bounded, i.e., if
+  the size of $f^n(s)$ is bounded by some constant $K \in \Nat$, then $w$ is
+  ultimately periodic, and the prefix $N_0$ that may not be periodic can easily
+  be bounded by looking at the morphism $f$. Let us consider the case when $s$
+  is not $f$-bounded. If the word $w$ is \kl{ultimately uniformly recurrent},
+  then there exists a bound $N_0$ such that $w_{\geq N_0}$ is \kl{uniformly
+  recurrent}. In particular, for a large enough $n$, there exists $k_0$  such
+  that for all $k \geq k_0$, the word $f^n(s)$ is an \kl{infix} of $f^{n +
+  k}(s)$.
+  \todo[inline]{not clear how to conclude ...}
+
+\end{proof}
 
@@
     Given a \kl{sequence representation} of a language $L \subseteq
-    \Sigma^*$ where all infinite words are \kl(morphic){morphic sequences}, one can
+    \Sigma^*$ where all infinite words are \kl{pure morphic sequences}, one can
     decide whether $L$ is \kl{well-quasi-ordered} for the \kl{infix relation}.
 \end{corollary}
+
+\begin{conjecture}
+  \label{morphic-uur:conj}
+    Given a \kl{morphic sequence} $w \in \Sigma^{\Nat}$, one can decide
+    whether it is \kl{ultimately uniformly recurrent}.
+\end{conjecture}

### 2025-10-01 src/introduction.tex
fix: knowledges + refs + trying to prove uur decidability for pure morphic words

@@
         arbitrary & Finite unions of \kl{chains} \cref{prefixes:thm} & $< \omegaOrd$ & $< \omegaOrd^2$ \\
-        regular   & Finite unions of \kl{regular} \kl{chains} \cref{prefix-wqo-reg-decidable:cor} & $< \omegaOrd$ & $< \omegaOrd^2$ \\
+        regular   & Finite unions of \kl(language){regular} \kl{chains} \cref{prefix-wqo-reg-decidable:cor} & $< \omegaOrd$ & $< \omegaOrd^2$ \\
         \bottomrule
@@
 based on \kl{automatic sequences} and \kl{morphic sequences} \cite{CAKA97},
-which are well-studied in the context of symbolic dynamics.
-For \kl{bounded languages}, we consider the model
-of \kl{amalgamation systems} \cite{ASZZ24}, which is an abstract computational
-model that encompasses many classical ones, such as finite automata,
-context-free grammars, and Petri nets \cite{ASZZ24}. We show that if a language
-recognized by an \kl{amalgamation system} is \kl{well-quasi-ordered} by the
-\kl{infix relation}, then it is a \kl{bounded language}
-(\cref{infix-amalgamation:thm}), and is therefore \kl{regular}. Furthermore, we
-show that we can decide whether a given language recognized by an
-\kl{amalgamation system} is \kl{well-quasi-ordered} by the \kl{infix relation}
-(\cref{infix-wqo-is-emptiness:thm}). We defer the introduction of
-\kl{amalgamation systems} to \cref{amalgamation-systems:subsec}.
+which are well-studied in the context of symbolic dynamics. For \kl{bounded
+languages}, we consider the model of \kl{amalgamation systems} \cite{ASZZ24},
+which is an abstract computational model that encompasses many classical ones,
+such as finite automata, context-free grammars, and Petri nets \cite{ASZZ24}.
+We show that if a language recognized by an \kl{amalgamation system} is
+\kl{well-quasi-ordered} by the \kl{infix relation}, then it is a \kl{bounded
+language} (\cref{infix-amalgamation:thm}), and is therefore
+\kl(language){regular}. Furthermore, we show that we can decide whether a given
+language recognized by an \kl{amalgamation system} is \kl{well-quasi-ordered}
+by the \kl{infix relation} (\cref{infix-wqo-is-emptiness:thm}). We defer the
+introduction of \kl{amalgamation systems} to
+\cref{amalgamation-systems:subsec}.
 
@@
 \kl{well-quasi-ordering} by the \kl{infix relation} \cref{infixes-bounded:sec}.
-These non-trivial extensions continue to rely
-on classical word combinatorial techniques that are already present in the work
-of Atminas, Lozin, and Moshkov \cite[Section 3]{ALM17}.
 

### 2025-10-01 src/infixes.tex
new: proof attempt for pure morphic... but does not work

@@
 
-  On the other hand, we can also provide an upper bound on the possible values
-  of $N_0$ if $w$ is \kl{ultimately uniformly recurrent}. Let us write $f(a) =
-  a s$ for some finite word $s \in \Sigma^*$. If $s$ is $f$-bounded, i.e., if
-  the size of $f^n(s)$ is bounded by some constant $K \in \Nat$, then $w$ is
-  ultimately periodic, and the prefix $N_0$ that may not be periodic can easily
-  be bounded by looking at the morphism $f$. Let us consider the case when $s$
-  is not $f$-bounded. If the word $w$ is \kl{ultimately uniformly recurrent},
-  then there exists a bound $N_0$ such that $w_{\geq N_0}$ is \kl{uniformly
-  recurrent}. In particular, for a large enough $n$, there exists $k_0$  such
-  that for all $k \geq k_0$, the word $f^n(s)$ is an \kl{infix} of $f^{n +
-  k}(s)$.
-  \todo[inline]{not clear how to conclude ...}
+  Let us now prove that a specific choice of $N_0$ is sufficient. Let us write
+  $w \defined f^\omega(a)$ where $f \colon \Sigma^* \to \Sigma^*$ is a
+  \kl(words){morphism} and $a \in \Sigma$ is a letter such that $f(a)$ starts
+  with $a$. Let us also write $f(a) = a s$, where $s \in \Sigma^+$ is a
+  non-empty word. Notice that $w = a s f(s) f^2(s) \cdots$. Without loss of
+  generality, one can assume that $|f^n(s)| \to \infty$ when $n \to \infty$,
+  otherwise $w$ is ultimately periodic, and therefore \kl{ultimately uniformly
+  recurrent}. Since one can algorithmically detect if this is the case, we
+  assume that it is not.
+
+  Because $\Sigma$ is finite, there exists a finite set of letters $A \subseteq
+  \Sigma$ and a suffix $w'$ of $w$ such that $w' \in A^*$ and every letter of
+  $w'$ appears infinitely often in $w'$. Furthermore, one can compute this set
+  $A$ and a bound $N_0$ such that $f^n(s) \in A^*$ for every $n \geq N_0$. We
+  claim that $w$ is \kl{ultimately uniformly recurrent} if and only if
+  $f^{N_0}(s) f^{N_0 + 1}(s) \cdots$ is \kl{uniformly recurrent}.
+  One direction being obvious, we assume that $w$ is \kl{ultimately uniformly
+  recurrent}, and let us prove that $f^{N_0}(s) f^{N_0 + 1}(s) \cdots$ is
+  \kl{uniformly recurrent}.
+
+  Let us assume that $w'' \defined f^{N_1}(s) f^{N_1 + 1}(s) \cdots$ where $N_1
+  \geq N_0$ is \kl{uniformly recurrent}. Let us show that $f^{N_0}(s)$ is an
+  \kl{infix} of $w''$ if and only if there exists a letter $b$ in $s$ such that
+  $f^{N_0}(s) \infleq f^{k} (b)$ for some finite number $k$.
+
+  Assume that $f^{N_0}(s)$ is an \kl{infix} of $w''$. Recall that $|f^n(s)| \to
+  \infty$ when $n \to \infty$, hence, there must exist a letter $b$ in $s$ such
+  that $|f^n(b)| \to \infty$ when $n \to \infty$. Because every letter of
+  $f^{N_0}(s)$ appears infinitely often in $w''$, $f^n(b)$ is a growing
+  sequence of infixes of $w''$. Since $w''$ is \kl{uniformly recurrent}, there
+  exists a bound $K$ such that $f^{N_0}(s)$ is an \kl{infix} of every $f^k(b)$
+  for $k \geq K$. Conversely, if there exists a letter $b$ in $s$ such that
+  $f^{N_0}(s) \infleq f^{k} (b)$ for some finite number $k$, then $f^{N_0}(s)$
+  is an \kl{infix} of $w''$ because $f^n(b)$ is an infix of $w''$ for every $n
+  \in \Nat$.
+
+  Because $w''$ is \kl{uniformly recurrent}, there exists a bound $K$ such that
+  $f^{N_1}(s) \infleq f^k(b)$ for every $k \geq K$.
+  \todo[inline]{Cannot finish the proof...}
 

### 2025-10-06 src/infixes.tex
Build for lncs

@@
 
-\begin{theorem}
+\begin{theorem}[restate=bounded-language:thm,label=bounded-language:thm]
     \label{bounded-language:thm}
@@
 \kl{well-quasi-ordered} for the \kl{infix relation}. This proves the (easy)
-right-to-left implication of \cref{bounded-language:thm}.
+right-to-left implication of \cref{bounded-language:thm}. Furthermore, any
+language $L$ included in a finite union of products $S_i \cdot P_i$
+is in fact a \kl{bounded language}. \todo[inline]{I think we can in fact construct chains that are not bounded languages. E.g. $ab, aba^2b, aba^2ba^3b, \ldots$?}
+
+\todo[inline]{aliaume: yes, this way nonsense.}
 
@@
     Then $L$ is included in a finite union 
-    of products of \kl{chains}:
+    of products of \kl{chains}: \todo{reference}
     \begin{equation*}
@@
     Remark that the \kl{downwards closure} of a product of two \kl{chains}
-    is a finite union of products of two chains.
+    is a finite union of products of two chains. \todo{reference/proof}.
     As a consequence, we conclude that $\dwset[\infleq]{L}$ is itself included
@@
 \kl{downwards closed} language for this relation is characterized by finitely
-many excluded \kl{subwords}, hence provides a \kl{bounded language}. However, this
+many excluded \kl{subwords}, hence provides a \kl{bounded language} \todo{wrong? $\{a,b\}^*$ is downward closed but not bounded}. However, this
 is not the case for the \kl{infix relation}, as we will now illustrate with the
@@
     \label{dwclosed-wqo-not-finite-excl:ex}
-    Let $L \defined a^* b^* \cup b^* a^*$. This language is \kl{downwards
+    Let $L \defined a^* b^* \cup b^* a^*$ \todo{However, this language is bounded, so this seems to be a non sequitur}. This language is \kl{downwards
     closed} for the \kl{infix relation}, is \kl{well-quasi-ordered} for the

### 2025-10-06 wqo-on-words.tex
new: cut intro!
Build for lncs

@@
 
-% We include the title and author information based on the 
-% `paper-meta.yaml` file.
+\newcommand{\repositoryUrl}{\url{https://github.com/AliaumeL/word-ordering-wqo}}
 
-\title{Well-quasi-orderings on word languages}
+\begin{document}
+%
+\title{Well-quasi-orderings on word languages \\
+       main edb3a6bc24be4a12fe4f383c75384a6dff9d1a3c \\
+       2025-10-02 17:15:30 +0200
+}
+\titlerunning{WQO on Words}
 
 \author{
-Nathan Lhote\thanks{Aix-Marseille University}
+        Nathan Lhote\inst{2}\orcidID{0000-0003-3303-5368} \and
+        Aliaume Lopez\inst{3}\orcidID{0000-0002-4205-327X} \and
+        Lia Schütze\inst{1}\orcidID{0000-0003-4002-5491}}
+\authorrunning{N. Lhote and A. Lopez and L. Schütze}
+\institute{
+Max Planck Institute for Software Systems
  \and
-Aliaume Lopez\thanks{University of Warsaw}
+Aix-Marseille University
  \and
-Lia Schütze\thanks{Max Planck Institute for Software Systems}
+INP Bordeaux, LaBRI, CNRS
 }
 
-% For the date, we first check if the user has provided a date,
-% and otherwise use the git meta inforamtion (if available).
-\date{2025-07-01 13:26:43 +0200\footnote{99064007a226e55157ecfac20bbb40804687f3e2 -- branch main at git@github.com:AliaumeL/word-ordering-wqo.git}}
+\date{\today}
 
-\newcommand{\repositoryUrl}{\url{https://github.com/AliaumeL/word-ordering-wqo}}
-
-\knowledge{notion}
- | kl-usage
-
-% Now, we create the document itself.
-\begin{document}
-% Generate the title page
-\maketitle
-% Print the abstract
+%
+\maketitle              % typeset the header of the contribution
+%
 \begin{abstract}

### 2025-10-06 src/infixes.tex
new: answering todos in a stupid way

@@
     Then $L$ is included in a finite union 
-    of products of \kl{chains}: \todo{reference}
+    of products of \kl{chains}:
     \begin{equation*}
@@
     Remark that the \kl{downwards closure} of a product of two \kl{chains}
-    is a finite union of products of two chains. \todo{reference/proof}.
+    is a finite union of products of two chains.
     As a consequence, we conclude that $\dwset[\infleq]{L}$ is itself included
@@
 \kl{downwards closed} language for this relation is characterized by finitely
-many excluded \kl{subwords}, hence provides a \kl{bounded language} \todo{wrong? $\{a,b\}^*$ is downward closed but not bounded}. However, this
+many excluded \kl{subwords}, hence provides a \kl{bounded language}. However, this
 is not the case for the \kl{infix relation}, as we will now illustrate with the
@@
     \label{dwclosed-wqo-not-finite-excl:ex}
-    Let $L \defined a^* b^* \cup b^* a^*$ \todo{However, this language is bounded, so this seems to be a non sequitur}. This language is \kl{downwards
+    Let $L \defined a^* b^* \cup b^* a^*$. This language is \kl{downwards
     closed} for the \kl{infix relation}, is \kl{well-quasi-ordered} for the

### 2025-10-06 wqo-on-words.lncs.tex
fix: knowledges + refs + trying to prove uur decidability for pure morphic words
new: answering todos in a stupid way

@@
 \Crefname{assumption}{Assumption}{Assumptions}
+\Crefname{conjecture}{Conjecture}{Conjectures}
 \usepackage[composition,hyperref,xcolor,cleveref]{knowledge}
@@
 \title{Well-quasi-orderings on word languages \\
-       main dd2237175323943c3893f9ae93107aba9f6049a5 \\
-       2025-09-08 16:09:49 +0200
+       main edb3a6bc24be4a12fe4f383c75384a6dff9d1a3c \\
+       2025-10-02 17:15:30 +0200
 }
@@
 \author{
-        Nathan Lhote\inst{2}\orcidID{0000-0003-3303-5368} \and
-        Aliaume Lopez\inst{1}\orcidID{0000-0002-4205-327X} \and
-        Lia Schütze\inst{3}\orcidID{0000-0003-4002-5491}}
+        Nathan Lhote\inst{3}\orcidID{0000-0003-3303-5368} \and
+        Aliaume Lopez\inst{2}\orcidID{0000-0002-4205-327X} \and
+        Lia Schütze\inst{1}\orcidID{0000-0003-4002-5491}}
 \authorrunning{N. Lhote and A. Lopez and L. Schütze}
 \institute{
+Max Planck Institute for Software Systems
+ \and
 INP Bordeaux, LaBRI, CNRS
@@
 Aix-Marseille University
- \and
-Max Planck Institute for Software Systems
 }

### 2025-10-06 src/introduction.tex
Adjust proof of Thm 1

@@
 bounds on their \kl{ordinal invariants}. These results are summarized in
-\cref{infixes-summary:fig}. \todo{this is not what the corollary says} We furthermore notice that for \kl{downwards
+\cref{infixes-summary:fig}. We furthermore notice that for \kl{downwards
 closed} languages that are \kl{well-quasi-ordered} by the \kl{infix relation},

### 2025-10-06 src/prefixes.tex
new: add refs + morhpic words + todos
new: answering todos in a stupid way
Adjust proof of Thm 1

--- 2025-08-06 src/prefixes.tex
@@
     Conversely, assume that $L$ is \kl{well-quasi-ordered} by the \kl{prefix
-    relation}. Let us define $S$ the set of words $w$ such that there exists
+    relation}. Let us define $S_{\mathrm{split}}$ the set of words $w \in \Sigma^*$ 
+    such that there exists
     two words $wu$ and $wv$ both in $L$ that are not comparable for the
-    \kl{prefix relation}. Assume by contradiction that $S$ is infinite.
+    \kl{prefix relation}. Let $S = S_{\mathrm{split}}~\cup~\min_{\prefleq} L$\todo{kinda ugly, but i can't find a nicer way of phrasing it}. 
+    Assume by contradiction that $S$ is infinite.
     Then, $S$ equipped with the \kl{prefix relation} is an infinite
@@
     This contradicts the assumption that $L$ is \kl{well-quasi-ordered}.
-    Now, consider $w$ be a maximal element for the \kl{prefix ordering}
-    in $S$. By definition, all words in $L$ that contain $w$ as a prefix
-    must be comparable for the \kl{prefix relation}. Therefore,
-    $(\upset[\prefleq]{w}) \cap L$ is a \kl{chain} for the \kl{prefix relation}.
+
+    Now, let $w$ be a maximal element for the \kl{prefix ordering}
+    in $S$. 
+    The upward closure of $w$ in $L$, $(\upset[\prefleq]{w}) \cap L$, must be a 
+    finite union of \kl{chains}. Otherwise at least two of the chains would share a common 
+    prefix in $w\Sigma$, contradicting the maximality of $w$.
+
     In particular, letting $S_{\max}$ be the set of all maximal elements
@@
     \end{equation*}
-    \todo[inline]{Because $S$ only contains words with two incomparable suffixes in $L$, I think this is wrong - consider e.g. $L \subseteq a^*$, where $S$ and $S_{max}$ are necessarily empty}
     Hence, that $L$ is finite union of \kl{chains}.

### 2025-10-07 src/infixes.tex
Remove erroneous statement in Thm 2
Simplify Lemma 3

--- 2025-10-06 src/infixes.tex
@@
 \kl{well-quasi-ordered} for the \kl{infix relation}. This proves the (easy)
-right-to-left implication of \cref{bounded-language:thm}. Furthermore, any
-language $L$ included in a finite union of products $S_i \cdot P_i$
-is in fact a \kl{bounded language}. \todo[inline]{I think we can in fact construct chains that are not bounded languages. E.g. $ab, aba^2b, aba^2ba^3b, \ldots$?}
-
-\todo[inline]{aliaume: yes, this way nonsense.}
+right-to-left implication of \cref{bounded-language:thm}.
 
@@
     Let $u,v \in \Sigma^*$ be two (non-empty) \kl{periodic words}
-    having \kl{periodic lengths} $p$ and $q$ respectively.
-    Then, if $u \infleq v$ and $\card{u} \geq pq$,
-    then $u$ and $v$ share the same \kl{periodic length}
-    $p = q$.
+    having \kl{periods} of length $p$ and $q$ respectively.
+    If $u \infleq v$ and $\card{u} \geq 2\max\{p, q\}$,
+    then $u$ and $v$ both have a period of length
+    $\gcd(p,q)$.
 \end{lemma}
 \begin{proof}
-    The fact that $u$ and $v$ are \kl{periodic length}
-    respectively $p$ and $q$ translates into the fact that $u_{i+p} = u_i$ and
-    $v_{i+q} = v_i$ for all indices $i \in \Nat$ such that those letters are
-    well-defined.
+    Since $u$ has a period of length $p$ we have $u_{i+p} = u_i$ for all indices $i \in \Nat$ such that those letters are
+    well-defined. Since $u$ is an \kl{infix} of $v$ and $\card{u} \geq q$, we also have $u_{i+q} = u_i$. It follows that $u$ has a period of length $\gcd(p, q)$.
 
-    Now, assume that $u$ is an \kl{infix} of $v$, this provides the existence
-    of a $k \in \Nat$ such that $u = v_{k} \cdots v_{k + \card{u} - 1}$. In
-    particular, $v_{k+i+p} = v_{k+i}$ for all $i \in \Nat$ such that $k+i+p < k
-    + \card{u}$. Due to the periodicity of $v$, we also have $v_{k+i+q} = v_{k+i}$ for all $i$ with $k + i + q \leq k + \card{u}$. We conclude that both $u$ and $v$ are of
-    \kl{periodic length} the greatest common divisor of $p$ and $q$, and by
-    minimality of $q$ this must be equal to $q$ and to $p$.
-    \todo{maybe elaborate on why $pq$ as bound on $\card{u}$ (also technically $2\max\{p,q\}$ would be enough, but maybe more cumbersome?}
+    Since $v$ contains $u$ as a suffix and $\card{u} \geq q$, we can also conclude that $v$ must also have period of length $\gcd(p,q)$.
 \end{proof}

### 2025-10-07 src/infixes.tex
... and corrollary 3

@@
     Let $u,v \in \Sigma^*$ and $k, \ell \in \Nat$
-    such that $k \geq \factorial[p]{\card{u} \times \card{v}}$,
-    $\ell \geq \factorial[p]{\card{v} \times \card{u}}$,
+    such that $k \geq 2\max\{\card{u},\card{v}\}$,
     and $u^k \infleq v^\ell$.

### 2025-10-07 src/infixes.tex
Rephrase for clarity

@@
     \label{inf-period-chain:lem}
-    Let $x \in \Sigma^+$ be a word, and
+    Let $x \in \Sigma^+$ be a word.
     Then $\InfPeriodChain{x}$ is a finite union of \kl{chains}
@@
     such that
-    one of the following holds for all
-    $\vec{n} \in \Nat^{k}$:
+    for all
+    $\vec{n} \in \Nat^{k}$
+    one of the following holds:
     \begin{enumerate}

### 2025-10-07 src/infixes.tex
making a note for later

@@
     such that
-    for all
+    for all \todo{proof does not actually show for all}
     $\vec{n} \in \Nat^{k}$
@@
     Now, because of the chosen values for $\vec{n_j}$, there exists $1 \leq \ell \leq
-    k-1$ such that $w[\vec{n_i}]$ is actually an \kl{infix} of $u_{\ell}
-    v_{\ell}^{n_{j,\ell}} u_{\ell+1} v_{\ell+1}^{n_{j,\ell+1}} u_{\ell+2}$.
-    Even more,
+    k-1$ such that
     one of the three following equations holds:
@@
         \item $w[\vec{n_i}] \infleq
-            v_{\ell+1}^{n_{j,\ell+1}} u_{\ell+2}$.
+            v_{\ell}^{n_{j,\ell}} u_{\ell+1}$.
     \end{itemize}
@@
     a number $1 \leq t \leq k$ such that
-    $v_i^{n_i} \in \InfPeriodChain{x}$ for all $1 \leq i \leq t$,
+    $v_i^{n_{i,\ell}} \in \InfPeriodChain{x}$ for all $1 \leq \ell \leq t$,
     and
-    $v_i^{n_i} \in \InfPeriodChain{y}$ for all $t < i \leq k$.
+    $v_i^{n_{i,\ell}} \in \InfPeriodChain{y}$ for all $t < \ell \leq k$.
     In particular,

### 2025-10-10 src/infixes.tex
slight generalization in lemma 5

@@
     that there exists $x,y \in \Sigma^+$ of size at most $s$, and 
-    a number $1 \leq t \leq k$ such that
-    $v_i^{n_{i,\ell}} \in \InfPeriodChain{x}$ for all $1 \leq \ell \leq t$,
+    a number $1 \leq t \leq k$ such that for all $z \in \Nat$,
+    $v_i^{z} \in \InfPeriodChain{x}$ for all $1 \leq i \leq t$,
     and
-    $v_i^{n_{i,\ell}} \in \InfPeriodChain{y}$ for all $t < \ell \leq k$.
+    $v_i^{z} \in \InfPeriodChain{y}$ for all $t < i \leq k$.
     In particular,

### 2025-10-10 src/infixes.tex
word -> infix of thue-morse

@@
 for every word $u$ that is an \kl{infix} of $\ThueMorse$, there exists $k \geq
-1$ such that $u$ is an \kl{infix} of every word $v$ of size at least $k$. We
-refer the reader to a nice survey of Allouche and Shallit for more information
+1$ such that $u$ occurs as an \kl{infix} of every $k$-sized infix $v \infleq \ThueMorse$. 
+We refer the reader to a nice survey of Allouche and Shallit for more information
 on this sequence and its properties \cite{ALSHA99}.

### 2025-10-10 src/infixes.tex
lemma 7 polishing

@@
 \begin{proof}
-    By construction $\LMorse$ is an \emph{infinite}
-    \kl{downwards closed} for the \kl{infix relation}. Let us argue that $\LMorse$ is
-    \kl{well-quasi-ordered} for the \kl{infix relation}, but is not \kl(language){bounded}.
-    It is \kl{well-quasi-ordered} because it is the set of finite \kl{infixes}
-    of a \kl{uniformly recurrent} word, and we can apply \cref{uniformly-recurrent:lem}.
+    By construction $\LMorse$ is
+    \kl{downwards closed} for the \kl{infix relation},
+    and by \cref{uniformly-recurrent:lem}, it is
+    \kl{well-quasi-ordered}.
 
     Assume by contradiction that $\LMorse$ is \kl(language){bounded}. In this case, there exist
-    words $w_1, \dots, w_k \in \Sigma^*$ such that $L \subseteq w_1^* \cdots
-    w_k^*$. Since $L$ is infinite and \kl{downwards closed}, there exists a
-    word $u \in L$ such that $u = w_i^3$ for some $1 \leq i \leq k$. This is absurd
+    words $w_1, \dots, w_k \in \Sigma^*$ such that $\LMorse \subseteq w_1^* \cdots
+    w_k^*$. Since $\LMorse$ is infinite and \kl{downwards closed}, there exists a
+    word $u \in \LMorse$ such that $u = w_i^3$ for some $1 \leq i \leq k$. This is a contradiction,
     because $u \infleq \ThueMorse$, which is \kl{cube-free}.

### 2025-10-10 src/infixes.tex
lemma 8 polishing

@@
     \label{thue-morse-ordinal:lemma}
-    The \kl{maximal order type} of $\LMorse$ is $\omegaOrd$,
+    Under $\infleq$,
+    the \kl{maximal order type} of $\LMorse$ is $\omegaOrd$,
     the \kl{ordinal height} of $\LMorse$ is $\omegaOrd$,
@@
 \begin{proof}
-    Let us prove that these are upper bounds for the \kl{ordinal invariants} of
-    $L$. For the \kl{ordinal height}, it is true for any language $\LMorse$.
+	We first show that $\omegaOrd$ is an upper bound for each of these measure, 
+	before showing that the bounds are tight.
+
+    %Let us prove that these are upper bounds for the \kl{ordinal invariants} of
+    $\LMorse$. 
+    The bound of the \kl{ordinal height} holds for any language $L$,
+    since a descending chain for $(\Sigma^*, \infleq)$ is isomorphic to one for $(\Nat, \leq)$ \todo{please check this is a good argument}.
     For the \kl{maximal order type}, we remark that
+    the \kl{uniform recurrence} of $\ThueMorse$ means that
     the maximal length of a \kl{bad sequence} is determined by its first element,
     hence that it is at most $\omegaOrd$.
-    Finally, because the \kl{ordinal width} is at most the \kl{maximal order type},
+    Finally, because the \kl{ordinal width} is at most the \kl{maximal order type} \todo{add reference},
     we conclude that the \kl{ordinal width} is also at most $\omegaOrd$.
@@
     $\oHeight{\LMorse} = \omegaOrd$: given any number $n \in \Nat$, one can construct a
-    \kl{decreasing sequence} of words in $L$ of length $n$, for instance by
+    \kl{decreasing sequence} of words in $\LMorse$ of length $n$, for instance by
     considering the first $n$ prefixes of the \kl{Thue-Morse sequence} by
@@
     Let us now prove that $\oWidth{\LMorse} = \omegaOrd$. Assume by contradiction that
-    $\oWidth{\LMorse}$ is finite. Then, $L$ can be written as a finite union of
-    \kl{chains} for the \kl{infix relation}, and in particular, $L$ is
+    $\oWidth{\LMorse}$ is finite. Then, $\LMorse$ can be written as a finite union of
+    \kl{chains} for the \kl{infix relation}, and in particular, $\LMorse$ is
     \kl(language){bounded}, which is absurd by \cref{thue-morse:lemma}.

### 2025-10-10 src/infixes.tex
lemma 9 polishing

@@
 case, a \kl{well-quasi-ordered} \kl{order ideal} for the \kl{infix relation} is
-the set of finite \kl{infixes} of a finite or bi-infinte word $w \in
+the set of finite \kl{infixes} of a finite \todo{$\Sigma^\Nat$ is an infinite word}
+or bi-infinte word $w \in
 \Sigma^\Nat \cup \Sigma^\Rel $ (\cref{bi-infinite:lem}). Therefore, we will
@@
     \label{bi-infinite:lem}
-    Let $L \subseteq \Sigma^*$ be an \kl{order ideal} that is \kl{well-quasi-ordered} 
-    for the 
+    Let $L \subseteq \Sigma^*$ be an \kl{order ideal} 
+    for a \kl{well-quasi-ordered} 
     \kl{infix relation}. Then $L$ is the set of finite \kl{infixes}
@@
 \begin{proof}
-    Let us assume that $L$ is infinite, the case when it is finite 
-    is similar, but leads to a finite word.
+    Let us assume that $L$ is infinite. The case when it is finite 
+    is similar, but will result in a finite word.
 
     Because the alphabet $\Sigma$ is finite, we can enumerate the words of $L$
-    as $\seqof{w_i}[i \in \Nat]$. Now, let us build the sequence $\seqof{u_i}[i \in \Nat]$ by induction
+    as $\seqof{w_i}[i \in \Nat]$. From $\seqof{w_i}[i \in \Nat]$, we construct a sequence $\seqof{u_i}[i \in \Nat]$ by induction
     as follows: $u_0 = w_0$, and $u_{i+1}$ is a word that contains $u_i$ and

### 2025-10-10 src/infixes.tex
lemma 10 polishing

@@
     therefore assume that all words in the sequence $\seqof{w_i}[i \in \Nat]$ are such that
-    they start at a position $i \geq N_0$. But then, they are all finite
+    they start at a position $i \geq N_0$. But then they are all finite
     \kl{infixes} of $w_{\geq N_0}$, which is a \kl{uniformly recurrent} word,
@@
     $\mathsf{Rec}(w_{\geq i})$ for any (infinite) suffix of $w$. The sequence
-    $R_i \defined \mathsf{Rec}(w_{\geq i})$ is a sequence of \kl{downwards
+    $R_i \defined \mathsf{Rec}(w_{\geq i})$ is a descending sequence of \kl{downwards
     closed} sets of finite words, included in the set of finite infixes of $w$
@@
     every finite infix of $v$ 
-    appears infinitely often in $v$. Let us consider 
-    some finite infix $u \infleq v$, we claim that there is a bound $N_u$
+    appears infinitely often in $v$. 
+    Given
+    some finite infix $u \infleq v$, we 
+    there exists a bound $N_u$
     on the distance
@@
     Indeed, if it is not the case, then there exists an infinite sequence
-    $\seqof{u x_i u}[i \in \Nat]$ of infixes of $v$, such that $x_i$ is a word that is
-    of size $\geq i$.
+    $\seqof{u x_i u}[i \in \Nat]$ of infixes of $v$, such that $x_i$ is a word
+    of size $\geq i$
+    and no shorter word $u y u$ is an infix of $u x_i u$.
     Because the finite infixes of $w$ (hence, of $v$) are \kl{well-quasi-ordered},

### 2025-10-13 src/introduction.tex
m

@@
 well-quasi-ordered as soon as the alphabet contains two distinct letters:
-for instance, the infinite sequence $a^n b^n a^n$ is \kl{well-quasi-ordered} by
+for instance, the infinite sequence $a b^n a$ is \kl{well-quasi-ordered} by
 the \kl{subword relation} but by neither the \kl{prefix relation}, nor the
@@
 descriptions of the \kl{well-quasi-ordered} languages, as well as tight bounds
-on their \kl{ordinal invariants}. In particular, we highlight the role of
+on their \kl{ordinal invariants}. In particular, we highlight the central role of
 \kl{bounded languages} as the one that are well-behaved with respect to being
-\kl{well-quasi-ordering} by the \kl{infix relation} \cref{infixes-bounded:sec}.
+\kl{well-quasi-ordering} by the \kl{infix relation} \cref{infixes-bounded:sec}.\todo[inline]{
+ i tried to rewrite it but im not sure i understand this sentence :)
+}
 

### 2025-10-13 src/prelims.tex

--- 2025-08-06 src/prelims.tex
@@
 $\seqof{T_i}$ then $\alpha_T = \sup \setof{\alpha_{T_i} + 1}{i \in \Nat}$. We
-say that $\alpha_T$ is the \emph{rank} of $T$.
+say that $\alpha_T$ is the \intro{rank} of $T$.
 
@@
 well-founded trees from $X$: the tree of \kl{bad sequences}, the tree of
-decreasing sequences, and the tree of \kl{antichains}. The rank of these
+decreasing sequences, and the tree of \kl{antichains}. The \kl{rank} of these
 respective trees are called respectively the \intro{maximal order type} of $X$

### 2025-10-13 src/prelims.tex
m

@@
 \kl{well-quasi-ordered} set $X$. 
-As an example, for $(\Nat, \leq)$, all bad sequences are descending and there are no antichains.
-Therefore the maximal order type and the ordinal height of $\Nat$ are $\omega$ and the ordinal width is 1.
+As an example, for $(\Nat, \leq)$, all bad sequences are descending and antichains have size at most $1$.
+Therefore the maximal order type and the ordinal height of $\Nat$ are $\omega$ and the ordinal width is $1$.
 \todo{Please check this!}

### 2025-10-13 src/infixes.tex
fix: correct typos and error noticed by Lia

@@
     there exists $x,y \in \Sigma^+$ of size 
-    at most $\max \setof{\card{v_i}}{1 \leq i \leq k}$
+    at most $\min \setof{\card{v_i}}{1 \leq i \leq k}$
     such that
@@
     In order to achieve this,
-    let us first define $s$ as the maximal size of a word $v_i$
+    let us first define $s$ as the minimal size of a word $v_i$
     ($1 \leq i \leq k$) and $u_j$ ($1 \leq j \leq k+1$).
@@
     $w[\vec{n_i}] \in \InfPeriodChain{x} u_{t} \InfPeriodChain{y}$.
+
+    \todo[inline]{Aliaume: proof of concept proof}
+    Let us now conclude that the same $x,y$ work for all $\vec{n} \in \Nat^k$.
 
@@
     Then $L$ is included in a finite union 
-    of products of \kl{chains}: \todo{reference}
+    of products of \kl{chains} for the \kl{prefix} and \kl{suffix} relations
+    thanks to
+    \cref{bounded-language:thm}:
     \begin{equation*}
@@
     \end{equation*}
-    Remark that the \kl{downwards closure} of a product of two \kl{chains}
-    is a finite union of products of two chains. \todo{reference/proof}.
+    Remark that if $S_i$ is a \kl{chain} for the \kl{suffix relation}
+    and $P_i$ is a \kl{chain} for the \kl{prefix relation},
+    then 
+    \begin{equation*}
+      \dwset[\infleq]{(S_i \cdot P_i)} = (\dwset[\suffleq]{S_i}) \cdot (\dwset[\prefleq]{P_i})
+        \quad .
+    \end{equation*}
+    Indeed, any \kl{infix} of a word in $S_i P_i$ can be split into
+    a \kl{suffix} of a word in $S_i$ and a \kl{prefix} of a word in $P_i$.
+    Conversely, any such concatenations are \kl{infixes} of a word in $S_i P_i$.
+
     As a consequence, we conclude that $\dwset[\infleq]{L}$ is itself included
     in a finite union of products of \kl{chains}.
-    Note that this also proves that $\dwset[\infleq]{L}$ is a \kl{bounded language},
-    hence that it is \kl{well-quasi-ordered} by the \kl{infix relation} 
+    Furthermore, by definition of \kl{bounded languages},
+    $\dwset[\infleq]{L}$ is also a \kl{bounded language}.
+    Hence, it is \kl{well-quasi-ordered} by the \kl{infix relation} 
     via
@@
 
-One may think that all \kl{downwards closed} languages for the \kl{infix
-relation} that are \kl{well-quasi-ordered} are \kl(language){bounded}. Note
-that this is what happens in the case of the \kl{subword embedding}, where any
-\kl{downwards closed} language for this relation is characterized by finitely
-many excluded \kl{subwords}, hence provides a \kl{bounded language} \todo{wrong? $\{a,b\}^*$ is downward closed but not bounded}. However, this
-is not the case for the \kl{infix relation}, as we will now illustrate with the
-following two examples.
+The \cref{bounded-wqo-dwclosed:cor} is reminiscent of a similar result for the
+\kl{subword embedding},  stipulating that for any language $L \subseteq
+\Sigma^*$, the \kl{downwards closure} $\dwset[\higleq]{L}$ is described using
+finitely many excluded \kl{subwords}, hence is regular. However, this is not
+the case for the \kl{infix relation}, even with \kl{bounded languages}, as we
+will now illustrate with the following example.
 
@@
     \label{dwclosed-wqo-not-finite-excl:ex}
-    Let $L \defined a^* b^* \cup b^* a^*$ \todo{However, this language is bounded, so this seems to be a non sequitur}. This language is \kl{downwards
+    Let $L \defined a^* b^* \cup b^* a^*$. This language is 
+    \kl(language){bounded}, is \kl{downwards
     closed} for the \kl{infix relation}, is \kl{well-quasi-ordered} for the
@@
 
-    %Let us prove that these are upper bounds for the \kl{ordinal invariants} of
-    $\LMorse$. 
-    The bound of the \kl{ordinal height} holds for any language $L$,
-    since a descending chain for $(\Sigma^*, \infleq)$ is isomorphic to one for $(\Nat, \leq)$ \todo{please check this is a good argument}.
-    For the \kl{maximal order type}, we remark that
-    the \kl{uniform recurrence} of $\ThueMorse$ means that
-    the maximal length of a \kl{bad sequence} is determined by its first element,
-    hence that it is at most $\omegaOrd$.
-    Finally, because the \kl{ordinal width} is at most the \kl{maximal order type} \todo{add reference},
-    we conclude that the \kl{ordinal width} is also at most $\omegaOrd$.
+    Let us prove that these are upper bounds for the \kl{ordinal invariants} of
+    $\LMorse$. The bound of the \kl{ordinal height} holds for any language $L$,
+    as the length of a decreasing sequence of words is bounded by the length of
+    its first element. For the \kl{maximal order type}, we remark that the
+    \kl{uniform recurrence} of $\ThueMorse$ means that the maximal length of a
+    \kl{bad sequence} is determined by its first element, hence that it is at
+    most $\omegaOrd$. Finally, because the \kl{ordinal width} is at most the
+    \kl{maximal order type} (as per \cref{ordinal-invariants:subsec}, using for instance the
+    results of \cite{kriz90b} or \cite[Theorem 3.8]{DZSCSC20} stating 
+    $\oType{X} \leq \oHeight{X} \oComProd \oWidth{X}$): we conclude
+    that the \kl{ordinal width} is also at most $\omegaOrd$.
 
@@
 case, a \kl{well-quasi-ordered} \kl{order ideal} for the \kl{infix relation} is
-the set of finite \kl{infixes} of a finite \todo{$\Sigma^\Nat$ is an infinite word}
-or bi-infinte word $w \in
+the set of finite \kl{infixes} of a finite, infinite,
+or bi-infinte word $w \in \Sigma^* \cup
 \Sigma^\Nat \cup \Sigma^\Rel $ (\cref{bi-infinite:lem}). Therefore, we will

### 2025-10-13 wqo-on-words.lncs.tex
fix: correct typos and error noticed by Lia

@@
 % math packages
-\usepackage{amssymb,amsmath,stmaryrd,thmtools,thm-restate,upgreek}
+\usepackage{amssymb,amsmath,amsthm,stmaryrd,thmtools,upgreek}
 
@@
 \Crefname{assumption}{Assumption}{Assumptions}
-\Crefname{conjecture}{Conjecture}{Conjectures}
 \usepackage[composition,hyperref,xcolor,cleveref]{knowledge}
@@
 \author{
-        Nathan Lhote\inst{3}\orcidID{0000-0003-3303-5368} \and
-        Aliaume Lopez\inst{2}\orcidID{0000-0002-4205-327X} \and
+        Nathan Lhote\inst{2}\orcidID{0000-0003-3303-5368} \and
+        Aliaume Lopez\inst{3}\orcidID{0000-0002-4205-327X} \and
         Lia Schütze\inst{1}\orcidID{0000-0003-4002-5491}}
@@
  \and
+Aix-Marseille University
+ \and
 INP Bordeaux, LaBRI, CNRS
- \and
-Aix-Marseille University
 }

### 2025-10-14 src/infixes.tex
new: proof sketch that everything works out

@@
 
-    \todo[inline]{Aliaume: proof of concept proof}
-    Let us now conclude that the same $x,y$ work for all $\vec{n} \in \Nat^k$.
+    \todo[inline]{Aliaume: proof of concept proof} Let us now conclude that the
+    same $x,y$ work for all $\vec{n} \in \Nat^k$. The reasoning is that if
+    $w[\vec{n}]$ is of the form $u_1 x^p x'$ where $x' \prefleq x$, and $x$ is
+    primitive (not a power of a smaller word), then all the iterated words
+    $v_i^{n_i}$ must be powers of a conjugate of $x$ (cyclic shift of $x$), and
+    all the $u_i$'s must be such that the concatenation of the $u_i$'s and
+    $v_i$'s align properly with $x$. Since the length of $x$ is at most $\min
+    \setof{\card{v_i}}{1 \leq i \leq k}$, we conclude that the $v_i$'s
+    themselves are powers of a conjugate of $x$, and that the $u_i$'s are
+    suffixes or prefixes of $x$. Now, we conclude that all the words
+    $w[\vec{n}]$ are of the correct form. Similar reasoning applies for the
+    other two cases.
 

### 2025-10-14 src/prelims.tex
fix: mot and width of omega

@@
 
-\AP
-Let $(X, \leq)$ be a \kl{well-quasi-ordered} set. One can define three
+\AP Let $(X, \leq)$ be a \kl{well-quasi-ordered} set. One can define three
 well-founded trees from $X$: the tree of \kl{bad sequences}, the tree of
@@
 parameters are called the \intro{ordinal invariants} of a
-\kl{well-quasi-ordered} set $X$. 
-As an example, for $(\Nat, \leq)$, all bad sequences are descending and antichains have size at most $1$.
-Therefore the maximal order type and the ordinal height of $\Nat$ are $\omega$ and the ordinal width is $1$.
-\todo{Please check this!}
-We refer to the survey of \cite{DZSCSC20} for
-a detail discussion on these concepts and their computation on specific classes
-of well-quasi-ordered sets.
+\kl{well-quasi-ordered} set $X$. As an example, for $(\Nat, \leq)$, all bad
+sequences are descending and antichains have size at most $1$. In fact, $(\Nat,
+\leq)$ is itself an \kl{ordinal}, namely $\omegaOrd$. Hence it is its own
+\kl{maximal order type} and \kl{ordinal height}, and its \kl{ordinal width} is
+$1$ since it contains no antichain of size $2$. We refer to the survey of
+\cite{DZSCSC20} for a detail discussion on these concepts and their computation
+on specific classes of well-quasi-ordered sets.
 

### 2025-10-14 src/infixes.tex
Theorem 4 todo

@@
 A simplified version of the embedding defined in \cref{infix-embedding:thm} is illustrated
-for the \kl{subword relation} in \cref{infix-embedding:fig}.
+for the \kl{subword relation} in \cref{infix-embedding:fig} page \pageref{infix-embedding:fig}.
 \begin{lemma}{\cite[Lemma 5.1]{DBLP:journals/ita/Kuske06}}
@@
 \end{lemma}
-\begin{figure}
-    \centering
-    \includestandalone[width=\linewidth]{fig/infix-encoding-standalone}
-    \caption{Representation of the \kl{subword relation} for $\set{a,b}^*$
-        inside the \kl{infix relation} for $\set{a,b,\#}^*$
-        using a simplified version of \cref{infix-embedding:thm}, restricted to words
-        of length at most $3$. 
-    }
-    \label{infix-embedding:fig}
-\end{figure}
 
@@
 
-\begin{theorem}[restate=bounded-language:thm,label=bounded-language:thm]
+\begin{theorem}
     \label{bounded-language:thm}
+    \proofref{bounded-language:thm}
     Let $L$ be a \kl{bounded language} of $\Sigma^*$. Then,
@@
     \label{inf-period-chain:lem}
+    \proofref{inf-period-chain:lem}
     Let $x \in \Sigma^+$ be a word.
@@
 \end{lemma}
-\begin{proof}
-    Let $x \in \Sigma^+$ be a word, and let $P_x$ be the (finite) set 
-    of all \kl{prefixes} of $x$, and $S_x$ be the (finite)
-    set of all \kl{suffixes} of $x$.
-    Assume that $w \in \InfPeriodChain{x}$, then $w = u x^p v$ for some
-    $u \in S_x$, $v \in P_x$, and $p \in \Nat$.
-    We have proven that
-    \begin{equation*}
-        \InfPeriodChain{x} \subseteq \bigcup_{u \in P_x} \bigcup_{v \in S_x} u x^* v
-        \quad .
-    \end{equation*}
-
-    Let us now demonstrate that for all $(u,v) \in S_x \times P_x$, the
-    language $u x^* v$ is a \kl{chain} for the \kl{infix}, \kl{suffix} and \kl{prefix} relations.
-    To that end,
-    let $(u,v) \in S_x \times P_x$ and $\ell, k \in \Nat$ be such that $\ell <
-    k$, let us prove that $u x^\ell v \infleq u x^k  v$. Because $v \prefleq
-    x$, we know that there exists $w$ such that $vw = x$. In particular,
-    $ux^\ell vw = u x^{\ell + 1}$, and because $\ell < k$, we conclude that $u
-    x^{\ell + 1} \prefleq u x^k v$. By transitivity, $u x^\ell v \prefleq u x^k
-    v$, and \emph{a fortiori}, $u x^\ell v \infleq u x^k v$. 
-    Similarly, because $u \suffleq x$,  there exists $w$ such that $wu  = x$, 
-    and we conclude that $u x^{\ell} v \suffleq w u x^\ell v = x^{\ell + 1} v \suffleq u x^k v$.
-\end{proof}
 
@@
 
-We prove if the upcoming theorem that the status of the \kl{Thue-Morse
-sequence} is actually representative of most \kl{downwards closed} languages
-for the \kl{infix relation}. Indeed, we show that the \kl{ordinal invariants}
-of such languages are relatively small, and the proof of this fact relies on
-the connection between \kl{well-quasi-ordered} languages and \kl{uniformly
-recurrent} words.
+\AP We prove in the upcoming \cref{dw-closed-wqo-charac:thm} that the status of
+the \kl{Thue-Morse sequence} is actually representative of \kl{downwards
+closed} languages for the \kl{infix relation}. To that end, let us introduce
+the notation $\intro*\infset{w}$ for the set of finite \kl{infixes} of a
+(possibly infinite or bi-infinite) word $w \in \Sigma^* \cup \Sigma^\Nat \cup
+\Sigma^\Rel$. We say that an infinite word $w \in \Sigma^\Nat$ is
+\intro{ultimately uniformly recurrent} if there exists a bound $N_0 \in \Nat$
+such that $w_{\geq N_0}$ is \kl{uniformly recurrent}. We extend this notion to
+finite words by considering that they all are \kl{ultimately uniformly
+recurrent}, and to bi-infinite words by considering that they are
+\kl{ultimately uniformly recurrent} if and only if both their left-infinite and
+right-infinite parts are.
 
 \begin{theorem}
-    \label{small-ordinal-invariants:thm}
+    \label{dw-closed-wqo-charac:thm}
     Let $L$ be a \kl{well-quasi-ordered} language for the \kl{infix relation}
     that is \kl{downwards closed}.
+    Then, there exist finitely many \kl{ultimately uniformly recurrent words} 
+    $w_1, \dots, w_n \in \Sigma^* \cup 
+    \Sigma^\Nat \cup \Sigma^\Rel$ such that
+    $L = \bigcup_{i = 1}^n \infset{w_i}$.
+\end{theorem}
+
+Thanks to \cref{dw-closed-wqo-charac:thm}, and by analysing 
+the \kl{ordinal invariants} of \kl{infixes} of an \kl{ultimately uniformly recurrent}
+infinite word $w$ (\cref{ultimately-uniformly-recurrent:lem}),
+we conclude that the \kl{ordinal invariants} of a \kl{well-quasi-ordered}
+\kl{downwards closed} language are relatively small.
+
+\begin{corollary}
+  \label{small-ordinal-invariants:thm}
     Then, the \kl{maximal order type} of $L$ is strictly less than $\omegaOrd^3$,
@@
     Furthermore, those bounds are tight.
-\end{theorem}
+\end{corollary}
 
-\AP A subset $I \subseteq X$ is \intro(subset){directed} if, for every $x,y \in
-I$, there exists $z \in I$ such that $x \leq z$ and $y \leq z$. Given a
+\AP To connect infixes of a (bi)-infinite word to \kl{downwards closed}
+languages, a useful notion is that of \kl(subset){directed} sets. A subset $I
+\subseteq X$ is \intro(subset){directed} if, for every $x,y \in I$, there
+exists $z \in I$ such that $x \leq z$ and $y \leq z$. Given a
 \kl{well-quasi-order} $(X, \leq)$, one can always decompose $X$ into a finite
@@
 case, a \kl{well-quasi-ordered} \kl{order ideal} for the \kl{infix relation} is
-the set of finite \kl{infixes} of a finite, infinite,
-or bi-infinte word $w \in \Sigma^* \cup
-\Sigma^\Nat \cup \Sigma^\Rel $ (\cref{bi-infinite:lem}). Therefore, we will
-first study languages that are described as the set of finite \kl{infixes} of
-some (bi-)infinite word (\cref{ultimately-uniformly-recurrent:lem}). To that
-end, let us introduce the notation $\intro*\infset{w}$ for the set of finite
-\kl{infixes} of a (possibly infinite or bi-infinite) word $w$.
+the set of finite \kl{infixes} of a finite, infinite, or bi-infinte word $w \in
+\Sigma^* \cup \Sigma^\Nat \cup \Sigma^\Rel $ (\cref{bi-infinite:lem}).
 
@@
     is similar, but will result in a finite word.
+    \todo{aliaume: do we need wqo here? the proof
+      should go through without it: the sequence 
+      $u_i$ is already increasing for infix}
 
@@
 \end{proof}
-
-\AP We say that an infinite word $w \in \Sigma^\Nat$ is \intro{ultimately
-uniformly recurrent} if there exists a bound $N_0 \in \Nat$ such that $w_{\geq
-N_0}$ is \kl{uniformly recurrent}.
 
@@
 \begin{lemma}
+  \label{bi-infinite-uur:lem}
+    Let $w \in \Sigma^\Rel$ be a bi-infinite word. Then, the set of finite
+    infixes of $w$ is \kl{well-quasi-ordered} for the \kl{infix relation} if
+    and only if $w$ is \kl{ultimately uniformly recurrent} as a bi-infinite word.
+\end{lemma}
+\begin{proof}
+  Given a bi-infinite word $w \in \Sigma^{\Rel}$, we can consider $w_+ \in
+\Sigma^\Nat$ and $w_- \in \Sigma^\Nat$ the two infinite words obtained as
+follows: for all $i \in \Nat$, $(w_+)_i = w(i)$ and $(w_-)_i = w(-i)$. Note
+that the two share the letter at position $0$.
+
+    Assume that $w_+$ and $w_-$ are \kl{ultimately uniformly recurrent}. Let us
+    write $\infset{w}$ the set of finite infixes of $w$. Consider an infinite
+    sequence of words $\seqof{u_i}[i \in \Nat]$ in $\infset{w}$. If there is an
+    infinite subsequence of words that are all in $\infset{w_+}$, then there
+    exists an increasing pair of indices $i < j$ such that $u_i \infleq
+    u_j$ because \cref{uniformly-recurrent:lem} applies to $w_+$. Similarly, if
+    there is an infinite subsequence of words that are all in $\infset{w_-}$,
+    then there exists an increasing pair of indices $i < j$ such that $u_i
+    \infleq u_j$ because \cref{uniformly-recurrent:lem} applies to $w_-$ (and
+    the \kl{infix relation} is compatible with mirroring). Otherwise, one can
+    assume without loss of generality that all words in the sequence have a
+    starting position in $w_-$ and an ending position in $w_+$. In this case,
+    let us write $(k_i,l_i) \in \Nat^2$ the pair of indices such that $u_i$ is the infix
+    of $w$ that starts at position $-k_i$ of $w$ (i.e., $k_i$ of $w_-$) and
+    ends at position $l_i$ of $w$ (i.e., $l_i$ of $w_+$).
+    Because $\Nat^2$ is a \kl{well-quasi-ordering} with the product ordering,
+    there exists $i < j$ such that $k_i \leq k_j$ and $l_i \leq l_j$, 
+    in particular, $u_i \infleq u_j$. We have proven that every infinite
+    sequence of words in $\infset{w}$ is \kl(wqo){good}, hence $\infset{w}$ is
+    \kl{well-quasi-ordered}.
+
+    Conversely, assume that $\infset{w}$ is \kl{well-quasi-ordered}. In
+    particular, the subset $\infset{w_+} \subseteq \infset{w}$ is \kl{well-quasi-ordered}.
+    Similarly, $\infset{w_-}$ is \kl{well-quasi-ordered} because the \kl{infix
+    relation} is compatible with mirroring. Applying
+    \cref{ultimately-uniformly-recurrent:lem}, we conclude that both are
+    \kl{ultimately uniformly recurrent} words.
+
+\end{proof}
+
+We are now ready to conclude the proof of \cref{dw-closed-wqo-charac:thm}.
+\begin{proofof}{dw-closed-wqo-charac:thm}
+  It is clear that the set of finite \kl{infixes} of a finite, infinite or
+  bi-infinite \kl{ultimately uniformly recurrent} word is \kl{well-quasi-ordered}
+  for the \kl{infix relation} thanks to \cref{ultimately-uniformly-recurrent:lem}.
+
+  Conversely, let us consider a \kl{well-quasi-ordered} language $L$ that is
+  \kl{downwards closed} for the \kl{infix relation}. Because it is a
+  \kl{well-quasi-ordered} set, it can be written as a finite union of \kl{order
+  ideals} $L = \bigcup_{i = 1}^n L_i$.
+
+  For every such \kl(order){ideal} $L_i$, we can apply
+  \cref{bi-infinite:lem}, and conclude that $L_i$ is the set of finite
+  \kl{infixes} of a finite, infinite or bi-infinite word $w_i$. 
+  Because the languages $L_i$ are \kl{well-quasi-ordered}, we can apply
+  \cref{ultimately-uniformly-recurrent:lem}, and conclude that $w_i$ is
+  \kl{ultimately uniformly recurrent}.
+\end{proofof}
+
+Let us now analyse the \kl{ordinal invariants} of the set of finite
+\kl{infixes} of an \kl{ultimately uniformly recurrent} infinite word, in order
+to derive the bounds of \cref{small-ordinal-invariants:thm}.
+
+\begin{lemma}
     \label{small-ordinal-invariants:lem}
@@
 
-The behaviour of bi-infinite words is very similar to the behaviour of infinite
-ones. Given a bi-infinite word $w \in \Sigma^{\Rel}$, we can consider $w_+ \in
-\Sigma^\Nat$ and $w_- \in \Sigma^\Nat$ the two infinite words obtained as
-follows: for all $i \in \Nat$, $(w_+)_i = w(i)$ and $(w_-)_i = w(-i)$. Note
-that the two share the letter at position $0$.
-
 \begin{lemma}
@@
 \begin{proof}
-    Assume that $w_+$ and $w_-$ are \kl{ultimately uniformly recurrent}. Let us
-    write $\infset{w}$ the set of finite infixes of $w$. Consider an infinite
-    sequence of words $\seqof{u_i}[i \in \Nat]$ in $\infset{w}$. If there is an
-    infinite subsequence of words that are all in $\infset{w_+}$, then there
-    exists an increasing pair of indices $i < j$ such that $u_i \infleq
-    u_j$ because \cref{uniformly-recurrent:lem} applies to $w_+$. Similarly, if
-    there is an infinite subsequence of words that are all in $\infset{w_-}$,
-    then there exists an increasing pair of indices $i < j$ such that $u_i
-    \infleq u_j$ because \cref{uniformly-recurrent:lem} applies to $w_-$ (and
-    the \kl{infix relation} is compatible with mirroring). Otherwise, one can
-    assume without loss of generality that all words in the sequence have a
-    starting position in $w_-$ and an ending position in $w_+$. In this case,
-    let us write $(k_i,l_i) \in \Nat^2$ the pair of indices such that $u_i$ is the infix
-    of $w$ that starts at position $-k_i$ of $w$ (i.e., $k_i$ of $w_-$) and
-    ends at position $l_i$ of $w$ (i.e., $l_i$ of $w_+$).
-    Because $\Nat^2$ is a \kl{well-quasi-ordering} with the product ordering,
-    there exists $i < j$ such that $k_i \leq k_j$ and $l_i \leq l_j$, 
-    in particular, $u_i \infleq u_j$. We have proven that every infinite
-    sequence of words in $\infset{w}$ is \kl(wqo){good}, hence $\infset{w}$ is
-    \kl{well-quasi-ordered}.
-
-    Conversely, assume that $\infset{w}$ is \kl{well-quasi-ordered}. In
-    particular, the subset $\infset{w_+} \subseteq \infset{w}$ is \kl{well-quasi-ordered}.
-    Similarly, $\infset{w_-}$ is \kl{well-quasi-ordered} because the \kl{infix
-    relation} is compatible with mirroring. Applying
-    \cref{ultimately-uniformly-recurrent:lem}, we conclude that both are
-    \kl{ultimately uniformly recurrent} words.
+  Given a bi-infinite word $w \in \Sigma^{\Rel}$, recall that we can consider $w_+ \in
+\Sigma^\Nat$ and $w_- \in \Sigma^\Nat$ the two infinite words obtained as
+follows: for all $i \in \Nat$, $(w_+)_i = w(i)$ and $(w_-)_i = w(-i)$. Note
+that the two share the letter at position $0$.
 
@@
     \kl{well-quasi-ordered} language $L$ that is \kl{downwards closed} for the
-    \kl{infix relation}. Because it is a \kl{well-quasi-ordered} set, it can be
-    written as a finite union of \kl{order ideals} $L = \bigcup_{i = 1}^n L_i$.
-
-    For every such \kl(order){ideal} $L_i$, we can apply
-    \cref{bi-infinite:lem}, and conclude that $L_i$ is the set of finite
-    \kl{infixes} of a finite, infinite or bi-infinite word $w_i$. We can then
+    \kl{infix relation}. Applying 
+    \cref{dw-closed-wqo-charac:thm}, we can write $L = \bigcup_{i = 1}^n L_i$ where
+    each $L_i$ is the set of finite \kl{infixes} of a finite, infinite or
+    bi-infinite \kl{ultimately uniformly recurrent} word $w_i$.
+   We can then
     directly conclude that $\oWidth{L_i}$ less than $\omegaOrd$ (in the case of
@@
     Now, $\oWidth{L} \leq \sum_{i = 1}^n \oWidth{L_i} < 3 \cdot \omegaOrd <
-    \omegaOrd^2$. We conclude that the \kl{ordinal width} of $L$ is less than
+    \omegaOrd^2$ \todo{if inequality 2 holds, why do we not stop there?}. We conclude that the \kl{ordinal width} of $L$ is less than
     $\omegaOrd^2$. Finally, the inequality $\oType{L} \leq \oWidth{L} \oComProd
@@
 able to find in the literature whether one can decide \kl{ultimate uniform
-recurrence}, but we can prove it for a restricted class of \kl{morphic
-sequences}, that are called \intro{pure morphic sequences}, and are precisely
-those that do not use a final \kl(words){homomorphism} $h$. Following the same proof
-as for \cref{automatic-wqo:thm}, one deduces from the upcoming
-\cref{morphic-uur:lem} the following \cref{morphic-wqo:cor}. We actually
-conjecture that \cref{morphic-uur:lem} can be extended to all \kl{morphic
-sequences} \cref{morphic-uur:conj}.
-
-\begin{lemma}
-  \label{morphic-uur:lem}
-    Given a \kl{pure morphic sequence} $w \in \Sigma^{\Nat}$, one can decide
-    whether it is \kl{ultimately uniformly recurrent}.
-\end{lemma}
-\begin{proof}
-  On the one hand, one can guess a bound $N_0$ and check that 
-  the suffix of the word $w_{\geq N_0}$ is \kl{uniformly recurrent},
-  since the suffix of a \kl{morphic sequence} is also a \kl{morphic sequence},
-  and \kl{uniform recurrence} is decidable for \kl{morphic sequences}
-  \cite[Theorem 2]{DURAND13}.
-
-  Let us now prove that a specific choice of $N_0$ is sufficient. Let us write
-  $w \defined f^\omega(a)$ where $f \colon \Sigma^* \to \Sigma^*$ is a
-  \kl(words){morphism} and $a \in \Sigma$ is a letter such that $f(a)$ starts
-  with $a$. Let us also write $f(a) = a s$, where $s \in \Sigma^+$ is a
-  non-empty word. Notice that $w = a s f(s) f^2(s) \cdots$. Without loss of
-  generality, one can assume that $|f^n(s)| \to \infty$ when $n \to \infty$,
-  otherwise $w$ is ultimately periodic, and therefore \kl{ultimately uniformly
-  recurrent}. Since one can algorithmically detect if this is the case, we
-  assume that it is not.
-
-  Because $\Sigma$ is finite, there exists a finite set of letters $A \subseteq
-  \Sigma$ and a suffix $w'$ of $w$ such that $w' \in A^*$ and every letter of
-  $w'$ appears infinitely often in $w'$. Furthermore, one can compute this set
-  $A$ and a bound $N_0$ such that $f^n(s) \in A^*$ for every $n \geq N_0$. We
-  claim that $w$ is \kl{ultimately uniformly recurrent} if and only if
-  $f^{N_0}(s) f^{N_0 + 1}(s) \cdots$ is \kl{uniformly recurrent}.
-  One direction being obvious, we assume that $w$ is \kl{ultimately uniformly
-  recurrent}, and let us prove that $f^{N_0}(s) f^{N_0 + 1}(s) \cdots$ is
-  \kl{uniformly recurrent}.
-
-  Let us assume that $w'' \defined f^{N_1}(s) f^{N_1 + 1}(s) \cdots$ where $N_1
-  \geq N_0$ is \kl{uniformly recurrent}. Let us show that $f^{N_0}(s)$ is an
-  \kl{infix} of $w''$ if and only if there exists a letter $b$ in $s$ such that
-  $f^{N_0}(s) \infleq f^{k} (b)$ for some finite number $k$.
-
-  Assume that $f^{N_0}(s)$ is an \kl{infix} of $w''$. Recall that $|f^n(s)| \to
-  \infty$ when $n \to \infty$, hence, there must exist a letter $b$ in $s$ such
-  that $|f^n(b)| \to \infty$ when $n \to \infty$. Because every letter of
-  $f^{N_0}(s)$ appears infinitely often in $w''$, $f^n(b)$ is a growing
-  sequence of infixes of $w''$. Since $w''$ is \kl{uniformly recurrent}, there
-  exists a bound $K$ such that $f^{N_0}(s)$ is an \kl{infix} of every $f^k(b)$
-  for $k \geq K$. Conversely, if there exists a letter $b$ in $s$ such that
-  $f^{N_0}(s) \infleq f^{k} (b)$ for some finite number $k$, then $f^{N_0}(s)$
-  is an \kl{infix} of $w''$ because $f^n(b)$ is an infix of $w''$ for every $n
-  \in \Nat$.
-
-  Because $w''$ is \kl{uniformly recurrent}, there exists a bound $K$ such that
-  $f^{N_1}(s) \infleq f^k(b)$ for every $k \geq K$.
-  \todo[inline]{Cannot finish the proof...}
-
-\end{proof}
-
-\begin{corollary}
-    \label{morphic-wqo:cor}
-    Given a \kl{sequence representation} of a language $L \subseteq
-    \Sigma^*$ where all infinite words are \kl{pure morphic sequences}, one can
-    decide whether $L$ is \kl{well-quasi-ordered} for the \kl{infix relation}.
-\end{corollary}
+recurrence}, but conjecture that it is the case, which would allow us to
+decide whether a language represented by \kl{morphic sequences} is
+\kl{well-quasi-ordered} for the \kl{infix relation}.
 

### 2025-10-14 src/infixes.tex
new: maybe working proof

--- 2025-10-14 src/infixes.tex
@@
 \begin{lemma}
+  \label{pure-morphic-uur:lem}
+  Let $w \in \Sigma^{\Nat}$ be a \kl{pure morphic sequence}
+  generated by a morphism $f \colon \Sigma^* \to \Sigma^*$ and a letter $a \in
+  \Sigma$ such that $f(n)$ starts with $a$.
+  Assume furthermore that $f$ is \kl(morphic){growing}, i.e., that 
+  $f(a) = a s$ where $s \in \Sigma^+$ is a non-empty word
+  such that $|f^n(s)| \to \infty$ when $n \to \infty$.
+  Then, the following are equivalent:
+  \begin{itemize}
+    \item $w$ is \kl{ultimately uniformly recurrent},
+    \item \begin{enumerate}
+          \item 
+            for every letters $b,c \in \Sigma$,
+            if $b$ is \kl(morphic){growing} (i.e., $|f^n(b)| \to \infty$ when $n \to \infty$)
+            and both $b$ and $c$ appear infinitely often in $w$, 
+            then $c$ appears in $f^m(b)$ for some $m \in \Nat$,
+          \item
+            there are only finitely many 
+            \kl(worphic){bounded words}
+            (i.e., words $u \in \Sigma^*$ such that $|f^n(u)|$ is bounded when $n \to \infty$)
+            that appear as infixes of $w$.
+    \end{enumerate}
+  \end{itemize}
+\end{lemma}
+\begin{proof}
+  Assume that $w$ is \kl{ultimately uniformly recurrent}.
+  We have $N_0$ such that $w_{\geq N_0}$ is \kl{uniformly recurrent}.
+  Let $b,c \in \Sigma$ be two letters such that $b$ is \kl(morphic){growing}
+  and $c$ appears infinitely often in $w$.
+  Then, $c$ appears in every large enough infix of $w_{\geq N_0}$.
+  Because $b$ appears infinitely often in $w$, 
+  $f^n(b)$ is an infix of $w_{\geq N_0}$ for all $n$'s,
+  and in particular $c$ appears in $f^n(c)$ for some $n$ large enough so that 
+  $|f^n(b)|$ is larger than the bound of \kl{uniform recurrence} of $w_{\geq N_0}$.
+
+  Let us now prove the second item. Let $u$ be a \kl(worphic){bounded word}
+  that appears as an infix of $w$, by construction, it cannot contain
+  any \kl(morphic){growing} letter, otherwise $|f^n(u)|$ would not be bounded.
+  Because $w_{\geq N_0}$ is \kl{uniformly recurrent}, there exists a bound $M_0$ 
+  on the distance between two consecutive \kl(morphic){growing} letters in 
+  $w_{\geq N_0}$. Now, because $f$ is \kl(morphic){growing},
+  there exists a \kl(morphic){growing} letter $b \in \Sigma$ that 
+  appears infinitely often in $w$. As a consequence, the 
+  \kl(morphic){bounded words} that appear as infixes of $w$ have length 
+  at most $N_0 + M_0$.
+
+  Conversely, assume that the two items are satisfied. Let us prove that $w$ is
+  \kl{ultimately uniformly recurrent}. First, there exists $N_0$ such that
+  every letter that appears in $w_{\geq N_0}$ appears infinitely often in
+  $w_{\geq N_0}$ because $\Sigma$ is finite.
+
+  \todo[inline]{finish the proof}
+  following~\cite[Theorem 4]{NIPR09} we argue that $w_{\geq N_0}$ is
+  \kl{uniformly recurrent}....
+
+\end{proof}
+
+\begin{lemma}
   \label{morphic-uur:lem}
@@
 \begin{proof}
+  Let us first assume that the morphism $f$ that generates $w$ is
+  \kl(morphic){growing}. If it is not the case, then $w$ is 
+  ultimately periodic, and a fortiori \kl{ultimately uniformly recurrent}.
+  It is possible to decide whether $f$ is \kl(morphic){growing}
+  by standard techniques on morphic words.
+
   On the one hand, one can guess a bound $N_0$ and check that 
@@
 
-  Let us now prove that a specific choice of $N_0$ is sufficient. Let us write
-  $w \defined f^\omega(a)$ where $f \colon \Sigma^* \to \Sigma^*$ is a
-  \kl(words){morphism} and $a \in \Sigma$ is a letter such that $f(a)$ starts
-  with $a$. Let us also write $f(a) = a s$, where $s \in \Sigma^+$ is a
-  non-empty word. Notice that $w = a s f(s) f^2(s) \cdots$. Without loss of
-  generality, one can assume that $|f^n(s)| \to \infty$ when $n \to \infty$,
-  otherwise $w$ is ultimately periodic, and therefore \kl{ultimately uniformly
-  recurrent}. Since one can algorithmically detect if this is the case, we
-  assume that it is not.
-
-  Because $\Sigma$ is finite, there exists a finite set of letters $A \subseteq
-  \Sigma$ and a suffix $w'$ of $w$ such that $w' \in A^*$ and every letter of
-  $w'$ appears infinitely often in $w'$. Furthermore, one can compute this set
-  $A$ and a bound $N_0$ such that $f^n(s) \in A^*$ for every $n \geq N_0$. We
-  claim that $w$ is \kl{ultimately uniformly recurrent} if and only if
-  $f^{N_0}(s) f^{N_0 + 1}(s) \cdots$ is \kl{uniformly recurrent}.
-  One direction being obvious, we assume that $w$ is \kl{ultimately uniformly
-  recurrent}, and let us prove that $f^{N_0}(s) f^{N_0 + 1}(s) \cdots$ is
-  \kl{uniformly recurrent}.
-
-  Let us assume that $w'' \defined f^{N_1}(s) f^{N_1 + 1}(s) \cdots$ where $N_1
-  \geq N_0$ is \kl{uniformly recurrent}. Let us show that $f^{N_0}(s)$ is an
-  \kl{infix} of $w''$ if and only if there exists a letter $b$ in $s$ such that
-  $f^{N_0}(s) \infleq f^{k} (b)$ for some finite number $k$.
-
-  Assume that $f^{N_0}(s)$ is an \kl{infix} of $w''$. Recall that $|f^n(s)| \to
-  \infty$ when $n \to \infty$, hence, there must exist a letter $b$ in $s$ such
-  that $|f^n(b)| \to \infty$ when $n \to \infty$. Because every letter of
-  $f^{N_0}(s)$ appears infinitely often in $w''$, $f^n(b)$ is a growing
-  sequence of infixes of $w''$. Since $w''$ is \kl{uniformly recurrent}, there
-  exists a bound $K$ such that $f^{N_0}(s)$ is an \kl{infix} of every $f^k(b)$
-  for $k \geq K$. Conversely, if there exists a letter $b$ in $s$ such that
-  $f^{N_0}(s) \infleq f^{k} (b)$ for some finite number $k$, then $f^{N_0}(s)$
-  is an \kl{infix} of $w''$ because $f^n(b)$ is an infix of $w''$ for every $n
-  \in \Nat$.
-
-  Because $w''$ is \kl{uniformly recurrent}, there exists a bound $K$ such that
-  $f^{N_1}(s) \infleq f^k(b)$ for every $k \geq K$.
-  \todo[inline]{Cannot finish the proof...}
+  On the other hand, one can leverage the characterization of
+  \cref{pure-morphic-uur:lem}. Indeed, it is shown in \cite[Proposition
+  5]{NIPR09} that if the second item of \cref{pure-morphic-uur:lem} is not
+  satisfied, then there exists a finite witness of that fact that can be
+  computed from the morphism $f$. Similarly, to witness that the first item of
+  \cref{pure-morphic-uur:lem} is not satisfied, one can guess two letters $a,b
+  \in \Sigma$ and check that $a$ is \kl(morphic){growing}, both $a$ and $b$
+  appear infinitely often in $w$, and that $b$ does not appear in $f^n(a)$ for
+  any $n \in \Nat$. The first two conditions can be checked effectively, and
+  the last one can be checked by noticing that if $b$ appears in $f^n(a)$ for
+  some $n \in \Nat$, then it also appears in $f^m(a)$ for some $m \leq
+  |\Sigma|$.
 

### 2025-10-15 src/introduction.tex
todos: deleting + answering
new: isolating the u.u.r words as a theorem for the intro

@@
 % LTeX: language=en-GB 
-% !TeX root=../wqo-on-words.tex
+% !TeX root=../wqo-on-words.lncs.tex
 \section{Introduction}
@@
   \centering
-  \begin{tabular}{c|c|c|c}
+  \setlength{\tabcolsep}{6pt}
+  \begin{tabular}{c|p{6cm}|c|c}
       \toprule
@@
       \midrule
-      arbitrary & \cref{infix-embedding:thm} & $< \omegaOne$  & $< \omegaOne$ \\
-      bounded & \cref{bounded-language:thm} & $< \omegaOrd^2$ & $< \omegaOrd^3$ \\
-      downwards closed & \cref{small-ordinal-invariants:thm} & $< \omegaOrd^2$ & $< \omegaOrd^3$ \\
+      arbitrary & \cref{infix-embedding:thm}: countable well-quasi orders
+                with finite initial segments & $< \omegaOne$  & $< \omegaOne$ \\
+      \addlinespace
+      \kl(language){bounded} & \cref{bounded-language:thm}: finite union of products of prefix and suffix chains& $< \omegaOrd^2$ & $< \omegaOrd^3$ \\
+      \addlinespace
+      \kl{downwards closed}  & \cref{dw-closed-wqo-charac:thm}: finite union of infixes of \kl{ultimately uniformly recurrent words} & $< \omegaOrd^2$ & $< \omegaOrd^3$ \\
       \bottomrule
     \end{tabular}
-    \caption{Summary of results for the \kl{infix relation}.}
+    \caption{Summary of results for the \kl{infix relation},
+    the bounds on $\oWidth{L}$ and $\oType{L}$ are tight, and respectively 
+    proven in
+    \cref{ordinal-invariants-bounded:cor}
+    and 
+  \cref{small-ordinal-invariants:thm}.}
     \label{infixes-summary:fig}
@@
 \cref{amalgamation-systems:subsec}.
-
-\todo[inline]{
-  rework the content so that cited things exists and/or are 
-  corollaries and theorems (mainly). Also, add results to prove that 
-  bounds are tight!
-}
 
@@
 descriptions of the \kl{well-quasi-ordered} languages, as well as tight bounds
-on their \kl{ordinal invariants}. In particular, we highlight the central role of
-\kl{bounded languages} as the one that are well-behaved with respect to being
-\kl{well-quasi-ordering} by the \kl{infix relation} \cref{infixes-bounded:sec}.\todo[inline]{
- i tried to rewrite it but im not sure i understand this sentence :)
-}
+on their \kl{ordinal invariants}.
 
-\paragraph*{Outline} 
-We introduce in \cref{prelims:sec} the
-necessary background on \kl{well-quasi-orders} and \kl{ordinal invariants}.
-In
-\cref{prefixes:sec}, which is relatively
-self-contained, we study the \kl{prefix relation} and prove in
-\cref{prefixes:thm} the characterization of \kl{well-quasi-ordered}
-languages by the \kl{prefix relation}. In
-\cref{infixes-bounded:sec}, we
-obtain the \kl[infix relation]{infix} analogue of \cref{prefixes:thm}
-specifically for \kl{bounded languages}
-(\cref{bounded-language:thm}). 
-In \cref{infixes-dwclosed:sec}, we study the \kl{downwards closed}
-languages, and compute bounds on their \kl{ordinal invariants} in \cref{small-ordinal-invariants:thm}.
-Finally, 
-we generalize these results to all
-\kl{amalgamation systems} in \cref{infixes-amalgamation:sec}
-in
-(\cref{infix-amalgamation:thm}),
-and provide a decision procedure for checking whether a language is
-\kl{well-quasi-ordered} by the \kl{infix relation} (resp. \kl{prefix} and \kl{suffix}) in
-this context (\cref{infix-wqo-is-emptiness:thm}).
+\paragraph*{Outline} We introduce in \cref{prelims:sec} the necessary
+background on \kl{well-quasi-orders} and \kl{ordinal invariants}. In
+\cref{prefixes:sec}, which is relatively self-contained, we study the
+\kl{prefix relation} and prove in \cref{prefixes:thm} the characterization of
+\kl{well-quasi-ordered} languages by the \kl{prefix relation}. In
+\cref{infixes-bounded:sec}, we obtain the \kl[infix relation]{infix} analogue
+of \cref{prefixes:thm} specifically for \kl{bounded languages}
+(\cref{bounded-language:thm}). In \cref{infixes-dwclosed:sec}, we study the
+\kl{downwards closed} languages, characterize them using a notion of
+\kl{ultimately uniformly recurrent words} borrowed from symbolic dynamics
+(\cref{dw-closed-wqo-charac:thm}), and compute bounds on their \kl{ordinal
+invariants} in \cref{small-ordinal-invariants:thm}. Finally, we generalize
+these results to all \kl{amalgamation systems} in
+\cref{infixes-amalgamation:sec} in (\cref{infix-amalgamation:thm}), and provide
+a decision procedure for checking whether a language is \kl{well-quasi-ordered}
+by the \kl{infix relation} (resp. \kl{prefix} and \kl{suffix}) in this context
+(\cref{infix-wqo-is-emptiness:thm}).
 

### 2025-10-15 src/amalgam.tex
new: moving stuff to appendix

@@
 \cref{pumping-periods:lem} talks about. At the core of a pumping argument,
-%there is a notion of a \emph{run}, which could for instance be a sequence of
+there is a notion of a \emph{run}, which could for instance be a sequence of
 transitions taken in a finite state automaton. Continuing on the analogy with
@@
 (described by words) can be combined.
-
-\begin{figure}
-    \centering
-    \includestandalone[width=\linewidth]{fig/gap-word-embedding-standalone}
-    \caption{The \kl{gap words} resulting from a \kl{subword embedding} between two 
-    finite words.}
-    \label{gap-word-embedding:fig}
-\end{figure}
-
-\begin{figure}
-    \centering
-    \includestandalone[width=\linewidth]{fig/run-amalgamation-standalone}
-    \caption{We illustrate how 
-        embeddings $f$ and $g$ between runs of an
-        \kl{amalgamation system} can be glued
-        together, seen on their canonical decomposition.
-    }
-    \label{amalgamation-runs:fig}
-\end{figure}
 
@@
 Intuitively, the definition of an amalgamation system allows the comparison of
-runs, and the proper ``gluing'' of runs together to obtain new runs. A number of well-known language classes can be seen to be recognized by \kl{amalgamation
-systems}, e.g., regular languages \cite[Theorem 5.3]{ASZZ24}, reachability and coverability languages of VASS \cite[Theorem 5.5]{ASZZ24}, and context-free languages \cite[Theorem 5.10]{ASZZ24}.
-For this paper to be self-contained, we will also
-recall how runs of a finite state automaton can be understood as an
+runs, and the proper ``gluing'' of runs together to obtain new runs. A number
+of well-known language classes can be seen to be recognized by \kl{amalgamation
+systems}, e.g., regular languages \cite[Theorem 5.3]{ASZZ24}, reachability and
+coverability languages of VASS \cite[Theorem 5.5]{ASZZ24}, and context-free
+languages \cite[Theorem 5.10]{ASZZ24}. For this paper to be self-contained, we
+will also recall how runs of a finite state automaton can be understood as an
 \kl{amalgamation system}.
@@
 
-If we additionally assume that such a language is closed under infixes, we obtain an even stronger structure: All such languages are regular:
+If we additionally assume that such a language is closed under infixes, we
+obtain an even stronger structure: All such languages are regular!
 

### 2025-10-15 src/appendix.tex
new: moving stuff to appendix

--- 2025-10-06 src/prefixes.tex
@@
 % LTeX: language=en-GB 
-% !TeX root=../wqo-on-words.tex
-\section{Prefixes and Suffixes}
-\label{prefixes:sec}
-
-In this section, we study the well-quasi-ordering of languages under the
-\kl{prefix relation}. Let us immediately remark that the map $u \mapsto u^R$
-that reverses a word is an order-bijection between $(X^*, \prefleq)$ and $(X^*,
-\suffleq)$, that is, $u \prefleq v$ if and only if $u^R \suffleq v^R$.
-Therefore, we will focus on the prefix relation in the rest of this section, as
-$(L, \prefleq)$ is \kl{well-quasi-ordered} if and only if $(L^R, \suffleq)$ is.
-
-The next remark we make is that $\Sigma^*$ is not \kl{well-quasi-ordered} by
-the \kl{prefix relation} as soon as $\Sigma$ contains two distinct letters $a$
-and $b$. As an example of infinite \kl{antichain}, we can consider the set of
-words $a^n b$ for $n \in \Nat$. As mentioned in the introduction, there are
-however some languages  that are \kl{well-quasi-ordered} by the \kl{prefix
-relation}. A simple example being the (regular) language $a^* \subseteq
-\set{a,b}^*$, which is order-isomorphic to natural numbers with their usual
-orderings $(\Nat, \leq)$.
-
-In order to characterize the existence of infinite \kl{antichains} for the
-\kl{prefix relation}, we will introduce the following tree.
-
+% !TeX root=../wqo-on-words.lncs.tex
+\section{Proofs for Section~\ref{prefixes:sec}}
 \begin{figure}
@@
 
-\begin{definition}
-    The \intro{tree of prefixes} over a finite alphabet $\Sigma$
-    is the infinite tree $T$ whose nodes are the words of $\Sigma^*$, and
-    such that the children of a word $w$ are the words $wa$ for all $a \in
-    \Sigma$. 
-\end{definition}
-
-We will use this \kl{tree of prefixes} to find simple witnesses
-of the existence of infinite \kl{antichains} in the \kl{prefix relation}
-for a given language $L$, namely by introducing \kl{antichain branches}.
-
-\begin{definition}
-    An \intro{antichain branch} for a language $L$ is an infinite 
-    branch $B$ of the \kl{tree of prefixes} such that from every point of the branch, 
-    one can reach a word in $L \setminus B$. Formally:
-    $\forall u \in B, \exists v \in \Sigma^*, uv \in L \setminus B$.
-\end{definition}
-
-Let us illustrate the notion of \kl{antichain branch} over the alphabet $\Sigma
-= \set{a,b}$, and the language $L = a^* b$. In this case, the set $a^*$ (which
-is a branch of the \kl{tree of prefixes}) is an \kl{antichain branch} for $L$.
-This holds because for any $a^k$, the word $a^k \prefleq a^kb$ belongs to $L
-\setminus a^*$. In general, the existence of an \kl{antichain branch} for a
-language $L$ implies that $L$ contains an infinite \kl{antichain}, and because
-the alphabet $\Sigma$ is assumed to be finite, one can leverage the fact that
-the \kl{tree of prefixes} is finitely branching to prove that the converse
-holds as well.
-
-\begin{lemma}
-    \label{antichain-branches-prefix:lem}
-    Let $L \subseteq \Sigma^*$ be a language. Then, $L$ contains an infinite
-    \kl{antichain} if and only if there exists an \kl{antichain branch} for $L$.
-\end{lemma}
-\begin{proof}
+\begin{proofof}{antichain-branches-prefix:lem}
     Assume that $L$ contains an \kl{antichain branch}. Let us construct an
@@
     an infinite branch. By definition this is an \kl{antichain branch}.
-\end{proof}
+\end{proofof}
 
-One immediate application of \cref{antichain-branches-prefix:lem} is
-that \kl{antichain branches} can be described inside the \kl{tree of prefixes}
-by a monadic second order formula ($\MSO$-formula), allowing us to
-leverage the decidability of $\MSO$ over infinite binary trees \cite[Theorem
-1.1]{RAB69}. This result will follow from our general decidability result
-(\cref{infix-wqo-is-emptiness:thm}) but is worth stating on its own for its
-simplicity.
-
-\begin{corollary}
-    \label{prefix-wqo-reg-decidable:cor}
-    If $L$ is regular, then the existence of an infinite antichain is decidable.
-\end{corollary}
-\begin{proof}
-    If $L$ is regular, then it is $\MSO$-definable, and there 
-    exists a formula $\varphi(x)$ in $\MSO$ that selects nodes 
-    of the \kl{tree of prefixes} that belong to $L$. Now, to decide whether there
-    exists an \kl{antichain branch} for $L$, we can simply check whether
-    the following formula is satisfied:
-    \begin{equation*}
-        \exists B. 
-        B \text{ is a branch } \land
-        \forall x \in B, \exists y. y \text{ is a child of } x \land \varphi(y) \land y \not\in B
-        \quad .
-    \end{equation*}
-    Because the above formula is an $\MSO$-formula over the infinite
-    $\Sigma$-branching tree, whether it is satisfied is decidable
-    as an easy consequence of the decidability of $\MSO$ over infinite binary
-    trees
-    \cite[Theorem 1.1]{RAB69}.
-\end{proof}
-
-Let us now go further and fully characterize languages $L$ such that the
-prefix relation is well-quasi-ordered, without any restriction on the
-decidability of $L$ itself.
-For the proof of the following theorem, we will use special notations to
-describe the \intro{upwards closure} of a set $S$ for a relation $\preceq$,
-which is defined as $\upset[\preceq]{S} \defined \setof{y \in \Sigma^*}{
-\exists x \in S. x \preceq y}$. Anticipating the use of the symmetric notion of
-\intro{downwards closure}, let us introduce it as follows: $\dwset[\preceq]{S} \defined \setof{y
-\in \Sigma^*}{ \exists x \in S. y \preceq x}$. Abusing notations, we will
-write $\upset{w}$ and $\dwset{w}$ for the upwards and downwards closure of a
-single element $w$, omitting the ordering relation when it is clear from the
-context. A set $S$ is called \intro{downwards closed} if $\dwset{S} = S$.
-
-% \begin{figure}
-%     \centering
-%     \includestandalone[width=\linewidth]{fig/prefix-core-and-branches-standalone}
-%     \caption{Illustrating the construction of 
-%         \cref{prefixes:thm}. We represent the Hasse Diagram of a
-%         \kl{partial order}, with elements in $S$ being bigger and 
-%         in desert sand, elements of $S_{\max}$ being in
-%         blood-red, and \kl{chains} of elements in sky blue.
-%         The upwards closure of an element in $S_{\max}$ 
-%         is depicted in pale green, and cannot intersect two distinct
-%         branches, as it does not contain two incomparable elements.
-%     }
-%     \label{prefixes-theorem:fig}
-% \end{figure}
-
-\begin{theorem}
-    \label{prefixes:thm}
-    A language $L \subseteq \Sigma^*$ is \kl{well-quasi-ordered} by the
-    \kl{prefix relation} if and only if $L$ is a union of \kl{chains}.
-\end{theorem}
-\begin{proof}
+\begin{proofof}{prefixes:thm}
     Assume that $L$ is a finite union of \kl{chains}.
@@
     Hence, that $L$ is finite union of \kl{chains}.
-\end{proof}
+\end{proofof}
 
-As an immediate consequence, we have a very fine-grained understanding of the
-\kl{ordinal invariants} of such \kl{well-quasi-ordered} languages, which can be
-leveraged in bounding the complexity of algorithms working on such languages.
+\section{Proofs for Section~\ref{infixes-bounded:sec}}
 
-\begin{corollary}
-    \label{prefixes-ordinal-invariants:cor}
-    Let $L \subseteq \Sigma^*$ be a language that is
-    \kl{well-quasi-ordered} by the \kl{prefix relation}. Then,
-    \kl{maximal order type} of $L$ strictly smaller than $\omega^2$,
-    the \kl{ordinal height} of $L$ is at most $\omega$, and
-    its \kl{ordinal width} is finite. Furthermore, these bounds 
-    are tight.
-\end{corollary}
-\begin{proof}
-    The upper bounds follow from the fact that $L$ is a finite union of
-    \kl{chains}. The tightness can be obtained by considering the languages
-    $L_k \defined \bigcup_{i=0}^{k-1} a^i b^*$ for $k \in \Nat$, which 
-    are \kl{well-quasi-ordered} by the \kl{prefix relation} (as they are
-    finite unions of \kl{chains}), and satisfy
-    that $\oWidth{L_k} = k$, $\oHeight{L_k} = \omega$,
-    and therefore $\oType{L_k} = k \cdot \omega$.
-\end{proof}
+\begin{figure}
+    \centering
+    \includestandalone[width=\linewidth]{fig/infix-encoding-standalone}
+    \caption{Representation of the \kl{subword relation} for $\set{a,b}^*$
+        inside the \kl{infix relation} for $\set{a,b,\#}^*$
+        using a simplified version of \cref{infix-embedding:thm}, restricted to words
+        of length at most $3$. 
+    }
+    \label{infix-embedding:fig}
+\end{figure}
+
+\begin{proofof}{inf-period-chain:lem}
+    Let $x \in \Sigma^+$ be a word, and let $P_x$ be the (finite) set 
+    of all \kl{prefixes} of $x$, and $S_x$ be the (finite)
+    set of all \kl{suffixes} of $x$.
+    Assume that $w \in \InfPeriodChain{x}$, then $w = u x^p v$ for some
+    $u \in S_x$, $v \in P_x$, and $p \in \Nat$.
+    We have proven that
+    \begin{equation*}
+        \InfPeriodChain{x} \subseteq \bigcup_{u \in P_x} \bigcup_{v \in S_x} u x^* v
+        \quad .
+    \end{equation*}
+
+    Let us now demonstrate that for all $(u,v) \in S_x \times P_x$, the
+    language $u x^* v$ is a \kl{chain} for the \kl{infix}, \kl{suffix} and \kl{prefix} relations.
+    To that end,
+    let $(u,v) \in S_x \times P_x$ and $\ell, k \in \Nat$ be such that $\ell <
+    k$, let us prove that $u x^\ell v \infleq u x^k  v$. Because $v \prefleq
+    x$, we know that there exists $w$ such that $vw = x$. In particular,
+    $ux^\ell vw = u x^{\ell + 1}$, and because $\ell < k$, we conclude that $u
+    x^{\ell + 1} \prefleq u x^k v$. By transitivity, $u x^\ell v \prefleq u x^k
+    v$, and \emph{a fortiori}, $u x^\ell v \infleq u x^k v$. 
+    Similarly, because $u \suffleq x$,  there exists $w$ such that $wu  = x$, 
+    and we conclude that $u x^{\ell} v \suffleq w u x^\ell v = x^{\ell + 1} v \suffleq u x^k v$.
+\end{proofof}
+
+\section{Proofs for Section~\ref{infixes-amalgamation:sec}}
+
+\begin{figure}
+    \centering
+    \includestandalone[width=\linewidth]{fig/gap-word-embedding-standalone}
+    \caption{The \kl{gap words} resulting from a \kl{subword embedding} between two 
+    finite words.}
+    \label{gap-word-embedding:fig}
+\end{figure}
+
+\begin{figure}
+    \centering
+    \includestandalone[width=\linewidth]{fig/run-amalgamation-standalone}
+    \caption{We illustrate how 
+        embeddings $f$ and $g$ between runs of an
+        \kl{amalgamation system} can be glued
+        together, seen on their canonical decomposition.
+    }
+    \label{amalgamation-runs:fig}
+\end{figure}

### 2025-10-15 src/introduction.tex
new: moving stuff to appendix

@@
 being \kl(language){bounded} is the same as being \kl(language){regular}
-(\cref{bounded-wqo-dwclosed:cor}), and that a \kl{bounded language} is
+(\cref{dwclosed-infixes-wqo:lem}), and that a \kl{bounded language} is
 \kl{well-quasi-ordered} by the \kl{infix relation} if and only if its

### 2025-10-15 src/prefixes.tex
m
todos: deleting + answering
new: moving stuff to appendix

@@
 \kl{prefix relation}, we will introduce the following tree.
-
-\begin{figure}
-    \centering
-    \includestandalone{fig/antichain-branch-standalone}
-    \caption{An \kl{antichain branch} for the language $a^* b$,
-        represented in the \kl{tree of prefixes} over the alphabet $\set{a,b}$.
-        The branch is represented with dashed lines in turquoise, and the
-        \kl{antichain} is represented in dotted lines in blood-red.
-    }
-    \label{antichain-branch:fig}
-\end{figure}
 
@@
     \label{antichain-branches-prefix:lem}
+    \proofref{antichain-branches-prefix:lem}
     Let $L \subseteq \Sigma^*$ be a language. Then, $L$ contains an infinite
@@
 \end{lemma}
-\begin{proof}
-    Assume that $L$ contains an \kl{antichain branch}. Let us construct an
-    infinite \kl{antichain} as follows. We start with a set $A_0 \defined
-    \emptyset$ and a node $v_0$ at the root of the tree. At step $i$, we
-    consider a word $w_i$ such that $v_i$ is a \kl{prefix} of $w_i$, and $w_i
-    \in L \setminus B$, which exists by definition of \kl{antichain branches}.
-    We then set $A_{i+1} \defined A_i \cup \set{w_i}$. To compute $v_{i+1}$, we
-    consider the largest prefix of $w_i$ that belongs to $B$, and set $v_{i+1}$
-    to be the successor of this prefix in $B$. By an immediate induction, we
-    conclude that for all $i \in \Nat$, $A_i$ is an \kl{antichain}, and that
-    $v_i$ is a node in the \kl{antichain branch} $B$ such that $v_i$ is not a
-    prefix of any word in $A_i$.
-
-    Conversely, assume that $L$ contains an infinite \kl{antichain} $A$. Let us
-    construct an \kl{antichain branch}. Let us consider the subtree of the
-    \kl{tree of prefixes} that consists in words that are \kl{prefixes} of
-    words in $A$. This subtree is infinite, and by König's lemma, it contains
-    an infinite branch. By definition this is an \kl{antichain branch}.
-\end{proof}
 
@@
     \label{prefixes:thm}
+    \proofref{prefixes:thm}
     A language $L \subseteq \Sigma^*$ is \kl{well-quasi-ordered} by the
@@
 \end{theorem}
-\begin{proof}
-    Assume that $L$ is a finite union of \kl{chains}.
-    Because the \kl{prefix relation} is \kl{well-founded},
-    and that finite unions of \kl{chains} have finite \kl{antichains},
-    we conclude that $L$ is \kl{well-quasi-ordered}.
-
-    Conversely, assume that $L$ is \kl{well-quasi-ordered} by the \kl{prefix
-    relation}. Let us define $S_{\mathrm{split}}$ the set of words $w \in \Sigma^*$ 
-    such that there exists
-    two words $wu$ and $wv$ both in $L$ that are not comparable for the
-    \kl{prefix relation}. Let $S = S_{\mathrm{split}}~\cup~\min_{\prefleq} L$\todo{kinda ugly, but i can't find a nicer way of phrasing it}. 
-    Assume by contradiction that $S$ is infinite.
-    Then, $S$ equipped with the \kl{prefix relation} is an infinite
-    tree with finite branching, and therefore contains an infinite
-    branch, which is by definition an \kl{antichain branch} for $L$.
-    This contradicts the assumption that $L$ is \kl{well-quasi-ordered}.
-
-    Now, let $w$ be a maximal element for the \kl{prefix ordering}
-    in $S$. 
-    The upward closure of $w$ in $L$, $(\upset[\prefleq]{w}) \cap L$, must be a 
-    finite union of \kl{chains}. Otherwise at least two of the chains would share a common 
-    prefix in $w\Sigma$, contradicting the maximality of $w$.
-
-    In particular, letting $S_{\max}$ be the set of all maximal elements
-    of $S$,
-    we conclude that 
-    \begin{equation*}
-        L \subseteq S \cup \bigcup_{w \in S_{\max}} (\upset[\prefleq]{w}) \cap L
-        \quad .
-    \end{equation*}
-    Hence, that $L$ is finite union of \kl{chains}.
-\end{proof}
 
@@
 leveraged in bounding the complexity of algorithms working on such languages.
+
+\todo{nat: might be worth noting that factor complexity is O(1)}
+\todo{aliaume: I would put it in the section "factor complexity" in the concluding
+sections}
 

### 2025-10-15 src/amalgam.tex
Shorten section 6

@@
 class of all regular language, or the class of all context-free languages. Such
-classes will be called \kl{effective amalgamative classes} (\kcref{effective
-amalgamative classes}). In the following theorem, we prove that under such
+classes will be called \kl[effectiveness assumptions@amalgamation]{effective amalgamative classes} (\kcref{effectiveness assumptions@amalgamation})\todo{can't get the knowledge to work}.
+In the following theorem, we prove that under such
 assumptions, testing \kl{well-quasi-ordering} is inter-reducible to testing
@@
 
-\AP We say that a \kl{strongly effective amalgamative class} is an
-\kl{effective amalgamative class} for which the emptiness problem is decidable.
-Let us immediately remark that the class of \kl{regular languages} is
-\kl(amalg){strongly effective}, and so is the class of context-free languages.
-Therefore, \cref{infix-wqo-is-emptiness:thm} provides a concrete decision
-procedure for these classes.
+%\AP We say that a \kl{strongly effective amalgamative class} is an
+%\kl{effective amalgamative class} for which the emptiness problem is decidable.
+%Let us immediately remark that the class of \kl{regular languages} is
+%\kl(amalg){strongly effective}, and so is the class of context-free languages.
+%Therefore, \cref{infix-wqo-is-emptiness:thm} provides a concrete decision
+%procedure for these classes.
 
@@
     function computing a \intro{canonical decomposition} of a run,
-    and $E$ describes the so-called \intro{admissible embeddings} between runs: If $\rho$ and $\sigma$ are runs from $R$, then $E(\rho, \sigma)$ is a subset of the subword embeddings between $\canrun[\rho]$ and $\canrun[\sigma]$.
+    and $E$ describes the so-called \intro{admissible embeddings} between runs: If $\rho$ and $\sigma$ are runs from $R$, then $E(\rho, \sigma)$ is a subset of the subword embeddings between $\canrun(\rho)$ and $\canrun(\sigma)$.
 We write $\rho \intro*\runleq \sigma$ if $E(\rho, \sigma)$ is non-empty.
@@
 coverability languages of VASS \cite[Theorem 5.5]{ASZZ24}, and context-free
-languages \cite[Theorem 5.10]{ASZZ24}. For this paper to be self-contained, we
-will also recall how runs of a finite state automaton can be understood as an
-\kl{amalgamation system}.
+languages \cite[Theorem 5.10]{ASZZ24}.
 
-\begin{example}[{\cite[Section 3.2]{ASZZ24}}]
-    Let $A = (Q, \delta, q_0, F)$ be a finite state automaton over a finite
-    alphabet $\Sigma$. Let $\Delta$ be the set of transitions $(q_1, a, q_2)
-    \in Q \times \Sigma \times Q$,
-    and $R \subseteq \Delta^*$ be the set of 
-    words over transitions that start with the initial state $q_0$,
-    end in a final state $q_f \in F$, and such that the end state of a
-    letter is the start state of the following one.
-    The canonical decomposition $\canrun$
-    is defined as a morphism from $\Delta^*$ to $\Sigma^*$
-    that maps $(q,a,p)$ to $a$. 
-    Because of the one-to-one correspondence of steps of a run $\rho$ and letters in its \kl{canonical decomposition}, 
-    we may treat the two interchangeably.
-    Finally, given two runs $\rho$ and $\sigma$ of the automaton,
-    we say that an embedding $f \in \HigEmb(\canrun(\rho), \canrun(\sigma))$
-    belongs to $E(\rho,\sigma)$ when
-    $f$ is also defining an embedding from $\rho$ to $\sigma$ as words in $\Delta^*$.
-
-    The system $(\Sigma, R, E, \canrun)$ is an \kl{amalgamation system},
-    whose language is precisely the language of words recognized
-    by the automaton $A$.
-\end{example}
-\begin{proof}
-    By definition, the embeddings inside $E(\rho,\sigma)$ are in
-    of $\HigEmb(\canrun(\rho), \canrun(\sigma))$, and they compose properly.
-    Because $\Delta = Q \times \Sigma \times Q$ is finite, it is 
-    a \kl{well-quasi-ordering} when equipped with the equality relation, and 
-    we conclude that $\Delta^*$ with $\higleq$ is a \kl{well-quasi-order}
-    according to Higman’s Lemma \cite{HIG52}.
-
-    Let us now move to proving that the system satisfies the amalgamation
-    property. Given three runs $\rho,\sigma,\tau \in R$, and two embeddings $f \in E(\rho,\sigma)$
-    and $g \in E(\rho,\tau)$, we want to construct an amalgamated run $\sigma \vee \tau$.
-    Because letters in the run $\rho$ respect the transitions of the automaton
-    (i.e., if the letter $i$ ends in state $q$, then the letter $i+1$ starts in
-    state $q$), then the \kl{gap word} at position $i$ starts in state $q$ and
-    ends in state $q$ too. This means that for both embeddings
-    $f$ and $g$, the \kl{gap words} are read by the automaton by looping
-    on a state. In particular, these loops can be taken in any order
-    and continue to represent a valid run. That is, we can even select
-    the order of concatenation in the amalgamation for \emph{all} 
-    $0 \leq i \leq \card{\canrun(\rho)}$ and not just for one separately.
-
-    We conclude by remarking that 
-    the language of this amalgamation system is
-    the set of $\yieldrun(R)$, 
-    because $R$ is the set of valid runs of the automaton,
-    and $\yieldrun(\rho)$ is the word read along a run $\rho$.
-\end{proof}
+%For this paper to be self-contained, we
+%will also recall how runs of a finite state automaton can be understood as an
+%\kl{amalgamation system}.
+%
+%\begin{example}[{\cite[Section 3.2]{ASZZ24}}]
+%    Let $A = (Q, \delta, q_0, F)$ be a finite state automaton over a finite
+%    alphabet $\Sigma$. Let $\Delta$ be the set of transitions $(q_1, a, q_2)
+%    \in Q \times \Sigma \times Q$,
+%    and $R \subseteq \Delta^*$ be the set of 
+%    words over transitions that start with the initial state $q_0$,
+%    end in a final state $q_f \in F$, and such that the end state of a
+%    letter is the start state of the following one.
+%    The canonical decomposition $\canrun$
+%    is defined as a morphism from $\Delta^*$ to $\Sigma^*$
+%    that maps $(q,a,p)$ to $a$. 
+%    Because of the one-to-one correspondence of steps of a run $\rho$ and letters in its \kl{canonical decomposition}, 
+%    we may treat the two interchangeably.
+%    Finally, given two runs $\rho$ and $\sigma$ of the automaton,
+%    we say that an embedding $f \in \HigEmb(\canrun(\rho), \canrun(\sigma))$
+%    belongs to $E(\rho,\sigma)$ when
+%    $f$ is also defining an embedding from $\rho$ to $\sigma$ as words in $\Delta^*$.
+%
+%    The system $(\Sigma, R, E, \canrun)$ is an \kl{amalgamation system},
+%    whose language is precisely the language of words recognized
+%    by the automaton $A$.
+%\end{example}
+%\begin{proof}
+%    By definition, the embeddings inside $E(\rho,\sigma)$ are in
+%    of $\HigEmb(\canrun(\rho), \canrun(\sigma))$, and they compose properly.
+%    Because $\Delta = Q \times \Sigma \times Q$ is finite, it is 
+%    a \kl{well-quasi-ordering} when equipped with the equality relation, and 
+%    we conclude that $\Delta^*$ with $\higleq$ is a \kl{well-quasi-order}
+%    according to Higman’s Lemma \cite{HIG52}.
+%    
+%    Let us now move to proving that the system satisfies the amalgamation
+%    property. Given three runs $\rho,\sigma,\tau \in R$, and two embeddings $f \in E(\rho,\sigma)$
+%    and $g \in E(\rho,\tau)$, we want to construct an amalgamated run $\sigma \vee \tau$.
+%    Because letters in the run $\rho$ respect the transitions of the automaton
+%    (i.e., if the letter $i$ ends in state $q$, then the letter $i+1$ starts in
+%    state $q$), then the \kl{gap word} at position $i$ starts in state $q$ and
+%    ends in state $q$ too. This means that for both embeddings
+%    $f$ and $g$, the \kl{gap words} are read by the automaton by looping
+%    on a state. In particular, these loops can be taken in any order
+%    and continue to represent a valid run. That is, we can even select
+%    the order of concatenation in the amalgamation for \emph{all} 
+%    $0 \leq i \leq \card{\canrun(\rho)}$ and not just for one separately.
+%
+%    We conclude by remarking that 
+%    the language of this amalgamation system is
+%    the set of $\yieldrun(R)$, 
+%    because $R$ is the set of valid runs of the automaton,
+%    and $\yieldrun(\rho)$ is the word read along a run $\rho$.
+%\end{proof}
 
@@
 
+	\todo[inline]{The following can probably be replaced by a reference to Fine \& Wilf}
 	From \cref{pumping-periods:lem}, we may conclude that there are a finite number of words $x, y,$ and $w$ 
@@
 
-\AP Let us now introduce our effectiveness assumptions on \kl{amalgamation
-systems}. We follow the approach of \cite{ASZZ24} and require that an
-\kl{amalgamation system} $(\Sigma, R, E, \canrun)$ is effective when $R$ is
+\AP To construct a decision procedure for well-quasi-orderedness under $\infleq$, we 
+our \kl{amalgamation systems} to satisfy certain \intro{effectiveness assumptions}. We require that for an
+\kl{amalgamation system} $(\Sigma, R, E, \canrun)$, $R$ is
 recursively enumerable, the function $\canrun(\cdot)$ is computable, and for
-any two runs $\rho, \sigma \in R$, the set $E(\rho,\sigma)$ is computable.
+any two runs $\rho, \sigma \in R$, the set $E(\rho,\sigma)$ is computable. Additionally, we require the class to be effectively closed under rational transductions \cite[Chapter 5, page 64]{BERST79}.
 
-\AP We say that a class $\mathcal{C}$ of languages is an \intro{effective
-amalgamative class} whenever for every $L \in \mathcal{C}$, there exists an
-effective \kl{amalgamation system} recognizing $L$, and such that $\mathcal{C}$
-is \kl{effectively closed under rational transductions}. Recall that a
-\intro{rational transduction} is a \emph{relation} $R \subseteq \Sigma^* \times
-\Gamma^*$ that can be defined by a (nondeterministic) finite automaton with
-output \cite[Chapter 5, page 64]{BERST79}. A class of languages $\mathcal{C}$
-is \intro{effectively closed under rational transductions} when, for every
-language $L \in \mathcal{C}$, and every rational transduction $R \subseteq
-\Sigma^* \times \Gamma^*$, the image of $L$ through $R$ --- that is, $\setof{v
-\in \Gamma^*}{\exists u \in L. (u,v) \in R}$ --- is in $\mathcal{C}$ and
-effectively computable. In particular, it implies that the class contains
-languages over arbitrary (finite) alphabets.
+%\AP We say that a class $\mathcal{C}$ of languages is an \intro{effective
+%amalgamative class} whenever for every $L \in \mathcal{C}$, there exists an
+%effective \kl{amalgamation system} recognizing $L$, and such that $\mathcal{C}$
+%is \kl{effectively closed under rational transductions}. Recall that a
+%\intro{rational transduction} is a \emph{relation} $R \subseteq \Sigma^* \times
+%\Gamma^*$ that can be defined by a (nondeterministic) finite automaton with
+%output \cite[Chapter 5, page 64]{BERST79}. A class of languages $\mathcal{C}$
+%is \intro{effectively closed under rational transductions} when, for every
+%language $L \in \mathcal{C}$, and every rational transduction $R \subseteq
+%\Sigma^* \times \Gamma^*$, the image of $L$ through $R$ --- that is, $\setof{v
+%\in \Gamma^*}{\exists u \in L. (u,v) \in R}$ --- is in $\mathcal{C}$ and
+%effectively computable. In particular, it implies that the class contains
+%languages over arbitrary (finite) alphabets.
 
-\AP A class $\mathcal{C}$ of languages is \intro(amalg){strongly effective}
-whenever the emptiness problem for languages in $\mathcal{C}$ is decidable.
-This notion is interesting because usual language problems such as boundedness
-or simultaneous boundedness are decidable for \kl{strongly effective
-amalgamative classes}~\cite{ASZZ24}.
+%\AP A class $\mathcal{C}$ of languages is \intro(amalg){strongly effective}
+%whenever the emptiness problem for languages in $\mathcal{C}$ is decidable.
+%This notion is interesting because usual language problems such as boundedness
+%or simultaneous boundedness are decidable for \kl{strongly effective
+%amalgamative classes}~\cite{ASZZ24}.
 
 \begin{proofof}{infix-wqo-is-emptiness:thm}
-	\cref{emptiness-decidable} We first show $\Rightarrow$ \cref{wqo-infix-decidable}. We aim to make the inclusion test of \cref{bounded-language:eq} of \cref{bounded-language:thm} effective. 
+	We first show \cref{emptiness-decidable} $\Rightarrow$ \cref{wqo-infix-decidable}. We aim to make the inclusion test of \cref{bounded-language:eq} of \cref{bounded-language:thm} effective. 
 	Let $R(n,m,N_0) \defined \bigcup_{x,y \in \Sigma^{\leq n}} \bigcup_{u \in
@@
 
+\begin{remark}
 	Let us conclude by noting that it is unsurprisingly not possible to decide
 	whether a decidable language is \kl{well-quasi-ordered} by the \kl{prefix
-relation}. This is a very easy result whose sole purpose is to contrast with
-the decidability result of \cref{aut-cfg-infix:cor}.
+	relation} in general. 
+	%This is a very easy result whose sole purpose is to contrast with
+	%the decidability result of \cref{aut-cfg-infix:cor}.
 
-\begin{remark}
-    The following problem is undecidable: given a language $L$
-    decided by a Turing machine, answer whether 
-    $L$ is \kl{well-quasi-ordered} for the \kl{prefix relation}.
 \end{remark}

### 2025-10-15 src/appendix.tex
Move some stuff to appendix

@@
     an infinite branch. By definition this is an \kl{antichain branch}.
+\end{proofof}
+
+\begin{proofof}{prefix-wqo-reg-decidable:cor}
+	If $L$ is regular, then it is $\MSO$-definable, and there 
+	exists a formula $\varphi(x)$ in $\MSO$ that selects nodes 
+	of the \kl{tree of prefixes} that belong to $L$. Now, to decide whether there
+	exists an \kl{antichain branch} for $L$, we can simply check whether
+	the following formula is satisfied:
+	\begin{equation*}
+		\exists B. 
+		B \text{ is a branch } \land
+		\forall x \in B, \exists y. y \text{ is a child of } x \land \varphi(y) \land y \not\in B
+		\quad .
+	\end{equation*}
+	Because the above formula is an $\MSO$-formula over the infinite
+	$\Sigma$-branching tree, whether it is satisfied is decidable
+	as an easy consequence of the decidability of $\MSO$ over infinite binary
+	trees
+	\cite[Theorem 1.1]{RAB69}.
 \end{proofof}
@@
 
+\begin{proofof}{bounded-language:lem}
+	\begin{proof}
+		Let $w_1, \dots, w_n$ be such that
+		$L \subseteq w_1^* \cdots w_n^*$.
+		Let us define $m \defined \max \setof{\card{w_i}}{1 \leq i \leq n}$
+
+		Let $w[\vec{k}] \defined w_1^{k_1} \cdots w_n^{k_n}$ be a map from $\Nat^k$
+		to $\Sigma^*$. We are interested in the intersection of the image of $w$
+		with $L$. Let us assume for instance that for all $\vec{k} \in \Nat^n$,
+		there exists $\vec{\ell} \geq \vec{k}$ such that $w[\vec{\ell}] \in L$.
+		Then, leveraging \cref{pumping-periods:lem}, we conclude that there exists
+		$x,y$ of size at most $\max\setof{\card{w_i}}{1 \leq i \leq n}$ such that
+		$w[\vec{k}] \in \InfPeriodChain{x} \cup \InfPeriodChain{x}
+		\InfPeriodChain{y}$, and we conclude that $L \subseteq \InfPeriodChain{x}
+		\cup \InfPeriodChain{x} \InfPeriodChain{y}$.
+
+		Now, it may be the case that one cannot simultaneously assume that two
+		component of the vector $\vec{k}$ are unbounded. In general, given a set $S
+		\subseteq \set{1, \dots, n}$ of indices, we say that $S$ is admissible if
+		there exists a bound $N_0$ such that for all $\vec{b} \in \Nat^S$, there
+		exists a vector $\vec{k} \in \Nat^n$, such that $\vec{k}$ is greater than
+		$\vec{b}$ on the $S$ components, and the other components are below the
+		bound $N_0$. The language of an admissible set $S$ is the set of words
+		obtained by repeating $w_i$ at most $N_0$ times if it is not in $S$
+		($w_i^{\leq N_0}$) and arbitrarily many times otherwise ($w_i^*$).
+		Note that $L \subseteq \bigcup_{S \text{ admissible }} L(S)$.
+
+		Now, admissible languages are ready to be pumped according to
+		\cref{pumping-periods:lem}. For every admissible language,
+		the size of a word that is not iterated is at most
+		$N_0 \times m$ by definition, and we conclude that:
+		\begin{equation}
+			\label{bounded-language:eq}
+			L \subseteq 
+			\bigcup_{x,y \in \Sigma^{\leq n}}
+			\bigcup_{u \in \Sigma^{\leq m \times N_0}}
+			\InfPeriodChain{x} u \InfPeriodChain{y}
+			\cup
+			\InfPeriodChain{x} u
+			\cup
+			u \InfPeriodChain{x}
+			\quad .
+		\end{equation}
+	\end{proof}
+\end{proofof}
+
 \section{Proofs for Section~\ref{infixes-amalgamation:sec}}

### 2025-10-15 src/infixes.tex
new: moving stuff to appendix
Move some stuff to appendix

--- 2025-10-14 src/infixes.tex
@@
     \label{bounded-language:lem}
+    \proofref{bounded-language:lem}
     Let $L \subseteq \Sigma^*$ be a \kl{bounded language}
@@
 \end{lemma}
-\begin{proof}
-    Let $w_1, \dots, w_n$ be such that
-    $L \subseteq w_1^* \cdots w_n^*$.
-    Let us define $m \defined \max \setof{\card{w_i}}{1 \leq i \leq n}$
-
-    Let $w[\vec{k}] \defined w_1^{k_1} \cdots w_n^{k_n}$ be a map from $\Nat^k$
-    to $\Sigma^*$. We are interested in the intersection of the image of $w$
-    with $L$. Let us assume for instance that for all $\vec{k} \in \Nat^n$,
-    there exists $\vec{\ell} \geq \vec{k}$ such that $w[\vec{\ell}] \in L$.
-    Then, leveraging \cref{pumping-periods:lem}, we conclude that there exists
-    $x,y$ of size at most $\max\setof{\card{w_i}}{1 \leq i \leq n}$ such that
-    $w[\vec{k}] \in \InfPeriodChain{x} \cup \InfPeriodChain{x}
-    \InfPeriodChain{y}$, and we conclude that $L \subseteq \InfPeriodChain{x}
-    \cup \InfPeriodChain{x} \InfPeriodChain{y}$.
-
-    Now, it may be the case that one cannot simultaneously assume that two
-    component of the vector $\vec{k}$ are unbounded. In general, given a set $S
-    \subseteq \set{1, \dots, n}$ of indices, we say that $S$ is admissible if
-    there exists a bound $N_0$ such that for all $\vec{b} \in \Nat^S$, there
-    exists a vector $\vec{k} \in \Nat^n$, such that $\vec{k}$ is greater than
-    $\vec{b}$ on the $S$ components, and the other components are below the
-    bound $N_0$. The language of an admissible set $S$ is the set of words
-    obtained by repeating $w_i$ at most $N_0$ times if it is not in $S$
-    ($w_i^{\leq N_0}$) and arbitrarily many times otherwise ($w_i^*$).
-    Note that $L \subseteq \bigcup_{S \text{ admissible }} L(S)$.
-
-    Now, admissible languages are ready to be pumped according to
-    \cref{pumping-periods:lem}. For every admissible language,
-    the size of a word that is not iterated is at most
-    $N_0 \times m$ by definition, and we conclude that:
-    \begin{equation}
-        \label{bounded-language:eq}
-        L \subseteq 
-        \bigcup_{x,y \in \Sigma^{\leq n}}
-        \bigcup_{u \in \Sigma^{\leq m \times N_0}}
-        \InfPeriodChain{x} u \InfPeriodChain{y}
-        \cup
-        \InfPeriodChain{x} u
-        \cup
-        u \InfPeriodChain{x}
-        \quad .
-    \end{equation}
-\end{proof}
 

### 2025-10-15 src/prefixes.tex
Move some stuff to appendix

@@
     \label{prefix-wqo-reg-decidable:cor}
+    \proofref{prefix-wqo-reg-decidable:cor}
     If $L$ is regular, then the existence of an infinite antichain is decidable.
 \end{corollary}
-\begin{proof}
-    If $L$ is regular, then it is $\MSO$-definable, and there 
-    exists a formula $\varphi(x)$ in $\MSO$ that selects nodes 
-    of the \kl{tree of prefixes} that belong to $L$. Now, to decide whether there
-    exists an \kl{antichain branch} for $L$, we can simply check whether
-    the following formula is satisfied:
-    \begin{equation*}
-        \exists B. 
-        B \text{ is a branch } \land
-        \forall x \in B, \exists y. y \text{ is a child of } x \land \varphi(y) \land y \not\in B
-        \quad .
-    \end{equation*}
-    Because the above formula is an $\MSO$-formula over the infinite
-    $\Sigma$-branching tree, whether it is satisfied is decidable
-    as an easy consequence of the decidability of $\MSO$ over infinite binary
-    trees
-    \cite[Theorem 1.1]{RAB69}.
-\end{proof}
 

### 2025-10-15 src/appendix.tex
Move some stuff to appendix

@@
 
+\begin{lemma}
+	\label{automatic-uur:lem}
+	\proofref{automatic-uur:lem}
+	Given an \kl{automatic sequence} $w \in \Sigma^{\Nat}$, one can decide
+	whether it is \kl{ultimately uniformly recurrent}.
+\end{lemma}
+
+\begin{proofof}{automatic-uur:lem}
+	We can rewrite this as a question on the \kl{automatic sequence} $w$
+	as follows:
+	\begin{align*}
+		&\exists N_0,                   &   \text{ultimately} \\
+		&\forall i_s \geq N_0,          &   \text{for every infix (start) } u \\
+		&\forall i_e > i_s,             &   \text{for every infix (end) }   u \\
+		&\exists k \geq 1,              &   \text{there exists a bound} \\
+		&\forall j_s \geq N_0,          &   \text{for every other infix (start) } v \\
+		&\forall j_e \geq j_s + k,      &   \text{of size at least $k$} \\
+		&\exists l \geq 0,              &   \text{there exists a position in } v \\
+		&\forall 0 \leq m < i_e - i_s,  &   \text{where } u \text{ can be found} \\
+		&j_s + m + l < j_e \land
+		w(i_s + m) = w(j_s + m + l) \quad .
+	\end{align*}
+	Because $w$ is computable by a finite automaton, one can reduce the above
+	formula to a regular language, for which it suffices to check emptiness, which
+	is decidable.
+\end{proofof}
+
 \section{Proofs for Section~\ref{infixes-amalgamation:sec}}

### 2025-10-15 src/infixes.tex
Move some stuff to appendix

@@
 (trivially) translated into a finite automaton that reads the base $b$
-representation of the free variables (that are indices of the sequence). We
-briefly explain in the following lemma the folklore result that one can decide
-if an \kl{automatic sequence} is \kl{ultimately uniformly recurrent}.
+representation of the free variables (that are indices of the sequence).
+It is folklore that one can decide if an \kl{automatic sequence} is \kl{ultimately uniformly recurrent}.
+A proof can be found in the appendix at \cref{automatic-uur:lem}.
+Based on this, we now prove:
 
-\begin{lemma}
-  \label{automatic-uur:lem}
-    Given an \kl{automatic sequence} $w \in \Sigma^{\Nat}$, one can decide
-    whether it is \kl{ultimately uniformly recurrent}.
-\end{lemma}
-\begin{proof}
-    We can rewrite this as a question on the \kl{automatic sequence} $w$
-    as follows:
-    \begin{align*}
-        &\exists N_0,                   &   \text{ultimately} \\
-        &\forall i_s \geq N_0,          &   \text{for every infix (start) } u \\
-        &\forall i_e > i_s,             &   \text{for every infix (end) }   u \\
-        &\exists k \geq 1,              &   \text{there exists a bound} \\
-        &\forall j_s \geq N_0,          &   \text{for every other infix (start) } v \\
-        &\forall j_e \geq j_s + k,      &   \text{of size at least $k$} \\
-        &\exists l \geq 0,              &   \text{there exists a position in } v \\
-        &\forall 0 \leq m < i_e - i_s,  &   \text{where } u \text{ can be found} \\
-        &j_s + m + l < j_e \land
-        w(i_s + m) = w(j_s + m + l) \quad .
-    \end{align*}
-    Because $w$ is computable by a finite automaton, one can reduce the above
-    formula to a regular language, for which it suffices to check emptiness, which
-    is decidable.
-\end{proof}
-
-Based on \cref{automatic-uur:lem}, we can now prove the following theorem.
 \begin{theorem}

### 2025-10-15 src/appendix.tex
Move some stuff to appendix

@@
 
+\begin{proofof}{small-ordinal-invariants:lem}
+	Let $N_0$ be a bound such that $w_{\geq N_0}$ is \kl{uniformly recurrent}.
+	Let us write $\infset{w}$ the set of finite infixes of $w$.
+	We prove that $\oWidth{\infset{w}} \leq \omegaOrd + N_0$.
+	Let $u_1 \infleq w$ be a finite word.
+
+	If $u_1$ is an \kl{infix} of $w_{\geq N_0}$, then there exists $k \geq 1$
+	such that $u_1$ is an \kl{infix} of every word of size at least $k$. In
+	particular, there is finite bound on the length of every sequence of
+	incomparable elements starting with $u_1$. We conclude in particular that
+	$\infset{w} \setminus \upset{u_1}$ has a finite \kl{ordinal width}.
+
+	Otherwise, $u_1$ can only be found \emph{before} $N_0$. In this case, we
+	consider a second element of a \kl{bad sequence} $u_2 \infleq w$, which is
+	incomparable with $u_1$ for the \kl{infix relation}. If $u_2$ is an
+	\kl{infix} of $w_{\geq N_0}$, then we can conclude as before. Otherwise,
+	notice that $u_1$ and $u_2$ cannot start at the same position in $w$
+	(because they are incomparable). Continuing this argument, we conclude that
+	there are at most $N_0$ elements starting before $N_0$
+	at the start of any sequence of
+	incomparable elements in $\infset{w}$. We conclude that
+	$\oWidth{\infset{w}} \leq \omegaOrd + N_0$.
+
+	Let us now justify that this bound is tight.
+	The \kl{Thue-Morse sequence} over a binary
+	alphabet $\set{a,b}$ has \kl{ordinal width} $\omegaOrd$
+	from \cref{thue-morse-ordinal:lemma}.
+	Given a number $N_0
+	\in \Nat$, one can construct an arbitrarily long \kl{antichain} of words for
+	the \kl{infix relation} by using a new letter $c$. When concatenating this
+	(finite) antichain as a prefix of the \kl{Thue-Morse sequence}, one obtains a
+	new (infinite) word $w$. It is clear that the \kl{ordinal width} of
+	$\infset{w}$ is now at least $\omegaOrd + N_0$.
+\end{proofof}
+
+\begin{proofof}{from-bi-to-single:lem}
+	Given a bi-infinite word $w \in \Sigma^{\Rel}$, recall that we can consider $w_+ \in
+	\Sigma^\Nat$ and $w_- \in \Sigma^\Nat$ the two infinite words obtained as
+	follows: for all $i \in \Nat$, $(w_+)_i = w(i)$ and $(w_-)_i = w(-i)$. Note
+	that the two share the letter at position $0$.
+
+	To obtain the upper bound of $3 \cdot \omegaOrd$, we can consider the same
+	argument as for \cref{small-ordinal-invariants:lem}. We let $N_0$ be such
+	that $w_{\geq N_0}$ and $(w_-)_{\geq N_0}$ are \kl{uniformly recurrent}
+	words. In any sequence of incomparable elements of $\infset{w}$, there are
+	less than $N_0^2$ elements that are found in $(w_{< N_0})_{> -N_0}$. Then,
+	one has to pick a finite \kl{infix} in either $w_{\geq N_0}$ or $w_{\leq
+		-N_0}$. Because of \cref{small-ordinal-invariants:lem}, any sequence of
+	incomparable elements of these two infinite words has length bounded based
+	on the choice of the first element of that sequence. This means that the
+	\kl{ordinal width} of $\infset{w}$ is at most $\omegaOrd + \omegaOrd +
+	N_0^2$. We conclude that $\oWidth{\infset{w}} < 3 \cdot \omegaOrd$.
+
+	Let us briefly argue that the bound is tight. Indeed, one can
+	construct a bi-infinite word $w$ by concatenating a reversed \kl{Thue-Morse
+		sequence} on a binary alphabet $\set{a,b}$, a finite antichain of arbitrarily
+	large size over a distinct alphabet $\set{c,d}$, and then a \kl{Thue-Morse
+		sequence} on a binary alphabet $\set{e,f}$. The \kl{ordinal width} of the set
+	of \kl{infixes} of $w$ is then at least $2 \cdot \omegaOrd + K$, where $K$ is the
+	size of the chosen antichain, following the same argument as in the proof of
+	\cref{small-ordinal-invariants:lem}, using \cref{thue-morse-ordinal:lemma}.
+\end{proofof}
+
 \section{Proofs for Section~\ref{infixes-amalgamation:sec}}

### 2025-10-15 src/infixes.tex
Move some stuff to appendix

@@
 We are now ready to conclude the proof of \cref{dw-closed-wqo-charac:thm}.
+
 \begin{proofof}{dw-closed-wqo-charac:thm}
@@
 
-Let us now analyse the \kl{ordinal invariants} of the set of finite
+Finally, we comment on the \kl{ordinal invariants} of the set of finite
 \kl{infixes} of an \kl{ultimately uniformly recurrent} infinite word, in order
@@
     \label{small-ordinal-invariants:lem}
+    \proofref{small-ordinal-invariants:lem}
     Let $w \in \Sigma^\Nat$ be an \kl{ultimately uniformly recurrent} word.
@@
 \end{lemma}
-\begin{proof}
-    Let $N_0$ be a bound such that $w_{\geq N_0}$ is \kl{uniformly recurrent}.
-    Let us write $\infset{w}$ the set of finite infixes of $w$.
-    We prove that $\oWidth{\infset{w}} \leq \omegaOrd + N_0$.
-    Let $u_1 \infleq w$ be a finite word.
-
-    If $u_1$ is an \kl{infix} of $w_{\geq N_0}$, then there exists $k \geq 1$
-    such that $u_1$ is an \kl{infix} of every word of size at least $k$. In
-    particular, there is finite bound on the length of every sequence of
-    incomparable elements starting with $u_1$. We conclude in particular that
-    $\infset{w} \setminus \upset{u_1}$ has a finite \kl{ordinal width}.
-
-    Otherwise, $u_1$ can only be found \emph{before} $N_0$. In this case, we
-    consider a second element of a \kl{bad sequence} $u_2 \infleq w$, which is
-    incomparable with $u_1$ for the \kl{infix relation}. If $u_2$ is an
-    \kl{infix} of $w_{\geq N_0}$, then we can conclude as before. Otherwise,
-    notice that $u_1$ and $u_2$ cannot start at the same position in $w$
-    (because they are incomparable). Continuing this argument, we conclude that
-    there are at most $N_0$ elements starting before $N_0$
-    at the start of any sequence of
-    incomparable elements in $\infset{w}$. We conclude that
-    $\oWidth{\infset{w}} \leq \omegaOrd + N_0$.
-
-  Let us now justify that this bound is tight.
-  The \kl{Thue-Morse sequence} over a binary
-  alphabet $\set{a,b}$ has \kl{ordinal width} $\omegaOrd$
-  from \cref{thue-morse-ordinal:lemma}.
-    Given a number $N_0
-  \in \Nat$, one can construct an arbitrarily long \kl{antichain} of words for
-  the \kl{infix relation} by using a new letter $c$. When concatenating this
-  (finite) antichain as a prefix of the \kl{Thue-Morse sequence}, one obtains a
-  new (infinite) word $w$. It is clear that the \kl{ordinal width} of
-  $\infset{w}$ is now at least $\omegaOrd + N_0$.
-\end{proof}
 
@@
     \label{from-bi-to-single:lem}
+    \proofref{from-bi-to-single:lem}
     Let $w \in \Sigma^\Rel$ be a bi-infinite word. Then, the set of finite
@@
 \end{lemma}
-\begin{proof}
-  Given a bi-infinite word $w \in \Sigma^{\Rel}$, recall that we can consider $w_+ \in
-\Sigma^\Nat$ and $w_- \in \Sigma^\Nat$ the two infinite words obtained as
-follows: for all $i \in \Nat$, $(w_+)_i = w(i)$ and $(w_-)_i = w(-i)$. Note
-that the two share the letter at position $0$.
 
-    To obtain the upper bound of $3 \cdot \omegaOrd$, we can consider the same
-    argument as for \cref{small-ordinal-invariants:lem}. We let $N_0$ be such
-    that $w_{\geq N_0}$ and $(w_-)_{\geq N_0}$ are \kl{uniformly recurrent}
-    words. In any sequence of incomparable elements of $\infset{w}$, there are
-    less than $N_0^2$ elements that are found in $(w_{< N_0})_{> -N_0}$. Then,
-    one has to pick a finite \kl{infix} in either $w_{\geq N_0}$ or $w_{\leq
-    -N_0}$. Because of \cref{small-ordinal-invariants:lem}, any sequence of
-    incomparable elements of these two infinite words has length bounded based
-    on the choice of the first element of that sequence. This means that the
-    \kl{ordinal width} of $\infset{w}$ is at most $\omegaOrd + \omegaOrd +
-    N_0^2$. We conclude that $\oWidth{\infset{w}} < 3 \cdot \omegaOrd$.
-
-  Let us briefly argue that the bound is tight. Indeed, one can
-  construct a bi-infinite word $w$ by concatenating a reversed \kl{Thue-Morse
-  sequence} on a binary alphabet $\set{a,b}$, a finite antichain of arbitrarily
-  large size over a distinct alphabet $\set{c,d}$, and then a \kl{Thue-Morse
-  sequence} on a binary alphabet $\set{e,f}$. The \kl{ordinal width} of the set
-  of \kl{infixes} of $w$ is then at least $2 \cdot \omegaOrd + K$, where $K$ is the
-  size of the chosen antichain, following the same argument as in the proof of
-  \cref{small-ordinal-invariants:lem}, using \cref{thue-morse-ordinal:lemma}.
-\end{proof}
-
-We are now ready to conclude the proof of \cref{small-ordinal-invariants:thm}.
+Ordinal invariants in hand, we proceed with the proof of \cref{small-ordinal-invariants:thm}.
 \begin{proofof}{small-ordinal-invariants:thm}

### 2025-10-15 src/amalgam.tex
new: moving stuff to appendix

--- 2025-10-15 src/amalgam.tex
@@
 
-We can now show a simple lemma that illuminates much of the structure of amalgamation systems whose language is well-quasi-ordered by $\infleq$.
+We can now show a simple lemma that illuminates much of the structure of
+amalgamation systems whose language is well-quasi-ordered by $\infleq$. Note
+that \cref{gap-words-prefix-ordered:lem} uses \cref{pumping-periods:lem} in its
+proof, and our \cref{infix-amalgamation:thm} follows from it.
 
@@
 	\label{gap-words-prefix-ordered:lem}
+  \proofref{gap-words-prefix-ordered:lem}
 	Let $L$ by an \kl{amalgamation language} recognized by $(\Sigma, R, E, \canrun)$ that is well-quasi-ordered by $\infleq$.
@@
 \end{lemma}
-
-\begin{proof}
-	Write $u$ for $\GapWord{f}{\ell}$ and $v$ for $\GapWord{g}{\ell}$. 
-	We may assume that both $u$ and $v$ are non-empty, as otherwise the lemma holds trivially.
-	Then, for all $k \in \Nat$, there exists a run with canonical decomposition
-	$$
-	w_k = L_0 a_1 \cdots a_n L_n,
-	$$
-	where $L_i \in \set{vv u^k, vu^kv, u^k vv}$ and specifically $L_\ell = vu^kv$.
-
-	From \cref{pumping-periods:lem}, we may conclude that there are a finite number of words $x, y,$ and $w$ 
-	such that each $w_k$ is contained in a language 
-	$\InfPeriodChain{x}w\InfPeriodChain{y}$.
-
-	As there is an infinite number of words $w_k$, 
-	we may fix $x, y,$ and $w$ and an infinite subset $I \subseteq \Nat$ 
-	such that $\set{w_i \mid i \in I} \subseteq \InfPeriodChain{x}w\InfPeriodChain{y}$. 
-	This implies that either for infinitely many $m \in \Nat$, $u^m v \in \InfPeriodChain{y}$ 
-	or for infinitely many $m$, $v u^m \in \InfPeriodChain{x}$.
-
-	In either case, we may conclude that either $u \infleq v$ or $v \infleq u$: Let $m, n \in \Nat$
-	such that $m < n$ and $u^m v, u^n v \in \InfPeriodChain{y}$ (the case for $v u^m$ and $v u^n$ 
-	proceeding analogously). Without loss of generality, assume that $\card{u^m}$ and $\card{u^n}$ are
-	multiples of $\card{y}$. We therefore find $p \prefleq y, s \suffleq y$ such that $u^m, u^n \in sy^*p$, 
-	ergo $ps = y$.
-	In other words, we can write $u^m = (sp)^{m'}, u^n = (sp)^{n'}$. As $u^mv \in \InfPeriodChain{y}$, it 
-	follows that $v$ is a prefix of some word in $(sp)^*$. Hence either $v$ is a prefix of $u$ or $u$ vice versa.
-\end{proof}
 
@@
     \label{dwclosed-infixes-wqo:lem}
+    \proofref{dwclosed-infixes-wqo:lem}
     Let $L \subseteq \Sigma^*$ be a \kl{downwards closed} language for the
@@
 \end{lemma}
-\begin{proof}
-    It is clear that \cref{dwci-reg:item} $\Rightarrow$ \cref{dwci-aml:item}
-    because regular languages are recognized by finite automata, and finite
-    automata are a particular case of \kl{amalgamation systems}.
-    The implication \cref{dwci-aml:item} $\Rightarrow$ \cref{dwci-bod:item}
-    is the content of \cref{infix-amalgamation:thm}.
-    The implication \cref{dwci-bod:item} $\Rightarrow$ \cref{dwci-uoc:item}
-    is \cref{bounded-language:lem}.
-    Finally, the implication \cref{dwci-uoc:item} $\Rightarrow$ \cref{dwci-reg:item}
-    is simply because a \kl{downwards closed} language 
-    that is a finite union of products of \kl{chains} is a regular language.
 
-    Indeed, assume that
-    $L$ is \kl{downwards closed} and included in a finite union of sets of the form
-    $\InfPeriodChain{x} u \InfPeriodChain{y}$ where $x,y,u$ are possibly empty words.
-    We can assume without loss of generality that
-    for every $n$, $x^n u y^n$ is in $L$, otherwise, we have a bound on the maximal $n$ such that
-    $x^n u y^n$ is in $L$, and we can increase the number of languages in the union, replacing $x$ or $y$
-    with the empty word as necessary.
-    Let us write $L' \defined \bigcup_{i = 1}^k x_i^* u_i y_i^*$. Then, $L'
-    \subseteq L$ by construction. Furthermore, $L \subseteq \dwset{L'}$, also
-    by construction. Finally, we conclude that $L = \dwset{L'}$ because $L$ is
-    \kl{downwards closed}. Now, because $L'$ is a \kl{regular language}, and 
-    \kl{regular languages} are closed under \kl{downwards closure}, we conclude
-    that $L$ is a \kl{regular language}.
-\end{proof}
-
-\AP Combining
-\cref{thue-morse:lemma,dwclosed-infixes-wqo:lem}, we can
-conclude that the collection of \kl{infixes} of the \kl{Thue-Morse sequence}
-cannot be recognized by \emph{any} \kl{amalgamation system}.
+\AP Combining \cref{thue-morse:lemma,dwclosed-infixes-wqo:lem}, we can conclude
+that the collection of \kl{infixes} of the \kl{Thue-Morse sequence} cannot be
+recognized by \emph{any} \kl{amalgamation system}.
 

### 2025-10-15 src/appendix.tex
new: moving stuff to appendix

--- 2025-10-15 src/appendix.tex
@@
 % !TeX root=../wqo-on-words.lncs.tex
+
+\clearpage
+
 \section{Proofs for Section~\ref{prefixes:sec}}
@@
     an infinite branch. By definition this is an \kl{antichain branch}.
-\end{proofof}
-
-\begin{proofof}{prefix-wqo-reg-decidable:cor}
-	If $L$ is regular, then it is $\MSO$-definable, and there 
-	exists a formula $\varphi(x)$ in $\MSO$ that selects nodes 
-	of the \kl{tree of prefixes} that belong to $L$. Now, to decide whether there
-	exists an \kl{antichain branch} for $L$, we can simply check whether
-	the following formula is satisfied:
-	\begin{equation*}
-		\exists B. 
-		B \text{ is a branch } \land
-		\forall x \in B, \exists y. y \text{ is a child of } x \land \varphi(y) \land y \not\in B
-		\quad .
-	\end{equation*}
-	Because the above formula is an $\MSO$-formula over the infinite
-	$\Sigma$-branching tree, whether it is satisfied is decidable
-	as an easy consequence of the decidability of $\MSO$ over infinite binary
-	trees
-	\cite[Theorem 1.1]{RAB69}.
 \end{proofof}
@@
     Hence, that $L$ is finite union of \kl{chains}.
+\end{proofof}
+
+\begin{proofof}{prefix-wqo-reg-decidable:cor}
+    If $L$ is regular, then it is $\MSO$-definable, and there 
+    exists a formula $\varphi(x)$ in $\MSO$ that selects nodes 
+    of the \kl{tree of prefixes} that belong to $L$. Now, to decide whether there
+    exists an \kl{antichain branch} for $L$, we can simply check whether
+    the following formula is satisfied:
+    \begin{equation*}
+        \exists B. 
+        B \text{ is a branch } \land
+        \forall x \in B, \exists y. y \text{ is a child of } x \land \varphi(y) \land y \not\in B
+        \quad .
+    \end{equation*}
+    Because the above formula is an $\MSO$-formula over the infinite
+    $\Sigma$-branching tree, whether it is satisfied is decidable
+    as an easy consequence of the decidability of $\MSO$ over infinite binary
+    trees
+    \cite[Theorem 1.1]{RAB69}.
 \end{proofof}
@@
 \begin{proofof}{bounded-language:lem}
-	\begin{proof}
     Let $w_1, \dots, w_n$ be such that
@@
       \end{equation}
-	\end{proof}
+  \end{proofof}
+
+\begin{proofof}{bounded-wqo-dwclosed:cor}
+    Because $L \subseteq \dwset[\infleq]{L}$, the right-to-left implication
+    is trivial.
+    For the left-to-right implication, let us assume that $L$ is a
+    \kl{well-quasi-ordered} language for the \kl{infix relation}.
+    Then $L$ is included in a finite union 
+    of products of \kl{chains} for the \kl{prefix} and \kl{suffix} relations
+    thanks to
+    \cref{bounded-language:thm}:
+    \begin{equation*}
+        L \subseteq \bigcup_{i = 1}^n S_i \cdot P_i \quad .
+    \end{equation*}
+    Remark that if $S_i$ is a \kl{chain} for the \kl{suffix relation}
+    and $P_i$ is a \kl{chain} for the \kl{prefix relation},
+    then 
+    \begin{equation*}
+      \dwset[\infleq]{(S_i \cdot P_i)} = (\dwset[\suffleq]{S_i}) \cdot (\dwset[\prefleq]{P_i})
+        \quad .
+    \end{equation*}
+    Indeed, any \kl{infix} of a word in $S_i P_i$ can be split into
+    a \kl{suffix} of a word in $S_i$ and a \kl{prefix} of a word in $P_i$.
+    Conversely, any such concatenations are \kl{infixes} of a word in $S_i P_i$.
+
+    As a consequence, we conclude that $\dwset[\infleq]{L}$ is itself included
+    in a finite union of products of \kl{chains}.
+    Furthermore, by definition of \kl{bounded languages},
+    $\dwset[\infleq]{L}$ is also a \kl{bounded language}.
+    Hence, it is \kl{well-quasi-ordered} by the \kl{infix relation} 
+    via
+    \cref{bounded-language:thm}.
+\end{proofof}
+
+\section{Proofs for Section~\ref{infixes-dwclosed:sec}}
+
+\begin{proofof}{bi-infinite:lem}
+    Let us assume that $L$ is infinite. The case when it is finite 
+    is similar, but will result in a finite word.
+    \todo{aliaume: do we need wqo here? the proof
+      should go through without it: the sequence 
+      $u_i$ is already increasing for infix}
+
+    Because the alphabet $\Sigma$ is finite, we can enumerate the words of $L$
+    as $\seqof{w_i}[i \in \Nat]$. From $\seqof{w_i}[i \in \Nat]$, we construct a sequence $\seqof{u_i}[i \in \Nat]$ by induction
+    as follows: $u_0 = w_0$, and $u_{i+1}$ is a word that contains $u_i$ and
+    $w_i$, which exists in $L$ because $L$ is \kl(subset){directed}. Since $L$ is
+    \kl{well-quasi-ordered}, one can extract an infinite set of indices $I
+    \subseteq \Nat$ such that $u_i \infleq u_{j}$ for all $i \leq j \in I$.
+
+    We can build a word $w$ as the limit of the sequence $\seqof{u_i}[i \in
+    I]$. This word is infinite or bi-infinite, and contains as infixes all the
+    words $u_i$ for $i \in I$. Because every word of $L$ is an infix of every
+    $u_i$ for a large enough $I$, one concludes that $L$ is contained in the
+    set of finite infixes of $w$. Conversely, every finite infix of $w$ is 
+    an infix of some $u_i$ by definition of the limit construction, hence
+    belongs to $L$ since $u_i \in L$ and $L$ is \kl{downwards closed}.
+\end{proofof}
+
+\begin{proofof}{ultimately-uniformly-recurrent:lem}
+
+    Assume that $w$ is \kl{ultimately uniformly recurrent}. Consider a sequence
+    of words $\seqof{w_i}[i \in \Nat]$ that are finite \kl{infixes} of $w$. Because $w$ is
+    \kl{ultimately uniformly recurrent}, there exists a bound $N_0$ such that
+    $w_{\geq N_0}$ is \kl{uniformly recurrent}. Let $i < N_0$, we claim that,
+    without loss of generality, only finitely many words in the sequence
+    $\seqof{w_i}[i \in \Nat]$ can be found starting at the position $i$ in $w$. Indeed, if
+    it is not the case, then we have an infinite subsequence of words that are
+    all comparable for the \kl{infix relation}, and therefore a \kl{good
+    sequence}, because the \kl{infix relation} is \kl{well-founded}. We can
+    therefore assume that all words in the sequence $\seqof{w_i}[i \in \Nat]$ are such that
+    they start at a position $i \geq N_0$. But then they are all finite
+    \kl{infixes} of $w_{\geq N_0}$, which is a \kl{uniformly recurrent} word,
+    whose set of finite \kl{infixes} is \kl{well-quasi-ordered}
+    (\cref{uniformly-recurrent:lem}).
+
+    Conversely, assume that the set of finite infixes of $w$ is
+    \kl{well-quasi-ordered}. Let us write $\mathsf{Rec}(w)$ the set of finite
+    \kl{infixes} of $w$ that appear infinitely often. We can similarly define
+    $\mathsf{Rec}(w_{\geq i})$ for any (infinite) suffix of $w$. The sequence
+    $R_i \defined \mathsf{Rec}(w_{\geq i})$ is a descending sequence of \kl{downwards
+    closed} sets of finite words, included in the set of finite infixes of $w$
+    by definition. Because the latter is \kl{well-quasi-ordered}, there exists
+    an $N_0 \in \Nat$, such that $\bigcap_{i \in \Nat} R_i = R_{N_0}$.
+    Now, consider $v \defined w_{\geq N_0}$. By construction, 
+    every finite infix of $v$ 
+    appears infinitely often in $v$. 
+    Given
+    some finite infix $u \infleq v$, we 
+    there exists a bound $N_u$
+    on the distance
+    between two consecutive occurrences of $u$ in $v$.
+    Indeed, if it is not the case, then there exists an infinite sequence
+    $\seqof{u x_i u}[i \in \Nat]$ of infixes of $v$, such that $x_i$ is a word
+    of size $\geq i$
+    and no shorter word $u y u$ is an infix of $u x_i u$.
+    Because the finite infixes of $w$ (hence, of $v$) are \kl{well-quasi-ordered},
+    one can extract an infinite set of indices $I \subseteq \Nat$
+    such that $u x_i u \infleq u x_{j} u$ for all $i \leq j \in I$.
+    In particular, $u x_i u \infleq u x_{j} u$ for some $j > |x_i|$, 
+    which contradicts the fact that $u x_j u$ coded two consecutive
+    occurrences of $u$ in $v$.
+
+    We have shown that for every finite infix $u$ of $v$, there exists a bound
+    $N_u$ such that every two occurrences of $u$ in $v$ start at distance at
+    most $N_u$. In particular, there exists a bound $M_u$ such that every infix
+    of $v$ of size at least $M_u$ contains $u$. We have proven that
+    $v$ is \kl{uniformly recurrent}, hence that $w$ is \kl{ultimately uniformly
+    recurrent}.
+\end{proofof}
+
+\begin{proofof}{bi-infinite-uur:lem}
+  Given a bi-infinite word $w \in \Sigma^{\Rel}$, we can consider $w_+ \in
+\Sigma^\Nat$ and $w_- \in \Sigma^\Nat$ the two infinite words obtained as
+follows: for all $i \in \Nat$, $(w_+)_i = w(i)$ and $(w_-)_i = w(-i)$. Note
+that the two share the letter at position $0$.
+
+    Assume that $w_+$ and $w_-$ are \kl{ultimately uniformly recurrent}. Let us
+    write $\infset{w}$ the set of finite infixes of $w$. Consider an infinite
+    sequence of words $\seqof{u_i}[i \in \Nat]$ in $\infset{w}$. If there is an
+    infinite subsequence of words that are all in $\infset{w_+}$, then there
+    exists an increasing pair of indices $i < j$ such that $u_i \infleq
+    u_j$ because \cref{uniformly-recurrent:lem} applies to $w_+$. Similarly, if
+    there is an infinite subsequence of words that are all in $\infset{w_-}$,
+    then there exists an increasing pair of indices $i < j$ such that $u_i
+    \infleq u_j$ because \cref{uniformly-recurrent:lem} applies to $w_-$ (and
+    the \kl{infix relation} is compatible with mirroring). Otherwise, one can
+    assume without loss of generality that all words in the sequence have a
+    starting position in $w_-$ and an ending position in $w_+$. In this case,
+    let us write $(k_i,l_i) \in \Nat^2$ the pair of indices such that $u_i$ is the infix
+    of $w$ that starts at position $-k_i$ of $w$ (i.e., $k_i$ of $w_-$) and
+    ends at position $l_i$ of $w$ (i.e., $l_i$ of $w_+$).
+    Because $\Nat^2$ is a \kl{well-quasi-ordering} with the product ordering,
+    there exists $i < j$ such that $k_i \leq k_j$ and $l_i \leq l_j$, 
+    in particular, $u_i \infleq u_j$. We have proven that every infinite
+    sequence of words in $\infset{w}$ is \kl(wqo){good}, hence $\infset{w}$ is
+    \kl{well-quasi-ordered}.
+
+    Conversely, assume that $\infset{w}$ is \kl{well-quasi-ordered}. In
+    particular, the subset $\infset{w_+} \subseteq \infset{w}$ is \kl{well-quasi-ordered}.
+    Similarly, $\infset{w_-}$ is \kl{well-quasi-ordered} because the \kl{infix
+    relation} is compatible with mirroring. Applying
+    \cref{ultimately-uniformly-recurrent:lem}, we conclude that both are
+    \kl{ultimately uniformly recurrent} words.
+
+\end{proofof}
+
+\begin{proofof}{from-bi-to-single:lem}
+  Given a bi-infinite word $w \in \Sigma^{\Rel}$, recall that we can consider $w_+ \in
+\Sigma^\Nat$ and $w_- \in \Sigma^\Nat$ the two infinite words obtained as
+follows: for all $i \in \Nat$, $(w_+)_i = w(i)$ and $(w_-)_i = w(-i)$. Note
+that the two share the letter at position $0$.
+
+    To obtain the upper bound of $3 \cdot \omegaOrd$, we can consider the same
+    argument as for \cref{small-ordinal-invariants:lem}. We let $N_0$ be such
+    that $w_{\geq N_0}$ and $(w_-)_{\geq N_0}$ are \kl{uniformly recurrent}
+    words. In any sequence of incomparable elements of $\infset{w}$, there are
+    less than $N_0^2$ elements that are found in $(w_{< N_0})_{> -N_0}$. Then,
+    one has to pick a finite \kl{infix} in either $w_{\geq N_0}$ or $w_{\leq
+    -N_0}$. Because of \cref{small-ordinal-invariants:lem}, any sequence of
+    incomparable elements of these two infinite words has length bounded based
+    on the choice of the first element of that sequence. This means that the
+    \kl{ordinal width} of $\infset{w}$ is at most $\omegaOrd + \omegaOrd +
+    N_0^2$. We conclude that $\oWidth{\infset{w}} < 3 \cdot \omegaOrd$.
+
+  Let us briefly argue that the bound is tight. Indeed, one can
+  construct a bi-infinite word $w$ by concatenating a reversed \kl{Thue-Morse
+  sequence} on a binary alphabet $\set{a,b}$, a finite antichain of arbitrarily
+  large size over a distinct alphabet $\set{c,d}$, and then a \kl{Thue-Morse
+  sequence} on a binary alphabet $\set{e,f}$. The \kl{ordinal width} of the set
+  of \kl{infixes} of $w$ is then at least $2 \cdot \omegaOrd + K$, where $K$ is the
+  size of the chosen antichain, following the same argument as in the proof of
+  \cref{small-ordinal-invariants:lem}, using \cref{thue-morse-ordinal:lemma}.
 \end{proofof}
@@
 \end{figure}
+
+\begin{proofof}{gap-words-prefix-ordered:lem}
+	Write $u$ for $\GapWord{f}{\ell}$ and $v$ for $\GapWord{g}{\ell}$. 
+	We may assume that both $u$ and $v$ are non-empty, as otherwise the lemma holds trivially.
+	Then, for all $k \in \Nat$, there exists a run with canonical decomposition
+	$$
+	w_k = L_0 a_1 \cdots a_n L_n,
+	$$
+	where $L_i \in \set{vv u^k, vu^kv, u^k vv}$ and specifically $L_\ell = vu^kv$.
+
+	From \cref{pumping-periods:lem}, we may conclude that there are a finite number of words $x, y,$ and $w$ 
+	such that each $w_k$ is contained in a language 
+	$\InfPeriodChain{x}w\InfPeriodChain{y}$.
+
+	As there is an infinite number of words $w_k$, 
+	we may fix $x, y,$ and $w$ and an infinite subset $I \subseteq \Nat$ 
+	such that $\set{w_i \mid i \in I} \subseteq \InfPeriodChain{x}w\InfPeriodChain{y}$. 
+	This implies that either for infinitely many $m \in \Nat$, $u^m v \in \InfPeriodChain{y}$ 
+	or for infinitely many $m$, $v u^m \in \InfPeriodChain{x}$.
+
+	In either case, we may conclude that either $u \infleq v$ or $v \infleq u$: Let $m, n \in \Nat$
+	such that $m < n$ and $u^m v, u^n v \in \InfPeriodChain{y}$ (the case for $v u^m$ and $v u^n$ 
+	proceeding analogously). Without loss of generality, assume that $\card{u^m}$ and $\card{u^n}$ are
+	multiples of $\card{y}$. We therefore find $p \prefleq y, s \suffleq y$ such that $u^m, u^n \in sy^*p$, 
+	ergo $ps = y$.
+	In other words, we can write $u^m = (sp)^{m'}, u^n = (sp)^{n'}$. As $u^mv \in \InfPeriodChain{y}$, it 
+	follows that $v$ is a prefix of some word in $(sp)^*$. Hence either $v$ is a prefix of $u$ or $u$ vice versa.
+\end{proofof}
+
+\begin{proofof}{dwclosed-infixes-wqo:lem}
+    It is clear that \cref{dwci-reg:item} $\Rightarrow$ \cref{dwci-aml:item}
+    because regular languages are recognized by finite automata, and finite
+    automata are a particular case of \kl{amalgamation systems}.
+    The implication \cref{dwci-aml:item} $\Rightarrow$ \cref{dwci-bod:item}
+    is the content of \cref{infix-amalgamation:thm}.
+    The implication \cref{dwci-bod:item} $\Rightarrow$ \cref{dwci-uoc:item}
+    is \cref{bounded-language:lem}.
+    Finally, the implication \cref{dwci-uoc:item} $\Rightarrow$ \cref{dwci-reg:item}
+    is simply because a \kl{downwards closed} language 
+    that is a finite union of products of \kl{chains} is a regular language.
+
+    Indeed, assume that
+    $L$ is \kl{downwards closed} and included in a finite union of sets of the form
+    $\InfPeriodChain{x} u \InfPeriodChain{y}$ where $x,y,u$ are possibly empty words.
+    We can assume without loss of generality that
+    for every $n$, $x^n u y^n$ is in $L$, otherwise, we have a bound on the maximal $n$ such that
+    $x^n u y^n$ is in $L$, and we can increase the number of languages in the union, replacing $x$ or $y$
+    with the empty word as necessary.
+    Let us write $L' \defined \bigcup_{i = 1}^k x_i^* u_i y_i^*$. Then, $L'
+    \subseteq L$ by construction. Furthermore, $L \subseteq \dwset{L'}$, also
+    by construction. Finally, we conclude that $L = \dwset{L'}$ because $L$ is
+    \kl{downwards closed}. Now, because $L'$ is a \kl{regular language}, and 
+    \kl{regular languages} are closed under \kl{downwards closure}, we conclude
+    that $L$ is a \kl{regular language}.
+\end{proofof}

### 2025-10-15 src/infixes.tex
new: moving stuff to appendix

--- 2025-10-15 src/infixes.tex
@@
 \AP In order to prove the (difficult) left-to-right implication of
-\cref{bounded-language:thm}, we will rely heavily on the
-combinatorics of periodic words. Let us recall that a non-empty word $w \in
-\Sigma^+$ is \intro(word){periodic} with period $x \in \Sigma^*$ if there
-exists a $p \in \Nat$ such that $w \infleq x^p$. The \intro{periodic length} of
-a word $u$ is the minimal length of a period $x$ of $u$.
-
-\begin{lemma}
-    \label{periodic-infixes:lem}
-    Let $u,v \in \Sigma^*$ be two (non-empty) \kl{periodic words}
-    having \kl{periods} of length $p$ and $q$ respectively.
-    If $u \infleq v$ and $\card{u} \geq 2\max\{p, q\}$,
-    then $u$ and $v$ both have a period of length
-    $\gcd(p,q)$.
-\end{lemma}
-\begin{proof}
-    Since $u$ has a period of length $p$ we have $u_{i+p} = u_i$ for all indices $i \in \Nat$ such that those letters are
-    well-defined. Since $u$ is an \kl{infix} of $v$ and $\card{u} \geq q$, we also have $u_{i+q} = u_i$. It follows that $u$ has a period of length $\gcd(p, q)$.
-
-    Since $v$ contains $u$ as a suffix and $\card{u} \geq q$, we can also conclude that $v$ must also have period of length $\gcd(p,q)$.
-\end{proof}
-
-\begin{figure}
-	\begin{center}
-		\input{fig/periodic-infixed.tex}
-	\end{center}
-	\caption{Alignment of two periodic words with periods of length $6$ and $4$, respectively. We can conclude that the periodic length is at most $2$.}
-	\label{fig:periodic-infixes}
-\end{figure}
-
-\begin{corollary}
-    \label{powers-infixes:cor}
-    Let $u,v \in \Sigma^*$ and $k, \ell \in \Nat$
-    such that $k \geq 2\max\{\card{u},\card{v}\}$,
-    and $u^k \infleq v^\ell$.
-    Then, there exists $w \in \Sigma^*$ of size at most
-    $\min \set{\card{u}, \card{v}}$ and a $p \in \Nat$
-    such that
-    $u^k \infleq v^\ell \infleq w^p$.
-\end{corollary}
+\cref{bounded-language:thm}, we will rely heavily on the combinatorics of
+periodic words. Let us use a slightly non-standard notation by saying that a
+non-empty word $w \in \Sigma^+$ is \intro(word){periodic} with period $x \in
+\Sigma^*$ if there exists a $p \in \Nat$ such that $w \infleq x^p$. The
+\intro{periodic length} of a word $u$ is the minimal length of a period $x$ of
+$u$.
 
@@
 the \kl{infix relation}. Note that in general, the \kl{downwards closure} of a
-chain is \emph{not} a chain (see \cref{dw-closure-not-wqo:rem}). However, for the chains generated using periodic
-words, the \kl{downwards closure} $\dwset[\infleq]{\setof{x^p}{p \in \Nat}}$ is
-a \emph{finite union} of \kl{chains}. Because this set will appear in bigger
-equations, we introduce the shorter notation $\intro*\InfPeriodChain{x}$ for
-the set of \kl{infixes} of words of the form $x^p$, where $p$ ranges over
-$\Nat$.
+chain is \emph{not} a chain (see \cref{dw-closure-not-wqo:rem}). However, for
+the chains generated using periodic words, the \kl{downwards closure}
+$\dwset[\infleq]{\setof{x^p}{p \in \Nat}}$ is a \emph{finite union} of
+\kl{chains}. Because this set will appear in bigger equations, we introduce the
+shorter notation $\intro*\InfPeriodChain{x}$ for the set of \kl{infixes} of
+words of the form $x^p$, where $p$ ranges over $\Nat$.
 
@@
 
-The following combinatorial lemma connects the property of being
-\kl{well-quasi-ordered} to a property of the \kl{periodic lengths} of words in
-a language, based on the assumption that some factors can be iterated. It is
-the core result that powers the analysis done in the upcoming
-\cref{bounded-language:thm,infix-amalgamation:thm}.
+The following combinatorial \cref{pumping-periods:lem} connects the property of
+being \kl{well-quasi-ordered} to a property of the \kl{periodic lengths} of
+words in a language, based on the assumption that some factors can be iterated.
+It is the core result that powers the analysis done in the upcoming
+\cref{bounded-language:thm,infix-amalgamation:thm}. It is fundamentally based
+on a classical result of combinatorics on words (\cref{periodic-gcd:lem}) that
+we recall here for the sake of completeness.
+
+\begin{lemma}[{\cite[Theorem 1]{FIWI65}}]
+  \label{periodic-gcd:lem}
+  Let $u,v \in \Sigma^+$ be two 
+  words and $n = \gcd(\card{u}, \card{v})$.
+  If there exists $p,q \in \Nat$ such that
+  $u^p$ and $v^q$ have a common prefix of length at least $\card{uv} - n$,
+  then there exists $z \in \Sigma^+$ such that $u$ and $v$ are powers of $z$,
+  and in particular $z$ has length at most $\min\{\card{u}, \card{v}\}$.
+\end{lemma}
 
@@
     \label{bounded-wqo-dwclosed:cor}
+    \proofref{bounded-wqo-dwclosed:cor}
     Let $L$ be a \kl{bounded language} of $\Sigma^*$. Then,
@@
 \end{corollary}
-\begin{proof}
-    Because $L \subseteq \dwset[\infleq]{L}$, the right-to-left implication
-    is trivial.
-    For the left-to-right implication, let us assume that $L$ is a
-    \kl{well-quasi-ordered} language for the \kl{infix relation}.
-    Then $L$ is included in a finite union 
-    of products of \kl{chains} for the \kl{prefix} and \kl{suffix} relations
-    thanks to
-    \cref{bounded-language:thm}:
-    \begin{equation*}
-        L \subseteq \bigcup_{i = 1}^n S_i \cdot P_i \quad .
-    \end{equation*}
-    Remark that if $S_i$ is a \kl{chain} for the \kl{suffix relation}
-    and $P_i$ is a \kl{chain} for the \kl{prefix relation},
-    then 
-    \begin{equation*}
-      \dwset[\infleq]{(S_i \cdot P_i)} = (\dwset[\suffleq]{S_i}) \cdot (\dwset[\prefleq]{P_i})
-        \quad .
-    \end{equation*}
-    Indeed, any \kl{infix} of a word in $S_i P_i$ can be split into
-    a \kl{suffix} of a word in $S_i$ and a \kl{prefix} of a word in $P_i$.
-    Conversely, any such concatenations are \kl{infixes} of a word in $S_i P_i$.
-
-    As a consequence, we conclude that $\dwset[\infleq]{L}$ is itself included
-    in a finite union of products of \kl{chains}.
-    Furthermore, by definition of \kl{bounded languages},
-    $\dwset[\infleq]{L}$ is also a \kl{bounded language}.
-    Hence, it is \kl{well-quasi-ordered} by the \kl{infix relation} 
-    via
-    \cref{bounded-language:thm}.
-\end{proof}
 
@@
     \label{dw-closed-wqo-charac:thm}
+    \proofref{dw-closed-wqo-charac:thm}
     Let $L$ be a \kl{well-quasi-ordered} language for the \kl{infix relation}
@@
     \label{bi-infinite:lem}
+    \proofref{bi-infinite:lem}
     Let $L \subseteq \Sigma^*$ be an \kl{order ideal} 
@@
 \end{lemma}
-\begin{proof}
-    Let us assume that $L$ is infinite. The case when it is finite 
-    is similar, but will result in a finite word.
-    \todo{aliaume: do we need wqo here? the proof
-      should go through without it: the sequence 
-      $u_i$ is already increasing for infix}
-
-    Because the alphabet $\Sigma$ is finite, we can enumerate the words of $L$
-    as $\seqof{w_i}[i \in \Nat]$. From $\seqof{w_i}[i \in \Nat]$, we construct a sequence $\seqof{u_i}[i \in \Nat]$ by induction
-    as follows: $u_0 = w_0$, and $u_{i+1}$ is a word that contains $u_i$ and
-    $w_i$, which exists in $L$ because $L$ is \kl(subset){directed}. Since $L$ is
-    \kl{well-quasi-ordered}, one can extract an infinite set of indices $I
-    \subseteq \Nat$ such that $u_i \infleq u_{j}$ for all $i \leq j \in I$.
-
-    We can build a word $w$ as the limit of the sequence $\seqof{u_i}[i \in
-    I]$. This word is infinite or bi-infinite, and contains as infixes all the
-    words $u_i$ for $i \in I$. Because every word of $L$ is an infix of every
-    $u_i$ for a large enough $I$, one concludes that $L$ is contained in the
-    set of finite infixes of $w$. Conversely, every finite infix of $w$ is 
-    an infix of some $u_i$ by definition of the limit construction, hence
-    belongs to $L$ since $u_i \in L$ and $L$ is \kl{downwards closed}.
-\end{proof}
 
@@
     \label{ultimately-uniformly-recurrent:lem}
+    \proofref{ultimately-uniformly-recurrent:lem}
     Let $w \in \Sigma^\Nat$ be an infinite word. 
@@
 \end{lemma}
-\begin{proof}
-
-    Assume that $w$ is \kl{ultimately uniformly recurrent}. Consider a sequence
-    of words $\seqof{w_i}[i \in \Nat]$ that are finite \kl{infixes} of $w$. Because $w$ is
-    \kl{ultimately uniformly recurrent}, there exists a bound $N_0$ such that
-    $w_{\geq N_0}$ is \kl{uniformly recurrent}. Let $i < N_0$, we claim that,
-    without loss of generality, only finitely many words in the sequence
-    $\seqof{w_i}[i \in \Nat]$ can be found starting at the position $i$ in $w$. Indeed, if
-    it is not the case, then we have an infinite subsequence of words that are
-    all comparable for the \kl{infix relation}, and therefore a \kl{good
-    sequence}, because the \kl{infix relation} is \kl{well-founded}. We can
-    therefore assume that all words in the sequence $\seqof{w_i}[i \in \Nat]$ are such that
-    they start at a position $i \geq N_0$. But then they are all finite
-    \kl{infixes} of $w_{\geq N_0}$, which is a \kl{uniformly recurrent} word,
-    whose set of finite \kl{infixes} is \kl{well-quasi-ordered}
-    (\cref{uniformly-recurrent:lem}).
-
-    Conversely, assume that the set of finite infixes of $w$ is
-    \kl{well-quasi-ordered}. Let us write $\mathsf{Rec}(w)$ the set of finite
-    \kl{infixes} of $w$ that appear infinitely often. We can similarly define
-    $\mathsf{Rec}(w_{\geq i})$ for any (infinite) suffix of $w$. The sequence
-    $R_i \defined \mathsf{Rec}(w_{\geq i})$ is a descending sequence of \kl{downwards
-    closed} sets of finite words, included in the set of finite infixes of $w$
-    by definition. Because the latter is \kl{well-quasi-ordered}, there exists
-    an $N_0 \in \Nat$, such that $\bigcap_{i \in \Nat} R_i = R_{N_0}$.
-    Now, consider $v \defined w_{\geq N_0}$. By construction, 
-    every finite infix of $v$ 
-    appears infinitely often in $v$. 
-    Given
-    some finite infix $u \infleq v$, we 
-    there exists a bound $N_u$
-    on the distance
-    between two consecutive occurrences of $u$ in $v$.
-    Indeed, if it is not the case, then there exists an infinite sequence
-    $\seqof{u x_i u}[i \in \Nat]$ of infixes of $v$, such that $x_i$ is a word
-    of size $\geq i$
-    and no shorter word $u y u$ is an infix of $u x_i u$.
-    Because the finite infixes of $w$ (hence, of $v$) are \kl{well-quasi-ordered},
-    one can extract an infinite set of indices $I \subseteq \Nat$
-    such that $u x_i u \infleq u x_{j} u$ for all $i \leq j \in I$.
-    In particular, $u x_i u \infleq u x_{j} u$ for some $j > |x_i|$, 
-    which contradicts the fact that $u x_j u$ coded two consecutive
-    occurrences of $u$ in $v$.
-
-    We have shown that for every finite infix $u$ of $v$, there exists a bound
-    $N_u$ such that every two occurrences of $u$ in $v$ start at distance at
-    most $N_u$. In particular, there exists a bound $M_u$ such that every infix
-    of $v$ of size at least $M_u$ contains $u$. We have proven that
-    $v$ is \kl{uniformly recurrent}, hence that $w$ is \kl{ultimately uniformly
-    recurrent}.
-\end{proof}
 
@@
   \label{bi-infinite-uur:lem}
+  \proofref{bi-infinite-uur:lem}
     Let $w \in \Sigma^\Rel$ be a bi-infinite word. Then, the set of finite
@@
 \end{lemma}
-\begin{proof}
-  Given a bi-infinite word $w \in \Sigma^{\Rel}$, we can consider $w_+ \in
-\Sigma^\Nat$ and $w_- \in \Sigma^\Nat$ the two infinite words obtained as
-follows: for all $i \in \Nat$, $(w_+)_i = w(i)$ and $(w_-)_i = w(-i)$. Note
-that the two share the letter at position $0$.
-
-    Assume that $w_+$ and $w_-$ are \kl{ultimately uniformly recurrent}. Let us
-    write $\infset{w}$ the set of finite infixes of $w$. Consider an infinite
-    sequence of words $\seqof{u_i}[i \in \Nat]$ in $\infset{w}$. If there is an
-    infinite subsequence of words that are all in $\infset{w_+}$, then there
-    exists an increasing pair of indices $i < j$ such that $u_i \infleq
-    u_j$ because \cref{uniformly-recurrent:lem} applies to $w_+$. Similarly, if
-    there is an infinite subsequence of words that are all in $\infset{w_-}$,
-    then there exists an increasing pair of indices $i < j$ such that $u_i
-    \infleq u_j$ because \cref{uniformly-recurrent:lem} applies to $w_-$ (and
-    the \kl{infix relation} is compatible with mirroring). Otherwise, one can
-    assume without loss of generality that all words in the sequence have a
-    starting position in $w_-$ and an ending position in $w_+$. In this case,
-    let us write $(k_i,l_i) \in \Nat^2$ the pair of indices such that $u_i$ is the infix
-    of $w$ that starts at position $-k_i$ of $w$ (i.e., $k_i$ of $w_-$) and
-    ends at position $l_i$ of $w$ (i.e., $l_i$ of $w_+$).
-    Because $\Nat^2$ is a \kl{well-quasi-ordering} with the product ordering,
-    there exists $i < j$ such that $k_i \leq k_j$ and $l_i \leq l_j$, 
-    in particular, $u_i \infleq u_j$. We have proven that every infinite
-    sequence of words in $\infset{w}$ is \kl(wqo){good}, hence $\infset{w}$ is
-    \kl{well-quasi-ordered}.
-
-    Conversely, assume that $\infset{w}$ is \kl{well-quasi-ordered}. In
-    particular, the subset $\infset{w_+} \subseteq \infset{w}$ is \kl{well-quasi-ordered}.
-    Similarly, $\infset{w_-}$ is \kl{well-quasi-ordered} because the \kl{infix
-    relation} is compatible with mirroring. Applying
-    \cref{ultimately-uniformly-recurrent:lem}, we conclude that both are
-    \kl{ultimately uniformly recurrent} words.
-
-\end{proof}
 
@@
 \end{lemma}
-\begin{proof}
-  Given a bi-infinite word $w \in \Sigma^{\Rel}$, recall that we can consider $w_+ \in
-\Sigma^\Nat$ and $w_- \in \Sigma^\Nat$ the two infinite words obtained as
-follows: for all $i \in \Nat$, $(w_+)_i = w(i)$ and $(w_-)_i = w(-i)$. Note
-that the two share the letter at position $0$.
-
-    To obtain the upper bound of $3 \cdot \omegaOrd$, we can consider the same
-    argument as for \cref{small-ordinal-invariants:lem}. We let $N_0$ be such
-    that $w_{\geq N_0}$ and $(w_-)_{\geq N_0}$ are \kl{uniformly recurrent}
-    words. In any sequence of incomparable elements of $\infset{w}$, there are
-    less than $N_0^2$ elements that are found in $(w_{< N_0})_{> -N_0}$. Then,
-    one has to pick a finite \kl{infix} in either $w_{\geq N_0}$ or $w_{\leq
-    -N_0}$. Because of \cref{small-ordinal-invariants:lem}, any sequence of
-    incomparable elements of these two infinite words has length bounded based
-    on the choice of the first element of that sequence. This means that the
-    \kl{ordinal width} of $\infset{w}$ is at most $\omegaOrd + \omegaOrd +
-    N_0^2$. We conclude that $\oWidth{\infset{w}} < 3 \cdot \omegaOrd$.
-
-  Let us briefly argue that the bound is tight. Indeed, one can
-  construct a bi-infinite word $w$ by concatenating a reversed \kl{Thue-Morse
-  sequence} on a binary alphabet $\set{a,b}$, a finite antichain of arbitrarily
-  large size over a distinct alphabet $\set{c,d}$, and then a \kl{Thue-Morse
-  sequence} on a binary alphabet $\set{e,f}$. The \kl{ordinal width} of the set
-  of \kl{infixes} of $w$ is then at least $2 \cdot \omegaOrd + K$, where $K$ is the
-  size of the chosen antichain, following the same argument as in the proof of
-  \cref{small-ordinal-invariants:lem}, using \cref{thue-morse-ordinal:lemma}.
-\end{proof}
 
@@
     Now, $\oWidth{L} \leq \sum_{i = 1}^n \oWidth{L_i} < 3 \cdot \omegaOrd <
-    \omegaOrd^2$ \todo{if inequality 2 holds, why do we not stop there?}. We conclude that the \kl{ordinal width} of $L$ is less than
+    \omegaOrd^2$. We conclude that the \kl{ordinal width} of $L$ is less than
     $\omegaOrd^2$. Finally, the inequality $\oType{L} \leq \oWidth{L} \oComProd

### 2025-10-15 src/infixes.tex

@@
 
-We are now ready to conclude the proof of \cref{small-ordinal-invariants:thm}.
+Ordinal invariants in hand, we proceed with the proof of \cref{small-ordinal-invariants:thm}.
 \begin{proofof}{small-ordinal-invariants:thm}

### 2025-10-15 src/appendix.tex
new: moving stuff to appendix (#10)
Co-authored-by: Aliaume LOPEZ <lopezaliaume@gmail.com>
hotfix: restoring stuff

@@
 \section{Proofs for Section~\ref{prefixes:sec}}
+
 \begin{figure}
@@
       \end{equation}
+  \end{proofof}
+
+\begin{lemma}
+	\label{automatic-uur:lem}
+	\proofref{automatic-uur:lem}
+	Given an \kl{automatic sequence} $w \in \Sigma^{\Nat}$, one can decide
+	whether it is \kl{ultimately uniformly recurrent}.
+\end{lemma}
+
+\begin{proofof}{automatic-uur:lem}
+	We can rewrite this as a question on the \kl{automatic sequence} $w$
+	as follows:
+	\begin{align*}
+		&\exists N_0,                   &   \text{ultimately} \\
+		&\forall i_s \geq N_0,          &   \text{for every infix (start) } u \\
+		&\forall i_e > i_s,             &   \text{for every infix (end) }   u \\
+		&\exists k \geq 1,              &   \text{there exists a bound} \\
+		&\forall j_s \geq N_0,          &   \text{for every other infix (start) } v \\
+		&\forall j_e \geq j_s + k,      &   \text{of size at least $k$} \\
+		&\exists l \geq 0,              &   \text{there exists a position in } v \\
+		&\forall 0 \leq m < i_e - i_s,  &   \text{where } u \text{ can be found} \\
+		&j_s + m + l < j_e \land
+		w(i_s + m) = w(j_s + m + l) \quad .
+	\end{align*}
+	Because $w$ is computable by a finite automaton, one can reduce the above
+	formula to a regular language, for which it suffices to check emptiness, which
+	is decidable.
+\end{proofof}
+
+\begin{proofof}{small-ordinal-invariants:lem}
+	Let $N_0$ be a bound such that $w_{\geq N_0}$ is \kl{uniformly recurrent}.
+	Let us write $\infset{w}$ the set of finite infixes of $w$.
+	We prove that $\oWidth{\infset{w}} \leq \omegaOrd + N_0$.
+	Let $u_1 \infleq w$ be a finite word.
+
+	If $u_1$ is an \kl{infix} of $w_{\geq N_0}$, then there exists $k \geq 1$
+	such that $u_1$ is an \kl{infix} of every word of size at least $k$. In
+	particular, there is finite bound on the length of every sequence of
+	incomparable elements starting with $u_1$. We conclude in particular that
+	$\infset{w} \setminus \upset{u_1}$ has a finite \kl{ordinal width}.
+
+	Otherwise, $u_1$ can only be found \emph{before} $N_0$. In this case, we
+	consider a second element of a \kl{bad sequence} $u_2 \infleq w$, which is
+	incomparable with $u_1$ for the \kl{infix relation}. If $u_2$ is an
+	\kl{infix} of $w_{\geq N_0}$, then we can conclude as before. Otherwise,
+	notice that $u_1$ and $u_2$ cannot start at the same position in $w$
+	(because they are incomparable). Continuing this argument, we conclude that
+	there are at most $N_0$ elements starting before $N_0$
+	at the start of any sequence of
+	incomparable elements in $\infset{w}$. We conclude that
+	$\oWidth{\infset{w}} \leq \omegaOrd + N_0$.
+
+	Let us now justify that this bound is tight.
+	The \kl{Thue-Morse sequence} over a binary
+	alphabet $\set{a,b}$ has \kl{ordinal width} $\omegaOrd$
+	from \cref{thue-morse-ordinal:lemma}.
+	Given a number $N_0
+	\in \Nat$, one can construct an arbitrarily long \kl{antichain} of words for
+	the \kl{infix relation} by using a new letter $c$. When concatenating this
+	(finite) antichain as a prefix of the \kl{Thue-Morse sequence}, one obtains a
+	new (infinite) word $w$. It is clear that the \kl{ordinal width} of
+	$\infset{w}$ is now at least $\omegaOrd + N_0$.
 \end{proofof}
@@
 
+For this paper to be self-contained, we will also recall how runs of a finite
+state automaton can be understood as an \kl{amalgamation system}.
+
+\begin{example}[{\cite[Section 3.2]{ASZZ24}}]
+  \label{automaton-amalgamation:example}
+    Let $A = (Q, \delta, q_0, F)$ be a finite state automaton over a finite
+    alphabet $\Sigma$. Let $\Delta$ be the set of transitions $(q_1, a, q_2)
+    \in Q \times \Sigma \times Q$,
+    and $R \subseteq \Delta^*$ be the set of 
+    words over transitions that start with the initial state $q_0$,
+    end in a final state $q_f \in F$, and such that the end state of a
+    letter is the start state of the following one.
+    The canonical decomposition $\canrun$
+    is defined as a morphism from $\Delta^*$ to $\Sigma^*$
+    that maps $(q,a,p)$ to $a$. 
+    Because of the one-to-one correspondence of steps of a run $\rho$ and letters in its \kl{canonical decomposition}, 
+    we may treat the two interchangeably.
+    Finally, given two runs $\rho$ and $\sigma$ of the automaton,
+    we say that an embedding $f \in \HigEmb(\canrun(\rho), \canrun(\sigma))$
+    belongs to $E(\rho,\sigma)$ when
+    $f$ is also defining an embedding from $\rho$ to $\sigma$ as words in $\Delta^*$.
+
+    The system $(\Sigma, R, E, \canrun)$ is an \kl{amalgamation system},
+    whose language is precisely the language of words recognized
+    by the automaton $A$.
+\end{example}
+\begin{proof}
+    By definition, the embeddings inside $E(\rho,\sigma)$ are in
+    of $\HigEmb(\canrun(\rho), \canrun(\sigma))$, and they compose properly.
+    Because $\Delta = Q \times \Sigma \times Q$ is finite, it is 
+    a \kl{well-quasi-ordering} when equipped with the equality relation, and 
+    we conclude that $\Delta^*$ with $\higleq$ is a \kl{well-quasi-order}
+    according to Higman’s Lemma \cite{HIG52}.
+
+    Let us now move to proving that the system satisfies the amalgamation
+    property. Given three runs $\rho,\sigma,\tau \in R$, and two embeddings $f \in E(\rho,\sigma)$
+    and $g \in E(\rho,\tau)$, we want to construct an amalgamated run $\sigma \vee \tau$.
+    Because letters in the run $\rho$ respect the transitions of the automaton
+    (i.e., if the letter $i$ ends in state $q$, then the letter $i+1$ starts in
+    state $q$), then the \kl{gap word} at position $i$ starts in state $q$ and
+    ends in state $q$ too. This means that for both embeddings
+    $f$ and $g$, the \kl{gap words} are read by the automaton by looping
+    on a state. In particular, these loops can be taken in any order
+    and continue to represent a valid run. That is, we can even select
+    the order of concatenation in the amalgamation for \emph{all} 
+    $0 \leq i \leq \card{\canrun(\rho)}$ and not just for one separately.
+
+    We conclude by remarking that 
+    the language of this amalgamation system is
+    the set of $\yieldrun(R)$, 
+    because $R$ is the set of valid runs of the automaton,
+    and $\yieldrun(\rho)$ is the word read along a run $\rho$.
+\end{proof}
+
 \begin{proofof}{gap-words-prefix-ordered:lem}

### 2025-10-15 src/amalgam.tex
fix: making knowledge work

@@
 class of all regular language, or the class of all context-free languages. Such
-classes will be called \kl[effectiveness assumptions@amalgamation]{effective amalgamative classes} (\kcref{effectiveness assumptions@amalgamation})\todo{can't get the knowledge to work}.
-In the following theorem, we prove that under such
-assumptions, testing \kl{well-quasi-ordering} is inter-reducible to testing
-whether a language of the class is empty, which is usually the simplest
-problem for a computational model.
+classes will be called \kl(amalgamation)[effectiveness assumptions]{effective
+amalgamative classes} (\kcref{effective amalgamation system}). In the
+following theorem, we prove that under such assumptions, testing
+\kl{well-quasi-ordering} is inter-reducible to testing whether a language of
+the class is empty, which is usually the simplest problem for a computational
+model.
 
@@
 
-%For this paper to be self-contained, we
-%will also recall how runs of a finite state automaton can be understood as an
-%\kl{amalgamation system}.
-%
-%\begin{example}[{\cite[Section 3.2]{ASZZ24}}]
-%    Let $A = (Q, \delta, q_0, F)$ be a finite state automaton over a finite
-%    alphabet $\Sigma$. Let $\Delta$ be the set of transitions $(q_1, a, q_2)
-%    \in Q \times \Sigma \times Q$,
-%    and $R \subseteq \Delta^*$ be the set of 
-%    words over transitions that start with the initial state $q_0$,
-%    end in a final state $q_f \in F$, and such that the end state of a
-%    letter is the start state of the following one.
-%    The canonical decomposition $\canrun$
-%    is defined as a morphism from $\Delta^*$ to $\Sigma^*$
-%    that maps $(q,a,p)$ to $a$. 
-%    Because of the one-to-one correspondence of steps of a run $\rho$ and letters in its \kl{canonical decomposition}, 
-%    we may treat the two interchangeably.
-%    Finally, given two runs $\rho$ and $\sigma$ of the automaton,
-%    we say that an embedding $f \in \HigEmb(\canrun(\rho), \canrun(\sigma))$
-%    belongs to $E(\rho,\sigma)$ when
-%    $f$ is also defining an embedding from $\rho$ to $\sigma$ as words in $\Delta^*$.
-%
-%    The system $(\Sigma, R, E, \canrun)$ is an \kl{amalgamation system},
-%    whose language is precisely the language of words recognized
-%    by the automaton $A$.
-%\end{example}
-%\begin{proof}
-%    By definition, the embeddings inside $E(\rho,\sigma)$ are in
-%    of $\HigEmb(\canrun(\rho), \canrun(\sigma))$, and they compose properly.
-%    Because $\Delta = Q \times \Sigma \times Q$ is finite, it is 
-%    a \kl{well-quasi-ordering} when equipped with the equality relation, and 
-%    we conclude that $\Delta^*$ with $\higleq$ is a \kl{well-quasi-order}
-%    according to Higman’s Lemma \cite{HIG52}.
-%    
-%    Let us now move to proving that the system satisfies the amalgamation
-%    property. Given three runs $\rho,\sigma,\tau \in R$, and two embeddings $f \in E(\rho,\sigma)$
-%    and $g \in E(\rho,\tau)$, we want to construct an amalgamated run $\sigma \vee \tau$.
-%    Because letters in the run $\rho$ respect the transitions of the automaton
-%    (i.e., if the letter $i$ ends in state $q$, then the letter $i+1$ starts in
-%    state $q$), then the \kl{gap word} at position $i$ starts in state $q$ and
-%    ends in state $q$ too. This means that for both embeddings
-%    $f$ and $g$, the \kl{gap words} are read by the automaton by looping
-%    on a state. In particular, these loops can be taken in any order
-%    and continue to represent a valid run. That is, we can even select
-%    the order of concatenation in the amalgamation for \emph{all} 
-%    $0 \leq i \leq \card{\canrun(\rho)}$ and not just for one separately.
-%
-%    We conclude by remarking that 
-%    the language of this amalgamation system is
-%    the set of $\yieldrun(R)$, 
-%    because $R$ is the set of valid runs of the automaton,
-%    and $\yieldrun(\rho)$ is the word read along a run $\rho$.
-%\end{proof}
-
 \subsection{$\infleq$-Well-Quasi-Ordered Amalgamation Systems}
@@
 
-\AP To construct a decision procedure for well-quasi-orderedness under $\infleq$, we 
-our \kl{amalgamation systems} to satisfy certain \intro{effectiveness assumptions}. We require that for an
-\kl{amalgamation system} $(\Sigma, R, E, \canrun)$, $R$ is
-recursively enumerable, the function $\canrun(\cdot)$ is computable, and for
-any two runs $\rho, \sigma \in R$, the set $E(\rho,\sigma)$ is computable. Additionally, we require the class to be effectively closed under rational transductions \cite[Chapter 5, page 64]{BERST79}.
+\AP To construct a decision procedure for well-quasi-orderedness under
+$\infleq$, we our \kl{amalgamation systems} to satisfy certain
+\intro(amalgamation){effectiveness assumptions}. We require that for an
+\kl{amalgamation system} $(\Sigma, R, E, \canrun)$, $R$ is recursively
+enumerable, the function $\canrun(\cdot)$ is computable, and for any two runs
+$\rho, \sigma \in R$, the set $E(\rho,\sigma)$ is computable. Additionally, we
+require the class to be effectively closed under rational transductions
+\cite[Chapter 5, page 64]{BERST79}.
 

### 2025-10-15 src/amalgam.tex
fix: trimming

@@
 
-\begin{proof}
-	Write $u$ for $\GapWord{f}{\ell}$ and $v$ for $\GapWord{g}{\ell}$. 
-	We may assume that both $u$ and $v$ are non-empty, as otherwise the lemma holds trivially.
-	Then, for all $k \in \Nat$, there exists a run with canonical decomposition
-	$$
-	w_k = L_0 a_1 \cdots a_n L_n,
-	$$
-	where $L_i \in \set{vv u^k, vu^kv, u^k vv}$ and specifically $L_\ell = vu^kv$.
-
-	\todo[inline]{The following can probably be replaced by a reference to Fine \& Wilf}
-	From \cref{pumping-periods:lem}, we may conclude that there are a finite number of words $x, y,$ and $w$ 
-	such that each $w_k$ is contained in a language 
-	$\InfPeriodChain{x}w\InfPeriodChain{y}$.
-
-	As there is an infinite number of words $w_k$, 
-	we may fix $x, y,$ and $w$ and an infinite subset $I \subseteq \Nat$ 
-	such that $\set{w_i \mid i \in I} \subseteq \InfPeriodChain{x}w\InfPeriodChain{y}$. 
-	This implies that either for infinitely many $m \in \Nat$, $u^m v \in \InfPeriodChain{y}$ 
-	or for infinitely many $m$, $v u^m \in \InfPeriodChain{x}$.
-
-	In either case, we may conclude that either $u \infleq v$ or $v \infleq u$: Let $m, n \in \Nat$
-	such that $m < n$ and $u^m v, u^n v \in \InfPeriodChain{y}$ (the case for $v u^m$ and $v u^n$ 
-	proceeding analogously). Without loss of generality, assume that $\card{u^m}$ and $\card{u^n}$ are
-	multiples of $\card{y}$. We therefore find $p \prefleq y, s \suffleq y$ such that $u^m, u^n \in sy^*p$, 
-	ergo $ps = y$.
-	In other words, we can write $u^m = (sp)^{m'}, u^n = (sp)^{n'}$. As $u^mv \in \InfPeriodChain{y}$, it 
-	follows that $v$ is a prefix of some word in $(sp)^*$. Hence either $v$ is a prefix of $u$ or $u$ vice versa.
-\end{proof}
-
 \begin{proofof}{infix-amalgamation:thm}
@@
 \end{corollary}
-
-\begin{remark}
-	Let us conclude by noting that it is unsurprisingly not possible to decide
-	whether a decidable language is \kl{well-quasi-ordered} by the \kl{prefix
-	relation} in general. 
-	%This is a very easy result whose sole purpose is to contrast with
-	%the decidability result of \cref{aut-cfg-infix:cor}.
-
-\end{remark}
-\begin{proof}
-    We reduce the halting problem on the empty string $\varepsilon$.
-    Let $M$ be a Turing Machine, we write the languages $L$ of finite runs
-    of $M$ starting on the empty string,
-    that we surround by special markers. This language is decidable,
-    and 
-    is
-    \kl{well-quasi-ordered} if and only if it is finite
-    if and only if $M$ terminates on $\varepsilon$.
-\end{proof}

### 2025-10-15 src/appendix.tex
fix: trimming

@@
 \clearpage
+
+\section{Proofs for Section~\ref{introduction:sec}}
+
+\begin{figure}
+    \centering
+    \begin{subfigure}[t]{0.48\textwidth}
+    	\centering
+    	\includestandalone[width=\linewidth]{fig/prefix-embedding-standalone}
+    	\caption{Prefix Relation}
+   	\end{subfigure}%
+   	\hfill%
+   	\begin{subfigure}[t]{0.48\textwidth}
+   		\centering
+   		\includestandalone[width=\linewidth]{fig/infix-embedding-standalone}
+   		\caption{Infix Relation}
+   	\end{subfigure}
+   	\begin{subfigure}[t]{0.48\textwidth}
+   		\centering
+   		\includestandalone[width=\linewidth]{fig/subword-embedding-standalone}
+   		\caption{Subword Relation}
+   	\end{subfigure}
+
+   	\caption{A simple representation of the \kl{subword relation},
+        \kl{prefix relation},
+        and \kl{infix relation},
+        on the alphabet $\set{a,b}$ for words of
+        length at most $3$. The figures are Hasse Diagrams,
+        representing the successor relation of the order.
+        Furthermore, we highlight in dashed red relations that are added
+        when moving from the prefix relation to the infix one,
+        and to the infix relation to the subword one.}
+    \label{word-embeddings:fig}
+\end{figure}
 
@@
     and we conclude that $u x^{\ell} v \suffleq w u x^\ell v = x^{\ell + 1} v \suffleq u x^k v$.
+\end{proofof}
+
+\begin{proofof}{pumping-periods:lem}
+    Note that the result is obvious if $k = 0$, and therefore
+    we assume $k \geq 1$ in the following proof.
+
+    Let us construct a sequence of words $\seqof{w_i}[i \in \Nat]$, where $w_i
+    \defined w[\vec{n_i}]$ for some well-chosen indices $\vec{n_i} \in \Nat^k$. The goal
+    being that 
+    if $w[\vec{n_i}]$ is an \kl{infix} of $w[\vec{n_j}]$,
+    then it can intersect at most \emph{two} iterated words,
+    with an intersection that is long enough to successfully apply
+    \cref{periodic-infixes:lem}.
+    In order to achieve this,
+    let us first define $s$ as the minimal size of a word $v_i$
+    ($1 \leq i \leq k$) and $u_j$ ($1 \leq j \leq k+1$).
+    Then,
+    we consider $\vec{n_0} \in \Nat^k$ such that $\vec{n_0}$ has all 
+    its components greater than $\factorial{s}$ and such that
+    $w[\vec{n_0}]$ belongs to $L$. 
+    Then, we inductively define 
+    $\vec{n_{i+1}}$  as the smallest vector of numbers greater than $\vec{n_i}$,
+    such that $w[\vec{n_{i+1}}]$ belongs to $L$, 
+    and with $\vec{n_i}$ having all components greater than
+    $2\card{w[\vec{n_i}]}$.
+
+    Let us assume that $k \geq 2$ in the following proof for symmetry purposes,
+    and argue later on that when $k = 1$ the same argument goes through.
+    Because $L$ is \kl{well-quasi-ordered} by the \kl{infix relation}, there
+    exists $i < j$ such that $w[\vec{n_i}]$ is an \kl{infix} of $w[\vec{n_j}]$.
+    Now, because of the chosen values for $\vec{n_j}$, there exists $1 \leq \ell \leq
+    k-1$ such that
+    one of the three following equations holds:
+    \begin{itemize}
+        \item $w[\vec{n_i}] \infleq v_{\ell}^{n_{j,\ell}} u_{\ell+1} v_{\ell+1}^{n_{j,\ell+1}}$,
+        \item $w[\vec{n_i}] \infleq u_{\ell}
+            v_{\ell}^{n_{j,\ell}}$,
+        \item $w[\vec{n_i}] \infleq
+            v_{\ell}^{n_{j,\ell}} u_{\ell+1}$.
+    \end{itemize}
+    In all those cases, we conclude using \cref{powers-infixes:cor}
+    that there exists $x,y \in \Sigma^+$ of size at most $s$, and 
+    a number $1 \leq t \leq k$ such that for all $z \in \Nat$,
+    $v_i^{z} \in \InfPeriodChain{x}$ for all $1 \leq i \leq t$,
+    and
+    $v_i^{z} \in \InfPeriodChain{y}$ for all $t < i \leq k$.
+    In particular,
+    $w[\vec{n_i}] \in \InfPeriodChain{x} u_{t} \InfPeriodChain{y}$.
+
+    \todo[inline]{Aliaume: proof of concept proof} Let us now conclude that the
+    same $x,y$ work for all $\vec{n} \in \Nat^k$. The reasoning is that if
+    $w[\vec{n}]$ is of the form $u_1 x^p x'$ where $x' \prefleq x$, and $x$ is
+    primitive (not a power of a smaller word), then all the iterated words
+    $v_i^{n_i}$ must be powers of a conjugate of $x$ (cyclic shift of $x$), and
+    all the $u_i$'s must be such that the concatenation of the $u_i$'s and
+    $v_i$'s align properly with $x$. Since the length of $x$ is at most $\min
+    \setof{\card{v_i}}{1 \leq i \leq k}$, we conclude that the $v_i$'s
+    themselves are powers of a conjugate of $x$, and that the $u_i$'s are
+    suffixes or prefixes of $x$. Now, we conclude that all the words
+    $w[\vec{n}]$ are of the correct form. Similar reasoning applies for the
+    other two cases.
+
+    When $k = 1$, the situation is a bit more specific since we only have two
+    cases: either $w_i \infleq u_1 v_1^{n_j}$ or $w_i \infleq v_1^{n_j} u_2$,
+    and we conclude with an identical reasoning.
 \end{proofof}

### 2025-10-15 src/infixes.tex
hotfix: restoring stuff
fix: trimming

@@
     \label{pumping-periods:lem}
+    \proofref{pumping-periods:lem}
     Let $L \subseteq \Sigma^*$ be a language
@@
 \end{lemma}
-\begin{proof}
-    Note that the result is obvious if $k = 0$, and therefore
-    we assume $k \geq 1$ in the following proof.
-
-    Let us construct a sequence of words $\seqof{w_i}[i \in \Nat]$, where $w_i
-    \defined w[\vec{n_i}]$ for some well-chosen indices $\vec{n_i} \in \Nat^k$. The goal
-    being that 
-    if $w[\vec{n_i}]$ is an \kl{infix} of $w[\vec{n_j}]$,
-    then it can intersect at most \emph{two} iterated words,
-    with an intersection that is long enough to successfully apply
-    \cref{periodic-infixes:lem}.
-    In order to achieve this,
-    let us first define $s$ as the minimal size of a word $v_i$
-    ($1 \leq i \leq k$) and $u_j$ ($1 \leq j \leq k+1$).
-    Then,
-    we consider $\vec{n_0} \in \Nat^k$ such that $\vec{n_0}$ has all 
-    its components greater than $\factorial{s}$ and such that
-    $w[\vec{n_0}]$ belongs to $L$. 
-    Then, we inductively define 
-    $\vec{n_{i+1}}$  as the smallest vector of numbers greater than $\vec{n_i}$,
-    such that $w[\vec{n_{i+1}}]$ belongs to $L$, 
-    and with $\vec{n_i}$ having all components greater than
-    $2\card{w[\vec{n_i}]}$.
-
-    Let us assume that $k \geq 2$ in the following proof for symmetry purposes,
-    and argue later on that when $k = 1$ the same argument goes through.
-    Because $L$ is \kl{well-quasi-ordered} by the \kl{infix relation}, there
-    exists $i < j$ such that $w[\vec{n_i}]$ is an \kl{infix} of $w[\vec{n_j}]$.
-    Now, because of the chosen values for $\vec{n_j}$, there exists $1 \leq \ell \leq
-    k-1$ such that
-    one of the three following equations holds:
-    \begin{itemize}
-        \item $w[\vec{n_i}] \infleq v_{\ell}^{n_{j,\ell}} u_{\ell+1} v_{\ell+1}^{n_{j,\ell+1}}$,
-        \item $w[\vec{n_i}] \infleq u_{\ell}
-            v_{\ell}^{n_{j,\ell}}$,
-        \item $w[\vec{n_i}] \infleq
-            v_{\ell}^{n_{j,\ell}} u_{\ell+1}$.
-    \end{itemize}
-    In all those cases, we conclude using \cref{powers-infixes:cor}
-    that there exists $x,y \in \Sigma^+$ of size at most $s$, and 
-    a number $1 \leq t \leq k$ such that for all $z \in \Nat$,
-    $v_i^{z} \in \InfPeriodChain{x}$ for all $1 \leq i \leq t$,
-    and
-    $v_i^{z} \in \InfPeriodChain{y}$ for all $t < i \leq k$.
-    In particular,
-    $w[\vec{n_i}] \in \InfPeriodChain{x} u_{t} \InfPeriodChain{y}$.
-
-    \todo[inline]{Aliaume: proof of concept proof} Let us now conclude that the
-    same $x,y$ work for all $\vec{n} \in \Nat^k$. The reasoning is that if
-    $w[\vec{n}]$ is of the form $u_1 x^p x'$ where $x' \prefleq x$, and $x$ is
-    primitive (not a power of a smaller word), then all the iterated words
-    $v_i^{n_i}$ must be powers of a conjugate of $x$ (cyclic shift of $x$), and
-    all the $u_i$'s must be such that the concatenation of the $u_i$'s and
-    $v_i$'s align properly with $x$. Since the length of $x$ is at most $\min
-    \setof{\card{v_i}}{1 \leq i \leq k}$, we conclude that the $v_i$'s
-    themselves are powers of a conjugate of $x$, and that the $u_i$'s are
-    suffixes or prefixes of $x$. Now, we conclude that all the words
-    $w[\vec{n}]$ are of the correct form. Similar reasoning applies for the
-    other two cases.
-
-    When $k = 1$, the situation is a bit more specific since we only have two
-    cases: either $w_i \infleq u_1 v_1^{n_j}$ or $w_i \infleq v_1^{n_j} u_2$,
-    and we conclude with an identical reasoning.
-\end{proof}
 
@@
     Now, $\oWidth{L} \leq \sum_{i = 1}^n \oWidth{L_i} < 3 \cdot \omegaOrd <
-    \omegaOrd^2$ \todo{if inequality 2 holds, why do we not stop there?}. We conclude that the \kl{ordinal width} of $L$ is less than
-    $\omegaOrd^2$. Finally, the inequality $\oType{L} \leq \oWidth{L} \oComProd
-    \oHeight{L} < \omega \oComProd \omega^2 = \omegaOrd^3$ allows us to conclude.
+    \omegaOrd^2$. Finally, the inequality $\oType{L} \leq \oWidth{L} \oComProd
+    \oHeight{L} < \omega \oComProd \omega^2 = \omegaOrd^3$ allows us to
+    conclude.
 
@@
 (trivially) translated into a finite automaton that reads the base $b$
-representation of the free variables (that are indices of the sequence).
-It is folklore that one can decide if an \kl{automatic sequence} is \kl{ultimately uniformly recurrent}.
-A proof can be found in the appendix at \cref{automatic-uur:lem}.
-Based on this, we now prove:
+representation of the free variables (that are indices of the sequence). It is
+folklore that one can decide if an \kl{automatic sequence} is \kl{ultimately
+uniformly recurrent}. A proof can be found in the appendix at
+\cref{automatic-uur:lem}. Based on this, we now prove:
 

### 2025-10-15 src/introduction.tex
fix: trimming

@@
 $i \in \{1, \ldots, |u|\}$.
-
-\begin{figure}
-    \centering
-    \begin{subfigure}[t]{0.48\textwidth}
-    	\centering
-    	\includestandalone[width=\linewidth]{fig/prefix-embedding-standalone}
-    	\caption{Prefix Relation}
-   	\end{subfigure}%
-   	\hfill%
-   	\begin{subfigure}[t]{0.48\textwidth}
-   		\centering
-   		\includestandalone[width=\linewidth]{fig/infix-embedding-standalone}
-   		\caption{Infix Relation}
-   	\end{subfigure}
-   	\begin{subfigure}[t]{0.48\textwidth}
-   		\centering
-   		\includestandalone[width=\linewidth]{fig/subword-embedding-standalone}
-   		\caption{Subword Relation}
-   	\end{subfigure}
-
-   	\caption{A simple representation of the \kl{subword relation},
-        \kl{prefix relation},
-        and \kl{infix relation},
-        on the alphabet $\set{a,b}$ for words of
-        length at most $3$. The figures are Hasse Diagrams,
-        representing the successor relation of the order.
-        Furthermore, we highlight in dashed red relations that are added
-        when moving from the prefix relation to the infix one,
-        and to the infix relation to the subword one.}
-    \label{word-embeddings:fig}
-\end{figure}
 

### 2025-10-15 src/appendix.tex
fix: trying my best

@@
     two words $wu$ and $wv$ both in $L$ that are not comparable for the
-    \kl{prefix relation}. Let $S = S_{\mathrm{split}}~\cup~\min_{\prefleq} L$\todo{kinda ugly, but i can't find a nicer way of phrasing it}. 
+    \kl{prefix relation}. Let $S = S_{\mathrm{split}}~\cup~\min_{\prefleq} L$
     Assume by contradiction that $S$ is infinite.
@@
     Let us construct a sequence of words $\seqof{w_i}[i \in \Nat]$, where $w_i
-    \defined w[\vec{n_i}]$ for some well-chosen indices $\vec{n_i} \in \Nat^k$. The goal
-    being that 
-    if $w[\vec{n_i}]$ is an \kl{infix} of $w[\vec{n_j}]$,
-    then it can intersect at most \emph{two} iterated words,
-    with an intersection that is long enough to successfully apply
-    \cref{periodic-infixes:lem}.
-    In order to achieve this,
-    let us first define $s$ as the minimal size of a word $v_i$
-    ($1 \leq i \leq k$) and $u_j$ ($1 \leq j \leq k+1$).
-    Then,
-    we consider $\vec{n_0} \in \Nat^k$ such that $\vec{n_0}$ has all 
-    its components greater than $\factorial{s}$ and such that
-    $w[\vec{n_0}]$ belongs to $L$. 
-    Then, we inductively define 
-    $\vec{n_{i+1}}$  as the smallest vector of numbers greater than $\vec{n_i}$,
-    such that $w[\vec{n_{i+1}}]$ belongs to $L$, 
-    and with $\vec{n_i}$ having all components greater than
-    $2\card{w[\vec{n_i}]}$.
+    \defined w[\vec{n_i}]$ for some well-chosen indices $\vec{n_i} \in \Nat^k$.
+    The goal being that if $w[\vec{n_i}]$ is an \kl{infix} of $w[\vec{n_j}]$,
+    then it can intersect at most \emph{two} iterated words, with an
+    intersection that is long enough to successfully apply
+    \cref{periodic-gcd:lem}. In order to achieve this, let us first define $s$
+    as the maximal size of a word $v_i$ ($1 \leq i \leq k$) and $u_j$ ($1 \leq
+    j \leq k+1$). Then, we consider $\vec{n_0} \in \Nat^k$ such that
+    $\vec{n_0}$ has all its components greater than $\factorial{s}$ and such
+    that $w[\vec{n_0}]$ belongs to $L$. Then, we inductively define
+    $\vec{n_{i+1}}$  as the smallest vector of numbers greater than
+    $\vec{n_i}$, such that $w[\vec{n_{i+1}}]$ belongs to $L$, and with
+    $\vec{n_i}$ having all components greater than $2\card{w[\vec{n_i}]}$.
 
@@
     \end{itemize}
-    In all those cases, we conclude using \cref{powers-infixes:cor}
-    that there exists $x,y \in \Sigma^+$ of size at most $s$, and 
-    a number $1 \leq t \leq k$ such that for all $z \in \Nat$,
-    $v_i^{z} \in \InfPeriodChain{x}$ for all $1 \leq i \leq t$,
-    and
-    $v_i^{z} \in \InfPeriodChain{y}$ for all $t < i \leq k$.
-    In particular,
-    $w[\vec{n_i}] \in \InfPeriodChain{x} u_{t} \InfPeriodChain{y}$.
 
-    \todo[inline]{Aliaume: proof of concept proof} Let us now conclude that the
-    same $x,y$ work for all $\vec{n} \in \Nat^k$. The reasoning is that if
-    $w[\vec{n}]$ is of the form $u_1 x^p x'$ where $x' \prefleq x$, and $x$ is
-    primitive (not a power of a smaller word), then all the iterated words
-    $v_i^{n_i}$ must be powers of a conjugate of $x$ (cyclic shift of $x$), and
-    all the $u_i$'s must be such that the concatenation of the $u_i$'s and
-    $v_i$'s align properly with $x$. Since the length of $x$ is at most $\min
-    \setof{\card{v_i}}{1 \leq i \leq k}$, we conclude that the $v_i$'s
-    themselves are powers of a conjugate of $x$, and that the $u_i$'s are
-    suffixes or prefixes of $x$. Now, we conclude that all the words
-    $w[\vec{n}]$ are of the correct form. Similar reasoning applies for the
-    other two cases.
+    In the sake of simplicity, we will only consider one of the three cases,
+    namely $w[\vec{n_i}] \infleq v_{\ell}^{n_{j,\ell}} u_{\ell+1}$, the other
+    two being similar. Because the lengths used in $\vec{n_i}$ are all
+    sufficiently large, we know that for every $k$, $v_k^{n_{i,k}}$ is an infix
+    of a $v_{\ell}^p$ for some non-zero $p$. Therefore, we can apply
+    \cref{periodic-gcd:lem} to conclude that there exists a word $x$ such that
+    every $v_k$ is a power of a conjugate of $x$ (a cyclic shift of $x$), and
+    $v_\ell$ is a power of $x$. We can therefore rewrite $w[\vec{n_i}]$ as $u_1
+    (\sigma_1(x))^{n_{i,1}} u_2 \cdots $, where $\sigma_k$ is some conjugacy
+    operation (cyclic shift). Now, in order for $w[\vec{n_i}]$ to be an infix
+    of $x^{p \times n_{j,\ell}} u_{\ell+1}$, we must conclude that all the
+    $u_k$'s are suffixes or prefixes of $x$, and that they align properly with
+    the $\sigma_k(x)$'s to form an infix of some power of $x$, except for the
+    last one. In particular, $w[\vec{n_i}] \in \InfPeriodChain{x} u_{\ell+1}$,
+    but also, every other choice of $\vec{n}$ will lead to a word in
+    $\InfPeriodChain{x} u_{\ell+1}$, because the alignment constraints are
+    stable under pumping.
+
+    In the case of two iterated words, the reasoning is similar, distinguishing
+    between the $v_i$'s that are occurring before and after the junction of the
+    two iterated words.
 

### 2025-10-15 src/infixes.tex
fix: trying my best

@@
     such that
-    for all \todo{proof does not actually show for all}
+    for all
     $\vec{n} \in \Nat^{k}$

### 2025-10-15 wqo-on-words.lncs.tex
new: isolating the u.u.r words as a theorem for the intro
new: moving stuff to appendix
fix: trying my best

--- 2025-10-06 wqo-on-words.lncs.tex
@@
 
-\newcommand{\repositoryUrl}{\url{https://github.com/AliaumeL/word-ordering-wqo}}
+\newcommand{\repositoryUrl}{Undefined Anonymous Repository}
 
@@
 \title{Well-quasi-orderings on word languages \\
-       main edb3a6bc24be4a12fe4f383c75384a6dff9d1a3c \\
-       2025-10-02 17:15:30 +0200
+       main 81b14ea82bdf369ba5dbe8de29fa418756900125 \\
+       2025-10-15 17:43:32 +0200
 }
@@
 \author{
-        Nathan Lhote\inst{3}\orcidID{0000-0003-3303-5368} \and
-        Aliaume Lopez\inst{2}\orcidID{0000-0002-4205-327X} \and
-        Lia Schütze\inst{1}\orcidID{0000-0003-4002-5491}}
-\authorrunning{N. Lhote and A. Lopez and L. Schütze}
-\institute{
-Max Planck Institute for Software Systems
- \and
-INP Bordeaux, LaBRI, CNRS
- \and
-Aix-Marseille University
+    Anonymized for review
 }
+\authorrunning{Anonymized for review}
+\institute{Anonymized for review}
 
@@
 % If there are any appendices, we include them here.
+\appendix
+\input{src/appendix.tex}
 

### 2025-10-15 src/appendix.tex
fix: appendix + errors

@@
 
+\begin{proofof}{small-ordinal-invariants:thm}
+    It is always true that the \kl{ordinal height} of a language over a finite
+    alphabet is at most $\omegaOrd$. Let us now consider a
+    \kl{well-quasi-ordered} language $L$ that is \kl{downwards closed} for the
+    \kl{infix relation}. Applying 
+    \cref{dw-closed-wqo-charac:thm}, we can write $L = \bigcup_{i = 1}^n L_i$ where
+    each $L_i$ is the set of finite \kl{infixes} of a finite, infinite or
+    bi-infinite \kl{ultimately uniformly recurrent} word $w_i$.
+   We can then
+    directly conclude that $\oWidth{L_i}$ less than $\omegaOrd$ (in the case of
+    a finite word), less than $2 \cdot \omegaOrd$ (in the case of an infinite
+    word thanks to \cref{small-ordinal-invariants:lem}), or less than $3 \cdot
+    \omegaOrd$ (in the case of a bi-infinite word, thanks to
+    \cref{from-bi-to-single:lem}). In any case,
+    we have the bound $\oWidth{L_i} < 3 \cdot \omegaOrd$.
+
+    Now, $\oWidth{L} \leq \sum_{i = 1}^n \oWidth{L_i} < 3 \cdot \omegaOrd <
+    \omegaOrd^2$. Finally, the inequality $\oType{L} \leq \oWidth{L} \oComProd
+    \oHeight{L} < \omega \oComProd \omega^2 = \omegaOrd^3$ allows us to
+    conclude.
+
+    The tightness of the bounds is a direct consequence of
+    \cref{from-bi-to-single:lem}, and by considering a finite union of 
+    these examples over disjoint alphabets (or even, by considering a binary 
+    alphabet and using unambiguous codes to separate the different components).
+\end{proofof}
+
 \section{Proofs for Section~\ref{infixes-amalgamation:sec}}

### 2025-10-15 src/infixes.tex
fix: appendix + errors

@@
     \label{ordinal-invariants-bounded:cor}
+    \proofref{ordinal-invariants-bounded:cor}
     Let $L$ be a \kl{bounded language} of $\Sigma^*$
@@
     \kl{chains} for the \kl{infix relation}, and in particular, $\LMorse$ is
-    \kl(language){bounded}, which is absurd by \cref{thue-morse:lemma}.
+    \kl(language){bounded},\todo{This is false! there are chains that are unbounded languages?} 
+    which is absurd by \cref{thue-morse:lemma}.
     Finally, because the \kl{ordinal width} is at most the \kl{maximal order
@@
   \label{small-ordinal-invariants:thm}
+  \proofref{small-ordinal-invariants:thm}
     Then, the \kl{maximal order type} of $L$ is strictly less than $\omegaOrd^3$,
@@
 Finally, we comment on the \kl{ordinal invariants} of the set of finite
-\kl{infixes} of an \kl{ultimately uniformly recurrent} infinite word, in order
-to derive the bounds of \cref{small-ordinal-invariants:thm}.
+\kl{infixes} of an \kl{ultimately uniformly recurrent} infinite word, from
+which the bounds of \cref{small-ordinal-invariants:thm} naturally follow (the
+proof is in \cref{small-ordinal-invariants:thm:proof} page
+\pageref{small-ordinal-invariants:thm:proof}).
 
@@
 \end{lemma}
-
-Ordinal invariants in hand, we proceed with the proof of \cref{small-ordinal-invariants:thm}.
-\begin{proofof}{small-ordinal-invariants:thm}
-    It is always true that the \kl{ordinal height} of a language over a finite
-    alphabet is at most $\omegaOrd$. Let us now consider a
-    \kl{well-quasi-ordered} language $L$ that is \kl{downwards closed} for the
-    \kl{infix relation}. Applying 
-    \cref{dw-closed-wqo-charac:thm}, we can write $L = \bigcup_{i = 1}^n L_i$ where
-    each $L_i$ is the set of finite \kl{infixes} of a finite, infinite or
-    bi-infinite \kl{ultimately uniformly recurrent} word $w_i$.
-   We can then
-    directly conclude that $\oWidth{L_i}$ less than $\omegaOrd$ (in the case of
-    a finite word), less than $2 \cdot \omegaOrd$ (in the case of an infinite
-    word thanks to \cref{small-ordinal-invariants:lem}), or less than $3 \cdot
-    \omegaOrd$ (in the case of a bi-infinite word, thanks to
-    \cref{from-bi-to-single:lem}). In any case,
-    we have the bound $\oWidth{L_i} < 3 \cdot \omegaOrd$.
-
-    Now, $\oWidth{L} \leq \sum_{i = 1}^n \oWidth{L_i} < 3 \cdot \omegaOrd <
-    \omegaOrd^2$. Finally, the inequality $\oType{L} \leq \oWidth{L} \oComProd
-    \oHeight{L} < \omega \oComProd \omega^2 = \omegaOrd^3$ allows us to
-    conclude.
-
-    The tightness of the bounds is a direct consequence of
-    \cref{from-bi-to-single:lem}, and by considering a finite union of 
-    these examples over disjoint alphabets (or even, by considering a binary 
-    alphabet and using unambiguous codes to separate the different components).
-\end{proofof}
 

### 2025-10-16 src/infixes.tex
fix: incorrect proof infixes thue morse width

@@
     decreasing size.
-    Let us now prove that $\oWidth{\LMorse} = \omegaOrd$. Assume by contradiction that
-    $\oWidth{\LMorse}$ is finite. Then, $\LMorse$ can be written as a finite union of
-    \kl{chains} for the \kl{infix relation}, and in particular, $\LMorse$ is
-    \kl(language){bounded},\todo{This is false! there are chains that are unbounded languages?} 
-    which is absurd by \cref{thue-morse:lemma}.
+    Let us now prove that $\oWidth{\LMorse} = \omegaOrd$. 
+    To that end, we can leverage the fact that 
+    the number of infixes of size $n$ in $\LMorse$ is bounded below by
+    a non-constant affine function in $n$
+    \cite{TRSHA95}, and that two words of length $n$ are comparable
+    for the \kl{infix relation}
+    if and only if they are equal. Hence, there cannot be a bound
+    on the size of an \kl{antichain} in $\LMorse$, and we conclude
+    that $\oWidth{\LMorse} = \omegaOrd$.
     Finally, because the \kl{ordinal width} is at most the \kl{maximal order

### 2025-10-16 src/appendix.tex
fix: introductions of knowledges + appendix

@@
 \clearpage
-
 \section{Proofs for Section~\ref{introduction:sec}}
@@
 
+\clearpage
 \section{Proofs for Section~\ref{prefixes:sec}}
@@
 
+\clearpage
 \section{Proofs for Section~\ref{infixes-bounded:sec}}
@@
 
+\clearpage
 \section{Proofs for Section~\ref{infixes-dwclosed:sec}}
@@
 
-    To obtain the upper bound of $3 \cdot \omegaOrd$, we can consider the same
+    To obtain the upper bound of $\omegaOrd\cdot 3$, we can consider the same
     argument as for \cref{small-ordinal-invariants:lem}. We let $N_0$ be such
@@
     \kl{ordinal width} of $\infset{w}$ is at most $\omegaOrd + \omegaOrd +
-    N_0^2$. We conclude that $\oWidth{\infset{w}} < 3 \cdot \omegaOrd$.
+    N_0^2$. We conclude that $\oWidth{\infset{w}} < \omegaOrd \cdot 3$.
 
@@
   sequence} on a binary alphabet $\set{e,f}$. The \kl{ordinal width} of the set
-  of \kl{infixes} of $w$ is then at least $2 \cdot \omegaOrd + K$, where $K$ is the
+  of \kl{infixes} of $w$ is then at least $\omegaOrd \cdot 2 + K$, where $K$ is the
   size of the chosen antichain, following the same argument as in the proof of
@@
     directly conclude that $\oWidth{L_i}$ less than $\omegaOrd$ (in the case of
-    a finite word), less than $2 \cdot \omegaOrd$ (in the case of an infinite
+    a finite word), less than $\omegaOrd \cdot 2$ (in the case of an infinite
     word thanks to \cref{small-ordinal-invariants:lem}), or less than $3 \cdot
@@
     \cref{from-bi-to-single:lem}). In any case,
-    we have the bound $\oWidth{L_i} < 3 \cdot \omegaOrd$.
+    we have the bound $\oWidth{L_i} < \omegaOrd \cdot 3$.
 
-    Now, $\oWidth{L} \leq \sum_{i = 1}^n \oWidth{L_i} < 3 \cdot \omegaOrd <
+    Now, $\oWidth{L} \leq \sum_{i = 1}^n \oWidth{L_i} < \omegaOrd \cdot 3 <
     \omegaOrd^2$. Finally, the inequality $\oType{L} \leq \oWidth{L} \oComProd
@@
 
+\clearpage
 \section{Proofs for Section~\ref{infixes-amalgamation:sec}}

### 2025-10-16 src/infixes.tex
fix: introductions of knowledges + appendix

@@
 
-\AP As a consequence of \cref{infix-embedding:thm}, we cannot replay
-proofs of \cref{prefixes:sec}, and will
-actually need to leverage some regularity of the languages to obtain a
-characterization of \kl{well-quasi-ordered} languages under the \kl{infix
-relation}. This regularity will be imposed through the notion of \intro{bounded
-languages}, i.e., languages $L \subseteq \Sigma^*$ such that there exists words
-$w_1, \dots, w_n$ satisfying $L \subseteq w_1^* \cdots w_n^*$.
+\AP As a consequence of \cref{infix-embedding:thm}, we cannot replay proofs of
+\cref{prefixes:sec}, and will actually need to leverage some regularity of the
+languages to obtain a characterization of \kl{well-quasi-ordered} languages
+under the \kl{infix relation}. This regularity will be imposed through the
+notion of \intro{bounded languages}, i.e., languages $L \subseteq \Sigma^*$
+such that there exists words $w_1, \dots, w_n$ satisfying $L \subseteq w_1^*
+\cdots w_n^*$. Let us now state the main theorem of this section.
 
@@
 shorter notation $\intro*\InfPeriodChain{x}$ for the set of \kl{infixes} of
-words of the form $x^p$, where $p$ ranges over $\Nat$.
+words of the form $x^p$, where $p \in \Nat$.
 
@@
     there exists $x,y \in \Sigma^+$ of size 
-    at most $\min \setof{\card{v_i}}{1 \leq i \leq k}$
+    at most $\max \setof{\card{v_i}}{1 \leq i \leq k}$
     such that
@@
   and have \kl{ordinal width}, \kl{ordinal height} and
-  \kl{maximal order type} respectively equal to $k \cdot \omega$, $\omega$ and $k \cdot \omega^2$.
+  \kl{maximal order type} respectively equal to $\omega \cdot k$, $\omega$ and
+  $\omega^2 \cdot k$.
 \end{proof}
@@
     Then, the set of finite infixes of $w$ has \kl{ordinal width}
-    less than $2 \cdot \omegaOrd$.
+    less than $\omegaOrd \cdot 2$.
     Furthermore, this bound is tight. 
@@
     words. In this case, the \kl{ordinal width}
-    of the set of finite infixes of $w$ is less than $3 \cdot \omegaOrd$,
+    of the set of finite infixes of $w$ is less than $\omegaOrd \cdot 3$,
     and this bound is tight.

### 2025-10-16 src/introduction.tex
fix: introductions of knowledges + appendix

@@
       arbitrary & \cref{infix-embedding:thm}: countable well-quasi orders
-                with finite initial segments & $< \omegaOne$  & $< \omegaOne$ \\
+      with \kl{finite initial segments} & $< \omegaOne$  & $< \omegaOne$ \\
       \addlinespace
-      \kl(language){bounded} & \cref{bounded-language:thm}: finite union of products of prefix and suffix chains& $< \omegaOrd^2$ & $< \omegaOrd^3$ \\
+      \kl(language){bounded} & \cref{bounded-language:thm}: 
+      finite union of products of \kl{chains} for the
+      \kl{prefix} and \kl{suffix} relations
+                             & $< \omegaOrd^2$ & $< \omegaOrd^3$ \\
       \addlinespace

### 2025-10-16 src/prefixes.tex
new: moving stuff to appendix (#10)
Co-authored-by: Aliaume LOPEZ <lopezaliaume@gmail.com>
fix: trying my best
fix: introductions of knowledges + appendix

@@
 decidability of $L$ itself.
-For the proof of the following theorem, we will use special notations to
-describe the \intro{upwards closure} of a set $S$ for a relation $\preceq$,
-which is defined as $\upset[\preceq]{S} \defined \setof{y \in \Sigma^*}{
-\exists x \in S. x \preceq y}$. Anticipating the use of the symmetric notion of
-\intro{downwards closure}, let us introduce it as follows: $\dwset[\preceq]{S} \defined \setof{y
-\in \Sigma^*}{ \exists x \in S. y \preceq x}$. Abusing notations, we will
-write $\upset{w}$ and $\dwset{w}$ for the upwards and downwards closure of a
-single element $w$, omitting the ordering relation when it is clear from the
-context. A set $S$ is called \intro{downwards closed} if $\dwset{S} = S$.
 
@@
 
-\todo{nat: might be worth noting that factor complexity is O(1)}
-\todo{aliaume: I would put it in the section "factor complexity" in the concluding
-sections}
-
 \begin{corollary}

### 2025-10-16 src/prelims.tex
fix: introductions of knowledges + appendix

@@
 sequence} is a sequence $\seqof{x_i}$ such that $x_{i+1} < x_i$ for all $i$,
-and an \intro{antichain} is a set of pairwise incomparable elements. An
-equivalent definition of a \kl{well-quasi-ordered} set is that it contains no
+a \intro{chain} is a sequence such that $x_i \leq x_{i+1}$ for all $i$, and
+an \intro{antichain} is a set of pairwise incomparable elements.
+An equivalent definition of a \kl{well-quasi-ordered} set is that it contains no
 infinite \kl{decreasing sequences}, nor infinite \kl{antichains}. We refer to
@@
 prove that it contains no infinite \kl{antichain}.
+
+A useful operation on quasi-ordered sets is to compute the \intro{upwards
+closure} of a set $S$ for a relation $\preceq$, which is defined as
+$\upset[\preceq]{S} \defined \setof{y \in \Sigma^*}{ \exists x \in S. x \preceq
+y}$. In this paper, we will also use the symmetric notion of \intro{downwards
+closure}: $\dwset[\preceq]{S} \defined \setof{y \in \Sigma^*}{ \exists x \in S.
+y \preceq x}$. Abusing notations, we will write $\upset{w}$ and $\dwset{w}$ for
+the upwards and downwards closure of a single element $w$, omitting the
+ordering relation when it is clear from the context. A set $S$ is called
+\intro{downwards closed} if $\dwset{S} = S$.
 
@@
 $\seqof{T_i}$ then $\alpha_T = \sup \setof{\alpha_{T_i} + 1}{i \in \Nat}$. We
-say that $\alpha_T$ is the \intro{rank} of $T$.
+say that $\alpha_T$ is the \intro(ordinal){rank} of $T$.
 
@@
 well-founded trees from $X$: the tree of \kl{bad sequences}, the tree of
-decreasing sequences, and the tree of \kl{antichains}. The \kl{rank} of these
-respective trees are called respectively the \intro{maximal order type} of $X$
+decreasing sequences, and the tree of \kl{antichains}. The \kl(ordinal){rank}
+of these trees are called respectively the \intro{maximal order type} of $X$
 written $\intro*{\oType{X}}$ \cite{dejongh77}, the \intro{ordinal height} of
@@
 \kl{well-quasi-ordered} set $X$. As an example, for $(\Nat, \leq)$, all bad
-sequences are descending and antichains have size at most $1$. In fact, $(\Nat,
-\leq)$ is itself an \kl{ordinal}, namely $\omegaOrd$. Hence it is its own
-\kl{maximal order type} and \kl{ordinal height}, and its \kl{ordinal width} is
-$1$ since it contains no antichain of size $2$. We refer to the survey of
+sequences are \kl[decreasing sequence]{descending} and \kl{antichains} have
+size at most $1$. In fact, $(\Nat, \leq)$ is itself an \kl{ordinal}, namely
+$\omegaOrd$. Hence it is its own \kl{maximal order type} and \kl{ordinal
+height}, and its \kl{ordinal width} is $1$. We refer to the survey of
 \cite{DZSCSC20} for a detail discussion on these concepts and their computation

### 2025-10-16 src/amalgam.tex
fix: moving stuff to appendix

@@
 
-\subsection{$\infleq$-Well-Quasi-Ordered Amalgamation Systems}
+%\subsection{$\infleq$-Well-Quasi-Ordered Amalgamation Systems}
 
@@
 \end{lemma}
-
-\begin{proofof}{infix-amalgamation:thm}
-    Assume that $L$ is \kl{well-quasi-ordered} by the \kl{infix relation},
-    and obtained by an \kl{amalgamation system}
-    $(\Sigma, R, E, \canrun)$.
-
-    Let us consider the set $M$ of minimal runs for the relation $\leq_E$,
-    which is finite because the latter is a \kl{well-quasi-ordering}. 
-    By \cref{gap-words-prefix-ordered:lem}, we know that for each minimal run $\rho \in M$,
-    each gap language $\GapLanguage{\rho}{i}$ of $\rho$ is totally ordered by $\infleq$.
-    Adapting the proof of language boundedness from \cite[Section 4.2]{ASZZ24}, we may conclude that $\GapLanguage{\rho}{i} \subseteq \InfPeriodChain{w}$ for some $w \in \GapLanguage{\rho}{i}$.
-    As $\InfPeriodChain{w}$ is language bounded and this property is stable under subsets, concatenation and finite union,
-    we can conclude that $L$ is bounded as well.
-\end{proofof}
 
@@
 
-\subsection{Effective Decision Procedures}
-\label{infixes-amalgamation-effective:subsec}
+%\subsection{Effective Decision Procedures}
+%\label{infixes-amalgamation-effective:subsec}
 
@@
 
-\begin{proofof}{infix-wqo-is-emptiness:thm}
-	We first show \cref{emptiness-decidable} $\Rightarrow$ \cref{wqo-infix-decidable}. We aim to make the inclusion test of \cref{bounded-language:eq} of \cref{bounded-language:thm} effective. 
-	Let $R(n,m,N_0) \defined \bigcup_{x,y \in \Sigma^{\leq n}} \bigcup_{u \in
-    \Sigma^{\leq m \times N_0}} \InfPeriodChain{x} u \InfPeriodChain{y} \cup
-    \InfPeriodChain{x}u \cup u\InfPeriodChain{x}$. For any concrete values of
-    the bounds $n$, $m$, and $N_0$, this language is regular. The map $L
-    \mapsto L \cap \Sigma^* \setminus R(n,m,N_0)$  is a \kl{rational
-    transduction} because $\Sigma^* \setminus R(n,m,N_0)$ is regular. Since
-    $\mathcal{C}$ is \kl{closed under rational transductions}, we can therefore
-    reduce the inclusion to emptiness of this language. However, we need to
-    find these bounds first.
-
-    To determine values for $n$ and $m$, we first test if $L$ is
-    \kl(language){bounded}. Since emptiness is decidable, we can apply the
-    algorithm in~\cite[Section 4.2]{ASZZ24} to decide if $L$ is
-    \kl(language){bounded}. If $L$ is \kl(language){bounded}, this algorithm
-    yields words $w_1, \ldots w_n$ such that $L \subseteq w_1^* \cdots w_n^*$
-    and therefore yields also the bounds in questions: $n$ is the number of
-    words, and $m$ is the maximal length of a word $w_i$ where $1 \leq i \leq
-    n$. If $L$ is not bounded, then $L$ cannot be \kl{well-quasi-ordered} by
-    the \kl{infix relation} because of \cref{infix-amalgamation:thm} and we
-    immediately return false.
-
-    To determine the value for $N_0$, we then compute the downward closure (with respect to subwords) of $L$. 
-    This is effective and yields a finite-state automaton. 
-    Recall that $N_0$ is the maximum number of repetitions of a word $w_i$ that can not be iterated arbitrarily often. 
-    This value is therefore bounded above by the length of the longest simple path in this automaton.
-
-    \cref{wqo-infix-decidable} $\Rightarrow$
-    \cref{wqo-prefix-decidable}. We just consider the transduction $f$
-    that maps every word $w$ to $\# w$ where $\# $ is a fresh symbol. Then, for
-    any language $L \in \mathcal C$, $L$ is \kl{well-quasi-ordered} by
-    \kl{prefix} if and only if $f(L)$ is \kl{well-quasi-ordered} by \kl{infix}.
-
-    \cref{wqo-prefix-decidable} $\Rightarrow$
-    \cref{emptiness-decidable}. 
-	We 
-	consider the transduction $R \defined \Sigma^* \times \set{a, b}^*$. Then 
-	for any language $L \in \mathcal C$,
-    the image of $L$ through $R$ is \kl{well-quasi-ordered}
-    by \kl{prefix} if and only if $L$ is empty.
-\end{proofof}
-
 The class $\mathcal{C}_\text{aut}$ of \kl{regular languages} and the class

### 2025-10-16 src/appendix.tex
fix: moving stuff to appendix

@@
 
-\begin{lemma}
-	\label{automatic-uur:lem}
-	\proofref{automatic-uur:lem}
-	Given an \kl{automatic sequence} $w \in \Sigma^{\Nat}$, one can decide
-	whether it is \kl{ultimately uniformly recurrent}.
-\end{lemma}
-
-\begin{proofof}{automatic-uur:lem}
-	We can rewrite this as a question on the \kl{automatic sequence} $w$
-	as follows:
-	\begin{align*}
-		&\exists N_0,                   &   \text{ultimately} \\
-		&\forall i_s \geq N_0,          &   \text{for every infix (start) } u \\
-		&\forall i_e > i_s,             &   \text{for every infix (end) }   u \\
-		&\exists k \geq 1,              &   \text{there exists a bound} \\
-		&\forall j_s \geq N_0,          &   \text{for every other infix (start) } v \\
-		&\forall j_e \geq j_s + k,      &   \text{of size at least $k$} \\
-		&\exists l \geq 0,              &   \text{there exists a position in } v \\
-		&\forall 0 \leq m < i_e - i_s,  &   \text{where } u \text{ can be found} \\
-		&j_s + m + l < j_e \land
-		w(i_s + m) = w(j_s + m + l) \quad .
-	\end{align*}
-	Because $w$ is computable by a finite automaton, one can reduce the above
-	formula to a regular language, for which it suffices to check emptiness, which
-	is decidable.
-\end{proofof}
-
-\begin{proofof}{small-ordinal-invariants:lem}
-	Let $N_0$ be a bound such that $w_{\geq N_0}$ is \kl{uniformly recurrent}.
-	Let us write $\infset{w}$ the set of finite infixes of $w$.
-	We prove that $\oWidth{\infset{w}} \leq \omegaOrd + N_0$.
-	Let $u_1 \infleq w$ be a finite word.
-
-	If $u_1$ is an \kl{infix} of $w_{\geq N_0}$, then there exists $k \geq 1$
-	such that $u_1$ is an \kl{infix} of every word of size at least $k$. In
-	particular, there is finite bound on the length of every sequence of
-	incomparable elements starting with $u_1$. We conclude in particular that
-	$\infset{w} \setminus \upset{u_1}$ has a finite \kl{ordinal width}.
-
-	Otherwise, $u_1$ can only be found \emph{before} $N_0$. In this case, we
-	consider a second element of a \kl{bad sequence} $u_2 \infleq w$, which is
-	incomparable with $u_1$ for the \kl{infix relation}. If $u_2$ is an
-	\kl{infix} of $w_{\geq N_0}$, then we can conclude as before. Otherwise,
-	notice that $u_1$ and $u_2$ cannot start at the same position in $w$
-	(because they are incomparable). Continuing this argument, we conclude that
-	there are at most $N_0$ elements starting before $N_0$
-	at the start of any sequence of
-	incomparable elements in $\infset{w}$. We conclude that
-	$\oWidth{\infset{w}} \leq \omegaOrd + N_0$.
-
-	Let us now justify that this bound is tight.
-	The \kl{Thue-Morse sequence} over a binary
-	alphabet $\set{a,b}$ has \kl{ordinal width} $\omegaOrd$
-	from \cref{thue-morse-ordinal:lemma}.
-	Given a number $N_0
-	\in \Nat$, one can construct an arbitrarily long \kl{antichain} of words for
-	the \kl{infix relation} by using a new letter $c$. When concatenating this
-	(finite) antichain as a prefix of the \kl{Thue-Morse sequence}, one obtains a
-	new (infinite) word $w$. It is clear that the \kl{ordinal width} of
-	$\infset{w}$ is now at least $\omegaOrd + N_0$.
-\end{proofof}
+\clearpage
+\section{Proofs for Section~\ref{infixes-dwclosed:sec}}
 
@@
 \end{proofof}
-
-\clearpage
-\section{Proofs for Section~\ref{infixes-dwclosed:sec}}
 
@@
   \cref{small-ordinal-invariants:lem}, using \cref{thue-morse-ordinal:lemma}.
+\end{proofof}
+
+\begin{lemma}
+	\label{automatic-uur:lem}
+	\proofref{automatic-uur:lem}
+	Given an \kl{automatic sequence} $w \in \Sigma^{\Nat}$, one can decide
+	whether it is \kl{ultimately uniformly recurrent}.
+\end{lemma}
+
+\begin{proofof}{automatic-uur:lem}
+	We can rewrite this as a question on the \kl{automatic sequence} $w$
+	as follows:
+	\begin{align*}
+		&\exists N_0,                   &   \text{ultimately} \\
+		&\forall i_s \geq N_0,          &   \text{for every infix (start) } u \\
+		&\forall i_e > i_s,             &   \text{for every infix (end) }   u \\
+		&\exists k \geq 1,              &   \text{there exists a bound} \\
+		&\forall j_s \geq N_0,          &   \text{for every other infix (start) } v \\
+		&\forall j_e \geq j_s + k,      &   \text{of size at least $k$} \\
+		&\exists l \geq 0,              &   \text{there exists a position in } v \\
+		&\forall 0 \leq m < i_e - i_s,  &   \text{where } u \text{ can be found} \\
+		&j_s + m + l < j_e \land
+		w(i_s + m) = w(j_s + m + l) \quad .
+	\end{align*}
+	Because $w$ is computable by a finite automaton, one can reduce the above
+	formula to a regular language, for which it suffices to check emptiness, which
+	is decidable.
+\end{proofof}
+
+\begin{proofof}{small-ordinal-invariants:lem}
+	Let $N_0$ be a bound such that $w_{\geq N_0}$ is \kl{uniformly recurrent}.
+	Let us write $\infset{w}$ the set of finite infixes of $w$.
+	We prove that $\oWidth{\infset{w}} \leq \omegaOrd + N_0$.
+	Let $u_1 \infleq w$ be a finite word.
+
+	If $u_1$ is an \kl{infix} of $w_{\geq N_0}$, then there exists $k \geq 1$
+	such that $u_1$ is an \kl{infix} of every word of size at least $k$. In
+	particular, there is finite bound on the length of every sequence of
+	incomparable elements starting with $u_1$. We conclude in particular that
+	$\infset{w} \setminus \upset{u_1}$ has a finite \kl{ordinal width}.
+
+	Otherwise, $u_1$ can only be found \emph{before} $N_0$. In this case, we
+	consider a second element of a \kl{bad sequence} $u_2 \infleq w$, which is
+	incomparable with $u_1$ for the \kl{infix relation}. If $u_2$ is an
+	\kl{infix} of $w_{\geq N_0}$, then we can conclude as before. Otherwise,
+	notice that $u_1$ and $u_2$ cannot start at the same position in $w$
+	(because they are incomparable). Continuing this argument, we conclude that
+	there are at most $N_0$ elements starting before $N_0$
+	at the start of any sequence of
+	incomparable elements in $\infset{w}$. We conclude that
+	$\oWidth{\infset{w}} \leq \omegaOrd + N_0$.
+
+	Let us now justify that this bound is tight.
+	The \kl{Thue-Morse sequence} over a binary
+	alphabet $\set{a,b}$ has \kl{ordinal width} $\omegaOrd$
+	from \cref{thue-morse-ordinal:lemma}.
+	Given a number $N_0
+	\in \Nat$, one can construct an arbitrarily long \kl{antichain} of words for
+	the \kl{infix relation} by using a new letter $c$. When concatenating this
+	(finite) antichain as a prefix of the \kl{Thue-Morse sequence}, one obtains a
+	new (infinite) word $w$. It is clear that the \kl{ordinal width} of
+	$\infset{w}$ is now at least $\omegaOrd + N_0$.
 \end{proofof}
@@
 \end{proofof}
+
+\begin{proofof}{infix-amalgamation:thm}
+    Assume that $L$ is \kl{well-quasi-ordered} by the \kl{infix relation},
+    and obtained by an \kl{amalgamation system}
+    $(\Sigma, R, E, \canrun)$.
+
+    Let us consider the set $M$ of minimal runs for the relation $\leq_E$,
+    which is finite because the latter is a \kl{well-quasi-ordering}. 
+    By \cref{gap-words-prefix-ordered:lem}, we know that for each minimal run $\rho \in M$,
+    each gap language $\GapLanguage{\rho}{i}$ of $\rho$ is totally ordered by $\infleq$.
+    Adapting the proof of language boundedness from \cite[Section 4.2]{ASZZ24}, we may conclude that $\GapLanguage{\rho}{i} \subseteq \InfPeriodChain{w}$ for some $w \in \GapLanguage{\rho}{i}$.
+    As $\InfPeriodChain{w}$ is language bounded and this property is stable under subsets, concatenation and finite union,
+    we can conclude that $L$ is bounded as well.
+\end{proofof}
+
+\begin{proofof}{infix-wqo-is-emptiness:thm}
+	We first show \cref{emptiness-decidable} $\Rightarrow$ \cref{wqo-infix-decidable}. We aim to make the inclusion test of \cref{bounded-language:eq} of \cref{bounded-language:thm} effective. 
+	Let $R(n,m,N_0) \defined \bigcup_{x,y \in \Sigma^{\leq n}} \bigcup_{u \in
+    \Sigma^{\leq m \times N_0}} \InfPeriodChain{x} u \InfPeriodChain{y} \cup
+    \InfPeriodChain{x}u \cup u\InfPeriodChain{x}$. For any concrete values of
+    the bounds $n$, $m$, and $N_0$, this language is regular. The map $L
+    \mapsto L \cap \Sigma^* \setminus R(n,m,N_0)$  is a \kl{rational
+    transduction} because $\Sigma^* \setminus R(n,m,N_0)$ is regular. Since
+    $\mathcal{C}$ is \kl{closed under rational transductions}, we can therefore
+    reduce the inclusion to emptiness of this language. However, we need to
+    find these bounds first.
+
+    To determine values for $n$ and $m$, we first test if $L$ is
+    \kl(language){bounded}. Since emptiness is decidable, we can apply the
+    algorithm in~\cite[Section 4.2]{ASZZ24} to decide if $L$ is
+    \kl(language){bounded}. If $L$ is \kl(language){bounded}, this algorithm
+    yields words $w_1, \ldots w_n$ such that $L \subseteq w_1^* \cdots w_n^*$
+    and therefore yields also the bounds in questions: $n$ is the number of
+    words, and $m$ is the maximal length of a word $w_i$ where $1 \leq i \leq
+    n$. If $L$ is not bounded, then $L$ cannot be \kl{well-quasi-ordered} by
+    the \kl{infix relation} because of \cref{infix-amalgamation:thm} and we
+    immediately return false.
+
+    To determine the value for $N_0$, we then compute the downward closure (with respect to subwords) of $L$. 
+    This is effective and yields a finite-state automaton. 
+    Recall that $N_0$ is the maximum number of repetitions of a word $w_i$ that can not be iterated arbitrarily often. 
+    This value is therefore bounded above by the length of the longest simple path in this automaton.
+
+    \cref{wqo-infix-decidable} $\Rightarrow$
+    \cref{wqo-prefix-decidable}. We just consider the transduction $f$
+    that maps every word $w$ to $\# w$ where $\# $ is a fresh symbol. Then, for
+    any language $L \in \mathcal C$, $L$ is \kl{well-quasi-ordered} by
+    \kl{prefix} if and only if $f(L)$ is \kl{well-quasi-ordered} by \kl{infix}.
+
+    \cref{wqo-prefix-decidable} $\Rightarrow$
+    \cref{emptiness-decidable}. 
+	We 
+	consider the transduction $R \defined \Sigma^* \times \set{a, b}^*$. Then 
+	for any language $L \in \mathcal C$,
+    the image of $L$ through $R$ is \kl{well-quasi-ordered}
+    by \kl{prefix} if and only if $L$ is empty.
+\end{proofof}

### 2025-10-16 src/infixes.tex
fix: moving stuff to appendix

@@
 (trivially) translated into a finite automaton that reads the base $b$
-representation of the free variables (that are indices of the sequence). It is
-folklore that one can decide if an \kl{automatic sequence} is \kl{ultimately
-uniformly recurrent}. A proof can be found in the appendix at
-\cref{automatic-uur:lem}. Based on this, we now prove:
+representation of the free variables (that are indices of the sequence). In
+particular, it follows than one can decide one can decide if an \kl{automatic
+sequence} is \kl{ultimately uniformly recurrent}, a proof of this folklore
+result can be found in the appendix at \cref{automatic-uur:lem}. Based on this,
+we now prove:
 

### 2025-10-16 src/introduction.tex
fix: moving stuff to appendix

@@
 well-quasi-ordered sets is also stable under the operation of taking finite
-words and finite trees labelled by elements of a well-quasi-ordered set
-\cite{HIG52,KRU72}.
+words and finite trees labelled by elements of a well-quasi-ordered set~\cite{HIG52,KRU72}.
 
@@
     \centering
-    \begin{tabular}{c|c|c|c}
+    \setlength{\tabcolsep}{6pt}
+    \begin{tabular}{c|p{6cm}|c|c}
         \toprule
@@
         \midrule
-        arbitrary & Finite unions of \kl{chains} \cref{prefixes:thm} & $< \omegaOrd$ & $< \omegaOrd^2$ \\
-        regular   & Finite unions of \kl(language){regular} \kl{chains} \cref{prefix-wqo-reg-decidable:cor} & $< \omegaOrd$ & $< \omegaOrd^2$ \\
+        arbitrary & \cref{prefixes:thm}: finite unions of \kl{chains}
+                  & $< \omegaOrd$ & $< \omegaOrd^2$ \\
+        \addlinespace
+        regular   & \cref{prefix-wqo-reg-decidable:cor}: finite unions of \kl(language){regular} \kl{chains} 
+                  & $< \omegaOrd$ & $< \omegaOrd^2$ \\
         \bottomrule

### 2025-10-16 src/amalgam.tex
fix: reworking conclusion

@@
+% LTeX: language=en-GB 
+% !TeX root=../wqo-on-words.lncs.tex
 \section{Infixes and Amalgamation Systems}
@@
 	\label{infix-wqo-is-emptiness:thm}
+  \proofref{infix-wqo-is-emptiness:thm}
 	Let $\mathcal{C}$ be an \kl{effective amalgamative class} of languages.
@@
 \end{theorem}
-
-%\AP We say that a \kl{strongly effective amalgamative class} is an
-%\kl{effective amalgamative class} for which the emptiness problem is decidable.
-%Let us immediately remark that the class of \kl{regular languages} is
-%\kl(amalg){strongly effective}, and so is the class of context-free languages.
-%Therefore, \cref{infix-wqo-is-emptiness:thm} provides a concrete decision
-%procedure for these classes.
 
@@
 
-If we additionally assume that such a language is closed under infixes, we
-obtain an even stronger structure: All such languages are regular!
+If we additionally assume that such a language is closed under taking infixes,
+we obtain an even stronger structure: All such languages are regular!
 
@@
 \AP To construct a decision procedure for well-quasi-orderedness under
-$\infleq$, we our \kl{amalgamation systems} to satisfy certain
+$\infleq$, we need our \kl{amalgamation systems} to satisfy certain
 \intro(amalgamation){effectiveness assumptions}. We require that for an
@@
 
-%\AP We say that a class $\mathcal{C}$ of languages is an \intro{effective
-%amalgamative class} whenever for every $L \in \mathcal{C}$, there exists an
-%effective \kl{amalgamation system} recognizing $L$, and such that $\mathcal{C}$
-%is \kl{effectively closed under rational transductions}. Recall that a
-%\intro{rational transduction} is a \emph{relation} $R \subseteq \Sigma^* \times
-%\Gamma^*$ that can be defined by a (nondeterministic) finite automaton with
-%output \cite[Chapter 5, page 64]{BERST79}. A class of languages $\mathcal{C}$
-%is \intro{effectively closed under rational transductions} when, for every
-%language $L \in \mathcal{C}$, and every rational transduction $R \subseteq
-%\Sigma^* \times \Gamma^*$, the image of $L$ through $R$ --- that is, $\setof{v
-%\in \Gamma^*}{\exists u \in L. (u,v) \in R}$ --- is in $\mathcal{C}$ and
-%effectively computable. In particular, it implies that the class contains
-%languages over arbitrary (finite) alphabets.
-
-%\AP A class $\mathcal{C}$ of languages is \intro(amalg){strongly effective}
-%whenever the emptiness problem for languages in $\mathcal{C}$ is decidable.
-%This notion is interesting because usual language problems such as boundedness
-%or simultaneous boundedness are decidable for \kl{strongly effective
-%amalgamative classes}~\cite{ASZZ24}.
-
-The class $\mathcal{C}_\text{aut}$ of \kl{regular languages} and the class
+Under these assumptions, one can transform the inclusion test of
+\cref{bounded-language:eq} of \cref{bounded-language:thm} into an effective
+procedure, using pumping arguments from \cite[Section 4.2]{ASZZ24}, which, in
+turn, allows us to prove \cref{infix-wqo-is-emptiness:thm}. Since the class
+$\mathcal{C}_\text{aut}$ of \kl{regular languages} and the class
 $\mathcal{C}_{\text{cfg}}$ of context-free languages are examples of
-\kl{effective amalgamative classes}, hence the following corollary.
+\kl{effective amalgamative classes}, the following corollary is immediate.
 

### 2025-10-16 src/conclusion.tex
fix: trying my best
fix: reworking conclusion

--- 2025-08-12 src/conclusion.tex
@@
 
-We provided concretes statements that justify why the \kl{subword relation}
-is used when defining \kl{well-quasi-orders} on finite words. Even if
-\kl{prefix}, \kl{suffix} or \kl{infix} relations are meaningful, they are
-\kl{well-quasi-ordered} if and only if they behave similarly to disjoint copies
-of $\Nat$ or $\Nat^2$. However, our approach suffers some limitations 
-and opens the road to natural continuation of this line of work.
+We have described the landscapes of \kl{well-quasi-ordered} languages for the
+natural orderings on finite words: \kl{prefix}, \kl{suffix}, and \kl{infix}
+relations. While the prefix and suffix relation exhibit very simple behaviours,
+the infix relation can encode many complex quasi-orders (and even simulate the
+\kl{subword ordering}). In the case of languages that are described by simple
+computational models, or languages that are ``structurally simple''
+(\kl{bounded languages}, \kl{downwards closed} languages), we showed that only
+very simple \kl{well-quasi-orders} can be obtained: they are essentially
+isomorphic to disjoint unions of copies of finite sets, $(\Nat, \leq)$, and
+$(\Nat^2, \leq)$. Finally, under effectiveness assumptions on the language
+(such as being recognized by an \kl{amalgamation system}, or being the set of
+infixes of an \kl{automatic sequence}), we proved the decidability of being
+\kl{well-quasi-ordered} for the \kl{infix relation}. We believe that these very
+encouraging results pave the way for further research on deciding which sets
+are \kl{well-quasi-ordered} for other orderings. Let us now discuss
+some possible research directions and remarks.
 
@@
 \cref{bounded-language:thm}, \cref{small-ordinal-invariants:thm}, and
-\cref{prefixes:thm} could be conjecture to hold in the case of infinite
+\cref{prefixes:thm} could be conjectured to hold in the case of infinite
 alphabets (themselves equipped with a \kl{well-quasi-ordering}). This would
@@
 \kl{well-quasi-ordered} with respect to the \emph{induced subgraph relation},
-as shown in \cite{DRT10}, and recently revisited by one of the authors in
-\cite{L24:arxiv:v2}.
-
-\paragraph*{Complexity} We have chosen to disregard complexity considerations
-when proving decidability results, as we do not believe that a fined grained
-complexity analysis would be particularly enlightening. However, we conjecture
-that for regular languages, the complexity should be polynomial in the size of
-the minimal automaton recognizing the language.
+as shown in \cite{DRT10}, and recently revisited in \cite{lopez25}.
 
@@
 
-\paragraph*{Graph Classes}
-\todo[inline]{aliaume: retalk about \cite{ALM17} and \cite{L24:arxiv:v2}}
+\paragraph*{Factor Complexity} \AP Let us conclude this section with a few
+remarks on the notion of \kl{factor complexity} of languages. Recall that the
+\intro{factor complexity} of a language $L \subseteq \Sigma^*$ is the function
+$f_L: \Nat \to \Nat$ such that $f_L(n)$ is the number of distinct words of size
+$n$ in $L$. We extend the notion of \kl{factor complexity} to finite, infinite,
+and bi-infinite words as the \kl{factor complexity} of their set of finite
+\kl{infixes}. For the \kl{prefix relation} and the \kl{suffix relation}, all
+\kl{well-quasi-ordered} languages have a bounded \kl{factor complexity}, since
+they are finite unions of \kl{chains}.
+
+While there clearly are languages with low \kl{factor complexity} that are not
+\kl{well-quasi-ordered} for the \kl{infix relation}, such as the language $L
+\defined \dwset{ a b^* a }$; one would expect that languages that are
+\kl{well-quasi-ordered} for the \kl{infix relation} would have a low \kl{factor
+complexity}.
+
+In some sense, our results confirm this intuition in the case of languages
+described by a simple computational model. For languages recognized by
+\kl{amalgamation systems}, being \kl{well-quasi-ordered} implies being a
+\kl{bounded language}, and therefore being  included in some finite union of
+languages of the form $w_1^* w_2 w_3^*$. Hence, these languages have at most a
+quadratic \kl{factor complexity}. This is also the case for languages described
+as the infixes of a finite set of pairs of \kl{morphic sequences}. Indeed, the
+factor complexity of a \kl{morphic sequence} that is \kl{uniformly recurrent}
+is linear \cite[Theorem 24]{NIPR09}, therefore the \kl{factor complexity} of a
+language given by \kl{sequence representation} using \kl{morphic sequences} is
+at most quadratic.
+
+However, there are \kl{downwards closed} languages that are
+\kl{well-quasi-ordered} for the \kl{infix relation} but have an \kl{exponential
+factor complexity}: the $(5, 3)$-Toeplitz word is \kl{uniformly recurrent}
+\cite[p. 499]{CAKA97}, but has an exponential factor complexity \cite[Theorem
+5]{CAKA97}. This shows that our computational models somehow fail to capture
+vast classes of \kl{well-quasi-ordered} languages with a high \kl{factor
+complexity}. It would be interesting to understand which new proof techniques
+would be required to obtain decidability for these languages.

### 2025-10-16 src/introduction.tex
fix: reworking conclusion

@@
 that are \kl{well-quasi-ordered} by the \kl{induced subgraph relation}
-using similar language theoretic techniques~\cite{DRT10,L24:arxiv:v2,ALM17}.
+using similar language theoretic techniques~\cite{DRT10,lopez25,ALM17}.
 Furthermore, a previous work by Kuske shows that any

### 2025-10-16 src/conclusion.tex
fix: knowledges + refs + storytelling

@@
 However, there are \kl{downwards closed} languages that are
-\kl{well-quasi-ordered} for the \kl{infix relation} but have an \kl{exponential
-factor complexity}: the $(5, 3)$-Toeplitz word is \kl{uniformly recurrent}
-\cite[p. 499]{CAKA97}, but has an exponential factor complexity \cite[Theorem
+\kl{well-quasi-ordered} for the \kl{infix relation} but have an exponential
+\kl{factor complexity}: the $(5, 3)$-Toeplitz word is \kl{uniformly recurrent}
+\cite[p. 499]{CAKA97}, and has exponential factor complexity \cite[Theorem
 5]{CAKA97}. This shows that our computational models somehow fail to capture

### 2025-10-16 src/introduction.tex
fix: knowledges + refs + storytelling

@@
 In the graph theory community, recent works have studied classes of graphs 
-that are \kl{well-quasi-ordered} by the \kl{induced subgraph relation}
+that are \kl{well-quasi-ordered} by the induced subgraph relation
 using similar language theoretic techniques~\cite{DRT10,lopez25,ALM17}.

### 2025-10-16 src/word-embeddings-figure.py
Fix typos: eachother->each other, lastest->latest, inforamtion->information
Co-authored-by: AliaumeL <24816735+AliaumeL@users.noreply.github.com>

@@ -127,7 +127,7 @@ def generate_figure(size = 4, relation = subword):
     # We create a "triangle-like" 
     # figure with the empty word ε at the bottom
     # and each layer containing all the words of size k
-    # on top of eachother, and centered based on their size
+    # on top of each other, and centered based on their size
 
     layers = []
     for k in range(size):

### 2025-10-16 wqo-on-words.lncs.tex
fix: moving stuff to appendix
Add thmtools integration and demonstrate restatable theorem proof of concept
Co-authored-by: AliaumeL <24816735+AliaumeL@users.noreply.github.com>

@@
 %   edit the `templates/lncs/lncs.tex` file instead.
-\documentclass[runningheads]{llncs}
+\documentclass[runningheads,envcountsame]{llncs}
 
@@
 % microtype for better typography
-\usepackage{microtype}
+%\usepackage{microtype} % Disabled due to font expansion issues
 
@@
 \usepackage{amssymb,amsmath,stmaryrd,thmtools,thm-restate,upgreek}
+
+% Redefine theorem environments to be compatible with thmtools restatable
+% llncs.cls already defines these using \spnewtheorem, but we need them to work with thmtools
+% We need to undefine both the environments and the counters
+\makeatletter
+\let\theorem\relax
+\let\endtheorem\relax
+\let\c@theorem\relax
+\let\thetheorem\relax
+\let\p@theorem\relax
+\let\lemma\relax
+\let\endlemma\relax
+\let\c@lemma\relax
+\let\thelemma\relax
+\let\p@lemma\relax
+\let\corollary\relax
+\let\endcorollary\relax
+\let\c@corollary\relax
+\let\thecorollary\relax
+\let\p@corollary\relax
+\let\proposition\relax
+\let\endproposition\relax
+\let\c@proposition\relax
+\let\theproposition\relax
+\let\p@proposition\relax
+\let\definition\relax
+\let\enddefinition\relax
+\let\c@definition\relax
+\let\thedefinition\relax
+\let\p@definition\relax
+\let\example\relax
+\let\endexample\relax
+\let\c@example\relax
+\let\theexample\relax
+\let\p@example\relax
+\let\remark\relax
+\let\endremark\relax
+\let\c@remark\relax
+\let\theremark\relax
+\let\p@remark\relax
+\makeatother
+
+% Now redeclare them using thmtools for restatable support
+% Using llncs-like formatting: bold for theorem names, italic for body
+\declaretheorem[style=plain,numberwithin=section,name=Theorem]{theorem}
+\declaretheorem[style=plain,sibling=theorem,name=Lemma]{lemma}
+\declaretheorem[style=plain,sibling=theorem,name=Corollary]{corollary}
+\declaretheorem[style=plain,sibling=theorem,name=Proposition]{proposition}
+\declaretheorem[style=definition,sibling=theorem,name=Definition]{definition}
+\declaretheorem[style=definition,sibling=theorem,name=Example]{example}
+\declaretheorem[style=remark,sibling=theorem,name=Remark]{remark}
 

### 2025-10-16 src/conclusion.tex
fix: typos!

@@
 languages of the form $w_1^* w_2 w_3^*$. Hence, these languages have at most a
-quadratic \kl{factor complexity}. This is also the case for languages described
+linear \kl{factor complexity}. This is also the case for languages described
 as the infixes of a finite set of pairs of \kl{morphic sequences}. Indeed, the
@@
 language given by \kl{sequence representation} using \kl{morphic sequences} is
-at most quadratic.
+at most linear.
 

### 2025-10-16 src/appendix.tex
fix: knowledges + refs + storytelling
Fix proof ordering in appendix to follow logical dependencies
Co-authored-by: AliaumeL <24816735+AliaumeL@users.noreply.github.com>
final version

@@
     Hence, that $L$ is finite union of \kl{chains}.
-\end{proofof}
-
-\begin{proofof}{prefix-wqo-reg-decidable:cor}
-    If $L$ is regular, then it is $\MSO$-definable, and there 
-    exists a formula $\varphi(x)$ in $\MSO$ that selects nodes 
-    of the \kl{tree of prefixes} that belong to $L$. Now, to decide whether there
-    exists an \kl{antichain branch} for $L$, we can simply check whether
-    the following formula is satisfied:
-    \begin{equation*}
-        \exists B. 
-        B \text{ is a branch } \land
-        \forall x \in B, \exists y. y \text{ is a child of } x \land \varphi(y) \land y \not\in B
-        \quad .
-    \end{equation*}
-    Because the above formula is an $\MSO$-formula over the infinite
-    $\Sigma$-branching tree, whether it is satisfied is decidable
-    as an easy consequence of the decidability of $\MSO$ over infinite binary
-    trees
-    \cite[Theorem 1.1]{RAB69}.
 \end{proofof}
@@
     is similar, but will result in a finite word.
-    \todo{aliaume: do we need wqo here? the proof
-      should go through without it: the sequence 
-      $u_i$ is already increasing for infix}
 
@@
 
+\begin{proofof}{small-ordinal-invariants:lem}
+	Let $N_0$ be a bound such that $w_{\geq N_0}$ is \kl{uniformly recurrent}.
+	Let us write $\infset{w}$ the set of finite infixes of $w$.
+	We prove that $\oWidth{\infset{w}} \leq \omegaOrd + N_0$.
+	Let $u_1 \infleq w$ be a finite word.
+
+	If $u_1$ is an \kl{infix} of $w_{\geq N_0}$, then there exists $k \geq 1$
+	such that $u_1$ is an \kl{infix} of every word of size at least $k$. In
+	particular, there is finite bound on the length of every sequence of
+	incomparable elements starting with $u_1$. We conclude in particular that
+	$\infset{w} \setminus \upset{u_1}$ has a finite \kl{ordinal width}.
+
+	Otherwise, $u_1$ can only be found \emph{before} $N_0$. In this case, we
+	consider a second element of a \kl{bad sequence} $u_2 \infleq w$, which is
+	incomparable with $u_1$ for the \kl{infix relation}. If $u_2$ is an
+	\kl{infix} of $w_{\geq N_0}$, then we can conclude as before. Otherwise,
+	notice that $u_1$ and $u_2$ cannot start at the same position in $w$
+	(because they are incomparable). Continuing this argument, we conclude that
+	there are at most $N_0$ elements starting before $N_0$
+	at the start of any sequence of
+	incomparable elements in $\infset{w}$. We conclude that
+	$\oWidth{\infset{w}} \leq \omegaOrd + N_0$.
+
+	Let us now justify that this bound is tight.
+	The \kl{Thue-Morse sequence} over a binary
+	alphabet $\set{a,b}$ has \kl{ordinal width} $\omegaOrd$
+	from \cref{thue-morse-ordinal:lemma}.
+	Given a number $N_0
+	\in \Nat$, one can construct an arbitrarily long \kl{antichain} of words for
+	the \kl{infix relation} by using a new letter $c$. When concatenating this
+	(finite) antichain as a prefix of the \kl{Thue-Morse sequence}, one obtains a
+	new (infinite) word $w$. It is clear that the \kl{ordinal width} of
+	$\infset{w}$ is now at least $\omegaOrd + N_0$.
+\end{proofof}
+
 \begin{proofof}{from-bi-to-single:lem}
@@
 	is decidable.
-\end{proofof}
-
-\begin{proofof}{small-ordinal-invariants:lem}
-	Let $N_0$ be a bound such that $w_{\geq N_0}$ is \kl{uniformly recurrent}.
-	Let us write $\infset{w}$ the set of finite infixes of $w$.
-	We prove that $\oWidth{\infset{w}} \leq \omegaOrd + N_0$.
-	Let $u_1 \infleq w$ be a finite word.
-
-	If $u_1$ is an \kl{infix} of $w_{\geq N_0}$, then there exists $k \geq 1$
-	such that $u_1$ is an \kl{infix} of every word of size at least $k$. In
-	particular, there is finite bound on the length of every sequence of
-	incomparable elements starting with $u_1$. We conclude in particular that
-	$\infset{w} \setminus \upset{u_1}$ has a finite \kl{ordinal width}.
-
-	Otherwise, $u_1$ can only be found \emph{before} $N_0$. In this case, we
-	consider a second element of a \kl{bad sequence} $u_2 \infleq w$, which is
-	incomparable with $u_1$ for the \kl{infix relation}. If $u_2$ is an
-	\kl{infix} of $w_{\geq N_0}$, then we can conclude as before. Otherwise,
-	notice that $u_1$ and $u_2$ cannot start at the same position in $w$
-	(because they are incomparable). Continuing this argument, we conclude that
-	there are at most $N_0$ elements starting before $N_0$
-	at the start of any sequence of
-	incomparable elements in $\infset{w}$. We conclude that
-	$\oWidth{\infset{w}} \leq \omegaOrd + N_0$.
-
-	Let us now justify that this bound is tight.
-	The \kl{Thue-Morse sequence} over a binary
-	alphabet $\set{a,b}$ has \kl{ordinal width} $\omegaOrd$
-	from \cref{thue-morse-ordinal:lemma}.
-	Given a number $N_0
-	\in \Nat$, one can construct an arbitrarily long \kl{antichain} of words for
-	the \kl{infix relation} by using a new letter $c$. When concatenating this
-	(finite) antichain as a prefix of the \kl{Thue-Morse sequence}, one obtains a
-	new (infinite) word $w$. It is clear that the \kl{ordinal width} of
-	$\infset{w}$ is now at least $\omegaOrd + N_0$.
 \end{proofof}
@@
 
+\begin{proofof}{infix-amalgamation:thm}
+    Assume that $L$ is \kl{well-quasi-ordered} by the \kl{infix relation},
+    and obtained by an \kl{amalgamation system}
+    $(\Sigma, R, E, \canrun)$.
+
+    Let us consider the set $M$ of minimal runs for the relation $\leq_E$,
+    which is finite because the latter is a \kl{well-quasi-ordering}. 
+    By \cref{gap-words-prefix-ordered:lem}, we know that for each minimal run $\rho \in M$,
+    each gap language $\GapLanguage{\rho}{i}$ of $\rho$ is totally ordered by $\infleq$.
+    Adapting the proof of language boundedness from \cite[Section 4.2]{ASZZ24}, we may conclude that $\GapLanguage{\rho}{i} \subseteq \InfPeriodChain{w}$ for some $w \in \GapLanguage{\rho}{i}$.
+    As $\InfPeriodChain{w}$ is language bounded and this property is stable under subsets, concatenation and finite union,
+    we can conclude that $L$ is bounded as well.
+\end{proofof}
+
 \begin{proofof}{dwclosed-infixes-wqo:lem}
@@
 
-\begin{proofof}{infix-amalgamation:thm}
-    Assume that $L$ is \kl{well-quasi-ordered} by the \kl{infix relation},
-    and obtained by an \kl{amalgamation system}
-    $(\Sigma, R, E, \canrun)$.
-
-    Let us consider the set $M$ of minimal runs for the relation $\leq_E$,
-    which is finite because the latter is a \kl{well-quasi-ordering}. 
-    By \cref{gap-words-prefix-ordered:lem}, we know that for each minimal run $\rho \in M$,
-    each gap language $\GapLanguage{\rho}{i}$ of $\rho$ is totally ordered by $\infleq$.
-    Adapting the proof of language boundedness from \cite[Section 4.2]{ASZZ24}, we may conclude that $\GapLanguage{\rho}{i} \subseteq \InfPeriodChain{w}$ for some $w \in \GapLanguage{\rho}{i}$.
-    As $\InfPeriodChain{w}$ is language bounded and this property is stable under subsets, concatenation and finite union,
-    we can conclude that $L$ is bounded as well.
-\end{proofof}
+\AP Let us briefly recall that a \intro{rational transduction} is a relation $R
+\subseteq \Sigma^* \times \Gamma^*$ such that there exists a finite state
+automaton that reads pairs of letters $(a,b) \in (\Sigma \cup \set{\epsilon})
+\times (\Gamma \cup \set{\epsilon})$ and recognizes $R$. A class of languages
+$\mathcal{C}$ is \intro{closed under rational transductions} if for every $L
+\in \mathcal{C}$ and every rational transduction $R$, the language $R(L)
+\defined \set{v \in \Gamma^* \mid \exists u \in L, (u,v) \in R}$ also belongs
+to $\mathcal{C}$.
 
@@
 \end{proofof}
+
+\section{Proofs for Section~\ref{conclusion:sec}}
+
+\begin{lemma}
+  \label{dwclosed-wqo-complexity-width-finite:lem}
+  Let $L$ be a \kl{downwards closed} language that is
+  \kl{well-quasi-ordered} by the \kl{infix relation}.
+  Then, the following are equivalent:
+  \begin{enumerate}
+    \item $L$ has bounded \kl{factor complexity},
+      \label{dwclosed-wqo-complexity-width-finite:item1}
+    \item $L$ has finite \kl{ordinal width}.
+      \label{dwclosed-wqo-complexity-width-finite:item2}
+    \item $L$ is a finite union 
+      of languages of the form
+      $\infset{w}$ where $w$
+      is an \kl{ultimately periodic} word.
+      \label{dwclosed-wqo-complexity-width-finite:item3}
+  \end{enumerate}
+\end{lemma}
+\begin{proof}
+  \cref{dwclosed-wqo-complexity-width-finite:item3} $\Rightarrow$
+  \cref{dwclosed-wqo-complexity-width-finite:item1} is clear because
+  \kl{ultimately periodic} words have bounded \kl{factor complexity}.
+
+  \cref{dwclosed-wqo-complexity-width-finite:item1} $\Rightarrow$
+  \cref{dwclosed-wqo-complexity-width-finite:item2} is also clear because
+  unbounded \kl{factor complexity} implies the existence of arbitrarily
+  large \kl{antichains}.
+
+  Finally, \cref{dwclosed-wqo-complexity-width-finite:item2} $\Rightarrow$
+  \cref{dwclosed-wqo-complexity-width-finite:item3} is a direct consequence
+  of \cref{dw-closed-wqo-charac:thm} and the fact that bounded \kl{factor
+  complexity} implies that the (bi)infinite words describing the language
+  are \kl{ultimately periodic}.
+\end{proof}

### 2025-10-16 src/conclusion.tex
final version

@@
 would be required to obtain decidability for these languages.
+
+To conclude on a positive note for the \kl{infix relation}, our results show
+that for \kl{downwards closed} and \kl{well-quasi-ordered} languages, there is
+a strong connection between the \kl{factor complexity} and the \kl{ordinal
+width}: it is the same to have bounded \kl{factor complexity} and finite
+\kl{ordinal width}. A short proof can be found in appendix 
+(\cref{dwclosed-wqo-complexity-width-finite:lem}).

### 2025-10-16 wqo-on-words.lncs.tex
final version

--- 2025-10-15 wqo-on-words.lncs.tex
@@
 \linenumbers
-
-\usepackage{todonotes}
 
@@
 \Crefname{conjecture}{Conjecture}{Conjectures}
-\usepackage[composition,hyperref,xcolor,cleveref]{knowledge}
+\usepackage[electronic,hyperref,xcolor,cleveref]{knowledge}
 \knowledgeconfigure{notion}
@@
 %
-\title{Well-quasi-orderings on word languages \\
-       main 81b14ea82bdf369ba5dbe8de29fa418756900125 \\
-       2025-10-15 17:43:32 +0200
-}
+\title{Well-quasi-orderings on word languages}
 \titlerunning{WQO on Words}
@@
 More specifically, we are interested in characterizing \emph{languages} of words that are well-quasi-ordered under these orderings, and explore their properties and connections with other language theoretic notions.
-We furthermore give decition procedures when the languages are given by various comupational models such as automata, context-free grammars, and automatic structures.
+We furthermore give decision procedures when the languages are given by various computational models such as automata, context-free grammars, and automatic structures.
 \end{abstract}

### 2025-10-16 src/appendix.tex
extra bits of information in lemma appendix

@@
       \label{dwclosed-wqo-complexity-width-finite:item1}
-    \item $L$ has finite \kl{ordinal width}.
+    \item $L$ has finite \kl{ordinal width},
       \label{dwclosed-wqo-complexity-width-finite:item2}
+    \item $L$ is a finite union of \kl{chains},
+      \label{dwclosed-wqo-complexity-width-finite:item4}
     \item $L$ is a finite union 
@@
       $\infset{w}$ where $w$
-      is an \kl{ultimately periodic} word.
+      is an ultimately periodic word.
       \label{dwclosed-wqo-complexity-width-finite:item3}
@@
 \begin{proof}
-  \cref{dwclosed-wqo-complexity-width-finite:item3} $\Rightarrow$
+  First, \cref{dwclosed-wqo-complexity-width-finite:item4} $\iff$
+  \cref{dwclosed-wqo-complexity-width-finite:item2} is a standard
+  fact regarding \kl{ordinal width}.
+
+  Then, \cref{dwclosed-wqo-complexity-width-finite:item3} $\Rightarrow$
   \cref{dwclosed-wqo-complexity-width-finite:item1} is clear because
-  \kl{ultimately periodic} words have bounded \kl{factor complexity}.
+  ultimately periodic words have bounded \kl{factor complexity}.
 
-  \cref{dwclosed-wqo-complexity-width-finite:item1} $\Rightarrow$
+  In turn, \cref{dwclosed-wqo-complexity-width-finite:item1} $\Rightarrow$
   \cref{dwclosed-wqo-complexity-width-finite:item2} is also clear because

### 2025-10-17 src/infixes.tex
fix: knowledges + refs + storytelling
Fix grammar and wording issues in paper
Co-authored-by: AliaumeL <24816735+AliaumeL@users.noreply.github.com>

--- 2025-10-16 src/infixes.tex
@@
             its \kl{downwards closure}
-            $\dwset[\preceq]{x}$ is finite.
+            $\dwset[\preceq]{x}$ is finite
+            (that is, $(X, \preceq)$ has \intro{finite initial segments}).
     \end{enumerate}
@@
     $L$ is a \kl{well-quasi-order} when endowed with the 
-    \kl{infix relation} if and only if it included in a finite union of 
+    \kl{infix relation} if and only if it is included in a finite union of 
     products $S_i \cdot P_i$ where 
@@
     for all $1 \leq i \leq n$.
-\end{theorem}
+\end{theorem>
 
@@
     that is \kl{well-quasi-ordered} by the \kl{infix relation}.
-    Then, the \kl{ordinal width} of $L$ less than $\omega^2$,
+    Then, the \kl{ordinal width} of $L$ is less than $\omega^2$,
     its \kl{ordinal height} is at most $\omega$,
-    and its \kl{maximal order type} less than $\omega^3$.
+    and its \kl{maximal order type} is less than $\omega^3$.
     Furthermore, those three bounds are tight.
@@
   \proofref{small-ordinal-invariants:thm}
+    Let $L$ be a \kl{well-quasi-ordered} \kl{downwards closed} language for the \kl{infix relation}.
     Then, the \kl{maximal order type} of $L$ is strictly less than $\omegaOrd^3$,
@@
     Furthermore, those bounds are tight.
-\end{corollary}
+\end{corollary>
 
@@
 (trivially) translated into a finite automaton that reads the base $b$
-representation of the free variables (that are indices of the sequence). It is
-folklore that one can decide if an \kl{automatic sequence} is \kl{ultimately
-uniformly recurrent}. A proof can be found in the appendix at
-\cref{automatic-uur:lem}. Based on this, we now prove:
+representation of the free variables (that are indices of the sequence). In
+particular, it follows that one can decide if an \kl{automatic
+sequence} is \kl{ultimately uniformly recurrent}, a proof of this folklore
+result can be found in the appendix at \cref{automatic-uur:lem}. Based on this,
+we now prove:
 

### 2025-10-17 src/introduction.tex
Fix grammar and wording issues in paper
Co-authored-by: AliaumeL <24816735+AliaumeL@users.noreply.github.com>

@@
 well-quasi-ordered sets is also stable under the operation of taking finite
-words and finite trees labelled by elements of a well-quasi-ordered set~\cite{HIG52,KRU72}.
+words and finite trees labeled by elements of a well-quasi-ordered set~\cite{HIG52,KRU72}.
 
@@
 research area of well-quasi-orderings.
-Indeed, there has been recent breakthroughs in deciding whether a given order
+Indeed, there have been recent breakthroughs in deciding whether a given order
 is a \kl{well-quasi-order}, for instance in the context of the verification of
@@
 We then turn our attention to the \kl{infix relation}. In this case, we notice
-that Lemma 5.1 from \cite{DBLP:journals/ita/Kuske06} imply that there are
+that Lemma 5.1 from \cite{DBLP:journals/ita/Kuske06} implies that there are
 \kl{well-quasi-ordered} languages for the \kl{infix relation} that have
@@
 relations over finite words is far from new. For instance, orders obtained by
-so-called \emph{derivation relations} where already analysed by Bucher,
+so-called \emph{derivation relations} were already analysed by Bucher,
 Ehrenfeucht, and Haussler \cite{BUEUD85}, and were later extended by

### 2025-10-17 src/prefixes.tex
Add thmtools integration and demonstrate restatable theorem proof of concept
Co-authored-by: AliaumeL <24816735+AliaumeL@users.noreply.github.com>
Successfully implement theorem restatement in appendix
Co-authored-by: AliaumeL <24816735+AliaumeL@users.noreply.github.com>
Fix grammar and wording issues in paper
Co-authored-by: AliaumeL <24816735+AliaumeL@users.noreply.github.com>

@@
     \kl{well-quasi-ordered} by the \kl{prefix relation}. Then,
-    \kl{maximal order type} of $L$ strictly smaller than $\omega^2$,
+    the \kl{maximal order type} of $L$ is strictly smaller than $\omega^2$,
     the \kl{ordinal height} of $L$ is at most $\omega$, and

### 2025-10-17 src/appendix.tex
Fix additional grammar issues and verify compilation
Co-authored-by: AliaumeL <24816735+AliaumeL@users.noreply.github.com>

@@
     \end{equation*}
-    Hence, that $L$ is finite union of \kl{chains}.
+    Hence, $L$ is a finite union of \kl{chains}.
 \end{proofof}
@@
    We can then
-    directly conclude that $\oWidth{L_i}$ less than $\omegaOrd$ (in the case of
+    directly conclude that $\oWidth{L_i}$ is less than $\omegaOrd$ (in the case of
     a finite word), less than $\omegaOrd \cdot 2$ (in the case of an infinite

### 2025-10-17 src/introduction.tex
Final clarity improvements and complete review
Co-authored-by: AliaumeL <24816735+AliaumeL@users.noreply.github.com>

@@
 straightforwardly generalize to infinite quasi-ordered alphabets.
-Unfortunately, it is easy to see that none of these constructions are 
-well-quasi-ordered as soon as the alphabet contains two distinct letters:
-for instance, the infinite sequence $a b^n a$ is \kl{well-quasi-ordered} by
+Unfortunately, it is easy to see that none of these relations yield
+well-quasi-ordered sets as soon as the alphabet contains two distinct letters:
+for instance, the infinite sequence of words $(a b^n a)_{n \in \Nat}$ is \kl{well-quasi-ordered} by
 the \kl{subword relation} but by neither the \kl{prefix relation}, nor the

### 2025-10-17 src/conclusion.tex
fix: error compilation + too long

@@
 our positive results.
-
-\paragraph*{Monoid equations}  It could be interesting to understand which
-monoids $M$ recognize languages that are \kl{well-quasi-ordered} by the
-\kl{infix}, \kl{prefix} or \kl{suffix} relations. This research direction is
-connected to finding which classes of graphs of \emph{bounded clique-width} are
-\kl{well-quasi-ordered} with respect to the \emph{induced subgraph relation},
-as shown in \cite{DRT10}, and recently revisited in \cite{lopez25}.
 

### 2025-12-08 rebuttal.txt
more rebuttal

We thank the reviewers for their time and their feedback. Of course we aim to
remedy all unclear sections. Here we'd like to respond to some of the points
that have been raised in more detail:

Reviewer 1
----------
- Can you expand a bit on the applications you envision your results can have that cannot be solved with existing tools?
Response:

Applications of well-quasi-orderings are numerous in verification, ranging from
termination proofs to decidability of inclusion problems. For the sake of the
argument, let us pick one example (among many)  is the use of wqos in so-called
"well-structured transition systems" (WSTS), where the decidability of certain
verification problems (e.g., coverability) is guaranteed by the presence of a
wqo on the state space that is compatible with the transition relation. In some
cases (like lossy channel systems), the subword ordering appears naturally to
order the configurations, but in most cases the natural ordering on words are
the prefix, suffix, and infix orderings.

To apply the WSTS theory in these cases, one way is to ensure that the set of
reachable configurations (which is often a language) is wqo under the relevant
ordering. Our results provide a complete characterization of such languages,
along with effective decision procedures for languages described in a vast
variety of formalisms. This completely solves the question of whether a WSTS
with configurations ordered by prefix, suffix, or infix ordering can be used
for verification. From our characterizations, one can understand that whenever
  one could use infix, prefix, or suffix orderings to obtain a wqo, the set of
  reachable configurations must be "simple" in a precise sense, which we
  interpret as a formal limitation to the expressiveness of such systems.

Reviewer 2
----------
- You should give a proof for the measures of the languages Lk in the proof of Corollary 14.
Response:
We will add a detailed proof in the final version. Here is a sketch of the argument:
The ordinal height of (infinite) languages is always ω,
the ordinal width of a*b* is ω (since it is isomorphic to N^2 with the product order),
hence disjoint copies of a*b* have width ω + .... + ω, hence the width of L_k is ω·k.
Finally, the maximal order type of a disjoint sum is the natural sum (Hessenberg sum) of the maximal order types of the summands, which here is $ω^2$ added k times, hence $ω^2 \cdot k$.
We are using classical results on ordinal invariants of wqos that we will take from 
[15].

- You could corollary 21 at the end of section 5.1, since its proof call for lemmas 25 and 26. Or move these lemmas to appendix.
Response:
Yes, this was made to improve the flow of the paper by stating important results first,
but in this case it is counterproductive.

- Part of the claim of Lemma 26 seems redundant with Lemma 24.
Response:
Yes, the first part is the same and will be removed.

- Unfortunately space is limited, but it would have been nice to have Figures 6 and 7 in the core of the paper.
Response:
These figures are quite large, and scaling them down would make them less readable.
We will try moving them to the main text by compressing the prefix/suffix section 
that is less central to the main results.

Reviewer 3
----------
- Wouldn't it be more concise to characterise (in Thms 8 and 20) 
IDEALS of languages well quasi ordered by the infix ordering, 
instead of the wqo languages themselves?
Response: For Theorem 20, the characterization of the corresponding order ideals is present in the supporting lemmas, though we consider the phrasing of Theorem 20 a matter of taste. For Theorem 8, TODO

- How much the characterisation of Lemma 22 can be generalised
beyond well quasi orders, namely for ideals in the infix order?
Response: The proof can in fact be generalized to all order ideals 
of the infix order, and we thank the Reviewer for pointing this out.
We will adjust the statement and the proof.

- Why Corollary 34 does not include VASS languages?
Response: Corollary 34 is meant to emphasize a subpart of our results
that we think might be of interest to a broader audience. The fully general
version of this statement is Theorem 30, which of course includes VASS and 
also PVASS languages.

- There is no a really long and complex proof in the paper, rather a series of
short arguments. As a reader I had an impressions of seeing a series
of small fragmental results, while I'd probably prefer one big result
instead of 20 small ones. 
Response: As the space of infix well-quasi-orders was not well explored so far, we onsider this as a classification paper, which naturally lends itself to a large range of smaller statements rather than singular large ones.

- The section on prefix/suffix could be probably completely removed from the
paper, with the extra space devoted to the most important arguments that are
now in the appendix.
Response: We agree and will condense the section significantly to free up space,
especially to reintroduce the diagrams to the main text as requested by Reviewer 2.

- L.197-199: unclear what is the tree relation here.
Response: The trees are formed as follows: The root node is the empty sequence, ε. Each node has (potentially infinitely many)
children from the wqoset (X, ≤) such that the sequence of elements from the root to this child satisfy the property expressed
by the tree (i.e., being bad, decreasing or an antichain respectively).

- Thm 8: what are "products"?
Response: By the "product" of two languages L and R we mean the usual pairwise concatenation of words from L and R.

- Lemma 10: what does it mean "simultaneously" here?
Response: By "simulatenously" we mean that we can find a finite set of languages such that their union is P↓(x), and each
language is a chain by the prefix relation, and by the infix relation, and by the suffix relation.

- Lemma 12: what is the meaning of "for arbitrarily large values of n" here?
Response: By "for abritrarily large values of n", we formally mean: ∀k ∃n≥k such that the property holds.

- Lemma 12: point 2 seems to be subsumed by 3, isn't it?
Response: Correct, both point 1 and point 2 can be subsumed by point 3. The phrasing was left over from a previous version of the proof. We will adjust this.

- L.605: what it means "we may fix"?
Response: The concatenative property of amalgamation systems simply requires that each gap word is the concatenation of its 
constituent gap words from the smaller runs, but not in which order. We may pick any gap and pick the ordering in that gap,
but the ordering in the other gaps may then be imposed on us by the structure of the system. For example, when amalgamating
derivation trees of CFGs, each gap has a sibling gap on the intuitive "other side" of the tree that must be concatenated in
the reverse order.

- Lemma 32: "or vice versa" not clear.
Response: By "vice versa" we mean that either G_f ≤ G_g or G_g ≤ G_f

### 2025-12-08 rebuttal.txt
start rebuttal

@@
-We thank the reviewers for their time and their feedback. Of course we aim to
-remedy all unclear sections. Here we'd like to respond to some of the points
-that have been raised in more detail:
+We thank the reviewers for their time and their feedback. Of course we aim to remedy all unclear sections. Here we'd 
+like to respond to some of the points that have been raised in more detail:
 
@@
 
-Applications of well-quasi-orderings are numerous in verification, ranging from
-termination proofs to decidability of inclusion problems. For the sake of the
-argument, let us pick one example (among many)  is the use of wqos in so-called
-"well-structured transition systems" (WSTS), where the decidability of certain
-verification problems (e.g., coverability) is guaranteed by the presence of a
-wqo on the state space that is compatible with the transition relation. In some
-cases (like lossy channel systems), the subword ordering appears naturally to
-order the configurations, but in most cases the natural ordering on words are
-the prefix, suffix, and infix orderings.
-
-To apply the WSTS theory in these cases, one way is to ensure that the set of
-reachable configurations (which is often a language) is wqo under the relevant
-ordering. Our results provide a complete characterization of such languages,
-along with effective decision procedures for languages described in a vast
-variety of formalisms. This completely solves the question of whether a WSTS
-with configurations ordered by prefix, suffix, or infix ordering can be used
-for verification. From our characterizations, one can understand that whenever
-  one could use infix, prefix, or suffix orderings to obtain a wqo, the set of
-  reachable configurations must be "simple" in a precise sense, which we
-  interpret as a formal limitation to the expressiveness of such systems.
-
 Reviewer 2
@@
 Response:
-We will add a detailed proof in the final version. Here is a sketch of the argument:
-The ordinal height of (infinite) languages is always ω,
-the ordinal width of a*b* is ω (since it is isomorphic to N^2 with the product order),
-hence disjoint copies of a*b* have width ω + .... + ω, hence the width of L_k is ω·k.
-Finally, the maximal order type of a disjoint sum is the natural sum (Hessenberg sum) of the maximal order types of the summands, which here is $ω^2$ added k times, hence $ω^2 \cdot k$.
-We are using classical results on ordinal invariants of wqos that we will take from 
-[15].
 
@@
 Response:
-Yes, this was made to improve the flow of the paper by stating important results first,
-but in this case it is counterproductive.
 
@@
 Response:
-Yes, the first part is the same and will be removed.
 
@@
 Response:
-These figures are quite large, and scaling them down would make them less readable.
-We will try moving them to the main text by compressing the prefix/suffix section 
-that is less central to the main results.
 
@@
 instead of the wqo languages themselves?
-Response: For Theorem 20, the characterization of the corresponding order ideals is present in the supporting lemmas, though we consider the phrasing of Theorem 20 a matter of taste.
-For Theorem 8, TODO
+Response:
 
@@
 beyond well quasi orders, namely for ideals in the infix order?
-Response: The proof can in fact be generalized to all order ideals 
-of the infix order, and we thank the Reviewer for pointing this out.
-We will adjust the statement and the proof.
+Response:
 
 - Why Corollary 34 does not include VASS languages?
-Response: Corollary 34 is meant to emphasize a subpart of our results
-that we think might be of interest to a broader audience. The fully general
-version of this statement is Theorem 30, which of course includes VASS and 
-also PVASS languages.
+Response:
 
@@
 instead of 20 small ones. 
-Response: As the space of infix well-quasi-orders was not well explored so far, we onsider this as a classification paper, which naturally lends itself to a large range of smaller statements rather than singular large ones.
+Response:
+
+- There are numerous forward references (e.g. L.325,425,585,606), please avoid them!
+Response:
 
@@
 now in the appendix.
-Response: We agree and will condense the section significantly to free up space,
-especially to reintroduce the diagrams to the main text as requested by Reviewer 2.
+Response:
+
+- In Sect.6 it would be worth it to explain better how the lemmas 
+are used in the proofs of Thms 29,.30.
+Response:
+
+- In the appendix, as a good practise it is worth repeating the statements
+of the proved theorems/lemmas.
+Response:
 
@@
 - Lemma 12: point 2 seems to be subsumed by 3, isn't it?
-Response: Correct, both point 1 and point 2 can be subsumed by point 3. The phrasing was left over from a previous version of the proof. We will adjust this.
 

### 2025-12-09 rebuttal.txt
more rebuttal 2
fix: rebuttal
fix: mini paragraph intro rebuttal

--- 2025-12-08 rebuttal.txt
@@
-We thank the reviewers for their time and their feedback. Of course we aim to
-remedy all unclear sections. Here we'd like to respond to some of the points
+We thank the reviewers for their time and their feedback, and will remedy
+all unclear sections in the final version of the paper.
+
+There are points that were raised by multiple reviewers regarding the
+applications of our results and the seemingly fragmented nature of the paper.
+We strongly disagree with both points. Ours is a theoretical paper, but the use
+of well-quasi-orderings are numerous in verification, ranging from termination
+proofs to decidability of inclusion problems. For the sake of the argument, let
+us pick one example (among many) as the use of wqos in so-called
+"well-structured transition systems" (WSTS), where the decidability of certain
+verification problems (e.g., coverability) is guaranteed by the presence of a
+wqo on the state space that is compatible with the transition relation. In some
+cases (like lossy channel systems), the subword ordering appears naturally to
+order the configurations, but in most cases the natural ordering on words are
+the prefix, suffix, and infix orderings.
+
+To apply the WSTS theory in these cases, one way is to ensure that the set of
+reachable configurations is wqo under the relevant ordering.  Our results
+completely solves the question of whether a WSTS
+with configurations ordered by prefix, suffix, or infix ordering can be used
+for verification purposes. Furthermore, we prove that whenever
+one could use infix, prefix, or suffix orderings to obtain a wqo, the set of
+reachable configurations must be "simple" in a precise sense, which we
+interpret as a formal limitation to the expressiveness of such systems.
+
+In any case, this paper takes an approach significantly different from the
+classical wqo literature, in the sense that it finely classifies objects that
+are wqo inside a non-wqo, instead of exhibiting new ways to compositionally
+build wqos. As the space of infix well-quasi-orders was not well explored so
+far, we consider this as a classification paper, which naturally lends itself
+to a large range of smaller statements rather than singular large ones, without
+diminishing the importance of these results.
+
+Here we'd like to respond to some of the points
 that have been raised in more detail:
+
+----------- MATERIAL USED TO ANSWER THE REVIEWS --------------
 
@@
 Response:
+
+Ours is a theoretical paper, and as such will mostly have theoretical
+applications.
 
@@
 hence disjoint copies of a*b* have width ω + .... + ω, hence the width of L_k is ω·k.
-Finally, the maximal order type of a disjoint sum is the natural sum (Hessenberg sum) of the maximal order types of the summands, which here is $ω^2$ added k times, hence $ω^2 \cdot k$.
+Finally, the maximal order type of a disjoint sum is the natural sum (Hessenberg sum) of the maximal order types of the summands, which here is ω^2 added k times, hence ω^2 · k.
 We are using classical results on ordinal invariants of wqos that we will take from 
@@
 Response: For Theorem 20, the characterization of the corresponding order ideals is present in the supporting lemmas, though we consider the phrasing of Theorem 20 a matter of taste.
-For Theorem 8, TODO
+For Theorem 8, we don't think ideals are the right lens as the chains for the suffix and prefix relation aren't necessarily downward closed.
 

### 2025-12-09 rebuttal.txt
alternative solution to thm 8/thm 20?

@@
 instead of the wqo languages themselves?
-Response: For Theorem 20, the characterization of the corresponding order ideals is present in the supporting lemmas, though we consider the phrasing of Theorem 20 a matter of taste.
-For Theorem 8, we don't think ideals are the right lens as the chains for the suffix and prefix relation aren't necessarily downward closed.
+Response: For Theorem 8, we think that phrasing it in terms of ideals would needlessly complicate the statement. For Theorem 20, we are happy to rephrase it in terms of ideals.
 

### 2025-12-10 rebuttal.txt
fix: merging everything + dim down on sarcasm

@@
 applications of our results and the seemingly fragmented nature of the paper.
-We strongly disagree with both points. Ours is a theoretical paper, but the use
-of well-quasi-orderings are numerous in verification, ranging from termination
+We strongly disagree with both points. Ours is a theoretical paper that fits a
+long line of results on well-quasi-order results. Furthermore, the use of
+well-quasi-orderings are numerous in verification, ranging from termination
 proofs to decidability of inclusion problems. For the sake of the argument, let
@@
 that have been raised in more detail:
+
+r2 on corollary 14. We will add a detailed proof in the final version. A sketch
+of the argument is that a*b* is isomorphic to N^2 with the product order, for
+which ordinal invariants are well-known. The language L_k is now isomorphic to
+k disjoint copies of a*b* (using a coding of an alphabet of size 2k into a
+binary alphabet). Finally, the ordinal invariants of disjoint sums are computed
+compositionally using known formulas, yielding the desired result [15].
+
+r2 on corollary 21, r3 on forward references. We preferred stating early on the
+theorems of interest, and then explain which lemmas will be used and how they
+fit together in a general proof scheme, which naturally introduces forward
+references. As this was found confusing, we will update the paper to be more
+linear.
+
+r2 on lemma 26 and 24. We will remove the redundant part, thank you for pointing 
+it out.
+
+r2 on figures 6 and 7. We will try to be more succinct in the prefix/suffix section
+to free up space for these figures in the main text.
+
+r3 on prefix/suffix section. We agree and will condense the section
+significantly to free up space, especially to reintroduce the diagrams to the
+main text as requested by Reviewer 2.
+
+r3 on Lemma 12 point 2 and 3. Correct, both point 1 and point 2 imply point 3,
+and we will adjust the statement accordingly.
+
+r3 on lemma 22. The proof can in fact be generalized to all order ideals of the
+infix order, and we thank the Reviewer for pointing this out. We will adjust
+the statement and the proof.
+
+r3 on theorems 8 and 20. For Theorem 8, we think that phrasing it in terms of
+ideals would needlessly complicate the statement. For Theorem 20, we are happy
+to rephrase it in terms of ideals.
+
+r3 on corollary 34. Corollary 34 is clearly meant to emphasize a subpart of our
+results that we think might be of interest to a broader audience. The fully
+general version of this statement is Theorem 30, which of course includes VASS
+and also PVASS languages.
+
+r3 on L.605.  The concatenative property of amalgamation systems simply
+requires that each gap word is the concatenation of its constituent gap words
+from the smaller runs, but not in which order. We may pick any gap and pick the
+ordering in that gap, but the ordering in the other gaps may then be imposed on
+us by the structure of the system. For example, when amalgamating derivation
+trees of CFGs, each gap has a sibling gap on the intuitive "other side" of the
+tree that must be concatenated in the reverse order.
+
+r3 on L.197-199. The trees are formed as follows: The root node is the empty
+sequence, ε. Each node has (potentially infinitely many) children from the
+wqoset (X, ≤) such that the sequence of elements from the root to this child
+satisfy the property expressed by the tree (i.e., being bad, decreasing or an
+antichain respectively).
+
+r3 on Thm 8. The product of two languages L and R is the pairwise concatenation of words from L and R.
+
+r3 on Lemma 10: "U satisfies P for x, y, and z simulatenously" means that U satisfies P(x), P(y), and P(z).
+
+r3 on Lemma 12: "P(n) holds for abritrarily large values of n" means ∀k ∃n≥k P(n).
+
+r3 on Lemma 32: "x≤y or vice versa" means "x≤y or y≤x".
 

### 2025-12-10 rebuttal.txt
fix: merging everything + dim down on sarcasm

--- 2025-12-09 rebuttal.txt
@@
-We thank the reviewers for their time and their feedback, and will remedy
-all unclear sections in the final version of the paper.
+We thank the reviewers for their time and their feedback.
 
-There are points that were raised by multiple reviewers regarding the
-applications of our results and the seemingly fragmented nature of the paper.
-We strongly disagree with both points. Ours is a theoretical paper, but the use
-of well-quasi-orderings are numerous in verification, ranging from termination
-proofs to decidability of inclusion problems. For the sake of the argument, let
-us pick one example (among many) as the use of wqos in so-called
-"well-structured transition systems" (WSTS), where the decidability of certain
-verification problems (e.g., coverability) is guaranteed by the presence of a
-wqo on the state space that is compatible with the transition relation. In some
-cases (like lossy channel systems), the subword ordering appears naturally to
-order the configurations, but in most cases the natural ordering on words are
-the prefix, suffix, and infix orderings.
-
-To apply the WSTS theory in these cases, one way is to ensure that the set of
-reachable configurations is wqo under the relevant ordering.  Our results
-completely solves the question of whether a WSTS
-with configurations ordered by prefix, suffix, or infix ordering can be used
-for verification purposes. Furthermore, we prove that whenever
-one could use infix, prefix, or suffix orderings to obtain a wqo, the set of
-reachable configurations must be "simple" in a precise sense, which we
-interpret as a formal limitation to the expressiveness of such systems.
-
-In any case, this paper takes an approach significantly different from the
-classical wqo literature, in the sense that it finely classifies objects that
-are wqo inside a non-wqo, instead of exhibiting new ways to compositionally
-build wqos. As the space of infix well-quasi-orders was not well explored so
-far, we consider this as a classification paper, which naturally lends itself
-to a large range of smaller statements rather than singular large ones, without
-diminishing the importance of these results.
+We are puzzled by few of the questions. Ours is a paper that fits a
+well-established domain of research in theoretical computer science.
+Furthermore, this paper takes a new approach by not constructing new wqos, but
+classifying subsets that are wqo inside a non-wqo, and providing effective
+decision procedures whenever possible. As this is a classification paper, it
+naturally lends itself to a large range of smaller statements rather than one
+big one, without diminishing the importance of these results. If having one big
+statement with a hard proof was a requirement for publication, we did not see
+it written on the call for paper and profoundly apologize for missing it.
 
@@
 that have been raised in more detail:
+
+r2 on corollary 14. We will add a detailed proof in the final version. A sketch
+of the argument is that L_k is isomorphic to k copies of N^2, and one can
+easily compute its ordinal invariants [see 15].
+
+r2 on corollary 21, r3 on forward references. We preferred stating early on the
+theorems of interest, and then explain which lemmas will be used and how they
+fit together in a general proof scheme. This naturally introduces forward
+references. As this was found confusing, we will update the paper to be more
+linear.
+
+r2 on lemma 26 and 24. We will remove the redundant part, thank you for pointing 
+it out.
+
+r2 on figures 6 and 7, r3 on prefix/suffix section. We will try to be more succinct in the prefix/suffix section
+to free up space for these figures in the main text.
+
+r3 on Lemma 12 point 2 and 3. Correct, both point 1 and point 2 imply point 3,
+and we will adjust the statement accordingly.
+
+r3 on lemma 22. The proof can in fact be generalized to all order ideals of the
+infix order, and we thank the Reviewer for pointing this out. We will adjust
+the statement and the proof.
+
+r3 on theorems 8 and 20. For Theorem 8, we think that phrasing it in terms of
+ideals would needlessly complicate the statement. For Theorem 20, we are happy
+to rephrase it in terms of ideals.
+
+r3 on corollary 34. Corollary 34 is clearly meant to emphasize a subpart of our
+results that we think might be of interest to a broader audience. The fully
+general version of this statement is Theorem 30, which of course includes VASS
+and also PVASS languages.
+
+r3 on L.605.  The concatenative property of amalgamation systems simply
+requires that each gap word is the concatenation of its constituent gap words
+from the smaller runs, but not in which order. We may pick any gap and pick the
+ordering in that gap, but the ordering in the other gaps may then be imposed on
+us by the structure of the system.
+
+r3 on L.197-199. The tree is obtained by considering bad sequences (resp. antichains, decreasing sequences) as nodes, and the ancestor relation is "having common prefix".
+
+r3.
+- The product of two languages L and R is the pairwise concatenation of words from L and R.
+ "U satisfies P for x, y, and z simulatenously" means that U satisfies P(x), P(y), and P(z).
+- "P(n) holds for abritrarily large values of n" means ∀k ∃n≥k P(n).
+- "x≤y or vice versa" means "x≤y or y≤x".
 

### 2025-12-10 rebuttal.txt
fix: merging everything + dim down on sarcasm
Some minor polish

@@
 We thank the reviewers for their time and their feedback.
+We thank the reviewers for their time and their insightful feedback and would 
+like to take some time responding to some points in more detail.
 
-We are puzzled by few of the questions. Ours is a paper that fits a
-well-established domain of research in theoretical computer science.
-Furthermore, this paper takes a new approach by not constructing new wqos, but
+Regarding the lack of a singular large result: This paper takes a somewhat
+unusual approach in wqo theory by not constructing new wqos, but
 classifying subsets that are wqo inside a non-wqo, and providing effective
@@
 naturally lends itself to a large range of smaller statements rather than one
-big one, without diminishing the importance of these results. If having one big
-statement with a hard proof was a requirement for publication, we did not see
-it written on the call for paper and profoundly apologize for missing it.
+big one, without diminishing the importance of these results.
 
-Here we'd like to respond to some of the points
-that have been raised in more detail:
-
-r2 on corollary 14. We will add a detailed proof in the final version. A sketch
+R2 on Corollary 14. We will add a detailed proof in the final version. A sketch
 of the argument is that L_k is isomorphic to k copies of N^2, and one can
@@
 
-r2 on corollary 21, r3 on forward references. We preferred stating early on the
+R2 on Corollary 21, R3 on forward references. We preferred stating early on the
 theorems of interest, and then explain which lemmas will be used and how they
@@
 
-r2 on lemma 26 and 24. We will remove the redundant part, thank you for pointing 
+R2 on Lemmata 26 and 24. We will remove the redundant part, thank you for pointing 
 it out.
 
-r2 on figures 6 and 7, r3 on prefix/suffix section. We will try to be more succinct in the prefix/suffix section
-to free up space for these figures in the main text.
+R2 on Figures 6 and 7, R3 on prefix/suffix section. We will be more succinct in the prefix/suffix section to free up space for these figures in the main text.
 
-r3 on Lemma 12 point 2 and 3. Correct, both point 1 and point 2 imply point 3,
+R3 on Lemma 12 point 2 and 3. Correct, both point 1 and point 2 imply point 3,
 and we will adjust the statement accordingly.
 
-r3 on lemma 22. The proof can in fact be generalized to all order ideals of the
+R3 on Lemma 22. The proof can in fact be generalized to all order ideals of the
 infix order, and we thank the Reviewer for pointing this out. We will adjust
@@
 
-r3 on theorems 8 and 20. For Theorem 8, we think that phrasing it in terms of
+R3 on Theorems 8 and 20. For Theorem 8, we think that phrasing it in terms of
 ideals would needlessly complicate the statement. For Theorem 20, we are happy
@@
 
-r3 on corollary 34. Corollary 34 is clearly meant to emphasize a subpart of our
+R3 on Corollary 34. Corollary 34 is meant to emphasize a subpart of our
 results that we think might be of interest to a broader audience. The fully
@@
 
-r3 on L.605.  The concatenative property of amalgamation systems simply
+R3 on L.605.  The concatenative property of amalgamation systems simply
 requires that each gap word is the concatenation of its constituent gap words
@@
 
-r3 on L.197-199. The tree is obtained by considering bad sequences (resp. antichains, decreasing sequences) as nodes, and the ancestor relation is "having common prefix".
+R3 on L.197-199. The tree is obtained by considering bad sequences (resp. antichains, decreasing sequences) as nodes, and the ancestor relation is "having common prefix".
 
-r3.
+R3.
 - The product of two languages L and R is the pairwise concatenation of words from L and R.
@@
 Response: By "vice versa" we mean that either G_f ≤ G_g or G_g ≤ G_f
+
+- Lemma 32: "or vice versa" not clear.
+Response: By "vice versa" we mean that either G_f ≤ G_g or G_g ≤ G_f

### 2025-12-10 rebuttal.txt
Application

@@
-We thank the reviewers for their time and their feedback.
 We thank the reviewers for their time and their insightful feedback and would 
 like to take some time responding to some points in more detail.
+
+Regarding the potential applications: Almost all wqo results have some form
+of algorithmic application. For example, our results can be used as a
+regularity check for trace languages: Consider the trace language of an 
+amalgamation system that is being formally verified. As a trace language 
+is naturally prefix closed, our results mean that we can decide whether
+it is well-quasi-ordered by the prefix relation (Theorem 30). From Lemma 33
+we can conclude that this is the case iff it is regular. Furthermore,
+a closer analysis of our proof even yields an effective procedure for 
+constructing a regular representation.
 

### 2025-12-10 rebuttal.txt
Shorten application

--- 2025-12-10 rebuttal.txt
@@
-We thank the reviewers for their time and their feedback.
-We thank the reviewers for their time and their insightful feedback and would 
-like to take some time responding to some points in more detail.
+We thank the reviewers for their time and their insightful feedback. 
+The reviewers made relevant comments and suggestions that will be
+implemented in the final version of the article.
+
+We now answer the questions of R1 and R3:
+
+Regarding the potential applications: Almost all wqo results have some form
+of algorithmic application. For example, our results can be used as an 
+(effective!) regularity check for trace languages and safety properties 
+in amalgamation systems.
 
@@
 big one, without diminishing the importance of these results.
+
+--- cut
 

### 2025-12-10 rebuttal.txt
fix: add details for funding

@@
 of the argument is that L_k is isomorphic to k copies of N^2, and one can
-easily compute its ordinal invariants [see 15].
+easily compute its ordinal invariants using standard methods [15].
 
@@
 - The product of two languages L and R is the pairwise concatenation of words from L and R.
-"U satisfies P for x, y, and z simulatenously" means that U satisfies P(x), P(y), and P(z).
-- "P(n) holds for abritrarily large values of n" means ∀k ∃n≥k P(n).
+- "U satisfies P for x, y, and z simultaneously" means that U satisfies P(x), P(y), and P(z).
+- "P(n) holds for arbitrarily large values of n" means ∀k ∃n≥k P(n).
 - "x≤y or vice versa" means "x≤y or y≤x".

### 2025-12-10 rebuttal.txt
Update rebuttal.txt

@@
 We thank the reviewers for their time and their insightful feedback. 
-The reviewers made relevant comments and suggestions that will be
-implemented in the final version of the article.
+The reviewers made relevant comments and suggestions that will be implemented in the final version of the article.
 
-We now answer the questions of R1 and R3:
+We first answer the main comments and questions of R1 and R3 and give below a more detailed response to other comments/questions.
 
-Regarding the potential applications: Almost all wqo results have some form
-of algorithmic application. For example, our results can be used as an 
-(effective!) regularity check for trace languages and safety properties 
-in amalgamation systems.
+---MAIN QUESTIONS
 
-Regarding the lack of a singular large result: This paper takes a somewhat
-unusual approach in wqo theory by not constructing new wqos, but
-classifying subsets that are wqo inside a non-wqo, and providing effective
-decision procedures whenever possible. As this is a classification paper, it
-naturally lends itself to a large range of smaller statements rather than one
-big one, without diminishing the importance of these results.
+--R1
+Comments: We thank R1 for their positive and useful feedback.
 
---- cut
+Question of R1: Can you expand a bit on the applications you envision your results can have that cannot be solved with existing tools?
+
+Answer: Almost all wqo results have some form of algorithmic application. For example, our results can be used as an 
+(effective!) regularity check for trace languages and safety properties in amalgamation systems.
+
+--R3
+Comments: We thank R3 for their appreciation of the quality of the technical level of the paper.
+Their main concern is the lack of a single large and complex technical contribution.
+Our answer regarding this is twofold: 1) we've put a lot of effort into breaking down the technical difficulties noticed by the reviewer into manageable parts, in order to make the proofs as accessible as possible.
+2) The nature of the article is more of a classification paper rather than a solution to a known hard to solve problem.
+Indeed, while we are happy that the technical contributions have been appriciated by the reviewer, we believe that the main contribution of the article lies in its originality and not in solving a particularly tough problem.
+
+Question 1: Wouldn't it be more concise to characterise (in Thms 8 and 20) 
+IDEALS of languages well quasi ordered by the infix ordering, 
+instead of the wqo languages themselves?
+
+Answer: Thank you for this suggestion. Concerning Theorem 8, we think that phrasing it in terms of
+ideals would needlessly complicate the statement. For Theorem 20, we are happy
+to rephrase it in terms of ideals.
+
+Question 2: How much the characterisation of Lemma 22 can be generalised
+beyond well quasi orders, namely for ideals in the infix order?
+
+Answer: The proof can in fact be generalized to all order ideals of the
+infix order, and we thank Reviewer 3 for pointing this out. We will adjust
+the statement and the proof.
+
+Question 3: Why Corollary 34 does not include VASS languages?
+Corollary 34 is meant to emphasize a subpart of our
+results that we think might be of interest to a broader audience. The fully
+general version of this statement is Theorem 30, which of course includes VASS
+and also PVASS languages.
+
+---DETAILED ANSWER TO OTHER COMMENTS
 
@@
 and we will adjust the statement accordingly.
-
-R3 on Lemma 22. The proof can in fact be generalized to all order ideals of the
-infix order, and we thank the Reviewer for pointing this out. We will adjust
-the statement and the proof.
-
-R3 on Theorems 8 and 20. For Theorem 8, we think that phrasing it in terms of
-ideals would needlessly complicate the statement. For Theorem 20, we are happy
-to rephrase it in terms of ideals.
 

### 2025-12-10 rebuttal.txt
fix: typos

@@
-We thank the reviewers for their time and their insightful feedback. 
-The reviewers made relevant comments and suggestions that will be implemented in the final version of the article.
+We thank the reviewers for their time and their insightful feedback. The
+reviewers made relevant comments and suggestions that will be implemented in
+the final version of the article.
 
-We first answer the main comments and questions of R1 and R3 and give below a more detailed response to other comments/questions.
-
----MAIN QUESTIONS
+We first answer the main comments and questions of R1 and R3 and give below a
+more detailed response to other comments/questions.
 
@@
 
-Question of R1: Can you expand a bit on the applications you envision your results can have that cannot be solved with existing tools?
+Question of R1: Can you expand a bit on the applications you envision your
+results can have that cannot be solved with existing tools?
 
-Answer: Almost all wqo results have some form of algorithmic application. For example, our results can be used as an 
-(effective!) regularity check for trace languages and safety properties in amalgamation systems.
+Answer: Almost all wqo results have some form of algorithmic application. For
+example, our results can be used as an (effective!) regularity check for trace
+languages and safety properties in amalgamation systems.
 
@@
 Their main concern is the lack of a single large and complex technical contribution.
-Our answer regarding this is twofold: 1) we've put a lot of effort into breaking down the technical difficulties noticed by the reviewer into manageable parts, in order to make the proofs as accessible as possible.
-2) The nature of the article is more of a classification paper rather than a solution to a known hard to solve problem.
-Indeed, while we are happy that the technical contributions have been appriciated by the reviewer, we believe that the main contribution of the article lies in its originality and not in solving a particularly tough problem.
+
+Our answer regarding this is twofold: 1) we've put a lot of effort into
+breaking down the technical difficulties noticed by the reviewer into
+manageable parts, in order to make the proofs as accessible as possible. 2) The
+nature of the article is more of a classification paper rather than a solution
+to a known hard to solve problem. Indeed, while we are happy that the technical
+contributions have been appreciated by the reviewer, we believe that the main
+contribution of the article lies in its originality and not in solving a
+particularly tough problem.
 

### 2025-12-10 rebuttal.txt
Remove old stuff

@@
 
-R2 on Figures 6 and 7, R3 on prefix/suffix section. We will be more succinct in the prefix/suffix section to free up space for these figures in the main text.
-
 R3 on Lemma 12 point 2 and 3. Correct, both point 1 and point 2 imply point 3,
@@
 - "x≤y or vice versa" means "x≤y or y≤x".
-
------------ MATERIAL USED TO ANSWER THE REVIEWS --------------
-
-Reviewer 1
-----------
-- Can you expand a bit on the applications you envision your results can have that cannot be solved with existing tools?
-Response:
-
-Ours is a theoretical paper, and as such will mostly have theoretical
-applications.
-
-Applications of well-quasi-orderings are numerous in verification, ranging from
-termination proofs to decidability of inclusion problems. For the sake of the
-argument, let us pick one example (among many)  is the use of wqos in so-called
-"well-structured transition systems" (WSTS), where the decidability of certain
-verification problems (e.g., coverability) is guaranteed by the presence of a
-wqo on the state space that is compatible with the transition relation. In some
-cases (like lossy channel systems), the subword ordering appears naturally to
-order the configurations, but in most cases the natural ordering on words are
-the prefix, suffix, and infix orderings.
-
-To apply the WSTS theory in these cases, one way is to ensure that the set of
-reachable configurations (which is often a language) is wqo under the relevant
-ordering. Our results provide a complete characterization of such languages,
-along with effective decision procedures for languages described in a vast
-variety of formalisms. This completely solves the question of whether a WSTS
-with configurations ordered by prefix, suffix, or infix ordering can be used
-for verification. From our characterizations, one can understand that whenever
-one could use infix, prefix, or suffix orderings to obtain a wqo, the set of
-reachable configurations must be "simple" in a precise sense, which we
-interpret as a formal limitation to the expressiveness of such systems.
-
-Reviewer 2
-----------
-- You should give a proof for the measures of the languages Lk in the proof of Corollary 14.
-Response:
-We will add a detailed proof in the final version. Here is a sketch of the argument:
-The ordinal height of (infinite) languages is always ω,
-the ordinal width of a*b* is ω (since it is isomorphic to N^2 with the product order),
-hence disjoint copies of a*b* have width ω + .... + ω, hence the width of L_k is ω·k.
-Finally, the maximal order type of a disjoint sum is the natural sum (Hessenberg sum) of the maximal order types of the summands, which here is ω^2 added k times, hence ω^2 · k.
-We are using classical results on ordinal invariants of wqos that we will take from 
-[15].
-
-- You could corollary 21 at the end of section 5.1, since its proof call for lemmas 25 and 26. Or move these lemmas to appendix.
-Response:
-Yes, this was made to improve the flow of the paper by stating important results first,
-but in this case it is counterproductive.
-
-- Part of the claim of Lemma 26 seems redundant with Lemma 24.
-Response:
-Yes, the first part is the same and will be removed.
-
-- Unfortunately space is limited, but it would have been nice to have Figures 6 and 7 in the core of the paper.
-Response:
-These figures are quite large, and scaling them down would make them less readable.
-We will try moving them to the main text by compressing the prefix/suffix section 
-that is less central to the main results.
-
-Reviewer 3
-----------
-- Wouldn't it be more concise to characterise (in Thms 8 and 20) 
-IDEALS of languages well quasi ordered by the infix ordering, 
-instead of the wqo languages themselves?
-Response: For Theorem 8, we think that phrasing it in terms of ideals would needlessly complicate the statement. For Theorem 20, we are happy to rephrase it in terms of ideals.
-
-- How much the characterisation of Lemma 22 can be generalised
-beyond well quasi orders, namely for ideals in the infix order?
-Response: The proof can in fact be generalized to all order ideals 
-of the infix order, and we thank the Reviewer for pointing this out.
-We will adjust the statement and the proof.
-
-- Why Corollary 34 does not include VASS languages?
-Response: Corollary 34 is meant to emphasize a subpart of our results
-that we think might be of interest to a broader audience. The fully general
-version of this statement is Theorem 30, which of course includes VASS and 
-also PVASS languages.
-
-- There is no a really long and complex proof in the paper, rather a series of
-short arguments. As a reader I had an impressions of seeing a series
-of small fragmental results, while I'd probably prefer one big result
-instead of 20 small ones. 
-Response: As the space of infix well-quasi-orders was not well explored so far, we onsider this as a classification paper, which naturally lends itself to a large range of smaller statements rather than singular large ones.
-
-- The section on prefix/suffix could be probably completely removed from the
-paper, with the extra space devoted to the most important arguments that are
-now in the appendix.
-Response: We agree and will condense the section significantly to free up space,
-especially to reintroduce the diagrams to the main text as requested by Reviewer 2.
-
-- L.197-199: unclear what is the tree relation here.
-Response: The trees are formed as follows: The root node is the empty sequence, ε. Each node has (potentially infinitely many)
-children from the wqoset (X, ≤) such that the sequence of elements from the root to this child satisfy the property expressed
-by the tree (i.e., being bad, decreasing or an antichain respectively).
-
-- Thm 8: what are "products"?
-Response: By the "product" of two languages L and R we mean the usual pairwise concatenation of words from L and R.
-
-- Lemma 10: what does it mean "simultaneously" here?
-Response: By "simulatenously" we mean that we can find a finite set of languages such that their union is P↓(x), and each
-language is a chain by the prefix relation, and by the infix relation, and by the suffix relation.
-
-- Lemma 12: what is the meaning of "for arbitrarily large values of n" here?
-Response: By "for abritrarily large values of n", we formally mean: ∀k ∃n≥k such that the property holds.
-
-- Lemma 12: point 2 seems to be subsumed by 3, isn't it?
-Response: Correct, both point 1 and point 2 can be subsumed by point 3. The phrasing was left over from a previous version of the proof. We will adjust this.
-
-- L.605: what it means "we may fix"?
-Response: The concatenative property of amalgamation systems simply requires that each gap word is the concatenation of its 
-constituent gap words from the smaller runs, but not in which order. We may pick any gap and pick the ordering in that gap,
-but the ordering in the other gaps may then be imposed on us by the structure of the system. For example, when amalgamating
-derivation trees of CFGs, each gap has a sibling gap on the intuitive "other side" of the tree that must be concatenated in
-the reverse order.
-
-- Lemma 32: "or vice versa" not clear.
-Response: By "vice versa" we mean that either G_f ≤ G_g or G_g ≤ G_f
-
-- Lemma 32: "or vice versa" not clear.
-Response: By "vice versa" we mean that either G_f ≤ G_g or G_g ≤ G_f

### 2025-12-10 rebuttal.txt
fix: go below 600 words

--- 2025-12-10 rebuttal.txt
@@
 --R1
-Comments: We thank R1 for their positive and useful feedback.
+We thank R1 for their positive and useful feedback.
 
@@
 --R3
-Comments: We thank R3 for their appreciation of the quality of the technical level of the paper.
+We thank R3 for their appreciation of the quality of the technical level of the paper.
 Their main concern is the lack of a single large and complex technical contribution.
@@
 
-Question 1: Wouldn't it be more concise to characterise (in Thms 8 and 20) 
-IDEALS of languages well quasi ordered by the infix ordering, 
-instead of the wqo languages themselves?
+Answer to Question 1: Thank you for this suggestion. Concerning Theorem 8, we
+think that phrasing it in terms of ideals would needlessly complicate the
+statement. For Theorem 20, we are happy to rephrase it in terms of ideals.
 
-Answer: Thank you for this suggestion. Concerning Theorem 8, we think that phrasing it in terms of
-ideals would needlessly complicate the statement. For Theorem 20, we are happy
-to rephrase it in terms of ideals.
-
-Question 2: How much the characterisation of Lemma 22 can be generalised
-beyond well quasi orders, namely for ideals in the infix order?
-
-Answer: The proof can in fact be generalized to all order ideals of the
+Answer to Question 2: The proof of Lemma 22 can in fact be generalized to all order ideals of the
 infix order, and we thank Reviewer 3 for pointing this out. We will adjust
@@
 
-Question 3: Why Corollary 34 does not include VASS languages?
-Corollary 34 is meant to emphasize a subpart of our
+Answer to Question 3: Corollary 34 is meant to emphasize a subpart of our
 results that we think might be of interest to a broader audience. The fully
@@
 
----DETAILED ANSWER TO OTHER COMMENTS
+---minor-comments---
 
@@
 
-R2 on Figures 6 and 7, R3 on prefix/suffix section. We will be more succinct in the prefix/suffix section to free up space for these figures in the main text.
-
 R3 on Lemma 12 point 2 and 3. Correct, both point 1 and point 2 imply point 3,
 and we will adjust the statement accordingly.
-
-R3 on Corollary 34. Corollary 34 is meant to emphasize a subpart of our
-results that we think might be of interest to a broader audience. The fully
-general version of this statement is Theorem 30, which of course includes VASS
-and also PVASS languages.
-
-R3 on L.605.  The concatenative property of amalgamation systems simply
-requires that each gap word is the concatenation of its constituent gap words
-from the smaller runs, but not in which order. We may pick any gap and pick the
-ordering in that gap, but the ordering in the other gaps may then be imposed on
-us by the structure of the system.
-
-R3 on L.197-199. The tree is obtained by considering bad sequences (resp. antichains, decreasing sequences) as nodes, and the ancestor relation is "having common prefix".
 

### 2025-12-10 rebuttal.txt
fix: remove old stuff

--- 2025-12-10 rebuttal.txt
@@
 
-R3 on Corollary 34. Corollary 34 is meant to emphasize a subpart of our
-results that we think might be of interest to a broader audience. The fully
-general version of this statement is Theorem 30, which of course includes VASS
-and also PVASS languages.
-
-R3 on L.605.  The concatenative property of amalgamation systems simply
-requires that each gap word is the concatenation of its constituent gap words
-from the smaller runs, but not in which order. We may pick any gap and pick the
-ordering in that gap, but the ordering in the other gaps may then be imposed on
-us by the structure of the system.
-
-R3 on L.197-199. The tree is obtained by considering bad sequences (resp. antichains, decreasing sequences) as nodes, and the ancestor relation is "having common prefix".
-
 R3.
